### 2023-03-17 manuscript/main.tex
set up latex template

\documentclass[12pt, titlepage]{article}

%% Make sure every package in preamble is used
\usepackage{siunitx} % for unit
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\numberwithin{equation}{section}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{underscore}
\usepackage{physics}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{bm}
\usepackage{threeparttable}
\usepackage{setspace}
\usepackage[width=.95\textwidth, skip=0.1\baselineskip]{caption}

\allowdisplaybreaks

% \usepackage{adjustbox}
\usepackage[colorlinks, citecolor=blue,  urlcolor=blue,
linkcolor=blue]{hyperref}
\usepackage{xcolor}

% \usepackage[pagewise]{lineno}
\usepackage[]{lineno}
\linenumbers*[1]
%% patches to make lineno work better with amsmath
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
	\expandafter\let\csname old#1\expandafter\endcsname\csname
	#1\endcsname
	\expandafter\let\csname oldend#1\expandafter\endcsname\csname
	end#1\endcsname
	\renewenvironment{#1}%
	{\linenomath\csname old#1\endcsname}%
	{\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
	\patchAmsMathEnvironmentForLineno{#1}%
	\patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
	\patchBothAmsMathEnvironmentsForLineno{equation}%
	\patchBothAmsMathEnvironmentsForLineno{align}%
	\patchBothAmsMathEnvironmentsForLineno{flalign}%
	\patchBothAmsMathEnvironmentsForLineno{alignat}%
	\patchBothAmsMathEnvironmentsForLineno{gather}%
	\patchBothAmsMathEnvironmentsForLineno{multline}%
}

% control floats
\renewcommand\floatpagefraction{.9}
\renewcommand\topfraction{.9}
\renewcommand\bottomfraction{.9}
\renewcommand\textfraction{.2}
\setcounter{totalnumber}{50}
\setcounter{topnumber}{50}
\setcounter{bottomnumber}{50}

\allowdisplaybreaks % \usepackage{rotating}

\newcommand{\jy}[1]{\textcolor{orange}{JY: (#1)}}
\newcommand{\yh}[1]{\textcolor{purple}{YH: (#1)}}
\newcommand{\yx}[1]{\textcolor{cyan}{YX: (#1)}}
\let\proglang=\textsf
\newcommand{\pkg}[1]{{\fontseries{m}\selectfont #1}}
\newcommand\code[2][black]{\textcolor{#1}{\texttt{#2}}}

\usepackage{multirow, makecell}
\usepackage{booktabs,array}
\usepackage[affil-it]{authblk}
\renewcommand\Authfont{\fontsize{12}{14.4}\selectfont}
\renewcommand\Affilfont{\fontsize{12}{10.8}\itshape}
% \AddToHook{cmd/section/before}{\clearpage} % will remove

\title{Tentative: On Open Data Releasing Protocols}
\date{}

\author[1]{David Tussey}
\author[2]{Jun Yan\thanks{Tentatively in alphabetical order}}
\affil[1]{Independent Researcher?}
\affil[2]{Department of Statistics, University of Connecticut}

\begin{document}
\maketitle

\begin{abstract}
  Open data is important ...
  Quality control is important ...
  We use New York City 311 requests data  ...
  We suggest some open data quality control protocols ...

\bigskip

\noindent
{\it Keywords:}
consistency check;    
data release protocol;
quality control
\end{abstract}

\doublespacing

\section{Introduction} \label{sec:intro}

Open data is the trend ... many kinds \citep[e.g.,][]{cdc2022} ...

NYC is a forerunner in open data ... The 311 requests data is one of the most
popular open data ...

A clean, efficiently designed data is important ...

The rest of the paper is organized as follows.
Section~\ref{sec:data} presents the history and current status of the 311
requests data. Section~\ref{sec:issues} lists potential issues on data quality
and efficiency.
Section~\ref{sec:improve} provides suggestions on how to remove or minimize the
issues.
Section~\ref{sec:protocol} summarizes a few general principals on open data
release.
A discussion concludes in Section~\ref{sec:disc}

\section{NYC 311 Requests Data} \label{sec:data}

\section{Potential Issues} \label{sec:issues}

Invalid data: closed data earlier than creation date;
latitude/longitude not in NYC; invalid zip code;

Inconsistent data: upper/lower cases; extra spaces; 
open text entries: provide enough major categories and sub categories;

Inefficiently stored data: lat/long versus (lat, long);
intersection street 1 versus cross street 1;
borough could be inferred from zipcode;
full name of agency not needed

\section{Improvements} \label{sec:improve}

Data validation: identify invalid data

Data cleaning: convert to all lower/upper cases;

Data organization: put zipcode/borough in a separate table; verify intersection
== cross; remove redundant columns (and compare how much space saved)

\section{Protocol Suggestions} \label{sec:protocol}

\section{Discussion} \label{sec:disc}

\bibliographystyle{asa}
\bibliography{ref}

\end{document}

### 2023-05-13 talk/slides.Rmd

---
title: "Curation of NYC 311 Request Open Data"
author: ""     # to be supplied through headers-includes below
institute: ""  # to be supplied through headers-includes below
output:
  beamer_presentation:
    theme: "AnnArbor"
    colortheme: "dolphin"
    slide_level: 2
#    toc: true
    number_sections: false
classoption: "aspectratio=169"
header-includes:
  - \usepackage{amsmath, amsfonts, bm}
  - \usepackage{hyperref}\hypersetup{colorlinks=true, citecolor=blue}
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \usepackage{multirow}
  - \usepackage{adjustbox}
  - \setbeamercolor{frametitle}{bg=white}
  - \AtBeginSection[]
    {
    \begin{frame}<beamer>
      \frametitle{Outline}
      \tableofcontents[currentsection]
    \end{frame}
    }
  - \AtBeginDocument
    {
    \title[NYC 311 Requests]{Curation of NYC 311 Request Open Data}
    \author[Jun Yan]{David Tussey\inst{1} and Jun Yan\inst{2}}
    \institute[University of Connecticut]{\inst{1} Independent Researcher\\
    \inst{2} University of Connecticut}
    }
bibliography: ["../manuscript/ref.bib"]
link-citations: true
link-color: true
# csl: apa.csl
---

# Introduction

## NYC 311 Open Data

- Open data is great ... trend ...
- NYC is a forerunner in open data ...
- 311 request data is one of the most popular open data at NYC Open Data
- Found some issues from teaching ...

# Issues with NYC 311 Requests Data

## Data Structure

Demonstration with data for the week of Januare 15, 2023, 
downloaded on February 22, 2023.
```{r read}
d311 <- read.csv("../data/nyc311_011523-012123_by022023.csv", header = TRUE)
names(d311) <- tolower(gsub("\\.", "_", names(d311)))

names(d311)
```

## Invalid Created/Closed Date
```{r date}
t1 <- strptime(d311$closed_date, format = '%m/%d/%Y %I:%M:%S %p')
t0 <- strptime(d311$created_date, format = '%m/%d/%Y %I:%M:%S %p')
tt <- as.numeric(difftime(t1, t0, units =  "secs"))
sum(tt < 0, na.rm = TRUE)
head(subset(d311, tt < 0, select = c("created_date", "closed_date")), 5)
```

## Invalid Zipcode

## Inconsistent Data

# Suggestions on Open Data Releasing Protocols

# References {.allowframebreaks} 
\scriptsize

### 2023-06-01 talk/slides.Rmd
major topics included in slides

@@
     \author[Jun Yan]{David Tussey\inst{1} and Jun Yan\inst{2}}
-    \institute[University of Connecticut]{\inst{1} Independent Researcher\\
+    \institute[University of Connecticut]{\inst{1} Former Executive Director, NYC DoITT\\
     \inst{2} University of Connecticut}
@@
 
+```{r, echo = FALSE}
+library(stringr)
+source("../code/functions.R")
+## Load the 311 SR data file
+```
+
 # Introduction
@@
 
-- Open data is great ... trend ...
-- NYC is a forerunner in open data ...
-- 311 request data is one of the most popular open data at NYC Open Data
-- Found some issues from teaching ...
+- Open Data is viewed as a right for all.
+- Cities lead open data efforts wtih >50 cities with formal programs.
+- NYC is a forerunner with Open Data Law in 2012.
+- Today NYC Open Data has 3511 datasets from 98 Agencies, accesible online.
+- 311 Service Request data is the most viewed dataset at NYC Open Data.
+- Used in STAT 3255/5255 Introduction to Data Science at UConn
+- Exported 1-31 March 2023 311 SRs for analysis (~250K records).
 
@@
 
-## Data Structure
+## Data Columns
+<!-- - Column names: Remove trailing "."s.  Remove extra "."s -->
+<!-- - Convert "." separator to "_".  Covert column names to lower cases -->
+- Tidied column names (lower cases; underscore delimited; no trailing spaces)
+- Last five columns not referenced in the Data Dictionary (not the case last year)
+    + The NYC-OD Team advised these were "computed" fields derived from other data.
 
-Demonstration with data for the week of Januare 15, 2023, 
-downloaded on February 22, 2023.
 ```{r read}
-d311 <- read.csv("../data/nyc311_011523-012123_by022023.csv", header = TRUE)
-names(d311) <- tolower(gsub("\\.", "_", names(d311)))
-
-names(d311)
+d311 <- read.csv("../data/test_sample4.csv", header = TRUE)
+names(d311) <- makeColNamesUserFriendly(names(d311))
+dim(d311)
+tail(names(d311))
 ```
 
-## Invalid Created/Closed Date
+## Contributions by Agencies
+
+```{r byAgency}
+sort(table(d311$agency), decreasing = TRUE)
+```
+
+## Missing data: Character columns
+- Various forms: "", "unspecified", and "unknown"
+- Results show 3 groupings: all blanks (>90%), a few blanks (12-44%), or nearly no blanks (1-5%)
+
+```{r blanks}
+cMissingDataPerColumn <- charColumnsMissingData(d311)
+cMissingDataPerColumn[order(-rowSums(cMissingDataPerColumn)), ]
+```
+
+## Missing data: Numeric Colummns
+
+```{r missing}
+nMissingDataPerColumn <- numColumnsMissingData(d311)
+sort(nMissingDataPerColumn, decreasing = TRUE)
+```
+
+<!-- ## Unspecified/Unknown data columns -->
+<!-- - Several columns have many "Unspecified" or "Unknown" values. These are deemed valid values. -->
+<!-- ```{r uknown} -->
+<!-- tail(missingDataPerColumn[ order( -missingDataPerColumn[ , 2] ), ], 4) -->
+<!-- ``` -->
+
+## Invalid Data: Created/Closed Date
 ```{r date}
@@
 tt <- as.numeric(difftime(t1, t0, units =  "secs"))
-sum(tt < 0, na.rm = TRUE)
-head(subset(d311, tt < 0, select = c("created_date", "closed_date")), 5)
+
+badTime <- subset(d311, tt <= 0, select = c("created_date", "closed_date", "agency"))
+nrow(badTime)
+sum(tt == 0, na.rm = TRUE)
+summary(tt[tt < 0 & (!is.na(tt))])
+table(badTime$agency)
 ```
 
-## Invalid Zipcode
+## Invalid Data: Incident Zipcodes
+```{r zip1}
+validZipcodes <- scan("../data/NYCzipcodes.csv")
 
-## Inconsistent Data
+badIncident_zip <- subset(d311, ! incident_zip %in% validZipcodes,
+                          select = c("incident_zip", "agency"))
+head(sort(table(badIncident_zip$incident_zip), decreasing = TRUE))
+```
+
+## Invalid Data: Zipcodes (One of the last 5 columns)
+```{r zip2}
+badZipcode  <- subset(d311, ! zip_codes %in% validZipcodes,
+                      select = c("zip_codes", "agency"))
+nrow(badZipcode)  # mostly invalid!
+```
+
+## Invalid Data: Borough
+```{r borough}
+validBorough <- c("BRONX", "BROOKLYN", "MANHATTAN", "QUEENS", "STATEN ISLAND")
+badBorough <- subset(d311, ! borough %in% validBorough,
+                     select = c("borough", "agency"))
+head(table(badBorough$agency))
+
+validBB <- c("1", "2", "3", "4", "5")
+badBB <- subset(d311, ! borough_boundaries %in% validBB,
+                select = c("borough_boundaries", "agency"))
+head(table(badBB$agency))
+
+badPark_borough <- subset(d311, ! park_borough %in% validBorough,
+                          select = c("park_borough", "agency"))
+head(table(badPark_borough$agency))
+```
+
+## Invalid Data: Open Data Channel
+```{r channel}
+validChannel <- c("MOBILE", "ONLINE", "PHONE")
+badChannel_type <- subset(d311, ! open_data_channel_type %in% validChannel,
+                          select = c("open_data_channel_type", "agency"))
+head(table(badChannel_type$agency))
+```
+
+## Invalid Data: Police Precincts
+
+```{r nypd}
+validPolice_precincts <- scan("../data/NYPDPrecincts2023.csv", skip = 1)
+badPolice_precincts <- subset(d311, ! police_precincts %in% validPolice_precincts,
+                              select = c("police_precincts", "agency"))
+sort(table(badPolice_precincts$agency), decreasing = TRUE)
+```
+
+## Redundant Data: Borough
+```{r redundant_borough}
+## 'borough' and 'park_borough' are identical
+any(d311$borough != d311$park_borough)
+
+## 'borough' and 'borough_boundaries' are almost identical
+nonMatchingBorough <- subset(
+    d311, borough_boundaries !=
+          as.integer(str_replace_all(borough,
+                                     c("STATEN ISLAND" = "1",
+                                       "BROOKLYN"      = "2",
+                                       "QUEENS"        = "3",
+                                       "MANHATTAN"     = "4",
+                                       "BRONX"         = "5",
+                                       "UNSPECIFIED"   = NA))),
+    select = c("borough", "borough_boundaries", "agency"))
+nrow(nonMatchingBorough)
+```
+
+## Redundant Data: Latitude, Longitude, and Location
+```{r location}
+head(d311[, c("latitude", "longitude", "location")])
+```
+
+## Redundant Data: Cross Street and Intersection Street
+```{r cross_intersection}
+## one empty but the other not
+nrow(subset(d311, cross_street_1 == "" & intersection_street_1 != ""))
+nrow(subset(d311, cross_street_1 != "" & intersection_street_1 == ""))
+
+## if both not empty, they are almost the same
+head(subset(d311, cross_street_1 != "" & intersection_street_1 != "" &
+                  cross_street_1 != intersection_street_1,
+            select = c("cross_street_1", "intersection_street_1", "agency")))
+```
+
+## Redundant Data: Cross Street and Intersection Street (cont'd)
+```{r extra_space}
+## after removing extra spaces, most of them are the same
+dupXStreets <-subset(
+    d311, gsub(" ", "", cross_street_1) ==
+          gsub(" ", "", intersection_street_1),
+    select = c("cross_street_1", "intersection_street_1", "agency"))
+head(dupXStreets)
+table(dupXStreets$agency)
+head(subset(dupXStreets, cross_street_1 != intersection_street_1))
+```
+
+## Less-Efficient Storage: Borough, Agency
+```{r storage}
+## code borough by integer
+object.size(d311$borough_boundaries)
+object.size(d311$borough)
+
+## code agency by integer
+object.size(d311$agency)
+agency_code <- rep(1:14, table(d311$agency))
+object.size(agency_code)
+```
 
@@
 
+## Suggestions on Data Curation
+
++ Validity check 
+    - date (logical check)
+    - descriptive fields (drop-down menu)
+    - legitimate values
++ Efficiency
+    - Eliminate redundant columns
+    - Reduce storage with no information loss
++ Feedback: open review for open data
+
 # References {.allowframebreaks} 

### 2023-06-01 code/cleansing.R

@@ -1,0 +1,142 @@
+library(stringr)
+
+source("../code/functions.R")
+##  Create the path to the file containing the 311 Service Request data.
+data1File <- file.path("../data", "test_sample4.csv") # 311_202303.csv")
+
+#########################################################################
+## Load the 311 SR data file
+d311 <- read.csv(data1File, header = TRUE)
+names(d311) <- makeColNamesUserFriendly(names(d311))
+
+## number of rows
+numRows <- nrow( d311 )
+
+## counts by agency
+sort(table(d311$agency), decreasing = TRUE)
+
+#########################################################################
+## check for missing info
+
+cMissingDataPerColumn <- charColumnsMissingData(d311)
+cMissingDataPerColumn[order(-rowSums(cMissingDataPerColumn)), ]
+
+
+nMissingDataPerColumn <- numColumnsMissingData(d311)
+sort(nMissingDataPerColumn, decreasing = TRUE)
+
+
+#########################################################################
+## any creation date later than closing date
+t1 <- strptime(d311$closed_date, format = '%m/%d/%Y %I:%M:%S %p')
+t0 <- strptime(d311$created_date, format = '%m/%d/%Y %I:%M:%S %p')
+tt <- as.numeric(difftime(t1, t0, units =  "secs"))
+
+badTime <- subset(d311, tt <= 0, select = c("created_date", "closed_date", "agency"))
+nrow(badTime)
+sum(tt == 0, na.rm = TRUE)
+summary(tt[tt < 0 & (!is.na(tt))])
+table(badTime$agency)
+
+#########################################################################
+## Check for allowable values
+validBorough <- c("BRONX", "BROOKLYN", "MANHATTAN", "QUEENS", "STATEN ISLAND")
+badBorough <- subset(d311, ! borough %in% validBorough,
+                     select = c("borough", "agency"))
+head(table(badBorough$agency))
+
+validBB <- c("1", "2", "3", "4", "5")
+badBB <- subset(d311, ! borough_boundaries %in% validBB,
+                select = c("borough_boundaries", "agency"))
+head(table(badBB$agency))
+
+badPark_borough <- subset(d311, ! park_borough %in% validBorough,
+                          select = c("park_borough", "agency"))
+head(table(badPark_borough$agency))
+
+validChannel <- c("MOBILE", "ONLINE", "PHONE")
+badChannel_type <- subset(d311, ! open_data_channel_type %in% validChannel,
[…]

### 2023-06-01 talk/slides.Rmd
wrapped long lines

@@
 tt <- as.numeric(difftime(t1, t0, units =  "secs"))
-
-badTime <- subset(d311, tt <= 0, select = c("created_date", "closed_date", "agency"))
+badTime <- subset(d311, tt <= 0,
+                  select = c("created_date", "closed_date", "agency"))
 nrow(badTime)
@@
 head(table(badBorough$agency))
+```
 
+## Invalid Data: Borough_boundaries and Park_borough
+```{r borough2}
 validBB <- c("1", "2", "3", "4", "5")
@@
 ## Invalid Data: Police Precincts
-
 ```{r nypd}
 validPolice_precincts <- scan("../data/NYPDPrecincts2023.csv", skip = 1)
-badPolice_precincts <- subset(d311, ! police_precincts %in% validPolice_precincts,
+badPolice_precincts <- subset(
+    d311, ! police_precincts %in% validPolice_precincts,
     select = c("police_precincts", "agency"))

### 2023-06-05 talk/slides.Rmd

@@
 ```{r read}
-d311 <- read.csv("../data/test_sample4.csv", header = TRUE)
+d311 <- read.csv("../data/311_202303.csv", header = TRUE)
 names(d311) <- makeColNamesUserFriendly(names(d311))
@@
 ```{r byAgency}
-sort(table(d311$agency), decreasing = TRUE)
+d311freq_agency <- table(d311$agency)
+sort(d311freq_agency, decreasing = TRUE)
 ```
@@
 object.size(d311$agency)
-agency_code <- rep(1:14, table(d311$agency))
+agency_code <- rep(seq(length(d311freq_agency)), d311freq_agency)
 object.size(agency_code)
@@
 
-# Suggestions on Open Data Releasing Protocols
+# Conclusion
 
@@
 
+## Discussion
+
++ something
++ something else
+
 # References {.allowframebreaks} 

### 2023-06-05 talk/slides.Rmd
minor edits to slides

@@
     - date (logical check)
-    - descriptive fields (drop-down menu)
+    - descriptive fields (drop-down menu, possibly multi-step)
+    - missing values
     - legitimate values
+    - computed fields in the dictionary
 + Efficiency

### 2023-06-06 talk/slides.Rmd
reduced default r chunk font size and minor edits to slides

--- 2023-06-05 talk/slides.Rmd
@@
 
-```{r, echo = FALSE}
+```{r setup, include = FALSE, echo = FALSE}
 library(stringr)
 source("../code/functions.R")
+
+def.chunk.hook  <- knitr::knit_hooks$get("chunk")
+knitr::knit_hooks$set(chunk = function(x, options) {
+  x <- def.chunk.hook(x, options)
+  paste0("\n \\", "small","\n\n", x, "\n\n \\normalsize")
+})
 ## Load the 311 SR data file
@@
                   select = c("created_date", "closed_date", "agency"))
-nrow(badTime)
+
+## exactly the same created and closed date down to the second; a lot!
 sum(tt == 0, na.rm = TRUE)
+table(subset(badTime, created_date == closed_date, select = "agency"))
+```
+
+## Bad Closed Date
+```{r baddate}
+## quite some negative response times
+table(subset(badTime, created_date != closed_date, select = "agency"))
+
 summary(tt[tt < 0 & (!is.na(tt))])
-table(badTime$agency)
 ```
@@
     d311, borough_boundaries !=
-          as.integer(str_replace_all(borough,
-                                     c("STATEN ISLAND" = "1",
-                                       "BROOKLYN"      = "2",
-                                       "QUEENS"        = "3",
-                                       "MANHATTAN"     = "4",
-                                       "BRONX"         = "5",
-                                       "UNSPECIFIED"   = NA))),
+          as.integer(str_replace_all(
+              borough, c("STATEN ISLAND" = "1",  "BROOKLYN"      = "2",
+                         "QUEENS"        = "3",  "MANHATTAN"     = "4",
+                         "BRONX"         = "5",  "Unspecified"   = NA))),
     select = c("borough", "borough_boundaries", "agency"))
@@
 ```{r location}
-head(d311[, c("latitude", "longitude", "location")])
+print(head(d311[, c("latitude", "longitude", "location")]), digits = 16)
 ```
@@
 
-## Redundant Data: Cross Street and Intersection Street (cont'd)
+##  Cross Street and Intersection Street (cont'd)
 ```{r extra_space}
@@
 head(dupXStreets)
+```
+
+##  Cross Street and Intersection Street (cont'd) 
+```{r dup_streets}
 table(dupXStreets$agency)
@@
 
-## Less-Efficient Storage: Borough, Agency
+## Less-Efficient Storage: Download File Size
 ```{r storage}
@@
 
-## Suggestions on Data Curation
+## Suggestions on Open Data Curation
 
@@
     - date (logical check)
-    - descriptive fields (drop-down menu)
     - legitimate values 
+	- descriptive fields (drop-down menu, possibly multi-step)
+    - missing value: coding; validity
+    - computed fields in the dictionary
 + Efficiency
@@
     - Reduce storage with no information loss
-+ Feedback: open review for open data
++ Feedback: 
+    - open review for open data
+	- interaction with the maintenance team
 

### 2023-06-09 talk/slides.Rmd
went through slides with David

@@
 ```{r read}
-d311 <- read.csv("../data/311_202303.csv", header = TRUE)
+d311 <- read.csv("../data/311_202303by202305.csv", header = TRUE)
 names(d311) <- makeColNamesUserFriendly(names(d311))
@@
 t0 <- strptime(d311$created_date, format = '%m/%d/%Y %I:%M:%S %p')
-tt <- as.numeric(difftime(t1, t0, units =  "secs"))
+tt <- as.numeric(difftime(t1, t0, units =  "day"))
 badTime <- subset(d311, tt <= 0,
@@
 ```{r borough}
-validBorough <- c("BRONX", "BROOKLYN", "MANHATTAN", "QUEENS", "STATEN ISLAND")
+validBorough <- c("BRONX", "BROOKLYN", "MANHATTAN", "QUEENS", "STATEN ISLAND",
+                  "Unspecified")
 badBorough <- subset(d311, ! borough %in% validBorough,
@@
 head(table(badBorough$agency))
+head(table(subset(d311, borough == "Unspecified")$agency))
 ```
@@
 ```{r channel}
-validChannel <- c("MOBILE", "ONLINE", "PHONE")
+validChannel <- c("MOBILE", "ONLINE", "PHONE", "OTHER")
 badChannel_type <- subset(d311, ! open_data_channel_type %in% validChannel,
                           select = c("open_data_channel_type", "agency"))
-head(table(badChannel_type$agency))
+head(table(badChannel_type$agency)) ## UNKNOWN
 ```
@@
 ## after removing extra spaces, most of them are the same
-dupXStreets <-subset(
+dupXStreets <- subset(
     d311, gsub(" ", "", cross_street_1) ==
@@
     select = c("cross_street_1", "intersection_street_1", "agency"))
+nrow(dupXStreets)
 head(dupXStreets)
@@
 
-## Less-Efficient Storage: Download File Size
+## Less-Efficient Storage: Memory
 ```{r storage}
@@
 
+## Less-Efficient Storage: Disk
+```{r storage-disk}
+write.csv(subset(
+    d311, select = -c(park_borough, borough_boundaries, location,
+                      intersection_street_1, intersection_street_2,
+                      agency_name)),
+    quote=FALSE, row.names = FALSE,
+    file = "../data/311_202303_smaller.csv")
+
+(newsize <- file.size("../data/311_202303_smaller.csv"))
+(oldsize <- file.size("../data/311_202303by202305.csv"))
+(oldsize - newsize) / oldsize
+```
+
 # Conclusion
@@
 
-+ Validity check 
++ Improve dictionary
+    - Code categorical variables (e.g., agency) 
+	- Code descriptive fields (e.g., complain type; vehicle type)
+    - Make the descriptions informative (e.g., zip code)
++ Impose validity check
     - date (logical check)
@@
     - missing value: coding; validity
-    - computed fields in the dictionary
-+ Efficiency
++ Improve data storage/transportation efficiency
     - Eliminate redundant columns
@@
 + Feedback: 
-    - open review for open data
-	- interaction with the maintenance team
-
-## Discussion
-
-+ something
-+ something else
+    - Open review for open data
+	- Interaction with the maintenance team
 

### 2023-06-09 talk/slides.Rmd

@@
 
-## NYC 311 Open Data
+## NYC Open Data
 
-- Open Data is viewed as a right for all.
+- Open Data is viewed as a right for all [@janssen2012benefits; @wang2016adoption].
 - Cities lead open data efforts wtih >50 cities with formal programs.
-- NYC is a forerunner with Open Data Law in 2012.
+- NYC is a forerunner with Open Data Law in 2012 [@zuiderwijk2014open].
 - Today NYC Open Data has 3511 datasets from 98 Agencies, accesible online.
+- Used in many research
+    + Health [@cantor2018facets, shankar2021data]
+    + Smart city [@neves2020impacts]
+    + Transportation [@gerte2019understanding]
+
+## NYC 311 Service Requests
+
 - 311 Service Request data is the most viewed dataset at NYC Open Data.
 - Used in STAT 3255/5255 Introduction to Data Science at UConn
-- Exported 1-31 March 2023 311 SRs for analysis (~250K records).
+- In this study, we exported 1-31 March 2023 311 SRs for analysis (~251K records).
+- Data curation is imporant issue for open data such as the 311 SRs [@witt2009constructing; @borgman2012conundrum; @mclure2014data; @barns2016mine; @hart2016ten; @beheshti2019datasynapse; @shankar2021data].
 

### 2023-06-09 talk/slides.Rmd
title matching GASP program

@@
 ---
-title: "Curation of NYC 311 Request Open Data"
+title: "On Principles for Government Open Data Curation"
 author: ""     # to be supplied through headers-includes below
@@
     {
-    \title[NYC 311 Requests]{Curation of NYC 311 Request Open Data}
+    \title[NYC 311 Requests]{On Principles for Government Open Data Curation\\A Case Study with NYC 311 Servece Requests}
     \author[Jun Yan]{David Tussey\inst{1} and Jun Yan\inst{2}}

### 2023-06-10 talk/slides.Rmd
minor edits

@@
 - NYC is a forerunner with Open Data Law in 2012 [@zuiderwijk2014open].
-- Today NYC Open Data has 3511 datasets from 98 Agencies, accesible online.
-- Used in many research
+- Today NYC Open Data has 3511 datasets from 98 Agencies, accesible online 
+  (\url{https://opendata.cityofnewyork.us}).
+- Used by many researchers:
     + Health [@cantor2018facets, shankar2021data]
@@
 - 311 Service Request data is the most viewed dataset at NYC Open Data.
-- Used in STAT 3255/5255 Introduction to Data Science at UConn
+- NYC 311 is a unified frontline for all service requests to the NYC government agents.
+- The data was used in STAT 3255/5255 Introduction to Data Science at UConn, with student presentations at 2022/2023 NYC Open Data Week.
 - In this study, we exported 1-31 March 2023 311 SRs for analysis (~251K records).
-- Data curation is imporant issue for open data such as the 311 SRs [@witt2009constructing; @borgman2012conundrum; @mclure2014data; @barns2016mine; @hart2016ten; @beheshti2019datasynapse; @shankar2021data].
+- Data curation is critical for open data such as the 311 SRs [@witt2009constructing; @borgman2012conundrum; @mclure2014data; @barns2016mine; @hart2016ten; @beheshti2019datasynapse; @shankar2021data].
+- Issues
+    - Validity
+	- Efficiency
 
@@
                       select = c("zip_codes", "agency"))
-nrow(badZipcode)  # mostly invalid!
+nrow(badZipcode)  # mostly invalid! One possibility: debt collectors
 ```
@@
 ## Invalid Data: Open Data Channel
+"Due to pandemic call handling modifications, the ‘Open Data Channel Type’ values may not accurately indicate the channel the Service Request was submitted in for the period starting March 2020."
+
 ```{r channel}
@@
 print(head(d311[, c("latitude", "longitude", "location")]), digits = 16)
+```
+
+## Redundant Data: Agence, Agency Name, and Other Categorical Variables
+```{r agent}
+head(d311[, c("agency", "agency_name")])
+head(sort(table(d311$complaint_type), decreasing = TRUE)) # 174 types
 ```
@@
     - Reduce storage with no information loss
-+ Feedback: 
+
+## Outlook
+
 - Open review for open data
+
 - Interaction with the maintenance team
+
+- Open data curation protocol
+
+- Open data analytic portal
 

### 2023-06-16 talk/slides.Rmd
added date

@@
 title: "On Principles for Government Open Data Curation"
+date: "June 15, 2023 @ GASP23"
 author: ""     # to be supplied through headers-includes below
@@
     {
-    \title[NYC 311 Requests]{On Principles for Government Open Data Curation\\A Case Study with NYC 311 Servece Requests}
+    \title[NYC 311 Requests]{On Principles for Government Open Data Curation:\\A Case Study with NYC 311 Servece Requests}
     \author[Jun Yan]{David Tussey\inst{1} and Jun Yan\inst{2}}

### 2023-06-17 nyc311clean.qmd
Quarto file for writing Arrow from CSV

---
title: "Creating an Arrow IPC file from the NYC 311 data"
author: 
  - name: Jun Yan
    affiliation: University of Connecticut
  - name: Douglas Bates
    affiliation: University of Wisconsin - Madison
---

## Abstract {.unnumbered}

We show conversion of a CSV file containing data on 311 service calls during March, 2023 in New York City to the [Arrow](https://arrow.apache.org) IPC (Inter-Process Communications) file format (also called the Feather V2 format) which can be read quickly and easily into many different data science environments.
This conversion is done in [Julia](https://julialang.org) but it could equally well have been done in a number of different languages, whichever is most convenient for the person doing the conversion.

# Reading the CSV file as a DataFrame

First we attach the packages to be used,

```{julia}
#| output: false
using Arrow, CSV, Dates, DataFrames
```

The first two lines of the CSV file are

```{julia}
#| panel: fill
open("./data/311_half_Mar_2023.csv", "r") do io
    println(readline(io))
    println(readline(io))
    seekstart(io)
end;
```

::: {.callout-note collapse="false"}
### Need to scroll right to see the entire line

The output from this code block has very long lines.
You will need to scroll right in the output to read it all.

Note that the character on the right of the title bar for callout blocks like this is a toggle to show or hide the contents of the block.
:::

We now have enough information, such as the date-time format used, to do an initial conversion of the CSV file to a `DataFrame`, using several optional arguments to customize the process.
(Details about each of these optional arguments are given callout blocks below.)

```{julia}
df = CSV.read(
    "./data/311_half_Mar_2023.csv",  # file name
    DataFrame;                       # output table type
    normalizenames=true, # convert column names to valid symbols
    missingstring=[      # strings that are used to indicate missing vals
        "",
        "Unspecified",
        "unspecified",
        "Unknown",
        "unknown",
        "NA",
    ],
    dateformat=dateformat"m/d/Y I:M:S p", # DateTimes look like this
    downcast=true,       # use smallest type of Int that can represent the column 
    stripwhitespace=true # strip leading and trailing whitespace 
)
```

:::{.callout-note collapse="true"}
### normalize names
Convert the column names in the first row of the CSV file into a form suitable as a `Symbol` in `Julia`.
In particular, strip whitespace at the beginning and end of the name and replace embedded blanks or `.` characters by underscores.
:::

:::{.callout-note collapse="true"}
### missingstring
Enumerate the various forms of missing data indicators used in the file.
Completing this list is often a process of trial-and-error.
:::

:::{.callout-note collapse="true"}
### dateformat
Provide a `Dates.DateFormat` object that will be applied to text columns if they match the pattern, converting to `DateTime` values.

The construction `dateformat"..."` is a call to a "string macro" in Julia.
That is, it is just a shorthand for replacing the string with the `Dates.DateFormat` object created from the string.
:::

:::{.callout-note collapse="true"}
### downcast
For columns of integer values, choose the smallest type that can represent all the values in the column.
:::

# Validity checking

## Review the storage type for each column

In the output (you must scroll horizontally to see it all) under the column name is an indicator of the type.
A type name that ends in `?` indicates that missing values are present in the column.
Thus the `Unique_Key` is stored as 32-bit signed integers with no missing values.
The `Created_Date` is a `DateTime` with no missing values but the `Closed_Date`, which is also a `DateTime`, does have missing values.

Some columns are given are `String` whereas others are given as `String7` or `String15`.
These are storage-saving representations used when the whole column consists of short character strings.
There is an anomaly in these short strings for the `Agency` column.
The `Agency` acronyms should be 4 or 5 characters at most but at least one (it happens to be `NYC311-PRD`) must be greater than 7 characters to force `String15` storage.

```{julia}
show(unique(df.Agency))
```

The `Agency` and `Agency_Name` columns should be in one-to-one correspondence but they are not.

```{julia}
unique(select(df, :Agency, :Agency_Name))
```

At this point we would need to check with the data providers to determine why two different `Agency` designations are used for the "Department of Environmental Protection".

The other problem shown in the types is that the `Incident_Zip` column is stored as strings when they should be integers.

```{julia}
levels(df.Incident_Zip)
```

and, of course, the problem is that someone got creative and used `na` as missing value indicator.

Recreate the DataFrame with `"na"` in the `missingstring` vector

```{julia}
df = CSV.read(
    "./data/311_half_Mar_2023.csv",  # file name
    DataFrame;                       # output table type
    normalizenames=true, # convert column names to valid symbols
    missingstring=[      # strings that are used to indicate missing vals
        "",
        "Unspecified",
        "unspecified",
        "Unknown",
        "unknown",
        "NA",
        "na",
    ],
    dateformat=dateformat"m/d/Y I:M:S p", # DateTimes look like this
    downcast=true,       # use smallest type of Int that can represent the column 
    stripwhitespace=true # strip leading and trailing whitespace 
)
```

## Further checks

### Are the `Unique_Key` values unique?
```{julia}
length(unique(df.Unique_Key)) == nrow(df)
```

### Are the `Created_Date` values in the specified range?

```{julia}
extrema(df.Created_Date)
```

### Are the `Closed_Date` values in the expected range?

```{julia}
extrema(skipmissing(df.Closed_Date))
```

:::{.callout-note collapse="true"}
### use of skipmissing
There are missing values in the `Closed_Date` column which must be skipped when evaluating the extremes.
:::

Now we can see a problem with the record giving the earliest `Closed_Date` because it is before the `Created_Date` range.

These rows can be isolated for manual checking

```{julia}
earlyclosing = (!ismissing).(df.Closed_Date) .&& (df.Closed_Date .< df.Created_Date)
immediateclosing = (!ismissing).(df.Closed_Date) .&& (df.Closed_Date .== df.Created_Date)
(early = count(earlyclosing), immediate = count(immediateclosing))
```

:::{.callout-note collapse="true"}
### What are all those dots for?
Julia provides "syntactic sugar" to vectorize a scalar operation with a `.` following a function name (e.g. `(!ismissing).`) or preceding an operator (e.g. `.<`)
:::

If we want to examine the rows with these characteristics we index the rows with the `BitVector` created from the logical expression.
To see the first 3 rows with early closings

```{julia}
first(df[earlyclosing, :], 3)
```

:::{.callout-note collapse="true"}
### Use of `:` as an index indicates all values for that index
In this case the use of `:` as the second index indicates that all the columns should be included in the subset.
:::

## Is Location redundant if Latitude and Longitude are given?

This is a tricky question because the values of `Latitude` and `Longitude` are stored to 17 significant digits, which is much more accuracy than could possibly be measured.
To check for consistency we need to convert the `Location` to a `Latitude-Longitude` pair then compare these values to the recorded `Latitude` and `Longitude` with some tolerance.

# Saving the dataset as an Arrow IPC file

We use `Arrow.write` to write the dataframe in the Arrow IPC format.

```{julia}
#| output: false
fn = Arrow.write("./data/311_half_March_2023.arrow", df)
```

The Arrow file is less than half the size of the original CSV file but contains much more specificity about the data types and representations.
Furthermore it can be read into other data science environments without needing to tinker with optional arguments.

The Arrow IPC format and the `Arrow.write` function in the Julia package allow for optional compression of the contents using `lz4` or `zstd` if further reduction of the file size is desired.
The trade-off is that the uncompressed file can be memory-mapped by some systems to make for extremely fast input.

This file can be read into Julia

```{julia}
tbl = Arrow.Table(fn)
```

or R
```r
> tibble::glimpse(arrow::read_ipc_file("./data/311_half_March_2023.arrow"))
Rows: 125,826
Columns: 46
$ Unique_Key                     <int> 57000688, 56993837, 57049479, 57056381,…
$ Created_Date                   <dttm> 2023-03-09 09:05:25, 2023-03-08 15:27:…
$ Closed_Date                    <dttm> 2023-03-13 03:21:11, 2023-05-03 02:17:…
$ Agency                         <fct> DOE, TLC, HPD, NYPD, DOB, DPR, DOT, NYP…
$ Agency_Name                    <fct> Department of Education, Taxi and Limou…
$ Complaint_Type                 <fct> School Maintenance, For Hire Vehicle Co…
$ Descriptor                     <chr> "Other School Condition", "Driver Compl…
$ Location_Type                  <fct> School, Street, RESIDENTIAL BUILDING, S…
$ Incident_Zip                   <int> NA, 10075, 10459, 10455, 11365, 11691, …
$ Incident_Address               <chr> NA, "995 MADISON AVENUE", "1010 BRYANT …
$ Street_Name                    <chr> NA, "MADISON AVENUE", "BRYANT AVENUE", …
$ Cross_Street_1                 <chr> NA, "EAST   77 STREET", NA, "EAST  147 …
$ Cross_Street_2                 <chr> NA, "EAST   78 STREET", NA, "EAST  149 …
$ Intersection_Street_1          <chr> NA, "EAST   77 STREET", NA, "EAST  147 …
$ Intersection_Street_2          <chr> NA, "EAST   78 STREET", NA, "EAST  149 …
$ Address_Type                   <fct> NA, ADDRESS, ADDRESS, ADDRESS, ADDRESS,…
$ City                           <fct> NA, NEW YORK, BRONX, BRONX, FRESH MEADO…
$ Landmark                       <chr> NA, "MADISON AVENUE", NA, "UNION AVENUE…
$ Facility_Type                  <fct> NA, NA, NA, NA, NA, NA, N/A, NA, NA, NA…
$ Status                         <fct> Closed, Closed, Closed, Closed, Assigne…
$ Due_Date                       <dttm> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA…
$ Resolution_Description         <fct> "The Department of Education determined…
$ Resolution_Action_Updated_Date <dttm> 2023-03-13 03:21:17, 2023-05-03 02:17:…
$ Community_Board                <fct> Unspecified BROOKLYN, 08 MANHATTAN, 02 …
$ BBL                            <int64> NA, 1013927501, 2027560001, 202582004…
$ Borough                        <fct> BROOKLYN, MANHATTAN, BRONX, BRONX, QUEE…
$ X_Coordinate_State_Plane_      <int> 989013, 994507, 1015090, 1010420, 10388…
$ Y_Coordinate_State_Plane_      <int> 173052, 221687, 239359, 235260, 206195,…
$ Open_Data_Channel_Type         <fct> PHONE, ONLINE, PHONE, MOBILE, UNKNOWN, …
$ Park_Facility_Name             <fct> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA,…
$ Park_Borough                   <fct> BROOKLYN, MANHATTAN, BRONX, BRONX, QUEE…
$ Vehicle_Type                   <fct> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA,…
$ Taxi_Company_Borough           <fct> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA,…
$ Taxi_Pick_Up_Location          <chr> NA, "995 MADISON AVENUE, MANHATTAN (NEW…
$ Bridge_Highway_Name            <fct> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA,…
$ Bridge_Highway_Direction       <fct> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA,…
$ Road_Ramp                      <fct> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA,…
$ Bridge_Highway_Segment         <fct> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA,…
$ Latitude                       <dbl> 40.64166, 40.77515, 40.82361, 40.81237,…
$ Longitude                      <dbl> -73.98284, -73.96297, -73.88857, -73.90…
$ Location                       <chr> "(40.641663101476226, -73.9828374867858…
$ Zip_Codes                      <int> 17620, 10092, 10937, 10933, 14507, 2052…
$ Community_Districts            <int> 2, 23, 8, 49, 25, 51, NA, 26, 5, 24, 47…
$ Borough_Boundaries             <int> 2, 4, 5, 5, 3, 3, NA, 3, 2, 5, 4, 5, 3,…
$ City_Council_Districts         <int> 27, 51, 43, 35, 24, 47, NA, 16, 8, 29, …
$ Police_Precincts               <int> 39, 11, 24, 23, 65, 59, NA, 69, 38, 34,…
```
:::{.callout-note collapse="true"}
### R applies a time zone to the datetime values
The R `arrow` package appears to read the DateTime values as if they are in given in UTC then prints them in the local time zone.
:::

or Python (not that this format is called `feather` in the `pyarrow` package)
```python
>>> import pyarrow.feather as fea
>>> fea.read_table("./data/311_half_March_2023.arrow")
pyarrow.Table
Unique_Key: int32 not null
Created_Date: timestamp[ms] not null
Closed_Date: timestamp[ms]
Agency: dictionary<values=string, indices=int8, ordered=0> not null
Agency_Name: dictionary<values=string, indices=int8, ordered=0> not null
Complaint_Type: dictionary<values=string, indices=int16, ordered=0> not null
Descriptor: string
Location_Type: dictionary<values=string, indices=int8, ordered=0>
Incident_Zip: int32
Incident_Address: string
Street_Name: string
Cross_Street_1: string
Cross_Street_2: string
Intersection_Street_1: string
Intersection_Street_2: string
Address_Type: dictionary<values=string, indices=int8, ordered=0>
City: dictionary<values=string, indices=int8, ordered=0>
Landmark: string
Facility_Type: dictionary<values=string, indices=int8, ordered=0>
Status: dictionary<values=string, indices=int8, ordered=0>
Due_Date: timestamp[ms]
Resolution_Description: dictionary<values=string, indices=int16, ordered=0>
Resolution_Action_Updated_Date: timestamp[ms]
Community_Board: dictionary<values=string, indices=int8, ordered=0> not null
BBL: int64
Borough: dictionary<values=string, indices=int8, ordered=0>
X_Coordinate_State_Plane_: int32
Y_Coordinate_State_Plane_: int32
Open_Data_Channel_Type: dictionary<values=string, indices=int8, ordered=0> not null
Park_Facility_Name: dictionary<values=string, indices=int16, ordered=0>
Park_Borough: dictionary<values=string, indices=int8, ordered=0>
Vehicle_Type: dictionary<values=string, indices=int8, ordered=0>
Taxi_Company_Borough: dictionary<values=string, indices=int8, ordered=0>
Taxi_Pick_Up_Location: string
Bridge_Highway_Name: dictionary<values=string, indices=int8, ordered=0>
Bridge_Highway_Direction: dictionary<values=string, indices=int8, ordered=0>
Road_Ramp: dictionary<values=string, indices=int8, ordered=0>
Bridge_Highway_Segment: dictionary<values=string, indices=int16, ordered=0>
Latitude: double
Longitude: double
Location: string
Zip_Codes: int16
Community_Districts: int8
Borough_Boundaries: int8
City_Council_Districts: int8
Police_Precincts: int8
----
Unique_Key: [[57000688,56993837,57049479,57056381,57023097,...,57116487,56954525,57132547,57102740,56950149]]
Created_Date: [[2023-03-09 15:05:25.000,2023-03-08 21:27:34.000,2023-03-15 19:26:41.000,2023-03-16 23:58:53.000,2023-03-12 15:15:54.000,...,2023-03-22 17:47:08.000,2023-03-04 16:23:00.000,2023-03-24 21:57:44.000,2023-03-21 05:10:12.000,2023-03-04 00:27:00.000]]
Closed_Date: [[2023-03-13 08:21:11.000,2023-05-03 07:17:20.000,2023-03-21 21:16:53.000,2023-03-17 02:44:31.000,2023-04-19 00:00:00.000,...,null,2023-03-08 10:50:00.000,2023-03-24 23:18:11.000,2023-03-21 05:16:30.000,2023-03-04 10:15:00.000]]
Agency: [  -- dictionary:
["DOE","TLC","HPD","NYPD","DOB",...,"DHS","EDC","DCA","NYC311-PRD","OTI"]  -- indices:
[0,1,2,3,4,...,5,7,3,3,6]]
Agency_Name: [  -- dictionary:
["Department of Education","Taxi and Limousine Commission","Department of Housing Preservation and Development","New York City Police Department","Department of Buildings",...,"Department of Sanitation","Department of Homeless Services","Economic Development Corporation","Department of Consumer Affairs","Office of Technology and Innovation"]  -- indices:
[0,1,2,3,4,...,5,7,3,3,6]]
Complaint_Type: [  -- dictionary:
["School Maintenance","For Hire Vehicle Complaint","UNSANITARY CONDITION","Illegal Parking","Building/Use",...,"Building Drinking Water Tank","Taxi Licensee Complaint","Peeling Paint","Green Taxi Report","Bus Stop Shelter Placement"]  -- indices:
[0,1,2,3,4,...,30,80,3,7,16]]
Descriptor: [["Other School Condition","Driver Complaint - Non Passenger","PESTS","Posted Parking Sign Violation","Illegal Conversion Of Residential Building/Space",...,"For One Address","Lead Kit Request (Residential) (L10)","Commercial Overnight Parking","Loud Music/Party","Post"]]
Location_Type: [  -- dictionary:
["School","Street","RESIDENTIAL BUILDING","Street/Sidewalk",null,...,"Building Entrance","Public Stairs","Cafeteria - Public School","Commercial","Speed Reducer"]  -- indices:
[0,1,2,3,null,...,1,null,3,6,null]]
Incident_Zip: [[null,10075,10459,10455,11365,...,null,null,null,null,null]]
Incident_Address: [[null,"995 MADISON AVENUE","1010 BRYANT AVENUE","541 UNION AVENUE","69-35 166 STREET",...,"209 WEST   93 STREET","220 WEST   16 STREET","38 STERLING AVENUE","243 HENRY STREET",null]]
...
>>>
```

### 2023-06-17 nyc311clean.qmd
Add an example reading Arrow with polars

@@
 
-or Python (not that this format is called `feather` in the `pyarrow` package)
+or Python (note that this format is called `feather` in the `pyarrow` package)
 ```python
@@
 ...
+>>> import polars as pl
+>>> df = pl.read_ipc("./data/311_half_March_2023.arrow", use_pyarrow=True)
+>>> df
+shape: (125_826, 46)
+┌────────────┬────────────┬──────────────┬────────┬───┬────────────┬──────────────┬──────────────┬──────────────┐
+│ Unique_Key ┆ Created_Da ┆ Closed_Date  ┆ Agency ┆ … ┆ Community_ ┆ Borough_Boun ┆ City_Council ┆ Police_Preci │
+│ ---        ┆ te         ┆ ---          ┆ ---    ┆   ┆ Districts  ┆ daries       ┆ _Districts   ┆ ncts         │
+│ i32        ┆ ---        ┆ datetime[ms] ┆ cat    ┆   ┆ ---        ┆ ---          ┆ ---          ┆ ---          │
+│            ┆ datetime[m ┆              ┆        ┆   ┆ i8         ┆ i8           ┆ i8           ┆ i8           │
+│            ┆ s]         ┆              ┆        ┆   ┆            ┆              ┆              ┆              │
+╞════════════╪════════════╪══════════════╪════════╪═══╪════════════╪══════════════╪══════════════╪══════════════╡
+│ 57000688   ┆ 2023-03-09 ┆ 2023-03-13   ┆ DOE    ┆ … ┆ 2          ┆ 2            ┆ 27           ┆ 39           │
+│            ┆ 15:05:25   ┆ 08:21:11     ┆        ┆   ┆            ┆              ┆              ┆              │
+│ 56993837   ┆ 2023-03-08 ┆ 2023-05-03   ┆ TLC    ┆ … ┆ 23         ┆ 4            ┆ 51           ┆ 11           │
+│            ┆ 21:27:34   ┆ 07:17:20     ┆        ┆   ┆            ┆              ┆              ┆              │
+│ 57049479   ┆ 2023-03-15 ┆ 2023-03-21   ┆ HPD    ┆ … ┆ 8          ┆ 5            ┆ 43           ┆ 24           │
+│            ┆ 19:26:41   ┆ 21:16:53     ┆        ┆   ┆            ┆              ┆              ┆              │
+│ 57056381   ┆ 2023-03-16 ┆ 2023-03-17   ┆ NYPD   ┆ … ┆ 49         ┆ 5            ┆ 35           ┆ 23           │
+│            ┆ 23:58:53   ┆ 02:44:31     ┆        ┆   ┆            ┆              ┆              ┆              │
+│ …          ┆ …          ┆ …            ┆ …      ┆ … ┆ …          ┆ …            ┆ …            ┆ …            │
+│ 56954525   ┆ 2023-03-04 ┆ 2023-03-08   ┆ DEP    ┆ … ┆ 12         ┆ 4            ┆ 10           ┆ 6            │
+│            ┆ 16:23:00   ┆ 10:50:00     ┆        ┆   ┆            ┆              ┆              ┆              │
+│ 57132547   ┆ 2023-03-24 ┆ 2023-03-24   ┆ NYPD   ┆ … ┆ 30         ┆ 1            ┆ 14           ┆ 76           │
+│            ┆ 21:57:44   ┆ 23:18:11     ┆        ┆   ┆            ┆              ┆              ┆              │
+│ 57102740   ┆ 2023-03-21 ┆ 2023-03-21   ┆ NYPD   ┆ … ┆ 70         ┆ 4            ┆ 32           ┆ 4            │
+│            ┆ 05:10:12   ┆ 05:16:30     ┆        ┆   ┆            ┆              ┆              ┆              │
+│ 56950149   ┆ 2023-03-04 ┆ 2023-03-04   ┆ DOT    ┆ … ┆ 41         ┆ 3            ┆ 16           ┆ 61           │
+│            ┆ 00:27:00   ┆ 10:15:00     ┆        ┆   ┆            ┆              ┆              ┆              │
+└────────────┴────────────┴──────────────┴────────┴───┴────────────┴──────────────┴──────────────┴──────────────┘
 >>> 

### 2023-06-17 nyc311clean.qmd
Minor edits to clean up text

@@
 We now have enough information, such as the date-time format used, to do an initial conversion of the CSV file to a `DataFrame`, using several optional arguments to customize the process.
-(Details about each of these optional arguments are given callout blocks below.)
+(Details about each of these optional arguments are given in callout blocks below.)
 
@@
 
-Some columns are given are `String` whereas others are given as `String7` or `String15`.
+Some columns are described as `String` whereas others are described as `String7` or `String15`.
 These are storage-saving representations used when the whole column consists of short character strings.
@@
 
-The other problem shown in the types is that the `Incident_Zip` column is stored as strings when they should be integers.
+The other problem shown in the types is that the values in the `Incident_Zip` column are stored as strings when they should be integers or missing.
 
@@
 
-or Python (note that this format is called `feather` in the `pyarrow` package)
+or Python (note that the Arrow IPC format is called `feather` in the `pyarrow` package) or in [Polars](https://pola.rs)
+
 ```python

### 2023-06-17 nyc311clean.qmd
Even more tweaking.

--- 2023-06-17 nyc311clean.qmd
@@
     affiliation: University of Wisconsin - Madison
+jupyter: julia-1.9
+execute:
+  cache: true
+  freeze: auto
 ---
@@
 ::: {.callout-note collapse="false"}
-### Need to scroll right to see the entire line
+### Need to scroll right to see all of the output lines
 
@@
 
-and, of course, the problem is that someone got creative and used `na` as missing value indicator.
+and, of course, the problem is that someone got creative and used `na` as a missing value indicator.
 
@@
 
-These rows can be isolated for manual checking
+We save the indicators of rows with early closing or immediate closing for later manual checking.
 
@@
 ```{julia}
-first(df[earlyclosing, :], 3)
+earlyfrm = df[earlyclosing, :]
+first(earlyfrm, 3)
 ```
@@
 
+These three are from the "Department of Transport" and, in fact, all rows with early closing dates are from the DOT
+
+```{julia}
+unique(select(earlyfrm, :Agency, :Agency_Name))
+```
+
 ## Is Location redundant if Latitude and Longitude are given?
@@
 This is a tricky question because the values of `Latitude` and `Longitude` are stored to 17 significant digits, which is much more accuracy than could possibly be measured.
-To check for consistency we need to convert the `Location` to a `Latitude-Longitude` pair then compare these values to the recorded `Latitude` and `Longitude` with some tolerance.
+To check for consistency we need to convert the `Location` to a `Latitude-Longitude` pair then compare these values to the recorded `Latitude` and `Longitude` by employing a tolerance in the comparison of floating-point values.
+We'll leave that as an exercise for the reader and just assure you that the `Location` column is redundant and can be dropped.
 
@@
 #| output: false
-fn = Arrow.write("./data/311_half_March_2023.arrow", df)
+fn = Arrow.write(
+    "./data/311_half_Mar_2023.arrow",
+    select(df, Not(:Location)),
+)
 ```
 
-The Arrow file is less than half the size of the original CSV file but contains much more specificity about the data types and representations.
-Furthermore it can be read into other data science environments without needing to tinker with optional arguments.
+The Arrow file is less than one third the size of the original CSV file but contains much more specificity about the data types and representations.
+Furthermore it can be read into other data science environments without needing to tinker with arguments to CSV file-reading functions.
 
-The Arrow IPC format and the `Arrow.write` function in the Julia package allow for optional compression of the contents using `lz4` or `zstd` if further reduction of the file size is desired.
-The trade-off is that the uncompressed file can be memory-mapped by some systems to make for extremely fast input.
+The Arrow IPC format and the `Arrow.write` function in the Julia package allow for optional compression of the contents using `lz4` or `zstd` compression, if further reduction of the file size is desired.
+The downside of compression is that an uncompressed file can be memory-mapped, allowing for extremely fast input, whereas the compressed files must be decompressed into memory.
 
@@
 ```r
-> tibble::glimpse(arrow::read_ipc_file("./data/311_half_March_2023.arrow"))
+> tibble::glimpse(arrow::read_ipc_file("./data/311_half_Mar_2023.arrow"))
 Rows: 125,826
-Columns: 46
+Columns: 45
 $ Unique_Key                     <int> 57000688, 56993837, 57049479, 57056381,…
@@
 $ Longitude                      <dbl> -73.98284, -73.96297, -73.88857, -73.90…
-$ Location                       <chr> "(40.641663101476226, -73.9828374867858…
 $ Zip_Codes                      <int> 17620, 10092, 10937, 10933, 14507, 2052…
@@
 ### R applies a time zone to the datetime values
-The R `arrow` package appears to read the DateTime values as if they are in given in UTC then prints them in the local time zone.
+The R `arrow` package appears to read the DateTime values as if they are UTC values then prints them in the local time zone.
 :::
 
-or Python (not that this format is called `feather` in the `pyarrow` package)
+or in Python with `pyarrow` or with [polars](https://pola.rs).
+(Note that the Arrow IPC format is called `feather` in the `pyarrow` package.)
 ```python
 >>> import pyarrow.feather as fea
->>> fea.read_table("./data/311_half_March_2023.arrow")
+>>> import polars as pl
+>>> fea.read_table('./data/311_half_Mar_2023.arrow')
 pyarrow.Table
@@
 Longitude: double
-Location: string
 Zip_Codes: int16
@@
 ...
->>>
+>>> pl.read_ipc('./data/311_half_Mar_2023.arrow', use_pyarrow=True)
+shape: (125_826, 45)
+┌──────────┬────────────┬───────────┬────────┬───┬────────────┬────────────┬────────────┬────────────┐
+│ Unique_K ┆ Created_Da ┆ Closed_Da ┆ Agency ┆ … ┆ Community_ ┆ Borough_Bo ┆ City_Counc ┆ Police_Pre │
+│ ey       ┆ te         ┆ te        ┆ ---    ┆   ┆ Districts  ┆ undaries   ┆ il_Distric ┆ cincts     │
+│ ---      ┆ ---        ┆ ---       ┆ cat    ┆   ┆ ---        ┆ ---        ┆ ts         ┆ ---        │
+│ i32      ┆ datetime[m ┆ datetime[ ┆        ┆   ┆ i8         ┆ i8         ┆ ---        ┆ i8         │
+│          ┆ s]         ┆ ms]       ┆        ┆   ┆            ┆            ┆ i8         ┆            │
+╞══════════╪════════════╪═══════════╪════════╪═══╪════════════╪════════════╪════════════╪════════════╡
+│ 57000688 ┆ 2023-03-09 ┆ 2023-03-1 ┆ DOE    ┆ … ┆ 2          ┆ 2          ┆ 27         ┆ 39         │
+│          ┆ 15:05:25   ┆ 3         ┆        ┆   ┆            ┆            ┆            ┆            │
+│          ┆            ┆ 08:21:11  ┆        ┆   ┆            ┆            ┆            ┆            │
+│ 56993837 ┆ 2023-03-08 ┆ 2023-05-0 ┆ TLC    ┆ … ┆ 23         ┆ 4          ┆ 51         ┆ 11         │
+│          ┆ 21:27:34   ┆ 3         ┆        ┆   ┆            ┆            ┆            ┆            │
+│          ┆            ┆ 07:17:20  ┆        ┆   ┆            ┆            ┆            ┆            │
+│ 57049479 ┆ 2023-03-15 ┆ 2023-03-2 ┆ HPD    ┆ … ┆ 8          ┆ 5          ┆ 43         ┆ 24         │
+│          ┆ 19:26:41   ┆ 1         ┆        ┆   ┆            ┆            ┆            ┆            │
+│          ┆            ┆ 21:16:53  ┆        ┆   ┆            ┆            ┆            ┆            │
+│ 57056381 ┆ 2023-03-16 ┆ 2023-03-1 ┆ NYPD   ┆ … ┆ 49         ┆ 5          ┆ 35         ┆ 23         │
+│          ┆ 23:58:53   ┆ 7         ┆        ┆   ┆            ┆            ┆            ┆            │
+│          ┆            ┆ 02:44:31  ┆        ┆   ┆            ┆            ┆            ┆            │
+│ …        ┆ …          ┆ …         ┆ …      ┆ … ┆ …          ┆ …          ┆ …          ┆ …          │
+│ 56954525 ┆ 2023-03-04 ┆ 2023-03-0 ┆ DEP    ┆ … ┆ 12         ┆ 4          ┆ 10         ┆ 6          │
+│          ┆ 16:23:00   ┆ 8         ┆        ┆   ┆            ┆            ┆            ┆            │
+│          ┆            ┆ 10:50:00  ┆        ┆   ┆            ┆            ┆            ┆            │
+│ 57132547 ┆ 2023-03-24 ┆ 2023-03-2 ┆ NYPD   ┆ … ┆ 30         ┆ 1          ┆ 14         ┆ 76         │
+│          ┆ 21:57:44   ┆ 4         ┆        ┆   ┆            ┆            ┆            ┆            │
+│          ┆            ┆ 23:18:11  ┆        ┆   ┆            ┆            ┆            ┆            │
+│ 57102740 ┆ 2023-03-21 ┆ 2023-03-2 ┆ NYPD   ┆ … ┆ 70         ┆ 4          ┆ 32         ┆ 4          │
+│          ┆ 05:10:12   ┆ 1         ┆        ┆   ┆            ┆            ┆            ┆            │
+│          ┆            ┆ 05:16:30  ┆        ┆   ┆            ┆            ┆            ┆            │
+│ 56950149 ┆ 2023-03-04 ┆ 2023-03-0 ┆ DOT    ┆ … ┆ 41         ┆ 3          ┆ 16         ┆ 61         │
+│          ┆ 00:27:00   ┆ 4         ┆        ┆   ┆            ┆            ┆            ┆            │
+│          ┆            ┆ 10:15:00  ┆        ┆   ┆            ┆            ┆            ┆            │
+└──────────┴────────────┴───────────┴────────┴───┴────────────┴────────────┴────────────┴────────────┘
 ```
+
+The columns that are described as `<fct>` in the `glimpse` output from R are characterized as `dictionary` in the output from `fea.read_table` and listed as `cat` (for "categorical") in the Polars data frame.
+The conversion of a column of character strings, many of which are repetitions, to a combination of a vector of the unique values and a small-sized integer index into that vector, can result in considerable storage saving, at the expense of a trivial amount of processing.
+For example, storing the `Agency_Name` as a vector of strings takes about 34 MB but, after dictionary encoding, it takes up about 1/8 MB.

### 2023-06-18 nyc311clean.qmd
Clean up some of the explanations.

@@
 
-We show conversion of a CSV file containing data on 311 service calls during March, 2023 in New York City to the [Arrow](https://arrow.apache.org) IPC (Inter-Process Communications) file format (also called the Feather V2 format) which can be read quickly and easily into many different data science environments.
-This conversion is done in [Julia](https://julialang.org) but it could equally well have been done in a number of different languages, whichever is most convenient for the person doing the conversion.
+We show conversion of a CSV file containing data on 311 service calls in New York City during March, 2023 to the [Arrow](https://arrow.apache.org) IPC (Inter-Process Communication) file format (also called the Feather V2 format), which can be read quickly and easily into several different data science environments.
+This conversion is done in [Julia](https://julialang.org) but it could equally well have been done in a different language, if that is more convenient for the person doing the conversion.
 
@@
 #| panel: fill
-open("./data/311_half_Mar_2023.csv", "r") do io
-    println(readline(io))
-    println(readline(io))
-    seekstart(io)
-end;
+open("./data/311_half_Mar_2023.csv", "r") do io # open the file
+    println(readline(io))     # read and print the first line
+    println(readline(io))     # read and print the next line
+end;  # this open/do/end construction automatically closes the file when done
 ```
@@
 ::: {.callout-note collapse="false"}
-### Need to scroll right to see all of the output lines
+### Need to scroll right to see the whole output line
 
@@
 
-Note that the character on the right of the title bar for callout blocks like this is a toggle to show or hide the contents of the block.
+Note that the character on the right of the title bar for callout blocks like this, currently a `v`-shape for this block, is a toggle to show or collapse the contents of the block.
+Try clicking on that character to collapse the contents of the block.
+
+In what follows the callout blocks are initially collapsed and you must click on the `>` character on the right end of the title bar to show the contents.
 :::
@@
     "./data/311_half_Mar_2023.csv",  # file name
-    DataFrame;                       # output table type
+    DataFrame,           # output table type
     normalizenames=true, # convert column names to valid symbols
-    missingstring=[      # strings that are used to indicate missing vals
+    missingstring=[      # strings that are used to indicate missing values
         "",
@@
 :::{.callout-note collapse="true"}
-### normalize names
-Convert the column names in the first row of the CSV file into a form suitable as a `Symbol` in `Julia`.
+### normalizenames
+Convert the column names from the CSV file into a form that can be parsed as a `Symbol` in `Julia`.
 In particular, strip whitespace at the beginning and end of the name and replace embedded blanks or `.` characters by underscores.
+
+Columns of a DataFrame (and other type of tables) are usually extracted as "properties" of the table, either as, e.g.
+`getproperty(df, :Unique_Key)`, or, more commonly, with the dot operator as, e.g.
+`df.Unique_Key`.
+The expression `:Unique_Key` or the right operand of `df.Unique_Key` evaluates to a Symbol.
+If the column name has an embedded blank or `.` this form cannot be used and the name must be quoted, either as `Symbol("Unique.Key")` or `var"Unique.Key"`.
+
+Thus it is an advantage to rewrite the column names so that each name can be used "as is" with the dot operator.
 :::
@@
 ### downcast
-For columns of integer values, choose the smallest type that can represent all the values in the column.
+For columns of integer values, choose the smallest integer type, such as `Int8`, `Int16`, etc., that can represent all the values in the column.
 :::
@@
 
-In the output (you must scroll horizontally to see it all) under the column name is an indicator of the type.
+In the output (remember, you must scroll horizontally to see it all) there is a description of the type of the elements of the column directly under the column name.
 A type name that ends in `?` indicates that missing values are present in the column.
-Thus the `Unique_Key` is stored as 32-bit signed integers with no missing values.
-The `Created_Date` is a `DateTime` with no missing values but the `Closed_Date`, which is also a `DateTime`, does have missing values.
+Thus, for example, the `Unique_Key` is designated `Int32` meaning the values are stored as 32-bit signed integers with no missing values.
+The `Created_Date` column, designated `DateTime`, is a `DateTime` column with no missing values but the `Closed_Date`, designated `DateTime?`, is also a `DateTime`, but can (and does) have missing values.
 
-Some columns are described as `String` whereas others are described as `String7` or `String15`.
+Some columns are designated as `String` whereas others are designated as `String7` or `String15`.
 These are storage-saving representations used when the whole column consists of short character strings.
+
 There is an anomaly in these short strings for the `Agency` column.
-The `Agency` acronyms should be 4 or 5 characters at most but at least one (it happens to be `NYC311-PRD`) must be greater than 7 characters to force `String15` storage.
+The `Agency` acronyms should be 5 characters at most, which would be stored as `String7`, but they are stored as `String15`.
 
-```{julia}
-show(unique(df.Agency))
-```
-
-The `Agency` and `Agency_Name` columns should be in one-to-one correspondence but they are not.
+To see why, we check the unique combinations of the `Agency` and `Agency_Name` columns
 
@@
 
+These columns should be in one-to-one correspondence but they are not.
+
 At this point we would need to check with the data providers to determine why two different `Agency` designations are used for the "Department of Environmental Protection".
 
-The other problem shown in the types is that the values in the `Incident_Zip` column are stored as strings when they should be integers or missing.
+The other problem shown in the types is that the values in the `Incident_Zip` column are stored as `String7?` when they should be `Int32?`.
+
+Checking the unique values
 
 ```{julia}
-levels(df.Incident_Zip)
+sort(unique(df.Incident_Zip))
 ```
 
-and, of course, the problem is that someone got creative and used `na` as a missing value indicator.
+shows the problem is that someone got creative and used `na` as a missing value indicator.
 
@@
     "./data/311_half_Mar_2023.csv",  # file name
-    DataFrame;                       # output table type
+    DataFrame,           # output table type
     normalizenames=true, # convert column names to valid symbols
@@
 
-### Are the `Created_Date` values in the specified range?
+### Are the `Created_Date` values all from March, 2023?
 
@@
 ```{julia}
-earlyclosing = (!ismissing).(df.Closed_Date) .&& (df.Closed_Date .< df.Created_Date)
-immediateclosing = (!ismissing).(df.Closed_Date) .&& (df.Closed_Date .== df.Created_Date)
+earlyclosing = @. !ismissing(df.Closed_Date) && (df.Closed_Date < df.Created_Date)
+immediateclosing = @. !ismissing(df.Closed_Date) && (df.Closed_Date == df.Created_Date)
 (early = count(earlyclosing), immediate = count(immediateclosing))
@@
 :::{.callout-note collapse="true"}
-### What are all those dots for?
-Julia provides "syntactic sugar" to vectorize a scalar operation with a `.` following a function name (e.g. `(!ismissing).`) or preceding an operator (e.g. `.<`)
+### What does `@.` do?
+
+[Dot vectorization](https://docs.julialang.org/en/v1/manual/functions/#man-vectorized) in Julia provides "syntactic sugar" to vectorize a scalar operation.
+Prefacing an expression with `@.` applies dot vectorization to all the function calls and operators in the expression. 
 :::
@@
 :::{.callout-note collapse="true"}
-### Use of `:` as an index indicates all values for that index
+### Use of `:` as an index indicates all values for that axis
 In this case the use of `:` as the second index indicates that all the columns should be included in the subset.
@@
 
-This is a tricky question because the values of `Latitude` and `Longitude` are stored to 17 significant digits, which is much more accuracy than could possibly be measured.
-To check for consistency we need to convert the `Location` to a `Latitude-Longitude` pair then compare these values to the recorded `Latitude` and `Longitude` by employing a tolerance in the comparison of floating-point values.
+This is a tricky question because the values of `Latitude` and `Longitude` are stored to 17 significant digits, which is much more accuracy than could possibly be measured, but equality comparisons, either as floating-point numbers or as strings with 17 significant digits, would require an exact match.
+
+To check for consistency we need to convert the `Location` to a `Latitude-Longitude` pair of floating point numbers then check for approximate equality of these values to the recorded `Latitude` and `Longitude`.
 We'll leave that as an exercise for the reader and just assure you that the `Location` column is redundant and can be dropped.
@@
 #| output: false
-fn = Arrow.write(
+fn = Arrow.write(    # returns the file name
     "./data/311_half_Mar_2023.arrow",
@@
 
-The columns that are described as `<fct>` in the `glimpse` output from R are characterized as `dictionary` in the output from `fea.read_table` and listed as `cat` (for "categorical") in the Polars data frame.
-The conversion of a column of character strings, many of which are repetitions, to a combination of a vector of the unique values and a small-sized integer index into that vector, can result in considerable storage saving, at the expense of a trivial amount of processing.
+The columns that are described as `<fct>` in the `glimpse` output from R (and `dictionary` in the output from `fea.read_table` and `cat` (for "categorical") in the Polars data frame) are `factor` representations, consisting of a (short) vector of the unique values, called the `levels` of the factor, and an integer vector of indices into the levels.
+The conversion of a column of character strings, many of which are repetitions, to a
+factor or "dictionary encoded" representation, can result in considerable storage saving, at the expense of a trivial amount of processing.
 For example, storing the `Agency_Name` as a vector of strings takes about 34 MB but, after dictionary encoding, it takes up about 1/8 MB.
+
+The output from `fea.read_table` is often the most informative representation of the structure of the table.

### 2023-07-05 manuscript/main.Rnw

--- 2023-03-17 manuscript/main.tex
@@
 
-%% Make sure every package in preamble is used
-\usepackage{siunitx} % for unit
-\usepackage[margin=1in]{geometry}
+\usepackage[margin = 1in]{geometry}
 \usepackage{amsmath}
-\numberwithin{equation}{section}
-\usepackage{graphicx}
+\allowdisplaybreaks
+\usepackage{authblk}
+\usepackage{setspace}
 \usepackage{natbib}
-\usepackage{underscore}
-\usepackage{physics}
-\usepackage[ruled,vlined]{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{booktabs}
-\usepackage{amsfonts}
-\usepackage{bm}
-\usepackage{threeparttable}
-\usepackage{setspace}
-\usepackage[width=.95\textwidth, skip=0.1\baselineskip]{caption}
+\usepackage{hyperref}\hypersetup{colorlinks=true, citecolor=blue}
 
-\allowdisplaybreaks
-
-% \usepackage{adjustbox}
-\usepackage[colorlinks, citecolor=blue,  urlcolor=blue,
-linkcolor=blue]{hyperref}
-\usepackage{xcolor}
-
-% \usepackage[pagewise]{lineno}
 \usepackage[]{lineno}
@@
 
-\allowdisplaybreaks % \usepackage{rotating}
-
 \newcommand{\jy}[1]{\textcolor{orange}{JY: (#1)}}
@@
 \let\proglang=\textsf
-\newcommand{\pkg}[1]{{\fontseries{m}\selectfont #1}}
-\newcommand\code[2][black]{\textcolor{#1}{\texttt{#2}}}
-
-\usepackage{multirow, makecell}
-\usepackage{booktabs,array}
-\usepackage[affil-it]{authblk}
-\renewcommand\Authfont{\fontsize{12}{14.4}\selectfont}
-\renewcommand\Affilfont{\fontsize{12}{10.8}\itshape}
-% \AddToHook{cmd/section/before}{\clearpage} % will remove
+%% \newcommand{\pkg}[1]{{\fontseries{m}\selectfont #1}}
+%% \newcommand\code[2][black]{\textcolor{#1}{\texttt{#2}}}
 
@@
 
-\doublespacing
+% \doublespacing
+
+<<prep, echo=FALSE, cache=FALSE>>=
+library(knitr)
+## set font size
+def.chunk.hook  <- knitr::knit_hooks$get("chunk")
+knitr::knit_hooks$set(chunk = function(x, options) {
+  x <- def.chunk.hook(x, options)
+  paste0("\n \\", "small","\n\n", x, "\n\n \\normalsize")
+})
+## read in all the chunks; so only need to maintain one place
+read_chunk("../code/data_cleansing_311_V4.R")
+@
 
@@
 
-Open data is the trend ... many kinds \citep[e.g.,][]{cdc2022} ...
+Open data is the trend ... many kinds \citep{barns2016mine}
 
@@
 \section{NYC 311 Requests Data} \label{sec:data}
+
+We need to make the variable names user friendly:
+<<makeColNamesUserFriendly>>=
+
+@
 

### 2024-01-04 manuscript/main.tex
use tex instead of Rnw
tentative title

@@
 \let\proglang=\textsf
-\newcommand{\pkg}[1]{{\fontseries{m}\selectfont #1}}
-\newcommand\code[2][black]{\textcolor{#1}{\texttt{#2}}}
+%% \newcommand{\pkg}[1]{{\fontseries{m}\selectfont #1}}
+%% \newcommand\code[2][black]{\textcolor{#1}{\texttt{#2}}}
 
-\usepackage{multirow, makecell}
-\usepackage{booktabs,array}
-\usepackage[affil-it]{authblk}
-\renewcommand\Authfont{\fontsize{12}{14.4}\selectfont}
-\renewcommand\Affilfont{\fontsize{12}{10.8}\itshape}
-% \AddToHook{cmd/section/before}{\clearpage} % will remove
-
-\title{Tentative: On Open Data Releasing Protocols}
-\date{}
+\title{Principles for Open Data Curation: A Case Study with the New
+  York City 311 Service Request Data}
 
 \author[1]{David Tussey}
-\author[2]{Jun Yan\thanks{Tentatively in alphabetical order}}
-\affil[1]{Independent Researcher?}
+\author[2]{Jun Yan}
+\affil[1]{Former Executive Director, NYC DoITT}
 \affil[2]{Department of Statistics, University of Connecticut}
@@
 \section{NYC 311 Requests Data} \label{sec:data}
+
+We need to make the variable names user friendly:
 

### 2024-01-09 manuscript/main.tex
first draft for intro and suggested structure for data section

@@
 
-Open data is the trend ... many kinds \citep{barns2016mine}
+In the early 21st century, the open data movement began to take shape,
+driven by the fundamental belief that freely accessible data can
+transform societies. This movement champions the principles of
+transparency, innovation, and public engagement. A landmark in this
+journey was the launch of the United States'
+\href{https://www.data.gov}{Data.gov} portal in 2009, a pioneering
+platform in making government data widely accessible. Shortly after,
+the European Union followed suit, unveiling its
+\href{https://data.europa.eu/euodp}{Open Data Portal} in 2012, further
+cementing the movement's global reach. These initiatives represent
+significant strides in democratizing data, breaking barriers that once
+kept valuable information in silos. Furthermore, the World Bank's Open
+Data initiative, initiated in 2010, stands out as a comprehensive
+repository for global development data, available at
+\href{https://data.worldbank.org}{World Bank Open Data}. The
+collective impact of these platforms is profound, extending beyond
+mere data sharing to fostering a culture of openness that benefits
+individuals, communities, and economies worldwide
+\citep{janssen2012benefits, barns2016mine, wang2016adoption}.
 
-NYC is a forerunner in open data ... The 311 requests data is one of the most
-popular open data ...
+New York City (NYC) has emerged as a forerunner in the open data
+movement, marked by the enactment of the Open Data Law in 2012
+\citep{zuiderwijk2014open}. This landmark legislation led to the
+creation of the \href{https://opendata.cityofnewyork.us}{NYC Open Data
+  portal}, which today hosts an impressive array of 3,511 datasets
+from 98 different city agencies. This resource has become invaluable
+for researchers in various fields. In health, datasets have enabled
+significant studies on facets of health and healthcare delivery
+\citep{cantor2018facets, shankar2021data}. In the realm of urban
+development, data has been instrumental in advancing smart city
+initiatives \citep{neves2020impacts}. Additionally, transportation
+research has benefited greatly from this wealth of data, aiding in the
+understanding of urban mobility and infrastructure
+\citep{gerte2019understanding}. NYC's Open Data initiative not only
+exemplifies commitment to transparency and public engagement but also
+illustrates how open data can be a powerful tool in addressing complex
+urban challenges.
 
-A clean, efficiently designed data is important ...
+Data curation is fundamental in the open data ecosystem, ensuring the
+utility and reliability of datasets for diverse applications. Among
+the earliest discussions, \citet{witt2009constructing} focus on the
+development of data curation profiles tailored to specific contexts,
+setting a precedent for targeted data management
+strategies. Addressing broader challenges in data sharing and
+management, \citet{borgman2012conundrum} highlights the complexities
+of research data distribution, emphasizing the need for robust
+strategies. This is complemented by the work of \citet{hart2016ten},
+who outline essential principles for effective data management,
+particularly emphasizing the importance of meticulous curation
+practices. In the realm of collaborative data management,
+\citet{beheshti2019datasynapse} underscore the significance of
+cooperative environments for managing and sharing social data
+effectively. This aspect of data curation gains further relevance in
+the research by \citet{mclure2014data}, which delves into the specific
+practices and needs within data curation communities. The practical
+implications of data curation are vividly illustrated in the context
+of public health and global challenges. \citet{cantor2018facets}
+demonstrate the utility of curated open data in evaluating community
+health determinants. Furthermore, the COVID-19 pandemic serves as a
+real-world example, with \citet{shankar2021data} observing the
+critical role of collective data curation efforts in managing and
+responding to the crisis. Collectively, these studies not only
+highlight the multifaceted nature of data curation but also emphasize
+its indispensable role in enhancing the applicability and value of
+open data across various domains.
 
-The rest of the paper is organized as follows.
-Section~\ref{sec:data} presents the history and current status of the 311
-requests data. Section~\ref{sec:issues} lists potential issues on data quality
-and efficiency.
-Section~\ref{sec:improve} provides suggestions on how to remove or minimize the
-issues.
-Section~\ref{sec:protocol} summarizes a few general principals on open data
-release.
-A discussion concludes in Section~\ref{sec:disc}
+The contributions of this paper are twofold. Firstly, we delve into
+the specifics of data curation challenges using the NYC 311 Service
+Request Data as a case study. This renowned open dataset serves as a
+prime example for examining key issues in data curation, including
+data validity, consistency, and curation efficiency. We illustrate
+these points with live examples drawn from our processing of the 311
+data. Secondly, building upon insights gained from this case study, we
+propose a set of data curation principles tailored for
+government-released open data. These principles are designed to
+address the unique challenges and requirements observed in the
+curation of such datasets.
+
+The remainder of this paper is organized as
+follows. Section~\ref{sec:data} offers an in-depth review of the
+history and current status of the 311 requests data. In
+Section~\ref{sec:issues}, we identify and list potential issues
+impacting data quality and curation
+efficiency. Section~\ref{sec:improve} is dedicated to providing
+actionable suggestions for mitigating or resolving these identified
+issues. Following this, Section~\ref{sec:protocol} outlines a series
+of general principles for the release of open data, drawing from our
+findings. The paper concludes with a discussion in
+Section~\ref{sec:disc}, encapsulating the key insights and
+implications of our research.
 
@@
 
-We need to make the variable names user friendly:
+The NYC 311 service, a critical component of New York City's public
+engagement and service response framework, serves as a centralized hub
+for non-emergency inquiries and requests. Introduced in 2010, the NYC
+311 system was designed to streamline the city's response to
+non-emergency issues, ranging from noise complaints to street
+maintenance requests. The evolution of this system can be traced from
+its initial implementation as a simple inquiry channel to a
+comprehensive data management system that handles millions of requests
+annually. Key milestones in its development include [list any
+significant changes or technological upgrades].
+
+Today, the NYC 311 data system is a robust platform that manages a
+vast array of urban living-related inquiries. As of [current year or
+most recent data year], the system handles approximately [number]
+requests per year, covering categories like [list a few
+categories]. The data, managed through [describe the current data
+management system or software], is a valuable resource for city
+administration and policy-making. It is publicly accessible through
+[provide details of the data access platform, e.g., an online
+portal]. This open data initiative enables not only governmental
+transparency but also empowers researchers, civic developers, and the
+general public. Examples of its use include [cite specific projects or
+research that used this data].
+
+Despite its success, the system faces challenges such as [list
+challenges like data quality, privacy issues, or technological
+constraints]. For instance, issues with [specific challenge] have
+implications for [discuss the implications briefly].
+The impact of the NYC 311 data extends beyond operational efficiency;
+it has become instrumental in shaping city governance and community
+engagement. The data has been pivotal in [give examples of policy
+changes, community impacts, or improvements in city services].
+The NYC 311 service exemplifies the dynamic nature of urban data
+management and its critical role in modern governance. It provides an
+invaluable resource for data scientists to practice data science
+activities, including data curation.
+
+We extracted [give the data period; extraction time] for this case
+study. [justify the choices.]
 

### 2024-05-21 manuscript/311 manuscript.tex
Console output for latest code revisions

\documentclass[12pt, titlepage]{article}

\usepackage[margin = 1in]{geometry}
\usepackage{amsmath}
\allowdisplaybreaks
\usepackage{authblk}
\usepackage{setspace}
\usepackage{natbib}
\usepackage{hyperref}\hypersetup{colorlinks=true, citecolor=blue}
\usepackage{parskip}
\usepackage{textgreek}
\usepackage{graphicx}
\usepackage{booktabs} % For better looking tables
\usepackage{siunitx} % For alignment of numbers
\usepackage{placeins}
\usepackage{caption}  % For custom captions

\usepackage[]{lineno}
\linenumbers*[1]
%% patches to make lineno work better with amsmath
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
	\expandafter\let\csname old#1\expandafter\endcsname\csname
	#1\endcsname
	\expandafter\let\csname oldend#1\expandafter\endcsname\csname
	end#1\endcsname
	\renewenvironment{#1}%
	{\linenomath\csname old#1\endcsname}%
	{\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
	\patchAmsMathEnvironmentForLineno{#1}%
	\patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
	\patchBothAmsMathEnvironmentsForLineno{equation}%
	\patchBothAmsMathEnvironmentsForLineno{align}%
	\patchBothAmsMathEnvironmentsForLineno{flalign}%
	\patchBothAmsMathEnvironmentsForLineno{alignat}%
	\patchBothAmsMathEnvironmentsForLineno{gather}%
	\patchBothAmsMathEnvironmentsForLineno{multline}%
}

\sisetup{
  group-separator = {,},
  round-mode = places,
  round-precision = 0
}

% control floats
\renewcommand\floatpagefraction{.9}
\renewcommand\topfraction{.9}
\renewcommand\bottomfraction{.9}
\renewcommand\textfraction{.2}
\setcounter{totalnumber}{50}
\setcounter{topnumber}{50}
\setcounter{bottomnumber}{50}

\newcommand{\jy}[1]{\textcolor{orange}{JY: (#1)}}
\newcommand{\yh}[1]{\textcolor{purple}{YH: (#1)}}
\newcommand{\yx}[1]{\textcolor{cyan}{YX: (#1)}}
\let\proglang=\textsf
%% \newcommand{\pkg}[1]{{\fontseries{m}\selectfont #1}}
%% \newcommand\code[2][black]{\textcolor{#1}{\texttt{#2}}}

\title{Principles for Open Data Curation: A Case Study with the New
  York City 311 Service Request Data}

\author[1]{David Tussey}
\author[2]{Jun Yan}
\affil[1]{Former Executive Director, NYC DoITT}
\affil[2]{Department of Statistics, University of Connecticut}

\begin{document}
\maketitle

\tableofcontents % Optional: Table of Contents
\listoffigures % List of Figures
\listoftables % List of Tables

\begin{abstract}
  Open data is important ...
  Quality control is important ...
  We use New York City 311 requests data  ...
  We suggest some open data quality control protocols ...

\bigskip

\noindent
{\it Keywords:}
consistency check;    
data release protocol;
data science;
open data;
data cleansing;
quality control	
\end{abstract}

\doublespacing

\section{Introduction} \label{sec:intro}

In the early 21st century, the open data movement began to take shape, driven by the 
fundamental belief that freely accessible data can transform 
both societies and governments. This movement champions the principles of 
transparency, innovation, and public engagement.

A landmark in this journey was the launch of the United States'
\href{https://www.data.gov}{Data.gov} portal in 2009, a pioneering
platform in making government data widely accessible. Shortly after,
the European Union followed suit, unveiling its
\href{https://data.europa.eu/euodp}{Open Data Portal} in 2012, further
cementing the movement's global reach. Furthermore, the World Bank's Open
Data initiative, initiated in 2010, stands out as a comprehensive
repository for global development data, available at
\href{https://data.worldbank.org}{World Bank Open Data}.

These initiatives represent significant strides in democratizing data, in breaking barriers that once
kept valuable information on government performance in silos. Their collective impact is profound, extending beyond
mere data sharing to fostering a culture of openness that benefits
individuals, communities, governments, and economies worldwide
\citep{barns2016mine, wang2016adoption}.

New York City (NYC) has emerged as a forerunner in the open data
movement, marked by the enactment of the Open Data Law in 2012
\citep{zuiderwijk2014open}. This landmark legislation led to the
creation of the \href{https://opendata.cityofnewyork.us}{NYC Open Data
  portal}, which today hosts an impressive array of 2.700 datasets
from 80 different city agencies. This resource has become invaluable
for researchers in various fields as well an enabling a high degree of local government transparency

In health, datasets have enabled
significant studies on facets of health and healthcare delivery
\citep{cantor2018facets, shankar2021data}. In the realm of urban
development, data has been instrumental in advancing smart city
initiatives \citep{neves2020impacts}. Additionally, transportation
research has benefited greatly from this wealth of data, aiding in the
understanding of urban mobility and infrastructure
\citep{gerte2019understanding}.

NYC's Open Data initiative not only
exemplifies commitment to transparency and public engagement but also
illustrates how open data can be a powerful tool in addressing complex
urban challenges.

Data curation is fundamental in the open data ecosystem, ensuring the
utility and reliability of datasets for diverse applications. Among
the earliest discussions, \citet{witt2009constructing} focus on the
development of data curation profiles tailored to specific contexts,
setting a precedent for targeted data management
strategies. Addressing broader challenges in data sharing and
management, \citet{borgman2012conundrum} highlights the complexities
of research data distribution, emphasizing the need for robust
strategies. This is complemented by the work of \citet{hart2016ten},
who outline essential principles for effective data management,
particularly emphasizing the importance of meticulous curation
practices.

In the realm of collaborative data management,
\citet{beheshti2019datasynapse} underscore the significance of
cooperative environments for managing and sharing social data
effectively. This aspect of data curation gains further relevance in
the research by \citet{mclure2014data}, which delves into the specific
practices and needs within data curation communities. The practical
implications of data curation are vividly illustrated in the context
of public health and global challenges. \citet{cantor2018facets}
demonstrate the utility of curated open data in evaluating community
health determinants. Furthermore, the COVID-19 pandemic serves as a
real-world example, with \citet{shankar2021data} observing the
critical role of collective data curation efforts in managing and
responding to the crisis.

Collectively, these studies not only highlight the multifaceted nature of data curation but also emphasize
its indispensable role in enhancing the applicability and value of
open data across various domains.

The contributions of this paper are twofold. First, we delve into
the specifics of data curation challenges using the NYC 311 Service
Request Data as a case study. This renowned and frequently viewed 
dataset serves as a prime example for examining key issues in data curation, 
including data validity, consistency, and curation efficiency. 
We illustrate these points with live examples drawn from our processing 
of the 311 data.

Secondly, building upon insights gained from this case study, we 
propose a set of data curation principles tailored for government-released open data. 
These principles are designed to address the unique challenges 
and requirements observed in the curation of such datasets.

The remainder of this paper is organized as follows. 
Section~\ref{sec:data} offers a brie review of the 
history and current status of the 311 requests data. In 
Section~\ref{sec:issues}, we identify and list potential issues 
impacting data quality and curation efficiency. 
Section~\ref{sec:improve} is dedicated to providing 
actionable suggestions for mitigating or resolving these identified issue. 
Following this, Section~\ref{sec:protocol} outlines a series of general 
principles for the release of open data, drawing from our findings. 
The paper concludes with a discussion in Section~\ref{sec:disc}, encapsulating 
the key insights and implications of our research.

\section{NYC 311 Service Requests Data} \label{sec:data}

The NYC 311 service, a critical component of New York City's public
engagement and service response framework, serves as a centralized hub
for non-emergency inquiries and requests. Introduced in 2003, the NYC
311 system was designed to streamline the city's response to
non-emergency issues, ranging from noise complaints to street
maintenance requests. The evolution of this system can be traced from
its initial implementation as a simple inquiry channel to a
comprehensive data management system that handles millions of requests
annually. Key milestones in its development include:

\begin{itemize}
 	\item 2003 - NYC 311 system  goes live with phone-based call center only.
   	\item 2009 - 311 online \& mobile apps launched. Today 60\% of all Service Requests come from mobile \& online channels.
   	\item 2019 - Major software upgrade of 311 system with enhanced capabilities and cloud-based resilience. 	
      	\item 2020 - In August, driven by the COVID pandemic, monthly 311 Service Requests hit an all-time high of 348,463. Seven of the Top 10 busiest days occur in 2020.
      	\item 2021 - 311 System is expanded to into the MTA's city subway system, the largest expansion since its inception.
   	\item 2023 - Record high year for 311 with 3.23 million Service Requests.
\end{itemize}

Today, the NYC 311 data system is a robust platform that manages a
vast array of urban living-related inquiries. As of 2024, the system handles approximately 3.5 million service 
requests per year, covering categories such as noise complaints, illegal parking, heat/hot water complaints, abandoned vehicles,
and unsanitary conditions. The data, managed through a tailored application built on the Microsoft Dynamics platform, is a valuable resource for city
administration and policy-making. It is publicly accessible through the 
\href{https://opendata.cityofnewyork.us/}{NYC Open Data Portal}. The online Portal provides access
to all the data sets (approx. 2700) and provides a convenient method for querying, grouping, aggregating, geo-mapping, visualization, and exporting
results. This open data initiative enables not only governmental
transparency but also empowers researchers, civic developers, and the
general public. Examples of its use include:

\begin{itemize}
	\item NYC restaurant violations
	\item Popular baby names
	\item Mapping of car crashes involving pedestirans
	\item Mapping of sidewalk widths in NYC 
	\item Visualization of NYC High School and College enrollment
	\item Interactive dashboard to filter and view charges for jail inmates
	\item Internet access points.
	\item ...and 220 different complaint types. 
\end{itemize}

Additional Open Data projects can be found at \href{https://opendata.cityofnewyork.us/projects/}{NYC Open Data Project Gallery}
and at \href{https://beta.nyc/beta/products/}{\textbeta etaNYC} products and tools.

Despite its success, the system faces many challenges such as:
\begin{itemize}
	\item Data timeliness, accuracy, and consistency
	\item Difficulties correlating data over long time periods, e.g. over a 10-year period
	\item Data anonymization and handling Personally identifiable information (PII)
	\item Integration with stand-alone systems at selected NYC Agencies
	\item Managing API usage, authentication, and load for 3rd party users
	\item Incorporating ongoing technology changes and upgrades 
	\item Understanding the role of chatbots and AI to provide online customer service
\end{itemize}

These (and other) challenges are constantly being addressed by the various agency open data managers
as well as the \href{https://www.nyc.gov/content/oti/pages/}{NYC Office of Technology and Innovation} which
provides the technology support for the open data system.  
Many of the underlying datasets are updated daily, usually running 1-2 days behind from creation.
However, a number of datasets are updates monthly (such as restaurant inspections), and some are
not updated for one or more years (such as community board boundaries). This poses problems when 
merging these dataset with one another, as would be expected. Typically, the data must be
exported and manipulated in another program

Data accuracy and consistency is typically established between OTI and the other contributing
City Agencies, such as NYPD, Housing Preservation and Development (HPD), Parks \& Recreation, etc.,
via Service Level Agreements (SLAs). However, enforcement of those SLAs is often challenging. The 
result is that often data accuracy and logical consistency is occasionally compromised.

The NYC 311 organization, part of OTI, operates a highly tailored and sophisticated case
management system based, as noted, on Microsoft Dynamics. However, when the original 311
system was developed in 2003, a decision was made to allow select NYC Agencies to
maintain their own local computer systems, such as the mainframe systems at the NYC Department
of Sanitation (DSNY) and others. And when the 311 system was completely upgraded in 2019, those
same integration were `honored` in order to reduce the scope of the project. (Perhaps in
retrospect, while a good budgetary decision, perhaps not such a good technology decision.)

The result was that at least nine NYC Agencies have local systems that are integrated either via
an API or custom code. These Agencies include:

\begin{itemize}
	\item Department of Social Services (DSS), one of the largest City Agencies
	\item Department of Sanitation, although recently part of DSNY has moved to the core 311 software.
	\item Department of Parks \& Recreation (forestry complaints only)
	\item Housing, Preservation \& Development (fully integrated)
	\item Department of Environmental Protection (DEP) (fully integrated)
	\item Department of Healthy \& Mental Hygiene (fully integrated) 
	\item Department of Consumer \& Worker Protection (DCWP - formerly known as the Dept of Consumer Affairs)
	\item New York Police Department (NYPD) (integrated with their reporting system)
	\item Department of Business (DoB) (fully integrated)
\end{itemize}

One result of these integrations, and other Agencies that use the 311 APIs, is that some of the data accuracies
and data consistency checks inherent in the 311 system are not applied uniformly, and in some cases not at all.
Errors often creep in from other Agency systems and are not easily corrected in the 311 application.

One area that will almost certainly remain an ongoing challenge is the incorporation of new software, hardware,
and cloud services which are constantly emerging.  Telephonic integration recently was reconfigured due to a new
telephone vendor. Desktop software at the 311 Agency are subject to the same upgrades that home users feel. And while the 
311 system was designed to be easily configured with new features, some emerging business issues as well as
technologies simply require additional software engineering efforts to accommodate.

One area in particular which is heavily influencing customer outreach for the 311 system is the use of
online chatbots. Increasing these are being influenced by Artificial Intelligence (AI) advances. It is a given that
a number of citizen's interactions with City government could be handled accurately and quickly by AI-driven
applications. How to do that while preserving customer privacy and anonymization are challenging issues.

The impact of the NYC 311 data extends beyond operational efficiency;
it has become instrumental in shaping City governance and community
engagement. The data has been pivotal in such areas as:

\begin{itemize}
	\item Advice on shelters during hurricanes and winter blizzards
	\item Countless inquires during the COVID pandemic (testing centers, vaccine sites, mobile vaccine clinics, etc.)
	\item Enforcement of standards between landlords and tenants.
	\item Re-allocation of NYC's yellow taxi routes based upon an insightful analysis conducted by the Taxi \& Limousine Commission (TLC)
	\item Improved responsiveness for City Agencies with direct client-facing responsibilities such as Parks \& Rec, Dept of Transportation, DSNY, etc.
	\item Increased 311 SRs related to Homelessness resulted in the launch of the City's Homeless Outreach \& Mobile Engagement Street Action Teams (HOME-STAT)
	which includes street canvassers and increases in homeless housing facilities.
	\item A significant increase in graffiti complains in 2017 led directly to the `Graffiti-Free NYC'' campaigns and increased funding for removal of graffiti in public spaces
	\item Enhanced geospatial data regarding the location of incidents, the responding Agency, and actions taken.
	\item Establishment of Agency SLAs in regard to certain complaint types, such as residential noise, homelessness person assistance, street light outages, 
	rat sightings, and other neighborhood quality of life issues.
\end{itemize}

The NYC 311 service exemplifies the dynamic nature of urban data
management and its critical role in modern governance. It provides an
invaluable resource for data scientists to practice data science
activities, including data curation.

We extracted  311 Service Request (SR) data from the period of CY 2022-2023, a dataset consisting of
6.4 million records. This dataset serves as the primary source for the analysis effort. This is a large dataset which enables 
observation of some rarely seen events, data inaccuracies, and logical inconsistencies. It provides enough data to examine trends and data outlier analysis. 
We query the data directly from the NYC Open Data Portal and then export the data in CSV format for analysis in a custom written R programs.

Additionally, we extracted a 10-year dataset consisting of the years 2014-2023; some 27 million rows. Albeit with just a few columns, 
this very large dataset allows for long-term trend observations and confirms that the analysis using the 
2-year dataset used for the majority of the analysis.

\section{10-year Analysis \& Trends} \label{sec:history}

Since the NYC 311 System went live in 2003, the usage has grown at an annual
rate of 4.6\%. For this study, we also looked at a 10-year period (2014-2023) to
observe certain system performance characteristics over a longer time-frame . 
During that time-frame, the number of 311 SRs grew 50\% as shown in the chart below which shows the yearly SR counts.

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.65]{Yearly.png}
  \caption{SR counts by year for a 10-year period}
  \label{fig:yearly-counts}
\end{figure}

This 10-year timeframe (2014-2023) included the COVID pandemic impact. Note that the pandemic
impact reached a peak in the July-September 2020 timeframe, accompanied by additional spikes
in the 2\textsuperscript{nd}half of 2021 as well. The all-time daily high for SRs occurred on Tuesday, 2020-08-04, with a spike of
24,415 SRs, a full 9.8 \textsigma's from the mean.

This next chart shows the 10-year period by calendar month. Note that almost all the months before the COVID pandemic outbreak in Feb/Mar
of 2020 are ``below average'', while almost all the months after that period are ``above average'' months. And that trend of much higher usage
of the 311 system has continued into the early months of 2024.

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.65]{Monthly.png}
  \caption{SR counts by month for a 10-year period}
  \label{fig:monthly-counts}
\end{figure}

The daily tend over that same 10-year period is also interesting, as it shows both significant spikes
and significant dips; a noisy graph. While the spikes in 2020 and 2021 appear COVID related, we are unable to 
ascertain any reason for the significant spike on Saturday, 2023-09-02 where 115,205 SRs were created.

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.3]{Daily.png}
  \caption{SR counts by day for a 10-year period}
  \label{fig:daily-counts}
\end{figure}

Seasonal trends are also prevalent. The months of November-thru-April are ``below average'' months, while the months May-thru-October are ``above average''.
 This seasonal trend is observed throughout the 10-year period.

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.65]{Calendar-Month.png}
  \caption{SR counts by calendar month for a 10-year period}
  \label{fig:calendar-months-counts}
\end{figure}

Of note, the lowest SR counts-per-day occur on Christmas  period (+/- a day), in 5 of the Bottom 10 SR counts-per-day. The summer moths and early
fall months make up all of the Top 10 SR counts-per-day. As previously noted, the all-time daily high for SRs occurred on Tuesday, 2020-08-04 with the 10-year
low number occurs on Christmas Day in 2016, a Sunday.

\begin{table}[htbp]
	\begin{minipage}[t]{0.35\textwidth} % Adjust the width of the minipage as needed
	\centering
	\small % Adjust the font size
	\caption{Top 10 days}
	\label{tab:top_ten_counts}
		\begin{tabular}{@{}lll@{}}
			\toprule
			Date & Count & Rank \\
			\midrule
			2020-08-04 & 24415 & 1 \\
			2020-08-05 & 19560 & 2 \\
			2023-09-29 & 17962 & 3 \\
			2020-07-05 & 16916 & 4 \\
			2020-06-21 & 15883 & 5 \\
			2020-06-20 & 15825 & 6 \\
			2020-07-04 & 15794 & 7 \\
			2021-09-02 & 15205 & 8 \\
			2020-06-28 & 14057 & 9 \\
			2021-10-28 & 13575 & 10 \\
			\bottomrule
		\end{tabular}
	\end{minipage}\hfill % Adjust the space between the tables
	\begin{minipage}[t]{0.35\textwidth} % Adjust the width of the minipage as needed
	\centering
	\small % Adjust the font size
	\caption{Bottom 10 days}
	\label{tab:bottom_ten_counts}
		\begin{tabular}{@{}lll@{}}
			\toprule
			Date & Count & Rank \\
			\midrule
			2014-12-25 & 2965 & 1 \\
			2015-12-25 & 3247 & 2 \\
			2014-04-20 & 3287 & 3 \\
			2015-12-26 & 3405 & 4 \\
			2016-11-06 & 3419 & 5 \\
			2014-07-04 & 3455 & 6 \\
			2014-12-27 & 3534 & 7 \\
			2014-09-14 & 3563 & 8 \\
			2014-02-02 & 3565 & 9 \\
			2016-12-25 & 3587 & 10 \\
			\bottomrule\\
		\end{tabular}
	\end{minipage}
\end{table}

And although the highest number of SRs occur in July and the lowest in February,
when the \#-of-SRs-per-day metric is computed by adjusting for the differing number of days for each month -- the
busiest month is shown to be June (followed by July and September). And similarly the quietest months are April, March, and 
February. In general, the winter months see fewer SRs while the summer months show 
an increase in activity, as previously noted.

\begin{table}[ht]
    \centering
    \begin{tabular}{lS[table-format=7.0]S[table-format=6.0]} % 'l' for left-aligned, 'S' for siunitx number alignment
        \toprule
        \textbf{Month}     & \textbf{Count}   & \textbf{Count per Day} \\ 
        \midrule
        June  & 2454156 & 81805 \\ 
        July  & 2461525 & 79404 \\ 
        September & 2352822 & 78427 \\ 
        October & 2407238 & 77653 \\ 
        August & 2401731 & 77475 \\ 
        May  & 2334466 & 75305 \\ 
        November  & 2236037 & 74535 \\ 
        January   & 2271524 & 73275 \\ 
        December  & 2199765 & 70960 \\ 
        February  & 1996408 & 70795 \\ 
        March & 2174349 & 70140 \\ 
        April  & 2058169 & 68606 \\ 
       \bottomrule
    \end{tabular}
    \caption{Monthly counts sorted by counts-per-day}
    \label{tab:monthly_counts}
\end{table}

\FloatBarrier % Ensure all floats before this point are processed before moving on

Two  areas that are important in this data cleansing analysis effort are the responsible NYC Agency, and the actual type of complaint.
Given both the size of the City of New York government, it's typically necessary to identify the responsible Agency as well as the type of complaint in order to 
clarify the responsible area and assist in troubleshooting any discrepancy.

In discovering a data error, we typically observe two trends; either a single (or a small few) Agencies are responsible (such as Dept of Transportation), or it is a system-wide issue that foolows the general distribution
of  Service Requests (SRs) across the system. Let's take a brief look at how this plays out across the system, so when we identify a
issue, it will be useful to see how that issue flows back to a specific Agency, and thus establish responsibility.

Although the 311 SRs involve 16 different agencies, there are the ``big six'' Agencies that comprise
90\% of the complaints. These six are (in order):

\begin{itemize}
	\item New York Police Department (NYPD)
	\item Housing Preservation \& Development (HPD
	\item New York City Department of Sanitation (DSNY)
	\item Department of Transportation (DOT)
	\item Department of Environmental Protection (DEP)
	\item Department of Parks \& Recreation (DPR)
\end{itemize}

\begin{figure}[htbp]
  \centering
	  \includegraphics[scale=0.65]{SRs_by_Agency.png}
	  \caption{SR counts by Agency and Cumulative Percentage}
	  \label{fig:SR_counts_by_Agency}
\end{figure}

It is also interesting to note the distribution of the type of complaints in the 311 SR dataset; it is heavily skewed to  a select few complaints.

\begin{figure}[htbp]
  \centering
	  \includegraphics[scale=0.65]{SRs_by_Complaint_Type.png}
	  \caption{Top 20 SR Complaints and Cumulative Percentage}
	  \label{fig:SR_complaints}
\end{figure}

The top 20 complaints contain several of the eight different types of ``Noise'' complaints (Residential, Commercial, Street, Helicopter, etc.). These Noise-related
complaints total 1.435 million in the 2022-2023 dataset, a full 22\% of all complaint types, which is the most frequently occurring type. (Note: Ilegal Parking is 
second, followed by Heat/Hot Water.) Noise complaints are also handled by NYPD, as is Illegal Parking and Blocked Driveway complaints, which mostly
accounts for the top ranking of 311 SRs being routed to the NYPD.

The NYC HPD handles about 49\% of what NYPD handles. For  non-NYC readers, 
HPD is the City agency that manages the  177,569 NYC public housing units as well as monitoring all rental and leased apartment units. HPD collectively
serves 528,105 people, a population larger than Atlanta or Miami. Hence the large number of Heat/Hot Water and Unsanitary complaints.

\section{Potential Issues} \label{sec:issues}

What is data cleansing?  Wikipedia \href{https://en.wikipedia.org/wiki/Data_cleansing}{Data Cleansing} offers a good definition.

``Data cleansing or data cleaning is the process of detecting and correcting (or removing) corrupt or inaccurate records from a record set, 
table, or database and refers to identifying incomplete, incorrect, inaccurate or irrelevant parts of the data 
and then replacing, modifying, or deleting the dirty or coarse data.''

Many quality criteria are required in order to process high-quality data. These include:

\begin{itemize}
	\item Data validation - this effort can span a number of criteria
	\begin{itemize}
		\item Mandatory fields: Certain data fields cannot be empty
		\item Data-types: Certain fields must be of the correct type, e.g. numeric, character, date, 5 numeric digits, etc. Typically these are referenced in a Data Dictionary.
		\item Domain adherence:  Many data fields must adhere to a specific domain of values, e.g. statuses, state names, zip codes, gender. This includes data
		fields that are restricted to certain ranges, such as longitudes and latitudes.
	\end{itemize}   
	\item Structural errors to include naming conventions, extra fields no in the Data Dictionary, inconsistent data entry.
	For example intermixing blanks, space, NA, N/A. and <NA> to indicate the absence of data.
	\item Redundant, unnecessary, or inappropriate fields
	\item Logical inconsistencies such as two related fields that violate the nature of that relationship, such as a ``due date'' that is before the ``created date''.
	\item Data standardization. Often free-form entry fields can suffer from inconsistent data structures. Efforts to standardize these fields can greatly assist analysis.
	\item Accuracy and precision, which of course are not the same thing 
\end{itemize}

The effort of this analysis effort is to identify the presence of such errors; not to correct them. That effort would be undertaken after an investigation as to the why
and how such errors came about. Often this requires discussions with subject matter experts and liaison with various NYC Agency open data coordinators
to identify the original source of the error. For example, an invalid date field that is provided to the NYC Open Data data lake via an Agency integration could
be the result of selecting the wrong field from the source Agency's system, e.g. due date instead of closed date. In other cases, without a subject matter
expert, it is not possible to actually determine if the data is inaccurate or not. The R code used to analyze the 311 Service Request dataset identifies
both obvious inaccuracies as well as potential errors requiring further investigation.

This paper will focus on a set of specific data cleansing areas:

\begin{itemize}
	\item Structural issues with the data. Is it in the expected format?
	\item Missing/blank data
	\item Data type correctness
	\item Invalid values
	\item Logical inconsistencies and inconsistent patterns in the data
	\item Accuracy and precision
	\item Redundant data  
\end{itemize}

\subsection{Structural Issues}

Structural issues in the context of data cleansing refers to issues related to the how data is organized, formatted, or structured within a dataset. Structural issues 
can make it difficult to analyze the data effectively. Some common structural concerns in this 311 SR 2022-2023 dataset include:

\begin{itemize}
	\item Data structure (data fields, columns, formats, data types, etc.) not corresponding to the Data Dictionary
	\item Verification of correct data types (numeric, character, geospactial, dates, etc.)
	\item Embedded or combined values: data elements that contain multiple pieces of information 
\end{itemize}

In this dataset the writers discovered several structural issues.

	\subsubsection{Fields not in the Data Dictionary}
	The 311 SR \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} identifies 41 data columns (fields) along with some
	related information for each column. However, when you either download the data or explore it via the online portal,
	you will immediately note that there are 47 fields. The additional six fields are not addressed in the Data Dictionary, and show up on the Column Manager
	page as ``@computed\_region\_xxxx\_xxxx''. To some extent one can infer what the these @computed\_fields are, but it is not possible to know
	for certain. Here is a screenshot of the NYC Open Data Portal ``Column Manager'' page showing those columns:

	\begin{figure}[htbp]
	  \centering
		  \includegraphics[scale=0.65]{computed_columns_screenshot.png}
		  \caption{Screenshot of data portal showing the ``@computed'' columns}
		  \label{fig:computed-columns}
	\end{figure}

	This, of course, raises several questions:
	\begin{itemize}
		\item How are these fields computed? Using what source data? Using what computational method?
		\item How can these fields be verified or cross-checked?
		\item Why are their two fields that appear redundant ``Police Precinct'' and ``Police Precincts''?
		\item What are ``borough boundaries''? Are these the regular NYC boroughs?
		\item What are ``community districts''? Is this the same as the ``Community Boards'' that are part of the NYC government?
		\item Can these fields be included in the official 311 SR Data Dictionary?
		\item Can these fields be used for analytical purposes? 
	\end{itemize}

	The last bullet point is the key here. Are these fields reliable and accurate, and can they be used for subsequent analysis? Unfortunately,
	as of this date, those questions have not been fully addressed by the  NYC Open Data Team. This paper will address further issues with the
	accuracy and subsequent analytical usage for these six @computed fields.

	\subsubsection{Verifying data types}
	Fortunately, both the Data Dictionary and the portal column manager do a good job of indicating the data type.
	As indicated by the red underline in Figure \ref{fig:computed-columns}, there is a small icon next to the column name 
	which indicates the data type of the column. The stylized capital ``T'' indicates text, the ``\#'' sign indicates 
	a numeric, a calendar icon indicates a date field, and the ``pin on a map'' symbol indicates a geospatial field.

	The following fields were checked for their specified data types:
		\begin{itemize}
		\item created\_date, closed\_date, due\_date, and resolution\_action\_updated\_date are all in proper date format.
		\item zip\_codes and incident\_zip: All are 5 numeric digits except for 2 non-numeric entries for incident\_zip (``na'' and ``N/A'').
		\item x\_coordinate\_state\_plane \&  y\_coordinate\_state\_plane (a NY State geo-location system) are all numeric.
		\item latitude \& longitude are both numeric.
		\item community\_districts, borough\_boundaries, city\_council\_district, police\_precinct, and police\_precincts are all numeric.
		\item Free-form text fields that contain a ``,'' (e.g. incident\_address, resolution\_description) are enclosed in quotation marks.
		\end{itemize}

\subsection{Missing-Blank data}

\subsection{Data type correctness}

\subsection{Invalid values}

\subsection{Logical inconsistencies }

\subsection{Accuracy and precision}

\subsection{Redundant data}

\subsection{Case Study I - Response to Requests for Homeless Person Assistance}

\subsection{Case Study I - Studying Noise Complaints by NYC Zip Code}

Invalid values?
Logical inconsistencies? Concerning patterns?
Redundant columns?

Invalid data: closed data earlier than creation date;
latitude/longitude not in NYC; invalid zip code;

Inconsistent data: upper/lower cases; extra spaces; 
open text entries: provide enough major categories and sub categories;

Inefficiently stored data: lat/long versus (lat, long);
intersection street 1 versus cross street 1;
borough could be inferred from zip code;
full name of agency not needed

\section{Improvements} \label{sec:improve}

Data validation: identify invalid data

Data cleaning: convert to all lower/upper cases;

Data organization: put zipcode/borough in a separate table; verify intersection
== cross; remove redundant columns (and compare how much space saved)

\section{Protocol Suggestions} \label{sec:protocol}

\section{Discussion} \label{sec:disc}

\bibliographystyle{asa}
\bibliography{ref}

\end{document}

\end{document}

### 2024-05-22 manuscript/311 manuscript.tex
Daily updates

@@
 \usepackage{caption}  % For custom captions
+\usepackage{float}   % Required for the H specifier
 
@@
 The daily tend over that same 10-year period is also interesting, as it shows both significant spikes
-and significant dips; a noisy graph. While the spikes in 2020 and 2021 appear COVID related, we are unable to 
-ascertain any reason for the significant spike on Saturday, 2023-09-02 where 115,205 SRs were created.
+and significant dips; a noisy graph. While the spikes on 2020-08-04 (24,415 SRs) and 2020-08-05 (19,560 SRs) are COVID related, we are unable to 
+ascertain any reason for the significant spike on Saturday, 2023-09-29 where 17,962 SRs were created.
 
@@
   \centering
-  \includegraphics[scale=0.3]{Daily.png}
+  \includegraphics[scale=0.35]{Daily.png}
   \caption{SR counts by day for a 10-year period}
@@
 fall months make up all of the Top 10 SR counts-per-day. As previously noted, the all-time daily high for SRs occurred on Tuesday, 2020-08-04 with the 10-year
-low number occurs on Christmas Day in 2016, a Sunday.
+low number occurs on Christmas Day in 2016, a Sunday. Note also that the difference between the peak day (24,415) and the minimum day (2965) varies
+by a factor of 8X, and the aveage daily count of 7489 -- all of which creates a challenging  system capacity design issue.
 
@@
 
+What are the characteristics of the 311 SR data set. Here are some insights:
+
+\begin{itemize}
+	\item 47 columns of data for each row, exportable as a CSV file
+	\item There are four date fields (created, closed, updated, due)
+	\item There are three borough fields; two of which we believe to be duplicates
+	\item Two zip code fields, but not duplicates
+	\item Seven street fields; one pair of which we believe to be duplicates
+	\item Agency name and Agency abbreviation
+	\item Two Police Precinct fields; not duplicates
+	\item In addition to street/city, there are three other location fields (lat/long), NY State plane, and Block \#
+	\item Incident address (25 Grymes Hill Road) and Street Name (Grymes Hill Road)
+	\item One form tex fielt; resolution description, which supports 934 characters of input
+\end{itemize}
+
 In this dataset the writers discovered several structural issues.
@@
 	\subsubsection{Fields not in the Data Dictionary}
-	The 311 SR \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} identifies 41 data columns (fields) along with some
+	The 311 SR \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}
+	 identifies 41 data columns (fields) along with some
 	related information for each column. However, when you either download the data or explore it via the online portal,
@@
 	page as ``@computed\_region\_xxxx\_xxxx''. To some extent one can infer what the these @computed\_fields are, but it is not possible to know
-	for certain. Here is a screenshot of the NYC Open Data Portal ``Column Manager'' page showing those columns:
+	for certain. These fields are:
+	\begin{itemize}
+		\item zip\_codes
+		\item community\_districts
+		\item borough\_boundaries
+		\item city\_council\_districts
+		\item police\_precincts
+		\item police\_precinct 
+	\end{itemize}
+
+	Here is a screenshot of the NYC Open Data Portal ``Column Manager'' page showing those columns:
 
@@
 	The following fields were checked for their specified data types:
+
 		\begin{itemize}
@@
 
+Understanding the absence of data in each fields is an important factor when undertaking analysis. For example, if you wanted to see if the SRs were
+closed before or after their due\_date, you would be challenged as 99.57\% of the due\_date field is blank. (Although with such a large dataset
+that still leaves 24,411 values, all of which are N/A, which calls into question are these missing values, or is the specification of a
+due\_date simply not applicable for the vast majority of SRs?)
+
+When counting the various fields for blank or N/A values, the dataset appears to divide into three groups: Mostly Empty, Partially Empty, and Few/None Empty.
+The Mostly Empty category ranges from 93-99.9\% blank and includes such fields as taxi\_company due\_date, pickup\_location, and landmark. The Partially 
+Empty includes such fields as location\_type, borough, and cross\_street. And the Few/None Empty includes created\_date, complaint\_type, agency, and status.
+In some cases it may make sense to inquiry as to why some fields are frequently blank. Is the data difficult to capture? Does it pertain to only a small set
+of complaint\_type or Agencies? And if the data is truly that sparse, does it make sense to even collect it. For example, you would think that every complaint
+that every SR that has a geographical component, such as an address, would also have a corresponding police\_precinct and community\_district since these
+are all-inclusive. That it would have a lat/long.  But indeed these fields do have small, but nonetheless significant empty entries.
+
+During the analysis, it was determined that the following fields are mandatory, and any rows missing these fields is removed from the dataset (which very, very
+rarely occurred fortunately): created\_date, agency, complaint\_type, and unique\_key.
+
+\begin{table}[ht]
+\centering
+\caption{Blank and N/A entries by field}
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{lrrrl}
+	\toprule
+	\textbf{Field} & \textbf{Total Empty} & \textbf{Percent Empty} & \textbf{NA Count} & \textbf{Blanks Only} \\
+	\midrule
+		taxi\_company\_borough & 6391341 & 99.94 & 0 & 6391341 \\
+		road\_ramp & 6379080 & 99.75 & 0 & 6379080 \\
+		vehicle\_type & 6376266 & 99.71 & 0 & 6376266 \\
+		due\_date & 6370497 & 99.62 & 6370497 & 0 \\
+		bridge\_highway\_direction & 6367626 & 99.57 & 0 & 6367626 \\
+		bridge\_highway\_name & 6340197 & 99.14 & 0 & 6340197 \\
+		bridge\_highway\_segment & 6340190 & 99.14 & 0 & 6340190 \\
+		taxi\_pick\_up\_location & 6329954 & 98.98 & 0 & 6329954 \\
+		facility\_type & 5966588 & 93.30 & 0 & 5966588 \\
+		landmark & 2635736 & 41.22 & 0 & 2635736 \\
+		intersection\_street\_1 & 2143847 & 33.52 & 0 & 2143847 \\
+		intersection\_street\_2 & 2140707 & 33.48 & 0 & 2140707 \\
+		cross\_street\_1 & 1846594 & 28.88 & 0 & 1846594 \\
+		cross\_street\_2 & 1846008 & 28.87 & 0 & 1846008 \\
+		location\_type & 798775 & 12.49 & 0 & 798775 \\
+		bbl & 768109 & 12.01 & 0 & 768109 \\
+		city & 332086 & 5.19 & 0 & 332086 \\
+		street\_name & 280463 & 4.39 & 0 & 280463 \\
+		incident\_address & 280268 & 4.38 & 0 & 280268 \\
+		closed\_date & 248839 & 3.89 & 248839 & 0 \\
+		resolution\_description & 132333 & 2.07 & 0 & 132333 \\
+		zip\_codes & 123167 & 1.93 & 0 & 123167 \\
+		borough\_boundaries & 101453 & 1.59 & 0 & 101453 \\
+		community\_districts & 101445 & 1.59 & 0 & 101445 \\
+		city\_council\_districts & 101445 & 1.59 & 0 & 101445 \\
+		police\_precincts & 101445 & 1.59 & 0 & 101445 \\
+		police\_precinct & 101416 & 1.59 & 0 & 101416 \\
+		latitude & 99778 & 1.56 & 0 & 99778 \\
+		longitude & 99778 & 1.56 & 0 & 99778 \\
+		location & 99778 & 1.56 & 0 & 99778 \\
+		x\_coordinate\_state\_plane & 99668 & 1.56 & 0 & 99668 \\
+		y\_coordinate\_state\_plane & 98822 & 1.55 & 0 & 98822 \\
+		resolution\_action\_updated\_date & 91399 & 1.43 & 91399 & 0 \\
+		incident\_zip & 83220 & 1.30 & 2 & 83218 \\
+		descriptor & 74645 & 1.17 & 0 & 74645 \\
+		address\_type & 38655 & 0.60 & 0 & 38655 \\
+		unique\_key & 0 & 0.00 & 0 & 0 \\
+		agency & 0 & 0.00 & 0 & 0 \\
+		agency\_name & 0 & 0.00 & 0 & 0 \\
+		complaint\_type & 0 & 0.00 & 0 & 0 \\
+		status & 0 & 0.00 & 0 & 0 \\
+		community\_board & 0 & 0.00 & 0 & 0 \\
+		borough & 0 & 0.00 & 0 & 0 \\
+		open\_data\_channel\_type & 0 & 0.00 & 0 & 0 \\
+		park\_facility\_name & 0 & 0.00 & 0 & 0 \\
+		park\_borough & 0 & 0.00 & 0 & 0 \\
+		created\_date & 0 & 0.00 & 0 & 0 \\
+	\bottomrule
+\end{tabular}
+}
+\end{table}
+
+\FloatBarrier % Ensure all floats before this point are processed before moving on
+
+Here is a graphic depiction of total empty (blank + N/As) for each fields. You can see the natural grouping into the Most, Some, and Few categories.
+
+\begin{figure}[htbp]
+  \centering
+	  \includegraphics[scale=0.7]{BlankFields.png}
+	  \caption{Number and Percentage of Empty Entries}
+	  \label{fig:blank_fields}
+\end{figure}
+
 \subsection{Data type correctness}
+All data fields were subjected to a test of ``correctness'' according to their ``type'' as specified
+in the Data Dictionary.  Of the nearly 300 million data fields to check (6.5 million rows, 47 fields per row), there were only two  values that failed to test for the
+proper data type; those being two text entries in the incident\_zip filed (``na'', and N/A). All date fields were dates. All number fields were numeric. All text
+fields were character.
 
-\subsection{Invalid values}
+\subsection{Checking for Valid Values}
+Testing for valid and invalid values revealed several fairly serious issues associated with select fields. Accordingly, an analyst would need to be
+very diligent if using these fields for analysis. In all likelihood, the rows with the invalid field values would likely need to be removed from the
+analysis.
+
+There were a variety of tests for validity that were applied individually to specific fields. Here are the ``good'' results:
+
+\begin{itemize}
+		\item Latitude and Longitude fields were tested to ensure all fell within the geographic boundaries of the City of New York. All were.
+		\item The unique\_key field was in fact unique.
+		\item Many fields have a domain of acceptable values. Often these values are determined from common usage or by examining
+		larger datasets. Unfortunately, these domains of acceptable values are not specified in the Data Dictionary. These fields all tested as
+		compliant with their domain of acceptable values:
+			\begin{itemize}
+				\item address\_type
+				\item status
+				\item borough, borough\_boundaries, \& park\_borough 
+				\item data\_channel
+				\item vehicle\_type
+				\item city\_council\_district
+			\end{itemize}
+\end{itemize}
+
+ Unfortunately some fields did prove to be quite problematic when comparing the values to the domain of legal values.
+ For example, all zip codes (two fields: zip\_codes and incident\_zip) should be valid as defined by the USPS database
+ which contains over 44,000 valid zip codes.
+
+ The computed field zip\_codes proved especially problematic with 58\% (3.6 million) of the field entries being invalid.
+This high number indicates that the computation of this computed fields is highly inaccurate. We would recommend
+dropping this field for future analytic efforts.
+
+Furthermore, the field incident\_zip while having only .07\% invalid entries, that is still 4163 such errors. As zip code is
+a primary measure of many NYC city services, and with a freely available database for which to validate entries
+against, it seems an oversight that any such errors could creep into the system. Some invalid entries are 
+clearly incorrect just by observation, e.g. 10000, 12345, 11111, etc.
+
+Because zip codes play a significant role in analyzing City services, these errors can be quite serious. It's useful in 
+such cases to observe the breakdown of these errors by the responsible Agency.
+
+The breakdown of the invalid entries in the zip\_code, sorted by Agency shows that the distribution by percentage
+almost precisely mirrors the overall breakdown of SRs by Agency. This indicates a systemic problem, and since the
+zip\_code fields is one of the computed fields, it appears these errors are caused by incorrect computations rather
+than by any Agency mistake.
+
+The  errors in the incident\_zip field are more troubling, even though they are, percentage-wise, small. Here is a graphic 
+illustrating how these errors occur by Agency. Here we can see that the majority of these errors lie in the Dept of Transportation,
+NYPD, Taxi \& Limo Commission (TLC), and the Economic Development Council (EDC) among others. This representation
+does not follow the full SR distribution indicating that these errors are likely generated by an incorrect
+process or application residing at those particular Agencies; it is not a systemic problem.
+
+\begin{figure}[htbp]
+  \centering
+	  \includegraphics[scale=0.7]{invalid_incident_zip.png}
+	  \caption{Invalid incident-zip by Agency}
+	  \label{fig:invalid_incident_zip}
+\end{figure}
+
+A curious case also exists when examining the two nearly identical fields - police\_precincts and police\_precinct.
 

### 2024-05-22 code/311 data cleansing V2.R

@@ -24,6 +24,9 @@ sink("console_output.txt")
 
 setwd("C:/Users/david/OneDrive/Documents/nyc311clean/code")
 
+setwd("C:/Users/david/OneDrive/Documents/nyc311clean/code")
+data1File <- file.path("..", "data", "311_2022-2023_AS_OF_01-31-2024.csv")
+
 # Hard code the max_closed_date to be midnight of the date of the data export from NYC Open Data
 max_closed_date <- as.POSIXct("2024-01-31 23:59:59", format = "%Y-%m-%d %H:%M:%S")
 
@@ -88,7 +91,7 @@ create_bar_chart <- function(
   cat(
     "\n", paste("Average", print_out_title, ":"), format(mean_value, big.mark = ","), "  ",
     paste("Median", print_out_title, ":"), format(median_value, big.mark = ","),
-    paste("\nStd Dev", print_out_title, ":"), format(median_value, big.mark = ",")
+    paste("Std Dev", print_out_title, ":"), format(median_value, big.mark = ",")
   )
 
 
@@ -1237,6 +1240,9 @@ missingDataPerColumn <- cbind(missingDataPerColumn, NA_count = na_counts_per_col
 
 missingDataPerColumn$blanks_only <- missingDataPerColumn$total_empty - missingDataPerColumn$NA_count
 
+# Add the "non_blank" column
+#missingDataPerColumn <- missingDataPerColumn %>% mutate(non_blank = num_rows_d311 - total_empty)
+
 # Sort the data frame by the sum of NAs and blanks in descending order
 missingDataPerColumn <- missingDataPerColumn[order(missingDataPerColumn$total_empty, decreasing = TRUE), ]
 
@@ -1401,7 +1407,7 @@ SR_day_of_the_month <- create_bar_chart(
 # SRs created by the day of the year (1-365)
 SR_day_of_the_year <- create_bar_chart(
   dataset = sorted_by_day_of_year,
-  x_col = "calendar_date",
+  x_col = "day_of_year",
   chart_title = "Day of the year SR count w/average",
   sub_title = chart_sub_title,
   x_axis_title = NULL,

### 2024-05-22 manuscript/311 manuscript.tex
Nightly update
Created a new 5-year trend analysis and console file

@@
 \usepackage{float}   % Required for the H specifier
+\usepackage{adjustbox}
 
@@
 These (and other) challenges are constantly being addressed by the various agency open data managers
-as well as the \href{https://www.nyc.gov/content/oti/pages/}{NYC Office of Technology and Innovation} which
+as well as the \href{https://www.nyc.gov/content/oti/pages/}{NYC Office of Technology and Innovation (OTI)} which
 provides the technology support for the open data system.  
@@
 and data consistency checks inherent in the 311 system are not applied uniformly, and in some cases not at all.
-Errors often creep in from other Agency systems and are not easily corrected in the 311 application.
+Errors often creep in from other Agency systems and are not easily corrected in the 311 application itself,
+but rather require correction at the source Agency.
 
@@
 
-Additionally, we extracted a 10-year dataset consisting of the years 2014-2023; some 27 million rows. Albeit with just a few columns, 
-this very large dataset allows for long-term trend observations and confirms that the analysis using the 
-2-year dataset used for the majority of the analysis.
+Additionally, we extracted two longer-term datasets, a 5-year look (2019-2023) and a 10-year look (2014-2023) ; some 14  and 27 million rows respectively, albeit with just a few columns. 
+These larger datasets afford an opportunity for long-term trend observations, confirming the analysis using the 2-year dataset. The 2022-2023 dataset
+was used exclusively for the detailed, field-level analysis that comprises the majority of this effort.
 
@@
   \centering
-  \includegraphics[scale=0.65]{Yearly.png}
+  \includegraphics[width=\textwidth]{Yearly.png}
   \caption{SR counts by year for a 10-year period}
@@
 
-This next chart shows the 10-year period by calendar month. Note that almost all the months before the COVID pandemic outbreak in Feb/Mar
-of 2020 are ``below average'', while almost all the months after that period are ``above average'' months. And that trend of much higher usage
+This below chart shows the 10-year period by calendar month. Note that almost all the months before the COVID pandemic outbreak in Feb/Mar
+of 2020 are ``below average'', while most of the months after that period are ``above average'' months. And that trend of higher usage
 of the 311 system has continued into the early months of 2024.
@@
   \centering
-  \includegraphics[scale=0.65]{Monthly.png}
+  \includegraphics[width=\textwidth]{Monthly.png}
   \caption{SR counts by month for a 10-year period}
@@
 
-The daily tend over that same 10-year period is also interesting, as it shows both significant spikes
+A quick look at the corresponding  daily tend over that same 10-year period is also interesting, as it shows both significant spikes
 and significant dips; a noisy graph. While the spikes on 2020-08-04 (24,415 SRs) and 2020-08-05 (19,560 SRs) are COVID related, we are unable to 
@@
 
-\begin{figure}[htbp]
+\begin{figure}[H]
   \centering
-  \includegraphics[scale=0.35]{Daily.png}
+  \includegraphics[width=\textwidth]{Daily.png}
   \caption{SR counts by day for a 10-year period}
@@
 
-\begin{figure}[htbp]
+\begin{figure}[H]
   \centering
-  \includegraphics[scale=0.65]{Calendar-Month.png}
+  \includegraphics[width=\textwidth]{Calendar-Month.png}
   \caption{SR counts by calendar month for a 10-year period}
@@
     \centering
+    \small
     \begin{tabular}{lS[table-format=7.0]S[table-format=6.0]} % 'l' for left-aligned, 'S' for siunitx number alignment
@@
   \centering
-	  \includegraphics[scale=0.65]{SRs_by_Agency.png}
+	  \includegraphics[width=\textwidth]{SRs_by_Agency.png}
 	  \caption{SR counts by Agency and Cumulative Percentage}
@@
   \centering
-	  \includegraphics[scale=0.65]{SRs_by_Complaint_Type.png}
+	  \includegraphics[width=\textwidth]{SRs_by_Complaint_Type.png}
 	  \caption{Top 20 SR Complaints and Cumulative Percentage}
@@
 accounts for the top ranking of 311 SRs being routed to the NYPD.
+
+There are also some rather curious and unusual Service Requests in the 311 data: 
+tanning, tattooing, trans fat, unsanitary pigeon condition, illegal animal kept as pet, harboring bees/wasps, and radioactive material. 
+New Yorkers certainly live interesting lives.
 
@@
 	\item Structural issues with the data. Is it in the expected format?
+	\item Data type correctness
 	\item Missing/blank data
-	\item Data type correctness
 	\item Invalid values
@@
 
-\subsection{Structural Issues}
+\section{Structural Issues}
 
@@
 
-In this dataset the writers discovered several structural issues.
-
-	\subsubsection{Fields not in the Data Dictionary}
-	The 311 SR \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}
+\textbf{Fields not in the Data Dictionary} The 311 SR \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}
  identifies 41 data columns (fields) along with some
@@
   \centering
-		  \includegraphics[scale=0.65]{computed_columns_screenshot.png}
+	  \includegraphics[scale = 0.7]{computed_columns_screenshot.png}
 	  \caption{Screenshot of data portal showing the ``@computed'' columns}
@@
 
-	\subsubsection{Verifying data types}
+\section{Verifying data types}
 Fortunately, both the Data Dictionary and the portal column manager do a good job of indicating the data type.
@@
 
-\subsection{Missing-Blank data}
+\section{Blank \& N/A data}
 
@@
 \caption{Blank and N/A entries by field}
-\resizebox{\textwidth}{!}{%
+\resizebox{0.8\linewidth}{!}{%
+\tiny
 \begin{tabular}{lrrrl}
 	\toprule
-	\textbf{Field} & \textbf{Total Empty} & \textbf{Percent Empty} & \textbf{NA Count} & \textbf{Blanks Only} \\
+	\textbf{Field} & \textbf{Total Empty} & \textbf{Percent Empty} & \textbf{N/A Count} \\
 	\midrule
-		taxi\_company\_borough & 6391341 & 99.94 & 0 & 6391341 \\
-		road\_ramp & 6379080 & 99.75 & 0 & 6379080 \\
-		vehicle\_type & 6376266 & 99.71 & 0 & 6376266 \\
-		due\_date & 6370497 & 99.62 & 6370497 & 0 \\
-		bridge\_highway\_direction & 6367626 & 99.57 & 0 & 6367626 \\
-		bridge\_highway\_name & 6340197 & 99.14 & 0 & 6340197 \\
-		bridge\_highway\_segment & 6340190 & 99.14 & 0 & 6340190 \\
-		taxi\_pick\_up\_location & 6329954 & 98.98 & 0 & 6329954 \\
-		facility\_type & 5966588 & 93.30 & 0 & 5966588 \\
-		landmark & 2635736 & 41.22 & 0 & 2635736 \\
-		intersection\_street\_1 & 2143847 & 33.52 & 0 & 2143847 \\
-		intersection\_street\_2 & 2140707 & 33.48 & 0 & 2140707 \\
-		cross\_street\_1 & 1846594 & 28.88 & 0 & 1846594 \\
-		cross\_street\_2 & 1846008 & 28.87 & 0 & 1846008 \\
-		location\_type & 798775 & 12.49 & 0 & 798775 \\
-		bbl & 768109 & 12.01 & 0 & 768109 \\
-		city & 332086 & 5.19 & 0 & 332086 \\
-		street\_name & 280463 & 4.39 & 0 & 280463 \\
-		incident\_address & 280268 & 4.38 & 0 & 280268 \\
-		closed\_date & 248839 & 3.89 & 248839 & 0 \\
-		resolution\_description & 132333 & 2.07 & 0 & 132333 \\
-		zip\_codes & 123167 & 1.93 & 0 & 123167 \\
-		borough\_boundaries & 101453 & 1.59 & 0 & 101453 \\
-		community\_districts & 101445 & 1.59 & 0 & 101445 \\
-		city\_council\_districts & 101445 & 1.59 & 0 & 101445 \\
-		police\_precincts & 101445 & 1.59 & 0 & 101445 \\
-		police\_precinct & 101416 & 1.59 & 0 & 101416 \\
-		latitude & 99778 & 1.56 & 0 & 99778 \\
-		longitude & 99778 & 1.56 & 0 & 99778 \\
-		location & 99778 & 1.56 & 0 & 99778 \\
-		x\_coordinate\_state\_plane & 99668 & 1.56 & 0 & 99668 \\
-		y\_coordinate\_state\_plane & 98822 & 1.55 & 0 & 98822 \\
-		resolution\_action\_updated\_date & 91399 & 1.43 & 91399 & 0 \\
-		incident\_zip & 83220 & 1.30 & 2 & 83218 \\
-		descriptor & 74645 & 1.17 & 0 & 74645 \\
-		address\_type & 38655 & 0.60 & 0 & 38655 \\
-		unique\_key & 0 & 0.00 & 0 & 0 \\
-		agency & 0 & 0.00 & 0 & 0 \\
-		agency\_name & 0 & 0.00 & 0 & 0 \\
-		complaint\_type & 0 & 0.00 & 0 & 0 \\
-		status & 0 & 0.00 & 0 & 0 \\
-		community\_board & 0 & 0.00 & 0 & 0 \\
-		borough & 0 & 0.00 & 0 & 0 \\
-		open\_data\_channel\_type & 0 & 0.00 & 0 & 0 \\
-		park\_facility\_name & 0 & 0.00 & 0 & 0 \\
-		park\_borough & 0 & 0.00 & 0 & 0 \\
-		created\_date & 0 & 0.00 & 0 & 0 \\
+		taxi\_company\_borough & 6391341 & 99.94 & 0 \\
+		road\_ramp & 6379080 & 99.75 & 0 \\
+		vehicle\_type & 6376266 & 99.71 & 0 \\
+		due\_date & 6370497 & 99.62 & 6370497 \\
+		bridge\_highway\_direction & 6367626 & 99.57 & 0 \\
+		bridge\_highway\_name & 6340197 & 99.14 & 0 \\
+		bridge\_highway\_segment & 6340190 & 99.14 & 0 \\
+		taxi\_pick\_up\_location & 6329954 & 98.98 & 0 \\
+		facility\_type & 5966588 & 93.30 & 0 \\
+		landmark & 2635736 & 41.22 & 0 \\
+		intersection\_street\_1 & 2143847 & 33.52 & 0 \\
+		intersection\_street\_2 & 2140707 & 33.48 & 0 \\
+		cross\_street\_1 & 1846594 & 28.88 & 0 \\
+		cross\_street\_2 & 1846008 & 28.87 & 0 \\
+		location\_type & 798775 & 12.49 & 0 \\
+		bbl & 768109 & 12.01 & 0 \\
+		city & 332086 & 5.19 & 0 \\
+		street\_name & 280463 & 4.39 & 0 \\
+		incident\_address & 280268 & 4.38 & 0 \\
+		closed\_date & 248839 & 3.89 & 248839 \\
+		resolution\_description & 132333 & 2.07 & 0 \\
+		zip\_codes & 123167 & 1.93 & 0 \\
+		borough\_boundaries & 101453 & 1.59 & 0 \\
+		community\_districts & 101445 & 1.59 & 0 \\
+		city\_council\_districts & 101445 & 1.59 & 0 \\
+		police\_precincts & 101445 & 1.59 & 0 \\
+		police\_precinct & 101416 & 1.59 & 0 \\
+		latitude & 99778 & 1.56 & 0 \\
+		longitude & 99778 & 1.56 & 0 \\
+		location & 99778 & 1.56 & 0 \\
+		x\_coordinate\_state\_plane & 99668 & 1.56 & 0 \\
+		y\_coordinate\_state\_plane & 98822 & 1.55 & 0 \\
+		resolution\_action\_updated\_date & 91399 & 1.43 & 91399 \\
+		incident\_zip & 83220 & 1.30 & 2 \\
+		descriptor & 74645 & 1.17 & 0 \\
+		address\_type & 38655 & 0.60 & 0 \\
+		unique\_key & 0 & 0.00 & 0 \\
+		agency & 0 & 0.00 & 0 \\
+		agency\_name & 0 & 0.00 & 0 \\
+		complaint\_type & 0 & 0.00 & 0 \\
+		status & 0 & 0.00 & 0 \\
+		community\_board & 0 & 0.00 & 0 \\
+		borough & 0 & 0.00 & 0 \\
+		open\_data\_channel\_type & 0 & 0.00 & 0 \\
+		park\_facility\_name & 0 & 0.00 & 0 \\
+		park\_borough & 0 & 0.00 & 0 \\
+		created\_date & 0 & 0.00 & 0 \\
 	\bottomrule
@@
 
-Here is a graphic depiction of total empty (blank + N/As) for each fields. You can see the natural grouping into the Most, Some, and Few categories.
+Here is a graphic depiction of total empty (blank \& N/As) for each fields. You can see the natural grouping into the Most, Some, and Few categories.
 
-\begin{figure}[htbp]
+\begin{figure}[H]
   \centering
-	  \includegraphics[scale=0.7]{BlankFields.png}
+	  \includegraphics[width=\textwidth]{BlankFields.png}
 	  \caption{Number and Percentage of Empty Entries}
@@
 
-\subsection{Data type correctness}
+\section{Data type correctness}
+
 All data fields were subjected to a test of ``correctness'' according to their ``type'' as specified
@@
 
-\subsection{Checking for Valid Values}
-Testing for valid and invalid values revealed several fairly serious issues associated with select fields. Accordingly, an analyst would need to be
-very diligent if using these fields for analysis. In all likelihood, the rows with the invalid field values would likely need to be removed from the
-analysis.
+ \section{Validating data}
+ While the dataset may have fields populated with data that appears correct, it is necessary to inspect those fields to determine if the data is indeed valid. For
+ example a date field that contains the value ``February 30, 2024'' is clearly not valid. Nor is a latitude of +95 deg north.
 
-There were a variety of tests for validity that were applied individually to specific fields. Here are the ``good'' results:
+ In this study, testing for invalid values revealed several serious issues associated with select fields. Accordingly, an analyst would need to be
+extra diligent if using these fields for analysis. In all likelihood, rows with invalid field values would likely need to be removed from the
+analysis dataset.
+
+Some fields can be validated against either a acceptable values, often drawn from historical context. In this study we looked at a 10-year
+dataset to ascertain a domain for values what do not have a specific reference source. We would recommend that these domains be published
+in the Data Dictionary. Here are some results.
 
@@
 	\item Many fields have a domain of acceptable values. Often these values are determined from common usage or by examining
-		larger datasets. Unfortunately, these domains of acceptable values are not specified in the Data Dictionary. These fields all tested as
+	larger historical datasets. Unfortunately, these domains of acceptable values are not specified in the Data Dictionary. These fields all tested as
 	compliant with their domain of acceptable values:
@@
 
+	\subsection{Issues with Zip Codes}
 	 Unfortunately some fields did prove to be quite problematic when comparing the values to the domain of legal values.
@@
 
-\begin{figure}[htbp]
+	\begin{figure}[H]
 	  \centering
-	  \includegraphics[scale=0.7]{invalid_incident_zip.png}
+		  \includegraphics[width=\textwidth]{invalid_incident_zip.png}
 		  \caption{Invalid incident-zip by Agency}
@@
 
-A curious case also exists when examining the two nearly identical fields - police\_precincts and police\_precinct.
+		\subsubsection{Case Study: Noise Complaints by Zip Code}
+		\textbf{Scenario:} The NYC Office of Nightlife (ONL) wants to know ``What are the top 10 zip codes for Noise Complaints (all 8 types) over the last two years?''
+		The goal is to assess the impact of the recent NYC effort looking to promote a safe and vibrant nightlife scene in NYC while seeking to ease
+		strained relations between bar and club owners.
 
-\subsection{Logical inconsistencies }
+		It may come as a surprise to non-NYC residents that dancing at bars and restaurants has been illegal since 1926 (during prohibition), and was only just repealed in 2018.
 
-\subsection{Accuracy and precision}
+		On the surface this would seem like a simple effort.  The NYC Open Data Portal allows for selection of a timeframe (2022-2023), complaint-type
+		begins with ``Noise'', and group by Zip code, sorted descending by count.  You can do this analysis and grouping right inside the NYC Open Data Portal. \textit{Voila!}
 
-\subsection{Redundant data}
+	\begin{figure}[H]
+	    		\centering
+	    		\includegraphics[scale = 0.6]{zipcode_casestudy.png}
+	    		\caption{Top 10 zip codes for Noise complaints (Zip Codes)}
+	    	\label{fig:casestudy1-zipcodes}
+		\end{figure}
+
+		However, this analysis uses the zip\_codes field, one of the (six) computed fields that has shown to have validity problems. If we repeat the analysis with the 
+		incident\_zip field instead of the zip\_codes field, we obtain a different set of zip code results:
+
+		\begin{figure}[H]
+	    		\centering
+	    		\includegraphics[scale = 0.6]{incident_zip_casestudy.png}
+	    		\caption{Top 10 zip codes for Noise complaints(Incident Zip)}
+	    	\label{fig:casestudy1-incident-zip}
+		\end{figure}
+
+	Let's subject these two data sets to validation against the US Postal Service zip code database.
+
+	\begin{table}[H]
+	\centering
+	\footnotesize
+	\caption{Comparison of Top Ten Zip Codes Lists}
+	\begin{tabular}{lrr@{\hskip 0.5cm}|@{\hskip 0.5cm}lrr}
+	\toprule
+	\multicolumn{3}{c}{\textbf{zip\_codes}} & \multicolumn{3}{c}{\textbf{incident\_zip}} \\
+	\cmidrule(r){1-3} \cmidrule(l){4-6}
+	\textbf{Zip Code} & \textbf{Count} & \textbf{Valid?} & \textbf{Zip Code} & \textbf{Count} & \textbf{Valid?} \\
+		\midrule
+		11275 & 104,556 & FALSE & 10466 & 104,562 & TRUE \\
+		12420 & 27,503 & TRUE & 10023 & 27,972 & TRUE \\
+		12428 & 26,564 & TRUE & 10031 & 25,548 & TRUE \\
+		10935 & 25,508 & FALSE & 10457 & 25,066 & TRUE \\
+		10934 & 23,448 & FALSE & 10453 & 24,752 & TRUE \\
+		10931 & 22,381 & TRUE & 10456 & 24,751 & TRUE \\
+		10930 & 22,121 & TRUE & 10452 & 22,527 & TRUE \\
+		17613 & 21,963 & FALSE & 10025 & 21,705 & TRUE \\
+		10936 & 21,707 & FALSE & 10458 & 21,689 & TRUE \\
+		11606 & 21,435 & FALSE & 10032 & 20,622 & TRUE \\
+		\bottomrule
+	\end{tabular}
+	\end{table}
+
+	As indicated, six out of ten zip\_codes are invalid, which corresponds closely with what is observed in the overall dataset (58\%). Whereas the incident\_zip
+	field is completely valid, again in-line with the overall incident\_zip validation percentage (99.04\%)
+	Perhaps more curious than the differing zip codes on the two lists, is the fact that the numerical counts are nearly identical;  the overall counts are quite close. This is curious since
+	one field is computed while the other is via data entry; the computed protocol is again the suspect.
+
+	\subsection{Issues with Police Precincts}
+	A curious case also exists when examining the two nearly identical fields - police\_precincts and police\_precinct. Both of those fields are among
+	the ``computed'' fields in the dataset. 
+	Using  \href{https://www.nyc.gov/site/nypd/bureaus/patrol/precincts-landing.page}{NYPD Precinct listings} it's possible to determine the valid
+	police precincts; they are generally numeric or have numeric representations in the dataset.
+
+	However, what we find is that both fields police\_precinct and police\_precincts  have  35\% invalid entries.
+	Unfortunately, they're not the same invalid entries.
+
+	For the police\_precincts field, there are 2,171,864 invalid entries (35\%)., a consequential number of errors. Similarly for the police\_precinct field,
+	 there are 2,171,778 invalid entries ( also 35\%), however they are not the same counts for the two similar fields.
+
+	The top ten (by count) of invalid precincts are:
+
+	\begin{table}[ht]
+	\footnotesize
+	\centering
+		\begin{tabular}{ccccc}
+		\toprule
+		\textbf{Invalid Precinct} & \textbf{police\_precinct count} & \textbf{police\_precincts count} & \textbf{Delta}\\
+		\midrule
+			62 & 151808 & 151720 & 88 \\
+			72 & 148881 & 148796 & 85 \\
+			67 & 133171 & 133115 & 56 \\
+			64 & 111236 & 111221 & 15 \\
+			60 & 105252 & 105317 & -65 \\
+			65 & 103334 & 103306 & 28 \\
+			53 & 100097 & 100074 & 23 \\
+			73 & 97746 & 97755 & -9 \\
+			68 & 94204 & 94254 & -50 \\
+			66 & 93654 & 93848 & -194 \\
+		\bottomrule
+		\end{tabular}
+	\caption{Comparison of the fields police\_precinct and police\_precincts}
+	\label{tab:comparison-precincts-diff}
+	\end{table}
+
+	A graphic representation of the invalid police\_precincts plotted by Agency show a near perfect match to the
+	distribution of the entire 6.4 million SRs. Note that the ``big six'' Agencies account for 90\% of these invalid precinct errors, identical
+	to the overall distribution of all SRs. This strongly suggests that this is a systemic problem, almost certainly
+	caused by the computation method applied for these two fields.
+
+	\begin{figure}[H]
+	  \centering
+		  \includegraphics[width=\textwidth]{invalid_police_precincts.png}
+		  \caption{Invalid NYPD precincts by Agency}
+		  \label{fig:invalid_police_precincts_zip}
+	\end{figure}
+
+	\subsection{Issues with Community Boards}
+	Community Boards are an important aspect of NYC government. Community boards are the most local, grassroots form of City government, 
+	and the connection between communities and elected officials and City agencies. They are organized by the five Boroughs of NYC and represented
+	in this dataset as ``\#\#-Borough'', e.g. ``10 Bronx''. As such, they play an important role in the quality of life for all New Yorkers. There are 59 community boards 
+	throughout the City. There are 12 Community Districts in the Bronx, 18 in Brooklyn, 12 in Manhattan, 14 in Queens and 3 in Staten Island. The Community Boards are
+	used frequently as a way to measure City services throughout the five boroughs, so correctness is important. NYC Borough Presidents in particular are focused
+	on the services offered at the Community Board level.
+
+	In the 2022-2023 dataset there are 27,276 invalid community\_board entries which represents 0.43\% of non-blank data. There are a total of 12 different invalid community boards.
+	These include:
+
+		\begin{table}[H]
+		\centering
+		\small
+		\caption{Top Ten Invalid 'community\_board' Values}
+		\begin{tabular}{rlr}
+		\toprule
+		\textbf{Rank} & \textbf{Invalid CB} & \textbf{Count} \\
+			\midrule
+				1 & 64 MANHATTAN & 7144 \\
+				2 & 83 QUEENS & 5133 \\
+				3 & 55 BROOKLYN & 3327 \\
+				4 & 81 QUEENS & 3180 \\
+				5 & 80 QUEENS & 2407 \\
+				6 & 26 BRONX & 2364 \\
+				7 & 28 BRONX & 1128 \\
+				8 & 82 QUEENS & 1054 \\
+				9 & 95 STATEN ISLAND & 515 \\
+				10 & 27 BRONX & 514 \\
+			\bottomrule
+		\end{tabular}
+		\end{table}
+
+		The distribution of invalid Community Boards by Agency is not consistent with the overall SR Agency distribution. This indicates that there are likely
+		specific issues at some key Agencies, in this case Taxi \& Limo Commission (TLC), Parks \& Recreation, etc. It's not a large error, but the division of services
+		(and complaints) by Community Board is a well-tracked measure.
+
+		\begin{figure}[H]
+	  \centering
+		  \includegraphics[scale = 0.65]{invalid_community_boards.png}
+		  \caption{Invalid Community Board by Agency}
+		  \label{fig:invalid_community_boards}
+	\end{figure}
+
+	\subsection{Issues with Community Districts}
+	Another of the computed fields is community\_districts. Community Districts are the boundaries for the Community Boards, but unlike how Community Boards are
+	an arm of the government of New York City, the Community District is used by the Department of City Planning (DCP) for purposes of environmental, socio-economic, and
+	demographic purposes. DCP is the NYC''s primary land use agency and is instrumental in the City's physical framework.  It is a geographical division rather
+	than a local government division. And like Community Boards, the Community Districts are a frequently used means to breakdown zoning, housing, community facilities,
+	and waterfront/open/public spaces. It is also a well-tracked measure.
+
+	Due to how the community\_district data is formatted, it is not possible to establish validity. However, it is possible to determine that the dataset contains
+	72 unique entries, while there are only 59 valid Community Districts.
+
+\section{Logical inconsistencies }
+
+closed before created (negative duration)
+closed exactly as created (zero duration)
+closed in the future
+
+\subsubsection Case Study: Homeless Person Assistance
+
+due\_date before created
+long resolution\_action\_updated\_date
+
+\section{Accuracy and precision}
+There is one area in particular where the question of precision vs. accuracy immediately arises, and that is with the Latitude and Longitude fields.  Both the 
+Longitude and Latitude fields are expresses as a 14-decimal number, e.g. 40.86769186022511 (also the Location field which is a straight concatenation of
+latitude and longitude). Given that 1 degree of latitude at the equator is equal to 111.044736 kilometers, the ``1'' at the end of that number represents
+approximate 1.1104 nanometer or 1/1,000,000,000 of a meter. A DNA molecule is approximately 2nm in width.
+
+Clearly the representation of the Latitude and Longitude fields are a classic case of 14-digit precision, 
+but limited accuracy. It is more likely that the Lat/Long values are accurate to the 5\textsuperscript{th} decimal place, about 3.64 feet; even that might be
+overstating things.
 
@@
 
-\subsection{Case Study I - Studying Noise Complaints by NYC Zip Code}
-
-Invalid values?
-Logical inconsistencies? Concerning patterns?
-Redundant columns?
-
-Invalid data: closed data earlier than creation date;
-latitude/longitude not in NYC; invalid zip code;
-
-Inconsistent data: upper/lower cases; extra spaces; 
-open text entries: provide enough major categories and sub categories;
-
-Inefficiently stored data: lat/long versus (lat, long);
-intersection street 1 versus cross street 1;
-borough could be inferred from zip code;
-full name of agency not needed
+\section{Redundant \& Duplicate data}
+Shrinking file size by deleting these 11 redundant fields:
+     1 - agency\_name 
+     2 - park\_borough 
+     3 - borough\_boundaries 
+     4 - location 
+     5 - intersection\_street\_1 
+     6 - intersection\_street\_2 
+     7 - police\_precinct 
+     8 - duration 
+     9 - postClosedUpdateDuration 
+     10 - translated\_borough\_boundaries 
+     11 - zip\_codes 
+Original size: 3.2 Gb 
+Size after removing redundant columns: 2.8 Gb 
+Potential size reduction: 395.3 Mb or 12.3
 

### 2024-05-27 manuscript/311 manuscript.tex
Updated Tex file and accompanying png files

@@
 
+\hyphenpenalty=1000
+
 \begin{abstract}
@@
 creation of the \href{https://opendata.cityofnewyork.us}{NYC Open Data
-  portal}, which today hosts an impressive array of 2.700 datasets
+  portal}, which today hosts an impressive array of 2700 datasets
 from 80 different city agencies. This resource has become invaluable
-for researchers in various fields as well an enabling a high degree of local government transparency
+for researchers in various fields as well an enabling local government transparency
 
@@
 understanding of urban mobility and infrastructure
-\citep{gerte2019understanding}.
-
-NYC's Open Data initiative not only
+\citep{gerte2019understanding}. NYC's Open Data initiative not only
 exemplifies commitment to transparency and public engagement but also
 illustrates how open data can be a powerful tool in addressing complex
-urban challenges.
+urban challenges. Examples of some of the more popular NYC datasets include:
+
+\begin{itemize}
+	\item NYC restaurant violations
+	\item Popular baby names
+	\item Mapping of car crashes involving pedestrians
+	\item Mapping of sidewalk widths in NYC 
+	\item Visualization of NYC High School and College enrollment
+	\item An interactive dashboard to filter and view charges for jail inmates
+	\item Location of City-wide free Internet access points
+	\item ...and 220 other complaint types. 
+\end{itemize}
 
@@
 critical role of collective data curation efforts in managing and
-responding to the crisis.
-
-Collectively, these studies not only highlight the multifaceted nature of data curation but also emphasize
+responding to the crisis. Collectively, these studies not only highlight the 
+multifaceted nature of data curation but also emphasize
 its indispensable role in enhancing the applicability and value of
@@
 including data validity, consistency, and curation efficiency. 
-We illustrate these points with live examples drawn from our processing 
-of the 311 data.
+We illustrate these points with live examples drawn from our processing of the 311 SR data.
 
@@
 
-The remainder of this paper is organized as follows. 
-Section~\ref{sec:data} offers a brie review of the 
-history and current status of the 311 requests data. In 
-Section~\ref{sec:issues}, we identify and list potential issues 
-impacting data quality and curation efficiency. 
-Section~\ref{sec:improve} is dedicated to providing 
-actionable suggestions for mitigating or resolving these identified issue. 
-Following this, Section~\ref{sec:protocol} outlines a series of general 
-principles for the release of open data, drawing from our findings. 
-The paper concludes with a discussion in Section~\ref{sec:disc}, encapsulating 
-the key insights and implications of our research.
-
-\section{NYC 311 Service Requests Data} \label{sec:data}
+The paper is organized as follows:
+Section~\ref{sec:history} offers a brief review of the history of the 311 system. In 
+Section~\ref{sec:trends} we take a look at long-term trends presented via a 10-year analysis.
+Section~\ref{sec:issues} offers a general discussion of data cleansing issues
+impacting data quality and curation efficiency. Section~\ref{sec:Structural} examines
+the technical dimensions of the dataset, the structural issues. In Section~\ref{sec:datatypes} we examine the data fields
+for compliance with the stated data type as found in the Data Dictionary. Section~\ref{sec:blanks} looks
+at the data fields as regards missing, blank, or N/A entries. Section~\ref{sec:domain} explores
+how data fields comply with a domain of legal or acceptable values. Section~\ref{sec:inconsistencies}
+deals with the important issues surrounding logical inconsistencies and concerning patterns in the data. Section~\ref{sec:precision} 
+explores the age-old issue of precision versus accuracy. Section~\ref{sec:duplicates}
+identifies duplicate and redundant data fields. Section~\ref{sec:improvements} provides 
+actionable suggestions for mitigating or resolving identified issues. Following this, Section~\ref{sec:protocol} outlines a series of general 
+principles for the release of open data, drawing from our findings. The paper concludes with a discussion 
+in Section~\ref{sec:discussion}, encapsulating the key insights and implications of our research.
+
+A note on naming conventions: dataset field names are typically named using ``snake\_case'' where
+each word in the variable name is written in lowercase, and words are separated by underscores,
+e.g created\_date, complaint\_type, zip\_code, etc..
+
+\section{NYC 311 System History} \label{sec:history}
 
@@
 maintenance requests. The evolution of this system can be traced from
-its initial implementation as a simple inquiry channel to a
+its initial implementation as a simple inquiry channel via telephone to a
 comprehensive data management system that handles millions of requests
-annually. Key milestones in its development include:
-
-\begin{itemize}
- 	\item 2003 - NYC 311 system  goes live with phone-based call center only.
+annually. Each inquiry is logged as a complaint in the 311 system.
+Key milestones in the 311 system development include:
+
+\begin{itemize}
+ 	\item 2003 - NYC 311 system goes live with a phone-based call center only.
    	\item 2009 - 311 online \& mobile apps launched. Today 60\% of all Service Requests come from mobile \& online channels.
-   	\item 2019 - Major software upgrade of 311 system with enhanced capabilities and cloud-based resilience. 	
-      	\item 2020 - In August, driven by the COVID pandemic, monthly 311 Service Requests hit an all-time high of 348,463. Seven of the Top 10 busiest days occur in 2020.
+   	\item 2019 - Major software upgrade of 311 system with enhanced capabilities, dynamic load-balancing, and cloud-based resilience. 	
+      \item 2020 - In August 2020, driven by the COVID pandemic, monthly 311 Service Requests hit an all-time high of 348,463. Seven of the Top 10 busiest days of all time occur in 2020.
       \item 2021 - 311 System is expanded to into the MTA's city subway system, the largest expansion since its inception.
@@
 Today, the NYC 311 data system is a robust platform that manages a
-vast array of urban living-related inquiries. As of 2024, the system handles approximately 3.5 million service 
-requests per year, covering categories such as noise complaints, illegal parking, heat/hot water complaints, abandoned vehicles,
-and unsanitary conditions. The data, managed through a tailored application built on the Microsoft Dynamics platform, is a valuable resource for city
-administration and policy-making. It is publicly accessible through the 
-\href{https://opendata.cityofnewyork.us/}{NYC Open Data Portal}. The online Portal provides access
+vast array of urban living-related inquiries. The system handles over 3 million service 
+requests (SRs) per year, encompassing such complaints as street noise, illegal parking, heat/hot water, abandoned vehicles,
+and unsanitary conditions. The data, managed through a tailored application built on the Microsoft Dynamics and Azure platform, is a valuable resource for city
+administration and policy-making. The 311 data is publicly accessible through the 
+\href{https://opendata.cityofnewyork.us/}{NYC Open Data Portal} which provides access
 to all the data sets (approx. 2700) and provides a convenient method for querying, grouping, aggregating, geo-mapping, visualization, and exporting
-results. This open data initiative enables not only governmental
-transparency but also empowers researchers, civic developers, and the
-general public. Examples of its use include:
-
-\begin{itemize}
-	\item NYC restaurant violations
-	\item Popular baby names
-	\item Mapping of car crashes involving pedestirans
-	\item Mapping of sidewalk widths in NYC 
-	\item Visualization of NYC High School and College enrollment
-	\item Interactive dashboard to filter and view charges for jail inmates
-	\item Internet access points.
-	\item ...and 220 different complaint types. 
-\end{itemize}
+results. This open data initiative enables not only governmental transparency but also empowers 
+researchers, civic developers, and the general public.
 
@@
 
-Despite its success, the system faces many challenges such as:
+Despite its success, the system constantly faces challenges such as:
+
 \begin{itemize}
 	\item Data timeliness, accuracy, and consistency
-	\item Difficulties correlating data over long time periods, e.g. over a 10-year period
-	\item Data anonymization and handling Personally identifiable information (PII)
+	\item Difficulties correlating data over long time periods, e.g. over a 10-year period, e.g. Agency name changes, new complaint types.
+	\item Data anonymization and handling of Personally identifiable information (PII)
 	\item Integration with stand-alone systems at selected NYC Agencies
-	\item Managing API usage, authentication, and load for 3rd party users
+	\item Managing API usage, authentication, and load for 3\textsuperscript{rd} party users, of which there are many
 	\item Incorporating ongoing technology changes and upgrades 
-	\item Understanding the role of chatbots and AI to provide online customer service
-\end{itemize}
-
-These (and other) challenges are constantly being addressed by the various agency open data managers
+	\item Understanding the role of chatbots and AI to provide online customer service, along with other emergent technologies
+\end{itemize}
+
+These and other challenges are continually addressed by the various Agency open data managers
 as well as the \href{https://www.nyc.gov/content/oti/pages/}{NYC Office of Technology and Innovation (OTI)} which
@@
 Many of the underlying datasets are updated daily, usually running 1-2 days behind from creation.
-However, a number of datasets are updates monthly (such as restaurant inspections), and some are
+However, a number of datasets are updated monthly (such as restaurant inspections), and some are
 not updated for one or more years (such as community board boundaries). This poses problems when 
-merging these dataset with one another, as would be expected. Typically, the data must be
-exported and manipulated in another program
+merging dataset with others, as would be expected. Typically, the data must be
+exported and manipulated in another, external program. The authors found this to be the case with 
+the large 311 dataset employed.
 
@@
 via Service Level Agreements (SLAs). However, enforcement of those SLAs is often challenging. The 
-result is that often data accuracy and logical consistency is occasionally compromised.
+result is that often data accuracy and logical consistency is occasionally compromised. However, the Open Data
+Team is quite responsive to inquires from users. The authors have provided many such 
+suggestions during the course of this research effort, and the most current data shows improvements
+in many of the areas highlighted. Questions and suggestion  can be provide here: 
+\href{https://opendata.cityofnewyork.us/engage/}{Contact Us}.
 
 The NYC 311 organization, part of OTI, operates a highly tailored and sophisticated case
-management system based, as noted, on Microsoft Dynamics. However, when the original 311
-system was developed in 2003, a decision was made to allow select NYC Agencies to
-maintain their own local computer systems, such as the mainframe systems at the NYC Department
-of Sanitation (DSNY) and others. And when the 311 system was completely upgraded in 2019, those
-same integration were `honored` in order to reduce the scope of the project. (Perhaps in
-retrospect, while a good budgetary decision, perhaps not such a good technology decision.)
-
-The result was that at least nine NYC Agencies have local systems that are integrated either via
-an API or custom code. These Agencies include:
+management. However, one constraint was particularly challenging during the recent complete overhaul
+of the software and hardware in 2019. When the original 311system was developed in 2003, 
+a decision was made to allow select NYC Agencies to maintain their own local computer systems, 
+such as the mainframe systems at the NYC Department of Sanitation (DSNY) and others. 
+And when the 311 system was completely upgraded in 2019, those
+same integrations were honored in order to reduce the scope of the project. Perhaps in
+retrospect, while a good budgetary and policy decision, it was perhaps not such a good technology decision.
+
+At least nine NYC Agencies have local systems that are integrated directly to the 311 system either via
+an API or custom code. These include:
 
@@
 
-One result of these integrations, and other Agencies that use the 311 APIs, is that some of the data accuracies
-and data consistency checks inherent in the 311 system are not applied uniformly, and in some cases not at all.
-Errors often creep in from other Agency systems and are not easily corrected in the 311 application itself,
-but rather require correction at the source Agency.
-
-One area that will almost certainly remain an ongoing challenge is the incorporation of new software, hardware,
-and cloud services which are constantly emerging.  Telephonic integration recently was reconfigured due to a new
-telephone vendor. Desktop software at the 311 Agency are subject to the same upgrades that home users feel. And while the 
-311 system was designed to be easily configured with new features, some emerging business issues as well as
-technologies simply require additional software engineering efforts to accommodate.
-
-One area in particular which is heavily influencing customer outreach for the 311 system is the use of
-online chatbots. Increasing these are being influenced by Artificial Intelligence (AI) advances. It is a given that
+One unfortunate side-effect from these integrations, and other Agencies that use the 311 APIs, is that some of the data consistency
+checks which are inherent in the 311 system are not applied uniformly, and in some cases not at all, to the integrated data streams.
+Errors often creep in from other Agency systems and are not easily corrected or detected in the 311 application itself,
+but rather require correction at the source Agency; challenging.
+
+One area that will almost certainly remain an ongoing issue is the ongoing incorporation of new software, hardware,
+and cloud services which are constantly changing while new approaches emerge. The 311 telephonic integration recently was reconfigured due to a new
+telephonic vendor. Desktop software at the 311 Agency are subject to the same upgrades that home and business users experience. And while the 
+311 system was designed to be easily configured and updated with new features, some emerging business issues as well as
+technologies nonetheless require additional software engineering efforts to accommodate, along with the concomitant budgets.
+
+One area in particular which is heavily influencing customer outreach for the 311 system is the appropriate use of
+online chatbots and assistants. Increasing these features are influenced by artificial intelligence (AI) advances. It is a given that
 a number of citizen's interactions with City government could be handled accurately and quickly by AI-driven
-applications. How to do that while preserving customer privacy and anonymization are challenging issues.
+applications thus improving timeliness and accuracy. But how to do that while preserving customer privacy and anonymization
+and user engagement are challenging issues.
 
@@
 \begin{itemize}
-	\item Advice on shelters during hurricanes and winter blizzards
-	\item Countless inquires during the COVID pandemic (testing centers, vaccine sites, mobile vaccine clinics, etc.)
-	\item Enforcement of standards between landlords and tenants.
+	\item Providing advice on shelters during hurricanes, heat emergencies, and winter blizzards
+	\item Handling countless inquires during the COVID pandemic (testing centers, vaccine sites, mobile vaccine clinics, etc.)
+	\item Enforcement of standards between landlords and tenants
 	\item Re-allocation of NYC's yellow taxi routes based upon an insightful analysis conducted by the Taxi \& Limousine Commission (TLC)
-	\item Improved responsiveness for City Agencies with direct client-facing responsibilities such as Parks \& Rec, Dept of Transportation, DSNY, etc.
+	\item Improved responsiveness for City Agencies with regards to direct client-facing responsibilities such as Parks \& Rec, Dept of Transportation, DSNY, etc.
 	\item Increased 311 SRs related to Homelessness resulted in the launch of the City's Homeless Outreach \& Mobile Engagement Street Action Teams (HOME-STAT)
 	which includes street canvassers and increases in homeless housing facilities.
-	\item A significant increase in graffiti complains in 2017 led directly to the `Graffiti-Free NYC'' campaigns and increased funding for removal of graffiti in public spaces
-	\item Enhanced geospatial data regarding the location of incidents, the responding Agency, and actions taken.
+	\item A significant increase in graffiti complains in 2017 led directly to the `Graffiti-Free NYC'' campaigns and increased funding for removal of graffiti in public spaces.
+	\item Enhanced geospatial data regarding the location of incidents, the responding Agency, and actions taken across a myriad of complaint\_types
 	\item Establishment of Agency SLAs in regard to certain complaint types, such as residential noise, homelessness person assistance, street light outages, 
@@
 The NYC 311 service exemplifies the dynamic nature of urban data
-management and its critical role in modern governance. It provides an
-invaluable resource for data scientists to practice data science
-activities, including data curation.
-
-We extracted  311 Service Request (SR) data from the period of CY 2022-2023, a dataset consisting of
-6.4 million records. This dataset serves as the primary source for the analysis effort. This is a large dataset which enables 
+management and its critical role in modern governance. In addition it provides an
+invaluable resource for data scientists to practice data science activities, including data curation.
+
+We extracted 311 SR data for the period of CY 2022-2023, a dataset consisting of
+6.4 million records. That dataset serves as the primary source for the analysis effort. This is a large dataset which enables 
 observation of some rarely seen events, data inaccuracies, and logical inconsistencies. It provides enough data to examine trends and data outlier analysis. 
-We query the data directly from the NYC Open Data Portal and then export the data in CSV format for analysis in a custom written R programs.
-
-Additionally, we extracted two longer-term datasets, a 5-year look (2019-2023) and a 10-year look (2014-2023) ; some 14  and 27 million rows respectively, albeit with just a few columns. 
-These larger datasets afford an opportunity for long-term trend observations, confirming the analysis using the 2-year dataset. The 2022-2023 dataset
-was used exclusively for the detailed, field-level analysis that comprises the majority of this effort.
-
-\section{10-year Analysis \& Trends} \label{sec:history}
+We queried the data directly from the NYC Open Data Portal and then exported the data in CSV format for analysis using custom written R programs.
+
+Additionally, we extracted two longer-term datasets, a 5-year look (2019-2023) and a 10-year look (2014-2023) ; some 14 and 27 million rows respectively. 
+These larger datasets afford an opportunity for long-term trend observations, confirming and supporting the analysis using the 2-year dataset. The 2022-2023 dataset
+was used exclusively for the detailed, field-by-field level analysis which comprises the majority of this effort.
+
+\section{10-year Analysis \& Trends} \label{sec:trends}
 
@@
 observe certain system performance characteristics over a longer time-frame . 
-During that time-frame, the number of 311 SRs grew 50\% as shown in the chart below which shows the yearly SR counts.
-
-\begin{figure}[htbp]
+During that time-frame, the number of 311 SRs grew 50\% as shown in the chart below showing yearly SR counts.
+
+\begin{figure}[H]
   \centering
   \includegraphics[width=\textwidth]{Yearly.png}
-  \caption{SR counts by year for a 10-year period}
+  \caption{Yearly SR counts over 10-year period}
   \label{fig:yearly-counts}
@@
   \includegraphics[width=\textwidth]{Monthly.png}
-  \caption{SR counts by month for a 10-year period}
+  \caption{Monthly SR counts over 10-year period}
   \label{fig:monthly-counts}
@@
 
-A quick look at the corresponding  daily tend over that same 10-year period is also interesting, as it shows both significant spikes
-and significant dips; a noisy graph. While the spikes on 2020-08-04 (24,415 SRs) and 2020-08-05 (19,560 SRs) are COVID related, we are unable to 
-ascertain any reason for the significant spike on Saturday, 2023-09-29 where 17,962 SRs were created.
+A quick look at the corresponding  daily tend over a 5-year period (2019-2023) shows both significant spikes
+and significant dips; a noisy graph. While the spikes on 2020-08-04 (24,415 SRs) and 2020-08-05 (19,560 SRs) are certainly COVID related, we were unable to 
+ascertain any reason for the significant spike on Saturday, 2023-09-29 when 7,962 SRs were created.
 
@@
   \includegraphics[width=\textwidth]{Daily.png}
-  \caption{SR counts by day for a 10-year period}
+  \caption{Daily SR counts over 5-year period}
   \label{fig:daily-counts}
@@
   \includegraphics[width=\textwidth]{Calendar-Month.png}
-  \caption{SR counts by calendar month for a 10-year period}
+  \caption{Calendar Month SR counts over a 10-year period}
   \label{fig:calendar-months-counts}
@@
 
-Of note, the lowest SR counts-per-day occur on Christmas  period (+/- a day), in 5 of the Bottom 10 SR counts-per-day. The summer moths and early
-fall months make up all of the Top 10 SR counts-per-day. As previously noted, the all-time daily high for SRs occurred on Tuesday, 2020-08-04 with the 10-year
-low number occurs on Christmas Day in 2016, a Sunday. Note also that the difference between the peak day (24,415) and the minimum day (2965) varies
-by a factor of 8X, and the aveage daily count of 7489 -- all of which creates a challenging  system capacity design issue.
-
-\begin{table}[htbp]
-	\begin{minipage}[t]{0.35\textwidth} % Adjust the width of the minipage as needed
-	\centering
-	\small % Adjust the font size
-	\caption{Top 10 days}
-	\label{tab:top_ten_counts}
-		\begin{tabular}{@{}lll@{}}
+Of note, the lowest SR counts-per-day occur during the Christmas  holiday period encompassing fully 5 of the Bottom 10 SR counts-per-day. The summer months and early
+fall months make up all of the Bottom 10 SR counts-per-day. As noted, the all-time daily high for SRs occurred on Tuesday, 2020-08-04 while the 10-year
+low count occurs on Christmas Day in 2016, a Sunday. Note also that the difference between the peak day (24,415) and the minimum day (2965) varies
+by a factor of 8X, and well exceeds the average daily count of 7489; all of which creates a challenging  system capacity and performance issue.
+
+\begin{table}[H]
+    \centering
+    \small
+    \begin{tabular}{@{}lll|lll@{}}
         \toprule
-			Date & Count & Rank \\
+        \multicolumn{3}{c|}{Top 10 Days} & \multicolumn{3}{c}{Bottom 10 Days} \\
+        \cmidrule(r){1-3} \cmidrule(l){4-6}
+        Date & Count & Rank & Date & Count & Rank \\
         \midrule
-			2020-08-04 & 24415 & 1 \\
-			2020-08-05 & 19560 & 2 \\
-			2023-09-29 & 17962 & 3 \\
-			2020-07-05 & 16916 & 4 \\
-			2020-06-21 & 15883 & 5 \\
-			2020-06-20 & 15825 & 6 \\
-			2020-07-04 & 15794 & 7 \\
-			2021-09-02 & 15205 & 8 \\
-			2020-06-28 & 14057 & 9 \\
-			2021-10-28 & 13575 & 10 \\
+        2020-08-04 & 24415 & 1 & 2014-12-25 & 2965 & 1 \\
+        2020-08-05 & 19560 & 2 & 2015-12-25 & 3247 & 2 \\
+        2023-09-29 & 17962 & 3 & 2014-04-20 & 3287 & 3 \\
+        2020-07-05 & 16916 & 4 & 2015-12-26 & 3405 & 4 \\
+        2020-06-21 & 15883 & 5 & 2016-11-06 & 3419 & 5 \\
+        2020-06-20 & 15825 & 6 & 2014-07-04 & 3455 & 6 \\
+        2020-07-04 & 15794 & 7 & 2014-12-27 & 3534 & 7 \\
+        2021-09-02 & 15205 & 8 & 2014-09-14 & 3563 & 8 \\
+        2020-06-28 & 14057 & 9 & 2014-02-02 & 3565 & 9 \\
+        2021-10-28 & 13575 & 10 & 2016-12-25 & 3587 & 10 \\
         \bottomrule
     \end{tabular}
-	\end{minipage}\hfill % Adjust the space between the tables
-	\begin{minipage}[t]{0.35\textwidth} % Adjust the width of the minipage as needed
-	\centering
-	\small % Adjust the font size
-	\caption{Bottom 10 days}
-	\label{tab:bottom_ten_counts}
-		\begin{tabular}{@{}lll@{}}
-			\toprule
-			Date & Count & Rank \\
-			\midrule
-			2014-12-25 & 2965 & 1 \\
-			2015-12-25 & 3247 & 2 \\
-			2014-04-20 & 3287 & 3 \\
-			2015-12-26 & 3405 & 4 \\
-			2016-11-06 & 3419 & 5 \\
-			2014-07-04 & 3455 & 6 \\
-			2014-12-27 & 3534 & 7 \\
-			2014-09-14 & 3563 & 8 \\
-			2014-02-02 & 3565 & 9 \\
-			2016-12-25 & 3587 & 10 \\
-			\bottomrule\\
-		\end{tabular}
-	\end{minipage}
+    \caption{Top 10 and Bottom 10 Days}
+    \label{tab:combined_counts}
 \end{table}
@@
 when the \#-of-SRs-per-day metric is computed by adjusting for the differing number of days for each month -- the
-busiest month is shown to be June (followed by July and September). And similarly the quietest months are April, March, and 
+busiest month is shown to be June (followed by July and September). Similarly the quietest months are April, March, and 
 February. In general, the winter months see fewer SRs while the summer months show 
-an increase in activity, as previously noted.
-
-\begin{table}[ht]
+an increase in activity.
+
+\begin{table}[H]
     \centering
@@
     \end{tabular}
-    \caption{Monthly counts sorted by counts-per-day}
+    \caption{Monthly counts ordered by count-per-day}
     \label{tab:monthly_counts}
@@
 
-Two areas that are important in this data cleansing analysis effort are the responsible NYC Agency, and the actual type of complaint.
-Given both the size of the City of New York government, it's typically necessary to identify the responsible Agency as well as the type of complaint in order to 
-clarify the responsible area and assist in troubleshooting any discrepancy.
-
-In discovering a data error, we typically observe two trends; either a single (or a small few) Agencies are responsible (such as Dept of Transportation), or it is a system-wide issue that foolows the general distribution
-of  Service Requests (SRs) across the system. Let's take a brief look at how this plays out across the system, so when we identify a
-issue, it will be useful to see how that issue flows back to a specific Agency, and thus establish responsibility.
+Two important fields in this analysis effort are the responsible NYC Agency, and the type of complaint.
+Given both the size of the City of New York government, it's necessary to identify the responsible Agency as well as the type of complaint in order to 
+clarify the responsible party and assist in troubleshooting discrepancies.
+
+When we discovered a data error, we typically observe two trends; either a single (or a few) Agencies are responsible (such as Dept of Transportation), or it is a system-wide issue that follows the general distribution
+of  Service Requests (SRs) across the system. Thus we can often identify if it is an Agency specific problem, or a systemic effect affecting all Agencies.
 
 Although the 311 SRs involve 16 different agencies, there are the ``big six'' Agencies that comprise
-90\% of the complaints. These six are (in order):
+90\% of the complaints. These six are (in order of \% of complaints):
 
@@
 
-\begin{figure}[htbp]
-  \centering
-	  \includegraphics[width=\textwidth]{SRs_by_Agency.png}
+\begin{figure}[H]
+  \centering
+	  \includegraphics[width=\textwidth]{SRs\_by\_Agency.png}
 	  \caption{SR counts by Agency and Cumulative Percentage}
-	  \label{fig:SR_counts_by_Agency}
-\end{figure}
-
-It is also interesting to note the distribution of the type of complaints in the 311 SR dataset; it is heavily skewed to  a select few complaints.
-
-\begin{figure}[htbp]
+
+\end{figure}
+
+Note the distribution of the complaint\_type; it is heavily skewed to a few select complaints with an exceptionally long tail.
+This graph shows the top 20 complaint\_type(s) out of a total of 220.
+
+\begin{figure}[H]
   \centering
 	  \includegraphics[width=\textwidth]{SRs_by_Complaint_Type.png}
-	  \caption{Top 20 SR Complaints and Cumulative Percentage}
+	  \caption{Top 20 SR complaint\_type and Cumulative Percentage}
 	  \label{fig:SR_complaints}
@@
 The top 20 complaints contain several of the eight different types of ``Noise'' complaints (Residential, Commercial, Street, Helicopter, etc.). These Noise-related
-complaints total 1.435 million in the 2022-2023 dataset, a full 22\% of all complaint types, which is the most frequently occurring type. (Note: Ilegal Parking is 
-second, followed by Heat/Hot Water.) Noise complaints are also handled by NYPD, as is Illegal Parking and Blocked Driveway complaints, which mostly
-accounts for the top ranking of 311 SRs being routed to the NYPD.
-
-There are also some rather curious and unusual Service Requests in the 311 data: 
-tanning, tattooing, trans fat, unsanitary pigeon condition, illegal animal kept as pet, harboring bees/wasps, and radioactive material. 
-New Yorkers certainly live interesting lives.
+complaints total 1.435 million in the 2022-2023 dataset, a full 22\% of all complaint types, which is the most frequently occurring type.
+(Note: Ilegal Parking is second, followed by Heat/Hot Water.) Noise complaints are also handled by NYPD, as is Illegal Parking and Blocked Driveway complaints, which contributes
+significantly to the top ranking of the NYPD as the responsible Agency.
+
+\begin{table}[H]
+    \centering
+    \footnotesize
+    \begin{tabular}{lS[table-format=7]S[table-format=2.2]l} % 'l' for left-aligned, 'S' for siunitx number alignment
+        \toprule
+        \textbf{complaint\_type} & \textbf{Count} & \textbf{Percentage} & \textbf{Agency} \\ 
+        \midrule
+        NOISE - RESIDENTIAL        & 675502 & 10.56 & NYPD  \\ 
+        NOISE - STREET/SIDEWALK    & 300507 &  4.70 & NYPD  \\ 
+        NOISE - COMMERCIAL         & 128892 &  2.02 & NYPD  \\ 
+        NOISE - VEHICLE            & 115956 &  1.81 & NYPD  \\ 
+        NOISE                      & 100413 &  1.57 & DEP   \\ 
+        NOISE - HELICOPTER         &  85345 &  1.33 & EDC   \\ 
+        NOISE - PARK               &  16620 &  0.26 & NYPD  \\ 
+        NOISE - HOUSE OF WORSHIP   &   2757 &  0.04 & NYPD  \\ 
+        \bottomrule
+    \end{tabular}
+    \caption{Noise-related complaints\_type(s) by count with Agency}
+    \label{tab:noisecomplaints}
+\end{table}
+
+There are also some  curious and unusual Service Requests in the 311 data: 
+tanning, tattooing, trans fat, unsanitary pigeon condition, illegal animal kept as pet, harboring bees/wasps, and the author's favorite
+radioactive material. New Yorkers certainly live interesting lives.
 
 The NYC HPD handles about 49\% of what NYPD handles. For  non-NYC readers, 
-HPD is the City agency that manages the  177,569 NYC public housing units as well as monitoring all rental and leased apartment units. HPD collectively
-serves 528,105 people, a population larger than Atlanta or Miami. Hence the large number of Heat/Hot Water and Unsanitary complaints.
-
-\section{Potential Issues} \label{sec:issues}
+HPD is the City agency that manages the 177,569 NYC public housing units as well as monitoring
+all rental and leased apartment units. HPD collectively serves 528,105 people, a population larger than Atlanta or Miami. 
+Hence the large number of Heat/Hot Water and Unsanitary complaints handled by HPD.
+
+\section{Data Cleansing Issues} \label{sec:issues}
 
@@
 	\begin{itemize}
-		\item Mandatory fields: Certain data fields cannot be empty
-		\item Data-types: Certain fields must be of the correct type, e.g. numeric, character, date, 5 numeric digits, etc. Typically these are referenced in a Data Dictionary.
+		\item Mandatory fields: Certain data fields cannot be empty.
+		\item Data-types: Certain fields must be of the correct type, e.g. numeric, character, date, 5 numeric digits, etc. Typically these are identified in a Data Dictionary.
 		\item Domain adherence:  Many data fields must adhere to a specific domain of values, e.g. statuses, state names, zip codes, gender. This includes data
@@
 	\end{itemize}   
-	\item Structural errors to include naming conventions, extra fields no in the Data Dictionary, inconsistent data entry.
-	For example intermixing blanks, space, NA, N/A. and <NA> to indicate the absence of data.
-	\item Redundant, unnecessary, or inappropriate fields
-	\item Logical inconsistencies such as two related fields that violate the nature of that relationship, such as a ``due date'' that is before the ``created date''.
+	\item Structural errors to include naming conventions, extra fields not in the Data Dictionary, or inconsistent data entry.
+	For example intermixing blanks, spaces, NA, N/A, and <NA> all to indicate the absence of data.
+	\item Redundant, unnecessary, or irrelevant fields
+	\item Logical inconsistencies such as related fields that violate the nature of that relationship, e.g. a ``due date'' that is before the ``created date''
 	\item Data standardization. Often free-form entry fields can suffer from inconsistent data structures. Efforts to standardize these fields can greatly assist analysis.
@@
 
-The effort of this analysis effort is to identify the presence of such errors; not to correct them. That effort would be undertaken after an investigation as to the why
-and how such errors came about. Often this requires discussions with subject matter experts and liaison with various NYC Agency open data coordinators
-to identify the original source of the error. For example, an invalid date field that is provided to the NYC Open Data data lake via an Agency integration could
-be the result of selecting the wrong field from the source Agency's system, e.g. due date instead of closed date. In other cases, without a subject matter
-expert, it is not possible to actually determine if the data is inaccurate or not. The R code used to analyze the 311 Service Request dataset identifies
-both obvious inaccuracies as well as potential errors requiring further investigation.
+This analysis effort is to identify the presence of such errors; not to correct them. That effort would be undertaken only after an investigation as to the why
+and how such errors came about, and a discussion as to whether or not it is even an ``error''. Typically this requires discussions with subject matter experts (SMEs)
+ and liaison within the various NYC Agency open data coordinators who can help identify the underlying cause for the error.
+ For example, an invalid date field that is provided to the NYC Open Data data lake via an Agency integration could
+be the result of selecting the wrong field from the source Agency's system, e.g. due\_date instead of closed\_date. In other cases, without a subject matter
+expert, it is not possible to actually determine if the data is inaccurate or not. There are many empty values in the taxi\_company\_borough field. Are those
+entries missing or does the entry in that field only apply to a very small number of complaint\_type(s) or circumstances. Only a SME
+from the Taxi \& Limo Commission (TLC) can provide that answer. The R code used here to analyze the 311 Service Request dataset identifies
+both obvious inaccuracies as well as potential errors requiring further investigation. It is not a conclusive list of errors, but rather mostly identifies
+``a place to look'' for further evaluation and potential action.
 
@@
 \begin{itemize}
-	\item Structural issues with the data. Is it in the expected format?
-	\item Data type correctness
-	\item Missing/blank data
-	\item Invalid values
-	\item Logical inconsistencies and inconsistent patterns in the data
-	\item Accuracy and precision
-	\item Redundant data  
-\end{itemize}
-
-\section{Structural Issues}
+	\item Examining structural issues within the data
+	\item Validating data type correctness
+	\item Identifying missing, blank, or N/A data
+	\item Identifying invalid values
+	\item Exposing logical inconsistencies \& inconsistent or unusual patterns in the data
+	\item Accuracy and precision issues
+	\item Identifying potential redundant or irrelevant data  
+\end{itemize}
+
+\section{Structural Issues}\label{sec:structural}
 
@@
 
-What are the characteristics of the 311 SR data set. Here are some insights:
-
-\begin{itemize}
-	\item 47 columns of data for each row, exportable as a CSV file
-	\item There are four date fields (created, closed, updated, due)
-	\item There are three borough fields; two of which we believe to be duplicates
+Here are some characteristics of the 311 SR data set:
+
+\begin{itemize}
+	\item There are 47 columns of data for each row, exportable as a CSV file.
+	\item There are four date fields (created, closed, updated, due).
+	\item There are three borough fields; two of which we believe to be duplicates.
 	\item Two zip code fields, but not duplicates
@@
 	\item Two Police Precinct fields; not duplicates
-	\item In addition to street/city, there are three other location fields (lat/long), NY State plane, and Block \#
-	\item Incident address (25 Grymes Hill Road) and Street Name (Grymes Hill Road)
-	\item One form tex fielt; resolution description, which supports 934 characters of input
-\end{itemize}
-
-\textbf{Fields not in the Data Dictionary} The 311 SR \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}
- identifies 41 data columns (fields) along with some
-related information for each column. However, when you either download the data or explore it via the online portal,
-you will immediately note that there are 47 fields. The additional six fields are not addressed in the Data Dictionary, and show up on the Column Manager
-page as ``@computed\_region\_xxxx\_xxxx''. To some extent one can infer what the these @computed\_fields are, but it is not possible to know
-for certain. These fields are:
+	\item In addition to street/city, there are three other location fields: lat/long, NY State plane, and Block \#
+	\item incident\_address and street\_name, e.g. incident\_address, 25 Grymes Hill Road and street\_name (Grymes Hill Road)
+	\item One free-form text field, resolution\_description, which supports 934 characters of input, including commas and special characters
+\end{itemize}
+
+\textbf{Issue: Fields not in the Data Dictionary} The 311 SR \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}
+ identifies 41 data columns (fields) along with related information for each column. 
+ However, when you download the data or explore it via the online portal, you immediately note that there are 47 columns (fields).
+ he additional six fields are not addressed in the Data Dictionary, but do and show up on the Column Manager widget
+ as ``@computed\_region\_xxxx\_xxxx''. To some extent one can infer what the these @computed\_fields are, but it is not possible to know
+for certain. These six fields are:
+
 \begin{itemize}
@@
 
-\begin{figure}[htbp]
-  \centering
-	  \includegraphics[scale = 0.7]{computed_columns_screenshot.png}
+\begin{figure}[H]
+  \centering
+	  \includegraphics[scale = 0.65]{computed_columns_screenshot.png}
 	  \caption{Screenshot of data portal showing the ``@computed'' columns}
@@
 This, of course, raises several questions:
+
 \begin{itemize}
@@
 	\item How can these fields be verified or cross-checked?
-	\item Why are their two fields that appear redundant ``Police Precinct'' and ``Police Precincts''?
-	\item What are ``borough boundaries''? Are these the regular NYC boroughs?
-	\item What are ``community districts''? Is this the same as the ``Community Boards'' that are part of the NYC government?
-	\item Can these fields be included in the official 311 SR Data Dictionary?
-	\item Can these fields be used for analytical purposes? 
+	\item Why are there two fields that appear redundant ``police\_precinct' and ``police\_precincts''? Are these redundant?
+	\item What are ``borough boundaries''? They appear to be borough names. Why are they referred to as ``boundaries''?
+	\item What are ``community districts''? Is this the same as the ``community\_board(s)'' that are part of the NYC government?
+	\item Why are not these fields included in the official 311 SR Data Dictionary?
+	\item Can these fields be reliably used for analytical purposes? 
 \end{itemize}
@@
 The last bullet point is the key here. Are these fields reliable and accurate, and can they be used for subsequent analysis? Unfortunately,
-as of this date, those questions have not been fully addressed by the  NYC Open Data Team. This paper will address further issues with the
-accuracy and subsequent analytical usage for these six @computed fields.
-
-\section{Verifying data types}
+as of this date, those questions have not been fully addressed by the  NYC Open Data Team. This paper will further address that issue by
+exploring the accuracy, validity, and candidacy for subsequent analytical usage.
+
+\section{Validating data types}\label{sec:datatypes}
 Fortunately, both the Data Dictionary and the portal column manager do a good job of indicating the data type.
-As indicated by the red underline in Figure \ref{fig:computed-columns}, there is a small icon next to the column name 
-which indicates the data type of the column. The stylized capital ``T'' indicates text, the ``\#'' sign indicates 
-a numeric, a calendar icon indicates a date field, and the ``pin on a map'' symbol indicates a geospatial field.
+As indicated by the red underline in Figure \ref{fig:computed-columns}, there is a small icon next to the field name 
+which indicates the data type of the field. The stylized capital ``T'' indicates ``text'', the ``\#'' sign indicates 
+``numeric'', a calendar icon indicates a ``date'' field, and the pin on a map symbol indicates a ``geospatial'' field.
 
@@
 \begin{itemize}
-	\item created\_date, closed\_date, due\_date, and resolution\_action\_updated\_date are all in proper date format.
+	\item created\_date, closed\_date, due\_date, and resolution\_action\_updated\_date: All are all in proper date format.
 	\item zip\_codes and incident\_zip: All are 5 numeric digits except for 2 non-numeric entries for incident\_zip (``na'' and ``N/A'').
-	\item x\_coordinate\_state\_plane \&  y\_coordinate\_state\_plane (a NY State geo-location system) are all numeric.
-	\item latitude \& longitude are both numeric.
-	\item community\_districts, borough\_boundaries, city\_council\_district, police\_precinct, and police\_precincts are all numeric.
-	\item Free-form text fields that contain a ``,'' (e.g. incident\_address, resolution\_description) are enclosed in quotation marks.
-\end{itemize}
-
-\section{Blank \& N/A data}
-
-Understanding the absence of data in each fields is an important factor when undertaking analysis. For example, if you wanted to see if the SRs were
+	\item x\_coordinate\_state\_plane \&  y\_coordinate\_state\_plane (a NY State geo-location system): All are all numeric.
+	\item latitude \& longitude: Both are numeric.
+	\item community\_districts, borough\_boundaries, city\_council\_district, police\_precinct, and police\_precincts: All are numeric.
+	\item Free-form text fields that contain a ``,'' (e.g. incident\_address \& resolution\_description) are properly enclosed in quotation marks.
+\end{itemize}
+
+All data fields were subjected to a test of ``correctness'' according to their ``type'' as specified
+in the Data Dictionary. Of the nearly 300 million data fields to check (6.5 million rows, 47 fields per row), there were only two values that failed to test the
+test for appropriate data type; those being two text entries in the incident\_zip field (``na'', and ``N/A''). All date fields were dates. All numbers were numeric.
+All text fields were character. Data typing appears to not be an issue with this dataset.
+
+\section{Blank \& N/A data}\label{sec:blanks}
+
+Understanding the absence of data by field is an important factor when undertaking analysis. For example, if you wanted to see if the SRs were
 closed before or after their due\_date, you would be challenged as 99.57\% of the due\_date field is blank. (Although with such a large dataset
-that still leaves 24,411 values, all of which are N/A, which calls into question are these missing values, or is the specification of a
-due\_date simply not applicable for the vast majority of SRs?)
-
-When counting the various fields for blank or N/A values, the dataset appears to divide into three groups: Mostly Empty, Partially Empty, and Few/None Empty.
-The Mostly Empty category ranges from 93-99.9\% blank and includes such fields as taxi\_company due\_date, pickup\_location, and landmark. The Partially 
+that still leaves 24,411 values, all of which are ``N/A'. Which calls into question...are these missing values, or is the concept of a
+due\_date simply not applicable for the vast majority of SRs?
+
+When counting the various fields for blank or N/A values, the fields appear to divide into three groups: \textbf{Mostly Empty}, \textbf{Partially Empty}, 
+and \textbf{Few/None Empty}. he Mostly Empty category ranges from 93-99.9\% blank. It includes such fields as taxi\_company due\_date, pickup\_location, and landmark. The Partially 
 Empty includes such fields as location\_type, borough, and cross\_street. And the Few/None Empty includes created\_date, complaint\_type, agency, and status.
 In some cases it may make sense to inquiry as to why some fields are frequently blank. Is the data difficult to capture? Does it pertain to only a small set
-of complaint\_type or Agencies? And if the data is truly that sparse, does it make sense to even collect it. For example, you would think that every complaint
-that every SR that has a geographical component, such as an address, would also have a corresponding police\_precinct and community\_district since these
-are all-inclusive. That it would have a lat/long.  But indeed these fields do have small, but nonetheless significant empty entries.
-
-During the analysis, it was determined that the following fields are mandatory, and any rows missing these fields is removed from the dataset (which very, very
-rarely occurred fortunately): created\_date, agency, complaint\_type, and unique\_key.
+of complaint\_type(s) or Agencies? And if the data is truly that sparse, does it even make sense to collect it? Is it legacy data? For example, one would expect that every complaint\_type
+that has a geographical component, such as an address, would also have a corresponding police\_precinct, community\_board, and community\_district since these
+are all-inclusive throughout the five Boroughs of NYC. Addtionally, that it would have a lat/long. And generally that is the case, but these fields do have small number, but nonetheless significant, 
+of empty entries.
+
+During the analysis, it was determined that the following fields are mandatory, and any row missing these fields should be removed from the dataset (which fortunately
+very rarely occurred): created\_date, agency, complaint\_type, and unique\_key. The absence of any one of these four fields prohibits further analysis  as currently conducted. Again,
+it is very rare occurrence, happening only 1-2 times in 6.4 million records. Here is a breakdown of bank and N/A entries by field.
 
@@
 \centering
-\caption{Blank and N/A entries by field}
 \resizebox{0.8\linewidth}{!}{%
-\tiny
+\scriptsize
 \begin{tabular}{lrrrl}
@@
 }
+\caption{Blank and N/A entries by Field}
 \end{table}
@@
 
-Here is a graphic depiction of total empty (blank \& N/As) for each fields. You can see the natural grouping into the Most, Some, and Few categories.
+Here is a graphic depiction of total empty (blank \& N/As) for each fields. You can see the natural grouping into the Most, Some, and Few categories
+in the stair-type visualization.
 
@@
 
-\section{Data type correctness}
-
-All data fields were subjected to a test of ``correctness'' according to their ``type'' as specified
-in the Data Dictionary.  Of the nearly 300 million data fields to check (6.5 million rows, 47 fields per row), there were only two  values that failed to test for the
-proper data type; those being two text entries in the incident\_zip filed (``na'', and N/A). All date fields were dates. All number fields were numeric. All text
-fields were character.
-
- \section{Validating data}
- While the dataset may have fields populated with data that appears correct, it is necessary to inspect those fields to determine if the data is indeed valid. For
- example a date field that contains the value ``February 30, 2024'' is clearly not valid. Nor is a latitude of +95 deg north.
+ \section{Validating Data for Acceptable Values}\label{sec:domain}
+ While the dataset may have fields populated with data that appears correct, it is necessary to inspect those fields to determine if the data is indeed valid. 
+ For example a date field that contains the value ``February 30, 2024'' is clearly not valid. Nor is a latitude of +95 deg north.
 
  In this study, testing for invalid values revealed several serious issues associated with select fields. Accordingly, an analyst would need to be
-extra diligent if using these fields for analysis. In all likelihood, rows with invalid field values would likely need to be removed from the
-analysis dataset.
-
-Some fields can be validated against either a acceptable values, often drawn from historical context. In this study we looked at a 10-year
-dataset to ascertain a domain for values what do not have a specific reference source. We would recommend that these domains be published
-in the Data Dictionary. Here are some results.
-
-\begin{itemize}
-	\item Latitude and Longitude fields were tested to ensure all fell within the geographic boundaries of the City of New York. All were.
+extra diligent if using these fields for analysis. In all likelihood, rows with invalid field values would  need to be removed from the
+dataset before conduction further analysis. Here are some results.
+
+\begin{itemize}
+	\item Latitude and Longitude fields were tested to ensure all fell within the geographic boundaries of the City of New York. All did.
 	\item The unique\_key field was in fact unique.
@@
 	larger historical datasets. Unfortunately, these domains of acceptable values are not specified in the Data Dictionary. These fields all tested as
-	compliant with their domain of acceptable values:
+	compliant within their domain of acceptable values as determined by examining a 10-year dataset:
 		\begin{itemize}
@@
 	\subsection{Issues with Zip Codes}
-	 Unfortunately some fields did prove to be quite problematic when comparing the values to the domain of legal values.
+	 Unfortunately some fields proved to be  problematic when comparing the values to a domain of legal values.
 	 For example, all zip codes (two fields: zip\_codes and incident\_zip) should be valid as defined by the USPS database
-	 which contains over 44,000 valid zip codes.
+	 which contains 37,946 valid zip codes.
 
 	The computed field zip\_codes proved especially problematic with 58\% (3.6 million) of the field entries being invalid.
-	This high number indicates that the computation of this computed fields is highly inaccurate. We would recommend
+	That high number indicates that the computation of this field is highly inaccurate. We recommend
 	dropping this field for future analytic efforts.
 
-	Furthermore, the field incident\_zip while having only .07\% invalid entries, that is still 4163 such errors. As zip code is
-	a primary measure of many NYC city services, and with a freely available database for which to validate entries
-	against, it seems an oversight that any such errors could creep into the system. Some invalid entries are 
-	clearly incorrect just by observation, e.g. 10000, 12345, 11111, etc.
-
-	Because zip codes play a significant role in analyzing City services, these errors can be quite serious. It's useful in 
-	such cases to observe the breakdown of these errors by the responsible Agency.
+	Furthermore, the field incident\_zip while having only .07\% invalid entries, still is a problem with 4163 such errors.
+	As the zip code is a primary measure of many NYC city services, 
+	and with a freely available database for which to validate entries, it seems an oversight 
+	that any such errors could creep into the 311 system. Some invalid entries can  
+	clearly be identified as incorrect just by observation, e.g. 10000, 12345, 11111, etc.
 
 	The breakdown of the invalid entries in the zip\_code, sorted by Agency shows that the distribution by percentage
-	almost precisely mirrors the overall breakdown of SRs by Agency. This indicates a systemic problem, and since the
+	almost precisely mirrors the overall breakdown of SRs by Agency as shown in . This indicates a systemic problem, and since the
 	zip\_code fields is one of the computed fields, it appears these errors are caused by incorrect computations rather
@@
 	\footnotesize
-	\caption{Comparison of Top Ten Zip Codes Lists}
 	\begin{tabular}{lrr@{\hskip 0.5cm}|@{\hskip 0.5cm}lrr}
@@
 	\end{tabular}
+	\caption{Comparison of Top Ten Zip Codes Lists}
 	\end{table}
@@
 	As indicated, six out of ten zip\_codes are invalid, which corresponds closely with what is observed in the overall dataset (58\%). Whereas the incident\_zip
-	field is completely valid, again in-line with the overall incident\_zip validation percentage (99.04\%)
-	Perhaps more curious than the differing zip codes on the two lists, is the fact that the numerical counts are nearly identical;  the overall counts are quite close. This is curious since
-	one field is computed while the other is via data entry; the computed protocol is again the suspect.
+	field is completely valid, again in-line with the overall incident\_zip validation percentage (99.04\%). If the analysis performed using the zip\_codes field were to be
+	presented to the Drirectror, ONL, the conclusion would be in error. Perhaps more curious than the differing zip codes on the two lists, 
+	is the fact that the numerical counts are nearly identical;  the overall counts are quite close. This is curious since
+	one field is computed while the other is via data entry; the computed protocol is again the suspect. The algorithm appears to be counting corrrectly (or nearly so)
+	but mis-identifying the corresponding zip code.
 
@@
 
-	The top ten (by count) of invalid precincts are:
+	The top ten (by count) of invalid precincts for each field are:
 
@@
 	A graphic representation of the invalid police\_precincts plotted by Agency show a near perfect match to the
-	distribution of the entire 6.4 million SRs. Note that the ``big six'' Agencies account for 90\% of these invalid precinct errors, identical
-	to the overall distribution of all SRs. This strongly suggests that this is a systemic problem, almost certainly
-	caused by the computation method applied for these two fields.
+	distribution of the entire 6.4 million SRs as seen in Figure \ref{fig:SR_counts_by_Agency}. Note that the ``big six'' Agencies account for 90\% of these invalid precinct errors, identical
+	to the overall distribution of all SRs.
+(The same is true for a similar graph of by the police\_precint field.) This strongly suggests that this is a systemic problem, almost certainly caused by the computation method applied for these two fields.
 
@@
 	\subsection{Issues with Community Boards}
-	Community Boards are an important aspect of NYC government. Community boards are the most local, grassroots form of City government, 
-	and the connection between communities and elected officials and City agencies. They are organized by the five Boroughs of NYC and represented
-	in this dataset as ``\#\#-Borough'', e.g. ``10 Bronx''. As such, they play an important role in the quality of life for all New Yorkers. There are 59 community boards 
-	throughout the City. There are 12 Community Districts in the Bronx, 18 in Brooklyn, 12 in Manhattan, 14 in Queens and 3 in Staten Island. The Community Boards are
-	used frequently as a way to measure City services throughout the five boroughs, so correctness is important. NYC Borough Presidents in particular are focused
-	on the services offered at the Community Board level.
-
-	In the 2022-2023 dataset there are 27,276 invalid community\_board entries which represents 0.43\% of non-blank data. There are a total of 12 different invalid community boards.
-	These include:
+	Community Boards are an important aspect of NYC government. They are the most local, grassroots form of City government, 
+	and a vital connection between communities and elected officials and City agencies. They are organized by the five Boroughs of NYC and represented
+	in this dataset as ``\#\#-Borough'', e.g. ``10 Bronx'' and report to the various Borough Presidents in NYC. As such, they play an important role in the quality of life for all New Yorkers.
+	There are 59 community boards throughout the City: 12 Community Districts in the Bronx, 18 in Brooklyn, 12 in Manhattan, 14 in Queens and 3 in Staten Island.
+	The Community Boards are used frequently as a way to measure distribution of City services throughout the five boroughs, so correctness is important. 
+	NYC Borough Presidents in particular are focused on the services offered at the Community Board level within their respective Borough.
+
+	In the 2022-2023 dataset there are 27,276 invalid community\_board entries which represents 0.43\% of non-blank data. 
+	There are a total of 12 different invalid community boards, to include:
 
@@
 		\centering
-		\small
-		\caption{Top Ten Invalid 'community\_board' Values}
+		\footnotesize
 		\begin{tabular}{rlr}
@@
 		\end{tabular}
+		\caption{Top Ten Invalid 'community\_board' Values}
 		\end{table}
@@
 		specific issues at some key Agencies, in this case Taxi \& Limo Commission (TLC), Parks \& Recreation, etc. It's not a large error, but the division of services
-		(and complaints) by Community Board is a well-tracked measure.
+		(and complaints) by Community Board is a well-tracked measure so accuracy is again important.
 
@@
 	demographic purposes. DCP is the NYC''s primary land use agency and is instrumental in the City's physical framework.  It is a geographical division rather
-	than a local government division. And like Community Boards, the Community Districts are a frequently used means to breakdown zoning, housing, community facilities,
-	and waterfront/open/public spaces. It is also a well-tracked measure.
+	than a local government division. And like Community Boards, the Community Districts are frequently cited as a means to track zoning, housing, community facilities,
+	and waterfront/open/public spaces. It is another of the NYC well-tracked measures.
 
 	Due to how the community\_district data is formatted, it is not possible to establish validity. However, it is possible to determine that the dataset contains
-	72 unique entries, while there are only 59 valid Community Districts.
-
-\section{Logical inconsistencies }
-
-closed before created (negative duration)
-closed exactly as created (zero duration)
-closed in the future
-
-\subsubsection Case Study: Homeless Person Assistance
-
-due\_date before created
-long resolution\_action\_updated\_date
-
-\section{Accuracy and precision}
+	72 unique entries, while there are only 59 valid Community Districts. So it would appear that there are 13 invalid commuity\_board(s) contained in the dataset.
+
+\section{Inconsistent \& Unusual Patterns with Dates}\label{sec:inconsistencies}
+
+As previously mentioned, the dataset has four data fields:  created\_date, closed\_date, due\_date, and resolution\_action\_update\_date.
+The due\_date field has few entries to analyze; 99.62\% of due\_date entries are ``N/A''.  It is not known if the due\_date(s) are truely
+``not applicable'' or that the due\_date field is rarely used. The  resolution\_action\_update\_date field serves as a date-time marker
+showing when the SR record was last updated.  All the date fields are recorded to the second, e/g/ ``2022-02-11 13:43:05''.
+
+\subsection{created\_date and closed\_date(s) -- Negative Duration}
+Let's look first at the created\_date and closed\_date fields. You might expect that these two fields are automatically populated by the 
+SR application software, either the software used at the 311 organization or that used at the Agencies that are integrated with the 311 system;
+and typically that is indeed the case. In such a case, the act of creating an SR would automatically trigger an entry in created\_date based on a system-wide clock. Thus
+the created\_date is associated with setting an SR status to ``new'' or ``open''. And similarly when an SR's status is changed to
+``closed'' then the closed\_date would be populated.
+
+Unfortunately, our analysis found that that assumption is not always the case. When you analyze the various date fields, several
+anomalies become quickly apparent:
+
+	\begin{itemize}
+		    \item Some SRs have a closed\_date that occurs before the created\_date.
+		    \item Some created\_date(s) and closed\_date(s) appear either in the future or in the far distant past.
+		    \item There are a large number of SRs that are closed and created exactly at midnight or exactly at noon. 
+	\end{itemize}
+
+\textbf{Why are these date fields important?} They're important because citizens, NYC Government Officials, and NYC Agencies use these dates
+to measure the responsiveness for providing the services. How long does it take to replace an out-of-order street light?
+How rapid is the response to a report of domestic violence? Does the NYPD response to a noise complaint in Staten Island take
+longer than in Manhattan? How long does it take to repair a street pothole in Queens?
+
+The measure of ``duration'', of the response time of a 311 call is a closely watched, carefully scrutinized metric, 
+both in terms of overall performance and with a geographical area. And it can have broad political, organizational,
+and budgetary consequences. Duration is not a field in the dataset, but it is easily computed by closed\_date - created\_date.
+Let's look at some anomalies in this ``duration'' field.
+
+Closed-before-Created:  There are 12,251 SRs that are closed before they were created, thereby generating a nonsensical 
+``negative duration''. While this is a small percentage overall (0.2\%) it can have significant impact on response time analysis.
+In most cases, these ``negative durations'' appear to just be mistakes. Here is a sample of some of the largest and the
+smallest errors:
+
+\begin{table}[H]
+    \centering
+    \small
+    \begin{tabular}{>{\ttfamily}l l l r l}
+        \toprule
+        \multicolumn{5}{c}{\textbf{Largest errors (days) excluding extreme negative values}} \\
+        \midrule
+        \textbf{created\_date} & \textbf{closed\_date} & \textbf{duration} & \textbf{agency} \\
+	        \midrule
+	        2023-01-27 14:40:00 & 2022-01-14 14:40:00 & -378.000000 & DOT \\
+	        2023-01-18 10:06:00 & 2022-01-12 10:06:00 & -371.000000 & DOT \\
+	        2023-01-27 14:36:00 & 2022-01-22 14:35:00 & -370.000694 & DOT \\
+	        2023-01-11 11:10:00 & 2022-01-09 11:10:00 & -367.000000 & DOT \\
+	        2023-12-18 03:13:00 & 2023-01-16 13:10:00 & -335.585417 & DOT \\
+	        \midrule
+	        \multicolumn{5}{c}{\textbf{Smallest errors (days)}} \\
+	        \midrule
+	        \textbf{created\_date} & \textbf{closed\_date} & \textbf{duration} & \textbf{agency} \\
+	        \midrule
+	        2023-06-28 07:07:31 & 2023-06-28 07:07:00 & -0.000359 & DOT \\
+	        2023-06-29 09:10:20 & 2023-06-29 09:10:00 & -0.000231 & DOT \\
+	        2023-01-12 06:50:13 & 2023-01-12 06:50:00 & -0.000150 & DOT \\
+	        2023-06-26 08:09:07 & 2023-06-26 08:09:00 & -0.000081 & DOT \\
+	        2023-01-12 06:51:01 & 2023-01-12 06:51:00 & -0.000012 & DOT \\
+	        \bottomrule
+    \end{tabular}
+    \caption{Largest and Smallest errors (days)}
+    \label{tab:combined_errors}
+\end{table}
+
+The largest errors are shown ``excluding extreme negative values''. We found eight SRs with extremely large negative durations (< -730),
+all containing an entry of ``1900-01-01'' as the closed\_date, which generates extremely large negative durations exceeding 44,601 days (122 yrs).
+That large of an anomaly can skew statistical results, even though the number of occurrences is small. This is especially curious
+as the date ``1900-01-01'' is also the earliest date that Microsoft Excel can accommodate as well as how Excel treats a blank
+cell that is designated as a date. All raising suspicions that perhaps these dates were imported into the 311 system from Excel.
+
+According, these  SR rows are removed from the box \& whiskers plot. Note that these extremely large negative duration SRs all
+are from the Department of Homeland Services (DHS) Sample:
+
+\begin{table}[H]
+    \centering
+    \small
+    \begin{tabular}{l l r l}
+        \toprule
+        \textbf{created\_date} & \textbf{closed\_date} & \textbf{duration} & \textbf{agency} \\
+	        \midrule
+	        2022-02-11 13:43:05 & 1900-01-01 & -44601.571586 & DHS \\
+	        2022-02-11 11:03:38 & 1900-01-01 & -44601.460856 & DHS \\
+	        2022-02-11 08:29:49 & 1900-01-01 & -44601.354039 & DHS \\
+	        2022-02-11 14:08:08 & 1900-01-01 & -44601.588981 & DHS \\
+	        2022-02-11 14:05:16 & 1900-01-01 & -44601.586991 & DHS \\
+	        \bottomrule
+    \end{tabular}
+    \caption{Sample of SRs with extremely large negative durations}
+    \label{tab:extreme_negative_durations}
+\end{table}
+
+Here is a graphic visualization of the negative-duration SRs by Agency. It is obvious that this
+negative-duration issue is a problem at the Dept. of Transportation (DOT) where 95\% of these
+types of errors occur.
+
+\begin{figure}[H]
+ 	 \centering
+	  \includegraphics[scale = 0.65]{negative_duration_SR_barchart.png}
+	  \caption{SRs with a Negative Duration by Agency}
+	  \label{fig:negative-duration}
+\end{figure}
+
+This violin chart shows the broad spread of negative-duration SRs, albeit with the extremely large 
+negative-duration SR removed. While there are few outlier points, the magnitude of the negative-durations
+is troubling and can produce bizarre analytical results.
+
+\begin{figure}[H]
+ 	 \centering
+	  \includegraphics[scale = 0.65]{negative_duration_SR_violin.png}
+	  \caption{Negative duration distribution}
+	  \label{fig:negative-duration-violin}
+\end{figure}
+
+\subsubsection{Case Study: Homeless Person Assistance}
+		\textbf{Scenario:} The Dept. of Homeless Services wants to know ``How quickly are 311 calls for ``Homeless Person
+		Assistance over the past two years?'' resolved. This is a typical request made by both the public and City government. It's a performance metric along the lines
+		of ``How quickly is my Agency responding to ``critical'' requests, and does that performance vary by Borough, Zip Code, etc.
+
+		As an analyst, this appears to be a routine effort:
+
+		\begin{enumerate}
+		    \item Select data from 2022-2023 and filter the data by complaint\_type = ``Homeless Person Assistance'' (yields 55,000 SRs)
+		    \item Compute a new field ``duration'' (closed\_date – created\_date)
+		    \item Take an average of the ``duration'' field == \textbf{Answer:  -4.8 days}  
+		\end{enumerate}
+
+		Clearly that answer is nonsensical. How did such a simple task result in an absurd answer? The answer lies in the computation of the
+		``duration'' field. As it turns out, there are eight DHS SRs that have a closed\_date of ``1900-01-01''. Each of those SRs creates a 
+		negative duration of -44,602 days (-122 years). So just those eight SRs with extreme negative duration values is enough to drive
+		the average of the 75,000 Homeless Assistance SRs to a negative value, clearly incorrect.
+
+		As it turns out the distribution of the ``Homeless Person Assistance'' is quite skewed due to a small number of long duration
+		SRs. In this case, the median is a better measure of central tendency than the average. And the median of the duration
+		field is  0.2 days (approx 5 hrs). Here is a look at the distribution of durations for the Homeless Person Assistance SRs showing
+		the large outliers.
+
+		\begin{figure}[H]
+		 	 \centering
+			  \includegraphics[scale = 0.65]{homeless_response_time_clean.png}
+			  \caption{Homeless Assistance SR Durations}
+			  \label{fig:homeless}
+		\end{figure}
+
+	\subsection{created\_date and closed\_date(s) --  Zero Duration}		
+	An even more serious, and more prevalent problem with closed\_date(s) occurs when the closed\_date and created\_date
+	are exactly the same -- to the second -- which accordingly creates a \textbf{zero duration'}. Again this is nonsensical, but the
+	presence of such zero durations can again severely distort analytics. There are 191,141 such SRs representing 3.1\% of all 
+	non-blank data; not an insignificant number. As shown in the chart below, 99\% of the zero duration SRs appear to predominately
+	occur in five Agencies (Dept of Mental Health \&Hygine, Dept of Transportation, Dept of Business, Dept of Sanitation,
+	and Dept of Environmental Protection.) This is not in-line with the overall Agency distribution of SRs and indicates an
+	Agency-specific issue, and the solution likely lies at those Agencies.
+
+	\begin{figure}[H]
+		 \centering
+		 \includegraphics[scale = 0.65]{zero_duration_SR.png}
+		 \caption{SRs with Zero Durations by Agency}
+		 \label{fig:sero-duration}
+	\end{figure}
+
+	\subsection{created\_date and closed\_date(s) -- Midnight \&Noon}
+	One final problem discovered with the created\_date and closed\_date; there appears to be an unusual number of exactly
+	midnight and noon date-time values for these two fields. The distribution of SR creation largely follows the clock; lots of
+	SRs created during day-light hours and much fewer created in the early hours of the morning. However, there appears
+	to be significantly greater numbers of SR both created and closed exactly at midnight, and exactly at noon. The reason this
+	issue is worth investigating, goes back to the ``duration'' question -- if SRs are used as a source to measure responsiveness
+	for the delivery of services, and if those times are not an accurate reflection of the start/finish of the Service, then the
+	duration is in question. Observations during this 2-year period:
+
+	\begin{itemize}
+		    \item  there are 7210 SRs created exactly at midnight
+		    \item There are 99,779 SRs created exactly at noon.
+		    \item There are 105,505 SRs closed exactly at noon. 
+		    \item There are 235,347 SRs closed exactly at midnight
+	\end{itemize}
+
+	Let's look SRs are created and closed by the hour of the day :
+
+	\begin{figure}[H]
+		 \centering
+		 \includegraphics[scale = 0.65]{Created_Hourly_SR_count.png}
+		 \caption{SRs by Hourly created\_date(s)}
+		 \label{fig:hourly-created}
+	\end{figure}
+
+	\begin{figure}[H]
+		 \centering
+		 \includegraphics[scale = 0.65]{Closed_Hourly_SR_count.png}
+		 \caption{SRs by Hourly closed\_date(s)}
+		 \label{fig:hourly-closed}
+	\end{figure}
+
+	These ``closed'' charts clearly show significant spikes in the number of SRs at both midnight
+	and noon. The midnight spike is clearly out of line with the hours before and after midnight. And
+	both the closed-at-noon and closed-at-midnight are approaching 2 standard deviations from the average.
+
+	These unusual patterns of created \& closed SRs on exactly the hours of midnight and noon likely indicate
+	some kind of bulk process that perhaps automatically ``closes'' a large number of SRs with a provided
+	time-stamp of midnight 00:00:00 or noon 12:00:00. Perhaps there is a similar situation for the ``created'' 
+	SRs. So to correct this issue likely requires a close look at what Agencies are providing these suspect SR closed/created times?
+
+	It is not an obvious answer. A visualization of the ``SRs closed at midnight'' shows a distribution that closely mirrors
+	the overall assignment of SRs, as the so-called ``Big Six'' Agencies dominate.
+
+	\begin{figure}[H]
+		 \centering
+		 \includegraphics[scale = 0.65]{closed_at_midnight_chart.png}
+		 \caption{SRs closed at Midnight by Agency}
+		 \label{fig:midnight-closed}
+	\end{figure}
+
+	But the ``SR closed at noon'' visualization shows that just a single Agency, Dept of Sanitation (DSNY),
+	is responsible for > 99\% of the SRs closed exactly at noon. This indicates that it is necessary
+	to work with DSNY to fully understand what is happening.
+
+	\begin{figure}[H]
+		 \centering
+		 \includegraphics[scale = 0.65]{closed_at_noon_chart.png}
+		 \caption{SRs closed at Noon by Agency}
+		 \label{fig:noon-closed}
+	\end{figure}
+
+	\subsection{resolution\_action\_update\_date}
+
+\section{Accuracy and precision}\label{sec:precision}
 There is one area in particular where the question of precision vs. accuracy immediately arises, and that is with the Latitude and Longitude fields.  Both the 
@@
 
-\section{Redundant \& Duplicate data}
+\section{Redundant \& Duplicate data}\label{sec:duplicates}
 Shrinking file size by deleting these 11 redundant fields:
@@
 
-\section{Improvements} \label{sec:improve}
+\section{Improvements} \label{sec:improvements}
 
@@
 
-\section{Discussion} \label{sec:disc}
+\section{Discussion} \label{sec:discussion}
 

### 2024-05-30 manuscript/311 manuscript.tex
Cleaning up

@@
 the specifics of data curation challenges using the NYC 311 Service
-Request Data as a case study. This renowned and frequently viewed 
+Request (SR) Data as a case study. This renowned and frequently viewed 
 dataset serves as a prime example for examining key issues in data curation, 
@@
 \begin{itemize}
- 	\item 2003 - NYC 311 system goes live with a phone-based call center only.
+ 	\item 2003 - NYC 311 system goes live with a phone-based call center only. The ``311'' phone number replaces a myriad of numbers to call for each individual City Agency.
    	\item 2009 - 311 online \& mobile apps launched. Today 60\% of all Service Requests come from mobile \& online channels.
-   	\item 2019 - Major software upgrade of 311 system with enhanced capabilities, dynamic load-balancing, and cloud-based resilience. 	
-      \item 2020 - In August 2020, driven by the COVID pandemic, monthly 311 Service Requests hit an all-time high of 348,463. Seven of the Top 10 busiest days of all time occur in 2020.
+   	\item 2019 - Major software upgrade of 311 system with enhanced capabilities, dynamic load-balancing, and cloud-based resilience. Only major software upgrade since 2004.	
+      \item 2020 - In August 2020, driven by the COVID pandemic, monthly 311 Service Requests hit an all-time high of 348,463. Seven of the Top 10 busiest days of all-time occur in 2020.
       \item 2021 - 311 System is expanded to into the MTA's city subway system, the largest expansion since its inception.
@@
 Since the NYC 311 System went live in 2003, the usage has grown at an annual
-rate of 4.6\%. For this study, we also looked at a 10-year period (2014-2023) to
-observe certain system performance characteristics over a longer time-frame . 
-During that time-frame, the number of 311 SRs grew 50\% as shown in the chart below showing yearly SR counts.
+rate of 4.6\%. As part of this study, we also looked at a 10-year period (2014-2023) to
+observe system performance over a longer time-frame . During that time-frame, 
+the number of 311 SRs grew 50\%.
 
@@
 
-This 10-year timeframe (2014-2023) included the COVID pandemic impact. Note that the pandemic
+This 10-year timeframe (2014-2023) includes the COVID pandemic. Note that the pandemic
 impact reached a peak in the July-September 2020 timeframe, accompanied by additional spikes
-in the 2\textsuperscript{nd}half of 2021 as well. The all-time daily high for SRs occurred on Tuesday, 2020-08-04, with a spike of
-24,415 SRs, a full 9.8 \textsigma's from the mean.
-
-This below chart shows the 10-year period by calendar month. Note that almost all the months before the COVID pandemic outbreak in Feb/Mar
-of 2020 are ``below average'', while most of the months after that period are ``above average'' months. And that trend of higher usage
+in the 2\textsuperscript{nd}half of 2021 as well. The all-time daily high for SRs occurred on Tuesday, 2020-08-04, 
+with a spike of 24,415 SRs, a full 9.8 \textsigma's from the mean.
+
+This chart shows the 2-year period which is the basis for this analysis, by calendar month. 
+Note that almost all the months before the COVID pandemic outbreak in Feb/Mar
+of 2020 are ``below average'', while most of the months after that period are ``above average''. That level of higher usage
 of the 311 system has continued into the early months of 2024.
@@
 
-A quick look at the corresponding  daily tend over a 5-year period (2019-2023) shows both significant spikes
-and significant dips; a noisy graph. While the spikes on 2020-08-04 (24,415 SRs) and 2020-08-05 (19,560 SRs) are certainly COVID related, we were unable to 
-ascertain any reason for the significant spike on Saturday, 2023-09-29 when 7,962 SRs were created.
+A quick look at the corresponding  daily tend.  An average day is 8761 SRs.
 
@@
   \includegraphics[width=\textwidth]{Daily.png}
-  \caption{Daily SR counts over 5-year period}
+  \caption{Daily SR counts over 2022-2023 period}
   \label{fig:daily-counts}
@@
 Seasonal trends are also prevalent. The months of November-thru-April are ``below average'' months, while the months May-thru-October are ``above average''. 
-This seasonal trend is observed throughout the 10-year period.
+This seasonal trend is observed throughout the 10-year period as visualized here.
 
@@
 
-Of note, the lowest SR counts-per-day occur during the Christmas  holiday period encompassing fully 5 of the Bottom 10 SR counts-per-day. The summer months and early
-fall months make up all of the Bottom 10 SR counts-per-day. As noted, the all-time daily high for SRs occurred on Tuesday, 2020-08-04 while the 10-year
+It is not surprising that the lowest SR counts-per-day occur during the Christmas holiday period. The summer months and early
+fall months make up all of the Top 5 SR counts by day. (As a note, the all-time daily high for SRs occurred on Tuesday, 2020-08-04 while the 10-year
 low count occurs on Christmas Day in 2016, a Sunday. Note also that the difference between the peak day (24,415) and the minimum day (2965) varies
-by a factor of 8X, and well exceeds the average daily count of 7489; all of which creates a challenging  system capacity and performance issue.
+by a factor of 8X; all of which creates a challenging  system capacity and performance issue.
 
@@
         \toprule
-        \multicolumn{3}{c|}{Top 10 Days} & \multicolumn{3}{c}{Bottom 10 Days} \\
+        \multicolumn{3}{c|}{Top 5 Days} & \multicolumn{3}{c}{Bottom 5 Days} \\
         \cmidrule(r){1-3} \cmidrule(l){4-6}
@@
         \midrule
-        2020-08-04 & 24415 & 1 & 2014-12-25 & 2965 & 1 \\
-        2020-08-05 & 19560 & 2 & 2015-12-25 & 3247 & 2 \\
-        2023-09-29 & 17962 & 3 & 2014-04-20 & 3287 & 3 \\
-        2020-07-05 & 16916 & 4 & 2015-12-26 & 3405 & 4 \\
-        2020-06-21 & 15883 & 5 & 2016-11-06 & 3419 & 5 \\
-        2020-06-20 & 15825 & 6 & 2014-07-04 & 3455 & 6 \\
-        2020-07-04 & 15794 & 7 & 2014-12-27 & 3534 & 7 \\
-        2021-09-02 & 15205 & 8 & 2014-09-14 & 3563 & 8 \\
-        2020-06-28 & 14057 & 9 & 2014-02-02 & 3565 & 9 \\
-        2021-10-28 & 13575 & 10 & 2016-12-25 & 3587 & 10 \\
+        2020-08-04 & 24,415 & 1 & 2014-12-25 & 2965 & 1 \\
+        2020-08-05 & 19,560 & 2 & 2015-12-25 & 3247 & 2 \\
+        2023-09-29 & 17,962 & 3 & 2014-04-20 & 3287 & 3 \\
+        2020-07-05 & 16,916 & 4 & 2015-12-26 & 3405 & 4 \\
+        2020-06-21 & 15,883 & 5 & 2016-11-06 & 3419 & 5 \\
         \bottomrule
     \end{tabular}
-    \caption{Top 10 and Bottom 10 Days}
+    \caption{Top 5 and Bottom 5 Days}
     \label{tab:combined_counts}
@@
 
-And although the highest number of SRs occur in July and the lowest in February,
-when the \#-of-SRs-per-day metric is computed by adjusting for the differing number of days for each month -- the
-busiest month is shown to be June (followed by July and September). Similarly the quietest months are April, March, and 
-February. In general, the winter months see fewer SRs while the summer months show 
-an increase in activity.
-
-\begin{table}[H]
-    \centering
-    \small
-    \begin{tabular}{lS[table-format=7.0]S[table-format=6.0]} % 'l' for left-aligned, 'S' for siunitx number alignment
-        \toprule
-        \textbf{Month}     & \textbf{Count}   & \textbf{Count per Day} \\ 
-        \midrule
-        June  & 2454156 & 81805 \\ 
-        July  & 2461525 & 79404 \\ 
-        September & 2352822 & 78427 \\ 
-        October & 2407238 & 77653 \\ 
-        August & 2401731 & 77475 \\ 
-        May  & 2334466 & 75305 \\ 
-        November  & 2236037 & 74535 \\ 
-        January   & 2271524 & 73275 \\ 
-        December  & 2199765 & 70960 \\ 
-        February  & 1996408 & 70795 \\ 
-        March & 2174349 & 70140 \\ 
-        April  & 2058169 & 68606 \\ 
-       \bottomrule
-    \end{tabular}
-    \caption{Monthly counts ordered by count-per-day}
-    \label{tab:monthly_counts}
-\end{table}
-
 \FloatBarrier % Ensure all floats before this point are processed before moving on
@@
 Two important fields in this analysis effort are the responsible NYC Agency, and the type of complaint.
-Given both the size of the City of New York government, it's necessary to identify the responsible Agency as well as the type of complaint in order to 
-clarify the responsible party and assist in troubleshooting discrepancies.
-
-When we discovered a data error, we typically observe two trends; either a single (or a few) Agencies are responsible (such as Dept of Transportation), or it is a system-wide issue that follows the general distribution
-of  Service Requests (SRs) across the system. Thus we can often identify if it is an Agency specific problem, or a systemic effect affecting all Agencies.
+Given the size of the City of New York government, it is necessary to identify the responsible Agency as well as the type of complaint in order to 
+identify the responsible party and assist in troubleshooting discrepancies.
+
+When we discovered a data error, we typically observe one of two trends; either a single (or a few) Agencies are responsible (such as Dept of Transportation), 
+or it is a system-wide issue that follows the general distribution of  SRs across the system. 
+Thus we can often identify if it is an Agency specific problem, or a systemic issue.
 
@@
   \centering
-	  \includegraphics[width=\textwidth]{SRs\_by\_Agency.png}
+	  \includegraphics[width=\textwidth]{SRs_by_Agency.png}
 	  \caption{SR counts by Agency and Cumulative Percentage}
@@
 
-Note the distribution of the complaint\_type; it is heavily skewed to a few select complaints with an exceptionally long tail.
-This graph shows the top 20 complaint\_type(s) out of a total of 220.
+The distribution of complaint\_type skews to a few select complaints accompanied by a long tail.
+This graph shows the top 20 complaint\_type(s) comprising 70\% out of a total of 220.
 
@@
 The top 20 complaints contain several of the eight different types of ``Noise'' complaints (Residential, Commercial, Street, Helicopter, etc.). These Noise-related
-complaints total 1.435 million in the 2022-2023 dataset, a full 22\% of all complaint types, which is the most frequently occurring type.
-(Note: Ilegal Parking is second, followed by Heat/Hot Water.) Noise complaints are also handled by NYPD, as is Illegal Parking and Blocked Driveway complaints, which contributes
-significantly to the top ranking of the NYPD as the responsible Agency.
+complaints total 1.435 million in the 2022-2023 dataset, a full 22\% of all complaint types, which, when consolidated, is the most frequently occurring complaint\_type.
+Ilegal Parking is second, followed by Heat/Hot Water. Noise complaints are handled by NYPD, as is Illegal Parking and Blocked Driveway complaints, which contributes
+significantly to the top ranking of the NYPD as the lead responsible Agency, receiving a full 43\% of all SRs. 
+Housing, Preservation, \& Development are 2\textsuperscript{nd}, with 21\%
+
+For  non-NYC residents, HPD is the City agency that manages the 177,569 NYC public housing units as well as monitoring
+all City rental and leased apartment units. HPD collectively serves 528,105 people, a population larger than Atlanta or Miami. 
+Hence the large number of Heat/Hot Water and Unsanitary complaints handled by HPD.
 
@@
         \midrule
-        NOISE - RESIDENTIAL        & 675502 & 10.56 & NYPD  \\ 
-        NOISE - STREET/SIDEWALK    & 300507 &  4.70 & NYPD  \\ 
-        NOISE - COMMERCIAL         & 128892 &  2.02 & NYPD  \\ 
-        NOISE - VEHICLE            & 115956 &  1.81 & NYPD  \\ 
-        NOISE                      & 100413 &  1.57 & DEP   \\ 
-        NOISE - HELICOPTER         &  85345 &  1.33 & EDC   \\ 
-        NOISE - PARK               &  16620 &  0.26 & NYPD  \\ 
+        NOISE - RESIDENTIAL        & 675,502 & 10.56 & NYPD  \\ 
+        NOISE - STREET/SIDEWALK    & 300,507 &  4.70 & NYPD  \\ 
+        NOISE - COMMERCIAL         & 128,892 &  2.02 & NYPD  \\ 
+        NOISE - VEHICLE            & 115,956 &  1.81 & NYPD  \\ 
+        NOISE                      & 100,413 &  1.57 & DEP   \\ 
+        NOISE - HELICOPTER         &  85,345 &  1.33 & EDC   \\ 
+        NOISE - PARK               &  16,620 &  0.26 & NYPD  \\ 
         NOISE - HOUSE OF WORSHIP   &   2757 &  0.04 & NYPD  \\ 
@@
 There are also some curious and unusual Service Requests in the 311 data: 
-tanning, tattooing, trans fat, unsanitary pigeon condition, illegal animal kept as pet, harboring bees/wasps, and the author's favorite
+tanning, tattooing, trans fat, unsanitary pigeon condition, illegal animal kept as pet, harboring bees/wasps, and the authors' favorite
 radioactive material. New Yorkers certainly live interesting lives.
-
-The NYC HPD handles about 49\% of what NYPD handles. For  non-NYC readers, 
-HPD is the City agency that manages the 177,569 NYC public housing units as well as monitoring
-all rental and leased apartment units. HPD collectively serves 528,105 people, a population larger than Atlanta or Miami. 
-Hence the large number of Heat/Hot Water and Unsanitary complaints handled by HPD.
 
@@
 	\item Structural errors to include naming conventions, extra fields not in the Data Dictionary, or inconsistent data entry.
-	For example intermixing blanks, spaces, NA, N/A, and <NA> all to indicate the absence of data.
+	For example intermixing blanks, spaces, NA, N/A, and \textless{}NA\textgreater{} all to indicate the absence of data.
 	\item Redundant, unnecessary, or irrelevant fields
@@
 	\subsection{Issues with Zip Codes}
+	\label{sec:zip-codes}
 	 Unfortunately some fields proved to be  problematic when comparing the values to a domain of legal values.
@@
 	\subsection{Issues with Police Precincts}
+	\label{sec:police-precincts}
 	A curious case also exists when examining the two nearly identical fields - police\_precincts and police\_precinct. Both of those fields are among
@@
 
-The largest errors are shown ``excluding extreme negative values''. We found eight SRs with extremely large negative durations (< -730),
+The largest errors are shown ``excluding extreme negative values''. We found eight SRs with extremely large negative durations (\textless{} -730),
 all containing an entry of ``1900-01-01'' as the closed\_date, which generates extremely large negative durations exceeding 44,601 days (122 yrs).
@@
 	        \midrule
-	        2022-02-11 13:43:05 & 1900-01-01 & -44601.571586 & DHS \\
-	        2022-02-11 11:03:38 & 1900-01-01 & -44601.460856 & DHS \\
-	        2022-02-11 08:29:49 & 1900-01-01 & -44601.354039 & DHS \\
-	        2022-02-11 14:08:08 & 1900-01-01 & -44601.588981 & DHS \\
-	        2022-02-11 14:05:16 & 1900-01-01 & -44601.586991 & DHS \\
+	        2022-02-11 13:43:05 & 1900-01-01 & -44602 & DHS \\
+	        2022-02-11 11:03:38 & 1900-01-01 & -44602 & DHS \\
+	        2022-02-11 08:29:49 & 1900-01-01 & -44601 & DHS \\
+	        2022-02-11 14:08:08 & 1900-01-01 & -44602 & DHS \\
+	        2022-02-11 14:05:16 & 1900-01-01 & -44602 & DHS \\
 	        \bottomrule
@@
 	non-blank data; not an insignificant number. As shown in the chart below, 99\% of the zero duration SRs appear to predominately
-	occur in five Agencies (Dept of Mental Health \&Hygine, Dept of Transportation, Dept of Business, Dept of Sanitation,
+	occur in five Agencies (Dept of Mental Health \&Hygiene, Dept of Transportation, Dept of Business, Dept of Sanitation,
 	and Dept of Environmental Protection.) This is not in-line with the overall Agency distribution of SRs and indicates an
@@
 	\subsection{created\_date and closed\_date(s) -- Midnight \&Noon}
-	One final problem discovered with the created\_date and closed\_date; there appears to be an unusual number of exactly
-	midnight and noon date-time values for these two fields. The distribution of SR creation largely follows the clock; lots of
-	SRs created during day-light hours and much fewer created in the early hours of the morning. However, there appears
-	to be significantly greater numbers of SR both created and closed exactly at midnight, and exactly at noon. The reason this
-	issue is worth investigating, goes back to the ``duration'' question -- if SRs are used as a source to measure responsiveness
-	for the delivery of services, and if those times are not an accurate reflection of the start/finish of the Service, then the
-	duration is in question. Observations during this 2-year period:
-
-	\begin{itemize}
-		    \item  there are 7210 SRs created exactly at midnight
-		    \item There are 99,779 SRs created exactly at noon.
-		    \item There are 105,505 SRs closed exactly at noon. 
-		    \item There are 235,347 SRs closed exactly at midnight
-	\end{itemize}
-
-	Let's look SRs are created and closed by the hour of the day :
+	An additional  problem discovered with the created\_date and closed\_date fields; there appears to be an unusually large number of 
+	SRs created or closed at exactly midnight (00:00:00) and noon (12:00:00), to the second. The distribution of SR creation and closure 
+	largely follows the work-day clock with lots of SRs created during day-light hours, and fewer SRs 
+	created in the early hours of the morning. However, here there appears to be significantly greater numbers of SR closed exactly at midnight and exactly at noon,
+	as well as a significant number of SRs created exactly at noon. The reason for this anomaly is worth investigating 
+	as it affects the SR ``duration'' question; if SR durations are used to measure responsiveness
+	for the delivery of services, and if those times are not an accurate reflection of the start/finish of the Service, then the use of
+	duration as a service quality metric is called into question. Observations during this 2-year period include:
+
+	\begin{itemize}
+		    \item There were 235,347 SRs closed exactly at midnight
+		    \item There were 105,505 SRs closed exactly at noon. 
+		    \item There were 99,779 SRs created exactly at noon 
+	\end{itemize}
+
+	Here is a visualization of SRs closed and created by the hour of the day :
+
+	\begin{figure}[H]
+		 \centering
+		 \includegraphics[scale = 0.65]{Closed_Hourly_SR_count.png}
+		 \caption{SRs by Hourly closed\_date(s)}
+		 \label{fig:hourly-closed}
+	\end{figure}
 
@@
 
-	\begin{figure}[H]
-		 \centering
-		 \includegraphics[scale = 0.65]{Closed_Hourly_SR_count.png}
-		 \caption{SRs by Hourly closed\_date(s)}
-		 \label{fig:hourly-closed}
-	\end{figure}
-
-	These ``closed'' charts clearly show significant spikes in the number of SRs at both midnight
-	and noon. The midnight spike is clearly out of line with the hours before and after midnight. And
-	both the closed-at-noon and closed-at-midnight are approaching 2 standard deviations from the average.
-
-	These unusual patterns of created \& closed SRs on exactly the hours of midnight and noon likely indicate
-	some kind of bulk process that perhaps automatically ``closes'' a large number of SRs with a provided
-	time-stamp of midnight 00:00:00 or noon 12:00:00. Perhaps there is a similar situation for the ``created'' 
-	SRs. So to correct this issue likely requires a close look at what Agencies are providing these suspect SR closed/created times?
-
-	It is not an obvious answer. A visualization of the ``SRs closed at midnight'' shows a distribution that closely mirrors
-	the overall assignment of SRs, as the so-called ``Big Six'' Agencies dominate.
+	The closed\_date chart show significant spikes in the number of SRs at both midnight
+	and noon with significant counts at exactly those times, to the second. The midnight spike is certainly suspect compared to the hours just before and after midnight where
+	there is normally a pronounced slump in SR closures. And both the closed-at-noon and closed-at-midnight are 
+	approaching 2\textsigma's from the hourly average. The created\_date shows a maximum value at noon also with a significant count occurring exactly at noon, again to 
+	the second.
+
+	These unusual patterns of created \& closed SRs at exactly the hours of midnight and noon perhaps indicates
+	the presence of a bulk process; one that perhaps automatically ``closes'' (or ``creates'') a large number of SRs with a provided
+	time-stamp of midnight (00:00:00) or noon (12:00:00). If so, then addressing this issue likely requires a closer look at those 
+	Agencies providing these suspect SR closed/created times. Clearly the desirable state is that the created\_date and closed\_date fields
+	reflect the true opening and closing of the SRs, such that an accurate ``duration'' can be computed which accurately
+	represents the SR life-cycle.
+
+	Looking at the distribution by Agency for the ``closed-exactly-at-midnight'' shows that \textgreater{}90\% of these suspect SRs come from just two Agencies - Dept. of 
+	Buildings (DOB) and Dept. of Sanitation (DSNY). If you include HPD and DOT, that accounts for 99\% of the suspect SR closures.
 
@@
 		 \includegraphics[scale = 0.65]{closed_at_midnight_chart.png}
-		 \caption{SRs closed at Midnight by Agency}
+		 \caption{SRs Closed Exactly at Midnight by Agency}
 		 \label{fig:midnight-closed}
@@
 
-	But the ``SR closed at noon'' visualization shows that just a single Agency, Dept of Sanitation (DSNY),
-	is responsible for > 99\% of the SRs closed exactly at noon. This indicates that it is necessary
-	to work with DSNY to fully understand what is happening.
+	Hoever, the ``SR closed at noon'' visualization shows that  a single Agency, DSNY,
+	is responsible for \textgreater{}99\% of the SRs ``closed-exactly-at-noon''. Accordingly, any effort to resolve this issue
+	necessitates working with DSNY to fully understand what is occurring.
 
@@
 		 \includegraphics[scale = 0.65]{closed_at_noon_chart.png}
-		 \caption{SRs closed at Noon by Agency}
+		 \caption{SRs Closed Exactly at Noon by Agency}
 		 \label{fig:noon-closed}
@@
 
+	The 99,779 SRs created-exactly-at-noon also appears to originate with the Dept of Sanitation (DSNY) which
+	is responsible for \textgreater{}99\% of the suspect SRs.
+
+	\begin{figure}[H]
+		 \centering
+		 \includegraphics[scale = 0.65]{created_at_noon_chart.png}
+		 \caption{SRs Created Exactly at Noon by Agency}
+		 \label{fig:noon-created}
+	\end{figure}
+
+	It is difficult to determine if these observations are indeed anomalies. They could be the result of normal, expected
+	311 SR activity. But at least by observation that appears unlikely.
+
 	\subsection{resolution\_action\_update\_date}
+	When an SR is updated, by any means on any field, the software automatically populates the resolution\_action\_update\_date. Analyzing that field
+	it is apparent that some of the SR updates are happening long, long after the SR is closed. In this dataset there are eight SRs that have 
+	resolution\_action\_update\_date values of 44,601 days after the closed\_date. The problem lies not with the resolution\_action\_update\_date, but
+	rather with the closed\_date values for these 8 SRs, all of which are in the format of ``1900-01-01''. It has previously been discussed that these
+	inaccurate dates may possibly be coming from an Excel spreadsheet as that is the minimum date which Excel can handle. Here is a sample
+	of those SRs:
+
+\begin{table}[H]
+	    \centering
+	    \footnotesize
+	    \begin{tabular}{cccc}
+	        \hline
+	        Agency & Closed Date & Resolution Action Updated Date & Post Closed Update Duration (days) \\
+	        \hline
+	        DHS & 1900-01-01 & 2022-02-12 00:58:43 & 44602 \\
+	        DHS & 1900-01-01 & 2022-02-11 14:28:59 & 44601 \\
+	        DHS & 1900-01-01 & 2022-02-11 14:17:41 & 44601 \\
+	        DHS & 1900-01-01 & 2022-02-11 13:45:59 & 44601 \\
+	        DHS & 1900-01-01 & 2022-02-11 13:22:12 & 44601 \\
+	        \hline
+	    \end{tabular}
+	    \caption{Extremely long Resolution Updates}
+	    \label{tab:resolution-updates}
+	\end{table}
+
+	These obvious inaccurate date values are removed from the analysis. But even so, there are some very late
+	updates to SRs, with some updates occurring almost two years after the SR is closed.   There are a total of 7460 SRs that are 
+	updated \textgreater{}30 days and \textless{}730 days after the closed\_date. The median of these late (\textless{}30 days)
+	post-closed resolution\_action\_update\_date(s) is 84 days. However, the mean resolution\_action\_update\_date 
+	for all of the SRs is only 0.39 days (9.3 hours) after the SR is closed. It is not known if this is normal behavior
+	or an area that may require further investigation.
+
+	\begin{table}[H]
+    \centering
+    \footnotesize
+	    \begin{tabular}{cccc}
+	        \hline
+	        Agency & Closed Date & Resolution Action Updated Date & Post Closed Update Duration (days) \\
+	        \hline
+	        TLC & 2023-05-25 14:21:12 & 2023-10-19 10:57:02 & 147 \\
+	        TLC & 2023-07-12 11:03:46 & 2023-12-07 09:00:05 & 148 \\
+	        TLC & 2023-08-03 10:34:23 & 2024-01-29 10:00:37 & 179 \\
+	        TLC & 2022-04-04 10:54:39 & 2022-06-02 12:14:11 & 59 \\
+	        TLC & 2022-12-05 15:23:45 & 2023-02-02 09:06:45 & 59 \\
+	        \hline
+	    \end{tabular}
+    \caption{Long post-closed Resolution Updates}
+    \label{tab:resolution}
+	\end{table}
+
+	Here is a look at the spread of the post-closed resolution\_action\_update\_date values those SRs 
+	that are \textgreater{}30 days and \textless{}730 days. There are some significant outliers
+
+	\begin{figure}[H]
+		 \centering
+		 \includegraphics[scale = 0.65]{post_closed_violin.png}
+		 \caption{Post-Closed resolution\_action\_update\_date >30 days }
+		 \label{fig:resolution-violin}
+	\end{figure}
+
+	If we look at a breakout of these (late?) update dates by Agency, we can see that the vast majority li with two
+	Agencies: Taxi \& Limo Commission (TLC) and Dept of Sanitation (DSNY) which account for 96\% of the SRs in
+	question.  Is there a valid reason that TLC and DSNY would be updating an SR over month (or months) after
+	the SR has been closed? Perhaps, but it seems to merit closer investigation.
+
+	\begin{figure}[H]
+		 \centering
+		 \includegraphics[scale = 0.65]{postClosedBarChart.png}
+		 \caption{Post-Closed resolution\_action\_update\_date >30 days by Agency }
+		 \label{fig:resolution-by-agency}
+	\end{figure}
 
@@
 latitude and longitude). Given that 1 degree of latitude at the equator is equal to 111.044736 kilometers, the ``1'' at the end of that number represents
-approximate 1.1104 nanometer or 1/1,000,000,000 of a meter. A DNA molecule is approximately 2nm in width.
+approximate 1.1104 nanometer or 1/1,000,000,000 of a meter. For reference a DNA molecule is approximately 2nm in width.
 
@@
 
-\subsection{Case Study I - Response to Requests for Homeless Person Assistance}
-
-\section{Redundant \& Duplicate data}\label{sec:duplicates}
+\section{Redundant \& Duplicate fields}\label{sec:duplicates}
+During this analysis, several redundant fields were observed.
+
+\subsection{latitude \& longitude and the location fields}  The location field is a pure concatenation of the latitude and longitude fields, 
+with a comma an parenthesis added. Example:
+
+	\begin{itemize}
+		\item  latitude: 40.768456429488
+		\item  longitude: -73.9575661888774
+		\item  location: (40.768456429488, -73.95756618887745)
+	\end{itemize}
+
+The inclusion of the location field seems questionable as the data is arguably more difficult to extract than the two specific fields, especially for
+software programs.
+
+\subsection{borough and park\_borough fields}  These two fields are 100\% matches; fully redundant.
+
+\subsection{borough and borough\_boundaries fields}  The borough field is a text field for the five New York City boroughs (Manhattan, Queens, Brooklyn, etc.).
+The borough\_boundary field is numeric with values from 1-5 corresponding to the five boroughs (Staten Island, Brooklyn, Queens, Manhattan, Bronx).
+When the borough\_boundary field is translated and compared to the borough field, there is a 98.3\% match. Given the size of the dataset, that 
+still leaves 110,715 non-matching values, not an insignificant number.
+
+And that, in a nutshell, is the problem with two ``near-duplicate'' fields. It is like the  story of a man wearing two watches; he's never
+quite sure what time it is. And with ``near-duplicate'' fields, which field is correct when they disagree? 
+If one field is blank, can you use the other one? Which field can be used as the reference value.
+ And as we will see, there are several such examples in this dataset.
+
+We can tell where the non-matches occur. It is mostly in DOT, NYPD, and DSNY which represent 90\% of the non-matching values.
+
+	\begin{figure}[H]
+		 \centering
+		 \includegraphics[scale = 0.65]{non_matching_borough_boundaries_chart.png}
+		 \caption{Non-matching borough and borough\_boundary by Agency }
+		 \label{fig:borough-boundaries}
+	\end{figure}
+
+\subsection{borough and taxi\_company\_borough fields}
+The two fields borough and taxi\_company\_borough might appear to be related. In fact, despite the names, 
+the two data fields are nearly completely different with only 0.05\% matching. The mismatch is so
+high that it suggests the two fields are used very differently.  This field is used exclusively by the 
+Taxi \& Limo Commission which governs taxis and other cars for hire.
+There are a total of 3567 non-blank entries in the taxi\_company\_borough field, with 99.94\% blank.
+
+ \subsection{incident\_zip and zip\_codes fields}
+ The zip\_code field is one of the six computed fields and its validity has been explored in Section \ref{sec:zip-codes} and
+ found to be lacking with 55.55\% of the entries invalid according to the USPS, while the companion zipcode
+ field incident\_zip field had an accuracy rate of 99.93\%. It is not surprising that 
+ only 0.58\% of the two fields match (37,292 matches out of 6.4 million SRs.). While an Agency breakdown can be
+ determined, given the questionable computation of the zip\_code field, such an analysis would
+ seem to provide any useful information.
+
+ \subsection{police\_precinct and police\_precincts fields}  
+Both the police\_precinct and police\_precincts are among the computed fields not cited in the Data Dictionary.
+Their usage however, is easy to discern.
+Are these two fields duplicates?
+Nearly so:  99.94\% of the entries match.
+Since the source of the fields and the underlying computational process are not know, it remains a challenge to determine which field is
+the ``right'' field to use. (The validity of these fields is covered in Section \ref{sec:police-precincts}.)
+
+ \subsection{agency and agency\_name fields}
+ While not duplicates, these fields have a 1:1 correspondence.  The Agency field contains abbreviations for the City agencies such as NYPD, DOT, HPD, DEP, etc.
+ The agency\_name field contains the full name of the various organizations: New York Police Department, Department of Transportation, Department
+ of Housing Preservation and Development, Department of Environmental Protection, and so on.  Given how well known the NYC Agency abbreviations are
+it seems redundant to include both in the dataset.
+
+\subsection{cross\_street\_1 \& intersection\_street\_1 and cross\_street\_2 \& intersection\_street\_2}
+There are two sets of street pairs in the dataset:
+
+	\begin{itemize}
+		\item cross\_street\_1
+		\item cross\_street\_2
+		\item intersection\_street\_1
+		\item intersection\_street\_2
+	\end{itemize}
+
+These two pairs of streets are used to help identify the location of the reported incident, as in common in
+New York to provide the street address and the cross street, as in ``24 W 90th street between Columbus and
+Amsterdam''. The Data Dictionary states that these streets are
+ ``based on the geo validated incident location''. It then adds ``If the service request 
+pertains to a location in the middle of a block, cross street will be provided by geocoding...If the service request 
+pertains to a street corner, intersection will be provided by geocoding.''  One might think that the
+cross\_street(s) would apply for ``middle of a block'' incidents, while the intersection\_street(s) would
+apply to those incidents that happen on a street corner; that those two sets of fields would be
+exclusive to one another; and either/or situation.
+
+Unfortunately, that is not the case. For the majority of data rows, both pair of fields are populated.
+In fact, 88\% of the time, cross\_street\_1 exactly matches intersection\_street\_1 and 
+cross\_street\_2 exactly matches intersection\_street\_2, and only 11.9\% are non-matches.
+
+(Note address standardization was applied to the street addresses using the R \emph{campfin} package. 
+This was done so as to prevent the software from generating a non-match between 
+streets such as ``240 E 69\textsuperscript{th} ST'' and ``240 E 69\textsuperscript{th} Street'' 
+which are clearly the same address, just some slight formatting and spelling differences.)
+
+The problem occurs when the two pairs of streets do not match, or when one is blank and the
+other is not. Which set is correct? If one is blank, say cross\_street\_1, but  intersection\_street\_1
+is not blank, should you use the non-blank one?  Let's look at some examples:
+
+\begin{table}[H]
+    \centering
+    \footnotesize
+    \begin{tabular}{>{\small\ttfamily}l >{\small\ttfamily}l >{\small\ttfamily}l}
+        \toprule
+        \multicolumn{3}{c}{\textbf{Matching cross\_street\_1 and intersection\_street\_1}} \\
+        \midrule
+        \textbf{cross\_street\_1} & \textbf{intersection\_street\_1} & \textbf{agency} \\
+        \midrule
+        FORT HAMILTON PKWY & FORT HAMILTON PKWY & NYPD \\
+        87 ST              & 87 ST              & NYPD \\
+        EAST 169 ST        & EAST 169 ST        & NYPD \\
+        EAST 74 ST         & EAST 74 ST         & DOT  \\
+        CHERRY AVE         & CHERRY AVE         & NYPD \\
+        \midrule
+        \multicolumn{3}{c}{\textbf{Non-matching cross\_street\_1 and intersection\_street\_1}} \\
+        \midrule
+        \textbf{cross\_street\_1} & \textbf{intersection\_street\_1} & \textbf{agency} \\
+        \midrule
+        KAPPOCK ST     & NETHERLAND AVE   & DOT \\
+        GREENE AVE     & CLINTON AVE      & DOT \\
+        LEWIS AVE      & FULTON ST        & DOT \\
+        GERARD AVE     & EAST 161 ST      & DOT \\
+        8 AVE          & WEST 136 ST      & DOT \\
+        \bottomrule
+    \end{tabular}
+    \caption{Matching/Non-Matching cross\_street\_1 \& intersection\_street\_1}
+    \label{tab:streets1}
+\end{table}
+
+	\begin{figure}[H]
+		 \centering
+		 \includegraphics[scale = 0.65]{non-matchingcross_street_1andintersection_street_1.png}
+		 \caption{Non-matching X\_Street\_1 \& Intersection\_Street\_1}
+		 \label{fig:xstreet1}
+	\end{figure}
+
+	\begin{figure}[H]
+		 \centering
+		 \includegraphics[scale = 0.65]{non-matchingcross_street_2andintersection_street_2.png}
+		 \caption{Non-matching X\_Street\_2 \& Intersection\_Street\_2}
+		 \label{fig:xstreet2}
+	\end{figure}
+
+Near matches.
+
+\begin{table}[H]
+    \centering
+    \footnotesize
+    \begin{tabular}{l l l r}
+        \toprule
+        \textbf{cross\_street\_1} & \textbf{intersection\_street\_1} & \textbf{agency} & \textbf{hamming\_distance} \\
+        \midrule
+        WEST 168 ST    & WEST 167 ST           & DOT    & 1 \\
+        115 AVE        & 120 AVE               & DOT    & 2 \\
+        105 AVE        & 107 AVE               & DOT    & 1 \\
+        71 ST          & 80 ST                 & DOT    & 2 \\
+        145 ST         & 150 ST                & DOT    & 2 \\
+        138 ST         & 139 ST                & DOT    & 1 \\
+        19 AVE         & 20 AVE                & DOT    & 2 \\
+        25 AVE         & 30 AVE                & DOT    & 2 \\
+        76 ST          & 79 ST                 & DOT    & 1 \\
+        67 ST          & 68 ST                 & DOT    & 1 \\
+        \bottomrule
+    \end{tabular}
+    \caption{Near-matches for cross\_street\_1, intersection\_street\_1}
+    \label{tab:x1nearmatches}
+\end{table}
+
+\begin{table}[H]
+    \centering
+    \footnotesize
+    \begin{tabular}{l r r}
+        \toprule
+        \textbf{category} & \textbf{count} & \textbf{percentage} \\
+        \midrule
+        Matching                    & 5,637,182 & 88.15     \\
+        Both blank -- Matching      & 1,624,499 & 25.4      \\
+        Non-matching                &   757,726 & 11.85     \\
+        cross\_street\_1\_blank     &   222,232 & N/A       \\
+        intersection\_street\_1\_blank &   519,348 & N/A       \\
+        Near-match                  &       128 & 0.002002  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Summary of cross\_street\_1 and intersection\_street\_1}
+    \label{tab:summary1}
+\end{table}
+
+\begin{table}[H]
+    \centering
+    \footnotesize
+    \begin{tabular}{l r r}
+        \toprule
+        \textbf{category} & \textbf{count} & \textbf{percentage} \\
+        \midrule
+        Matching                    & 5,644,365 & 88.26     \\
+        Both blank -- Matching      & 1,623,276 & 25.38     \\
+        Non-matching                &   750,543 & 11.74     \\
+        cross\_street\_2\_blank     &   222,788 & N/A       \\
+        intersection\_street\_2\_blank &   517,431 & N/A       \\
+        Near-match                  &     1,500 & 0.023456  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Summary of cross\_street\_2 and intersection\_street\_2}
+    \label{tab:summary_cross_intersection_2}
+\end{table}
+
+\begin{table}[H]
+    \centering
+    \footnotesize
+    \begin{tabular}{l r r}
+        \toprule
+        \textbf{category} & \textbf{count} & \textbf{percentage} \\
+        \midrule
+        Matching                    & 5,644,365 & 88.26     \\
+        Both blank -- Matching      & 1,623,276 & 25.38     \\
+        Non-matching                &   750,543 & 11.74     \\
+        cross\_street\_2\_blank     &   222,788 & N/A       \\
+        intersection\_street\_2\_blank &   517,431 & N/A       \\
+        Near-match                  &     1,500 & 0.023456  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Summary of cross\_street\_2 and intersection\_street\_2}
+    \label{tab:summary_cross_intersection_2}
+\end{table}
+
+\begin{table}[H]
+    \centering
+    \footnotesize
+    \begin{tabular}{l l l r}
+        \toprule
+        \textbf{cross\_street\_2} & \textbf{intersection\_street\_2} & \textbf{agency} & \textbf{hamming\_distance} \\
+        \midrule
+        19 AVE         & 18 AVE                & DOT    & 1 \\
+        227 ST         & 225 ST                & DOT    & 1 \\
+        65 ST          & 65 PL                 & DOT    & 2 \\
+        18 AVE         & 17 AVE                & DOT    & 1 \\
+        BEACH 110 ST   & BEACH 109 ST          & DOT    & 2 \\
+        3 AVE          & 2 AVE                 & DOT    & 1 \\
+        88 ST          & 87 ST                 & DOT    & 1 \\
+        WEST 114 ST    & WEST 113 ST           & DOT    & 1 \\
+        192 ST         & 189 ST                & DOT    & 2 \\
+        5 AVE          & 4 AVE                 & DOT    & 1 \\
+        \bottomrule
+    \end{tabular}
+    \caption{Sample of near-matching cross\_street\_2 and intersection\_street\_2 (both non-blank)}
+    \label{tab:near_matching_cross_intersection_2}
+\end{table}
+
 Shrinking file size by deleting these 11 redundant fields:
@@
 Size after removing redundant columns: 2.8 Gb 
-Potential size reduction: 395.3 Mb or 12.3
+Potential size reduction: 395.3 Mb or 12.3\%
 

### 2024-05-30 code/311 data cleansing V2.R

@@ -30,7 +30,7 @@ data1File <- file.path("..", "data", "311_2022-2023_AS_OF_01-31-2024.csv")
 # Hard code the max_closed_date to be midnight of the date of the data export from NYC Open Data
 max_closed_date <- as.POSIXct("2024-01-31 23:59:59", format = "%Y-%m-%d %H:%M:%S")
 
-chart_directory_path <- file.path("..", "charts")
+chart_directory_path <- file.path("..", "charts", "2022-2023 study", "Core charts")
 
 writeFilePath <- file.path("..", "data", "smaller.csv")
 
@@ -277,7 +277,8 @@ cross_street_analysis <- function(
     chart_title <- paste0("Non-Matching '", cross_street, "' and '", intersection_street, "' by Agency & cumulative percentage", sep = "")
     chart_file_name <- paste0("non-matching", cross_street, "and", intersection_street, ".png", sep = "")
     create_combo_chart(
-      non_dup_streets_non_blank,
+#      non_dup_streets_non_blank,
+      non_dup_streets,
       chart_title,
       chart_file_name
     )
@@ -362,7 +363,7 @@ cross_street_analysis <- function(
     data.frame(
       category = c(
         "Matching",
-        paste("Both blank", "\u2208", "Matching", sep = ""),
+        "Both blank -- Matching",
         "Non-matching",
         paste(cross_street, "blank", sep = "_"),
         paste(intersection_street, "blank", sep = "_"),
@@ -1442,8 +1443,8 @@ SR_created_hourly_ <- create_bar_chart(
   x_axis_title = NULL,
   y_axis_title = "count",
   print_out_title = "Hourly created",
-  add_mean = FALSE,
-  add_median = TRUE,
+  add_mean = TRUE,
+  add_median = FALSE,
   add_sd = TRUE,
   chart_file_name = "Created_Hourly_SR_count.png"
 )
@@ -1456,8 +1457,8 @@ SR_closed_hourly_ <- create_bar_chart(
   x_axis_title = NULL,
   y_axis_title = "count",
   print_out_title = "Hourly closed",
-  add_mean = FALSE,
-  add_median = TRUE,
+  add_mean = TRUE,
+  add_median = FALSE,
   add_sd = TRUE,
   chart_file_name = "Closed_Hourly_SR_count.png"
 )
@@ -2157,15 +2158,10 @@ if (num_rows_future > 0) {
 
 #########################################################################
 # Identify SRs created at midnight and noon
[…]

### 2024-05-30 code/Calculate year-month-day SR counts.R

@@ -14,11 +14,11 @@ library(lubridate)
 
 # Set the working directory to the "nyc311clean/code" directory to enable relative codes.
 setwd("C:/Users/david/OneDrive/Documents/nyc311clean/code")
-data1File <- file.path("..", "data", "10-year 2014-2023.csv")
+data1File <- file.path("..", "data", "2022-2023 created and closed only.csv")
 
 #Define the chart directory
-chart_directory_path <- file.path("C:", "Users", "david", "OneDrive", 
-                                  "Documents", "nyc311clean", "charts", "10-year trends")
+chart_directory_path <- 
+  file.path("..", "charts", "2022-2023 study", "Time lapse charts")
 
 programStart <- as.POSIXct(Sys.time())
 formattedStartTime <- format(programStart, "%Y-%m-%d %H:%M:%S")
@@ -145,17 +145,6 @@ max_rows_yearly <- max(yearly_summary$count)
 yearly_summary <- na.omit(yearly_summary)
 
 #########################################################################
-# Aggregate at the quarterly level
-d311$Quarter <- cut(d311$created_date, breaks = "quarter", labels = FALSE)
-
-# Calculate quarterly counts with corresponding quarter dates
-quarterly_summary <- d311 %>%
-  group_by(Quarter) %>%
-  summarise(count = n(), Quarter = min(created_date))
-max_rows_quarterly <- max(quarterly_summary$count)
-quarterly_summary <- na.omit(quarterly_summary)
-
-########################################################################                       
 # Aggregate at the monthly level
 monthly_summary <- d311 %>%
   group_by(YearMonth) %>%
@@ -243,10 +232,6 @@ closed_counts <- d311 %>%
 #########################################################################
 yearly_df <- as.data.frame(yearly_summary)
 
-quarterly_df <- select(quarterly_summary,Quarter,  count)
-quarterly_df$Quarter <- format(quarterly_df$Quarter, "%Y-%m-%d")
-quarterly_df <- as.data.frame(quarterly_df)
-
 monthly_df <- as.data.frame(monthly_summary)
 
 daily_df <- as.data.frame(daily_summary)
@@ -285,7 +270,7 @@ earliest_year_count <- yearly_df %>% filter(Year == min(Year)) %>% select(count)
[…]

### 2024-06-06 manuscript/311 manuscript.tex
Weekly updates

@@
 	An additional problem discovered with the created\_date and closed\_date fields; there appears to be an unusually large number of 
-	SRs created or closed at exactly midnight (00:00:00) and noon (12:00:00), to the second. The distribution of SR creation and closure 
-	largely follows the work-day clock with lots of SRs created during day-light hours, and fewer SRs 
-	created in the early hours of the morning. However, here there appears to be significantly greater numbers of SR closed exactly at midnight and exactly at noon,
-	as well as a significant number of SRs created exactly at noon. The reason for this anomaly is worth investigating 
-	as it affects the SR ``duration'' question; if SR durations are used to measure responsiveness
+	SRs created or closed at exactly midnight (00:00:00) and exactly noon (12:00:00), to the second. The distribution of SR creation and closure 
+	largely follows the work-day clock with many SRs created during day-light hours, and fewer SRs 
+	created at night and in the early hours of the morning. 
+	However, here there appears to be significantly greater numbers of SR closed exactly at midnight and exactly at noon,
+	as well as a significant number of SRs created exactly at noon. The reason behind this anomaly is worth investigating 
+	as it affects the SR ``duration'' question. If SR durations are used to measure responsiveness
 	for the delivery of services, and if those times are not an accurate reflection of the start/finish of the Service, then the use of
@@
 	\begin{itemize}
-		    \item There were 235,347 SRs closed exactly at midnight
+		    \item There were 99,779 SRs created exactly at noon (12:00:00)
+		    \item There were 235,347 SRs closed exactly at midnight (00:00:00)
 		    \item There were 105,505 SRs closed exactly at noon. 
-		    \item There were 99,779 SRs created exactly at noon 
-	\end{itemize}
-
-	Here is a visualization of SRs closed and created by the hour of the day :
+	\end{itemize}
+
+	There is, of course, a broad distribution of the time that SRs are created by day and hour, by night and day, by weekday and weekend. 
+	One would think that that distribution would be (mostly) evenly distributed by the minute and second. That is, 
+	if there are normally 100 SRs created between 1-2pm, that the distribution to the minute and second 
+	would be close to random. For example, there is no reason to suspect that a larger number of SRs would be 
+	created at 6 minutes past the hour rather than at 5 minutes past the hour. And certainly, if you look at the second that these SRs
+	are created you would again expect a (more-or-less) random distribution; that there would be no reason to believe 
+	that a higher (or lower) number of SRs would arrive at 18 seconds past the minute versus at 39 seconds past the minute.
+	Unfortunately, that turned out not to be true.
+
+	Let's look at the SR creation distribution. We discovered that there are a significant number of SRs created exactly at noon. To reveal this
+	trend, we looked at the exact date-time stamp of the created\_date(s). We began by selected only SRs that were created exactly on the minute;
+	at 09:00:00 or 13:00:00 for example.  Here is a visualization of the number of SRs created and closed by hour-of-the-day:
+
+	\begin{figure}[H]
+		 \centering
+		 \includegraphics[scale = 0.65]{Created_Hourly_SR_count.png}
+		 \caption{SRs by Hourly created\_date(s)}
+		 \label{fig:hourly-created}
+	\end{figure}
 
@@
 
-	\begin{figure}[H]
-		 \centering
-		 \includegraphics[scale = 0.65]{Created_Hourly_SR_count.png}
-		 \caption{SRs by Hourly created\_date(s)}
-		 \label{fig:hourly-created}
-	\end{figure}
-
-	The closed\_date chart show significant spikes in the number of SRs at both midnight
-	and noon with significant counts at exactly those times, to the second. The midnight spike is certainly suspect compared to the hours just before and after midnight where
-	there is normally a pronounced slump in SR closures. And both the closed-at-noon and closed-at-midnight are 
-	approaching 2\textsigma's from the hourly average. The created\_date shows a maximum value at noon also with a significant count occurring exactly at noon, again to 
-	the second.
-
-	These unusual patterns of created \& closed SRs at exactly the hours of midnight and noon perhaps indicates
-	the presence of a bulk process; one that perhaps automatically ``closes'' (or ``creates'') a large number of SRs with a provided
-	time-stamp of midnight (00:00:00) or noon (12:00:00). If so, then addressing this issue likely requires a closer look at those 
-	Agencies providing these suspect SR closed/created times. Clearly the desirable state is that the created\_date and closed\_date fields
-	reflect the true opening and closing of the SRs, such that an accurate ``duration'' can be computed which accurately
-	represents the SR life-cycle.
+	The created\_date shows a maximum value at 12:00. The closed\_date chart show significant spikes at both midnight
+	and noon. The midnight (00:00) spike appears to be an anomaly especially as compared to the hours just 
+	before and after midnight where there is normally a pronounced slump in SR closures. 
+	And both the closed-at-noon and closed-at-midnight are approaching 2\textsigma's from the hourly average.
+
+	Let's drill down on each of these outlier points. First, we aggregate the created\_date over the 2-year period that fall exactly
+	on the top-of-the-hour with the minute and second values equal to 0, e.g. 09:00:00, 14:00:00. This visualization
+	reveals a clearer picture of where these spikes are:
+
+	\begin{figure}[H]
+		 \centering
+		 \includegraphics[scale = 0.65]{SR_created_by_top_of_hour.png}
+		 \caption{SRs Created at Top of the Hour}
+		 \label{fig:tophourcreated}
+	\end{figure}
+
+	One other way to visualize this anomaly is to look at the busiest day during this 2-year period (Friday, 2023-09-29).
+	Here we will aggregated the created\_date(s) by minute (with seconds equal to zero). This provides a 
+	minute-by-minute look at SR creation on the busiest day of this study. Here we see a clear spike at 
+	exactly noon (12:00), well beyond the 3\textsigma line, and well above any other hour of the day.
+
+	\begin{figure}[H]
+		 \centering
+		 \includegraphics[width=\textwidth]{SR_created_by_minute_of_busiest_day.png}
+		 \caption{SRs created minute-by-minute on busiest day}
+		 \label{fig:busiestcreated}
+	\end{figure}
+
+	Let's take a similar approach to the closed\_date distribution over the 2-year period. We aggregate 
+	SRs by close\_date to find the top-of-the-hour, where minute and second values equal to 0, e.g. 20:00:00. 
+	Here we see spikes at both midnight (00:00) and at noon (12:00), with the midnight spike being 
+	well beyond the 3\textsigma line. Also note how much lower (and nearly equal) all the other
+	top-of-the-hour values are, relative to the midnight and noon spikes.
+
+	\begin{figure}[H]
+		 \centering
+		 \includegraphics[scale = 0.65]{SR_closed_by_top_of_hour.png}
+		 \caption{SRs Closed at Top of the Hour}
+		 \label{fig:tophourclosed}
+	\end{figure}
+
+	As before, let's visualize this anomaly by looking at the closed\_date(s) on the busiest day
+	during this 2-year study period (Friday, 2023-09-29). We aggregate closed\_date(s) by the
+	minute (with seconds equal to zero). This gives a minute-by-minute view of SR closure
+	on the busiest day of this study. Clearly visual is the very large closure spikes occurring
+	at both midnight and noon. The chart highlights both of these anomalies with spikes
+	of 286 and 120 at 00:00:00 and 12:00:00 respectively. Again, note how these spikes
+	are so very much above the median values for all other minutes.
+
+ 	\begin{figure}[H]
+		 \centering
+		 \includegraphics[width=\textwidth]{SR_closed_by_minute_of_busiest_day.png}
+		 \caption{SRs closed minute-by-minute on busiest day}
+		 \label{fig:busiestclosed}
+	\end{figure}
+
+	These unusual patterns of created \& closed SRs at exactly the hours of midnight and noon likely indicates
+	the presence of a bulk create/close software process; a process that perhaps automatically ``closes'' (or ``creates'') 
+	a large number of SRs with a provided time-stamp of midnight (00:00:00) or noon (12:00:00), and does so in bulk. 
+	If so, then addressing this issue likely requires a closer look at those Agencies providing these suspect SR closed/created times.
+	 The desirable state is that the created\_date and closed\_date fields correctly reflect the true opening and closing 
+	 of the SRs, such that an accurate ``duration'' can be computed for the SR life-cycle. Otherwise, we are likely seeing 
+	 durations that are artificially manipulated by a software process performing bulk uploads to the 311 system.
 
 	Looking at the distribution by Agency for the ``closed-exactly-at-midnight'' shows that \textgreater{}90\% of these suspect SRs come from just two Agencies - Dept. of 
-	Buildings (DOB) and Dept. of Sanitation (DSNY). If you include HPD and DOT, that accounts for 99\% of the suspect SR closures.
+	Buildings (DOB) and Dept. of Sanitation (DSNY). If you include HPD and DOT, that accounts for 99\% of the suspect SR closures. This Agency
+	distribution does not follow the overall SR distribution and we can assume that it is Agency specific.
 
@@
 
-	Hoever, the ``SR closed at noon'' visualization shows that  a single Agency, DSNY,
+	However, the ``SR closed at noon'' visualization shows that  a single Agency, DSNY,
 	is responsible for \textgreater{}99\% of the SRs ``closed-exactly-at-noon''. Accordingly, any effort to resolve this issue
-	necessitates working with DSNY to fully understand what is occurring.
+	necessitates working with DSNY to fully understand how this is occurring.
 
@@
 
-	It is difficult to determine if these observations are indeed anomalies. They could be the result of normal, expected
-	311 SR activity. But at least by observation that appears unlikely.
+	While it is difficult to ultimately know if these observations are indeed anomalies; they could be the result of normal, expected
+	311 SR activity. But at least by observation that outcome appears very unlikely.
 

### 2024-06-06 code/Calculate year-month-day SR counts.R

@@ -101,21 +101,33 @@ numrows <- nrow(d311)
 #########################################################################
 cat("\n\n**********DATA SUMMARY**********\n")
 
-# # Extract the date part of the character string
-# d311$extracted_created_date <- substr(d311$created_date, 1,10)
-# d311$extracted_closed_date <- substr(d311$closed_date, 1,10)
-
-# Convert character date-time strings to datetime objects
 # Convert character date-time strings to datetime objects with America/New_York timezone
 d311$created_date <- as.POSIXct(d311$created_date, format = "%m/%d/%Y %I:%M:%S %p", tz = "America/New_York")
 d311$closed_date <- as.POSIXct(d311$closed_date, format = "%m/%d/%Y %I:%M:%S %p", tz = "America/New_York")
 
 d311 <- d311[!is.na(d311$created_date),]
 
-num_years <- d311 %>%
-  mutate(year = year(created_date)) %>%
-  summarise(distinct_years = n_distinct(year)) %>%
-  pull(distinct_years)
+# # Create new columns for Year, Year-Month, Month, Day of Week, Day of Year, Minute, and Second
+ d311 <- d311 %>%
+   mutate(
+     Year = format(created_date, "%Y"),
+     YearMonth = format(created_date, "%Y-%m"),
+#     YearMonthDay = format(created_date, "%Y-%m-%d"),  # Year-Month-Da
+#     YearMonthDayHour = format(created_date, "%Y-%m-%d %H"), 
+#     Month = format(created_date, "%b"),                # Month as abbreviated name (Jan, Feb, Mar, ...)
+#     DayOfWeek = format(created_date, "%a"),            # Day of the week as abbreviated name (Mon, Tue, ...)
+#     DayOfYear = yday(created_date),                    # Day of the year (1, 2, ..., 365 or 366)
+#     Minute = minute(created_date),
+#     Hour = hour(created_date),
+#     Second = second(created_date)
+   )
+
+# Extract the years from the created_date
+years <- d311 %>%
+  mutate(year = format(created_date, "%Y")) %>%
+  select(year)
+
+num_years <- length(years) + 1
 
 cat("\nTotal rows:", format(numrows, big.mark = ","), "covering", num_years,"years")
 
@@ -129,12 +141,9 @@ latest_date_formatted <- format(latest_date, format = "%Y-%m-%d %H:%M:%S")
[…]

### 2024-06-10 manuscript/311 manuscript.tex
Weekly update

@@
 \begin{abstract}
-  Open data is important ...
-  Quality control is important ...
-  We use New York City 311 requests data  ...
-  We suggest some open data quality control protocols ...
+  In the early 21st century, the open data movement began to transform societies and governments through principles of transparency, 
+  innovation, and public engagement. New York City (NYC) has emerged as a leader in this movement 
+  with the enactment of the Open Data Law in 2012, leading to the creation of the NYC Open Data portal, 
+  which now hosts 2700 datasets from 80 city agencies. This resource has proven invaluable for research 
+  across various domains, including health, urban development, and transportation. 
+  The success of these initiatives underscores the importance of data curation, ensuring the utility and reliability of datasets.
+
+This paper examines the data curation challenges using the NYC 311 Service Request (SR) Data as a case study, 
+  addressing issues of data validity, consistency, and curation efficiency. Based on insights from this case study, 
+  we propose a set of data curation principles tailored for government-released open data. 
+  These principles aim to enhance data management practices and ensure the ongoing utility of open data. 
+  The paper concludes with actionable suggestions for improving data curation 
+  and offers general principles for the release of open data.
 
@@
 {\it Keywords:}
-consistency check;    
-data release protocol;
-data science;
-open data;
-data cleansing;
-quality control	
+Data science;
+Open data;
+Data cleansing;
+Quality control;
+Open Data Movement;
+Data Democratization;
+Transparency;
+Data Curation;
+Government Data;
+NYC Open Data;
+311 Service Request Data;
+Data Validity;
+Data Consistency;
+Data Efficiency;
+Smart Cities;
+Public Engagement;
+Research Data Management;
+
 \end{abstract}
@@
 requests (SRs) per year, encompassing such complaints as street noise, illegal parking, heat/hot water, abandoned vehicles,
-and unsanitary conditions. The data, managed through a tailored application built on the Microsoft Dynamics and Azure platform, is a valuable resource for city
+and unsanitary conditions. The data, managed through a tailored application built on the 
+Microsoft Dynamics and Azure platform, is a valuable resource for city
 administration and policy-making. The 311 data is publicly accessible through the 
@@
 or it is a system-wide issue that follows the general distribution of  SRs across the system. 
-Thus we can often identify if it is an Agency specific problem, or a systemic issue.
+Thus we can often identify if it is an Agency specific problem, or a systemic issue.  (Note: a few key NYC Agencies are not 
+represented in the 311 dataset. These include the New York City Housing Authority (NYCHA) and the
+Department of Corrections (DOC).)
 
@@
   \centering
-	  \includegraphics[scale = 0.65]{computed_columns_screenshot.png}
+	  \includegraphics[width = \textwidth] {computed_columns_screenshot.png}
 	  \caption{Screenshot of data portal showing the ``@computed'' columns}
@@
 
-		\subsubsection{Case Study: Noise Complaints by Zip Code}
+		\subsubsection{Case Study: Noise Complaints by Zip Code}{label{sec:case-study-zip-codes}
 		\textbf{Scenario:} The NYC Office of Nightlife (ONL) wants to know ``What are the top 10 zip codes for Noise Complaints (all 8 types) over the last two years?''
@@
 	field is completely valid, again in-line with the overall incident\_zip validation percentage (99.04\%). If the analysis performed using the zip\_codes field were to be
-	presented to the Drirectror, ONL, the conclusion would be in error. Perhaps more curious than the differing zip codes on the two lists, 
+	presented to the Director, ONL, the conclusion would be in error. Perhaps more curious than the differing zip codes on the two lists, 
 	is the fact that the numerical counts are nearly identical;  the overall counts are quite close. This is curious since
@@
 	 	 \centering
-		  \includegraphics[scale = 0.65]{invalid_community_boards.png}
+		  \includegraphics[width = \textwidth]{invalid_community_boards.png}
 		  \caption{Invalid Community Board by Agency}
@@
 	Due to how the community\_district data is formatted, it is not possible to establish validity. However, it is possible to determine that the dataset contains
-	72 unique entries, while there are only 59 valid Community Districts. So it would appear that there are 13 invalid commuity\_board(s) contained in the dataset.
-
-\section{Inconsistent \& Unusual Patterns with Dates}\label{sec:inconsistencies}
+	72 unique entries, while there are only 59 valid Community Districts. So it would appear that there are 13 invalid community\_board(s) contained in the dataset.
 
@@
  	 \centering
-	  \includegraphics[scale = 0.65]{negative_duration_SR_barchart.png}
+	  \includegraphics[width = \textwidth]{negative_duration_SR_barchart.png}
 	  \caption{SRs with a Negative Duration by Agency}
@@
  	 \centering
-	  \includegraphics[scale = 0.65]{negative_duration_SR_violin.png}
+	  \includegraphics[width = \textwidth]{negative_duration_SR_violin.png}
 	  \caption{Negative duration distribution}
@@
 		 	 \centering
-			  \includegraphics[scale = 0.65]{homeless_response_time_clean.png}
+			  \includegraphics[width = \textwidth]{homeless_response_time_clean.png}
 			  \caption{Homeless Assistance SR Durations}
@@
 		 \centering
-		 \includegraphics[scale = 0.65]{zero_duration_SR.png}
+		 \includegraphics[width = \textwidth]{zero_duration_SR.png}
 		 \caption{SRs with Zero Durations by Agency}
@@
 		 \centering
-		 \includegraphics[scale = 0.65]{Created_Hourly_SR_count.png}
+		 \includegraphics[width = \textwidth]{Created_Hourly_SR_count.png}
 		 \caption{SRs by Hourly created\_date(s)}
@@
 		 \centering
-		 \includegraphics[scale = 0.65]{Closed_Hourly_SR_count.png}
+		 \includegraphics[width = \textwidth]{Closed_Hourly_SR_count.png}
 		 \caption{SRs by Hourly closed\_date(s)}
@@
 		 \centering
-		 \includegraphics[scale = 0.65]{SR_created_by_top_of_hour.png}
+		 \includegraphics[width = \textwidth]{SR_created_by_top_of_hour.png}
 		 \caption{SRs Created at Top of the Hour}
@@
 		 \centering
-		 \includegraphics[scale = 0.65]{SR_closed_by_top_of_hour.png}
+		 \includegraphics[width = \textwidth]{SR_closed_by_top_of_hour.png}
 		 \caption{SRs Closed at Top of the Hour}
@@
 		 \centering
-		 \includegraphics[scale = 0.65]{closed_at_midnight_chart.png}
+		 \includegraphics[width = \textwidth]{closed_at_midnight_chart.png}
 		 \caption{SRs Closed Exactly at Midnight by Agency}
@@
 		 \centering
-		 \includegraphics[scale = 0.65]{closed_at_noon_chart.png}
+		 \includegraphics[width = \textwidth]{closed_at_noon_chart.png}
 		 \caption{SRs Closed Exactly at Noon by Agency}
@@
 		 \centering
-		 \includegraphics[scale = 0.65]{created_at_noon_chart.png}
+		 \includegraphics[width = \textwidth]{created_at_noon_chart.png}
 		 \caption{SRs Created Exactly at Noon by Agency}
@@
 		 \centering
-		 \includegraphics[scale = 0.65]{post_closed_violin.png}
+		 \includegraphics[width = \textwidth]{post_closed_violin.png}
 		 \caption{Post-Closed resolution\_action\_update\_date >30 days }
@@
 		 \centering
-		 \includegraphics[scale = 0.65]{postClosedBarChart.png}
+		 \includegraphics[width = \textwidth]{postClosedBarChart.png}
 		 \caption{Post-Closed resolution\_action\_update\_date >30 days by Agency }
@@
 \section{Redundant \& Duplicate fields}\label{sec:duplicates}
-During this analysis, several redundant fields were observed.
+During this analysis, several redundant fields were observed. Listed below is a discussion for each of a pair
+of data fields that are redundant or near-redundant. These fields should be examined further for possible
+consolidation in the overall dataset.
 
@@
 		 \centering
-		 \includegraphics[scale = 0.65]{non_matching_borough_boundaries_chart.png}
+		 \includegraphics[width = \textwidth]{non_matching_borough_boundaries_chart.png}
 		 \caption{Non-matching borough and borough\_boundary by Agency }
@@
 
+\subsection{landmark and street\_name}
+The landmark field is listed in the Data Dictionary as ``Can refer to any noteworthy location, including but not 
+limited to, parks, hospitals, airports, sports facilities, performance spaces, etc.'' We found that was not the case. To be sure
+many of the entries in the landmark field do contain landmark names, e.g.``Pennsylvania Station'', ``La Guardia Airport'', 
+``Gramercy Park'', the vast majority of the entries are street names, e.g. ``Fenton Avenue'', ``Steinway Street'', ``Broadway''.
+These entries appear to be very similar to those in the street\_name field. And in fact, the values of street\_name
+and landmark exactly match 62\% of the time, and while this is not a full duplicate, it is a precentage of 
+matching that is certainly indicative of duplicate usage. Even the non-matches (excluding blanks) would appear
+to be matches except for minor spelling and nomenclature changes, e.g. ``NINTH AVE'' \& ``9 AVE''.
+
+\begin{table}[ht]
+    \centering
+    \footnotesize
+    \begin{tabular}{>{\raggedright\arraybackslash}p{5cm} >{\raggedright\arraybackslash}p{5cm}}
+        \toprule
+        \textbf{street\_name} & \textbf{landmark} \\
+        \midrule
+        MACDOUGAL ST & MAC DOUGAL ST \\
+        NINTH AVE & 9 AVE \\
+        NORTH SIXTH ST & NORTH 6 ST \\
+        SOUTH FOURTH ST & SOUTH 4 ST \\
+        SIXTH AVE & 6 AVE \\
+        THIRD AVE & 3 AVE \\
+        EAST FIRST ST & EAST 1 ST \\
+        FOURTH AVE & 4 AVE \\
+        SAINT LAWRENCE AVE & ST LAWRENCE AVE \\
+        SOUTH FIRST ST & SOUTH 1 ST \\
+        BRIGHTON SEVENTH ST & BRIGHTON 7 ST \\
+        WEST FIFTH ST & WEST 5 ST \\
+        MT HOPE PL & MOUNT HOPE PL \\
+        PENN STA & PENNSYLVANIA STA \\
+        \bottomrule
+    \end{tabular}
+    \caption{Non-matches between 'street\_name' and 'landmark' fields}
+    \label{landmark}
+\end{table}
+
+The distribution by Agency for the non-matches somewhat follows the overall SR distribution (NYPD, DPR, DSNY), 
+but with some notable exceptions such as Taxi \& Limo Commission, Dept of Homeless Services (DHS),
+ and Dept of Health \& Mental Hygiene (DOHMH).
+
+\begin{figure}[H]
+		 \centering
+		 \includegraphics[width = \textwidth]{landmark_dup.png}
+		 \caption{Non-matching street\_name \& landmark by Agency }
+		 \label{fig:landmarkchart}
+	\end{figure}
+
 \subsection{cross\_street\_1 \& intersection\_street\_1 and cross\_street\_2 \& intersection\_street\_2}
-There are two sets of street pairs in the dataset:
+There are two sets of street pairs in the dataset in addition to the complaint address:
 
@@
 
-These two pairs of streets are used to help identify the location of the reported incident, as in common in
-New York to provide the street address and the cross street, as in ``24 W 90th street between Columbus and
+These two pairs of streets are used to help identify the location of the reported incident, as it is common in
+New York City to provide the street address and the cross street, as in ``24 W 90th street between Columbus and
 Amsterdam''. The Data Dictionary states that these streets are
@@
 
-Unfortunately, that is not the case. For the majority of data rows, both pair of fields are populated.
+Unfortunately, that is not the case. For the majority of data rows, both pair of fields are populated and are duplicates.
 In fact, 88\% of the time, cross\_street\_1 exactly matches intersection\_street\_1 and 
-cross\_street\_2 exactly matches intersection\_street\_2, and only 11.9\% are non-matches.
-
-(Note address standardization was applied to the street addresses using the R \emph{campfin} package. 
+cross\_street\_2 exactly matches intersection\_street\_2, Only 11.9\% are non-matches.
+
+\textit{(Note: Address standardization was applied to the street addresses using the R \emph{campfin} package. 
 This was done so as to prevent the software from generating a non-match between 
 streets such as ``240 E 69\textsuperscript{th} ST'' and ``240 E 69\textsuperscript{th} Street'' 
-which are clearly the same address, just some slight formatting and spelling differences.)
-
-The problem occurs when the two pairs of streets do not match, or when one is blank and the
-other is not. Which set is correct? If one is blank, say cross\_street\_1, but  intersection\_street\_1
+which are clearly the same address, just slight formatting and spelling differences.)}
+
+The problem occurs when the two pairs of streets do not match, such as when one is blank and the
+other is not. Which street is correct? If one is blank, say cross\_street\_1, but  intersection\_street\_1
 is not blank, should you use the non-blank one?  Let's look at some examples:
@@
 		 \centering
-		 \includegraphics[scale = 0.65]{non-matchingcross_street_1andintersection_street_1.png}
+		 \includegraphics[width = \textwidth]{non-matchingcross_street_1andintersection_street_1.png}
 		 \caption{Non-matching X\_Street\_1 \& Intersection\_Street\_1}
@@
 		 \centering
-		 \includegraphics[scale = 0.65]{non-matchingcross_street_2andintersection_street_2.png}
+		 \includegraphics[width = \textwidth]{non-matchingcross_street_2andintersection_street_2.png}
 		 \caption{Non-matching X\_Street\_2 \& Intersection\_Street\_2}
@@
 
-Near matches.
+In addition to the matching/non-matching values, we encountered a number of what you might call
+\emph{near-matches}. There were numerous cases where the cross\_street and intersection\_street
+would be close to identical, e.g. HARMON DR and HARMON RD. To identify these near-matches
+we applied a common method of determining matches by measuring 
+the \href{https://en.wikipedia.org/wiki/Hamming_distance}{Hamming Distance} between
+the two fields. The Hamming Distance is a measure of how many letters have to be changed
+in order for the two fields to match. For this analysis, we chose a Hamming Distance of 2.
+Here are the results for both cross\_street and intersection\_street.
 
@@
         \midrule
-        Matching                    & 5,644,365 & 88.26     \\
-        Both blank -- Matching      & 1,623,276 & 25.38     \\
+        Matching -- non-blank      & 4,012,683 & 62.75     \\
+        Matching -- both blank     & 1,623,276 & 25.38     \\
         Non-matching                &   750,543 & 11.74     \\
@@
         \midrule
-        Matching                    & 5,644,365 & 88.26     \\
-        Both blank -- Matching      & 1,623,276 & 25.38     \\
+        Matching -- non-blank         & 4,021,089 & 62.88     \\
+        Matching -- both blank  & 1,623,276 & 25.38     \\
         Non-matching                &   750,543 & 11.74     \\
@@
 
-Shrinking file size by deleting these 11 redundant fields:
-     1 - agency\_name 
-     2 - park\_borough 
-     3 - borough\_boundaries 
-     4 - location 
-     5 - intersection\_street\_1 
-     6 - intersection\_street\_2 
-     7 - police\_precinct 
-     8 - duration 
-     9 - postClosedUpdateDuration 
-     10 - translated\_borough\_boundaries 
-     11 - zip\_codes 
-Original size: 3.2 Gb 
-Size after removing redundant columns: 2.8 Gb 
-Potential size reduction: 395.3 Mb or 12.3\%
-
-\section{Improvements} \label{sec:improvements}
-
-Data validation: identify invalid data
-
-Data cleaning: convert to all lower/upper cases;
-
-Data organization: put zipcode/borough in a separate table; verify intersection
-== cross; remove redundant columns (and compare how much space saved)
+ \subsection{Shrinking file size by removing duplicates}
+
+ By removing duplicate and ``near-duplicate'' fields, it is possible to shrink the file size by 12.3\%
+ which for this dataset equates to a reduction of 395 Mb. A smaller file size means faster downloads,
+ less storage impact, as well as simplifying data analysis efforts.
+
+ Below is a list of duplicate and near-duplicate fields. There are some challenges with these proposed deletions, 
+ mostly with the ``near duplicate'' fields.  So while the park\_borough is a complete duplicate of the borough
+ field, the borough\_boundary field is only 98.3\% duplicate. Admittedly, loosing 1.7\% of the data in those two
+ fields is quite small, for this large dataset that amounts to 110, 715 non-matching occurrences. Similarly with 
+ police\_precincts and police\_precincts where there is a 99.9\% match; close to 100\% but not a complete match.
+
+Ultimately, it will require a deeper dive into the relationship between these duplicate and near-duplicate fields to
+ determine if they can be removed or not. Pending that, the authors would propose deleting the following fields
+ from the overall data set, which explanation provided.
+
+ 	\begin{itemize}
+		    \item agency\_name: Each row of data contains the agency field, which is an abbreviation of the Agency's name.
+		    The agency\_name abbreviations are clear, simple, and well understood. The authors believe the agency\_name field could
+		    be eliminated without loss of data quality.
+
+		    \item park\_borough:  This field is a 100\% match with the borough field. Removal of this field will not result in any loss
+		    of data or data quality.
+
+		    \item location:  The location field is simply a concatenation of the latitude and longitude fields, with a comma and 
+		    parenthesis added. The authors feel this field actually hinders data analysis as the
+		    field must be re-split into the two components (latitude and longitude) to conduct geographic analysis.
+		     The data is a 100\% duplication of the latitude and longitude fields. Removal of this field will
+		    not result in any loss of data.
+
+		    \item police\_precinct: This field has a 99.95\% match with the police\_precincts fields. We are unable to determine which
+		    field is more correct, but feel that removal of one of the precinct fields is prudent and would result in very minimal 
+		    data loss. (We should note that both police\_precinct and police\_precincts fields each have 34.5\% invalid data.)
+
+		   \item borough\_boundaries (computed field): This field is one of the six computed fields. It has a 98.3 match with the
+		    borough field. We recommend deleting this fields despite the somewhat limited loss of data.
+
+		    \item cross\_street\_1 \& 2 and intersection\_street\_1 \& 2: These two pairs of fields each have an 88\% match. We would
+		    recommend deleting the two intersection\_street fields while acknowledging some loss of data in doing so.
+
+		    \item zip\_codes(computed field):  This field is one of the six computed fields.  However, this field has error rate of
+		    58\% and as shown in the \ref{sec:case-study-zip-codes} can lead to dangerous mistakes when used for analytical purposes.
+		    We would recommend deleting this field.
+	\end{itemize}
+
+Additionally, it would be worth investigating if certain data fields are in-fact useful, as the population of these fields is very, very 
+scarce. For example, the taxi\_company\_borough field is 99.94\% blank, and the remaining entries match the borough field only
+0.05\% field. A similar situation exists for the road\_ramp field (99.75\% blank), the vehicle\_type field (99.71\%blank), and possibly 
+due\_date which is 99.62\% blank. There are several others: bridge\_highway\_direction, bridge\_highway\_name,
+bridge\_highway\_segment, and taxi\_pick\_up\_location...all of which have \textgreater99\% blanks. Most of those fields appear to be 
+almost exclusively used by the Taxi \& Limo Commission (TLC), which should be consulted to 
+determine if those fields are used and properly populated.
+
+\section{Data Dictionary Observations} \label{sec:datadictionay}
+During this analysis, it became apparent that the published Data Dictionary could use an update. We found many discrepancies between
+the Data Dictionary and the actual data. While not actually, incorrect or dirty data, it nonetheless can lead analytic efforts astray.
+In data analysis, it's essential to accurately describe the nature and purpose of each field to ensure proper handling and interpretation. 
+Here are some of the noted discrepancies:
+
+\begin{itemize}
+	\item While the Open Data portal lists the basic data types of the various fields, the Data Dictionary does not. Nor
+	does it provide meaningful information on the specifics of various fields, such as constraints or domain of legal values.
+	For example, the incident\_zip field is specified as ``text''. That's probably not the best definition, perhaps
+	``categorical'' would be more appropriate with a description indicating that the data can contain only numeric characters
+	and is not subjected to arithmetic operations. We found two instances with the zip\_code field contained character strings (``na'', ``N/A'').
+
+	\item Domain of legal values as contained in the Data Dictionary are incomplete. For example, the Data Dictionary indicates
+	that the status field has values of assigned, canceled, closed, or pending. However, we found additional status values:
+	in progress, started, and unspecified. Additionally, we discovered no SRs with a status of canceled. Similarly, the 
+	address\_type field is indicated to have values address, blockface, intersection, latlong, and placename. We discovered
+	additional values of bbl and unrecognized. No values of latlong were observed. Other fields suffer the same
+	inaccuracies (facility\_type, vehicle\_type, taxi\_pick\_up\_location, road\_ramp, city).
+\end{itemize}
+
+\section{Recommendations} \label{sec:recommendations}
+Having identified a number of data cleanliness related issues with the 311 Service Request (SR) dataset, what follows
+are a series of recommendations as to how to resolve these issues. Below are those recommendations;
+
+\begin{itemize}
+	\item First, and most importantly, the Open Data Team will need to undertake a special effort to identify these data anomalies,
+	confirm the veracity of the issue, and engage with the appropriate Agency Open Data Representatives to resolve these issues.
+	That effort will not be easy, fast, or cheap as in some software changes will likely be required. Perhaps hardest will be
+	working with the external City Agencies that feed data to the 311 SR dataset, but are not using the core 311 system.
+	Integrations and API usage will need to be investigated, and in many cases the changes required to correct an issue will
+	be on the source Agency side. Coordination and mutual approval of any changes will be key. Strong support from senior
+	leadership will be a critical success factor.
+
+	\item Secondly, standards will be to be identified, agreed to, and commitment to support those standards must be
+	obtained. For example, what are the domain of legal values for certain fields, especially those with identified issues?
+
+	\item Obviously, many of the data integrity issues can be resolved through rule enforcement by software. 
+	For example, the software could easily prevent invalid values, such as zip codes, 
+	from entering the system. Similarly for an entire range of data issues. One key area is the
+	created\_date and closed\_date which are frequently cited when determining 
+	responsiveness of various Agency services. The authors would argue that such dates 
+	should be driven by change in ``status'', not manually entered. For example, the closed\_date is captured
+	when the SR ``status'' is changed to ``closed'', and not manually entered. Such a change, along with many others, could prevent errors such
+	nras negative and  zero durations which significantly affect measurements of responsiveness.
+
+	\item The large spikes in SR closure and creation occurring at midnight and noon, almost certainly point to an underlying automated
+	process from an external-to-311 Agency. Such processes again distort the accurate capture of an SR duration. Data rule enforcement
+	at the software integration or API level should be investigated.
+
+	\item One possible way to have software enforce certain data standards, would be to move more non-311 Agencies to the core 311
+	software system. That system was completely upgraded in 2019, and should be (relatively) easy to expand to other Agencies and usages.
+
+	\item As noted, there are six ``computed'' fields in the data export that are not identified in the Data Dictionary. While these fields may be
+	``experimental'', they are nonetheless present in any data portal presentation or export. As noted in this study, a number of those computed
+	fields suffer from data inaccuracies. It is difficult to know what to recommend for these six field without further understanding
+	of their origin, computational method, or intended usage. We would recommend updating the Data Dictionary at a minimum and
+	possibly hiding those fields from public display until such time as their usage and accuracy can be ascertained.
+
+	\item Removal of duplicate fields should be examined.  In some cases, such a the borough and park\_borough fields, the duplication is 100%. 
+	In other cases, the duplication is in the 80-90\% range. These fields and their usage/utility should be carefully examined to determine if their
+	inclusion in the 311 SR dataset is meaningful, adds value not found in other fields, and should be removed or continued to be collected.
+	Initial analysis indicates that as much as 12\% size reductions are easily achievable with the removal of duplicate fields. A more thorough
+	review and understanding of the fields would no doubt yield more savings.
+
+	\item Fields with extremely low data presence, such as taxi\_company\_borough which is populated only 0.06\% of the time. Similarly for several other fields 
+	evaluated in this study which have a high percentage of blank/unknown values. If the presence of a small percentage of SRs 
+	with these fields being populated is deemed useful and necessary, then so be it. But if the usage of these fields is so diminished, 
+	then perhaps its inclusion in the 311 SR dataset is not warranted.
+
+	\item The Data Dictionary is in need of an upgrade and update. The last revision is indicated at 6 March 2023, well over a year ago. 
+	As noted in \textbackslash nameref: \nameref{sec:datadictionary} there are many problems with the Data Dictionary, including domain
+	values, absence of the six ``computed'' fields, and inaccurate description of the usage of several fields. Some of the Data Dictionary
+	entries can be undertaken immediately, which further revision should be addressed as part of a deeper data clean-up initiative.
+
+	\item The excessive precision in the latitude, longitude, and location fields is simply nonsensical.  We would recommend no more than four
+	decimal places (0.000X) which represents a distance of approximately 36.4 feet. Five decimals (0.0000X) represents 3.6 feet, which 
+	although possible, would be an extremely sophisticated and accurate geo-coding system. A 14 decimal number is simply a case
+	of a lot of precision with limited accuracy, and should be avoided.
+
+	\item The 10 address-related address fields (incident\_address, street\_name, city, intersection\_street\_1 \& 2, cross\_street\_1 \& 2,  landmark,
+	road\_ramp, bridge\_highway\_segment, and taxi\_pick\_up\_location should be in standard USPS format. 
+	There are several software packages which can standardize addresses to these standards.
+	We believe such an effort should be automatically applied to these fields at data entry. This step would also greatly improve any underlying geo-coding routines which
+	may be used to determine data values such as BBL, latitude, longitude, and New York State X/Y-plane coordinates.
+
+	\item The use of the value ``UNKNOWN'' , ``N/A'', ``NA'', and ``na'' when entered directly as text into a field should be very carefully examined, and
+	in all likelihood prohibited.  There are some indeed a few complaint\_type(s) where an ``N/A'' value is appropriate, such as a consumer\_complaint
+	about certain non-geo centric issues, such as a travel agency or tax preparation service . But even in those cases a value of ``N/A'' should be 
+	entered only via a drop-down menu selection. During this analysis, there were far too many data representations of missing data including: <N/A>, ``na'', `` ``,
+	UNKNOWN, and ``N/A''. It's challenging for analyst to have to account for so the many different methods of representing empty or null data fields.
+
+\end{itemize}
 
@@
 \end{document}
-
-\end{document}

### 2024-06-10 manuscript/manuscript.tex
reorg and some cleaning
Moved images to new images folder. Changed from png to pdf format.

--- 2024-06-10 manuscript/311 manuscript.tex
@@
 
+% image path
+\graphicspath{{.}{./images}}
+
 % control floats
@@
 
-\tableofcontents % Optional: Table of Contents
-\listoffigures % List of Figures
-\listoftables % List of Tables
+% \tableofcontents % Optional: Table of Contents
+% \listoffigures % List of Figures
+% \listoftables % List of Tables
 
@@
 transparency, innovation, and public engagement. 
-
 A landmark in this journey was the launch of the United States'
@@
 \href{https://data.worldbank.org}{World Bank Open Data}. 
-
 These initiatives represent significant strides in democratizing data, in breaking barriers that once
@@
 We illustrate these points with live examples drawn from our processing of the 311 SR data. 
-
 Secondly, building upon insights gained from this case study, we 
@@
   \centering
-  \includegraphics[width=\textwidth]{Yearly.png}
+  \includegraphics[width=\textwidth]{Yearly.pdf}
   \caption{Yearly SR counts over 10-year period}
@@
   \centering
-  \includegraphics[width=\textwidth]{Monthly.png}
+  \includegraphics[width=\textwidth]{Monthly.pdf}
   \caption{Monthly SR counts over 10-year period}
@@
   \centering
-  \includegraphics[width=\textwidth]{Daily.png}
+  \includegraphics[width=\textwidth]{Daily.pdf}
   \caption{Daily SR counts over 2022-2023 period}
@@
   \centering
-  \includegraphics[width=\textwidth]{Calendar-Month.png}
+  \includegraphics[width=\textwidth]{Calendar-Month.pdf}
   \caption{Calendar Month SR counts over a 10-year period}
@@
   \centering
-	  \includegraphics[width=\textwidth]{SRs_by_Agency.png}
+	  \includegraphics[width=\textwidth]{SRs_by_Agency.pdf}
 	  \caption{SR counts by Agency and Cumulative Percentage}
@@
   \centering
-	  \includegraphics[width=\textwidth]{SRs_by_Complaint_Type.png}
+	  \includegraphics[width=\textwidth]{SR_by_Complaint_Type.pdf}
 	  \caption{Top 20 SR complaint\_type and Cumulative Percentage}
@@
   \centering
-	  \includegraphics[width=\textwidth]{BlankFields.png}
+	  \includegraphics[width=\textwidth]{BlankFields.pdf}
 	  \caption{Number and Percentage of Empty Entries}
@@
 	  \centering
-		  \includegraphics[width=\textwidth]{invalid_incident_zip.png}
+		  \includegraphics[width=\textwidth]{invalid_incident_zip.pdf}
 		  \caption{Invalid incident-zip by Agency}
@@
 	  \centering
-		  \includegraphics[width=\textwidth]{invalid_police_precincts.png}
+		  \includegraphics[width=\textwidth]{invalid_police_precincts.pdf}
 		  \caption{Invalid NYPD precincts by Agency}
@@
 	 	 \centering
-		  \includegraphics[width = \textwidth]{invalid_community_boards.png}
+		  \includegraphics[width = \textwidth]{invalid_community_boards.pdf}
 		  \caption{Invalid Community Board by Agency}
@@
  	 \centering
-	  \includegraphics[width = \textwidth]{negative_duration_SR_barchart.png}
+	  \includegraphics[width = \textwidth]{negative_duration_SR_barchart.pdf}
 	  \caption{SRs with a Negative Duration by Agency}
@@
  	 \centering
-	  \includegraphics[width = \textwidth]{negative_duration_SR_violin.png}
+	  \includegraphics[width = \textwidth]{negative_duration_SR_violin.pdf}
 	  \caption{Negative duration distribution}
@@
 		 	 \centering
-			  \includegraphics[width = \textwidth]{homeless_response_time_clean.png}
+			  \includegraphics[width = \textwidth]{homeless_response_time_clean.pdf}
 			  \caption{Homeless Assistance SR Durations}
@@
 		 \centering
-		 \includegraphics[width = \textwidth]{zero_duration_SR.png}
+		 \includegraphics[width = \textwidth]{zero_duration_SR.pdf}
 		 \caption{SRs with Zero Durations by Agency}
@@
 		 \centering
-		 \includegraphics[width = \textwidth]{Created_Hourly_SR_count.png}
+		 \includegraphics[width = \textwidth]{Created_Hourly_SR_count.pdf}
 		 \caption{SRs by Hourly created\_date(s)}
@@
 		 \centering
-		 \includegraphics[width = \textwidth]{Closed_Hourly_SR_count.png}
+		 \includegraphics[width = \textwidth]{Closed_Hourly_SR_count.pdf}
 		 \caption{SRs by Hourly closed\_date(s)}
@@
 		 \centering
-		 \includegraphics[width = \textwidth]{SR_created_by_top_of_hour.png}
+		 \includegraphics[width = \textwidth]{SR_created_by_top_of_hour.pdf}
 		 \caption{SRs Created at Top of the Hour}
@@
 		 \centering
-		 \includegraphics[width=\textwidth]{SR_created_by_minute_of_busiest_day.png}
+		 \includegraphics[width=\textwidth]{SR_created_by_minute_of_busiest_day.pdf}
 		 \caption{SRs created minute-by-minute on busiest day}
@@
 		 \centering
-		 \includegraphics[width = \textwidth]{SR_closed_by_top_of_hour.png}
+		 \includegraphics[width = \textwidth]{SR_closed_by_top_of_hour.pdf}
 		 \caption{SRs Closed at Top of the Hour}
@@
 		 \centering
-		 \includegraphics[width=\textwidth]{SR_closed_by_minute_of_busiest_day.png}
+		 \includegraphics[width=\textwidth]{SR_closed_by_minute_of_busiest_day.pdf}
 		 \caption{SRs closed minute-by-minute on busiest day}
@@
 		 \centering
-		 \includegraphics[width = \textwidth]{closed_at_midnight_chart.png}
+		 \includegraphics[width = \textwidth]{closed_at_midnight_chart.pdf}
 		 \caption{SRs Closed Exactly at Midnight by Agency}
@@
 		 \centering
-		 \includegraphics[width = \textwidth]{closed_at_noon_chart.png}
+		 \includegraphics[width = \textwidth]{closed_at_noon_chart.pdf}
 		 \caption{SRs Closed Exactly at Noon by Agency}
@@
 		 \centering
-		 \includegraphics[width = \textwidth]{created_at_noon_chart.png}
+		 \includegraphics[width = \textwidth]{created_at_noon_chart.pdf}
 		 \caption{SRs Created Exactly at Noon by Agency}
@@
 		 \centering
-		 \includegraphics[width = \textwidth]{post_closed_violin.png}
+		 \includegraphics[width = \textwidth]{post_closed_violin.pdf}
 		 \caption{Post-Closed resolution\_action\_update\_date >30 days }
@@
 		 \centering
-		 \includegraphics[width = \textwidth]{postClosedBarChart.png}
+		 \includegraphics[width = \textwidth]{postClosedBarChart.pdf}
 		 \caption{Post-Closed resolution\_action\_update\_date >30 days by Agency }
@@
 		 \centering
-		 \includegraphics[width = \textwidth]{non_matching_borough_boundaries_chart.png}
+		 \includegraphics[width = \textwidth]{non_matching_borough_boundaries_chart.pdf}
 		 \caption{Non-matching borough and borough\_boundary by Agency }
@@
 		 \centering
-		 \includegraphics[width = \textwidth]{non-matchingcross_street_1andintersection_street_1.png}
+		 \includegraphics[width = \textwidth]{non-matchingcross_street_1andintersection_street_1.pdf}
 		 \caption{Non-matching X\_Street\_1 \& Intersection\_Street\_1}
@@
 		 \centering
-		 \includegraphics[width = \textwidth]{non-matchingcross_street_2andintersection_street_2.png}
+		 \includegraphics[width = \textwidth]{non-matchingcross_street_2andintersection_street_2.pdf}
 		 \caption{Non-matching X\_Street\_2 \& Intersection\_Street\_2}

### 2024-06-10 code/311 data cleansing V2.R

@@ -1,2776 +1,0 @@
-#########################################################################
-install.packages("ggplot2")
-install.packages("campfin")
-install.packages("stringr")
-install.packages("stringdist")
-install.packages("dply(raw data")
-install.packages("styler")
-install.packages("ggpmisc")
-install.packages("lubridate")
-install.packages("data.table")
-
-library(ggplot2)
-library(campfin)
-library(stringr)
-library(stringdist)
-library(dplyr)
-library(scales)
-library(ggpmisc)
-library(lubridate)
-library(data.table)
-
-setwd("C:/Users/david/OneDrive/Documents/nyc311clean/console_output")
-sink("console_output.txt")
-
-setwd("C:/Users/david/OneDrive/Documents/nyc311clean/code")
-
-setwd("C:/Users/david/OneDrive/Documents/nyc311clean/code")
-data1File <- file.path("..", "data", "311_2022-2023_AS_OF_01-31-2024.csv")
-
-# Hard code the max_closed_date to be midnight of the date of the data export from NYC Open Data
-max_closed_date <- as.POSIXct("2024-01-31 23:59:59", format = "%Y-%m-%d %H:%M:%S")
-
-chart_directory_path <- file.path("..", "charts", "2022-2023 study", "Core charts")
-
-writeFilePath <- file.path("..", "data", "smaller.csv")
-
-#########################################################################
-# Create the bar chart
-create_bar_chart <- function(
-    dataset,
-    x_col,
-    chart_title,
-    sub_title,
-    x_axis_title,
-    y_axis_title,
-    print_out_title,
-    add_mean = FALSE,
-    add_median = FALSE,
-    add_sd = FALSE,
-    chart_file_name) {
-  # Find the row with the maximum count
-  max_row <- dataset[which.max(dataset[[y_axis_title]]), ]
-  max_count <- max_row[[y_axis_title]]
-  total_count <- sum(dataset[[y_axis_title]], na.rm = TRUE)
-
-  # Find the row with the minimum count
-  min_row <- dataset[which.min(dataset[[y_axis_title]]), ]
-  min_count <- min_row[[y_axis_title]]
-
-  result <- calculate_values(max_count)
-  starting_value <- result$starting_value
-  increment <- result$increment
-  scaling_factor <- result$scaling_factor
-
-  # Calculate mean, median, and standard deviation if requested
-
[…]

### 2024-06-10 code/Calculate year-month-day SR counts.R

@@ -1,1299 +1,0 @@
-# install.packages("ggplot2")
-# install.packages("scales")
-# install.packages("dplyr")
-# install.packages('zoo')
-# install.packages("ggpmisc")
-# install.packages("lubridate")
-
-library(ggplot2)
-library(scales)
-library(dplyr)
-library(zoo)
-library(ggpmisc)
-library(lubridate)
-
-# Set the working directory to the "nyc311clean/code" directory to enable relative codes.
-setwd("C:/Users/david/OneDrive/Documents/nyc311clean/code")
-data1File <- file.path("..", "data", "2022-2023 created and closed only.csv")
-
-#Define the chart directory
-chart_directory_path <- 
-  file.path("..", "charts", "2022-2023 study", "Time lapse charts")
-
-programStart <- as.POSIXct(Sys.time())
-formattedStartTime <- format(programStart, "%Y-%m-%d %H:%M:%S")
-cat("\nExecution begins at:", formattedStartTime)
-
-#########################################################################
-calculate_values <- function(max_count) {
-  # Set scipen option to a large value
-  options(scipen = 999)
-  thresholds <- c(10, 
-                  100, 
-                  250, 
-                  500, 
-                  1000, 
-                  2500, 
-                  5000, 
-                  10000,
-                  15000,
-                  25000, 
-                  50000,
-                  100000, 
-                  500000, 
-                  1000000, 
-                  2500000, 
-                  5000000, 
-                  10000000,
-                  50000000, 
-                  100000000)
-  
-  values <- matrix(c(0, 1, 1,
-                     0, 20, 1,
-                     50, 50, 1,
-                     100, 100, 1,
-                     100, 200, 1,
-                     250, 250, 1,
-                     1000, 1000, 1,
-                     1000, 2000, 1000,
-                     2500, 2500, 1000,
-                     5000,5000,1000,
-                     5000, 10000, 100,
-                     10000, 20000, 1000,
-                     100000, 100000, 1000,
-                     100000, 200000, 1000,
-                     250000, 250000, 1000,
-                     500000, 500000, 1000,
-                     1000000, 1000000, 1000000,
-                     5000000, 5000000, 1000000,
-                     10000000, 20000000, 1000000), ncol = 3, byrow = TRUE)
-  
[…]

### 2024-06-13 manuscript/manuscript.tex
Updates

@@
 \usepackage{adjustbox}
-
 \usepackage[]{lineno}
 \linenumbers*[1]
+
 %% patches to make lineno work better with amsmath
@@
     round-mode = places,
-  round-precision = 0
+    round-precision = 2,
+    output-decimal-marker = {.},
+    table-number-alignment = center,
+    table-figures-integer = 6,
+    table-figures-decimal = 2,
+    table-figures-uncertainty = 2
 }
@@
 
-% \tableofcontents % Optional: Table of Contents
-% \listoffigures % List of Figures
-% \listoftables % List of Tables
+\tableofcontents % Optional: Table of Contents
+\listoffigures % List of Figures
+\listoftables % List of Tables
 
@@
 \noindent
-{\it Keywords:}
+\textit{Keywords:}
+311 Service Request Data;
+Data cleansing;
+Data Consistency;
+Data Curation;
+Data Democratization;
+Data Efficiency;
 Data science;
-Open data;
-Data cleansing;
-Quality control;
-Open Data Movement;
-Data Democratization;
-Transparency;
-Data Curation;
+Data Validity;
 Government Data;
 NYC Open Data;
-311 Service Request Data;
-Data Validity;
-Data Consistency;
-Data Efficiency;
+Open Data Movement;
+Open data;
+Public Engagement;
+Quality control;
+Research Data Management;
 Smart Cities;
-Public Engagement;
-Research Data Management;
-
+Transparency;
 \end{abstract}
@@
 Section~\ref{sec:issues} offers a general discussion of data cleansing issues
-impacting data quality and curation efficiency. Section~\ref{sec:Structural} examines
+impacting data quality and curation efficiency. Section~\ref{sec:structural} examines
 the technical dimensions of the dataset, the structural issues. In Section~\ref{sec:datatypes} we examine the data fields
@@
 administration and policy-making. The 311 data is publicly accessible through the 
-\href{https://opendata.cityofnewyork.us/}{NYC Open Data Portal} which provides access
+NYC Open Data Portal which provides access
 to all the data sets (approx. 2700) and provides a convenient method for querying, grouping, aggregating, geo-mapping, visualization, and exporting
@@
 
-\begin{figure}[H]
-  \centering
+\begin{figure}[tbp]
+  \centering
+  \caption{Yearly SR counts over 10-years}
   \includegraphics[width=\textwidth]{Yearly.pdf}
-  \caption{Yearly SR counts over 10-year period}
   \label{fig:yearly-counts}
@@
 
-This 10-year timeframe (2014-2023) includes the COVID pandemic. Note that the pandemic
+This 10-year timeframe (2014-2023) includes the COVID pandemic. INote that the pandemic
 impact reached a peak in the July-September 2020 timeframe, accompanied by additional spikes
@@
   \centering
-  \includegraphics[width=\textwidth]{Monthly.pdf}
-  \caption{Monthly SR counts over 10-year period}
+   \caption{Monthly SR counts over 10-years}
+  \includegraphics[width=\textwidth]{10-year-monthly.pdf}
   \label{fig:monthly-counts}
-\end{figure}
-
-A quick look at the corresponding  daily tend.  An average day is 8761 SRs.
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\textwidth]{Daily.pdf}
-  \caption{Daily SR counts over 2022-2023 period}
-  \label{fig:daily-counts}
-\end{figure}
-
-Seasonal trends are also prevalent. The months of November-thru-April are ``below average'' months, while the months May-thru-October are ``above average''. 
-This seasonal trend is observed throughout the 10-year period as visualized here.
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\textwidth]{Calendar-Month.pdf}
-  \caption{Calendar Month SR counts over a 10-year period}
-  \label{fig:calendar-months-counts}
 \end{figure}
@@
 
-\begin{table}[H]
+\begin{table}[tbp]
     \centering
     \small
-    \begin{tabular}{@{}lll|lll@{}}
+    \caption{Top 5 and Bottom 5 Days for SRs}
+    \begin{tabular}{@{}lll lll@{}}
         \toprule
-        \multicolumn{3}{c|}{Top 5 Days} & \multicolumn{3}{c}{Bottom 5 Days} \\
+        \multicolumn{3}{c}{Top 5 Days} & \multicolumn{3}{c}{Bottom 5 Days} \\
         \cmidrule(r){1-3} \cmidrule(l){4-6}
@@
     \end{tabular}
-    \caption{Top 5 and Bottom 5 Days}
     \label{tab:combined_counts}
@@
 
-\FloatBarrier % Ensure all floats before this point are processed before moving on
+%\FloatBarrier % Ensure all floats before this point are processed before moving on
 
@@
 
-\begin{figure}[H]
-  \centering
+\begin{figure}[tbp]
+	\centering
+  	\caption{SR counts by Agency and Cumulative Percentage}
 	\includegraphics[width=\textwidth]{SRs_by_Agency.pdf}  
-	  \caption{SR counts by Agency and Cumulative Percentage}
-
 \end{figure}
@@
 
-\begin{figure}[H]
-  \centering
+\begin{figure}[tbp]
+  \centering
+  	\caption{Top 20 SR complaint\_type and Cumulative Percentage}
 	  \includegraphics[width=\textwidth]{SR_by_Complaint_Type.pdf}  
-	  \caption{Top 20 SR complaint\_type and Cumulative Percentage}
 	  \label{fig:SR_complaints}
@@
 
-\begin{table}[H]
-    \centering
+\begin{table}[tbp]
+    \centering
+    \caption{Noise-related complaints\_type(s) by count with Agency}
     \footnotesize
-    \begin{tabular}{lS[table-format=7]S[table-format=2.2]l} % 'l' for left-aligned, 'S' for siunitx number alignment
+    \begin{tabular}{@{}lS[table-format=7.0,round-mode=places,round-precision=0]S[table-format=2.2,round-mode=places,round-precision=2]l@{}} % 'l' for left-aligned, 'S' for siunitx number alignment
         \toprule
@@
         \midrule
-        NOISE - RESIDENTIAL        & 675,502 & 10.56 & NYPD  \\ 
-        NOISE - STREET/SIDEWALK    & 300,507 &  4.70 & NYPD  \\ 
-        NOISE - COMMERCIAL         & 128,892 &  2.02 & NYPD  \\ 
-        NOISE - VEHICLE            & 115,956 &  1.81 & NYPD  \\ 
-        NOISE                      & 100,413 &  1.57 & DEP   \\ 
-        NOISE - HELICOPTER         &  85,345 &  1.33 & EDC   \\ 
-        NOISE - PARK               &  16,620 &  0.26 & NYPD  \\ 
+        NOISE - RESIDENTIAL        & 675502 & 10.56 & NYPD  \\ 
+        NOISE - STREET/SIDEWALK    & 300507 &  4.70 & NYPD  \\ 
+        NOISE - COMMERCIAL         & 128892 &  2.02 & NYPD  \\ 
+        NOISE - VEHICLE            & 115956 &  1.81 & NYPD  \\ 
+        NOISE                      & 100413 &  1.57 & DEP   \\ 
+        NOISE - HELICOPTER         &  85345 &  1.33 & EDC   \\ 
+        NOISE - PARK               &  16620 &  0.26 & NYPD  \\ 
         NOISE - HOUSE OF WORSHIP   &   2757 &  0.04 & NYPD  \\ 
@@
     \end{tabular}
-    \caption{Noise-related complaints\_type(s) by count with Agency}
     \label{tab:noisecomplaints}
@@
 
-\begin{figure}[H]
-  \centering
-	  \includegraphics[width = \textwidth] {computed_columns_screenshot.png}
+\begin{figure}[tbp]
+  \centering
   \caption{Screenshot of data portal showing the ``@computed'' columns}
+	  \includegraphics[width = 0.65\textwidth] {computed_columns_screenshot.pdf}	  
 	  \label{fig:computed-columns}
@@
 
-\begin{table}[ht]
-\centering
+\begin{table}[tbp]
+\centering
+\caption{Blank and N/A entries by Field}
 \resizebox{0.8\linewidth}{!}{%
@@
 }
-\caption{Blank and N/A entries by Field}
 \end{table}
 
-\FloatBarrier % Ensure all floats before this point are processed before moving on
+%\FloatBarrier % Ensure all floats before this point are processed before moving on
 
@@
 
-\begin{figure}[H]
-  \centering
+\begin{figure}[tbp]
+  \centering
+  	\caption{Number and Percentage of Empty Entries}
 	\includegraphics[width=\textwidth]{BlankFields.pdf}
-	  \caption{Number and Percentage of Empty Entries}
 	\label{fig:blank_fields}
@@
 	\subsection{Issues with Zip Codes}
-	\label{sec:zip-codes}
+	\label{sec:zipcodesissues}
 	 Unfortunately some fields proved to be  problematic when comparing the values to a domain of legal values.
@@
 
-	\begin{figure}[H]
-	  \centering
+	\begin{figure}[tbp]
+	  \centering
+	  	\caption{Invalid incident-zip by Agency}
 		\includegraphics[width=\textwidth]{invalid_incident_zip.pdf} 
-		  \caption{Invalid incident-zip by Agency}
 		\label{fig:invalid_incident_zip}
@@
 
-	\begin{figure}[H]
-	    		\centering
-	    		\includegraphics[scale = 0.6]{zipcode_casestudy.png}
-	    		\caption{Top 10 zip codes for Noise complaints (Zip Codes)}
-	    	\label{fig:casestudy1-zipcodes}
-		\end{figure}
-
 		However, this analysis uses the zip\_codes field, one of the (six) computed fields that has shown to have validity problems. If we repeat the analysis with the 
@@
 
-		\begin{figure}[H]
-	    		\centering
-	    		\includegraphics[scale = 0.6]{incident_zip_casestudy.png}
-	    		\caption{Top 10 zip codes for Noise complaints(Incident Zip)}
-	    	\label{fig:casestudy1-incident-zip}
-		\end{figure}
-
 		Let's subject these two data sets to validation against the US Postal Service zip code database.
 
-	\begin{table}[H]
+\begin{table}[tbp]
     \centering
     \footnotesize
-	\begin{tabular}{lrr@{\hskip 0.5cm}|@{\hskip 0.5cm}lrr}
+    \caption{Comparison of Top Ten Zip Codes Lists}
+    \begin{tabular}{@{}lS[table-format=6.0,round-mode=places,round-precision=0]c@{\hskip 0.5cm}@{}lS[table-format=6.0,round-mode=places,round-precision=0]c@{}}
         \toprule
@@
         \midrule
-		11275 & 104,556 & FALSE & 10466 & 104,562 & TRUE \\
-		12420 & 27,503 & TRUE & 10023 & 27,972 & TRUE \\
-		12428 & 26,564 & TRUE & 10031 & 25,548 & TRUE \\
-		10935 & 25,508 & FALSE & 10457 & 25,066 & TRUE \\
-		10934 & 23,448 & FALSE & 10453 & 24,752 & TRUE \\
-		10931 & 22,381 & TRUE & 10456 & 24,751 & TRUE \\
-		10930 & 22,121 & TRUE & 10452 & 22,527 & TRUE \\
-		17613 & 21,963 & FALSE & 10025 & 21,705 & TRUE \\
-		10936 & 21,707 & FALSE & 10458 & 21,689 & TRUE \\
-		11606 & 21,435 & FALSE & 10032 & 20,622 & TRUE \\
+        11275 & 104556 & FALSE & 10466 & 104562 & TRUE \\
+        12420 & 27503 & TRUE & 10023 & 27972 & TRUE \\
+        12428 & 26564 & TRUE & 10031 & 25548 & TRUE \\
+        10935 & 25508 & FALSE & 10457 & 25066 & TRUE \\
+        10934 & 23448 & FALSE & 10453 & 24752 & TRUE \\
+        10931 & 22381 & TRUE & 10456 & 24751 & TRUE \\
+        10930 & 22121 & TRUE & 10452 & 22527 & TRUE \\
+        17613 & 21963 & FALSE & 10025 & 21705 & TRUE \\
+        10936 & 21707 & FALSE & 10458 & 21689 & TRUE \\
+        11606 & 21435 & FALSE & 10032 & 20622 & TRUE \\
         \bottomrule
     \end{tabular}
-	\caption{Comparison of Top Ten Zip Codes Lists}
+    \label{tab:zipcodes}
 \end{table}
@@
 
-	\begin{table}[ht]
+	\begin{table}[tbp]
 	\footnotesize
 	\centering
+	\caption{Comparison of the fields police\_precinct and police\_precincts}
 		\begin{tabular}{ccccc}
@@
 	\end{tabular}
-	\caption{Comparison of the fields police\_precinct and police\_precincts}
 	\label{tab:comparison-precincts-diff}
@@
 
-	\begin{figure}[H]
-	  \centering
+	\begin{figure}[tbp]
+	  \centering
+	  \caption{Invalid NYPD precincts by Agency}
 		  \includegraphics[width=\textwidth]{invalid_police_precincts.pdf}	  
-		  \caption{Invalid NYPD precincts by Agency}
 		  \label{fig:invalid_police_precincts_zip}
@@
 
-		\begin{table}[H]
-		\centering
+		\begin{table}[tbp]
+		\centering
+		\caption{Top Ten Invalid 'community\_board' Values}
 		\footnotesize
@@
 		\end{tabular}
-		\caption{Top Ten Invalid 'community\_board' Values}
 		\end{table}
@@
 
-		\begin{figure}[H]
+		\begin{figure}[tbp]
 	 	 \centering
@@
 
-\begin{table}[H]
-    \centering
+\begin{table}[tbp]
+    \centering
+    \caption{Largest and Smallest errors (days)}
     \small
@@
 	        \midrule
-	        2023-01-27 14:40:00 & 2022-01-14 14:40:00 & -378.000000 & DOT \\
-	        2023-01-18 10:06:00 & 2022-01-12 10:06:00 & -371.000000 & DOT \\
-	        2023-01-27 14:36:00 & 2022-01-22 14:35:00 & -370.000694 & DOT \\
-	        2023-01-11 11:10:00 & 2022-01-09 11:10:00 & -367.000000 & DOT \\
-	        2023-12-18 03:13:00 & 2023-01-16 13:10:00 & -335.585417 & DOT \\
+	        2023-01-27 14:40:00 & 2022-01-14 14:40:00 & -378.0 & DOT \\
+	        2023-01-18 10:06:00 & 2022-01-12 10:06:00 & -371.0 & DOT \\
+	        2023-01-27 14:36:00 & 2022-01-22 14:35:00 & -370.0 & DOT \\
+	        2023-01-11 11:10:00 & 2022-01-09 11:10:00 & -367.0 & DOT \\
+	        2023-12-18 03:13:00 & 2023-01-16 13:10:00 & -335.6 & DOT \\
 	        \midrule
@@
 	        \midrule
-	        2023-06-28 07:07:31 & 2023-06-28 07:07:00 & -0.000359 & DOT \\
-	        2023-06-29 09:10:20 & 2023-06-29 09:10:00 & -0.000231 & DOT \\
-	        2023-01-12 06:50:13 & 2023-01-12 06:50:00 & -0.000150 & DOT \\
-	        2023-06-26 08:09:07 & 2023-06-26 08:09:00 & -0.000081 & DOT \\
-	        2023-01-12 06:51:01 & 2023-01-12 06:51:00 & -0.000012 & DOT \\
+	        2023-06-28 07:07:31 & 2023-06-28 07:07:00 & -0.00036 & DOT \\
+	        2023-06-29 09:10:20 & 2023-06-29 09:10:00 & -0.00023 & DOT \\
+	        2023-01-12 06:50:13 & 2023-01-12 06:50:00 & -0.00015 & DOT \\
+	        2023-06-26 08:09:07 & 2023-06-26 08:09:00 & -0.00008 & DOT \\
+	        2023-01-12 06:51:01 & 2023-01-12 06:51:00 & -0.00001 & DOT \\
 	        \bottomrule
     \end{tabular}
-    \caption{Largest and Smallest errors (days)}
     \label{tab:combined_errors}
@@
 
-\begin{table}[H]
-    \centering
+\begin{table}[tbp]
+    \centering
+    \caption{Sample of SRs with extremely large negative durations}
     \small
@@
     \end{tabular}
-    \caption{Sample of SRs with extremely large negative durations}
     \label{tab:extreme_negative_durations}
@@
 
-\begin{figure}[H]
- 	 \centering
+\begin{figure}[tbp]
+ 	 \centering
+ 	 \caption{SRs with a Negative Duration by Agency}
 	  \includegraphics[width = \textwidth]{negative_duration_SR_barchart.pdf}
-	  \caption{SRs with a Negative Duration by Agency}
 	  \label{fig:negative-duration}
@@
 
-\begin{figure}[H]
- 	 \centering
+\begin{figure}[tbp]
+ 	 \centering
+ 	 \caption{Negative duration distribution}
 	  \includegraphics[width = \textwidth]{negative_duration_SR_violin.pdf}
-	  \caption{Negative duration distribution}
 	  \label{fig:negative-duration-violin}
@@
 
-		\begin{figure}[H]
-		 	 \centering
+		\begin{figure}[tbp]
+		 	 \centering
+		 	 \caption{Homeless Assistance SR Durations}
 			  \includegraphics[width = \textwidth]{homeless_response_time_clean.pdf}
-			  \caption{Homeless Assistance SR Durations}
 			  \label{fig:homeless}
@@
 
-	\begin{figure}[H]
-		 \centering
+	\begin{figure}[tbp]
+		 \centering
+		  \caption{SRs with Zero Durations by Agency}
 		 \includegraphics[width = \textwidth]{zero_duration_SR.pdf}
-		 \caption{SRs with Zero Durations by Agency}
 		 \label{fig:sero-duration}
@@
 
-	\begin{figure}[H]
-		 \centering
+	\begin{figure}[tbp]
+		 \centering
+		 \caption{SRs by Hourly created\_date(s)}
 		 \includegraphics[width = \textwidth]{Created_Hourly_SR_count.pdf}
-		 \caption{SRs by Hourly created\_date(s)}
 		 \label{fig:hourly-created}
@@
 
-	\begin{figure}[H]
-		 \centering
+	\begin{figure}[tbp]
+		 \centering
+		 \caption{SRs by Hourly closed\_date(s)}
 		 \includegraphics[width = \textwidth]{Closed_Hourly_SR_count.pdf}
-		 \caption{SRs by Hourly closed\_date(s)}
 		 \label{fig:hourly-closed}
@@
 
-	\begin{figure}[H]
-		 \centering
+	\begin{figure}[tbp]
+		 \centering
+ 		 \caption{SRs Created at Top of the Hour}
 		 \includegraphics[width = \textwidth]{SR_created_by_top_of_hour.pdf}
-		 \caption{SRs Created at Top of the Hour}
 		 \label{fig:tophourcreated}
@@
 
-	\begin{figure}[H]
-		 \centering
+	\begin{figure}[tbp]
+		 \centering
+		 \caption{SRs created minute-by-minute on busiest day}
 		 \includegraphics[width=\textwidth]{SR_created_by_minute_of_busiest_day.pdf}
-		 \caption{SRs created minute-by-minute on busiest day}
 		 \label{fig:busiestcreated}
@@
 
-	\begin{figure}[H]
-		 \centering
+	\begin{figure}[tbp]
+		 \centering
+		 \caption{SRs Closed at Top of the Hour}
 		 \includegraphics[width = \textwidth]{SR_closed_by_top_of_hour.pdf}
-		 \caption{SRs Closed at Top of the Hour}
 		 \label{fig:tophourclosed}
@@
 
- 	\begin{figure}[H]
-		 \centering
+ 	\begin{figure}[tbp]
+		 \centering
+		  \caption{SRs closed minute-by-minute on busiest day}
 		 \includegraphics[width=\textwidth]{SR_closed_by_minute_of_busiest_day.pdf}
-		 \caption{SRs closed minute-by-minute on busiest day}
 		 \label{fig:busiestclosed}
@@
 
-	\begin{figure}[H]
-		 \centering
+	\begin{figure}[tbp]
+		 \centering
+		 \caption{SRs Closed Exactly at Midnight by Agency}
 		 \includegraphics[width = \textwidth]{closed_at_midnight_chart.pdf}
-		 \caption{SRs Closed Exactly at Midnight by Agency}
 		 \label{fig:midnight-closed}
@@
 
-	\begin{figure}[H]
-		 \centering
+	\begin{figure}[tbp]
+		 \centering
+		  \caption{SRs Closed Exactly at Noon by Agency}
 		 \includegraphics[width = \textwidth]{closed_at_noon_chart.pdf}
-		 \caption{SRs Closed Exactly at Noon by Agency}
 		 \label{fig:noon-closed}
@@
 
-	\begin{figure}[H]
-		 \centering
+	\begin{figure}[tbp]
+		 \centering
+		 \caption{SRs Created Exactly at Noon by Agency}
 		 \includegraphics[width = \textwidth]{created_at_noon_chart.pdf}
-		 \caption{SRs Created Exactly at Noon by Agency}
 		 \label{fig:noon-created}
@@
 
-	\begin{table}[H]
-	    \centering
+	\begin{table}[tbp]
+	    \centering
+	    \caption{Extremely long Resolution Updates}
 	    \footnotesize
@@
 	    \end{tabular}
-	    \caption{Extremely long Resolution Updates}
 	    \label{tab:resolution-updates}
@@
 
-	\begin{table}[H]
-    \centering
+	\begin{table}[tbp]
+    \centering
+     \caption{Long post-closed Resolution Updates}
     \footnotesize
@@
 	    \end{tabular}
-    \caption{Long post-closed Resolution Updates}
     \label{tab:resolution}
@@
 
-	\begin{figure}[H]
-		 \centering
+	\begin{figure}[tbp]
+		 \centering
+		 \caption{Post-Closed resolution\_action\_update\_date >30 days }
 		 \includegraphics[width = \textwidth]{post_closed_violin.pdf}
-		 \caption{Post-Closed resolution\_action\_update\_date >30 days }
 		 \label{fig:resolution-violin}
@@
 
-	\begin{figure}[H]
-		 \centering
+	\begin{figure}[tbp]
+		 \centering
+		 \caption{Post-Closed resolution\_action\_update\_date >30 days by Agency }
 		 \includegraphics[width = \textwidth]{postClosedBarChart.pdf}
-		 \caption{Post-Closed resolution\_action\_update\_date >30 days by Agency }
 		 \label{fig:resolution-by-agency}
@@
 
-	\begin{figure}[H]
-		 \centering
+	\begin{figure}[tbp]
+		 \centering
+		  \caption{Non-matching borough and borough\_boundary by Agency }
 		 \includegraphics[width = \textwidth]{non_matching_borough_boundaries_chart.pdf}
-		 \caption{Non-matching borough and borough\_boundary by Agency }
 		 \label{fig:borough-boundaries}
@@
     \centering
+    \caption{Non-matches between 'street\_name' and 'landmark' fields}
     \footnotesize
@@
     \end{tabular}
-    \caption{Non-matches between 'street\_name' and 'landmark' fields}
     \label{landmark}
@@
 
-\begin{figure}[H]
-		 \centering
-		 \includegraphics[width = \textwidth]{landmark_dup.png}
+\begin{figure}[tbp]
+		 \centering
 		  \caption{Non-matching street\_name \& landmark by Agency }
+		 \includegraphics[width = \textwidth]{non-matchingstreet_nameandlandmark.pdf}
 		 \label{fig:landmarkchart}
@@
 
-\begin{table}[H]
-    \centering
+\begin{table}[tbp]
+    \centering
+     \caption{Matching/Non-Matching cross\_street\_1 \& intersection\_street\_1}
     \footnotesize
@@
     \end{tabular}
-    \caption{Matching/Non-Matching cross\_street\_1 \& intersection\_street\_1}
     \label{tab:streets1}
@@
 
-	\begin{figure}[H]
-		 \centering
+	\begin{figure}[tbp]
+		 \centering
+		 \caption{Non-matching X\_Street\_1 \& Intersection\_Street\_1}
 		 \includegraphics[width = \textwidth]{non-matchingcross_street_1andintersection_street_1.pdf}
-		 \caption{Non-matching X\_Street\_1 \& Intersection\_Street\_1}
 		 \label{fig:xstreet1}
@@
 
-	\begin{figure}[H]
-		 \centering
+	\begin{figure}[tbp]
+		 \centering
+		 \caption{Non-matching X\_Street\_2 \& Intersection\_Street\_2}
 		 \includegraphics[width = \textwidth]{non-matchingcross_street_2andintersection_street_2.pdf}
-		 \caption{Non-matching X\_Street\_2 \& Intersection\_Street\_2}
 		 \label{fig:xstreet2}
@@
 
-\begin{table}[H]
-    \centering
+\begin{table}[tbp]
+    \centering
+     \caption{Near-matches for cross\_street\_1, intersection\_street\_1}
     \footnotesize
@@
     \end{tabular}
-    \caption{Near-matches for cross\_street\_1, intersection\_street\_1}
      \label{tab:x1nearmatches}
@@
 
-\begin{table}[H]
-    \centering
+\begin{table}[tbp]
+    \centering
+    \caption{Summary of cross\_street\_1 and intersection\_street\_1}
     \footnotesize
@@
     \end{tabular}
-    \caption{Summary of cross\_street\_1 and intersection\_street\_1}
     \label{tab:summary1}
@@
 
-\begin{table}[H]
-    \centering
+\begin{table}[tbp]
+    \centering
+    \caption{Summary of cross\_street\_2 and intersection\_street\_2}
     \footnotesize
@@
     \end{tabular}
-    \caption{Summary of cross\_street\_2 and intersection\_street\_2}
     \label{tab:summary_cross_intersection_2}
@@
 
-\begin{table}[H]
-    \centering
-    \footnotesize
-    \begin{tabular}{l r r}
-        \toprule
-        \textbf{category} & \textbf{count} & \textbf{percentage} \\
-        \midrule
-        Matching -- non-blank         & 4,021,089 & 62.88     \\
-        Matching -- both blank  & 1,623,276 & 25.38     \\
-        Non-matching                &   750,543 & 11.74     \\
-        cross\_street\_2\_blank     &   222,788 & N/A       \\
-        intersection\_street\_2\_blank &   517,431 & N/A       \\
-        Near-match                  &     1,500 & 0.023456  \\
-        \bottomrule
-    \end{tabular}
-    \caption{Summary of cross\_street\_2 and intersection\_street\_2}
-    \label{tab:summary_cross_intersection_2}
-\end{table}
-
-\begin{table}[H]
-    \centering
+\begin{table}[tbp]
+    \centering
+     \caption{Sample of near-matching cross\_street\_2 and intersection\_street\_2 (both non-blank)}
     \footnotesize
@@
     \end{tabular}
-    \caption{Sample of near-matching cross\_street\_2 and intersection\_street\_2 (both non-blank)}
     \label{tab:near_matching_cross_intersection_2}

### 2024-06-13 code/datacleansing.R

@@ -19,20 +19,15 @@ library(ggpmisc)
 library(lubridate)
 library(data.table)
 
-setwd("C:/Users/david/OneDrive/Documents/datacleaningproject/nyc311clean/console_output")
 sink("console_output.txt")
 
-setwd("C:/Users/david/OneDrive/Documents/datacleaningproject/nyc311clean/code")
-data1File <- file.path("..", "data", "311_2022-2023_AS_OF_01-31-2024.csv")
-chart_directory_path <- file.path("..", "charts", "2022-2023 study", "Core charts")
+data1File <- file.path("..", "..", "data", "311_2022-2023_AS_OF_01-31-2024.csv")
+chart_directory_path <- file.path("..", "..", "charts", "2022-2023 study", "Core charts")
 
-writeFilePath <- file.path("..", "data", "smaller.csv")
-
+writeFilePath <- file.path("..", "..", "data", "smaller.csv")
 
 # Hard code the max_closed_date to be midnight of the date of the data export from NYC Open Data
 max_closed_date <- as.POSIXct("2024-01-31 23:59:59", format = "%Y-%m-%d %H:%M:%S")
-
-
 
 #########################################################################
 # Create the bar chart
@@ -97,7 +92,7 @@ create_bar_chart <- function(
 
   # Create the bar chart
   bar_chart <- ggplot(dataset, aes(x = .data[[x_col]], y = .data[[y_axis_title]])) +
-    geom_bar(stat = "identity", fill = "cadetblue") +
+    geom_bar(stat = "identity", fill = "#009E73") +
     theme(
       axis.title.x = element_text(vjust = 0, size = 11),
       axis.title.y = element_text(vjust = 1, size = 11),
@@ -119,41 +114,41 @@ create_bar_chart <- function(
     # Add annotations for min and max values
     annotate("text",
       x = min_row[[x_col]], y = min_count,
-      label = "Min", size = 4, color = "gray21", vjust = 0.4, hjust = -0.6
+      label = "Min", size = 4, color = "gray20", vjust = 0.4, hjust = -0.6
     ) +
     annotate("text",
       x = max_row[[x_col]], y = max_count,
-      label = "Max", size = 4, color = "gray21", vjust = 0.4, hjust = -0.6
+      label = "Max", size = 4, color = "gray20", vjust = 0.4, hjust = -0.6
     )
 
   # Add optional elements
[…]

### 2024-06-13 code/timelinecharts..R

@@ -12,13 +12,10 @@ library(zoo)
 library(ggpmisc)
 library(lubridate)
 
-# Set the working directory to the "nyc311clean/code" directory to enable relative codes.
-setwd("C:/Users/david/OneDrive/Documents/datacleaningproject/nyc311clean/code")
-data1File <- file.path("..", "data", "2022-2023 created and closed only.csv")
+data1File <- file.path("..", "..", "data", "10-year 2014-2023.csv")
 
 #Define the chart directory
-chart_directory_path <- 
-  file.path("..", "charts", "2022-2023 study", "Time lapse charts")
+chart_directory_path <- file.path("..", "..", "charts", "2022-2023 study", "10-year trends")
 
 programStart <- as.POSIXct(Sys.time())
 formattedStartTime <- format(programStart, "%Y-%m-%d %H:%M:%S")
@@ -302,7 +299,7 @@ earliest_year <- min(yearly_df$Year)
 
 # Create the bar chart with vertical X-axis labels
 SR_yearly <- ggplot(yearly_df, aes(x = Year, y = count)) +
-  geom_bar(stat = "identity", fill = "cadetblue") +
+  geom_bar(stat = "identity", fill = "#009E73") +
   scale_y_continuous(labels = scales::comma) +
   theme(
     axis.title.x = element_text(vjust = 0, size = 11),
@@ -325,7 +322,7 @@ SR_yearly <- ggplot(yearly_df, aes(x = Year, y = count)) +
   geom_point(color = "transparent") +
   scale_x_continuous(breaks = unique(yearly_df$Year)) +
   stat_poly_eq(use_label(c("R2"))) +
-  geom_smooth(method = "lm", span = 1, se = FALSE, color = "firebrick3", linetype = "dotted", linewidth = 1) +
+  geom_smooth(method = "lm", span = 1, se = FALSE, color = "black", linetype = "dotted", linewidth = 2) +
   labs(x=NULL, y=NULL)
 
 # Print the bar chart
@@ -394,7 +391,13 @@ start_date <- min(monthly_df$YearMonth)
 end_date <- max(monthly_df$YearMonth)
 
 # Create a sequence of every other month
-breaks_seq <- seq(from = start_date, to = end_date, by = "2 months")
+
+# Ensure YearMonth is a Date object
+monthly_df$YearMonth <- as.Date(monthly_df$YearMonth)
+
[…]

### 2024-06-14 manuscript/manuscript.tex
Update

@@
 
-One area in particular which is heavily influencing customer outreach for the 311 system is the appropriate use of
-online chatbots and assistants. Increasing these features are influenced by artificial intelligence (AI) advances. It is a given that
-a number of citizen's interactions with City government could be handled accurately and quickly by AI-driven
-applications thus improving timeliness and accuracy. But how to do that while preserving customer privacy and anonymization
-and user engagement are challenging issues.
+One area in particular which is heavily influencing customer outreach for the 311 system
+ is the appropriate use of online chatbots and assistants. Increasing these features are 
+ influenced by artificial intelligence (AI) advances. It is a given that a number of 
+ citizen's interactions with City government could be handled accurately and quickly by AI-driven
+applications thus improving timeliness and accuracy. But how to do that while 
+preserving customer privacy and anonymization and user engagement are challenging issues.
 
@@
 observe system performance over a longer time-frame . During that time-frame, 
-the number of 311 SRs grew 50\%.
-
-\begin{figure}[tbp]
-  \centering
-  \caption{Yearly SR counts over 10-years}
-  \includegraphics[width=\textwidth]{Yearly.pdf}
-  \label{fig:yearly-counts}
-\end{figure}
-
-This 10-year timeframe (2014-2023) includes the COVID pandemic. INote that the pandemic
-impact reached a peak in the July-September 2020 timeframe, accompanied by additional spikes
-in the 2\textsuperscript{nd}half of 2021 as well. The all-time daily high for SRs occurred on Tuesday, 2020-08-04, 
-with a spike of 24,415 SRs, a full 9.8 \textsigma's from the mean.
-
-This chart shows the 2-year period which is the basis for this analysis, by calendar month. 
-Note that almost all the months before the COVID pandemic outbreak in Feb/Mar
-of 2020 are ``below average'', while most of the months after that period are ``above average''. That level of higher usage
-of the 311 system has continued into the early months of 2024.
-
-\begin{figure}[htbp]
-  \centering
+the number of 311 SRs grew 50\%. The summer  and early fall months make up all of the Top 5 SR counts by day. 
+The all-time daily high for SRs, 24,415 SRs (a full 9.8 \textsigma's from the 10-year mean), occurred on Tuesday, 2020-08-04.
+The 10-year low count occurs on Christmas Day, 2016, a Sunday. Note also that the difference between 
+the peak day (24,415) and the minimum day (2965) varies by 8X, creating a 
+challenging system performance issue. This timeframe (2014-2023) also includes the COVID pandemic,  
+which reached a peak in the July-September 2020 timeframe, 
+accompanied by additional spikes in the 2\textsuperscript{nd}half of 2021 as well.
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=\textwidth]{10-year-trend-monthly.pdf}
   \caption{Monthly SR counts over 10-years}
-  \includegraphics[width=\textwidth]{10-year-monthly.pdf}
-  \label{fig:monthly-counts}
-\end{figure}
-
-It is not surprising that the lowest SR counts-per-day occur during the Christmas holiday period. The summer months and early
-fall months make up all of the Top 5 SR counts by day. (As a note, the all-time daily high for SRs occurred on Tuesday, 2020-08-04 while the 10-year
-low count occurs on Christmas Day in 2016, a Sunday. Note also that the difference between the peak day (24,415) and the minimum day (2965) varies
-by a factor of 8X; all of which creates a challenging  system capacity and performance issue.
+  \label{fig:10-yr-monthly}
+\end{figure}
 
@@
     \centering
-    \small
+    \normalsize
     \caption{Top 5 and Bottom 5 Days for SRs}
@@
     \end{tabular}
-    \label{tab:combined_counts}
+    \label{tab:top-10-counts}
 \end{table}
 
-%\FloatBarrier % Ensure all floats before this point are processed before moving on
+This chart shows the 2-year period which is the basis for this analysis, by calendar month. 
+Note that almost all the months before the COVID pandemic outbreak in Feb/Mar
+of 2020 are ``below average'', while most of the months after that period are ``above average''. That level of higher usage
+of the 311 system has continued into the early months of 2024. It is not surprising that the lowest SR counts-per-day occur during the Christmas holiday period.
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=\textwidth]{2-year-trend-monthly.pdf}
+  \caption{Monthly SR counts over years 2022-2023}
+  \label{fig:2-yr-monthly}
+\end{figure}
 
@@
 	\centering
+	\includegraphics[width=\textwidth]{SRs_by_Agency.pdf}
   	\caption{SR counts by Agency and Cumulative Percentage}
-	\includegraphics[width=\textwidth]{SRs_by_Agency.pdf}  
+	\label{fig:SRcountbyAgency}
 \end{figure}
@@
   \centering
+ 	\includegraphics[width=\textwidth]{SR_by_Complaint_Type.pdf} 
 	\caption{Top 20 SR complaint\_type and Cumulative Percentage} 
-	  \includegraphics[width=\textwidth]{SR_by_Complaint_Type.pdf}  
 	\label{fig:SR_complaints}
@@
     \caption{Noise-related complaints\_type(s) by count with Agency}
-    \footnotesize
+    \normalsize
     \begin{tabular}{@{}lS[table-format=7.0,round-mode=places,round-precision=0]S[table-format=2.2,round-mode=places,round-precision=2]l@{}} % 'l' for left-aligned, 'S' for siunitx number alignment
@@
   \centering
+ 	\includegraphics[width = 0.65\textwidth] {computed_fields.pdf}	
 	\caption{Screenshot of data portal showing the ``@computed'' columns}  
-	  \includegraphics[width = 0.65\textwidth] {computed_columns_screenshot.pdf}	  
 	\label{fig:computed-columns}
@@
 \resizebox{0.8\linewidth}{!}{%
-\scriptsize
+\normalsize
 \begin{tabular}{lrrrl}
@@
     \centering
-    \footnotesize
+    \normalsize
     \caption{Comparison of Top Ten Zip Codes Lists}
@@
 	\begin{table}[tbp]
-	\footnotesize
+	\normalsize
 	\centering
@@
 		\caption{Top Ten Invalid 'community\_board' Values}
-		\footnotesize
+		\normalsize
 		\begin{tabular}{rlr}
@@
     \caption{Largest and Smallest errors (days)}
-    \small
+    \normalsize
     \begin{tabular}{>{\ttfamily}l l l r l}
@@
     \caption{Sample of SRs with extremely large negative durations}
-    \small
+    \normalsize
     \begin{tabular}{l l r l}
@@
 		 \includegraphics[width = \textwidth]{zero_duration_SR.pdf}
-		 \label{fig:sero-duration}
+		 \label{fig:zero-duration}
 	\end{figure}
@@
 		\centering
+		\includegraphics[width = \textwidth]{Created_Hourly_SR_count.pdf}
 		\caption{SRs by Hourly created\_date(s)}
-		 \includegraphics[width = \textwidth]{Created_Hourly_SR_count.pdf}
 		\label{fig:hourly-created}
@@
 		\centering
+		\includegraphics[width = \textwidth]{Closed_Hourly_SR_count.pdf}
 		\caption{SRs by Hourly closed\_date(s)}
-		 \includegraphics[width = \textwidth]{Closed_Hourly_SR_count.pdf}
 		\label{fig:hourly-closed}
@@
 		\centering
+ 		\includegraphics[width = \textwidth]{2-year-trend-SR_created_by_top_of_hour.pdf}
 		\caption{SRs Created at Top of the Hour}
-		 \includegraphics[width = \textwidth]{SR_created_by_top_of_hour.pdf}
 		\label{fig:tophourcreated}
@@
 		\centering
+		\includegraphics[width=\textwidth]{2-year-trend-SR_created_by_minute_of_busiest_day.pdf}
 		\caption{SRs created minute-by-minute on busiest day}
-		 \includegraphics[width=\textwidth]{SR_created_by_minute_of_busiest_day.pdf}
 		\label{fig:busiestcreated}
@@
 		\centering
+		\includegraphics[width = \textwidth]{2-year-trend-SR_closed_by_top_of_hour.pdf}
 		\caption{SRs Closed at Top of the Hour}
-		 \includegraphics[width = \textwidth]{SR_closed_by_top_of_hour.pdf}
 		\label{fig:tophourclosed}
@@
 		\centering
+		\includegraphics[width=\textwidth]{2-year-trend-SR_closed_by_minute_of_busiest_day.pdf}
 		\caption{SRs closed minute-by-minute on busiest day}
-		 \includegraphics[width=\textwidth]{SR_closed_by_minute_of_busiest_day.pdf}
 		\label{fig:busiestclosed}
@@
 
-	Looking at the distribution by Agency for the ``closed-exactly-at-midnight'' shows that \textgreater{}90\% of these suspect SRs come from just two Agencies - Dept. of 
+	Looking at the distribution by Agency for the ``closed-exactly-at-midnight'' shows that \textgreater90\% of these suspect SRs come from just two Agencies - Dept. of 
 	Buildings (DOB) and Dept. of Sanitation (DSNY). If you include HPD and DOT, that accounts for 99\% of the suspect SR closures. This Agency
@@
 		\centering
+		\includegraphics[width = \textwidth]{closed_at_midnight_chart.pdf}
 		\caption{SRs Closed Exactly at Midnight by Agency}
-		 \includegraphics[width = \textwidth]{closed_at_midnight_chart.pdf}
 		\label{fig:midnight-closed}
@@
 	However, the ``SR closed at noon'' visualization shows that  a single Agency, DSNY,
-	is responsible for \textgreater{}99\% of the SRs ``closed-exactly-at-noon''. Accordingly, any effort to resolve this issue
+	is responsible for \textgreater99\% of the SRs ``closed-exactly-at-noon''. Accordingly, any effort to resolve this issue
 	necessitates working with DSNY to fully understand how this is occurring.
@@
 		\centering
+		\includegraphics[width = \textwidth]{closed_at_noon_chart.pdf}
 		\caption{SRs Closed Exactly at Noon by Agency}
-		 \includegraphics[width = \textwidth]{closed_at_noon_chart.pdf}
 		\label{fig:noon-closed}
@@
 	The 99,779 SRs created-exactly-at-noon also appears to originate with the Dept of Sanitation (DSNY) which
-	is responsible for \textgreater{}99\% of the suspect SRs.
-
-	\begin{figure}[tbp]
-		 \centering
+	is responsible for \textgreater99\% of the suspect SRs.
+
+	\begin{figure}[tbp]
+		\centering
+		\includegraphics[width = \textwidth]{created_at_noon_chart.pdf}
 		\caption{SRs Created Exactly at Noon by Agency}
-		 \includegraphics[width = \textwidth]{created_at_noon_chart.pdf}
 		\label{fig:noon-created}
@@
 	    \caption{Extremely long Resolution Updates}
-	    \footnotesize
+	    \normalsize
 	    \begin{tabular}{cccc}
@@
 	updates to SRs, with some updates occurring almost two years after the SR is closed.   There are a total of 7460 SRs that are 
-	updated \textgreater{}30 days and \textless{}730 days after the closed\_date. The median of these late (\textless{}30 days)
+	updated \textgreater30 days and \textless{}730 days after the closed\_date. The median of these late (\textless{}30 days)
 	post-closed resolution\_action\_update\_date(s) is 84 days. However, the mean resolution\_action\_update\_date 
@@
      \caption{Long post-closed Resolution Updates}
-    \footnotesize
+    \normalsize
 	    \begin{tabular}{cccc}
@@
 	Here is a look at the spread of the post-closed resolution\_action\_update\_date values those SRs 
-	that are \textgreater{}30 days and \textless{}730 days. There are some significant outliers
-
-	\begin{figure}[tbp]
-		 \centering
-		 \caption{Post-Closed resolution\_action\_update\_date >30 days }
+	that are \textgreater30 days and \textless{}730 days. There are some significant outliers
+
+	\begin{figure}[tbp]
+		\centering
 		\includegraphics[width = \textwidth]{post_closed_violin.pdf}
+		\caption{Post-Closed resolution\_action\_update\_date \textgreater30 days}
 		\label{fig:resolution-violin}
@@
 		\centering
-		 \caption{Post-Closed resolution\_action\_update\_date >30 days by Agency }
 		\includegraphics[width = \textwidth]{postClosedBarChart.pdf}
+		\caption{Post-Closed resolution\_action\_update\_date \textgreater30 days by Agency}
 		\label{fig:resolution-by-agency}
@@
 		\centering
-		  \caption{Non-matching borough and borough\_boundary by Agency }
 		\includegraphics[width = \textwidth]{non_matching_borough_boundaries_chart.pdf}
+		\caption{Non-matching borough and borough\_boundary by Agency}
 		\label{fig:borough-boundaries}
@@
     \caption{Non-matches between 'street\_name' and 'landmark' fields}
-    \footnotesize
+    \normalsize
     \begin{tabular}{>{\raggedright\arraybackslash}p{5cm} >{\raggedright\arraybackslash}p{5cm}}
@@
 		\centering
-		  \caption{Non-matching street\_name \& landmark by Agency }
 		\includegraphics[width = \textwidth]{non-matchingstreet_nameandlandmark.pdf}
+		\caption{Non-matching street\_name \& landmark by Agency}
 		\label{fig:landmarkchart}
@@
      \caption{Matching/Non-Matching cross\_street\_1 \& intersection\_street\_1}
-    \footnotesize
-    \begin{tabular}{>{\small\ttfamily}l >{\small\ttfamily}l >{\small\ttfamily}l}
+    \normalsize
+    \begin{tabular}{>{\normalsize\ttfamily}l >{\normalsize\ttfamily}l >{\normalsize\ttfamily}l}
         \toprule
@@
 		\centering
+		\includegraphics[width = \textwidth]{non-matchingcross_street_1andintersection_street_1.pdf}
 		\caption{Non-matching X\_Street\_1 \& Intersection\_Street\_1}
-		 \includegraphics[width = \textwidth]{non-matchingcross_street_1andintersection_street_1.pdf}
 		\label{fig:xstreet1}
@@
 		\centering
+		\includegraphics[width = \textwidth]{non-matchingcross_street_2andintersection_street_2.pdf}
 		\caption{Non-matching X\_Street\_2 \& Intersection\_Street\_2}
-		 \includegraphics[width = \textwidth]{non-matchingcross_street_2andintersection_street_2.pdf}
 		\label{fig:xstreet2}
@@
      \caption{Near-matches for cross\_street\_1, intersection\_street\_1}
-    \footnotesize
+    \normalsize
     \begin{tabular}{l l l r}
@@
     \caption{Summary of cross\_street\_1 and intersection\_street\_1}
-    \footnotesize
+    \normalsize
     \begin{tabular}{l r r}
@@
     \caption{Summary of cross\_street\_2 and intersection\_street\_2}
-    \footnotesize
+    \normalsize
     \begin{tabular}{l r r}
@@
      \caption{Sample of near-matching cross\_street\_2 and intersection\_street\_2 (both non-blank)}
-    \footnotesize
+    \normalsize
     \begin{tabular}{l l l r}

### 2024-06-14 manuscript/manuscript.tex
Clean up of figures and graphics
Standardizing figures and graphics, and file names in Images directory.  Ongoing editing and clean up of \captions.

@@
   \includegraphics[width=\textwidth]{10-year-trend-monthly.pdf}
-  \caption{Monthly SR counts over 10-years}
+  \caption{10-year (2014-2023) Monthly SR Counts}
   \label{fig:10-yr-monthly}
@@
   \includegraphics[width=\textwidth]{2-year-trend-monthly.pdf}
-  \caption{Monthly SR counts over years 2022-2023}
+  \caption{Monthly SR Counts for 2022-2023}
   \label{fig:2-yr-monthly}
@@
 	\includegraphics[width=\textwidth]{SRs_by_Agency.pdf}
-  	\caption{SR counts by Agency and Cumulative Percentage}
+  	\caption{SR counts by Agency with Cumulative Percentage}
 	\label{fig:SRcountbyAgency}
@@
  	\includegraphics[width=\textwidth]{SR_by_Complaint_Type.pdf} 
-	\caption{Top 20 SR complaint\_type and Cumulative Percentage} 
+	\caption{Top 20 complaint\_type(s) and Cumulative Percentage} 
 	\label{fig:SR_complaints}
@@
 complaints total 1.435 million in the 2022-2023 dataset, a full 22\% of all complaint types, which, when consolidated, is the most frequently occurring complaint\_type.
-Ilegal Parking is second, followed by Heat/Hot Water. Noise complaints are handled by NYPD, as is Illegal Parking and Blocked Driveway complaints, which contributes
+Illegal Parking is second, followed by Heat/Hot Water. Noise complaints are handled by NYPD, as is Illegal Parking and Blocked Driveway complaints, which contributes
 significantly to the top ranking of the NYPD as the lead responsible Agency, receiving a full 43\% of all SRs. 
@@
   \centering
- 	\includegraphics[width = 0.65\textwidth] {computed_fields.pdf}	
-	\caption{Screenshot of data portal showing the ``@computed'' columns}  
+ 	\includegraphics[width = \textwidth] {computed_fields.pdf}	
+	\caption{Screenshot of Data Portal with ``computed'' Columns}  
 	\label{fig:computed-columns}
@@
   \centering
-  	\caption{Number and Percentage of Empty Entries}
+  	\caption{Number and Percentage of Empty/Blank Entries}
 	\includegraphics[width=\textwidth]{BlankFields.pdf}
@@
 		\centering
-	  	\caption{Invalid incident-zip by Agency}
+		\caption{Invalid incident\_zip(s) by Agency}
 		\includegraphics[width=\textwidth]{invalid_incident_zip.pdf} 
@@
  	 \centering
- 	 \caption{SRs with a Negative Duration by Agency}
+ 	 \caption{Negative Duration SRs by Agency}
 	  \includegraphics[width = \textwidth]{negative_duration_SR_barchart.pdf}
@@
  	 \centering
- 	 \caption{Negative duration distribution}
+ 	 \caption{Distribution of Negative Durations}
 	  \includegraphics[width = \textwidth]{negative_duration_SR_violin.pdf}
@@
  		\includegraphics[width = \textwidth]{2-year-trend-SR_created_by_top_of_hour.pdf}
-		\caption{SRs Created at Top of the Hour}
+		\caption{SRs Created at Top-of-the-Hour}
 		\label{fig:tophourcreated}
@@
 		\includegraphics[width=\textwidth]{2-year-trend-SR_created_by_minute_of_busiest_day.pdf}
-		\caption{SRs created minute-by-minute on busiest day}
+		\caption{SRs Created Minute-by-Minute on Busiest Day}
 		\label{fig:busiestcreated}
@@
 		\includegraphics[width = \textwidth]{2-year-trend-SR_closed_by_top_of_hour.pdf}
-		\caption{SRs Closed at Top of the Hour}
+		\caption{SRs Closed at Top-of-the-Hour}
 		\label{fig:tophourclosed}
@@
 		\includegraphics[width=\textwidth]{2-year-trend-SR_closed_by_minute_of_busiest_day.pdf}
-		\caption{SRs closed minute-by-minute on busiest day}
+		\caption{SRs Closed Minute-by-Minute on Busiest Day}
 		\label{fig:busiestclosed}
@@
 		\includegraphics[width = \textwidth]{post_closed_violin.pdf}
-		\caption{Post-Closed resolution\_action\_update\_date \textgreater30 days}
+		\caption{Post-Closed resolution\_action\_update\_date(s) \textgreater30 days}
 		\label{fig:resolution-violin}
@@
 		\includegraphics[width = \textwidth]{postClosedBarChart.pdf}
-		\caption{Post-Closed resolution\_action\_update\_date \textgreater30 days by Agency}
+		\caption{Post-Closed resolution\_action\_update\_date(s) \textgreater30 days by Agency}
 		\label{fig:resolution-by-agency}
@@
 		\centering
-		\includegraphics[width = \textwidth]{non-matchingstreet_nameandlandmark.pdf}
+		\includegraphics[width = \textwidth]{non-matchingstreet\_nameandlandmark.pdf}
 		\caption{Non-matching street\_name \& landmark by Agency}

### 2024-06-14 manuscript/manuscript.tex
Added the one JPG file for figure
Deleted older PDF file which wasn't working and added JPG image for computed_fields in NYC Data Portal
Spacing and caption location fixes

@@
 impacting data quality and curation efficiency. Section~\ref{sec:structural} examines
-the technical dimensions of the dataset, the structural issues. In Section~\ref{sec:datatypes} we examine the data fields
-for compliance with the stated data type as found in the Data Dictionary. Section~\ref{sec:blanks} looks
+the technical dimensions of the dataset, the structural issues. 
+In Section~\ref{sec:datatypes} we examine the data fields for compliance 
+with the stated data type as found in the Data Dictionary. Section~\ref{sec:blanks} looks
 at the data fields as regards missing, blank, or N/A entries. Section~\ref{sec:domain} explores
 how data fields comply with a domain of legal or acceptable values. Section~\ref{sec:inconsistencies}
-deals with the important issues surrounding logical inconsistencies and concerning patterns in the data. Section~\ref{sec:precision} 
+deals with the important issues surrounding logical inconsistencies 
+and concerning patterns in the data. Section~\ref{sec:precision} 
 explores the age-old issue of precision versus accuracy. Section~\ref{sec:duplicates}
@@
 actionable suggestions for mitigating or resolving identified issues. Following this, Section~\ref{sec:protocol} outlines a series of general 
-principles for the release of open data, drawing from our findings. The paper concludes with a discussion 
+principles for the release of open data, drawing from our findings. 
+The paper concludes with a discussion and recommendations
 in Section~\ref{sec:discussion}, encapsulating the key insights and implications of our research.
@@
 \begin{itemize}
- 	\item 2003 - NYC 311 system goes live with a phone-based call center only. The ``311'' phone number replaces a myriad of numbers to call for each individual City Agency.
-   	\item 2009 - 311 online \& mobile apps launched. Today 60\% of all Service Requests come from mobile \& online channels.
-   	\item 2019 - Major software upgrade of 311 system with enhanced capabilities, dynamic load-balancing, and cloud-based resilience. Only major software upgrade since 2004.	
-      \item 2020 - In August 2020, driven by the COVID pandemic, monthly 311 Service Requests hit an all-time high of 348,463. Seven of the Top 10 busiest days of all-time occur in 2020.
-      \item 2021 - 311 System is expanded to into the MTA's city subway system, the largest expansion since its inception.
+ 	\item 2003 - NYC 311 system goes live with a phone-based call center only. 
+ 	The ``311'' phone number replaces a myriad of numbers to call for each individual City Agency.
+   	\item 2009 - 311 online \& mobile apps launched. Today 60\% of all 
+   	Service Requests come from mobile \& online channels.
+   	\item 2019 - Major software upgrade of 311 system with enhanced capabilities, 
+   	dynamic load-balancing, and cloud-based resilience. Only major software upgrade since 2004.	
+      \item 2020 - In August 2020, driven by the COVID pandemic, monthly 311 
+      Service Requests hit an all-time high of 348,463. Seven of the Top 10 busiest 
+      days of all-time occur in 2020.
+      \item 2021 - 311 System is expanded to into the MTA's city subway system, 
+      the largest expansion since its inception.
    	\item 2023 - Record high year for 311 with 3.23 million Service Requests.
@@
 vast array of urban living-related inquiries. The system handles over 3 million service 
-requests (SRs) per year, encompassing such complaints as street noise, illegal parking, heat/hot water, abandoned vehicles,
+requests (SRs) per year, encompassing such complaints as street noise, 
+illegal parking, heat/hot water, abandoned vehicles,
 and unsanitary conditions. The data, managed through a tailored application built on the 
@@
 
-Additional Open Data projects can be found at \href{https://opendata.cityofnewyork.us/projects/}{NYC Open Data Project Gallery}
+Additional Open Data projects can be found 
+at \href{https://opendata.cityofnewyork.us/projects/}{NYC Open Data Project Gallery}
 and at \href{https://beta.nyc/beta/products/}{\textbeta etaNYC} products and tools.
@@
 	\item Data timeliness, accuracy, and consistency
-	\item Difficulties correlating data over long time periods, e.g. over a 10-year period, e.g. Agency name changes, new complaint types.
+	\item Difficulties correlating data over long time periods, e.g. over a 10-year 
+	period, e.g. Agency name changes, new complaint types.
 	\item Data anonymization and handling of Personally identifiable information (PII)
 	\item Integration with stand-alone systems at selected NYC Agencies
-	\item Managing API usage, authentication, and load for 3\textsuperscript{rd} party users, of which there are many
+	\item Managing API usage, authentication, and load for 3\textsuperscript{rd} 
+	party users, of which there are many
 	\item Incorporating ongoing technology changes and upgrades 
-	\item Understanding the role of chatbots and AI to provide online customer service, along with other emergent technologies
+	\item Understanding the role of chatbots and AI to provide online 
+	customer service, along with other emergent technologies
 \end{itemize}
@@
 These and other challenges are continually addressed by the various Agency open data managers
-as well as the \href{https://www.nyc.gov/content/oti/pages/}{NYC Office of Technology and Innovation (OTI)} which
-provides the technology support for the open data system.  
+as well as the 
+\href{https://www.nyc.gov/content/oti/pages/}{NYC Office of Technology and Innovation (OTI)} 
+which provides the technology support for the open data system.  
 Many of the underlying datasets are updated daily, usually running 1-2 days behind from creation.
@@
 merging dataset with others, as would be expected. Typically, the data must be
-exported and manipulated in another, external program. The authors found this to be the case with 
-the large 311 dataset employed.
+exported and manipulated in another, external program. The authors found 
+this to be the case with the large 311 dataset employed.
 
@@
 via Service Level Agreements (SLAs). However, enforcement of those SLAs is often challenging. The 
-result is that often data accuracy and logical consistency is occasionally compromised. However, the Open Data
-Team is quite responsive to inquires from users. The authors have provided many such 
-suggestions during the course of this research effort, and the most current data shows improvements
-in many of the areas highlighted. Questions and suggestion  can be provide here: 
+result is that often data accuracy and logical consistency is occasionally compromised. 
+However, the Open Data Team is quite responsive to inquires from users. The authors have provided many such suggestions during the course of this research effort, 
+and the most current data shows improvements in many of the areas highlighted. 
+Questions and suggestion  can be provide here: 
 \href{https://opendata.cityofnewyork.us/engage/}{Contact Us}.
@@
 The NYC 311 organization, part of OTI, operates a highly tailored and sophisticated case
-management. However, one constraint was particularly challenging during the recent complete overhaul
-of the software and hardware in 2019. When the original 311system was developed in 2003, 
-a decision was made to allow select NYC Agencies to maintain their own local computer systems, 
+management. However, one constraint was particularly challenging 
+during the recent complete overhaul of the software and hardware in 2019. When the original 311system was developed in 2003, a decision was made to allow select 
+NYC Agencies to maintain their own local computer systems, 
 such as the mainframe systems at the NYC Department of Sanitation (DSNY) and others. 
 And when the 311 system was completely upgraded in 2019, those
-same integrations were honored in order to reduce the scope of the project. Perhaps in
-retrospect, while a good budgetary and policy decision, it was perhaps not such a good technology decision.
-
-At least nine NYC Agencies have local systems that are integrated directly to the 311 system either via
-an API or custom code. These include:
+same integrations were honored in order to reduce the scope of the project. 
+Perhaps in retrospect, while a good budgetary and policy decision, 
+it was perhaps not such a good technology decision.
+
+At least nine NYC Agencies have local systems that are integrated 
+directly to the 311 system either via an API or custom code. These include:
 
@@
 	\item Department of Social Services (DSS), one of the largest City Agencies
-	\item Department of Sanitation, although recently part of DSNY has moved to the core 311 software.
+	\item Department of Sanitation, although recently part of DSNY 
+	has moved to the core 311 software.
 	\item Department of Parks \& Recreation (forestry complaints only)
@@
 	\item Department of Healthy \& Mental Hygiene (fully integrated) 
-	\item Department of Consumer \& Worker Protection (DCWP - formerly known as the Dept of Consumer Affairs)
+	\item Department of Consumer \& Worker Protection 
+	(DCWP - formerly known as the Dept of Consumer Affairs)
 	\item New York Police Department (NYPD) (integrated with their reporting system)
@@
 
-One unfortunate side-effect from these integrations, and other Agencies that use the 311 APIs, is that some of the data consistency
-checks which are inherent in the 311 system are not applied uniformly, and in some cases not at all, to the integrated data streams.
-Errors often creep in from other Agency systems and are not easily corrected or detected in the 311 application itself,
+One unfortunate side-effect from these integrations, and other Agencies 
+that use the 311 APIs, is that some of the data consistency
+checks which are inherent in the 311 system are not applied uniformly, 
+and in some cases not at all, to the integrated data streams.
+Errors often creep in from other Agency systems and are not easily 
+corrected or detected in the 311 application itself,
 but rather require correction at the source Agency; challenging.
 
-One area that will almost certainly remain an ongoing issue is the ongoing incorporation of new software, hardware,
-and cloud services which are constantly changing while new approaches emerge. The 311 telephonic integration recently was reconfigured due to a new
-telephonic vendor. Desktop software at the 311 Agency are subject to the same upgrades that home and business users experience. And while the 
-311 system was designed to be easily configured and updated with new features, some emerging business issues as well as
-technologies nonetheless require additional software engineering efforts to accommodate, along with the concomitant budgets.
+One area that will almost certainly remain an ongoing issue is the 
+ongoing incorporation of new software, hardware,
+and cloud services which are constantly changing while new approaches emerge. 
+The 311 telephonic integration recently was reconfigured due to a new
+telephonic vendor. Desktop software at the 311 Agency are subject 
+to the same upgrades that home and business users experience. And while the 
+311 system was designed to be easily configured and updated with new features, 
+some emerging business issues as well as technologies nonetheless require additional software engineering efforts to accommodate, along with the concomitant budgets.
 
@@
 	\item Providing advice on shelters during hurricanes, heat emergencies, and winter blizzards
-	\item Handling countless inquires during the COVID pandemic (testing centers, vaccine sites, mobile vaccine clinics, etc.)
+	\item Handling countless inquires during the COVID pandemic 
+	(testing centers, vaccine sites, mobile vaccine clinics, etc.)
 	\item Enforcement of standards between landlords and tenants
-	\item Re-allocation of NYC's yellow taxi routes based upon an insightful analysis conducted by the Taxi \& Limousine Commission (TLC)
-	\item Improved responsiveness for City Agencies with regards to direct client-facing responsibilities such as Parks \& Rec, Dept of Transportation, DSNY, etc.
-	\item Increased 311 SRs related to Homelessness resulted in the launch of the City's Homeless Outreach \& Mobile Engagement Street Action Teams (HOME-STAT)
+	\item Re-allocation of NYC's yellow taxi routes based upon an insightful analysis 
+	conducted by the Taxi \& Limousine Commission (TLC)
+	\item Improved responsiveness for City Agencies with regards to direct 
+	client-facing responsibilities such as Parks \& Rec, Dept of Transportation, DSNY, etc.
+	\item Increased 311 SRs related to Homelessness resulted in the launch 
+	of the City's Homeless Outreach \& Mobile Engagement Street Action Teams (HOME-STAT)
 	which includes street canvassers and increases in homeless housing facilities.
-	\item A significant increase in graffiti complains in 2017 led directly to the `Graffiti-Free NYC'' campaigns and increased funding for removal of graffiti in public spaces.
-	\item Enhanced geospatial data regarding the location of incidents, the responding Agency, and actions taken across a myriad of complaint\_types
-	\item Establishment of Agency SLAs in regard to certain complaint types, such as residential noise, homelessness person assistance, street light outages, 
+	\item A significant increase in graffiti complains in 2017 led directly to 
+	the `Graffiti-Free NYC'' campaigns and increased funding for removal of graffiti in public spaces.
+	\item Enhanced geospatial data regarding the location of incidents, 
+	the responding Agency, and actions taken across a myriad of complaint\_types
+	\item Establishment of Agency SLAs in regard to certain complaint types, 
+	such as residential noise, homelessness person assistance, street light outages, 
 	rat sightings, and other neighborhood quality of life issues.
@@
 We extracted 311 SR data for the period of CY 2022-2023, a dataset consisting of
-6.4 million records. That dataset serves as the primary source for the analysis effort. This is a large dataset which enables 
-observation of some rarely seen events, data inaccuracies, and logical inconsistencies. It provides enough data to examine trends and data outlier analysis. 
-We queried the data directly from the NYC Open Data Portal and then exported the data in CSV format for analysis using custom written R programs.
-
-Additionally, we extracted two longer-term datasets, a 5-year look (2019-2023) and a 10-year look (2014-2023) ; some 14 and 27 million rows respectively. 
-These larger datasets afford an opportunity for long-term trend observations, confirming and supporting the analysis using the 2-year dataset. The 2022-2023 dataset
-was used exclusively for the detailed, field-by-field level analysis which comprises the majority of this effort.
+6.4 million records. That dataset serves as the primary source for the analysis effort. 
+This is a large dataset which enables 
+observation of some rarely seen events, data inaccuracies, and logical inconsistencies. 
+It provides enough data to examine trends and data outlier analysis. 
+We queried the data directly from the NYC Open Data Portal and then exported the 
+data in CSV format for analysis using custom written R programs.
+
+Additionally, we extracted two longer-term datasets, a 5-year look (2019-2023) 
+and a 10-year look (2014-2023) ; some 14 and 27 million rows respectively. 
+These larger datasets afford an opportunity for long-term trend observations, 
+confirming and supporting the analysis using the 2-year dataset. The 2022-2023 dataset
+was used exclusively for the detailed, field-by-field level analysis 
+which comprises the majority of this effort.
 
@@
 observe system performance over a longer time-frame . During that time-frame, 
-the number of 311 SRs grew 50\%. The summer  and early fall months make up all of the Top 5 SR counts by day. 
-The all-time daily high for SRs, 24,415 SRs (a full 9.8 \textsigma's from the 10-year mean), occurred on Tuesday, 2020-08-04.
-The 10-year low count occurs on Christmas Day, 2016, a Sunday. Note also that the difference between 
+the number of 311 SRs grew 50\%. The summer  and early fall months 
+make up all of the Top 5 SR counts by day. 
+The all-time daily high for SRs, 24,415 SRs (a full 9.8 \textsigma's from the 10-year mean), 
+occurred on Tuesday, 2020-08-04.
+The 10-year low count occurs on Christmas Day, 2016, a Sunday. 
+Note also that the difference between 
 the peak day (24,415) and the minimum day (2965) varies by 8X, creating a 
-challenging system performance issue. This timeframe (2014-2023) also includes the COVID pandemic,  
+challenging system performance issue. This timeframe (2014-2023) 
+also includes the COVID pandemic,  
 which reached a peak in the July-September 2020 timeframe, 
@@
 Note that almost all the months before the COVID pandemic outbreak in Feb/Mar
-of 2020 are ``below average'', while most of the months after that period are ``above average''. That level of higher usage
-of the 311 system has continued into the early months of 2024. It is not surprising that the lowest SR counts-per-day occur during the Christmas holiday period.
+of 2020 are ``below average'', while most of the months after that period 
+are ``above average''. That level of higher usage
+of the 311 system has continued into the early months of 2024. 
+It is not surprising that the lowest SR counts-per-day occur during the Christmas holiday period.
 
@@
 
-Two important fields in this analysis effort are the responsible NYC Agency, and the type of complaint.
-Given the size of the City of New York government, it is necessary to identify the responsible Agency as well as the type of complaint in order to 
+Two important fields in this analysis effort are the responsible NYC Agency, 
+and the type of complaint.
+Given the size of the City of New York government, it is necessary to 
+identify the responsible Agency as well as the type of complaint in order to 
 identify the responsible party and assist in troubleshooting discrepancies.
 
-When we discovered a data error, we typically observe one of two trends; either a single (or a few) Agencies are responsible (such as Dept of Transportation), 
+When we discovered a data error, we typically observe one of two trends; 
+either a single (or a few) Agencies are responsible (such as Dept of Transportation), 
 or it is a system-wide issue that follows the general distribution of  SRs across the system. 
-Thus we can often identify if it is an Agency specific problem, or a systemic issue.  (Note: a few key NYC Agencies are not 
+Thus we can often identify if it is an Agency specific problem, or a systemic issue.  
+(Note: a few key NYC Agencies are not 
 represented in the 311 dataset. These include the New York City Housing Authority (NYCHA) and the
@@
 
-Although the 311 SRs involve 16 different agencies, there are the ``big six'' Agencies that comprise
-90\% of the complaints. These six are (in order of \% of complaints):
+Although the 311 SRs involve 16 different agencies, there are the ``big six'' 
+Agencies that comprise 90\% of the complaints. 
+These six are (in order of \% of complaints):
 
@@
 
-The distribution of complaint\_type skews to a few select complaints accompanied by a long tail.
-This graph shows the top 20 complaint\_type(s) comprising 70\% out of a total of 220.
+The distribution of complaint\_type skews to a few select complaints 
+accompanied by a long tail. This graph shows the top 20 complaint\_type(s) 
+comprising 70\% out of a total of 220.
 
@@
 
-The top 20 complaints contain several of the eight different types of ``Noise'' complaints (Residential, Commercial, Street, Helicopter, etc.). These Noise-related
-complaints total 1.435 million in the 2022-2023 dataset, a full 22\% of all complaint types, which, when consolidated, is the most frequently occurring complaint\_type.
-Illegal Parking is second, followed by Heat/Hot Water. Noise complaints are handled by NYPD, as is Illegal Parking and Blocked Driveway complaints, which contributes
-significantly to the top ranking of the NYPD as the lead responsible Agency, receiving a full 43\% of all SRs. 
+The top 20 complaints contain several of the eight different types of ``Noise'' 
+complaints (Residential, Commercial, Street, Helicopter, etc.). These Noise-related
+complaints total 1.435 million in the 2022-2023 dataset, a full 22\% of 
+all complaint types, which, when consolidated, is the most frequently occurring complaint\_type.
+Illegal Parking is second, followed by Heat/Hot Water. Noise complaints 
+are handled by NYPD, as is Illegal Parking and Blocked Driveway complaints, 
+which contributes significantly to the top ranking of the NYPD as the lead 
+responsible Agency, receiving a full 43\% of all SRs. 
 Housing, Preservation, \& Development are 2\textsuperscript{nd}, with 21\%
 
-For  non-NYC residents, HPD is the City agency that manages the 177,569 NYC public housing units as well as monitoring
-all City rental and leased apartment units. HPD collectively serves 528,105 people, a population larger than Atlanta or Miami. 
-Hence the large number of Heat/Hot Water and Unsanitary complaints handled by HPD.
+For  non-NYC residents, HPD is the City agency that manages the 
+177,569 NYC public housing units as well as monitoring
+all City rental and leased apartment units. HPD collectively serves 
+528,105 people, a population larger than Atlanta or Miami. 
+Hence the large number of Heat/Hot Water and Unsanitary complaints 
+handled by HPD.
 
@@
     \centering
+    \normalsize
     \caption{Noise-related complaints\_type(s) by count with Agency}
-    \normalsize
-    \begin{tabular}{@{}lS[table-format=7.0,round-mode=places,round-precision=0]S[table-format=2.2,round-mode=places,round-precision=2]l@{}} % 'l' for left-aligned, 'S' for siunitx number alignment
+    \begin{tabular}{@{}lS[table-format=7.0,round-mode=places,
+    round-precision=0]S[table-format=2.2,round-mode=places,
+    round-precision=2]l@{}} % 'l' for left-aligned, 'S' for siunitx number alignment
         \toprule
@@
 There are also some curious and unusual Service Requests in the 311 data: 
-tanning, tattooing, trans fat, unsanitary pigeon condition, illegal animal kept as pet, harboring bees/wasps, and the authors' favorite
+tanning, tattooing, trans fat, unsanitary pigeon condition, illegal animal 
+kept as pet, harboring bees/wasps, and the authors' favorite
 radioactive material. New Yorkers certainly live interesting lives.
@@
 
-What is data cleansing?  Wikipedia \href{https://en.wikipedia.org/wiki/Data_cleansing}{Data Cleansing} offers a good definition.
-
-``Data cleansing or data cleaning is the process of detecting and correcting (or removing) corrupt or inaccurate records from a record set, 
-table, or database and refers to identifying incomplete, incorrect, inaccurate or irrelevant parts of the data 
+What is data cleansing?  Wikipedia 
+\href{https://en.wikipedia.org/wiki/Data_cleansing}{Data Cleansing} offers a good definition.
+
+``Data cleansing or data cleaning is the process of detecting and 
+correcting (or removing) corrupt or inaccurate records from a record set, 
+table, or database and refers to identifying incomplete, incorrect, 
+inaccurate or irrelevant parts of the data 
 and then replacing, modifying, or deleting the dirty or coarse data.''
 
-Many quality criteria are required in order to process high-quality data. These include:
+Many quality criteria are required in order to process high-quality data. 
+These include:
 
@@
 		\item Mandatory fields: Certain data fields cannot be empty.
-		\item Data-types: Certain fields must be of the correct type, e.g. numeric, character, date, 5 numeric digits, etc. Typically these are identified in a Data Dictionary.
-		\item Domain adherence:  Many data fields must adhere to a specific domain of values, e.g. statuses, state names, zip codes, gender. This includes data
-		fields that are restricted to certain ranges, such as longitudes and latitudes.
+		\item Data-types: Certain fields must be of the correct type, e.g. numeric, 
+		character, date, 5 numeric digits, etc. Typically these are identified 
+		in a Data Dictionary.
+		\item Domain adherence:  Many data fields must adhere to a specific 
+		domain of values, e.g. statuses, state names, zip codes, gender. 
+		This includes data fields that are restricted to certain ranges, 
+		such as longitudes and latitudes.
 	\end{itemize}   
-	\item Structural errors to include naming conventions, extra fields not in the Data Dictionary, or inconsistent data entry.
-	For example intermixing blanks, spaces, NA, N/A, and \textless{}NA\textgreater{} all to indicate the absence of data.
+	\item Structural errors to include naming conventions, 
+	extra fields not in the Data Dictionary, or inconsistent data entry.
+	For example intermixing blanks, spaces, NA, N/A, and \textless{}NA\textgreater{} 
+	all to indicate the absence of data.
 	\item Redundant, unnecessary, or irrelevant fields
-	\item Logical inconsistencies such as related fields that violate the nature of that relationship, e.g. a ``due date'' that is before the ``created date''
-	\item Data standardization. Often free-form entry fields can suffer from inconsistent data structures. Efforts to standardize these fields can greatly assist analysis.
+	\item Logical inconsistencies such as related fields that violate the 
+	nature of that relationship, e.g. a ``due date'' that is before the ``created date''
+	\item Data standardization. Often free-form entry fields can suffer 
+	from inconsistent data structures. Efforts to standardize these fields 
+	can greatly assist analysis.
 	\item Accuracy and precision, which of course are not the same thing 
@@
 
-This analysis effort is to identify the presence of such errors; not to correct them. That effort would be undertaken only after an investigation as to the why
-and how such errors came about, and a discussion as to whether or not it is even an ``error''. Typically this requires discussions with subject matter experts (SMEs)
- and liaison within the various NYC Agency open data coordinators who can help identify the underlying cause for the error.
- For example, an invalid date field that is provided to the NYC Open Data data lake via an Agency integration could
-be the result of selecting the wrong field from the source Agency's system, e.g. due\_date instead of closed\_date. In other cases, without a subject matter
-expert, it is not possible to actually determine if the data is inaccurate or not. There are many empty values in the taxi\_company\_borough field. Are those
-entries missing or does the entry in that field only apply to a very small number of complaint\_type(s) or circumstances. Only a SME
-from the Taxi \& Limo Commission (TLC) can provide that answer. The R code used here to analyze the 311 Service Request dataset identifies
-both obvious inaccuracies as well as potential errors requiring further investigation. It is not a conclusive list of errors, but rather mostly identifies
+This analysis effort is to identify the presence of such errors; not to correct them. 
+That effort would be undertaken only after an investigation as to the why
+and how such errors came about, and a discussion as to whether or not it 
+is even an ``error''. Typically this requires discussions with subject matter 
+experts (SMEs) and liaison within the various NYC Agency open data 
+coordinators who can help identify the underlying cause for the error. 
+For example, an invalid date field that is provided to the NYC Open Data 
+data lake via an Agency integration could
+be the result of selecting the wrong field from the source Agency's 
+system, e.g. due\_date instead of closed\_date. In other cases, without a subject matter
+expert, it is not possible to actually determine if the data is inaccurate or not. 
+There are many empty values in the taxi\_company\_borough field. Are those
+entries missing or does the entry in that field only apply to a 
+very small number of complaint\_type(s) or circumstances. Only a SME
+from the Taxi \& Limo Commission (TLC) can provide that answer. 
+The R code used here to analyze the 311 Service Request dataset identifies
+both obvious inaccuracies as well as potential errors requiring further 
+investigation. It is not a conclusive list of errors, but rather mostly identifies
 ``a place to look'' for further evaluation and potential action.
@@
 	\item Identifying invalid values
-	\item Exposing logical inconsistencies \& inconsistent or unusual patterns in the data
+	\item Exposing logical inconsistencies \& inconsistent or 
+	unusual patterns in the data
 	\item Accuracy and precision issues
@@
 
-Structural issues in the context of data cleansing refers to issues related to the how data is organized, formatted, or structured within a dataset. Structural issues 
-can make it difficult to analyze the data effectively. Some common structural concerns in this 311 SR 2022-2023 dataset include:
+Structural issues in the context of data cleansing refers to issues 
+related to the how data is organized, formatted, or structured within a dataset. Structural issues 
+can make it difficult to analyze the data effectively. Some common 
+structural concerns in this 311 SR 2022-2023 dataset include:
 
 \begin{itemize}
-	\item Data structure (data fields, columns, formats, data types, etc.) not corresponding to the Data Dictionary
+	\item Data structure (data fields, columns, formats, data types, etc.) 
+	not corresponding to the Data Dictionary
 	\item Verification of correct data types (numeric, character, geospactial, dates, etc.)
-	\item Embedded or combined values: data elements that contain multiple pieces of information 
+	\item Embedded or combined values: data elements that 
+	contain multiple pieces of information 
 \end{itemize}
@@
 	\item Two Police Precinct fields; not duplicates
-	\item In addition to street/city, there are three other location fields: lat/long, NY State plane, and Block \#
-	\item incident\_address and street\_name, e.g. incident\_address, 25 Grymes Hill Road and street\_name (Grymes Hill Road)
-	\item One free-form text field, resolution\_description, which supports 934 characters of input, including commas and special characters
+	\item In addition to street/city, there are three other location fields: 
+	lat/long, NY State plane, and Block \#
+	\item incident\_address and street\_name, e.g. incident\_address, 
+	25 Grymes Hill Road and street\_name (Grymes Hill Road)
+	\item One free-form text field, resolution\_description, which 
+	supports 934 characters of input, including commas and special characters
 \end{itemize}
 
-\textbf{Issue: Fields not in the Data Dictionary} The 311 SR \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}
- identifies 41 data columns (fields) along with related information for each column. 
- However, when you download the data or explore it via the online portal, you immediately note that there are 47 columns (fields).
- he additional six fields are not addressed in the Data Dictionary, but do and show up on the Column Manager widget
- as ``@computed\_region\_xxxx\_xxxx''. To some extent one can infer what the these @computed\_fields are, but it is not possible to know
+\textbf{Issue: Fields not in the Data Dictionary} The 311 SR 
+\href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}
+ identifies 41 data columns (fields) along with related information 
+ for each column. However, when you download the data or 
+ explore it via the online portal, you immediately note that there 
+ are 47 columns (fields). The additional six fields are not addressed in the 
+ Data Dictionary, but do and show up on the Column Manager widget
+ as ``@computed\_region\_xxxx\_xxxx''. To some extent one can infer 
+ what the these @computed\_fields are, but it is not possible to know 
  for certain. These six fields are:
@@
 
-Here is a screenshot of the NYC Open Data Portal ``Column Manager'' page showing those columns:
-
-\begin{figure}[tbp]
-  \centering
- 	\includegraphics[width = \textwidth] {computed_fields.pdf}	
-	\caption{Screenshot of Data Portal with ``computed'' Columns}  
+Here is a screenshot of the NYC Open Data Portal ``Column Manager'' 
+page showing those columns:
+
+\begin{figure}[tbp]
+  \centering
+ 	\includegraphics[width = .65\textwidth] {computed_fields_jpg.jpg}	
+	\caption{Screenshot of Data Portal with ``@computed'' Columns}  
 	\label{fig:computed-columns}
@@
 \begin{itemize}
-	\item How are these fields computed? Using what source data? Using what computational method?
+	\item How are these fields computed? Using what source data? 
+	Using what computational method?
 	\item How can these fields be verified or cross-checked?
-	\item Why are there two fields that appear redundant ``police\_precinct' and ``police\_precincts''? Are these redundant?
-	\item What are ``borough boundaries''? They appear to be borough names. Why are they referred to as ``boundaries''?
-	\item What are ``community districts''? Is this the same as the ``community\_board(s)'' that are part of the NYC government?
+	\item Why are there two fields that appear redundant ``police\_precinct' 
+	and ``police\_precincts''? Are these redundant?
+	\item What are ``borough boundaries''? They appear to be borough names. 
+	Why are they referred to as ``boundaries''?
+	\item What are ``community districts''? Is this the same as the
+	 ``community\_board(s)'' that are part of the NYC government?
 	\item Why are not these fields included in the official 311 SR Data Dictionary?
@@
 
-The last bullet point is the key here. Are these fields reliable and accurate, and can they be used for subsequent analysis? Unfortunately,
-as of this date, those questions have not been fully addressed by the  NYC Open Data Team. This paper will further address that issue by
+The last bullet point is the key here. Are these fields reliable and accurate, 
+and can they be used for subsequent analysis? Unfortunately,
+as of this date, those questions have not been fully addressed by the 
+NYC Open Data Team. This paper will further address that issue by
 exploring the accuracy, validity, and candidacy for subsequent analytical usage.
@@
 \section{Validating data types}\label{sec:datatypes}
-Fortunately, both the Data Dictionary and the portal column manager do a good job of indicating the data type.
-As indicated by the red underline in Figure \ref{fig:computed-columns}, there is a small icon next to the field name 
-which indicates the data type of the field. The stylized capital ``T'' indicates ``text'', the ``\#'' sign indicates 
-``numeric'', a calendar icon indicates a ``date'' field, and the pin on a map symbol indicates a ``geospatial'' field.
+Fortunately, both the Data Dictionary and the portal column manager 
+do a good job of indicating the data type.
+As indicated by the red underline in Figure \ref{fig:computed-columns}, 
+there is a small icon next to the field name which indicates the 
+data type of the field. The stylized capital ``T'' indicates ``text'', the ``\#'' sign indicates 
+``numeric'', a calendar icon indicates a ``date'' field, and the pin on a 
+map symbol indicates a ``geospatial'' field.
 
@@
 \begin{itemize}
-	\item created\_date, closed\_date, due\_date, and resolution\_action\_updated\_date: All are all in proper date format.
-	\item zip\_codes and incident\_zip: All are 5 numeric digits except for 2 non-numeric entries for incident\_zip (``na'' and ``N/A'').
-	\item x\_coordinate\_state\_plane \&  y\_coordinate\_state\_plane (a NY State geo-location system): All are all numeric.
+	\item created\_date, closed\_date, due\_date, and resolution\_action\_updated\_date: 
+	All are all in proper date format.
+	\item zip\_codes and incident\_zip: All are 5 numeric digits except 
+	for two non-numeric entries for incident\_zip (``na'' and ``N/A'').
+	\item x\_coordinate\_state\_plane \&  y\_coordinate\_state\_plane 
+	(a NY State geo-location system): All are all numeric.
 	\item latitude \& longitude: Both are numeric.
-	\item community\_districts, borough\_boundaries, city\_council\_district, police\_precinct, and police\_precincts: All are numeric.
-	\item Free-form text fields that contain a ``,'' (e.g. incident\_address \& resolution\_description) are properly enclosed in quotation marks.
+	\item community\_districts, borough\_boundaries, 
+	city\_council\_district, police\_precinct, and police\_precincts: All are numeric.
+	\item Free-form text fields that contain a ``,'' (e.g. incident\_address 
+	\& resolution\_description) are properly enclosed in quotation marks.
 \end{itemize}
 
-All data fields were subjected to a test of ``correctness'' according to their ``type'' as specified
-in the Data Dictionary. Of the nearly 300 million data fields to check (6.5 million rows, 47 fields per row), there were only two values that failed to test the
-test for appropriate data type; those being two text entries in the incident\_zip field (``na'', and ``N/A''). All date fields were dates. All numbers were numeric.
-All text fields were character. Data typing appears to not be an issue with this dataset.
+All data fields were subjected to a test of ``correctness'' 
+according to their ``type'' as specified
+in the Data Dictionary. Of the nearly 300 million data fields to check 
+(6.5 million rows, 47 fields per row), there were only two values that failed 
+the test for appropriate data type; those being two text entries in the 
+incident\_zip field (``na'', and ``N/A''). All date fields were dates. 
+All numbers were numeric. All text fields were character. 
+Data typing appears to not be an issue with this dataset.
 
@@
 
-Understanding the absence of data by field is an important factor when undertaking analysis. For example, if you wanted to see if the SRs were
-closed before or after their due\_date, you would be challenged as 99.57\% of the due\_date field is blank. (Although with such a large dataset
-that still leaves 24,411 values, all of which are ``N/A'. Which calls into question...are these missing values, or is the concept of a
+Understanding the absence of data by field is an important factor 
+when undertaking analysis. For example, if you wanted to see if the SRs were
+closed before or after their due\_date, you would be challenged 
+as 99.57\% of the due\_date field is blank. (Although with such a large dataset
+that still leaves 24,411 values, all of which are ``N/A'. Which calls 
+into question...are these missing values, or is the concept of a
 due\_date simply not applicable for the vast majority of SRs?
 
-When counting the various fields for blank or N/A values, the fields appear to divide into three groups: \textbf{Mostly Empty}, \textbf{Partially Empty}, 
-and \textbf{Few/None Empty}. he Mostly Empty category ranges from 93-99.9\% blank. It includes such fields as taxi\_company due\_date, pickup\_location, and landmark. The Partially 
-Empty includes such fields as location\_type, borough, and cross\_street. And the Few/None Empty includes created\_date, complaint\_type, agency, and status.
-In some cases it may make sense to inquiry as to why some fields are frequently blank. Is the data difficult to capture? Does it pertain to only a small set
-of complaint\_type(s) or Agencies? And if the data is truly that sparse, does it even make sense to collect it? Is it legacy data? For example, one would expect that every complaint\_type
-that has a geographical component, such as an address, would also have a corresponding police\_precinct, community\_board, and community\_district since these
-are all-inclusive throughout the five Boroughs of NYC. Addtionally, that it would have a lat/long. And generally that is the case, but these fields do have small number, but nonetheless significant, 
-of empty entries.
-
-During the analysis, it was determined that the following fields are mandatory, and any row missing these fields should be removed from the dataset (which fortunately
-very rarely occurred): created\_date, agency, complaint\_type, and unique\_key. The absence of any one of these four fields prohibits further analysis  as currently conducted. Again,
-it is very rare occurrence, happening only 1-2 times in 6.4 million records. Here is a breakdown of bank and N/A entries by field.
+When counting the various fields for blank or N/A values, the fields appear 
+to divide into three groups: \textbf{Mostly Empty}, \textbf{Partially Empty}, 
+and \textbf{Few/None Empty}. he Mostly Empty category ranges from 93-99.9\% blank. 
+It includes such fields as taxi\_company due\_date, pickup\_location, and 
+landmark. The Partially Empty includes such fields as location\_type, borough, 
+and cross\_street. And the Few/None Empty includes created\_date, 
+complaint\_type, agency, and status. In some cases it may make 
+sense to inquiry as to why some fields are frequently blank. Is the data 
+difficult to capture? Does it pertain to only a small set of 
+complaint\_type(s) or Agencies? And if the data is truly that sparse, 
+does it make sense to collect it? Is it legacy data? For example, 
+one would expect that every complaint\_type that has a geographical 
+component, such as an address, would also have a corresponding 
+police\_precinct, community\_board, and community\_district since these
+are all-inclusive throughout the five Boroughs of NYC. Addtionally, 
+that it would have a lat/long. And generally that is the case, but these fields 
+do have small number, but nonetheless significant, of empty entries.
+
+During the analysis, it was determined that the following fields are mandatory, 
+and any row missing these fields should be removed from the dataset (which fortunately
+very rarely occurred): created\_date, agency, complaint\_type, and unique\_key. 
+The absence of any one of these four fields prohibits further analysis  as currently 
+conducted. Again, it is very rare occurrence, happening only 1-2 times 
+in 6.4 million records. Here is a breakdown of bank and N/A entries by field.
 
@@
 \centering
+\normalsize
 \caption{Blank and N/A entries by Field}
 \resizebox{0.8\linewidth}{!}{%
-\normalsize
 \begin{tabular}{lrrrl}
 	\toprule
-	\textbf{Field} & \textbf{Total Empty} & \textbf{Percent Empty} & \textbf{N/A Count} \\
+	\textbf{Field} & \textbf{Total Empty} & \textbf{Percent Empty} 
+	& \textbf{N/A Count} \\
 	\midrule
@@
 
-Here is a graphic depiction of total empty (blank \& N/As) for each fields. You can see the natural grouping into the Most, Some, and Few categories
+Here is a graphic depiction of total empty (blank \& N/As) for each fields. 
+You can see the natural grouping into the Most, Some, and Few categories 
 in the stair-type visualization.
@@
 	\centering
+  	\includegraphics[width=\textwidth]{BlankFields.pdf}
 	\caption{Number and Percentage of Empty/Blank Entries}
-	\includegraphics[width=\textwidth]{BlankFields.pdf}
 	\label{fig:blank_fields}
@@
  \section{Validating Data for Acceptable Values}\label{sec:domain}
- While the dataset may have fields populated with data that appears correct, it is necessary to inspect those fields to determine if the data is indeed valid. 
- For example a date field that contains the value ``February 30, 2024'' is clearly not valid. Nor is a latitude of +95 deg north.
-
- In this study, testing for invalid values revealed several serious issues associated with select fields. Accordingly, an analyst would need to be
-extra diligent if using these fields for analysis. In all likelihood, rows with invalid field values would  need to be removed from the
+ While the dataset may have fields populated with data that appears correct, 
+ it is necessary to inspect those fields to determine if the data is indeed valid. 
+ For example a date field that contains the value ``February 30, 2024'' is 
+ clearly not valid. Nor is a latitude of +95 deg north.
+
+ In this study, testing for invalid values revealed several serious issues 
+ associated with select fields. Accordingly, an analyst would need to be 
+ extra diligent if using these fields for analysis. In all likelihood, 
+ rows with invalid field values would  need to be removed from the 
  dataset before conduction further analysis. Here are some results.
@@
 \begin{itemize}
-	\item Latitude and Longitude fields were tested to ensure all fell within the geographic boundaries of the City of New York. All did.
+	\item Latitude and Longitude fields were tested to ensure all fell 
+	within the geographic boundaries of the City of New York. All did.
 	\item The unique\_key field was in fact unique.
-	\item Many fields have a domain of acceptable values. Often these values are determined from common usage or by examining
-	larger historical datasets. Unfortunately, these domains of acceptable values are not specified in the Data Dictionary. These fields all tested as
-	compliant within their domain of acceptable values as determined by examining a 10-year dataset:
+	\item Many fields have a domain of acceptable values. Often these 
+	values are determined from common usage or by examining
+	larger historical datasets. Unfortunately, these domains of acceptable 
+	values are not specified in the Data Dictionary. These fields all tested as 
+	compliant within their domain of acceptable values as determined by 
+	examining a 10-year dataset:
 		\begin{itemize}
@@
 	\label{sec:zipcodesissues}
-	 Unfortunately some fields proved to be  problematic when comparing the values to a domain of legal values.
-	 For example, all zip codes (two fields: zip\_codes and incident\_zip) should be valid as defined by the USPS database
-	 which contains 37,946 valid zip codes.
-
-	The computed field zip\_codes proved especially problematic with 58\% (3.6 million) of the field entries being invalid.
-	That high number indicates that the computation of this field is highly inaccurate. We recommend
-	dropping this field for future analytic efforts.
-
-	Furthermore, the field incident\_zip while having only .07\% invalid entries, still is a problem with 4163 such errors.
-	As the zip code is a primary measure of many NYC city services, 
-	and with a freely available database for which to validate entries, it seems an oversight 
-	that any such errors could creep into the 311 system. Some invalid entries can  
-	clearly be identified as incorrect just by observation, e.g. 10000, 12345, 11111, etc.
-
-	The breakdown of the invalid entries in the zip\_code, sorted by Agency shows that the distribution by percentage
-	almost precisely mirrors the overall breakdown of SRs by Agency as shown in . This indicates a systemic problem, and since the
-	zip\_code fields is one of the computed fields, it appears these errors are caused by incorrect computations rather
-	than by any Agency mistake.
-
-	The  errors in the incident\_zip field are more troubling, even though they are, percentage-wise, small. Here is a graphic 
-	illustrating how these errors occur by Agency. Here we can see that the majority of these errors lie in the Dept of Transportation,
-	NYPD, Taxi \& Limo Commission (TLC), and the Economic Development Council (EDC) among others. This representation
-	does not follow the full SR distribution indicating that these errors are likely generated by an incorrect
-	process or application residing at those particular Agencies; it is not a systemic problem.
-
-	\begin{figure}[tbp]
-		\centering
+	 Unfortunately some fields proved to be  problematic when comparing 
+	 the values to a domain of legal values. For example, all zip codes 
+	 (two fields: zip\_codes and incident\_zip) should be valid as defined 
+	 by the USPS database which contains 37,946 valid zip codes.
+
+	The computed field zip\_codes proved especially problematic with 
+	58\% (3.6 million) of the field entries being invalid. That high number 
+	indicates that the computation of this field is highly inaccurate. We 
+	recommend dropping this field for future analytic efforts.
+
+	Furthermore, the field incident\_zip while having only .07\% invalid entries, 
+	still is a problem with 4163 such errors. As the zip code is a primary 
+	measure of many NYC city services, and with a freely available 
+	database for which to validate entries, it seems an oversight 
+	that any such errors could creep into the 311 system. 
+	Some invalid entries can clearly be identified as incorrect by 
+	observation, e.g. 10000, 12345, 11111, etc.
+
+	The breakdown of the invalid entries in the zip\_code, sorted by 
+	Agency shows that the distribution by percentage almost precisely 
+	mirrors the overall breakdown of SRs by Agency as shown in . 
+	This indicates a systemic problem, and since the zip\_code fields is 
+	one of the computed fields, it appears these errors are caused by 
+	incorrect computations rather than by any Agency mistake.
+
+	The  errors in the incident\_zip field are more troubling, even though 
+	they are, percentage-wise, small. Here is a graphic illustrating how 
+	these errors occur by Agency. Here we can see that the majority of 
+	the errors lie in the Dept of Transportation, NYPD, Taxi \& Limo 
+	Commission (TLC), and the Economic Development Council (EDC) 
+	among others. This representation does not follow the full SR 
+	distribution indicating that these errors are likely generated by 
+	an incorrect process or application residing at those particular 
+	Agencies; it is not a systemic problem.
+
+	\begin{figure}[tbp]
+		\centering
+		\includegraphics[width=\textwidth]{invalid_incident_zip.pdf}
 		\caption{Invalid incident\_zip(s) by Agency} 
-		\includegraphics[width=\textwidth]{invalid_incident_zip.pdf} 
 		\label{fig:invalid_incident_zip}
@@
 
-		\subsubsection{Case Study: Noise Complaints by Zip Code}{label{sec:case-study-zip-codes}
-		\textbf{Scenario:} The NYC Office of Nightlife (ONL) wants to know ``What are the top 10 zip codes for Noise Complaints (all 8 types) over the last two years?''
-		The goal is to assess the impact of the recent NYC effort looking to promote a safe and vibrant nightlife scene in NYC while seeking to ease
-		strained relations between bar and club owners.
-
-		It may come as a surprise to non-NYC residents that dancing at bars and restaurants has been illegal since 1926 (during prohibition), and was only just repealed in 2018.
-
-		On the surface this would seem like a simple effort.  The NYC Open Data Portal allows for selection of a timeframe (2022-2023), complaint-type
-		begins with ``Noise'', and group by Zip code, sorted descending by count.  You can do this analysis and grouping right inside the NYC Open Data Portal. \textit{Voila!}
-
-		However, this analysis uses the zip\_codes field, one of the (six) computed fields that has shown to have validity problems. If we repeat the analysis with the 
-		incident\_zip field instead of the zip\_codes field, we obtain a different set of zip code results:
-
-		Let's subject these two data sets to validation against the US Postal Service zip code database.
+		\subsubsection{Case Study: Noise Complaints by Zip Code}
+		{label{sec:case-study-zip-codes}
+		\textbf{Scenario:} The NYC Office of Nightlife (ONL) wants to 
+		know ``What are the top 10 zip codes for Noise Complaints (all 8 types) 
+		over the last two years?'' The goal is to assess the impact of the 
+		recent NYC effort looking to promote a safe and vibrant 
+		nightlife scene in NYC while seeking to ease strained 
+		relations between bar and club owners.
+
+		It may come as a surprise to non-NYC residents that dancing at 
+		bars and restaurants has been illegal since 1926 (during prohibition), 
+		and was only just repealed in 2018.
+
+		On the surface this would seem like a simple effort.  The NYC Open 
+		Data Portal allows for selection of a timeframe (2022-2023), 
+		complaint-type begins with ``Noise'', and group by Zip code, 
+		sorted descending by count. You can do this analysis and 
+		grouping right inside the NYC Open Data Portal. \textit{Voila!}
+
+		However, this analysis uses the zip\_codes field, one of the (six) 
+		computed fields that has shown to have validity problems. 
+		If we repeat the analysis with the incident\_zip field instead 
+		of the zip\_codes field, we obtain a different set of zip code results:
+
+		Let's subject these two data sets to validation against 
+		the US Postal Service zip code database.
 
@@
     \caption{Comparison of Top Ten Zip Codes Lists}
-    \begin{tabular}{@{}lS[table-format=6.0,round-mode=places,round-precision=0]c@{\hskip 0.5cm}@{}lS[table-format=6.0,round-mode=places,round-precision=0]c@{}}
+    \begin{tabular}{@{}lS[table-format=6.0,round-mode=places,
+    round-precision=0]c@{\hskip 0.5cm}@{}lS[table-format=6.0,
+    round-mode=places,round-precision=0]c@{}}
         \toprule
@@
         \cmidrule(r){1-3} \cmidrule(l){4-6}
-        \textbf{Zip Code} & \textbf{Count} & \textbf{Valid?} & \textbf{Zip Code} & \textbf{Count} & \textbf{Valid?} \\
+        \textbf{Zip Code} & \textbf{Count} & \textbf{Valid?} 
+        & \textbf{Zip Code} & \textbf{Count} & \textbf{Valid?} \\
         \midrule
@@
 
-	As indicated, six out of ten zip\_codes are invalid, which corresponds closely with what is observed in the overall dataset (58\%). Whereas the incident\_zip
-	field is completely valid, again in-line with the overall incident\_zip validation percentage (99.04\%). If the analysis performed using the zip\_codes field were to be
-	presented to the Director, ONL, the conclusion would be in error. Perhaps more curious than the differing zip codes on the two lists, 
-	is the fact that the numerical counts are nearly identical;  the overall counts are quite close. This is curious since
-	one field is computed while the other is via data entry; the computed protocol is again the suspect. The algorithm appears to be counting corrrectly (or nearly so)
-	but mis-identifying the corresponding zip code.
+As indicated, six out of ten zip\_codes are invalid, which corresponds closely 
+with what is observed in the overall dataset (58\%). Whereas the incident\_zip 
+field is completely valid, again in-line with the overall incident\_zip 
+validation percentage (99.04\%). If the analysis performed using the 
+zip\_codes field were to be 	presented to the Director, ONL, the conclusion 
+would be in error. Perhaps more curious than the differing zip codes 
+on the two lists, is the fact that the numerical counts are nearly identical; 
+the overall counts are quite close. This is curious since one field is 
+computed while the other is via data entry; the computed protocol is 
+again the suspect. The algorithm appears to be counting corrrectly 
+(or nearly so) 	but mis-identifying the corresponding zip code.
 
@@
 	\label{sec:police-precincts}
-	A curious case also exists when examining the two nearly identical fields - police\_precincts and police\_precinct. Both of those fields are among
-	the ``computed'' fields in the dataset. 
-	Using \href{https://www.nyc.gov/site/nypd/bureaus/patrol/precincts-landing.page}{NYPD Precinct listings} it's possible to determine the valid
-	police precincts; they are generally numeric or have numeric representations in the dataset.
-
-	However, what we find is that both fields police\_precinct and police\_precincts  have  35\% invalid entries.
-	Unfortunately, they're not the same invalid entries.
-
-	For the police\_precincts field, there are 2,171,864 invalid entries (35\%)., a consequential number of errors. Similarly for the police\_precinct field,
-	 there are 2,171,778 invalid entries ( also 35\%), however they are not the same counts for the two similar fields.
+	A curious case also exists when examining the two nearly identical 
+	fields - police\_precincts and police\_precinct. Both of those fields 
+	are among the ``computed'' fields in the dataset. Using 
+	\href{https://www.nyc.gov/site/nypd/bureaus/patrol/precincts-landing.page}
+	{NYPD Precinct listings} it's possible to determine the valid 
+	police precincts; they are generally numeric or have numeric 
+	representations in the dataset.
+
+	However, what we find is that both fields police\_precinct 
+	and police\_precincts  have  35\% invalid entries. Unfortunately, 
+	they're not the same invalid entries.
+
+	For the police\_precincts field, there are 2,171,864 invalid entries 
+	(35\%)., a consequential number of errors. Similarly for the 
+	police\_precinct field, there are 2,171,778 invalid entries 
+	(also 35\%), however they are not the same counts for the two similar fields.
 
@@
 		\toprule
-		\textbf{Invalid Precinct} & \textbf{police\_precinct count} & \textbf{police\_precincts count} & \textbf{Delta}\\
+		\textbf{Invalid Precinct} & \textbf{police\_precinct count} 
+		& \textbf{police\_precincts count} & \textbf{Delta}\\
 		\midrule
@@
 
-	A graphic representation of the invalid police\_precincts plotted by Agency show a near perfect match to the
-	distribution of the entire 6.4 million SRs as seen in Figure \ref{fig:SR_counts_by_Agency}. Note that the ``big six'' Agencies account for 90\% of these invalid precinct errors, identical
-	to the overall distribution of all SRs.
-(The same is true for a similar graph of by the police\_precint field.) This strongly suggests that this is a systemic problem, almost certainly caused by the computation method applied for these two fields.
-
-	\begin{figure}[tbp]
-	  \centering
+	A graphic representation of the invalid police\_precincts plotted 
+	by Agency show a near perfect match to the distribution of the 
+	entire 6.4 million SRs as seen in Figure \ref{fig:SR_counts_by_Agency}. 
+	Note that the ``big six'' Agencies account for 90\% of these invalid 
+	precinct errors, identical to the overall distribution of all SRs. 
+	(The same is true for a similar graph of by the police\_precint field.) 
+	This strongly suggests that this is a systemic problem, 
+	almost certainly caused by the computation method applied for these two fields.
+
+	\begin{figure}[tbp]
+	  \centering
+	  \includegraphics[width=\textwidth]{invalid_police_precincts.pdf}
 	  \caption{Invalid NYPD precincts by Agency}	  
-	  \includegraphics[width=\textwidth]{invalid_police_precincts.pdf}	  
 	  \label{fig:invalid_police_precincts_zip}
@@
 	\subsection{Issues with Community Boards}
-	Community Boards are an important aspect of NYC government. They are the most local, grassroots form of City government, 
-	and a vital connection between communities and elected officials and City agencies. They are organized by the five Boroughs of NYC and represented
-	in this dataset as ``\#\#-Borough'', e.g. ``10 Bronx'' and report to the various Borough Presidents in NYC. As such, they play an important role in the quality of life for all New Yorkers.
-	There are 59 community boards throughout the City: 12 Community Districts in the Bronx, 18 in Brooklyn, 12 in Manhattan, 14 in Queens and 3 in Staten Island.
-	The Community Boards are used frequently as a way to measure distribution of City services throughout the five boroughs, so correctness is important. 
-	NYC Borough Presidents in particular are focused on the services offered at the Community Board level within their respective Borough.
-
-	In the 2022-2023 dataset there are 27,276 invalid community\_board entries which represents 0.43\% of non-blank data. 
-	There are a total of 12 different invalid community boards, to include:
+	Community Boards are an important aspect of NYC government. They are the 
+	most local, grassroots form of City government, and a vital connection 
+	between communities and elected officials and City agencies. 
+	They are organized by the five Boroughs of NYC and represented	
+	in this dataset as ``\#\#-Borough'', e.g. ``10 Bronx'' and report to the 
+	various Borough Presidents in NYC. As such, they play an important 
+	role in the quality of life for all New Yorkers. There are 59 community 
+	boards throughout the City: 12 Community Districts in Bronx, 18 in Brooklyn, 
+	12 in Manhattan, 14 in Queens and 3 in Staten Island. The Community 
+	Boards are used frequently as a way to measure distribution of 
+	City services throughout the five boroughs, so correctness is important. 
+	NYC Borough Presidents in particular are focused on the services 
+	offered at the Community Board level within their respective Borough.
+
+	In the 2022-2023 dataset there are 27,276 invalid community\_board 
+	entries which represents 0.43\% of non-blank data. There are a total 
+	of 12 different invalid community boards, to include:
 
@@
 		\centering
+		\normalsize
 		\caption{Top Ten Invalid 'community\_board' Values}
-		\normalsize
 		\begin{tabular}{rlr}
@@
 
-		The distribution of invalid Community Boards by Agency is not consistent with the overall SR Agency distribution. This indicates that there are likely
-		specific issues at some key Agencies, in this case Taxi \& Limo Commission (TLC), Parks \& Recreation, etc. It's not a large error, but the division of services
-		(and complaints) by Community Board is a well-tracked measure so accuracy is again important.
+		The distribution of invalid Community Boards by Agency is not consistent 
+		with the overall SR Agency distribution. This indicates that there are likely 
+		specific issues at some key Agencies, in this case Taxi \& Limo Commission 
+		(TLC), Parks \& Recreation, etc. It's not a large error, but the 
+		division of services (and complaints) by Community Board is a 
+		well-tracked measure so accuracy is again important.
 
@@
 	\subsection{Issues with Community Districts}
-	Another of the computed fields is community\_districts. Community Districts are the boundaries for the Community Boards, but unlike how Community Boards are
-	an arm of the government of New York City, the Community District is used by the Department of City Planning (DCP) for purposes of environmental, socio-economic, and
-	demographic purposes. DCP is the NYC''s primary land use agency and is instrumental in the City's physical framework.  It is a geographical division rather
-	than a local government division. And like Community Boards, the Community Districts are frequently cited as a means to track zoning, housing, community facilities,
-	and waterfront/open/public spaces. It is another of the NYC well-tracked measures.
-
-	Due to how the community\_district data is formatted, it is not possible to establish validity. However, it is possible to determine that the dataset contains
-	72 unique entries, while there are only 59 valid Community Districts. So it would appear that there are 13 invalid community\_board(s) contained in the dataset.
-
-As previously mentioned, the dataset has four data fields:  created\_date, closed\_date, due\_date, and resolution\_action\_update\_date.
-The due\_date field has few entries to analyze; 99.62\% of due\_date entries are ``N/A''.  It is not known if the due\_date(s) are truely
-``not applicable'' or that the due\_date field is rarely used. The  resolution\_action\_update\_date field serves as a date-time marker
-showing when the SR record was last updated.  All the date fields are recorded to the second, e/g/ ``2022-02-11 13:43:05''.
+	Another of the computed fields is community\_districts. Community Districts 
+	are the boundaries for the Community Boards, but unlike how 
+	Community Boards are an arm of the government of New York City, 
+	the Community District is used by the Department of City Planning 
+	(DCP) for purposes of environmental, socio-economic, and demographic 
+	purposes. DCP is the NYC''s primary land use agency and is 
+	instrumental in the City's physical framework.  It is a geographical 
+	division rather than a local government division. And like Community 
+	Boards, the Community Districts are frequently cited as a means to 
+	track zoning, housing, community facilities, and waterfront/open/public spaces. 
+	It is another of the NYC well-tracked measures.
+
+	Due to how the community\_district data is formatted, it is not possible 
+	to establish validity. However, it is possible to determine that the 
+	dataset contains 72 unique entries, while there are only 59 valid Community 
+	Districts. So it would appear that there are 13 invalid 
+	community\_board(s) contained in the dataset.
+
+As previously mentioned, the dataset has four data fields:  created\_date, 
+closed\_date, due\_date, and resolution\_action\_update\_date. 
+The due\_date field has few entries to analyze; 99.62\% of due\_date entries 
+are ``N/A''.  It is not known if the due\_date(s) are truely ``not applicable'' 
+or that the due\_date field is rarely used. The  resolution\_action\_update\_date 
+field serves as a date-time marker showing when the SR record was last updated.  
+All the date fields are recorded to the second, e/g/ ``2022-02-11 13:43:05''.
 
 \subsection{created\_date and closed\_date(s) -- Negative Duration}
-Let's look first at the created\_date and closed\_date fields. You might expect that these two fields are automatically populated by the 
-SR application software, either the software used at the 311 organization or that used at the Agencies that are integrated with the 311 system;
-and typically that is indeed the case. In such a case, the act of creating an SR would automatically trigger an entry in created\_date based on a system-wide clock. Thus
-the created\_date is associated with setting an SR status to ``new'' or ``open''. And similarly when an SR's status is changed to
-``closed'' then the closed\_date would be populated.
-
-Unfortunately, our analysis found that that assumption is not always the case. When you analyze the various date fields, several
-anomalies become quickly apparent:
+Let's look first at the created\_date and closed\_date fields. You might 
+expect that these two fields are automatically populated by the 
+SR application software, either the software used at the 311 organization 
+or that used at the Agencies that are integrated with the 311 system; 
+and typically that is indeed the case. In such a case, the act of creating an 
+SR would automatically trigger an entry in created\_date based on a 
+system reference clock. Thus the created\_date is associated with 
+setting an SR status to ``new'' or ``open''. And similarly when an SR's 
+status is changed to ``closed'' then the closed\_date would be populated.
+
+Unfortunately, our analysis found that that assumption is not always 
+the case. When you analyze the various date fields, several anomalies 
+quickly become apparent:
 
@@
 		    \item Some SRs have a closed\_date that occurs before the created\_date.
-		    \item Some created\_date(s) and closed\_date(s) appear either in the future or in the far distant past.
-		    \item There are a large number of SRs that are closed and created exactly at midnight or exactly at noon. 
+		    \item Some created\_date(s) and closed\_date(s) appear either in 
+		    the future or in the far distant past.
+		    \item There are a large number of SRs that are closed and created 
+		    exactly at midnight or exactly at noon. 
 	\end{itemize}
 
-\textbf{Why are these date fields important?} They're important because citizens, NYC Government Officials, and NYC Agencies use these dates
-to measure the responsiveness for providing the services. How long does it take to replace an out-of-order street light?
-How rapid is the response to a report of domestic violence? Does the NYPD response to a noise complaint in Staten Island take
-longer than in Manhattan? How long does it take to repair a street pothole in Queens?
-
-The measure of ``duration'', of the response time of a 311 call is a closely watched, carefully scrutinized metric, 
-both in terms of overall performance and with a geographical area. And it can have broad political, organizational,
-and budgetary consequences. Duration is not a field in the dataset, but it is easily computed by closed\_date - created\_date.
-Let's look at some anomalies in this ``duration'' field.
-
-Closed-before-Created:  There are 12,251 SRs that are closed before they were created, thereby generating a nonsensical 
-``negative duration''. While this is a small percentage overall (0.2\%) it can have significant impact on response time analysis.
-In most cases, these ``negative durations'' appear to just be mistakes. Here is a sample of some of the largest and the
-smallest errors:
+\textbf{Why are these date fields important?} They're important because citizens, NYC Government Officials, and NYC Agencies use these dates to measure the responsiveness 
+for providing the services. How long does it take to replace an out-of-order 
+street light? How rapid is the response to a report of domestic violence? 
+Does the NYPD response to a noise complaint in Staten Island take longer 
+than in Manhattan? How long does it take to repair a street pothole in Queens?
+
+The measure of ``duration'', of the response time of a 311 call is a closely 
+watched, carefully scrutinized metric, both in terms of overall performance 
+and with a geographical area. And it can have broad political, organizational, 
+and budgetary consequences. Duration is not a field in the dataset, 
+but it is easily computed by closed\_date - created\_date. Let's look at 
+some anomalies in this ``duration'' field.
+
+Closed-before-Created:  There are 12,251 SRs that are closed before they 
+were created, thereby generating a nonsensical ``negative duration''. 
+While this is a small percentage overall (0.2\%) it can have significant impact 
+on response time analysis. In most cases, these ``negative durations'' 
+appear to just be mistakes. Here is a sample of some of the 
+largest and the smallest errors:
 
@@
     \centering
+    \normalsize
     \caption{Largest and Smallest errors (days)}
-    \normalsize
     \begin{tabular}{>{\ttfamily}l l l r l}
         \toprule
-        \multicolumn{5}{c}{\textbf{Largest errors (days) excluding extreme negative values}} \\
+        \multicolumn{5}{c}{\textbf{Largest errors (days) excluding 
+        extreme negative values}} \\
         \midrule
-        \textbf{created\_date} & \textbf{closed\_date} & \textbf{duration} & \textbf{agency} \\
+        \textbf{created\_date} & \textbf{closed\_date} & \textbf{duration} 
+        & \textbf{agency} \\
 	        \midrule
@@
 	        \midrule
-	        \textbf{created\_date} & \textbf{closed\_date} & \textbf{duration} & \textbf{agency} \\
+	        \textbf{created\_date} & \textbf{closed\_date} & \textbf{duration} 
+	        & \textbf{agency} \\
 	        \midrule
@@
 
-The largest errors are shown ``excluding extreme negative values''. We found eight SRs with extremely large negative durations (\textless{} -730),
-all containing an entry of ``1900-01-01'' as the closed\_date, which generates extremely large negative durations exceeding 44,601 days (122 yrs).
-That large of an anomaly can skew statistical results, even though the number of occurrences is small. This is especially curious
-as the date ``1900-01-01'' is also the earliest date that Microsoft Excel can accommodate as well as how Excel treats a blank
-cell that is designated as a date. All raising suspicions that perhaps these dates were imported into the 311 system from Excel.
-
-According, these  SR rows are removed from the box \& whiskers plot. Note that these extremely large negative duration SRs all
-are from the Department of Homeland Services (DHS) Sample:
+The largest errors are shown ``excluding extreme negative values''. We found 
+eight SRs with extremely large negative durations (\textless{} -730), all containing 
+an entry of ``1900-01-01'' as the closed\_date, which generates extremely large 
+negative durations exceeding 44,601 days (122 yrs). That large of an anomaly 
+can skew statistical results, even though the number of occurrences is small. 
+This is especially curious as the date ``1900-01-01'' is also the earliest date that 
+Microsoft Excel can accommodate as well as how Excel treats a blank cell 
+that is designated as a date. All raising suspicions that perhaps these dates 
+were imported into the 311 system from Excel.
+
+According, these  SR rows are removed from the box \& whiskers plot. 
+Note that these extremely large negative duration SRs are all from the 
+Department of Homeland Services (DHS) Sample:
 
@@
     \centering
+     \normalsize
     \caption{Sample of SRs with extremely large negative durations}
-    \normalsize
     \begin{tabular}{l l r l}
         \toprule
-        \textbf{created\_date} & \textbf{closed\_date} & \textbf{duration} & \textbf{agency} \\
+        \textbf{created\_date} & \textbf{closed\_date} & \textbf{duration} 
+        & \textbf{agency} \\
 	        \midrule
@@
 
-Here is a graphic visualization of the negative-duration SRs by Agency. It is obvious that this
-negative-duration issue is a problem at the Dept. of Transportation (DOT) where 95\% of these
-types of errors occur.
-
-\begin{figure}[tbp]
- 	 \centering
+Here is a graphic visualization of the negative-duration SRs by Agency. It is 
+obvious that this negative-duration issue is a problem at the 
+Dept. of Transportation (DOT) where 95\% of these types of errors occur.
+
+\begin{figure}[tbp]
+ 	 \centering
+ 	 \includegraphics[width = \textwidth]{negative_duration_SR_barchart.pdf}
 	  \caption{Negative Duration SRs by Agency}
-	  \includegraphics[width = \textwidth]{negative_duration_SR_barchart.pdf}
 	  \label{fig:negative-duration}
@@
 
-This violin chart shows the broad spread of negative-duration SRs, albeit with the extremely large 
-negative-duration SR removed. While there are few outlier points, the magnitude of the negative-durations
-is troubling and can produce bizarre analytical results.
-
-\begin{figure}[tbp]
- 	 \centering
+This violin chart shows the broad spread of negative-duration SRs, albeit with 
+the extremely large negative-duration SR removed. While there are few 
+outlier points, the magnitude of the negative-durations is troubling and can 
+produce bizarre analytical results.
+
+\begin{figure}[tbp]
+ 	 \centering
+ 	 \includegraphics[width = \textwidth]{negative_duration_SR_violin.pdf}
 	 \caption{Distribution of Negative Durations}
-	  \includegraphics[width = \textwidth]{negative_duration_SR_violin.pdf}
 	 \label{fig:negative-duration-violin}
@@
 \subsubsection{Case Study: Homeless Person Assistance}
-		\textbf{Scenario:} The Dept. of Homeless Services wants to know ``How quickly are 311 calls for ``Homeless Person
-		Assistance over the past two years?'' resolved. This is a typical request made by both the public and City government. It's a performance metric along the lines
-		of ``How quickly is my Agency responding to ``critical'' requests, and does that performance vary by Borough, Zip Code, etc.
-
+		\textbf{Scenario:} The Dept. of Homeless Services wants to 
+		know ``How quickly are 311 calls for ``Homeless Person Assistance 
+		over the past two years?'' resolved. This is a typical request made by 
+		both the public and City government. It's a performance metric along 
+		the lines of ``How quickly is my Agency responding to ``critical'' 
+		requests, and does that performance vary by Borough, Zip Code, etc. 
 		As an analyst, this appears to be a routine effort:
@@
 		\begin{enumerate}
-		    \item Select data from 2022-2023 and filter the data by complaint\_type = ``Homeless Person Assistance'' (yields 55,000 SRs)
+		    \item Select data from 2022-2023 and filter the data by 
+		    complaint\_type = ``Homeless Person Assistance'' (yields 55,000 SRs)
 		    \item Compute a new field ``duration'' (closed\_date – created\_date)
@@
 
-		Clearly that answer is nonsensical. How did such a simple task result in an absurd answer? The answer lies in the computation of the
-		``duration'' field. As it turns out, there are eight DHS SRs that have a closed\_date of ``1900-01-01''. Each of those SRs creates a 
-		negative duration of -44,602 days (-122 years). So just those eight SRs with extreme negative duration values is enough to drive
-		the average of the 75,000 Homeless Assistance SRs to a negative value, clearly incorrect.
-
-		As it turns out the distribution of the ``Homeless Person Assistance'' is quite skewed due to a small number of long duration
-		SRs. In this case, the median is a better measure of central tendency than the average. And the median of the duration
-		field is  0.2 days (approx 5 hrs). Here is a look at the distribution of durations for the Homeless Person Assistance SRs showing
-		the large outliers.
-
-		\begin{figure}[tbp]
-		 	 \centering
+		Clearly that answer is nonsensical. How did such a simple task result 
+		in an absurd answer? The answer lies in the computation of the ``duration'' 
+		field. As it turns out, there are eight DHS SRs that have a closed\_date 
+		of ``1900-01-01''. Each of those SRs creates a negative duration of -44,602 
+		days (-122 years). So just those eight SRs with extreme negative duration 
+		values is enough to drive the average of the 75,000 
+		Homeless Assistance SRs to a negative value, clearly incorrect.
+
+		As it turns out the distribution of the ``Homeless Person Assistance'' is 
+		quite skewed due to a small number of long duration SRs. In this case, 
+		the median is a better measure of central tendency than the average. 
+		And the median of the duration field is  0.2 days (approx 5 hrs). 
+		Here is a look at the distribution of durations for the Homeless Person 
+		Assistance SRs showing 	the large outliers.
+
+		\begin{figure}[tbp]
+		 	 \centering
+		 	 \includegraphics[width = \textwidth]{homeless_response_time_clean.pdf}
 			 \caption{Homeless Assistance SR Durations}
-			  \includegraphics[width = \textwidth]{homeless_response_time_clean.pdf}
 			 \label{fig:homeless}
@@
 	\subsection{created\_date and closed\_date(s) --  Zero Duration}		
-	An even more serious, and more prevalent problem with closed\_date(s) occurs when the closed\_date and created\_date
-	are exactly the same -- to the second -- which accordingly creates a \textbf{zero duration'}. Again this is nonsensical, but the
-	presence of such zero durations can again severely distort analytics. There are 191,141 such SRs representing 3.1\% of all 
-	non-blank data; not an insignificant number. As shown in the chart below, 99\% of the zero duration SRs appear to predominately
-	occur in five Agencies (Dept of Mental Health \&Hygiene, Dept of Transportation, Dept of Business, Dept of Sanitation,
-	and Dept of Environmental Protection.) This is not in-line with the overall Agency distribution of SRs and indicates an
+	An even more serious, and more prevalent problem with closed\_date(s) 
+	occurs when the closed\_date and created\_date are exactly the same -- to 
+	the second -- which accordingly creates a \textbf{zero duration'}. Again 
+	this is nonsensical, but the presence of such zero durations can again 
+	severely distort analytics. There are 191,141 such SRs representing 3.1\% of 
+	all non-blank data; not an insignificant number. As shown in the chart 
+	below, 99\% of the zero duration SRs appear to predominately occur in five 
+	Agencies (Dept of Mental Health \&Hygiene, Dept of Transportation, 
+	Dept of Business, Dept of Sanitation, and Dept of Environmental Protection.) 
+	This is not in-line with the overall Agency distribution of SRs and indicates an 
 	Agency-specific issue, and the solution likely lies at those Agencies.
@@
 		 \centering
+		 \includegraphics[width = \textwidth]{zero_duration_SR.pdf}
 		 \caption{SRs with Zero Durations by Agency}
-		 \includegraphics[width = \textwidth]{zero_duration_SR.pdf}
 		 \label{fig:zero-duration}
@@
 	\subsection{created\_date and closed\_date(s) -- Midnight \&Noon}
-	An additional problem discovered with the created\_date and closed\_date fields; there appears to be an unusually large number of 
-	SRs created or closed at exactly midnight (00:00:00) and exactly noon (12:00:00), to the second. The distribution of SR creation and closure 
-	largely follows the work-day clock with many SRs created during day-light hours, and fewer SRs 
-	created at night and in the early hours of the morning. 
-	However, here there appears to be significantly greater numbers of SR closed exactly at midnight and exactly at noon,
-	as well as a significant number of SRs created exactly at noon. The reason behind this anomaly is worth investigating 
-	as it affects the SR ``duration'' question. If SR durations are used to measure responsiveness
-	for the delivery of services, and if those times are not an accurate reflection of the start/finish of the Service, then the use of
-	duration as a service quality metric is called into question. Observations during this 2-year period include:
+	An additional problem discovered with the created\_date and 
+	closed\_date fields; there appears to be an unusually large number of SRs 
+	created or closed at exactly midnight (00:00:00) and exactly noon (12:00:00), 
+	to the second. The distribution of SR creation and closure largely follows the 
+	work-day clock with many SRs created during day-light hours, and 
+	fewer SRs 	created at night and in the early hours of the morning. However, 
+	here there appears to be significantly greater numbers of SR closed exactly 
+	at midnight and exactly at noon, as well as a significant number of SRs 
+	created exactly at noon. The reason behind this anomaly is worth investigating as 
+	it affects the SR ``duration'' question. If SR durations are used to measure 
+	responsiveness for the delivery of services, and if those times are not an 
+	accurate reflection of the start/finish of the Service, then the use of duration 
+	as a service quality metric is called into question. Observations during 
+	this 2-year period include:
 
@@
 
-	There is, of course, a broad distribution of the time that SRs are created by day and hour, by night and day, by weekday and weekend. 
-	One would think that that distribution would be (mostly) evenly distributed by the minute and second. That is, 
-	if there are normally 100 SRs created between 1-2pm, that the distribution to the minute and second 
-	would be close to random. For example, there is no reason to suspect that a larger number of SRs would be 
-	created at 6 minutes past the hour rather than at 5 minutes past the hour. And certainly, if you look at the second that these SRs
-	are created you would again expect a (more-or-less) random distribution; that there would be no reason to believe 
-	that a higher (or lower) number of SRs would arrive at 18 seconds past the minute versus at 39 seconds past the minute.
-	Unfortunately, that turned out not to be true.
-
-	Let's look at the SR creation distribution. We discovered that there are a significant number of SRs created exactly at noon. To reveal this
-	trend, we looked at the exact date-time stamp of the created\_date(s). We began by selected only SRs that were created exactly on the minute;
-	at 09:00:00 or 13:00:00 for example.  Here is a visualization of the number of SRs created and closed by hour-of-the-day:
+	There is, of course, a broad distribution of the time that SRs are 
+	created by day and hour, by night and day, by weekday and weekend.  
+	One would think that that distribution would be (mostly) evenly distributed 
+	by the minute and second. That is, if there are normally 100 SRs created 
+	between 1-2pm, that the distribution to the minute and second would be 
+	close to random. For example, there is no reason to suspect that a larger 
+	number of SRs would be created at 6 minutes past the hour rather 
+	than at 5 minutes past the hour. And certainly, if you look at the second 
+	that these SRs are created you would again expect a (more-or-less) 
+	random distribution; that there would be no reason to believe that a higher 
+	(or lower) number of SRs would arrive at 18 seconds past the minute 
+	versus at 39 seconds past the minute. Unfortunately, that turned out not to be true.
+
+	Let's look at the SR creation distribution. We discovered that there 
+	are a significant number of SRs created exactly at noon. To reveal this trend, 
+	we looked at the exact date-time stamp of the created\_date(s). We began 
+	by selected only SRs that were created exactly on the minute; at 09:00:00 
+	or 13:00:00 for example.  Here is a visualization of the number of SRs 
+	created and closed by hour-of-the-day:
 
@@
 
-	The created\_date shows a maximum value at 12:00. The closed\_date chart show significant spikes at both midnight
-	and noon. The midnight (00:00) spike appears to be an anomaly especially as compared to the hours just 
-	before and after midnight where there is normally a pronounced slump in SR closures. 
-	And both the closed-at-noon and closed-at-midnight are approaching 2\textsigma's from the hourly average.
-
-	Let's drill down on each of these outlier points. First, we aggregate the created\_date over the 2-year period that fall exactly
-	on the top-of-the-hour with the minute and second values equal to 0, e.g. 09:00:00, 14:00:00. This visualization
-	reveals a clearer picture of where these spikes are:
-
-	\begin{figure}[tbp]
-		\centering
- 		\includegraphics[width = \textwidth]{2-year-trend-SR_created_by_top_of_hour.pdf}
+	The created\_date shows a maximum value at 12:00. The closed\_date chart 
+	show significant spikes at both midnight and noon. The midnight (00:00) 
+	spike appears to be an anomaly especially as compared to the hours just 
+	before and after midnight where there is normally a pronounced slump 
+	in SR closures. And both the closed-at-noon and closed-at-midnight are 
+	approaching 2\textsigma's from the hourly average.
+
+	Let's drill down on each of these outlier points. First, we aggregate the 
+	created\_date over the 2-year period that fall exactly on the top-of-the-hour 
+	with the minute and second values equal to 0, e.g. 09:00:00, 14:00:00. 
+	This visualization reveals a clearer picture of where these spikes are:
+
+	\begin{figure}[tbp]
+		\centering
+ 		\includegraphics[width = \textwidth]
+ 		{2-year-trend-SR_created_by_top_of_hour.pdf}
 		\caption{SRs Created at Top-of-the-Hour}
@@
 
-	One other way to visualize this anomaly is to look at the busiest day during this 2-year period (Friday, 2023-09-29).
-	Here we will aggregated the created\_date(s) by minute (with seconds equal to zero). This provides a 
-	minute-by-minute look at SR creation on the busiest day of this study. Here we see a clear spike at 
-	exactly noon (12:00), well beyond the 3\textsigma line, and well above any other hour of the day.
-
-	\begin{figure}[tbp]
-		\centering
-		\includegraphics[width=\textwidth]{2-year-trend-SR_created_by_minute_of_busiest_day.pdf}
+	One other way to visualize this anomaly is to look at the busiest day 
+	during this 2-year period (Friday, 2023-09-29). Here we will aggregated 
+	the created\_date(s) by minute (with seconds equal to zero). This 
+	provides a minute-by-minute look at SR creation on the busiest day 
+	of this study. Here we see a clear spike at exactly noon (12:00), well 
+	beyond the 3\textsigma line, and well above any other hour of the day.
+
+	\begin{figure}[tbp]
+		\centering
+		\includegraphics[width=\textwidth]
+		{2-year-trend-SR_created_by_minute_of_busiest_day.pdf}
 		\caption{SRs Created Minute-by-Minute on Busiest Day}
@@
 
-	Let's take a similar approach to the closed\_date distribution over the 2-year period. We aggregate 
-	SRs by close\_date to find the top-of-the-hour, where minute and second values equal to 0, e.g. 20:00:00. 
-	Here we see spikes at both midnight (00:00) and at noon (12:00), with the midnight spike being 
-	well beyond the 3\textsigma line. Also note how much lower (and nearly equal) all the other
-	top-of-the-hour values are, relative to the midnight and noon spikes.
-
-	\begin{figure}[tbp]
-		\centering
-		\includegraphics[width = \textwidth]{2-year-trend-SR_closed_by_top_of_hour.pdf}
+	Let's take a similar approach to the closed\_date distribution over the 
+	2-year period. We aggregate SRs by close\_date to find the top-of-the-hour, 
+	where minute and second values equal to 0, e.g. 20:00:00. Here we see 
+	spikes at both midnight (00:00) and at noon (12:00), with the midnight 
+	spike being well beyond the 3\textsigma line. Also note how much lower 
+	(and nearly equal) all the other top-of-the-hour values are, 
+	relative to the midnight and noon spikes.
+
+	\begin{figure}[tbp]
+		\centering
+		\includegraphics[width = \textwidth]
+		{2-year-trend-SR_closed_by_top_of_hour.pdf}
 		\caption{SRs Closed at Top-of-the-Hour}
@@
 
-	As before, let's visualize this anomaly by looking at the closed\_date(s) on the busiest day
-	during this 2-year study period (Friday, 2023-09-29). We aggregate closed\_date(s) by the
-	minute (with seconds equal to zero). This gives a minute-by-minute view of SR closure
-	on the busiest day of this study. Clearly visual is the very large closure spikes occurring
-	at both midnight and noon. The chart highlights both of these anomalies with spikes
-	of 286 and 120 at 00:00:00 and 12:00:00 respectively. Again, note how these spikes
-	are so very much above the median values for all other minutes.
-
- 	\begin{figure}[tbp]
-		\centering
-		\includegraphics[width=\textwidth]{2-year-trend-SR_closed_by_minute_of_busiest_day.pdf}
+	As before, let's visualize this anomaly by looking at the closed\_date(s) 
+	on the busiest day during this 2-year study period (Friday, 2023-09-29). 
+	We aggregate closed\_date(s) by the minute (with seconds equal to zero). 
+	This gives a minute-by-minute view of SR closure on the busiest day 
+	of this study. Clearly visual is the very large closure spikes occurring at 
+	both midnight and noon. The chart highlights both of these anomalies 
+	with spikes of 286 and 120 at 00:00:00 and 12:00:00 respectively. 
+	Again, note how these spikes are so very much above the median 
+	values for all other minutes.
+
+ 	\begin{figure}[tbp]
+		\centering
+		\includegraphics[width=\textwidth]
+		{2-year-trend-SR_closed_by_minute_of_busiest_day.pdf}
 		\caption{SRs Closed Minute-by-Minute on Busiest Day}
@@
 
-	These unusual patterns of created \& closed SRs at exactly the hours of midnight and noon likely indicates
-	the presence of a bulk create/close software process; a process that perhaps automatically ``closes'' (or ``creates'') 
-	a large number of SRs with a provided time-stamp of midnight (00:00:00) or noon (12:00:00), and does so in bulk. 
-	If so, then addressing this issue likely requires a closer look at those Agencies providing these suspect SR closed/created times.
-	 The desirable state is that the created\_date and closed\_date fields correctly reflect the true opening and closing 
-	 of the SRs, such that an accurate ``duration'' can be computed for the SR life-cycle. Otherwise, we are likely seeing 
-	 durations that are artificially manipulated by a software process performing bulk uploads to the 311 system.
-
-	Looking at the distribution by Agency for the ``closed-exactly-at-midnight'' shows that \textgreater90\% of these suspect SRs come from just two Agencies - Dept. of 
-	Buildings (DOB) and Dept. of Sanitation (DSNY). If you include HPD and DOT, that accounts for 99\% of the suspect SR closures. This Agency
-	distribution does not follow the overall SR distribution and we can assume that it is Agency specific.
+	These unusual patterns of created \& closed SRs at exactly the hours 
+	of midnight and noon likely indicates the presence of a bulk create/close 
+	software process; a process that perhaps automatically ``closes'' 
+	(or ``creates'') a large number of SRs with a provided time-stamp of 
+	midnight (00:00:00) or noon (12:00:00), and does so in bulk. 
+	If so, then addressing this issue likely requires a closer look at those 
+	Agencies providing these suspect SR closed/created times. The desirable 
+	state is that the created\_date and closed\_date fields correctly 
+	reflect the true opening and closing of the SRs, such that an 
+	accurate ``duration'' can be computed for the SR life-cycle. 
+	Otherwise, we are likely seeing durations that are artificially 
+	manipulated by a software process performing bulk uploads 
+	to the 311 system.
+
+	Looking at the distribution by Agency for the ``closed-exactly-at-midnight'' 
+	shows that \textgreater90\% of these suspect SRs come from just two 
+	Agencies - Dept. of Buildings (DOB) and Dept. of Sanitation (DSNY). 
+	If you include HPD and DOT, that accounts for 99\% of the suspect 
+	SR closures. This Agency distribution does not follow the overall 
+	SR distribution and we can assume that it is Agency specific.
 
@@
 
-	However, the ``SR closed at noon'' visualization shows that  a single Agency, DSNY,
-	is responsible for \textgreater99\% of the SRs ``closed-exactly-at-noon''. Accordingly, any effort to resolve this issue
-	necessitates working with DSNY to fully understand how this is occurring.
-
-	\begin{figure}[tbp]
-		\centering
-		\includegraphics[width = \textwidth]{closed_at_noon_chart.pdf}
+	However, the ``SR closed at noon'' visualization shows that  a single 
+	Agency, DSNY, is responsible for \textgreater99\% of the 
+	SRs ``closed-exactly-at-noon''. Accordingly, any effort to 
+	resolve this issue necessitates working with DSNY to fully understand 
+	how this is occurring.
+
+	\begin{figure}[tbp]
+		\centering
+		\includegraphics[width = \textwidth]
+		{closed_at_noon_chart.pdf}
 		\caption{SRs Closed Exactly at Noon by Agency}
@@
 
-	The 99,779 SRs created-exactly-at-noon also appears to originate with the Dept of Sanitation (DSNY) which
-	is responsible for \textgreater99\% of the suspect SRs.
+	The 99,779 SRs created-exactly-at-noon also appears to 
+	originate with the Dept of Sanitation (DSNY) which is responsible 
+	for \textgreater99\% of the suspect SRs.
 
@@
 
-	While it is difficult to ultimately know if these observations are indeed anomalies; they could be the result of normal, expected
-	311 SR activity. But at least by observation that outcome appears very unlikely.
+	While it is difficult to ultimately know if these observations 
+	are indeed anomalies; they could be the result of normal, 
+	expected 311 SR activity. But at least by observation that 
+	outcome appears very unlikely.
 
 	\subsection{resolution\_action\_update\_date}
-	When an SR is updated, by any means on any field, the software automatically populates the resolution\_action\_update\_date. Analyzing that field
-	it is apparent that some of the SR updates are happening long, long after the SR is closed. In this dataset there are eight SRs that have 
-	resolution\_action\_update\_date values of 44,601 days after the closed\_date. The problem lies not with the resolution\_action\_update\_date, but
-	rather with the closed\_date values for these 8 SRs, all of which are in the format of ``1900-01-01''. It has previously been discussed that these
-	inaccurate dates may possibly be coming from an Excel spreadsheet as that is the minimum date which Excel can handle. Here is a sample
-	of those SRs:
+	When an SR is updated, by any means on any field, the software 
+	automatically populates the resolution\_action\_update\_date. 
+	Analyzing that field it is apparent that some of the SR updates 
+	are happening long, long after the SR is closed. In this dataset 
+	there are eight SRs that have resolution\_action\_update\_date values 
+	of 44,601 days after the closed\_date. The problem lies not with 
+	the  resolution\_action\_update\_date, but rather with the closed\_date 
+	values for these 8 SRs, all of which are in the format of ``1900-01-01''. 
+	It has previously been discussed that these inaccurate dates may 
+	possibly be coming from an Excel spreadsheet as that is the minimum 
+	date which Excel can represent. Here is a sample of those SRs:
 
@@
 	    \centering
+	    \normalsize
 	    \caption{Extremely long Resolution Updates}
-	    \normalsize
 	    \begin{tabular}{cccc}
 	        \hline
-	        Agency & Closed Date & Resolution Action Updated Date & Post Closed Update Duration (days) \\
+	        Agency & Closed Date & Resolution Action Updated Date 
+	        & Post Closed Update Duration (days) \\
 	        \hline
@@
 
-	These obvious inaccurate date values are removed from the analysis. But even so, there are some very late
-	updates to SRs, with some updates occurring almost two years after the SR is closed.   There are a total of 7460 SRs that are 
-	updated \textgreater30 days and \textless{}730 days after the closed\_date. The median of these late (\textless{}30 days)
-	post-closed resolution\_action\_update\_date(s) is 84 days. However, the mean resolution\_action\_update\_date 
-	for all of the SRs is only 0.39 days (9.3 hours) after the SR is closed. It is not known if this is normal behavior
-	or an area that may require further investigation.
+	These obvious inaccurate date values are removed from the analysis. 
+	But even so, there are some very late updates to SRs, with some 
+	updates occurring almost two years after the SR is closed. There are a 
+	total of 7460 SRs that are updated \textgreater30 days and \textless{}730 
+	days after the closed\_date. The median of these late 
+	(\textless{}30 days) post-closed resolution\_action\_update\_date(s) is 
+	84 days. However, the mean resolution\_action\_update\_date for 
+	all of the SRs is only 0.39 days (9.3 hours) after the SR is closed. 
+	It is not known if this is normal behavior or an area that may 
+	require further investigation.
 
@@
     \centering
+    \normalsize
      \caption{Long post-closed Resolution Updates}
-    \normalsize
     \begin{tabular}{cccc}
 	        \hline
-	        Agency & Closed Date & Resolution Action Updated Date & Post Closed Update Duration (days) \\
+	        Agency & Closed Date & Resolution Action Updated Date 
+	        & Post Closed Update Duration (days) \\
 	        \hline
@@
 
-	Here is a look at the spread of the post-closed resolution\_action\_update\_date values those SRs 
-	that are \textgreater30 days and \textless{}730 days. There are some significant outliers
+	Here is a look at the spread of the post-closed resolution\_action\_update\_date 
+	values those SRs 	that are \textgreater30 days and \textless{}730 days. 
+	There are some significant outliers:
 
@@
 		\includegraphics[width = \textwidth]{post_closed_violin.pdf}
-		\caption{Post-Closed resolution\_action\_update\_date(s) \textgreater30 days}
+		\caption{Post-Closed resolution\_action\_update\_date(s) 
+		\textgreater30 days}
 		\label{fig:resolution-violin}
@@
 
-	If we look at a breakout of these (late?) update dates by Agency, we can see that the vast majority li with two
-	Agencies: Taxi \& Limo Commission (TLC) and Dept of Sanitation (DSNY) which account for 96\% of the SRs in
-	question.  Is there a valid reason that TLC and DSNY would be updating an SR over month (or months) after
+	If we look at a breakout of these (late?) update dates by Agency, 
+	we can see that the vast majority li with two Agencies: Taxi \& Limo 
+	Commission (TLC) and Dept of Sanitation (DSNY) which account 
+	for 96\% of the SRs in question.  Is there a valid reason that TLC and 
+	DSNY would be updating an SR over month (or months) after 
 	the SR has been closed? Perhaps, but it seems to merit closer investigation.
@@
 		\includegraphics[width = \textwidth]{postClosedBarChart.pdf}
-		\caption{Post-Closed resolution\_action\_update\_date(s) \textgreater30 days by Agency}
+		\caption{Post-Closed resolution\_action\_update\_date(s) 
+		\textgreater30 days by Agency}
 		\label{fig:resolution-by-agency}
@@
 \section{Accuracy and precision}\label{sec:precision}
-There is one area in particular where the question of precision vs. accuracy immediately arises, and that is with the Latitude and Longitude fields.  Both the 
-Longitude and Latitude fields are expresses as a 14-decimal number, e.g. 40.86769186022511 (also the Location field which is a straight concatenation of
-latitude and longitude). Given that 1 degree of latitude at the equator is equal to 111.044736 kilometers, the ``1'' at the end of that number represents
-approximate 1.1104 nanometer or 1/1,000,000,000 of a meter. For reference a DNA molecule is approximately 2nm in width.
-
-Clearly the representation of the Latitude and Longitude fields are a classic case of 14-digit precision, 
-but limited accuracy. It is more likely that the Lat/Long values are accurate to the 5\textsuperscript{th} decimal place, about 3.64 feet; even that might be
-overstating things.
+There is one area in particular where the question of precision vs. 
+accuracy immediately arises, and that is with the Latitude and Longitude 
+fields. Both the Longitude and Latitude fields are expresses as a 14-decimal 
+number, e.g. 40.86769186022511 (also the Location field which is a 
+straight concatenation of latitude and longitude). Given that 1 
+degree of latitude at the equator is equal to 111.044736 kilometers, 
+the ``1'' at the end of that number represents approximate 1.1104 nanometer 
+or 1/1,000,000,000 of a meter. For reference a DNA molecule is 
+approximately 2nm in width.
+
+Clearly the representation of the Latitude and Longitude fields are a 
+classic case of 14-digit precision, but limited accuracy. It is more likely 
+that the Lat/Long values are accurate to the 5\textsuperscript{th} decimal place, 
+about 3.64 feet; even that might be overstating things.
 
 \section{Redundant \& Duplicate fields}\label{sec:duplicates}
-During this analysis, several redundant fields were observed. Listed below is a discussion for each of a pair
-of data fields that are redundant or near-redundant. These fields should be examined further for possible
-consolidation in the overall dataset.
-
-\subsection{latitude \& longitude and the location fields}  The location field is a pure concatenation of the latitude and longitude fields, 
-with a comma an parenthesis added. Example:
+During this analysis, several redundant fields were observed. Listed below 
+is a discussion for each of a pair of data fields that are redundant or 
+near-redundant. These fields should be examined further for 
+possible consolidation in the overall dataset.
+
+\subsection{latitude \& longitude and the location fields}  The location field is a 
+pure concatenation of the latitude and longitude fields, with a comma and 
+parenthesis added. Example:
 
@@
 
-The inclusion of the location field seems questionable as the data is arguably more difficult to extract than the two specific fields, especially for
-software programs.
-
-\subsection{borough and park\_borough fields}  These two fields are 100\% matches; fully redundant.
-
-\subsection{borough and borough\_boundaries fields}  The borough field is a text field for the five New York City boroughs (Manhattan, Queens, Brooklyn, etc.).
-The borough\_boundary field is numeric with values from 1-5 corresponding to the five boroughs (Staten Island, Brooklyn, Queens, Manhattan, Bronx).
-When the borough\_boundary field is translated and compared to the borough field, there is a 98.3\% match. Given the size of the dataset, that 
-still leaves 110,715 non-matching values, not an insignificant number.
-
-And that, in a nutshell, is the problem with two ``near-duplicate'' fields. It is like the  story of a man wearing two watches; he's never
-quite sure what time it is. And with ``near-duplicate'' fields, which field is correct when they disagree? 
-If one field is blank, can you use the other one? Which field can be used as the reference value.
- And as we will see, there are several such examples in this dataset.
-
-We can tell where the non-matches occur. It is mostly in DOT, NYPD, and DSNY which represent 90\% of the non-matching values.
+The inclusion of the location field seems questionable as the data is 
+arguably more difficult to extract than the two specific fields, especially 
+for software programs.
+
+\subsection{borough and park\_borough fields}  These two fields are 
+100\% matches; fully redundant.
+
+\subsection{borough and borough\_boundaries fields}  The borough field 
+is a text field for the five New York City boroughs (Manhattan, Queens, 
+Brooklyn, etc.). The borough\_boundary field is numeric with values 
+from 1-5 corresponding to the five boroughs (Staten Island, Brooklyn, 
+Queens, Manhattan, Bronx). When the borough\_boundary field is 
+translated and compared to the borough field, there is a 
+98.3\% match. Given the size of the dataset, that still leaves 
+110,715 non-matching values, not an insignificant number.
+
+And that, in a nutshell, is the problem with two ``near-duplicate'' 
+fields. It is like the  story of a man wearing two watches; he's never 
+quite sure what time it is. And with ``near-duplicate'' fields, which 
+field is correct when they disagree? If one field is blank, can you 
+use the other one? Which field can be used as the reference value. And 
+as we will see, there are several such examples in this dataset.
+
+We can tell where the non-matches occur. It is mostly in DOT, NYPD, 
+and DSNY which represent 90\% of the non-matching values.
 
@@
 \subsection{borough and taxi\_company\_borough fields}
-The two fields borough and taxi\_company\_borough might appear to be related. In fact, despite the names, 
-the two data fields are nearly completely different with only 0.05\% matching. The mismatch is so
-high that it suggests the two fields are used very differently.  This field is used exclusively by the 
-Taxi \& Limo Commission which governs taxis and other cars for hire.
-There are a total of 3567 non-blank entries in the taxi\_company\_borough field, with 99.94\% blank.
+The two fields borough and taxi\_company\_borough might appear to be 
+related. In fact, despite the names, the two data fields are nearly 
+completely different with only 0.05\% matching. The mismatch is so 
+high that it suggests the two fields are used very differently. This field 
+is used exclusively by the Taxi \& Limo Commission which governs 
+taxis and other cars for hire. There are a total of 3567 non-blank 
+entries in the taxi\_company\_borough field, with 99.94\% blank.
 
  \subsection{incident\_zip and zip\_codes fields}
- The zip\_code field is one of the six computed fields and its validity has been explored in Section \ref{sec:zip-codes} and
- found to be lacking with 55.55\% of the entries invalid according to the USPS, while the companion zipcode
- field incident\_zip field had an accuracy rate of 99.93\%. It is not surprising that 
- only 0.58\% of the two fields match (37,292 matches out of 6.4 million SRs.). While an Agency breakdown can be
- determined, given the questionable computation of the zip\_code field, such an analysis would
- seem to provide any useful information.
+ The zip\_code field is one of the six computed fields and its validity 
+ has been explored in Section \ref{sec:zip-codes} and found to be 
+ lacking with 55.55\% of the entries invalid according to the 
+ USPS, while the companion zipcode field incident\_zip field had 
+ an accuracy rate of 99.93\%. It is not surprising that only 0.58\% of 
+ the two fields match (37,292 matches out of 6.4 million SRs.). While 
+ an Agency breakdown can be determined, given the questionable 
+ computation of the zip\_code field, such an analysis would seem 
+ to provide any useful information.
 
  \subsection{police\_precinct and police\_precincts fields}  
-Both the police\_precinct and police\_precincts are among the computed fields not cited in the Data Dictionary.
-Their usage however, is easy to discern.
-Are these two fields duplicates?
-Nearly so:  99.94\% of the entries match.
-Since the source of the fields and the underlying computational process are not know, it remains a challenge to determine which field is
-the ``right'' field to use. (The validity of these fields is covered in Section \ref{sec:police-precincts}.)
+Both the police\_precinct and police\_precincts are among the computed 
+fields not cited in the Data Dictionary. Their usage however, is easy to 
+discern. Are these two fields duplicates? Nearly so:  99.94\% of the entries 
+match. Since the source of the fields and the underlying computational 
+process are not know, it remains a challenge to determine which 
+field is the ``right'' field to use. (The validity of these fields is 
+covered in Section \ref{sec:police-precincts}.)
 
  \subsection{agency and agency\_name fields}
- While not duplicates, these fields have a 1:1 correspondence.  The Agency field contains abbreviations for the City agencies such as NYPD, DOT, HPD, DEP, etc.
- The agency\_name field contains the full name of the various organizations: New York Police Department, Department of Transportation, Department
- of Housing Preservation and Development, Department of Environmental Protection, and so on.  Given how well known the NYC Agency abbreviations are
-it seems redundant to include both in the dataset.
+ While not duplicates, these fields have a 1:1 correspondence.  The Agency 
+ field contains abbreviations for the City agencies such as NYPD, DOT, 
+ HPD, DEP, etc. The agency\_name field contains the full name of 
+ the various organizations: New York Police Department, Department 
+ of Transportation, Department of Housing Preservation and 
+ Development, Department of Environmental Protection, and so on. Given 
+ how well known the NYC Agency abbreviations are it seems redundant 
+ to include both in the dataset.
 
 \subsection{landmark and street\_name}
-The landmark field is listed in the Data Dictionary as ``Can refer to any noteworthy location, including but not 
-limited to, parks, hospitals, airports, sports facilities, performance spaces, etc.'' We found that was not the case. To be sure
-many of the entries in the landmark field do contain landmark names, e.g.``Pennsylvania Station'', ``La Guardia Airport'', 
-``Gramercy Park'', the vast majority of the entries are street names, e.g. ``Fenton Avenue'', ``Steinway Street'', ``Broadway''.
-These entries appear to be very similar to those in the street\_name field. And in fact, the values of street\_name
-and landmark exactly match 62\% of the time, and while this is not a full duplicate, it is a precentage of 
-matching that is certainly indicative of duplicate usage. Even the non-matches (excluding blanks) would appear
-to be matches except for minor spelling and nomenclature changes, e.g. ``NINTH AVE'' \& ``9 AVE''.
+The landmark field is listed in the Data Dictionary as ``Can refer to 
+any noteworthy location, including but not limited to, parks, 
+hospitals, airports, sports facilities, performance spaces, etc.'' We 
+found that was not the case. To be sure many of the entries in 
+the landmark field do contain landmark names, e.g.``Pennsylvania 
+Station'', ``La Guardia Airport'', ``Gramercy Park'', the vast majority 
+of the entries are street names, e.g. ``Fenton Avenue'', ``Steinway 
+Street'', ``Broadway''. These entries appear to be very similar to 
+those in the street\_name field. And in fact, the values of 
+street\_name and landmark exactly match 62\% of the time, and 
+while this is not a full duplicate, it is a precentage of matching 
+that is certainly indicative of duplicate usage. Even the non-matches 
+(excluding blanks) would appear to be matches except for minor 
+spelling and nomenclature changes, e.g. ``NINTH AVE'' \& ``9 AVE''.
 
@@
     \centering
+    \normalsize
     \caption{Non-matches between 'street\_name' and 'landmark' fields}
-    \normalsize
-    \begin{tabular}{>{\raggedright\arraybackslash}p{5cm} >{\raggedright\arraybackslash}p{5cm}}
+    \begin{tabular}{>{\raggedright\arraybackslash}p{5cm} >
+    {\raggedright\arraybackslash}p{5cm}}
         \toprule
@@
         SOUTH FIRST ST & SOUTH 1 ST \\
-        BRIGHTON SEVENTH ST & BRIGHTON 7 ST \\
         WEST FIFTH ST & WEST 5 ST \\
@@
 
-The distribution by Agency for the non-matches somewhat follows the overall SR distribution (NYPD, DPR, DSNY), 
-but with some notable exceptions such as Taxi \& Limo Commission, Dept of Homeless Services (DHS),
- and Dept of Health \& Mental Hygiene (DOHMH).
-
-\begin{figure}[tbp]
-		\centering
-		\includegraphics[width = \textwidth]{non-matchingstreet\_nameandlandmark.pdf}
+The distribution by Agency for the non-matches somewhat follows 
+the overall SR distribution (NYPD, DPR, DSNY), but with some notable 
+exceptions such as Taxi \& Limo Commission, Dept of Homeless 
+Services (DHS), and Dept of Health \& Mental Hygiene (DOHMH).
+
+\begin{figure}[tbp]
+		\centering
+		\includegraphics[width = \textwidth]
+		{non-matchingstreet\_nameandlandmark.pdf}
 		\caption{Non-matching street\_name \& landmark by Agency}
@@
 
-\subsection{cross\_street\_1 \& intersection\_street\_1 and cross\_street\_2 \& intersection\_street\_2}
-There are two sets of street pairs in the dataset in addition to the complaint address:
+\subsection{cross\_street\_1 \& intersection\_street\_1 and cross\_street\_2 
+\& intersection\_street\_2} There are two sets of street pairs in the 
+dataset in addition to the complaint address:
 
@@
 
-These two pairs of streets are used to help identify the location of the reported incident, as it is common in
-New York City to provide the street address and the cross street, as in ``24 W 90th street between Columbus and
-Amsterdam''. The Data Dictionary states that these streets are
- ``based on the geo validated incident location''. It then adds ``If the service request 
-pertains to a location in the middle of a block, cross street will be provided by geocoding...If the service request 
-pertains to a street corner, intersection will be provided by geocoding.''  One might think that the
-cross\_street(s) would apply for ``middle of a block'' incidents, while the intersection\_street(s) would
-apply to those incidents that happen on a street corner; that those two sets of fields would be
-exclusive to one another; and either/or situation.
-
-Unfortunately, that is not the case. For the majority of data rows, both pair of fields are populated and are duplicates.
-In fact, 88\% of the time, cross\_street\_1 exactly matches intersection\_street\_1 and 
-cross\_street\_2 exactly matches intersection\_street\_2, Only 11.9\% are non-matches.
-
-\textit{(Note: Address standardization was applied to the street addresses using the R \emph{campfin} package. 
-This was done so as to prevent the software from generating a non-match between 
-streets such as ``240 E 69\textsuperscript{th} ST'' and ``240 E 69\textsuperscript{th} Street'' 
-which are clearly the same address, just slight formatting and spelling differences.)}
-
-The problem occurs when the two pairs of streets do not match, such as when one is blank and the
-other is not. Which street is correct? If one is blank, say cross\_street\_1, but  intersection\_street\_1
-is not blank, should you use the non-blank one?  Let's look at some examples:
+These two pairs of streets are used to help identify the location of the 
+reported incident, as it is common in New York City to provide the 
+street address and the cross street, as in ``24 W 90th street between 
+Columbus and Amsterdam''. The Data Dictionary states that these 
+streets are ``based on the geo validated incident location''. It then 
+adds ``If the service request pertains to a location in the middle of a 
+block, cross street will be provided by geocoding...If the service 
+request pertains to a street corner, intersection will be provided by 
+geocoding.''  One might think that the cross\_street(s) would apply 
+for ``middle of a block'' incidents, while the intersection\_street(s) 
+would apply to those incidents that happen on a street corner; 
+that those two sets of fields would be exclusive to one another; 
+an either/or situation.
+
+Unfortunately, that is not the case. For the majority of data rows, both 
+pair of fields are populated and are duplicates. In fact, 88\% of the 
+time, cross\_street\_1 exactly matches intersection\_street\_1 and 
+cross\_street\_2 exactly matches intersection\_street\_2, Only 
+11.9\% are non-matches.
+
+\textit{(Note: Address standardization was applied to the street 
+addresses using the R \emph{campfin} package. This was done so as 
+to prevent the software from generating a non-match between 
+streets such as ``240 E 69\textsuperscript{th} ST'' and ``240 E 
+69\textsuperscript{th} Street'' which are clearly the same address, 
+just slight formatting and spelling differences.)}
+
+The problem occurs when the two pairs of streets do not match, such as 
+when one is blank and the other is not. Which street is correct? If one 
+is blank, say cross\_street\_1, but  intersection\_street\_1 is not blank, should 
+you use the non-blank one?  Let's look at some examples:
 
@@
     \centering
-     \caption{Matching/Non-Matching cross\_street\_1 \& intersection\_street\_1}
      \normalsize
-    \begin{tabular}{>{\normalsize\ttfamily}l >{\normalsize\ttfamily}l >{\normalsize\ttfamily}l}
+     \caption{Matching/Non-Matching cross\_street\_1 
+     \& intersection\_street\_1}
+    \begin{tabular}{>{\normalsize\ttfamily}l >
+    {\normalsize\ttfamily}l >{\normalsize\ttfamily}l}
         \toprule
-        \multicolumn{3}{c}{\textbf{Matching cross\_street\_1 and intersection\_street\_1}} \\
+        \multicolumn{3}{c}
+        {\textbf{Matching cross\_street\_1 and intersection\_street\_1}} \\
         \midrule
-        \textbf{cross\_street\_1} & \textbf{intersection\_street\_1} & \textbf{agency} \\
+        \textbf{cross\_street\_1} & \textbf{intersection\_street\_1} 
+        & \textbf{agency} \\
         \midrule
@@
         \midrule
-        \multicolumn{3}{c}{\textbf{Non-matching cross\_street\_1 and intersection\_street\_1}} \\
+        \multicolumn{3}{c}
+        {\textbf{Non-matching cross\_street\_1 and intersection\_street\_1}} \\
         \midrule
-        \textbf{cross\_street\_1} & \textbf{intersection\_street\_1} & \textbf{agency} \\
+        \textbf{cross\_street\_1} & \textbf{intersection\_street\_1} 
+        & \textbf{agency} \\
         \midrule
@@
 		\centering
-		\includegraphics[width = \textwidth]{non-matchingcross_street_1andintersection_street_1.pdf}
+		\includegraphics[width = \textwidth]
+		{non-matchingcross_street_1andintersection_street_1.pdf}
 		\caption{Non-matching X\_Street\_1 \& Intersection\_Street\_1}
@@
 		\centering
-		\includegraphics[width = \textwidth]{non-matchingcross_street_2andintersection_street_2.pdf}
+		\includegraphics[width = \textwidth]
+		{non-matchingcross_street_2andintersection_street_2.pdf}
 		\caption{Non-matching X\_Street\_2 \& Intersection\_Street\_2}
@@
 
-In addition to the matching/non-matching values, we encountered a number of what you might call
-\emph{near-matches}. There were numerous cases where the cross\_street and intersection\_street
-would be close to identical, e.g. HARMON DR and HARMON RD. To identify these near-matches
-we applied a common method of determining matches by measuring 
-the \href{https://en.wikipedia.org/wiki/Hamming_distance}{Hamming Distance} between
-the two fields. The Hamming Distance is a measure of how many letters have to be changed
-in order for the two fields to match. For this analysis, we chose a Hamming Distance of 2.
-Here are the results for both cross\_street and intersection\_street.
+In addition to the matching/non-matching values, we encountered a number 
+of what you might call \emph{near-matches}. There were numerous cases 
+where the cross\_street and intersection\_street would be close to 
+identical, e.g. HARMON DR and HARMON RD. To identify these 
+near-matches we applied a common method of determining matches 
+by measuring  the \href{https://en.wikipedia.org/wiki/Hamming_distance}
+{Hamming Distance} between the two fields. The Hamming Distance is a 
+measure of how many letters have to be changed in order for the two 
+fields to match. For this analysis, we chose a Hamming Distance of 2. Here 
+are the results for both cross\_street and intersection\_street.
 
@@
     \centering
+     \normalsize
      \caption{Near-matches for cross\_street\_1, intersection\_street\_1}
-    \normalsize
 	\begin{tabular}{l l l r}
         \toprule
-        \textbf{cross\_street\_1} & \textbf{intersection\_street\_1} & \textbf{agency} & \textbf{hamming\_distance} \\
+        \textbf{cross\_street\_1} & \textbf{intersection\_street\_1} 
+        & \textbf{agency} & \textbf{hamming\_distance} \\
         \midrule
@@
     \centering
+    \normalsize
     \caption{Summary of cross\_street\_1 and intersection\_street\_1}
-    \normalsize
     \begin{tabular}{l r r}
@@
     \centering
+    \normalsize
     \caption{Summary of cross\_street\_2 and intersection\_street\_2}
-    \normalsize
     \begin{tabular}{l r r}
@@
     \centering
-     \caption{Sample of near-matching cross\_street\_2 and intersection\_street\_2 (both non-blank)}
      \normalsize
+     \caption{Sample of near-matching cross\_street\_2 and 
+     intersection\_street\_2 (both non-blank)}
      \begin{tabular}{l l l r}
         \toprule
-        \textbf{cross\_street\_2} & \textbf{intersection\_street\_2} & \textbf{agency} & \textbf{hamming\_distance} \\
+        \textbf{cross\_street\_2} & \textbf{intersection\_street\_2} 
+        & \textbf{agency} & \textbf{hamming\_distance} \\
         \midrule
@@
 
- By removing duplicate and ``near-duplicate'' fields, it is possible to shrink the file size by 12.3\%
- which for this dataset equates to a reduction of 395 Mb. A smaller file size means faster downloads,
- less storage impact, as well as simplifying data analysis efforts.
-
- Below is a list of duplicate and near-duplicate fields. There are some challenges with these proposed deletions, 
- mostly with the ``near duplicate'' fields.  So while the park\_borough is a complete duplicate of the borough
- field, the borough\_boundary field is only 98.3\% duplicate. Admittedly, loosing 1.7\% of the data in those two
- fields is quite small, for this large dataset that amounts to 110, 715 non-matching occurrences. Similarly with 
- police\_precincts and police\_precincts where there is a 99.9\% match; close to 100\% but not a complete match.
-
-Ultimately, it will require a deeper dive into the relationship between these duplicate and near-duplicate fields to
- determine if they can be removed or not. Pending that, the authors would propose deleting the following fields
- from the overall data set, which explanation provided.
+ By removing duplicate and ``near-duplicate'' fields, it is possible to 
+ shrink the file size by 12.3\% which for this dataset equates to a 
+ reduction of 395 Mb. A smaller file size means faster downloads, less 
+ storage impact, as well as simplifying data analysis efforts.
+
+ Below is a list of duplicate and near-duplicate fields. There are some 
+ challenges with these proposed deletions, mostly with the ``near 
+ duplicate'' fields.  So while the park\_borough is a complete duplicate 
+ of the borough field, the borough\_boundary field is only 98.3\% 
+ duplicate. Admittedly, loosing 1.7\% of the data in those two fields 
+ is quite small, for this large dataset that amounts to 110, 715 
+ non-matching occurrences. Similarly with police\_precincts 
+ and police\_precincts where there is a 99.9\% match; close 
+ to 100\% but not a complete match.
+
+Ultimately, it will require a deeper dive into the relationship 
+between these duplicate and near-duplicate fields to determine 
+if they can be removed or not. Pending that, the authors would propose 
+deleting the following fields from the overall data set, 
+which explanation provided.
 
  	\begin{itemize}
-		    \item agency\_name: Each row of data contains the agency field, which is an abbreviation of the Agency's name.
-		    The agency\_name abbreviations are clear, simple, and well understood. The authors believe the agency\_name field could
-		    be eliminated without loss of data quality.
-
-		    \item park\_borough:  This field is a 100\% match with the borough field. Removal of this field will not result in any loss
-		    of data or data quality.
-
-		    \item location:  The location field is simply a concatenation of the latitude and longitude fields, with a comma and 
-		    parenthesis added. The authors feel this field actually hinders data analysis as the
-		    field must be re-split into the two components (latitude and longitude) to conduct geographic analysis.
-		     The data is a 100\% duplication of the latitude and longitude fields. Removal of this field will
-		    not result in any loss of data.
-
-		    \item police\_precinct: This field has a 99.95\% match with the police\_precincts fields. We are unable to determine which
-		    field is more correct, but feel that removal of one of the precinct fields is prudent and would result in very minimal 
-		    data loss. (We should note that both police\_precinct and police\_precincts fields each have 34.5\% invalid data.)
-
-		   \item borough\_boundaries (computed field): This field is one of the six computed fields. It has a 98.3 match with the
-		    borough field. We recommend deleting this fields despite the somewhat limited loss of data.
-
-		    \item cross\_street\_1 \& 2 and intersection\_street\_1 \& 2: These two pairs of fields each have an 88\% match. We would
-		    recommend deleting the two intersection\_street fields while acknowledging some loss of data in doing so.
-
-		    \item zip\_codes(computed field):  This field is one of the six computed fields.  However, this field has error rate of
-		    58\% and as shown in the \ref{sec:case-study-zip-codes} can lead to dangerous mistakes when used for analytical purposes.
-		    We would recommend deleting this field.
+		    \item agency\_name: Each row of data contains the agency field, 
+		    which is an abbreviation of the Agency's name. The agency\_name 
+		    abbreviations are clear, simple, and well understood. The 
+		    authors believe the agency\_name field could be eliminated 
+		    without loss of data quality.
+
+		    \item park\_borough:  This field is a 100\% match with the 
+		    borough field. Removal of this field will not result in any loss of 
+		    data or data quality.
+
+		    \item location:  The location field is simply a concatenation of 
+		    the latitude and longitude fields, with a comma and parenthesis 
+		    added. The authors feel this field actually hinders data analysis 
+		    as the field must be re-split into the two components 
+		    (latitude and longitude) to conduct geographic analysis. The data is 
+		    a 100\% duplication of the latitude and longitude fields. Removal 
+		    of this field will not result in any loss of data.
+
+		    \item police\_precinct: This field has a 99.95\% match with the 
+		    police\_precincts fields. We are unable to determine which field 
+		    is more correct, but feel that removal of one of the precinct fields 
+		    is prudent and would result in very minimal data loss. (We 
+		    should note that both police\_precinct and police\_precincts 
+		    fields each have 34.5\% invalid data.)
+
+		   \item borough\_boundaries (computed field): This field is one of 
+		   the six computed fields. It has a 98.3 match with the borough 
+		   field. We recommend deleting this fields despite the somewhat 
+		   limited loss of data.
+
+		    \item cross\_street\_1 \& 2 and intersection\_street\_1 \& 2: These 
+		    two pairs of fields each have an 88\% match. We would recommend 
+		    deleting the two intersection\_street fields while acknowledging 
+		    some loss of data in doing so.
+
+		    \item zip\_codes(computed field):  This field is one of the six 
+		    computed fields.  However, this field has error rate of 58\% and 
+		    as shown in the \ref{sec:case-study-zip-codes} can lead to 
+		    dangerous mistakes when used for analytical purposes. We would 
+		    recommend deleting this field.
 	\end{itemize}
 
-Additionally, it would be worth investigating if certain data fields are in-fact useful, as the population of these fields is very, very 
-scarce. For example, the taxi\_company\_borough field is 99.94\% blank, and the remaining entries match the borough field only
-0.05\% field. A similar situation exists for the road\_ramp field (99.75\% blank), the vehicle\_type field (99.71\%blank), and possibly 
-due\_date which is 99.62\% blank. There are several others: bridge\_highway\_direction, bridge\_highway\_name,
-bridge\_highway\_segment, and taxi\_pick\_up\_location...all of which have \textgreater99\% blanks. Most of those fields appear to be 
-almost exclusively used by the Taxi \& Limo Commission (TLC), which should be consulted to 
-determine if those fields are used and properly populated.
+Additionally, it would be worth investigating if certain data fields are 
+in-fact useful, as the population of these fields is very, very scarce. For 
+example, the taxi\_company\_borough field is 99.94\% blank, and the 
+remaining entries match the borough field only 0.05\% field. A similar 
+situation exists for the road\_ramp field (99.75\% blank), the 
+vehicle\_type field (99.71\%blank), and possibly due\_date which 
+is 99.62\% blank. There are several others: bridge\_highway\_direction, 
+bridge\_highway\_name, bridge\_highway\_segment, and 
+taxi\_pick\_up\_location...all of which have \textgreater99\% 
+blanks. Most of those fields appear to be almost exclusively used by 
+the Taxi \& Limo Commission (TLC), which should be consulted 
+to determine if those fields are used and properly populated.
 
 \section{Data Dictionary Observations} \label{sec:datadictionay}
-During this analysis, it became apparent that the published Data Dictionary could use an update. We found many discrepancies between
-the Data Dictionary and the actual data. While not actually, incorrect or dirty data, it nonetheless can lead analytic efforts astray.
-In data analysis, it's essential to accurately describe the nature and purpose of each field to ensure proper handling and interpretation. 
-Here are some of the noted discrepancies:
+During this analysis, it became apparent that the published Data 
+Dictionary could use an update. We found many discrepancies 
+between the Data Dictionary and the actual data. While not actually, 
+incorrect or dirty data, it nonetheless can lead analytic efforts astray. In 
+data analysis, it's essential to accurately describe the nature and purpose 
+of each field to ensure proper handling and interpretation. Here 
+are some of the noted discrepancies:
 
 \begin{itemize}
-	\item While the Open Data portal lists the basic data types of the various fields, the Data Dictionary does not. Nor
-	does it provide meaningful information on the specifics of various fields, such as constraints or domain of legal values.
-	For example, the incident\_zip field is specified as ``text''. That's probably not the best definition, perhaps
-	``categorical'' would be more appropriate with a description indicating that the data can contain only numeric characters
-	and is not subjected to arithmetic operations. We found two instances with the zip\_code field contained character strings (``na'', ``N/A'').
-
-	\item Domain of legal values as contained in the Data Dictionary are incomplete. For example, the Data Dictionary indicates
-	that the status field has values of assigned, canceled, closed, or pending. However, we found additional status values:
-	in progress, started, and unspecified. Additionally, we discovered no SRs with a status of canceled. Similarly, the 
-	address\_type field is indicated to have values address, blockface, intersection, latlong, and placename. We discovered
-	additional values of bbl and unrecognized. No values of latlong were observed. Other fields suffer the same
-	inaccuracies (facility\_type, vehicle\_type, taxi\_pick\_up\_location, road\_ramp, city).
+	\item While the Open Data portal lists the basic data types of the 
+	various fields, the Data Dictionary does not. Nor does it provide 
+	meaningful information on the specifics of various fields, such 
+	as constraints or domain of legal values. For example, the 
+	incident\_zip field is specified as ``text''. That's probably not the 
+	best definition, perhaps	``categorical'' would be more 
+	appropriate with a description indicating that the data can 
+	contain only numeric characters and is not subjected to arithmetic 
+	operations. We found two instances with the zip\_code field 
+	contained character strings (``na'', ``N/A'').
+
+	\item Domain of legal values as contained in the Data Dictionary 
+	are incomplete. For example, the Data Dictionary indicates that 
+	the status field has values of assigned, canceled, closed, or pending. 
+	However, we found additional status values: in progress, started, 
+	and unspecified. Additionally, we discovered no SRs with a status of 
+	canceled. Similarly, the address\_type field is indicated to have 
+	values address, blockface, intersection, latlong, and placename. We 
+	discovered additional values of bbl and unrecognized. No values of 
+	latlong were observed. Other fields suffer the same 	inaccuracies 
+	(facility\_type, vehicle\_type, taxi\_pick\_up\_location, road\_ramp, city).
 \end{itemize}
@@
 \section{Recommendations} \label{sec:recommendations}
-Having identified a number of data cleanliness related issues with the 311 Service Request (SR) dataset, what follows
-are a series of recommendations as to how to resolve these issues. Below are those recommendations;
+Having identified a number of data cleanliness related issues with the 
+311 Service Request (SR) dataset, what follows are a series of 
+recommendations as to how to resolve these issues. Below are 
+those recommendations;
 
 \begin{itemize}
-	\item First, and most importantly, the Open Data Team will need to undertake a special effort to identify these data anomalies,
-	confirm the veracity of the issue, and engage with the appropriate Agency Open Data Representatives to resolve these issues.
-	That effort will not be easy, fast, or cheap as in some software changes will likely be required. Perhaps hardest will be
-	working with the external City Agencies that feed data to the 311 SR dataset, but are not using the core 311 system.
-	Integrations and API usage will need to be investigated, and in many cases the changes required to correct an issue will
-	be on the source Agency side. Coordination and mutual approval of any changes will be key. Strong support from senior
-	leadership will be a critical success factor.
-
-	\item Secondly, standards will be to be identified, agreed to, and commitment to support those standards must be
-	obtained. For example, what are the domain of legal values for certain fields, especially those with identified issues?
-
-	\item Obviously, many of the data integrity issues can be resolved through rule enforcement by software. 
-	For example, the software could easily prevent invalid values, such as zip codes, 
-	from entering the system. Similarly for an entire range of data issues. One key area is the
-	created\_date and closed\_date which are frequently cited when determining 
-	responsiveness of various Agency services. The authors would argue that such dates 
-	should be driven by change in ``status'', not manually entered. For example, the closed\_date is captured
-	when the SR ``status'' is changed to ``closed'', and not manually entered. Such a change, along with many others, could prevent errors such
-	nras negative and  zero durations which significantly affect measurements of responsiveness.
-
-	\item The large spikes in SR closure and creation occurring at midnight and noon, almost certainly point to an underlying automated
-	process from an external-to-311 Agency. Such processes again distort the accurate capture of an SR duration. Data rule enforcement
-	at the software integration or API level should be investigated.
-
-	\item One possible way to have software enforce certain data standards, would be to move more non-311 Agencies to the core 311
-	software system. That system was completely upgraded in 2019, and should be (relatively) easy to expand to other Agencies and usages.
-
-	\item As noted, there are six ``computed'' fields in the data export that are not identified in the Data Dictionary. While these fields may be
-	``experimental'', they are nonetheless present in any data portal presentation or export. As noted in this study, a number of those computed
-	fields suffer from data inaccuracies. It is difficult to know what to recommend for these six field without further understanding
-	of their origin, computational method, or intended usage. We would recommend updating the Data Dictionary at a minimum and
-	possibly hiding those fields from public display until such time as their usage and accuracy can be ascertained.
-
-	\item Removal of duplicate fields should be examined.  In some cases, such a the borough and park\_borough fields, the duplication is 100%. 
-	In other cases, the duplication is in the 80-90\% range. These fields and their usage/utility should be carefully examined to determine if their
-	inclusion in the 311 SR dataset is meaningful, adds value not found in other fields, and should be removed or continued to be collected.
-	Initial analysis indicates that as much as 12\% size reductions are easily achievable with the removal of duplicate fields. A more thorough
-	review and understanding of the fields would no doubt yield more savings.
-
-	\item Fields with extremely low data presence, such as taxi\_company\_borough which is populated only 0.06\% of the time. Similarly for several other fields 
-	evaluated in this study which have a high percentage of blank/unknown values. If the presence of a small percentage of SRs 
-	with these fields being populated is deemed useful and necessary, then so be it. But if the usage of these fields is so diminished, 
-	then perhaps its inclusion in the 311 SR dataset is not warranted.
-
-	\item The Data Dictionary is in need of an upgrade and update. The last revision is indicated at 6 March 2023, well over a year ago. 
-	As noted in \textbackslash nameref: \nameref{sec:datadictionary} there are many problems with the Data Dictionary, including domain
-	values, absence of the six ``computed'' fields, and inaccurate description of the usage of several fields. Some of the Data Dictionary
-	entries can be undertaken immediately, which further revision should be addressed as part of a deeper data clean-up initiative.
-
-	\item The excessive precision in the latitude, longitude, and location fields is simply nonsensical.  We would recommend no more than four
-	decimal places (0.000X) which represents a distance of approximately 36.4 feet. Five decimals (0.0000X) represents 3.6 feet, which 
-	although possible, would be an extremely sophisticated and accurate geo-coding system. A 14 decimal number is simply a case
-	of a lot of precision with limited accuracy, and should be avoided.
-
-	\item The 10 address-related address fields (incident\_address, street\_name, city, intersection\_street\_1 \& 2, cross\_street\_1 \& 2,  landmark,
-	road\_ramp, bridge\_highway\_segment, and taxi\_pick\_up\_location should be in standard USPS format. 
-	There are several software packages which can standardize addresses to these standards.
-	We believe such an effort should be automatically applied to these fields at data entry. This step would also greatly improve any underlying geo-coding routines which
-	may be used to determine data values such as BBL, latitude, longitude, and New York State X/Y-plane coordinates.
-
-	\item The use of the value ``UNKNOWN'' , ``N/A'', ``NA'', and ``na'' when entered directly as text into a field should be very carefully examined, and
-	in all likelihood prohibited.  There are some indeed a few complaint\_type(s) where an ``N/A'' value is appropriate, such as a consumer\_complaint
-	about certain non-geo centric issues, such as a travel agency or tax preparation service . But even in those cases a value of ``N/A'' should be 
-	entered only via a drop-down menu selection. During this analysis, there were far too many data representations of missing data including: <N/A>, ``na'', `` ``,
-	UNKNOWN, and ``N/A''. It's challenging for analyst to have to account for so the many different methods of representing empty or null data fields.
+	\item First, and most importantly, the Open Data Team will need 
+	to undertake a special effort to identify these data anomalies, confirm 
+	the veracity of the issue, and engage with the appropriate Agency 
+	Open Data Representatives to resolve these issues. That effort will 
+	not be easy, fast, or cheap as in some software changes will likely 
+	be required. Perhaps hardest will be working with the external City 
+	Agencies that feed data to the 311 SR dataset, but are not using 
+	the core 311 system. Integrations and API usage will need to be 
+	investigated, and in many cases the changes required to correct 
+	an issue will be on the source Agency side. Coordination and 
+	mutual approval of any changes will be key. Strong support from 
+	senior 	leadership will be a critical success factor.
+
+	\item Secondly, standards will be to be identified, agreed to, and 
+	commitment to support those standards must be obtained. For 
+	example, what are the domain of legal values for certain fields, 
+	especially those with identified issues?
+
+	\item Obviously, many of the data integrity issues can be 
+	resolved through rule enforcement by software. For example, the 
+	software could easily prevent invalid values, such as zip codes, from 
+	entering the system. Similarly for an entire range of data issues. One 
+	key area is the created\_date and closed\_date which are frequently 
+	cited when determining responsiveness of various Agency services. The 
+	authors would argue that such dates should be driven by change 
+	in ``status'', not manually entered. For example, the closed\_date 
+	is captured when the SR ``status'' is changed to ``closed'', and 
+	not manually entered. Such a change, along with many others, could 
+	prevent errors such as negative and  zero durations which 
+	significantly affect measurements of responsiveness.
+
+	\item The large spikes in SR closure and creation occurring at 
+	midnight and noon, almost certainly point to an underlying automated 
+	process from an external-to-311 Agency. Such processes again distort 
+	the accurate capture of an SR duration. Data rule enforcement at the 
+	software integration or API level should be investigated.
+
+	\item One possible way to have software enforce certain data 
+	standards, would be to move more non-311 Agencies to the core 
+	311 software system. That system was completely upgraded in 
+	2019, and should be (relatively) easy to expand to other 
+	Agencies and usages.
+
+	\item As noted, there are six ``computed'' fields in the data export 
+	that are not identified in the Data Dictionary. While these fields may 
+	be ``experimental'', they are nonetheless present in any data portal 
+	presentation or export. As noted in this study, a number of those 
+	computed fields suffer from data inaccuracies. It is difficult to know 
+	what to recommend for these six field without further understanding of 
+	their origin, computational method, or intended usage. We would 
+	recommend updating the Data Dictionary at a minimum and 
+	possibly hiding those fields from public display until such time as 
+	their usage and accuracy can be ascertained.
+
+	\item Removal of duplicate fields should be examined.  In some cases, such 
+	as the borough and park\_borough fields, the duplication is 100\%. In 
+	other cases, the duplication is in the 80-90\% range. These fields and 
+	their usage/utility should be carefully examined to determine if 
+	their inclusion in the 311 SR dataset is meaningful, adds value not found 
+	in other fields, and should be removed or continued to be collected. Initial 
+	analysis indicates that as much as 12\% size reductions are achievable 
+	with the removal of duplicate fields. A more thorough review and 
+	understanding of the fields would no doubt yield more savings.
+
+	\item Fields with extremely low data presence, such as taxi\_company\_borough 
+	which is populated only 0.06\% of the time. Similarly for several other 
+	fields evaluated in this study which have a high percentage of blank/unknown 
+	values. If the presence of a small percentage of SRs with these fields 
+	being populated is deemed useful and necessary, then so be it. But if 
+	the usage of these fields is so diminished, then perhaps its inclusion 
+	in the 311 SR dataset is not warranted.
+
+	\item The Data Dictionary is in need of an upgrade and update. The last 
+	revision is indicated at 6 March 2023, well over a year ago. 	As noted 
+	in \textbackslash nameref: \nameref{sec:datadictionary} there are many 
+	problems with the Data Dictionary, including domain values, absence of 
+	the six ``computed'' fields, and inaccurate description of the usage of 
+	several fields. Some of the Data Dictionary entries can be undertaken 
+	immediately, which further revision should be addressed as part of 
+	a deeper data clean-up initiative.
+
+	\item The excessive precision in the latitude, longitude, and location 
+	fields is simply nonsensical.  We would recommend no more than 
+	four decimal places (0.000X) which represents a distance of approximately 
+	36.4 feet. Five decimals (0.0000X) represents 3.6 feet, which although 
+	possible, would be an extremely sophisticated and accurate 
+	geo-coding system. A 14 decimal number is simply a case of a lot 
+	of precision with limited accuracy, and should be avoided.
+
+	\item The 10 address-related address fields (incident\_address, 
+	street\_name, city, intersection\_street\_1 \& 2, cross\_street\_1 \& 2,  
+	landmark, road\_ramp, bridge\_highway\_segment, and 
+	taxi\_pick\_up\_location should be in standard USPS format. There 
+	are several software packages which can standardize addresses to 
+	these standards. We believe such an effort should be automatically 
+	applied to these fields at data entry. This step would also greatly 
+	improve any underlying geo-coding routines which 	may be used 
+	to determine data values such as BBL, latitude, longitude, and New 
+	York State X/Y-plane coordinates.
+
+	\item The use of the value ``UNKNOWN'' , ``N/A'', ``NA'', and ``na'' when 
+	entered directly as text into a field should be very carefully examined, 
+	and in all likelihood prohibited.  There are some indeed a few 
+	complaint\_type(s) where an ``N/A'' value is appropriate, such 
+	as a consumer\_complaint about certain non-geo centric issues, such 
+	as a travel agency or tax preparation service . But even in those cases 
+	a value of ``N/A'' should be entered only via a drop-down menu 
+	selection. During this analysis, there were far too many data 
+	representations of missing data including: <N/A>, ``na'', `` ``, UNKNOWN, 
+	and ``N/A''. It's challenging for analyst to have to account for so 
+	the many different methods of representing empty or null data fields.
 

### 2024-06-14 manuscript/manuscript.tex
Cleanup on line spacing on manuscript.tex

@@
 \begin{abstract}
-  In the early 21st century, the open data movement began to transform societies and governments through principles of transparency, 
-  innovation, and public engagement. New York City (NYC) has emerged as a leader in this movement 
-  with the enactment of the Open Data Law in 2012, leading to the creation of the NYC Open Data portal, 
-  which now hosts 2700 datasets from 80 city agencies. This resource has proven invaluable for research 
-  across various domains, including health, urban development, and transportation. 
-  The success of these initiatives underscores the importance of data curation, ensuring the utility and reliability of datasets.
-
-This paper examines the data curation challenges using the NYC 311 Service Request (SR) Data as a case study, 
-  addressing issues of data validity, consistency, and curation efficiency. Based on insights from this case study, 
-  we propose a set of data curation principles tailored for government-released open data. 
-  These principles aim to enhance data management practices and ensure the ongoing utility of open data. 
-  The paper concludes with actionable suggestions for improving data curation 
-  and offers general principles for the release of open data.
+  In the early 21st century, the open data movement began to transform societies 
+  and governments through principles of transparency, innovation, and 
+  public engagement. New York City (NYC) has emerged as a leader in this 
+  movement with the enactment of the Open Data Law in 2012, 
+  leading to the creation of the NYC Open Data portal, which now hosts 
+  2700 datasets from 80 city agencies. This resource has proven invaluable 
+  for research across various domains, including health, 
+  urban development, and transportation. The success of these initiatives 
+  underscores the importance of data curation, ensuring 
+  the utility and reliability of datasets.
+
+This paper examines the data curation challenges using the 
+NYC 311 Service Request (SR) Data as a case study, addressing issues 
+of data validity, consistency, and curation efficiency. Based on 
+insights from this case study, we propose a set of data curation 
+principles tailored for government-released open data. These principles 
+aim to enhance data management practices and ensure 
+the ongoing utility of open data. The paper concludes with 
+actionable suggestions for improving data curation and offers 
+general principles for the release of open data.
 
@@
 \href{https://data.worldbank.org}{World Bank Open Data}. 
-These initiatives represent significant strides in democratizing data, in breaking barriers that once
-kept valuable information on government performance in silos. Their collective impact is profound, extending beyond
-mere data sharing to fostering a culture of openness that benefits
-individuals, communities, governments, and economies worldwide
+These initiatives represent significant strides in democratizing data, 
+in breaking barriers that once kept valuable information 
+on government performance in silos. Their collective impact 
+is profound, extending beyond mere data sharing to 
+fostering a culture of openness that benefits individuals, 
+communities, governments, and economies worldwide 
 \citep{barns2016mine, wang2016adoption}.
@@
 The paper concludes with a discussion and recommendations
-in Section~\ref{sec:discussion}, encapsulating the key insights and implications of our research.
-
-A note on naming conventions: dataset field names are typically named using ``snake\_case'' where
-each word in the variable name is written in lowercase, and words are separated by underscores,
-e.g created\_date, complaint\_type, zip\_code, etc..
+in Section~\ref{sec:discussion}, encapsulating the key 
+insights and implications of our research.
+
+A note on naming conventions: dataset field names are typically 
+named using ``snake\_case'' where each word in the 
+variable name is written in lowercase, and 
+words are separated by underscores, e.g 
+created\_date, complaint\_type, zip\_code, etc..
 
@@
  	\item 2003 - NYC 311 system goes live with a phone-based call center only. 
- 	The ``311'' phone number replaces a myriad of numbers to call for each individual City Agency.
+ 	The ``311'' phone number replaces a myriad of numbers to 
+ 	call for each individual City Agency.
    	\item 2009 - 311 online \& mobile apps launched. Today 60\% of all 
@@
    	\item 2019 - Major software upgrade of 311 system with enhanced capabilities, 
-   	dynamic load-balancing, and cloud-based resilience. Only major software upgrade since 2004.	
+   	dynamic load-balancing, and cloud-based resilience. Only major 
+   	software upgrade since 2004.	
       \item 2020 - In August 2020, driven by the COVID pandemic, monthly 311 
@@
 which provides the technology support for the open data system.  
-Many of the underlying datasets are updated daily, usually running 1-2 days behind from creation.
-However, a number of datasets are updated monthly (such as restaurant inspections), and some are
-not updated for one or more years (such as community board boundaries). This poses problems when 
-merging dataset with others, as would be expected. Typically, the data must be
-exported and manipulated in another, external program. The authors found 
+Many of the underlying datasets are updated daily, usually running 
+1-2 days behind from creation. However, a number of datasets 
+are updated monthly (such as restaurant inspections), and some are 
+not updated for one or more years (such as community board 
+boundaries). This poses problems when merging datasets with 
+others, as would be expected. Typically, the data must be exported 
+and manipulated in another, external program. The authors found 
 this to be the case with the large 311 dataset employed.
 
-Data accuracy and consistency is typically established between OTI and the other contributing
-City Agencies, such as NYPD, Housing Preservation and Development (HPD), Parks \& Recreation, etc.,
-via Service Level Agreements (SLAs). However, enforcement of those SLAs is often challenging. The 
-result is that often data accuracy and logical consistency is occasionally compromised. 
-However, the Open Data Team is quite responsive to inquires from users. The authors have provided many such suggestions during the course of this research effort, 
+Data accuracy and consistency is typically established between OTI 
+and the other contributing City Agencies, such as NYPD, Housing 
+Preservation and Development (HPD), Parks \& Recreation, etc., via 
+Service Level Agreements (SLAs). However, enforcement of those 
+SLAs is often challenging. The result is that often data accuracy and 
+logical consistency is occasionally compromised. However, 
+the Open Data Team is quite responsive to inquires from users. The 
+authors have provided many such suggestions during the course of this research effort, 
 and the most current data shows improvements in many of the areas highlighted. 
@@
 	applied to these fields at data entry. This step would also greatly 
-	improve any underlying geo-coding routines which 	may be used 
+	improve any underlying geocoding routines which 	may be used 
 	to determine data values such as BBL, latitude, longitude, and New 
@@
 	complaint\_type(s) where an ``N/A'' value is appropriate, such 
-	as a consumer\_complaint about certain non-geo centric issues, such 
+	as a consumer\_complaint about certain non-geocentric issues, such 
 	as a travel agency or tax preparation service . But even in those cases 

### 2024-06-14 code/datacleansing.R

@@ -141,14 +141,14 @@ create_bar_chart <- function(
 
   if (add_sd) {
     bar_chart <- bar_chart +
-      # geom_hline(yintercept = mean_value + 1*sd_value, linetype = "longdash", color = "#gray20", linewidth = 0.5) +
-      # geom_hline(yintercept = mean_value - 1*sd_value, linetype = "longdash", color = "#gray20", linewidth = 0.5) +
-      geom_hline(yintercept = mean_value + 2 * sd_value, linetype = "longdash", color = "#gray20", linewidth = 0.4) +
-      geom_hline(yintercept = mean_value - 2 * sd_value, linetype = "longdash", color = "#gray20", linewidth = 0.4) +
-      # annotate("text", x = min(dataset[[x_col]]), y = mean_value - 1*sd_value, label = "-1 SD", size = 4, color = "#gray20", hjust = -0.5, vjust = -0.75) +
-      # annotate("text", x = min(dataset[[x_col]]), y = mean_value + 1*sd_value, label = "+1 SD", size = 4, color = "#gray20", hjust = -0.5, vjust = -0.75) +
-      annotate("text", x = min(dataset[[x_col]]), y = mean_value + 2 * sd_value, label = "+2 SD", size = 4, color = "#gray20", hjust = -0.5, vjust = -0.75) +
-      annotate("text", x = min(dataset[[x_col]]), y = mean_value - 2 * sd_value, label = "-2 SD", size = 4, color = "#gray20", hjust = -0.5, vjust = -0.75)
+      # geom_hline(yintercept = mean_value + 1*sd_value, linetype = "longdash", color = "gray20", linewidth = 0.5) +
+      # geom_hline(yintercept = mean_value - 1*sd_value, linetype = "longdash", color = "gray20", linewidth = 0.5) +
+      geom_hline(yintercept = mean_value + 2 * sd_value, linetype = "longdash", color = "gray20", linewidth = 0.4) +
+      geom_hline(yintercept = mean_value - 2 * sd_value, linetype = "longdash", color = "gray20", linewidth = 0.4) +
+      # annotate("text", x = min(dataset[[x_col]]), y = mean_value - 1*sd_value, label = "-1 SD", size = 4, color = "gray20", hjust = -0.5, vjust = -0.75) +
[…]

### 2024-06-17 manuscript/manuscript.tex
removed backslash in figure file names in includegraphics
Clean up of spaces and indentations for consistent presentation

@@
 
-\hyphenpenalty=1000
+\hyphenpenalty=750
 
 \begin{abstract}
-  In the early 21st century, the open data movement began to transform societies 
-  and governments through principles of transparency, innovation, and 
-  public engagement. New York City (NYC) has emerged as a leader in this 
-  movement with the enactment of the Open Data Law in 2012, 
-  leading to the creation of the NYC Open Data portal, which now hosts 
-  2700 datasets from 80 city agencies. This resource has proven invaluable 
-  for research across various domains, including health, 
-  urban development, and transportation. The success of these initiatives 
-  underscores the importance of data curation, ensuring 
-  the utility and reliability of datasets.
+In the early 21st century, the open data movement began to transform 
+societies and governments through principles of transparency, 
+innovation, and public engagement. New York City (NYC) has emerged 
+as a leader in this movement with the enactment of the Open 
+Data Law in 2012, leading to the creation of the NYC Open Data 
+portal, which now hosts 2700 datasets from 80 city agencies. This 
+resource has proven invaluable for research across various domains, 
+including health, urban development, and transportation. The 
+success of these initiatives underscores the importance of data 
+curation, ensuring the utility and reliability of datasets.
 
@@
 \section{Introduction} \label{sec:intro}
-
 In the early 21st century, the open data movement began to take shape, driven by the 
@@
 identifies duplicate and redundant data fields. Section~\ref{sec:improvements} provides 
-actionable suggestions for mitigating or resolving identified issues. Following this, Section~\ref{sec:protocol} outlines a series of general 
-principles for the release of open data, drawing from our findings. 
-The paper concludes with a discussion and recommendations
-in Section~\ref{sec:discussion}, encapsulating the key 
-insights and implications of our research.
+actionable suggestions for mitigating or resolving identified issues. 
+Following this, Section~\ref{sec:protocol} outlines a series of general 
+principles for the release of open data, drawing from our findings. The 
+paper concludes with a discussion and recommendations in 
+Section~\ref{sec:discussion}, encapsulating the key insights and 
+implications of our research.
 
 A note on naming conventions: dataset field names are typically 
-named using ``snake\_case'' where each word in the 
-variable name is written in lowercase, and 
-words are separated by underscores, e.g 
-created\_date, complaint\_type, zip\_code, etc..
+named using ``snake\_case'' where each word in the variable name 
+is written in lowercase, and words are separated by 
+underscores, e.g. created\_date, complaint\_type, zip\_code, etc..
 
 \section{NYC 311 System History} \label{sec:history}
-
 The NYC 311 service, a critical component of New York City's public
@@
 NYC Open Data Portal which provides access
-to all the data sets (approx. 2700) and provides a convenient method for querying, grouping, aggregating, geo-mapping, visualization, and exporting
-results. This open data initiative enables not only governmental transparency but also empowers 
-researchers, civic developers, and the general public.
+to all the data sets (approx. 2700) and provides a convenient method for querying, 
+grouping, aggregating, geo-mapping, visualization, and exporting 
+results. This open data initiative enables not only governmental transparency 
+but also empowers researchers, civic developers, and the general public.
 
@@
 and at \href{https://beta.nyc/beta/products/}{\textbeta etaNYC} products and tools. 
-
 Despite its success, the system constantly faces challenges such as:
@@
 
-These and other challenges are continually addressed by the various Agency open data managers
-as well as the 
-\href{https://www.nyc.gov/content/oti/pages/}{NYC Office of Technology and Innovation (OTI)} 
-which provides the technology support for the open data system.  
-Many of the underlying datasets are updated daily, usually running 
-1-2 days behind from creation. However, a number of datasets 
-are updated monthly (such as restaurant inspections), and some are 
-not updated for one or more years (such as community board 
+These and other challenges are continually addressed by the various 
+Agency open data managers as well as the 
+\href{https://www.nyc.gov/content/oti/pages/}
+{NYC Office of Technology and Innovation (OTI)} which provides the 
+technology support for the open data system. Many of the underlying 
+datasets are updated daily, usually running 1-2 days behind 
+from creation. However, a number of datasets are updated 
+monthly (such as restaurant inspections), and some are not 
+updated for one or more years (such as community board 
 boundaries). This poses problems when merging datasets with 
@@
 311 system was designed to be easily configured and updated with new features, 
-some emerging business issues as well as technologies nonetheless require additional software engineering efforts to accommodate, along with the concomitant budgets.
+some emerging business issues as well as technologies nonetheless 
+require additional software engineering efforts to accommodate, 
+along with the concomitant budgets.
 
@@
 6.4 million records. That dataset serves as the primary source for the analysis effort. 
-This is a large dataset which enables 
-observation of some rarely seen events, data inaccuracies, and logical inconsistencies. 
-It provides enough data to examine trends and data outlier analysis. 
-We queried the data directly from the NYC Open Data Portal and then exported the 
-data in CSV format for analysis using custom written R programs.
+This is a large dataset which enables observation of some rarely 
+seen events, data inaccuracies, and logical inconsistencies. It 
+provides enough data to examine trends and data outlier analysis. 
+We queried the data directly from the NYC Open Data Portal and then 
+exported the data in CSV format for analysis using custom written R programs.
 
@@
 \section{10-year Analysis \& Trends} \label{sec:trends}
-
 Since the NYC 311 System went live in 2003, the usage has grown at an annual
@@
 the number of 311 SRs grew 50\%. The summer  and early fall months 
-make up all of the Top 5 SR counts by day. 
-The all-time daily high for SRs, 24,415 SRs (a full 9.8 \textsigma's from the 10-year mean), 
-occurred on Tuesday, 2020-08-04.
-The 10-year low count occurs on Christmas Day, 2016, a Sunday. 
-Note also that the difference between 
+make up all of the Top 5 SR counts by day. The all-time daily 
+high for SRs, 24,415 SRs (a full 9.8 \textsigma's from the 10-year mean), 
+occurred on Tuesday, 2020-08-04. The 10-year low count occurs 
+on Christmas Day, 2016, a Sunday. Note also that the difference between 
 the peak day (24,415) and the minimum day (2965) varies by 8X, creating a 
 challenging system performance issue. This timeframe (2014-2023) 
-also includes the COVID pandemic,  
-which reached a peak in the July-September 2020 timeframe, 
-accompanied by additional spikes in the 2\textsuperscript{nd}half of 2021 as well.
+also includes the COVID pandemic, which reached a peak in 
+the July-September 2020 timeframe, accompanied by additional 
+spikes in the 2\textsuperscript{nd}half of 2021 as well.
 
@@
     \centering
-    \normalsize
     \caption{Top 5 and Bottom 5 Days for SRs}
@@
 of 2020 are ``below average'', while most of the months after that period 
-are ``above average''. That level of higher usage
-of the 311 system has continued into the early months of 2024. 
-It is not surprising that the lowest SR counts-per-day occur during the Christmas holiday period.
+are ``above average''. That level of higher usage of the 311 system 
+has continued into the early months of 2024. It is not surprising that 
+the lowest SR counts-per-day occur during the Christmas holiday period.
 
@@
 Two important fields in this analysis effort are the responsible NYC Agency, 
-and the type of complaint.
-Given the size of the City of New York government, it is necessary to 
-identify the responsible Agency as well as the type of complaint in order to 
-identify the responsible party and assist in troubleshooting discrepancies.
+and the type of complaint. Given the size of the City of New 
+York government, it is necessary to identify the responsible 
+Agency as well as the type of complaint in order to identify 
+the responsible party and assist in troubleshooting discrepancies.
 
@@
 Thus we can often identify if it is an Agency specific problem, or a systemic issue.  
-(Note: a few key NYC Agencies are not 
-represented in the 311 dataset. These include the New York City Housing Authority (NYCHA) and the
-Department of Corrections (DOC).)
-
-Although the 311 SRs involve 16 different agencies, there are the ``big six'' 
-Agencies that comprise 90\% of the complaints. 
-These six are (in order of \% of complaints):
+(Note: a few key NYC Agencies are not represented in the 311 
+dataset. These include the New York City Housing Authority (NYCHA) 
+and the Department of Corrections (DOC).)
+
+Although the 311 SRs involve 16 different agencies, there are 
+the ``big six'' Agencies that comprise 90\% of the complaints. These 
+six are (in order of \% of complaints):
 
@@
     \centering
-    \normalsize
     \caption{Noise-related complaints\_type(s) by count with Agency}
@@
 	        \midrule
-        NOISE - RESIDENTIAL        & 675502 & 10.56 & NYPD  \\ 
-        NOISE - STREET/SIDEWALK    & 300507 &  4.70 & NYPD  \\ 
-        NOISE - COMMERCIAL         & 128892 &  2.02 & NYPD  \\ 
-        NOISE - VEHICLE            & 115956 &  1.81 & NYPD  \\ 
-        NOISE                      & 100413 &  1.57 & DEP   \\ 
-        NOISE - HELICOPTER         &  85345 &  1.33 & EDC   \\ 
-        NOISE - PARK               &  16620 &  0.26 & NYPD  \\ 
-        NOISE - HOUSE OF WORSHIP   &   2757 &  0.04 & NYPD  \\ 
+	        Noise - Residential        & 675502 & 10.56 & NYPD  \\ 
+	        Noise - Street/Sidewalk    & 300507 &  4.70 & NYPD  \\ 
+	        Noise - Commercial         & 128892 &  2.02 & NYPD  \\ 
+	        Noise - Vehicle            & 115956 &  1.81 & NYPD  \\ 
+	        Noise                      & 100413 &  1.57 & DEP   \\ 
+	        Noise - Helicopter         &  85345 &  1.33 & EDC   \\ 
+	        Noise - Park               &  16620 &  0.26 & NYPD  \\ 
+	        Noise - House of Worship &   2757 &  0.04 & NYPD  \\ 
 	        \bottomrule
@@
 \section{Data Cleansing Issues} \label{sec:issues}
-
 What is data cleansing?  Wikipedia 
@@
 
-``Data cleansing or data cleaning is the process of detecting and 
+\begin{quote}\textit{Data cleansing or data cleaning is the process of detecting and 
 correcting (or removing) corrupt or inaccurate records from a record set, 
 table, or database and refers to identifying incomplete, incorrect, 
-inaccurate or irrelevant parts of the data 
-and then replacing, modifying, or deleting the dirty or coarse data.''
-
-Many quality criteria are required in order to process high-quality data. 
-These include:
+inaccurate or irrelevant parts of the data and then replacing, 
+modifying, or deleting the dirty or coarse data.}
+\end{quote}
+
+Many quality criteria are required in order to process high-quality 
+data. These include:
 
@@
 For example, an invalid date field that is provided to the NYC Open Data 
-data lake via an Agency integration could
-be the result of selecting the wrong field from the source Agency's 
-system, e.g. due\_date instead of closed\_date. In other cases, without a subject matter
+data lake via an Agency integration could be the result of selecting 
+the wrong field from the source Agency's system, e.g. 
+due\_date instead of closed\_date. In other cases, without a subject matter
 expert, it is not possible to actually determine if the data is inaccurate or not. 
@@
 \section{Structural Issues}\label{sec:structural}
-
 Structural issues in the context of data cleansing refers to issues 
@@
 Fortunately, both the Data Dictionary and the portal column manager 
-do a good job of indicating the data type.
-As indicated by the red underline in Figure \ref{fig:computed-columns}, 
-there is a small icon next to the field name which indicates the 
+do a good job of indicating the data type. As indicated by 
+the red underline in Figure \ref{fig:computed-columns}, there 
+is a small icon next to the field name which indicates the 
 data type of the field. The stylized capital ``T'' indicates ``text'', the ``\#'' sign indicates 
 ``numeric'', a calendar icon indicates a ``date'' field, and the pin on a 
-map symbol indicates a ``geospatial'' field.
-
-The following fields were checked for their specified data types:
+map symbol indicates a ``geospatial'' field. The following fields were 
+checked for their specified data types:
 
@@
 \section{Blank \& N/A data}\label{sec:blanks}
-
 Understanding the absence of data by field is an important factor 
@@
 \centering
-\normalsize
 \caption{Blank and N/A entries by Field}
@@
 
-%\FloatBarrier % Ensure all floats before this point are processed before moving on
-
 Here is a graphic depiction of total empty (blank \& N/As) for each fields. 
@@
 \subsubsection{Case Study: Noise Complaints by Zip Code}
-		{label{sec:case-study-zip-codes}
+\label{sec:case-study-zip-codes}
 		\textbf{Scenario:} The NYC Office of Nightlife (ONL) wants to 
@@
     \centering
-    \normalsize
     \caption{Comparison of Top Ten Zip Codes Lists}
@@
 	\begin{table}[tbp]
-	\normalsize
 	\centering
@@
 
-	A graphic representation of the invalid police\_precincts plotted 
-	by Agency show a near perfect match to the distribution of the 
-	entire 6.4 million SRs as seen in Figure \ref{fig:SR_counts_by_Agency}. 
-	Note that the ``big six'' Agencies account for 90\% of these invalid 
-	precinct errors, identical to the overall distribution of all SRs. 
-	(The same is true for a similar graph of by the police\_precint field.) 
-	This strongly suggests that this is a systemic problem, 
-	almost certainly caused by the computation method applied for these two fields.
+	A graphic representation of the invalid police\_precincts plotted by 
+	Agency show a near perfect match to the distribution of the entire 
+	6.4 million SRs as seen in Figure \ref{fig:SR_counts_by_Agency}. Note 
+	that the ``big six'' Agencies account for 90\% of these invalid precinct 
+	errors, identical to the overall distribution of all SRs. (The same is true 
+	for a similar graph of by the police\_precint field.) This strongly 
+	suggests that this is a systemic problem, most likely caused by the 
+	computation method applied for these two fields.
 
@@
 \subsection{Issues with Community Boards}
+\label{sec:communityboards}
 	Community Boards are an important aspect of NYC government. They are the 
@@
 		\centering
-		\normalsize
 		\caption{Top Ten Invalid 'community\_board' Values}
@@
 \subsection{Issues with Community Districts}
+\label{sec:communitydistrict}
 	Another of the computed fields is community\_districts. Community Districts 
@@
 \subsection{created\_date and closed\_date(s) -- Negative Duration}
+\label{sec:negativeduration}
 	Let's look first at the created\_date and closed\_date fields. You might 
@@
 
-\textbf{Why are these date fields important?} They're important because citizens, NYC Government Officials, and NYC Agencies use these dates to measure the responsiveness 
-for providing the services. How long does it take to replace an out-of-order 
-street light? How rapid is the response to a report of domestic violence? 
-Does the NYPD response to a noise complaint in Staten Island take longer 
-than in Manhattan? How long does it take to repair a street pothole in Queens?
+	\textbf{Why are these date fields important?} They're important because 
+	citizens, NYC Government Officials, and NYC Agencies use these 
+	dates to measure the responsiveness for providing the 
+	services. How long does it take to replace an out-of-order 
+	street light? How rapid is the response to a report of domestic 
+	violence? Does the NYPD response to a noise complaint in 
+	Staten Island take longer than in Manhattan? How long does it 
+	take to repair a street pothole in Queens?
 
@@
 	    \centering
-    \normalsize
 	    \caption{Largest and Smallest errors (days)}
-    \begin{tabular}{>{\ttfamily}l l l r l}
+		    \begin{tabular}{l l l r l}
 		        \toprule
@@
 	    \centering
-     \normalsize
 	    \caption{Sample of SRs with extremely large negative durations}
@@
 \subsubsection{Case Study: Homeless Person Assistance}
+\label{sec:homlessassistance}
 	\textbf{Scenario:} The Dept. of Homeless Services wants to 
@@
 \subsection{created\_date and closed\_date(s) --  Zero Duration}
+\label{sec: zeroduration}		
 	An even more serious, and more prevalent problem with closed\_date(s) 
@@
 \subsection{created\_date and closed\_date(s) -- Midnight \&Noon}
+\label{sec:midnightandnoon}
 	An additional problem discovered with the created\_date and 
@@
 \subsection{resolution\_action\_update\_date}
+\label{sec: resolutionaction}
 	When an SR is updated, by any means on any field, the software 
@@
 	    \centering
-	    \normalsize
 	    \caption{Extremely long Resolution Updates}
@@
     	\centering
-    \normalsize
      \caption{Long post-closed Resolution Updates}
@@
 
-\section{Accuracy and precision}\label{sec:precision}
+\section{Accuracy and precision}
+\label{sec:precision}
 There is one area in particular where the question of precision vs. 
@@
 
-\subsection{latitude \& longitude and the location fields}  The location field is a 
-pure concatenation of the latitude and longitude fields, with a comma and 
-parenthesis added. Example:
+\subsection{latitude \& longitude and the location fields}
+\label{sec:latlong}
+The location field is a pure concatenation of the latitude 
+and longitude fields, with a comma and parenthesis 
+added. Example:
 
@@
 
-\subsection{borough and park\_borough fields}  These two fields are 
-100\% matches; fully redundant.
-
-\subsection{borough and borough\_boundaries fields}  The borough field 
-is a text field for the five New York City boroughs (Manhattan, Queens, 
+\subsection{borough and park\_borough fields}
+\label{sec:parkborough}
+These two fields are 100\% matches; fully redundant.
+
+\subsection{borough and borough\_boundaries fields}
+\label{sec:boroughboundaries}
+The borough field is a text field for the five New York City boroughs (Manhattan, Queens, 
 Brooklyn, etc.). The borough\_boundary field is numeric with values 
@@
 \subsection{borough and taxi\_company\_borough fields}
+\label{sec:taxicompanyborough}
 The two fields borough and taxi\_company\_borough might appear to be 
@@
  \subsection{incident\_zip and zip\_codes fields}
+ \label{sec:zipcodes}
  The zip\_code field is one of the six computed fields and its validity 
@@
  \subsection{police\_precinct and police\_precincts fields} 
+ \label{sec:police} 
 Both the police\_precinct and police\_precincts are among the computed 
@@
  \subsection{agency and agency\_name fields}
+ \label{sec:agencyname}
  While not duplicates, these fields have a 1:1 correspondence.  The Agency 
@@
 \subsection{landmark and street\_name}
+\label{sec:landmark}
 The landmark field is listed in the Data Dictionary as ``Can refer to 
@@
     \centering
-    \normalsize
     \caption{Non-matches between 'street\_name' and 'landmark' fields}
@@
 \subsection{cross\_street\_1 \& intersection\_street\_1 and cross\_street\_2 
-\& intersection\_street\_2} There are two sets of street pairs in the 
-dataset in addition to the complaint address:
+\& intersection\_street\_2}
+\label{sec:street1}
+There are two sets of street pairs in the dataset in addition to 
+the complaint address:
 
@@
 	\centering
-     \normalsize
-     \caption{Matching/Non-Matching cross\_street\_1 
-     \& intersection\_street\_1}
-    \begin{tabular}{>{\normalsize\ttfamily}l >
-    {\normalsize\ttfamily}l >{\normalsize\ttfamily}l}
+     \caption{Matching/Non-Matching cross\_street\_1 \& intersection\_street\_1}
+     	\label{tab:streets1}
+	    	\begin{tabular}{l l l}
 	      	\toprule
 	      	\multicolumn{3}{c}
-        {\textbf{Matching cross\_street\_1 and intersection\_street\_1}} \\
+	      	{\textbf{Matching cross\_street\_1 \& intersection\_street\_1}} \\
 	      	\midrule
-        \textbf{cross\_street\_1} & \textbf{intersection\_street\_1} 
-        & \textbf{agency} \\
+	      	\textbf{cross\_street\_1} & \textbf{intersection\_street\_1} & \textbf{agency} \\
 	      	\midrule
@@
 	      	\midrule
-        \textbf{cross\_street\_1} & \textbf{intersection\_street\_1} 
-        & \textbf{agency} \\
+	      	\textbf{cross\_street\_1} & \textbf{intersection\_street\_1} & \textbf{agency} \\
 	      	\midrule
@@
 	   	\end{tabular}
-    \label{tab:streets1}
 \end{table}
@@
     \centering
-     \normalsize
      \caption{Near-matches for cross\_street\_1, intersection\_street\_1}
+     \label{tab:x1nearmatches}
 		\begin{tabular}{l l l r}
@@
 	    \end{tabular}
-     \label{tab:x1nearmatches}
 \end{table}
@@
     \centering
-    \normalsize
     \caption{Summary of cross\_street\_1 and intersection\_street\_1}
+    \label{tab:summary1}
 	    \begin{tabular}{l r r}
@@
 	    \end{tabular}
-    \label{tab:summary1}
 \end{table}
-
 \begin{table}[tbp]
     \centering
-    \normalsize
     \caption{Summary of cross\_street\_2 and intersection\_street\_2}
+    \label{tab:summary_cross_intersection_2}
 	    \begin{tabular}{l r r}
@@
 	    \end{tabular}
-    \label{tab:summary_cross_intersection_2}
 \end{table}
@@
 	\centering
-     \normalsize
 	\caption{Sample of near-matching cross\_street\_2 and 
      	intersection\_street\_2 (both non-blank)}
+     	\label{tab:near_matching_cross_intersection_2}
 	     	\begin{tabular}{l l l r}
@@
 	    	\end{tabular}
-    \label{tab:near_matching_cross_intersection_2}
 \end{table}
@@
  \subsection{Shrinking file size by removing duplicates}
-
+ \label{sec:filesize}
 By removing duplicate and ``near-duplicate'' fields, it is possible to 
 shrink the file size by 12.3\% which for this dataset equates to a 
- reduction of 395 Mb. A smaller file size means faster downloads, less 
- storage impact, as well as simplifying data analysis efforts.
-
- Below is a list of duplicate and near-duplicate fields. There are some 
- challenges with these proposed deletions, mostly with the ``near 
- duplicate'' fields.  So while the park\_borough is a complete duplicate 
- of the borough field, the borough\_boundary field is only 98.3\% 
- duplicate. Admittedly, loosing 1.7\% of the data in those two fields 
- is quite small, for this large dataset that amounts to 110, 715 
- non-matching occurrences. Similarly with police\_precincts 
- and police\_precincts where there is a 99.9\% match; close 
- to 100\% but not a complete match.
+reduction of 395 Mb. A smaller file size means faster downloads, 
+less storage impact, as well as simplifying data analysis efforts.
+
+Below is a list of duplicate and near-duplicate fields. There are 
+some challenges with these proposed deletions, mostly with 
+the ``near duplicate'' fields.  So while the park\_borough is 
+a complete duplicate of the borough field, the 
+borough\_boundary field is only 98.3\% duplicate. Admittedly, 
+loosing 1.7\% of the data in those two fields is quite small, for 
+this large dataset that amounts to 110, 715 non-matching 
+occurrences. Similarly with police\_precincts and police\_precincts 
+where there is a 99.9\% match; close to 100\% but 
+not a complete match.
 
@@
 
-\section{Data Dictionary Observations} \label{sec:datadictionay}
+\section{Data Dictionary Observations} 
+\label{sec:datadictionay}
 During this analysis, it became apparent that the published Data 
@@
 
-\section{Recommendations} \label{sec:recommendations}
+\section{Recommendations} 
+\label{sec:recommendations}
 Having identified a number of data cleanliness related issues with the 
@@
 	the many different methods of representing empty or null data fields. 
-
 \end{itemize}

### 2024-06-18 manuscript/manuscript.tex
Cleanup after mistake with github.
First attempt to reducing size of the document
Also, updated post_closed_violin chart (color change) which is reflected in datacleansing.R also.

@@
 
-\section{Introduction} \label{sec:intro}
+\section{Introduction} 
+\label{sec:intro}
 In the early 21st century, the open data movement began to take shape, driven by the 
@@
 \begin{itemize}
-	\item NYC restaurant violations
-	\item Popular baby names
+	\item NYC restaurant health inspection violations
 	\item Mapping of car crashes involving pedestrians
-	\item Mapping of sidewalk widths in NYC 
 	\item Visualization of NYC High School and College enrollment
@@
 including data validity, consistency, and curation efficiency. 
-We illustrate these points with live examples drawn from our processing of the 311 SR data. 
-Secondly, building upon insights gained from this case study, we 
-propose a set of data curation principles tailored for government-released open data. 
-These principles are designed to address the unique challenges 
+We illustrate these points with live examples drawn from our 
+processing of the 311 SR data. Secondly, building upon insights 
+gained from this case study, we propose a set of data curation 
+principles tailored for government-released open data. These 
+principles are designed to address the unique challenges 
 and requirements observed in the curation of such datasets.
@@
 Section~\ref{sec:history} offers a brief review of the history of the 311 system. In 
-Section~\ref{sec:trends} we take a look at long-term trends presented via a 10-year analysis.
-Section~\ref{sec:issues} offers a general discussion of data cleansing issues
-impacting data quality and curation efficiency. Section~\ref{sec:structural} examines
+Section~\ref{sec:trends} we take a look at long-term trends presented 
+via a 10-year analysis. Section~\ref{sec:issues} offers a 
+general discussion of data cleansing issues impacting data 
+quality and curation efficiency. Section~\ref{sec:structural} examines
 the technical dimensions of the dataset, the structural issues. 
 In Section~\ref{sec:datatypes} we examine the data fields for compliance 
-with the stated data type as found in the Data Dictionary. Section~\ref{sec:blanks} looks
-at the data fields as regards missing, blank, or N/A entries. Section~\ref{sec:domain} explores
-how data fields comply with a domain of legal or acceptable values. Section~\ref{sec:inconsistencies}
-deals with the important issues surrounding logical inconsistencies 
+with the Data Dictionary. Section~\ref{sec:blanks} looks at the 
+data fields as regards missing, blank, or N/A entries. Section~\ref{sec:domain} 
+explores how data fields comply with a domain of legal 
+or acceptable values. Section~\ref{sec:inconsistencies} deals 
+with the important issues surrounding logical inconsistencies 
 and concerning patterns in the data. Section~\ref{sec:precision} 
-explores the age-old issue of precision versus accuracy. Section~\ref{sec:duplicates}
+explores the issue of precision versus accuracy. Section~\ref{sec:duplicates}
 identifies duplicate and redundant data fields. Section~\ref{sec:improvements} provides 
@@
 
-A note on naming conventions: dataset field names are typically 
-named using ``snake\_case'' where each word in the variable name 
+A note on naming conventions: dataset field names are named 
+using ``snake\_case'' where each word in the variable name 
 is written in lowercase, and words are separated by 
@@
 
-\section{NYC 311 System History} \label{sec:history}
+\section{NYC 311 System History} 
+\label{sec:history}
 The NYC 311 service, a critical component of New York City's public
@@
 comprehensive data management system that handles millions of requests
-annually. Each inquiry is logged as a complaint in the 311 system.
-Key milestones in the 311 system development include:
+annually. Each inquiry is logged as a complaint, a Service Request (SR), 
+in the 311 system. Key milestones in the 311 system include:
 
@@
    	Service Requests come from mobile \& online channels.
-   	\item 2019 - Major software upgrade of 311 system with enhanced capabilities, 
-   	dynamic load-balancing, and cloud-based resilience. Only major 
-   	software upgrade since 2004.	
       \item 2020 - In August 2020, driven by the COVID pandemic, monthly 311 
-      Service Requests hit an all-time high of 348,463. Seven of the Top 10 busiest 
-      days of all-time occur in 2020.
-      \item 2021 - 311 System is expanded to into the MTA's city subway system, 
+      SRs hit an all-time high of 348,463. Seven of the Top 10 all-time busiest 
+      days occur in 2020.
+      \item 2021 - 311 System is expanded to include the MTA's city subway system, 
       the largest expansion since its inception.
@@
 illegal parking, heat/hot water, abandoned vehicles,
-and unsanitary conditions. The data, managed through a tailored application built on the 
-Microsoft Dynamics and Azure platform, is a valuable resource for city
+and unsanitary conditions. The data is a valuable resource for City
 administration and policy-making. The 311 data is publicly accessible through the 
-NYC Open Data Portal which provides access
-to all the data sets (approx. 2700) and provides a convenient method for querying, 
+NYC Open Data Portal which provides access to all the data 
+sets (approx. 2700) and provides a convenient method for querying, 
 grouping, aggregating, geo-mapping, visualization, and exporting 
-results. This open data initiative enables not only governmental transparency 
-but also empowers researchers, civic developers, and the general public.
-
-Additional Open Data projects can be found 
-at \href{https://opendata.cityofnewyork.us/projects/}{NYC Open Data Project Gallery}
-and at \href{https://beta.nyc/beta/products/}{\textbeta etaNYC} products and tools. 
-Despite its success, the system constantly faces challenges such as:
+results. This open data  enables not only governmental transparency 
+but empowers researchers, civic developers, and the general public.
+
+Despite its success, the 311 system faces challenges such as:
 
@@
 	party users, of which there are many
-	\item Incorporating ongoing technology changes and upgrades 
-	\item Understanding the role of chatbots and AI to provide online 
-	customer service, along with other emergent technologies
-\end{itemize}
-
-These and other challenges are continually addressed by the various 
+\end{itemize}
+
+These and other challenges are addressed by the various 
 Agency open data managers as well as the 
@@
 {NYC Office of Technology and Innovation (OTI)} which provides the 
-technology support for the open data system. Many of the underlying 
-datasets are updated daily, usually running 1-2 days behind 
-from creation. However, a number of datasets are updated 
-monthly (such as restaurant inspections), and some are not 
-updated for one or more years (such as community board 
-boundaries). This poses problems when merging datasets with 
-others, as would be expected. Typically, the data must be exported 
-and manipulated in another, external program. The authors found 
-this to be the case with the large 311 dataset employed.
-
-Data accuracy and consistency is typically established between OTI 
-and the other contributing City Agencies, such as NYPD, Housing 
-Preservation and Development (HPD), Parks \& Recreation, etc., via 
-Service Level Agreements (SLAs). However, enforcement of those 
-SLAs is often challenging. The result is that often data accuracy and 
-logical consistency is occasionally compromised. However, 
-the Open Data Team is quite responsive to inquires from users. The 
-authors have provided many such suggestions during the course of this research effort, 
-and the most current data shows improvements in many of the areas highlighted. 
-Questions and suggestion  can be provide here: 
-\href{https://opendata.cityofnewyork.us/engage/}{Contact Us}.
-
-The NYC 311 organization, part of OTI, operates a highly tailored and sophisticated case
-management. However, one constraint was particularly challenging 
-during the recent complete overhaul of the software and hardware in 2019. When the original 311system was developed in 2003, a decision was made to allow select 
-NYC Agencies to maintain their own local computer systems, 
-such as the mainframe systems at the NYC Department of Sanitation (DSNY) and others. 
-And when the 311 system was completely upgraded in 2019, those
-same integrations were honored in order to reduce the scope of the project. 
-Perhaps in retrospect, while a good budgetary and policy decision, 
-it was perhaps not such a good technology decision.
-
-At least nine NYC Agencies have local systems that are integrated 
-directly to the 311 system either via an API or custom code. These include:
-
-\begin{itemize}
-	\item Department of Social Services (DSS), one of the largest City Agencies
-	\item Department of Sanitation, although recently part of DSNY 
-	has moved to the core 311 software.
-	\item Department of Parks \& Recreation (forestry complaints only)
-	\item Housing, Preservation \& Development (fully integrated)
-	\item Department of Environmental Protection (DEP) (fully integrated)
-	\item Department of Healthy \& Mental Hygiene (fully integrated) 
-	\item Department of Consumer \& Worker Protection 
-	(DCWP - formerly known as the Dept of Consumer Affairs)
-	\item New York Police Department (NYPD) (integrated with their reporting system)
-	\item Department of Business (DoB) (fully integrated)
-\end{itemize}
-
-One unfortunate side-effect from these integrations, and other Agencies 
-that use the 311 APIs, is that some of the data consistency
-checks which are inherent in the 311 system are not applied uniformly, 
-and in some cases not at all, to the integrated data streams.
-Errors often creep in from other Agency systems and are not easily 
-corrected or detected in the 311 application itself,
-but rather require correction at the source Agency; challenging.
-
-One area that will almost certainly remain an ongoing issue is the 
-ongoing incorporation of new software, hardware,
-and cloud services which are constantly changing while new approaches emerge. 
-The 311 telephonic integration recently was reconfigured due to a new
-telephonic vendor. Desktop software at the 311 Agency are subject 
-to the same upgrades that home and business users experience. And while the 
-311 system was designed to be easily configured and updated with new features, 
-some emerging business issues as well as technologies nonetheless 
-require additional software engineering efforts to accommodate, 
-along with the concomitant budgets.
-
-One area in particular which is heavily influencing customer outreach for the 311 system
- is the appropriate use of online chatbots and assistants. Increasing these features are 
- influenced by artificial intelligence (AI) advances. It is a given that a number of 
- citizen's interactions with City government could be handled accurately and quickly by AI-driven
-applications thus improving timeliness and accuracy. But how to do that while 
-preserving customer privacy and anonymization and user engagement are challenging issues.
+technology support for the open data system. Typically, the data 
+will be exported and processed in an external program. The authors found 
+this to be the case with the large 311 dataset examined.
 
@@
 	which includes street canvassers and increases in homeless housing facilities.
-	\item A significant increase in graffiti complains in 2017 led directly to 
-	the `Graffiti-Free NYC'' campaigns and increased funding for removal of graffiti in public spaces.
-	\item Enhanced geospatial data regarding the location of incidents, 
-	the responding Agency, and actions taken across a myriad of complaint\_types
-	\item Establishment of Agency SLAs in regard to certain complaint types, 
-	such as residential noise, homelessness person assistance, street light outages, 
-	rat sightings, and other neighborhood quality of life issues.
-\end{itemize}
-
-The NYC 311 service exemplifies the dynamic nature of urban data
-management and its critical role in modern governance. In addition it provides an
-invaluable resource for data scientists to practice data science activities, including data curation.
-
-We extracted 311 SR data for the period of CY 2022-2023, a dataset consisting of
-6.4 million records. That dataset serves as the primary source for the analysis effort. 
-This is a large dataset which enables observation of some rarely 
-seen events, data inaccuracies, and logical inconsistencies. It 
-provides enough data to examine trends and data outlier analysis. 
-We queried the data directly from the NYC Open Data Portal and then 
-exported the data in CSV format for analysis using custom written R programs.
-
-Additionally, we extracted two longer-term datasets, a 5-year look (2019-2023) 
-and a 10-year look (2014-2023) ; some 14 and 27 million rows respectively. 
-These larger datasets afford an opportunity for long-term trend observations, 
-confirming and supporting the analysis using the 2-year dataset. The 2022-2023 dataset
-was used exclusively for the detailed, field-by-field level analysis 
-which comprises the majority of this effort.
-
-\section{10-year Analysis \& Trends} \label{sec:trends}
-Since the NYC 311 System went live in 2003, the usage has grown at an annual
-rate of 4.6\%. As part of this study, we also looked at a 10-year period (2014-2023) to
+\end{itemize}
+
+We extracted 311 SR data for the period of CY 2022-2023, a dataset of
+6.4 million SR records. This is a large dataset which enables 
+observation of some rarely seen events, data inaccuracies, 
+and logical inconsistencies. It provides enough data to 
+examine trends and data outlier analysis. We queried the data 
+directly from the NYC Open Data Portal and exported the 
+data for analysis with custom written R programs.
+
+\section{10-year Analysis \& Trends} 
+\label{sec:trends}
+As part of this study, we also looked at a 10-year period (2014-2023) to
 observe system performance over a longer time-frame . During that time-frame, 
-the number of 311 SRs grew 50\%. The summer  and early fall months 
+the number of 311 SRs grew by 50\%. The summer and early fall months 
 make up all of the Top 5 SR counts by day. The all-time daily 
@@
 occurred on Tuesday, 2020-08-04. The 10-year low count occurs 
-on Christmas Day, 2016, a Sunday. Note also that the difference between 
-the peak day (24,415) and the minimum day (2965) varies by 8X, creating a 
-challenging system performance issue. This timeframe (2014-2023) 
-also includes the COVID pandemic, which reached a peak in 
-the July-September 2020 timeframe, accompanied by additional 
-spikes in the 2\textsuperscript{nd}half of 2021 as well.
+on Christmas Day, 2016, a Sunday.
 
@@
 	        \toprule
-	        \multicolumn{3}{c}{Top 5 Days} & \multicolumn{3}{c}{Bottom 5 Days} \\
-	        \cmidrule(r){1-3} \cmidrule(l){4-6}
-	        Date & Count & Rank & Date & Count & Rank \\
+	        \multicolumn{2}{c}{Top 5 Days} & \multicolumn{2}{c}{Bottom 5 Days} \\
+	        \cmidrule(r){1-2} \cmidrule(l){3-4}
+	        Date & Count & Date & Count\\
 	        \midrule
-	        2020-08-04 & 24,415 & 1 & 2014-12-25 & 2965 & 1 \\
-	        2020-08-05 & 19,560 & 2 & 2015-12-25 & 3247 & 2 \\
-	        2023-09-29 & 17,962 & 3 & 2014-04-20 & 3287 & 3 \\
-	        2020-07-05 & 16,916 & 4 & 2015-12-26 & 3405 & 4 \\
-	        2020-06-21 & 15,883 & 5 & 2016-11-06 & 3419 & 5 \\
+	        2020-08-04 & 24,415 & 2014-12-25 & 2965\\
+	        2020-08-05 & 19,560 & 2015-12-25 & 3247\\
+	        2023-09-29 & 17,962 & 2014-04-20 & 3287\\
+	        2020-07-05 & 16,916 & 2015-12-26 & 3405\\
+	        2020-06-21 & 15,883 & 2016-11-06 & 3419\\
 	        \bottomrule
@@
 \end{table}
-
-This chart shows the 2-year period which is the basis for this analysis, by calendar month. 
-Note that almost all the months before the COVID pandemic outbreak in Feb/Mar
-of 2020 are ``below average'', while most of the months after that period 
-are ``above average''. That level of higher usage of the 311 system 
-has continued into the early months of 2024. It is not surprising that 
-the lowest SR counts-per-day occur during the Christmas holiday period.
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=\textwidth]{2-year-trend-monthly.pdf}
-  \caption{Monthly SR Counts for 2022-2023}
-  \label{fig:2-yr-monthly}
-\end{figure}
 
@@
 the responsible party and assist in troubleshooting discrepancies.
-
-When we discovered a data error, we typically observe one of two trends; 
-either a single (or a few) Agencies are responsible (such as Dept of Transportation), 
-or it is a system-wide issue that follows the general distribution of  SRs across the system. 
-Thus we can often identify if it is an Agency specific problem, or a systemic issue.  
-(Note: a few key NYC Agencies are not represented in the 311 
-dataset. These include the New York City Housing Authority (NYCHA) 
-and the Department of Corrections (DOC).)
-
-Although the 311 SRs involve 16 different agencies, there are 
-the ``big six'' Agencies that comprise 90\% of the complaints. These 
-six are (in order of \% of complaints):
+Thus we can often identify if it is an Agency specific problem, 
+or a systemic issue. Although 311 SRs involve 16 different agencies, 
+there are six Agencies that comprise 90\% of the complaints.
 
@@
 accompanied by a long tail. This graph shows the top 20 complaint\_type(s) 
-comprising 70\% out of a total of 220.
+comprising 70\% out of a total of 220 different types.
 
@@
 complaints (Residential, Commercial, Street, Helicopter, etc.). These Noise-related
-complaints total 1.435 million in the 2022-2023 dataset, a full 22\% of 
-all complaint types, which, when consolidated, is the most frequently occurring complaint\_type.
-Illegal Parking is second, followed by Heat/Hot Water. Noise complaints 
-are handled by NYPD, as is Illegal Parking and Blocked Driveway complaints, 
-which contributes significantly to the top ranking of the NYPD as the lead 
-responsible Agency, receiving a full 43\% of all SRs. 
-Housing, Preservation, \& Development are 2\textsuperscript{nd}, with 21\%
-
-For  non-NYC residents, HPD is the City agency that manages the 
-177,569 NYC public housing units as well as monitoring
-all City rental and leased apartment units. HPD collectively serves 
-528,105 people, a population larger than Atlanta or Miami. 
-Hence the large number of Heat/Hot Water and Unsanitary complaints 
-handled by HPD.
+complaints total a full 22\% of all complaint types, which, when consolidated, is the most frequently occurring complaint\_type.
 
@@
 
-There are also some curious and unusual Service Requests in the 311 data: 
-tanning, tattooing, trans fat, unsanitary pigeon condition, illegal animal 
-kept as pet, harboring bees/wasps, and the authors' favorite
-radioactive material. New Yorkers certainly live interesting lives.
-
-\section{Data Cleansing Issues} \label{sec:issues}
+\section{Data Cleansing Issues} 
+\label{sec:issues}
 What is data cleansing?  Wikipedia 
@@
 		\item Data-types: Certain fields must be of the correct type, e.g. numeric, 
-		character, date, 5 numeric digits, etc. Typically these are identified 
+		character, date, 5 numeric digits, etc. Typically these are specified 
 		in a Data Dictionary.
-		\item Domain adherence:  Many data fields must adhere to a specific 
+		\item Domain compliance:  Many data fields must adhere to a specific 
 		domain of values, e.g. statuses, state names, zip codes, gender. 
-		This includes data fields that are restricted to certain ranges, 
-		such as longitudes and latitudes.
 	\end{itemize}   
 	\item Structural errors to include naming conventions, 
-	extra fields not in the Data Dictionary, or inconsistent data entry.
-	For example intermixing blanks, spaces, NA, N/A, and \textless{}NA\textgreater{} 
+	 fields not in the Data Dictionary, or inconsistent data entry. For example 
+	 intermixing blanks, spaces, NA, N/A, and \textless{}NA\textgreater{} 
 	 all to indicate the absence of data.
@@
 	nature of that relationship, e.g. a ``due date'' that is before the ``created date''
-	\item Data standardization. Often free-form entry fields can suffer 
-	from inconsistent data structures. Efforts to standardize these fields 
-	can greatly assist analysis.
-	\item Accuracy and precision, which of course are not the same thing 
+	\item Accuracy versus precision
 \end{itemize}
@@
 and how such errors came about, and a discussion as to whether or not it 
-is even an ``error''. Typically this requires discussions with subject matter 
-experts (SMEs) and liaison within the various NYC Agency open data 
-coordinators who can help identify the underlying cause for the error. 
-For example, an invalid date field that is provided to the NYC Open Data 
-data lake via an Agency integration could be the result of selecting 
-the wrong field from the source Agency's system, e.g. 
-due\_date instead of closed\_date. In other cases, without a subject matter
-expert, it is not possible to actually determine if the data is inaccurate or not. 
-There are many empty values in the taxi\_company\_borough field. Are those
-entries missing or does the entry in that field only apply to a 
-very small number of complaint\_type(s) or circumstances. Only a SME
-from the Taxi \& Limo Commission (TLC) can provide that answer. 
-The R code used here to analyze the 311 Service Request dataset identifies
-both obvious inaccuracies as well as potential errors requiring further 
-investigation. It is not a conclusive list of errors, but rather mostly identifies
-``a place to look'' for further evaluation and potential action.
-
-This paper will focus on a set of specific data cleansing areas:
-
-\begin{itemize}
-	\item Examining structural issues within the data
-	\item Validating data type correctness
-	\item Identifying missing, blank, or N/A data
-	\item Identifying invalid values
-	\item Exposing logical inconsistencies \& inconsistent or 
-	unusual patterns in the data
-	\item Accuracy and precision issues
-	\item Identifying potential redundant or irrelevant data  
-\end{itemize}
-
-\section{Structural Issues}\label{sec:structural}
+is even an ``error''.
+
+\section{Structural Issues}
+\label{sec:structural}
 Structural issues in the context of data cleansing refers to issues 
-related to the how data is organized, formatted, or structured within a dataset. Structural issues 
-can make it difficult to analyze the data effectively. Some common 
-structural concerns in this 311 SR 2022-2023 dataset include:
-
-\begin{itemize}
-	\item Data structure (data fields, columns, formats, data types, etc.) 
-	not corresponding to the Data Dictionary
-	\item Verification of correct data types (numeric, character, geospactial, dates, etc.)
-	\item Embedded or combined values: data elements that 
-	contain multiple pieces of information 
-\end{itemize}
-
-Here are some characteristics of the 311 SR data set:
+related to the how data is organized, formatted, or structured within a 
+dataset. Structural issues can make it difficult to analyze the 
+data effectively. Here are some characteristics of the 311 SR data set:
 
@@
 	\item There are four date fields (created, closed, updated, due).
-	\item There are three borough fields; two of which we believe to be duplicates.
+	\item There are three borough fields; two of which appear to be duplicates.
 	\item Two zip code fields, but not duplicates
-	\item Seven street fields; one pair of which we believe to be duplicates
-	\item Agency name and Agency abbreviation
+	\item Seven street fields; one pair of which appear to be duplicates
 	\item Two Police Precinct fields; not duplicates
-	\item In addition to street/city, there are three other location fields: 
-	lat/long, NY State plane, and Block \#
-	\item incident\_address and street\_name, e.g. incident\_address, 
-	25 Grymes Hill Road and street\_name (Grymes Hill Road)
+	\item In addition to incident\_address, there are five additional location fields: 
+	lat/long, street\_name, landmark, NY State plane, and Block \#
 	\item One free-form text field, resolution\_description, which 
@@
  as ``@computed\_region\_xxxx\_xxxx''. To some extent one can infer 
- what the these @computed\_fields are, but it is not possible to know 
- for certain. These six fields are:
+ what the these @computed\_fields are and how they are determined, but 
+ it is not possible to know for certain. Most importantly, can these fields 
+ be reliably used for analytical purposes?  These six fields are:
 
@@
 
-Here is a screenshot of the NYC Open Data Portal ``Column Manager'' 
-page showing those columns:
-
-\begin{figure}[tbp]
- 	\centering
- 	\includegraphics[width = .65\textwidth] {computed_fields_jpg.jpg}	
-	\caption{Screenshot of Data Portal with ``@computed'' Columns}  
-	\label{fig:computed-columns}
-\end{figure}
-
-This, of course, raises several questions:
-
-\begin{itemize}
-	\item How are these fields computed? Using what source data? 
-	Using what computational method?
-	\item How can these fields be verified or cross-checked?
-	\item Why are there two fields that appear redundant ``police\_precinct' 
-	and ``police\_precincts''? Are these redundant?
-	\item What are ``borough boundaries''? They appear to be borough names. 
-	Why are they referred to as ``boundaries''?
-	\item What are ``community districts''? Is this the same as the
-	 ``community\_board(s)'' that are part of the NYC government?
-	\item Why are not these fields included in the official 311 SR Data Dictionary?
-	\item Can these fields be reliably used for analytical purposes? 
-\end{itemize}
-
-The last bullet point is the key here. Are these fields reliable and accurate, 
-and can they be used for subsequent analysis? Unfortunately,
-as of this date, those questions have not been fully addressed by the 
-NYC Open Data Team. This paper will further address that issue by
-exploring the accuracy, validity, and candidacy for subsequent analytical usage.
-
-\section{Validating data types}\label{sec:datatypes}
+\section{Validating data types}
+\label{sec:datatypes}
 Fortunately, both the Data Dictionary and the portal column manager 
-do a good job of indicating the data type. As indicated by 
-the red underline in Figure \ref{fig:computed-columns}, there 
-is a small icon next to the field name which indicates the 
-data type of the field. The stylized capital ``T'' indicates ``text'', the ``\#'' sign indicates 
-``numeric'', a calendar icon indicates a ``date'' field, and the pin on a 
-map symbol indicates a ``geospatial'' field. The following fields were 
+do a good job of indicating the data type. The following fields were 
 checked for their specified data types:
@@
 
-All data fields were subjected to a test of ``correctness'' 
-according to their ``type'' as specified
-in the Data Dictionary. Of the nearly 300 million data fields to check 
-(6.5 million rows, 47 fields per row), there were only two values that failed 
-the test for appropriate data type; those being two text entries in the 
-incident\_zip field (``na'', and ``N/A''). All date fields were dates. 
-All numbers were numeric. All text fields were character. 
-Data typing appears to not be an issue with this dataset.
-
-\section{Blank \& N/A data}\label{sec:blanks}
+\section{Blank \& N/A data}
+\label{sec:blanks}
 Understanding the absence of data by field is an important factor 
@@
 closed before or after their due\_date, you would be challenged 
-as 99.57\% of the due\_date field is blank. (Although with such a large dataset
-that still leaves 24,411 values, all of which are ``N/A'. Which calls 
-into question...are these missing values, or is the concept of a
+as 99.57\% of the due\_date field is blank. Which calls into 
+question...are these missing values, or is the concept of a 
 due\_date simply not applicable for the vast majority of SRs?
@@
 complaint\_type, agency, and status. In some cases it may make 
-sense to inquiry as to why some fields are frequently blank. Is the data 
-difficult to capture? Does it pertain to only a small set of 
-complaint\_type(s) or Agencies? And if the data is truly that sparse, 
-does it make sense to collect it? Is it legacy data? For example, 
-one would expect that every complaint\_type that has a geographical 
-component, such as an address, would also have a corresponding 
-police\_precinct, community\_board, and community\_district since these
-are all-inclusive throughout the five Boroughs of NYC. Addtionally, 
-that it would have a lat/long. And generally that is the case, but these fields 
-do have small number, but nonetheless significant, of empty entries.
-
-During the analysis, it was determined that the following fields are mandatory, 
-and any row missing these fields should be removed from the dataset (which fortunately
-very rarely occurred): created\_date, agency, complaint\_type, and unique\_key. 
-The absence of any one of these four fields prohibits further analysis  as currently 
-conducted. Again, it is very rare occurrence, happening only 1-2 times 
-in 6.4 million records. Here is a breakdown of bank and N/A entries by field.
-
-\begin{table}[tbp]
-\centering
-\caption{Blank and N/A entries by Field}
-\resizebox{0.8\linewidth}{!}{%
-	\begin{tabular}{lrrrl}
-		\toprule
-		\textbf{Field} & \textbf{Total Empty} & \textbf{Percent Empty} 
-		& \textbf{N/A Count} \\
-		\midrule
-			taxi\_company\_borough & 6391341 & 99.94 & 0 \\
-			road\_ramp & 6379080 & 99.75 & 0 \\
-			vehicle\_type & 6376266 & 99.71 & 0 \\
-			due\_date & 6370497 & 99.62 & 6370497 \\
-			bridge\_highway\_direction & 6367626 & 99.57 & 0 \\
-			bridge\_highway\_name & 6340197 & 99.14 & 0 \\
-			bridge\_highway\_segment & 6340190 & 99.14 & 0 \\
-			taxi\_pick\_up\_location & 6329954 & 98.98 & 0 \\
-			facility\_type & 5966588 & 93.30 & 0 \\
-			landmark & 2635736 & 41.22 & 0 \\
-			intersection\_street\_1 & 2143847 & 33.52 & 0 \\
-			intersection\_street\_2 & 2140707 & 33.48 & 0 \\
-			cross\_street\_1 & 1846594 & 28.88 & 0 \\
-			cross\_street\_2 & 1846008 & 28.87 & 0 \\
-			location\_type & 798775 & 12.49 & 0 \\
-			bbl & 768109 & 12.01 & 0 \\
-			city & 332086 & 5.19 & 0 \\
-			street\_name & 280463 & 4.39 & 0 \\
-			incident\_address & 280268 & 4.38 & 0 \\
-			closed\_date & 248839 & 3.89 & 248839 \\
-			resolution\_description & 132333 & 2.07 & 0 \\
-			zip\_codes & 123167 & 1.93 & 0 \\
-			borough\_boundaries & 101453 & 1.59 & 0 \\
-			community\_districts & 101445 & 1.59 & 0 \\
-			city\_council\_districts & 101445 & 1.59 & 0 \\
-			police\_precincts & 101445 & 1.59 & 0 \\
-			police\_precinct & 101416 & 1.59 & 0 \\
-			latitude & 99778 & 1.56 & 0 \\
-			longitude & 99778 & 1.56 & 0 \\
-			location & 99778 & 1.56 & 0 \\
-			x\_coordinate\_state\_plane & 99668 & 1.56 & 0 \\
-			y\_coordinate\_state\_plane & 98822 & 1.55 & 0 \\
-			resolution\_action\_updated\_date & 91399 & 1.43 & 91399 \\
-			incident\_zip & 83220 & 1.30 & 2 \\
-			descriptor & 74645 & 1.17 & 0 \\
-			address\_type & 38655 & 0.60 & 0 \\
-			unique\_key & 0 & 0.00 & 0 \\
-			agency & 0 & 0.00 & 0 \\
-			agency\_name & 0 & 0.00 & 0 \\
-			complaint\_type & 0 & 0.00 & 0 \\
-			status & 0 & 0.00 & 0 \\
-			community\_board & 0 & 0.00 & 0 \\
-			borough & 0 & 0.00 & 0 \\
-			open\_data\_channel\_type & 0 & 0.00 & 0 \\
-			park\_facility\_name & 0 & 0.00 & 0 \\
-			park\_borough & 0 & 0.00 & 0 \\
-			created\_date & 0 & 0.00 & 0 \\
-		\bottomrule
-	\end{tabular}
-}
-\end{table}
-
-Here is a graphic depiction of total empty (blank \& N/As) for each fields. 
-You can see the natural grouping into the Most, Some, and Few categories 
+sense to inquiry as to why some fields are frequently blank. Here is a graphic 
+depiction of total empty (blank \& N/As) for each fields. You can see 
+the natural grouping into the Most, Some, and Few categories 
 in the stair-type visualization.
@@
 
- \section{Validating Data for Acceptable Values}\label{sec:domain}
- While the dataset may have fields populated with data that appears correct, 
- it is necessary to inspect those fields to determine if the data is indeed valid. 
- For example a date field that contains the value ``February 30, 2024'' is 
- clearly not valid. Nor is a latitude of +95 deg north.
-
- In this study, testing for invalid values revealed several serious issues 
- associated with select fields. Accordingly, an analyst would need to be 
- extra diligent if using these fields for analysis. In all likelihood, 
+ \section{Validating Data for Acceptable Values}
+ \label{sec:domain}
+ In this study, testing for invalid values revealed several issues 
+ associated with select fields. Accordingly, an analyst needs to be 
+ extra diligent if using these fields for analysis, and in all likelihood, 
  rows with invalid field values would  need to be removed from the 
- dataset before conduction further analysis. Here are some results.
-
-\begin{itemize}
-	\item Latitude and Longitude fields were tested to ensure all fell 
-	within the geographic boundaries of the City of New York. All did.
+ analysis. Here are some results.
+
+\begin{itemize}
+	\item Latitude and Longitude fields fell within the geographic 
+	boundaries of New York City. 
 	\item The unique\_key field was in fact unique.
-	\item Many fields have a domain of acceptable values. Often these 
-	values are determined from common usage or by examining
-	larger historical datasets. Unfortunately, these domains of acceptable 
+	\item Many fields have a domain of acceptable values, often 
+	determined from common usage or by examining historical 
+	datasets. Unfortunately, these domains of acceptable 
 	values are not specified in the Data Dictionary. These fields all tested as 
-	compliant within their domain of acceptable values as determined by 
-	examining a 10-year dataset:
+	compliant within their domain:
 	\begin{itemize}
@@
 	 (two fields: zip\_codes and incident\_zip) should be valid as defined 
-	 by the USPS database which contains 37,946 valid zip codes.
-
-	The computed field zip\_codes proved especially problematic with 
-	58\% (3.6 million) of the field entries being invalid. That high number 
-	indicates that the computation of this field is highly inaccurate. We 
-	recommend dropping this field for future analytic efforts.
-
-	Furthermore, the field incident\_zip while having only .07\% invalid entries, 
-	still is a problem with 4163 such errors. As the zip code is a primary 
-	measure of many NYC city services, and with a freely available 
-	database for which to validate entries, it seems an oversight 
-	that any such errors could creep into the 311 system. 
-	Some invalid entries can clearly be identified as incorrect by 
-	observation, e.g. 10000, 12345, 11111, etc.
-
-	The breakdown of the invalid entries in the zip\_code, sorted by 
+	 by the USPS database which contains 37,946 valid zip codes. However,
+	 the computed field zip\_codes proved problematic with 
+	58\% (3.6 million) of the field entries being invalid. Even the 
+	field incident\_zip while having only .07\% invalid entries, exhibited
+	4163 such errors. The breakdown of the invalid entries in the zip\_code, sorted by 
 	Agency shows that the distribution by percentage almost precisely 
-	mirrors the overall breakdown of SRs by Agency as shown in . 
-	This indicates a systemic problem, and since the zip\_code fields is 
-	one of the computed fields, it appears these errors are caused by 
-	incorrect computations rather than by any Agency mistake.
-
-	The  errors in the incident\_zip field are more troubling, even though 
-	they are, percentage-wise, small. Here is a graphic illustrating how 
-	these errors occur by Agency. Here we can see that the majority of 
-	the errors lie in the Dept of Transportation, NYPD, Taxi \& Limo 
-	Commission (TLC), and the Economic Development Council (EDC) 
-	among others. This representation does not follow the full SR 
-	distribution indicating that these errors are likely generated by 
-	an incorrect process or application residing at those particular 
-	Agencies; it is not a systemic problem.
-
-	\begin{figure}[tbp]
-		\centering
-		\includegraphics[width=\textwidth]{invalid_incident_zip.pdf}
-		\caption{Invalid incident\_zip(s) by Agency} 
-		\label{fig:invalid_incident_zip}
-	\end{figure}
+	mirrors the overall breakdown of SRs by Agency, potentially indicating
+	a systemic problem.
 
@@
 		know ``What are the top 10 zip codes for Noise Complaints (all 8 types) 
-		over the last two years?'' The goal is to assess the impact of the 
-		recent NYC effort looking to promote a safe and vibrant 
-		nightlife scene in NYC while seeking to ease strained 
-		relations between bar and club owners.
-
-		It may come as a surprise to non-NYC residents that dancing at 
-		bars and restaurants has been illegal since 1926 (during prohibition), 
-		and was only just repealed in 2018.
-
-		On the surface this would seem like a simple effort.  The NYC Open 
-		Data Portal allows for selection of a timeframe (2022-2023), 
-		complaint-type begins with ``Noise'', and group by Zip code, 
-		sorted descending by count. You can do this analysis and 
-		grouping right inside the NYC Open Data Portal. \textit{Voila!}
-
-		However, this analysis uses the zip\_codes field, one of the (six) 
+		over the last two years?'' The intent is to assess the impact of the 
+		recent NYC effort promoting a vibrant nightlife 
+		in NYC while seeking to ease strained relations 
+		between bar and club owners. Let's assume that this analysis 
+		uses the zip\_codes field, one of the (six) 
 		computed fields that has shown to have validity problems. 
-		If we repeat the analysis with the incident\_zip field instead 
-		of the zip\_codes field, we obtain a different set of zip code results:
-
-		Let's subject these two data sets to validation against 
-		the US Postal Service zip code database.
+		We then repeat the analysis with the incident\_zip field instead. Let's 
+		subject these two data sets to validation against the USPS zip code database.
 
@@
 zip\_codes field were to be 	presented to the Director, ONL, the conclusion 
-would be in error. Perhaps more curious than the differing zip codes 
-on the two lists, is the fact that the numerical counts are nearly identical; 
-the overall counts are quite close. This is curious since one field is 
-computed while the other is via data entry; the computed protocol is 
-again the suspect. The algorithm appears to be counting corrrectly 
-(or nearly so) 	but mis-identifying the corresponding zip code.
+would be in error.
 
@@
 	police precincts; they are generally numeric or have numeric 
-	representations in the dataset.
-
-	However, what we find is that both fields police\_precinct 
-	and police\_precincts  have  35\% invalid entries. Unfortunately, 
-	they're not the same invalid entries.
+	representations in the dataset. However, what we find is that 
+	both fields police\_precinct and police\_precincts  have  35\% 
+	invalid entries. Unfortunately, it's not the same invalid entries.
 
@@
 	police\_precinct field, there are 2,171,778 invalid entries 
-	(also 35\%), however they are not the same counts for the two similar fields.
-
-	The top ten (by count) of invalid precincts for each field are:
-
-	\begin{table}[tbp]
-	\centering
-	\caption{Comparison of the fields police\_precinct and police\_precincts}
-		\begin{tabular}{ccccc}
-		\toprule
-		\textbf{Invalid Precinct} & \textbf{police\_precinct count} 
-		& \textbf{police\_precincts count} & \textbf{Delta}\\
-		\midrule
-				62 & 151808 & 151720 & 88 \\
-				72 & 148881 & 148796 & 85 \\
-				67 & 133171 & 133115 & 56 \\
-				64 & 111236 & 111221 & 15 \\
-				60 & 105252 & 105317 & -65 \\
-				65 & 103334 & 103306 & 28 \\
-				53 & 100097 & 100074 & 23 \\
-				73 & 97746 & 97755 & -9 \\
-				68 & 94204 & 94254 & -50 \\
-				66 & 93654 & 93848 & -194 \\
-		\bottomrule
-		\end{tabular}
-	\label{tab:comparison-precincts-diff}
-	\end{table}
-
-	A graphic representation of the invalid police\_precincts plotted by 
-	Agency show a near perfect match to the distribution of the entire 
-	6.4 million SRs as seen in Figure \ref{fig:SR_counts_by_Agency}. Note 
-	that the ``big six'' Agencies account for 90\% of these invalid precinct 
-	errors, identical to the overall distribution of all SRs. (The same is true 
-	for a similar graph of by the police\_precint field.) This strongly 
-	suggests that this is a systemic problem, most likely caused by the 
-	computation method applied for these two fields.
-
-	\begin{figure}[tbp]
-		\centering
-		\includegraphics[width=\textwidth]{invalid_police_precincts.pdf}
-		\caption{Invalid NYPD precincts by Agency}	  
-		\label{fig:invalid_police_precincts_zip}
-	\end{figure}
+	(also 35\%).
 
@@
 	most local, grassroots form of City government, and a vital connection 
-	between communities and elected officials and City agencies. 
-	They are organized by the five Boroughs of NYC and represented	
-	in this dataset as ``\#\#-Borough'', e.g. ``10 Bronx'' and report to the 
-	various Borough Presidents in NYC. As such, they play an important 
-	role in the quality of life for all New Yorkers. There are 59 community 
-	boards throughout the City: 12 Community Districts in Bronx, 18 in Brooklyn, 
-	12 in Manhattan, 14 in Queens and 3 in Staten Island. The Community 
-	Boards are used frequently as a way to measure distribution of 
-	City services throughout the five boroughs, so correctness is important. 
-	NYC Borough Presidents in particular are focused on the services 
-	offered at the Community Board level within their respective Borough.
-
-	In the 2022-2023 dataset there are 27,276 invalid community\_board 
-	entries which represents 0.43\% of non-blank data. There are a total 
-	of 12 different invalid community boards, to include:
-
-		\begin{table}[tbp]
-		\centering
-		\caption{Top Ten Invalid 'community\_board' Values}
-			\begin{tabular}{rlr}
-			\toprule
-			\textbf{Rank} & \textbf{Invalid CB} & \textbf{Count} \\
-				\midrule
-					1 & 64 MANHATTAN & 7144 \\
-					2 & 83 QUEENS & 5133 \\
-					3 & 55 BROOKLYN & 3327 \\
-					4 & 81 QUEENS & 3180 \\
-					5 & 80 QUEENS & 2407 \\
-					6 & 26 BRONX & 2364 \\
-					7 & 28 BRONX & 1128 \\
-					8 & 82 QUEENS & 1054 \\
-					9 & 95 STATEN ISLAND & 515 \\
-					10 & 27 BRONX & 514 \\
-				\bottomrule
-			\end{tabular}
-		\end{table}
-
-		The distribution of invalid Community Boards by Agency is not consistent 
-		with the overall SR Agency distribution. This indicates that there are likely 
-		specific issues at some key Agencies, in this case Taxi \& Limo Commission 
-		(TLC), Parks \& Recreation, etc. It's not a large error, but the 
-		division of services (and complaints) by Community Board is a 
-		well-tracked measure so accuracy is again important.
-
-		\begin{figure}[tbp]
-		 	 \centering
-			  \includegraphics[width = \textwidth]{invalid_community_boards.pdf}
-			  \caption{Invalid Community Board by Agency}
-			  \label{fig:invalid_community_boards}
-		\end{figure}
+	between communities and elected officials and City agencies. Community 
+	Boards are used as a measure of City services throughout the five 
+	boroughs. In the 2022-2023 dataset there are 27,276 invalid community\_board 
+	entries which represents 0.43\% of non-blank data. The distribution 
+	of invalid Community Boards by Agency is not consistent with 
+	the overall SR Agency distribution. This indicates that there are likely 
+	specific issues at some key Agencies, in this case Taxi \& Limo 
+	Commission (TLC), Parks \& Recreation, etc.
 
@@
 \label{sec:communitydistrict}
-	Another of the computed fields is community\_districts. Community Districts 
+	Another of the ``computed'' fields is community\_districts. Community Districts 
 	are the boundaries for the Community Boards, but unlike how 
@@
 	(DCP) for purposes of environmental, socio-economic, and demographic 
-	purposes. DCP is the NYC''s primary land use agency and is 
-	instrumental in the City's physical framework.  It is a geographical 
-	division rather than a local government division. And like Community 
-	Boards, the Community Districts are frequently cited as a means to 
-	track zoning, housing, community facilities, and waterfront/open/public spaces. 
-	It is another of the NYC well-tracked measures.
-
-	Due to how the community\_district data is formatted, it is not possible 
-	to establish validity. However, it is possible to determine that the 
-	dataset contains 72 unique entries, while there are only 59 valid Community 
-	Districts. So it would appear that there are 13 invalid 
-	community\_board(s) contained in the dataset.
-
-	As previously mentioned, the dataset has four data fields:  created\_date, 
-	closed\_date, due\_date, and resolution\_action\_update\_date. 
-	The due\_date field has few entries to analyze; 99.62\% of due\_date entries 
-	are ``N/A''.  It is not known if the due\_date(s) are truely ``not applicable'' 
-	or that the due\_date field is rarely used. The  resolution\_action\_update\_date 
-	field serves as a date-time marker showing when the SR record was last updated.  
-	All the date fields are recorded to the second, e/g/ ``2022-02-11 13:43:05''.
+	purposes. It is a geographical division rather than a local 
+	government division. Due to how the community\_district data 
+	is formatted, it is not possible to establish validity. However, 
+	it is possible to determine that the dataset contains 72 unique 
+	entries, while there are only 59 valid Community Districts.
 
@@
 	expect that these two fields are automatically populated by the 
-	SR application software, either the software used at the 311 organization 
-	or that used at the Agencies that are integrated with the 311 system; 
-	and typically that is indeed the case. In such a case, the act of creating an 
-	SR would automatically trigger an entry in created\_date based on a 
-	system reference clock. Thus the created\_date is associated with 
-	setting an SR status to ``new'' or ``open''. And similarly when an SR's 
-	status is changed to ``closed'' then the closed\_date would be populated.
-
-	Unfortunately, our analysis found that that assumption is not always 
-	the case. When you analyze the various date fields, several anomalies 
-	quickly become apparent:
-
-	\begin{itemize}
-		\item Some SRs have a closed\_date that occurs before the created\_date.
-	 	\item Some created\_date(s) and closed\_date(s) appear either in 
-	 	the future or in the far distant past.
-		\item There are a large number of SRs that are closed and created 
-		exactly at midnight or exactly at noon. 
-	\end{itemize}
-
-	\textbf{Why are these date fields important?} They're important because 
-	citizens, NYC Government Officials, and NYC Agencies use these 
-	dates to measure the responsiveness for providing the 
-	services. How long does it take to replace an out-of-order 
-	street light? How rapid is the response to a report of domestic 
-	violence? Does the NYPD response to a noise complaint in 
-	Staten Island take longer than in Manhattan? How long does it 
-	take to repair a street pothole in Queens?
-
-	The measure of ``duration'', of the response time of a 311 call is a closely 
+	SR application software. In such a case, creating an SR would 
+	automatically trigger a created\_date based on a system clock. 
+	Thus the created\_date is associated with setting an SR status 
+	to ``new'' or ``open''. And similarly when an SR's 	status is 
+	changed to ``closed'' then the closed\_date would be 
+	populated. Unfortunately, that is not always the case. There are
+	several anomalies with the various date fields.
+
+	\begin{itemize}
+		\item SRs with a closed\_date that occurs before the created\_date.
+	 	\item created\_date(s) and closed\_date(s) in the far distant past.
+	 	\item created\_date(s) and closed\_date(s) that match to the second.
+		\item A large number of SRs closed and created exactly at midnight 
+		or exactly at noon. 
+	\end{itemize}
+
+	These date fields are important because citizens, NYC Government Officials, 
+	and NYC Agencies use these dates to measure the 
+	responsiveness for providing services. The measure of ``duration'',	 
+	of the response time of a 311 call is a closely 
 	watched, carefully scrutinized metric, both in terms of overall performance 
-	and with a geographical area. And it can have broad political, organizational, 
-	and budgetary consequences. Duration is not a field in the dataset, 
-	but it is easily computed by closed\_date - created\_date. Let's look at 
-	some anomalies in this ``duration'' field.
+	and with a geographical area. Duration, while not a field in the dataset, 
+	it is easily computed by closed\_date - created\_date.
 
@@
 	While this is a small percentage overall (0.2\%) it can have significant impact 
-	on response time analysis. In most cases, these ``negative durations'' 
-	appear to just be mistakes. Here is a sample of some of the 
-	largest and the smallest errors:
+	on response time analysis. Here is a sample:
 
@@
 	negative durations exceeding 44,601 days (122 yrs). That large of an anomaly 
-	can skew statistical results, even though the number of occurrences is small. 
-	This is especially curious as the date ``1900-01-01'' is also the earliest date that 
-	Microsoft Excel can accommodate as well as how Excel treats a blank cell 
-	that is designated as a date. All raising suspicions that perhaps these dates 
-	were imported into the 311 system from Excel.
-
-	According, these  SR rows are removed from the box \& whiskers plot. 
-	Note that these extremely large negative duration SRs are all from the 
-	Department of Homeland Services (DHS) Sample:
-
-	\begin{table}[tbp]
-	    \centering
-	    \caption{Sample of SRs with extremely large negative durations}
-		    \begin{tabular}{l l r l}
-		        \toprule
-		        \textbf{created\_date} & \textbf{closed\_date} & \textbf{duration} 
-		        & \textbf{agency} \\
-			        \midrule
-			        2022-02-11 13:43:05 & 1900-01-01 & -44602 & DHS \\
-			        2022-02-11 11:03:38 & 1900-01-01 & -44602 & DHS \\
-			        2022-02-11 08:29:49 & 1900-01-01 & -44601 & DHS \\
-			        2022-02-11 14:08:08 & 1900-01-01 & -44602 & DHS \\
-			        2022-02-11 14:05:16 & 1900-01-01 & -44602 & DHS \\
-			        \bottomrule
-		    \end{tabular}
-	    \label{tab:extreme_negative_durations}
-	\end{table}
-
-	Here is a graphic visualization of the negative-duration SRs by Agency. It is 
-	obvious that this negative-duration issue is a problem at the 
-	Dept. of Transportation (DOT) where 95\% of these types of errors occur.
+	can skew statistical results, even though the number of occurrences is 
+	small. According, these  SR rows are removed from the analysis. These 
+	extremely large negative duration SRs are all from the 
+	Department of Homeland Services (DHS) A visualization of the 
+	negative-duration SRs by Agency shows that this negative-duration 
+	issue is a DOT problem, where 95\% of these types of errors occur.
 
@@
 \label{sec:homlessassistance}
-	\textbf{Scenario:} The Dept. of Homeless Services wants to 
+	\textbf{Scenario:} The Dept. of Homeless Services (DHS) wants to 
 	know ``How quickly are 311 calls for ``Homeless Person Assistance
-	over the past two years?'' resolved. This is a typical request made by 
+	resolved over the past two years?'' This is a typical request made by 
 	both the public and City government. It's a performance metric along 
@@
 
-	As it turns out the distribution of the ``Homeless Person Assistance'' is 
-	quite skewed due to a small number of long duration SRs. In this case, 
-	the median is a better measure of central tendency than the average. 
+	In this case, the median is a better measure of central tendency than the average. 
 	And the median of the duration field is  0.2 days (approx 5 hrs).
-	Here is a look at the distribution of durations for the Homeless Person 
-	Assistance SRs showing 	the large outliers.
 
@@
   	\includegraphics[width = \textwidth]
-	{non-matchingstreet\_nameandlandmark.pdf}
+  	{non-matchingstreet_nameandlandmark.pdf}
   	\caption{Non-matching street\_name \& landmark by Agency}

### 2024-06-18 code/datacleansing.R

@@ -619,7 +619,7 @@ create_violin_chart <- function(
   violin_chart <- ggplot(data = dataset, aes(x = !!rlang::sym(x_axis_field), y = factor(1))) +
     geom_jitter(width = 0.25, alpha = 0.4, color = "#2271B2", size = 1.9, shape = 17) +
     geom_violin(linewidth = 0.7, fill = "transparent", color = "gray20") +
-    geom_boxplot(width = 0.25, fill = "#E69F00", color = "gray20", alpha = 0.6, 
+    geom_boxplot(width = 0.25, fill = "#D55E00", color = "gray20", alpha = 0.6, 
                  outlier.colour = "gray20", outlier.size = 1) +
     labs(
       title = chart_title,

### 2024-06-20 manuscript/manuscript.tex
went over abstract and intro

@@
 \usepackage{hyperref}\hypersetup{colorlinks=true, citecolor=blue}
-\usepackage{parskip}
+% \usepackage{parskip}
 \usepackage{textgreek}
@@
 
-\newcommand{\jy}[1]{\textcolor{orange}{JY: (#1)}}
-\newcommand{\yh}[1]{\textcolor{purple}{YH: (#1)}}
-\newcommand{\yx}[1]{\textcolor{cyan}{YX: (#1)}}
+\usepackage{xcolor}
+\newcommand{\dt}[1]{\textcolor{purple}{DT: (#1)}}
+\newcommand{\jy}[1]{\textcolor{cyan}{JY: (#1)}}
+
 \let\proglang=\textsf
@@
 In the early 21st century, the open data movement began to transform 
-societies and governments through principles of transparency, 
-innovation, and public engagement. New York City (NYC) has emerged 
-as a leader in this movement with the enactment of the Open 
-Data Law in 2012, leading to the creation of the NYC Open Data 
-portal, which now hosts 2700 datasets from 80 city agencies. This 
-resource has proven invaluable for research across various domains, 
+societies and governments by promoting transparency,
+innovation, and public engagement. New York City (NYC) has been at
+the forefront of this movement since the enactment of the Open 
+Data Law in 2012, which led to the creation of the NYC Open Data
+portal. This portal now hosts 2,700 datasets from 80 city agencies,
+serving as a crucial resource for research across various domains, 
 including health, urban development, and transportation. The 
-success of these initiatives underscores the importance of data 
-curation, ensuring the utility and reliability of datasets.
-
-This paper examines the data curation challenges using the 
-NYC 311 Service Request (SR) Data as a case study, addressing issues 
+success of these initiatives highlights the importance of data 
+curation in ensuring the utility and reliability of open datasets.
+This paper examines the challenges of open data curation through a
+case study of the NYC 311 Service Request (SR), addressing issues 
 of data validity, consistency, and curation efficiency. Based on 
@@
 the ongoing utility of open data. The paper concludes with 
-actionable suggestions for improving data curation and offers 
-general principles for the release of open data.
+actionable recommendations for enhancing data curation and outlines
+general principles for the effective release of open data.
 
@@
 
+\jy{cut down to 6-8}
 \noindent
@@
 \label{sec:intro}
-In the early 21st century, the open data movement began to take shape, driven by the 
+In the early 21st century,
+the open data movement began to take shape, driven by the
 fundamental belief that freely accessible data can transform 
-both societies and governments. This movement champions the principles of 
-transparency, innovation, and public engagement. 
+both societies and governments. This movement champions the principles
+of transparency, innovation, and public engagement. 
 A landmark in this journey was the launch of the United States'
@@
 
-New York City (NYC) has emerged as a forerunner in the open data
-movement, marked by the enactment of the Open Data Law in 2012
+New York City (NYC) has emerged as a leader in the open data movement,
+marked by the enactment of the Open Data Law in 2012
 \citep{zuiderwijk2014open}. This landmark legislation led to the
 creation of the \href{https://opendata.cityofnewyork.us}{NYC Open Data
-  portal}, which today hosts an impressive array of 2700 datasets
+  portal}, which today hosts an impressive array of 2,700 datasets
 from 80 different city agencies. This resource has become invaluable
-for researchers in various fields as well an enabling local government transparency
-
-In health, datasets have enabled
-significant studies on facets of health and healthcare delivery
-\citep{cantor2018facets, shankar2021data}. In the realm of urban
-development, data has been instrumental in advancing smart city
-initiatives \citep{neves2020impacts}. Additionally, transportation
-research has benefited greatly from this wealth of data, aiding in the
-understanding of urban mobility and infrastructure
-\citep{gerte2019understanding}. NYC's Open Data initiative not only
-exemplifies commitment to transparency and public engagement but also
-illustrates how open data can be a powerful tool in addressing complex
-urban challenges. Examples of some of the more popular NYC datasets include:
-
-\begin{itemize}
-	\item NYC restaurant health inspection violations
-	\item Mapping of car crashes involving pedestrians
-	\item Visualization of NYC High School and College enrollment
-	\item An interactive dashboard to filter and view charges for jail inmates
-	\item Location of City-wide free Internet access points
-	\item ...and 220 other complaint types. 
-\end{itemize}
-
-Data curation is fundamental in the open data ecosystem, ensuring the
-utility and reliability of datasets for diverse applications. Among
-the earliest discussions, \citet{witt2009constructing} focus on the
-development of data curation profiles tailored to specific contexts,
-setting a precedent for targeted data management
-strategies. Addressing broader challenges in data sharing and
-management, \citet{borgman2012conundrum} highlights the complexities
-of research data distribution, emphasizing the need for robust
-strategies. This is complemented by the work of \citet{hart2016ten},
-who outline essential principles for effective data management,
-particularly emphasizing the importance of meticulous curation
-practices.
-
-In the realm of collaborative data management,
-\citet{beheshti2019datasynapse} underscore the significance of
-cooperative environments for managing and sharing social data
-effectively. This aspect of data curation gains further relevance in
-the research by \citet{mclure2014data}, which delves into the specific
+for researchers across various fields and has significantly enhanced
+local government transparency. Popular datasets include information on
+restaurant health inspection violations, car crashes, high school and
+college enrollment statistics, jail inmate charges, and the location
+of city-wide free Internet access points. These datasets have been
+applied in civil life in various ways, such as mapping car crashes
+involving pedestrians and visualizing high school and college
+enrollment trends. Furthermore, they have enabled significant research
+across multiple domains, including health \citep{cantor2018facets,
+  shankar2021data}, urban development \citep{neves2020impacts}, and
+transportation \citep{gerte2019understanding}, aiding in the
+understanding and addressing of complex urban challenges.
+
+The civil and research applications of open data critically depend on
+its quality, making data curation fundamental in the open data
+ecosystem. Among the earliest discussions,
+\citet{witt2009constructing} focused on developing data curation
+profiles tailored to specific contexts, setting a precedent for
+targeted data management strategies. Addressing broader challenges in
+data sharing and management, \citet{borgman2012conundrum} highlighted
+the complexities of research data distribution, emphasizing the need
+for robust strategies. This is complemented by \citet{hart2016ten},
+who outlined essential principles for effective data management,
+particularly emphasizing meticulous curation practices. In
+collaborative data management, \citet{beheshti2019datasynapse}
+underscored the significance of cooperative environments for managing
+and sharing social data effectively. This aspect gains further
+relevance in \citet{mclure2014data}, which delved into the specific
 practices and needs within data curation communities. The practical
-implications of data curation are vividly illustrated in the context
-of public health and global challenges. \citet{cantor2018facets}
-demonstrate the utility of curated open data in evaluating community
-health determinants. Furthermore, the COVID-19 pandemic serves as a
-real-world example, with \citet{shankar2021data} observing the
-critical role of collective data curation efforts in managing and
-responding to the crisis. Collectively, these studies not only highlight the 
-multifaceted nature of data curation but also emphasize
-its indispensable role in enhancing the applicability and value of
-open data across various domains.
+implications of data curation are vividly illustrated in public health
+and global challenges. \citet{cantor2018facets} demonstrated the
+utility of curated open data in evaluating community health
+determinants. The COVID-19 pandemic served as a real-world example,
+with \citet{shankar2021data} observing the critical role of collective
+data curation efforts in managing and responding to the
+crisis. Collectively, these studies highlight the multifaceted nature
+of data curation and emphasize its indispensable role in enhancing the
+applicability and value of open data across various domains.
 
@@
 
+\dt{
 A note on naming conventions: dataset field names are named 
@@
 underscores, e.g. created\_date, complaint\_type, zip\_code, etc..   
+}
 

### 2024-06-20 manuscript/manuscript.tex
streamlined data section

@@
 \label{sec:intro}
+
 In the early 21st century,
@@
 The paper is organized as follows:
-Section~\ref{sec:history} offers a brief review of the history of the 311 system. In 
-Section~\ref{sec:trends} we take a look at long-term trends presented 
-via a 10-year analysis. Section~\ref{sec:issues} offers a 
+Section~\ref{sec:data} offers a brief review of the history of the 311
+system and the long-term trends presented via a 10-year analysis.
+Section~\ref{sec:issues} offers a 
 general discussion of data cleansing issues impacting data 
@@
 
-\section{NYC 311 System History} 
-\label{sec:history}
+\section{NYC 311 Service Request Data} 
+\label{sec:data}
+
 The NYC 311 service, a critical component of New York City's public
@@
 non-emergency issues, ranging from noise complaints to street
-maintenance requests. The evolution of this system can be traced from
-its initial implementation as a simple inquiry channel via telephone to a
-comprehensive data management system that handles millions of requests
-annually. Each inquiry is logged as a complaint, a Service Request (SR), 
-in the 311 system. Key milestones in the 311 system include:
-
-\begin{itemize}
- 	\item 2003 - NYC 311 system goes live with a phone-based call center only. 
- 	The ``311'' phone number replaces a myriad of numbers to 
- 	call for each individual City Agency.
-   	\item 2009 - 311 online \& mobile apps launched. Today 60\% of all 
-   	Service Requests come from mobile \& online channels.
-      \item 2020 - In August 2020, driven by the COVID pandemic, monthly 311 
-      SRs hit an all-time high of 348,463. Seven of the Top 10 all-time busiest 
-      days occur in 2020.
-      \item 2021 - 311 System is expanded to include the MTA's city subway system, 
-      the largest expansion since its inception.
-   	\item 2023 - Record high year for 311 with 3.23 million Service Requests.
-\end{itemize}
-
-Today, the NYC 311 data system is a robust platform that manages a
-vast array of urban living-related inquiries. The system handles over 3 million service 
-requests (SRs) per year, encompassing such complaints as street noise, 
-illegal parking, heat/hot water, abandoned vehicles,
-and unsanitary conditions. The data is a valuable resource for City
-administration and policy-making. The 311 data is publicly accessible through the 
-NYC Open Data Portal which provides access to all the data 
-sets (approx. 2700) and provides a convenient method for querying, 
-grouping, aggregating, geo-mapping, visualization, and exporting 
-results. This open data  enables not only governmental transparency 
-but empowers researchers, civic developers, and the general public.
-
-Despite its success, the 311 system faces challenges such as:
-
-\begin{itemize}
-	\item Data timeliness, accuracy, and consistency
-	\item Difficulties correlating data over long time periods, e.g. over a 10-year 
-	period, e.g. Agency name changes, new complaint types.
-	\item Data anonymization and handling of Personally identifiable information (PII)
-	\item Integration with stand-alone systems at selected NYC Agencies
-	\item Managing API usage, authentication, and load for 3\textsuperscript{rd} 
-	party users, of which there are many
-\end{itemize}
-
-These and other challenges are addressed by the various 
-Agency open data managers as well as the 
-\href{https://www.nyc.gov/content/oti/pages/}
-{NYC Office of Technology and Innovation (OTI)} which provides the 
-technology support for the open data system. Typically, the data 
-will be exported and processed in an external program. The authors found 
-this to be the case with the large 311 dataset examined.
-
-The impact of the NYC 311 data extends beyond operational efficiency;
-it has become instrumental in shaping City governance and community
-engagement. The data has been pivotal in such areas as:
-
-\begin{itemize}
-	\item Providing advice on shelters during hurricanes, heat emergencies, and winter blizzards
-	\item Handling countless inquires during the COVID pandemic 
-	(testing centers, vaccine sites, mobile vaccine clinics, etc.)
-	\item Enforcement of standards between landlords and tenants
-	\item Re-allocation of NYC's yellow taxi routes based upon an insightful analysis 
-	conducted by the Taxi \& Limousine Commission (TLC)
-	\item Improved responsiveness for City Agencies with regards to direct 
-	client-facing responsibilities such as Parks \& Rec, Dept of Transportation, DSNY, etc.
-	\item Increased 311 SRs related to Homelessness resulted in the launch 
-	of the City's Homeless Outreach \& Mobile Engagement Street Action Teams (HOME-STAT)
-	which includes street canvassers and increases in homeless housing facilities.
-\end{itemize}
-
-We extracted 311 SR data for the period of CY 2022-2023, a dataset of
+maintenance requests. Initially a phone-based call center, the system
+evolved into a comprehensive data management platform handling
+millions of requests annually. Key milestones since its launch in 2003
+include the addition of online and mobile app channels in 2009, a
+record high of 348,463 monthly service requests in August 2020 due to
+the COVID-19 pandemic, the 2021 expansion to include the MTA's city
+subway system, and a record 3.23 million service requests in
+2023. Today, the NYC 311 data system manages over 3 million service
+requests per year, covering issues such as street noise, illegal
+parking, heat and hot water, abandoned vehicles, and unsanitary
+conditions. This data is publicly accessible through the NYC Open Data
+Portal, which provides tools for querying, grouping, aggregating,
+geo-mapping, visualizing, and exporting results.
+
+Despite its success, the 311 system faces several challenges, such as
+data timeliness, accuracy, and consistency, difficulties in
+correlating data over long periods, handling of personally
+identifiable information (PII), integration with stand-alone systems
+at selected NYC agencies, and managing API usage and load for numerous
+third-party users. These challenges are addressed by various agency
+open data managers and the
+\href{https://www.nyc.gov/content/oti/pages/}{NYC Office of Technology
+  and Innovation (OTI)}, which provides technological support for the
+open data system. Data often needs to be exported and processed in
+external programs to address these issues, as observed in the
+examination of the large 311 dataset. Furthermore, the system must
+adapt to changes such as agency name changes and the introduction of
+new complaint types, which can complicate long-term data analysis.
+
+The impact of NYC 311 data extends beyond operational efficiency; it
+has become instrumental in shaping city governance and community
+engagement. This open data not only ensures governmental transparency
+but also empowers researchers, civic developers, and the general
+public. The data has been pivotal in providing advice on shelters
+during emergencies, handling inquiries during the COVID-19 pandemic,
+enforcing standards between landlords and tenants, reallocating taxi
+routes based on analyses by the Taxi and Limousine Commission (TLC),
+improving responsiveness for city agencies like Parks and Recreation,
+the Department of Transportation, and the Department of Sanitation,
+and addressing homelessness through initiatives like the Homeless
+Outreach and Mobile Engagement Street Action Teams (HOME-STAT). These
+applications demonstrate the transformative potential of open data in
+enhancing urban living and fostering a more responsive and transparent
+city government.
+
+\jy{FixME: this paragraph's location will be revisited later.}
+We extracted 311 SR data for the period of CY 2022--2023, a dataset of
 6.4 million SR records. This is a large dataset which enables 
@@
 
-\section{10-year Analysis \& Trends} 
-\label{sec:trends}
-As part of this study, we also looked at a 10-year period (2014-2023) to
-observe system performance over a longer time-frame . During that time-frame, 
-the number of 311 SRs grew by 50\%. The summer and early fall months 
-make up all of the Top 5 SR counts by day. The all-time daily 
-high for SRs, 24,415 SRs (a full 9.8 \textsigma's from the 10-year mean), 
-occurred on Tuesday, 2020-08-04. The 10-year low count occurs 
-on Christmas Day, 2016, a Sunday.
-
-\begin{figure}[tbp]
-	\centering
-  	\includegraphics[width=\textwidth]{10-year-trend-monthly.pdf}
+\jy{Every float (figure/table) needs to be referred to in the text.}
+
+\begin{figure}[tbp]
+	\centering
+  	\includegraphics[width=0.8\textwidth]{10-year-trend-monthly.pdf}
   	\caption{10-year (2014-2023) Monthly SR Counts}
@@
 
+We also looked at a 10-year period (2014-2023) to
+observe system performance over a longer time-frame; see
+Figure~\ref{fig:10-yr-monthly}. During that time-frame,
+the number of 311 SRs grew by 50\%. The summer and early fall months 
+make up all of the Top 5 SR counts by day.
+
 \begin{table}[tbp]
@@
     \caption{Top 5 and Bottom 5 Days for SRs}
+    \label{tab:top-10-counts}
     \begin{tabular}{@{}lll lll@{}}
@@
     \end{tabular}
-    \label{tab:top-10-counts}
 \end{table}
 
-Two important fields in this analysis effort are the responsible NYC Agency, 
-and the type of complaint. Given the size of the City of New 
-York government, it is necessary to identify the responsible 
-Agency as well as the type of complaint in order to identify 
-the responsible party and assist in troubleshooting discrepancies.
-Thus we can often identify if it is an Agency specific problem, 
-or a systemic issue. Although 311 SRs involve 16 different agencies, 
-there are six Agencies that comprise 90\% of the complaints.
-
-\begin{itemize}
-	\item New York Police Department (NYPD)
-	\item Housing Preservation \& Development (HPD
-	\item New York City Department of Sanitation (DSNY)
-	\item Department of Transportation (DOT)
-	\item Department of Environmental Protection (DEP)
-	\item Department of Parks \& Recreation (DPR)
-\end{itemize}
-
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width=\textwidth]{SRs_by_Agency.pdf}
+Table~\ref{tab:top-10-counts} summarizes the top 5 and bottom 5 days
+in counts of SRs. The highest daily SR count, 24,415, occurred on
+August 4, 2020, coinciding with the aftermath of Tropical Storm
+Isaias, highlighting the system's critical role during crises. The
+second-highest count, 19,560, was the following day, reflecting the
+storm's continued impact. 
+Conversely, the lowest daily SR count, 2,965, was on Christmas Day,
+2014, with similar lows on other Christmas Days, indicating reduced
+activity during major holidays. The data also shows fewer requests on
+weekends, such as April 20, 2014 (Easter Sunday), and November 6, 2016
+(a non-holiday Sunday).
+
+\begin{figure}[tbp]
+	\centering
+	\includegraphics[width=0.8\textwidth]{SRs_by_Agency.pdf}
   	\caption{SR counts by Agency with Cumulative Percentage}
@@
 
-The distribution of complaint\_type skews to a few select complaints 
-accompanied by a long tail. This graph shows the top 20 complaint\_type(s) 
-comprising 70\% out of a total of 220 different types.
-
-\begin{figure}[tbp]
-	\centering
- 	\includegraphics[width=\textwidth]{SR_by_Complaint_Type.pdf} 
+Figure~\ref{fig:SRcountbyAgency} display the SR counts by agency with
+cumulative percentage. Two important fields in this analysis are the
+responsible NYC Agency and the type of complaint. Given the size of
+the City of New York government, identifying the responsible agency
+and the type of complaint is essential for pinpointing the accountable
+party and troubleshooting discrepancies. This identification helps
+determine whether an issue is specific to an agency or indicative of a
+systemic problem. Although 311 SRs involve 16 different agencies, six
+agencies account for 90\% of the complaints. These are the New York
+Police Department (NYPD), Housing Preservation and Development (HPD),
+New York City Department of Sanitation (DSNY), Department of
+Transportation (DOT), Department of Environmental Protection (DEP),
+and Department of Parks and Recreation (DPR).
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=0.8\textwidth]{SR_by_Complaint_Type.pdf} 
   \caption{Top 20 complaint\_type(s) and Cumulative Percentage} 
@@
 
-The top 20 complaints contain several of the eight different types of ``Noise'' 
-complaints (Residential, Commercial, Street, Helicopter, etc.). These Noise-related
-complaints total a full 22\% of all complaint types, which, when consolidated, is the most frequently occurring complaint\_type.
-
 \begin{table}[tbp]
@@
   \caption{Noise-related complaints\_type(s) by count with Agency}
+  \label{tab:noisecomplaints}
   \begin{tabular}{@{}lS[table-format=7.0,round-mode=places,
@@
   \end{tabular}
-    \label{tab:noisecomplaints}
 \end{table}
+
+The distribution of complaint types skews toward a few select
+categories, accompanied by a long tail. Figure~\ref{fig:SR_complaints}
+shows that the top 20 complaint types account for 70\% of the total
+220 different types. Notably, these top complaints include several
+variations of noise-related issues, which collectively represent 22\%
+of all complaints. When consolidated, noise complaints emerge as the
+most frequently occurring type. Table~\ref{tab:noisecomplaints}
+provides a detailed breakdown of noise-related complaint types by
+count and responsible agency. The most common noise complaint is
+``Noise - Residential,'' with 675,502 complaints, accounting for 10.56\%
+of all service requests, and primarily handled by the NYPD. Other
+significant noise complaints include ``Noise - Street/Sidewalk"
+(4.70\%), ``Noise - Commercial'' (2.02\%), and ``Noise - Vehicle"
+(1.81\%), all managed by the NYPD. Additional categories such as
+"Noise - Helicopter'' (1.33\%) fall under the jurisdiction of the EDC,
+and ``Noise'' (1.57\%) is managed by the DEP. The prevalence of these
+complaints underscores the critical role of the NYPD and other
+agencies in addressing noise-related issues, reflecting the
+significant impact of urban noise on residents' quality of life. This
+data highlights the importance of efficient response and management
+strategies to mitigate noise disturbances in New York City.
 

### 2024-06-21 manuscript/manuscript.tex
Review and editing through Part 10
Doing overall review, editing, and summarizing of text to reduce page size and # of figures/graphics. Making good progress for first sweep through. Have reviewed up to Section 10.

@@
 both fields police\_precinct and police\_precincts  have  35\% 
-	invalid entries. Unfortunately, it's not the same invalid entries.
-
-	For the police\_precincts field, there are 2,171,864 invalid entries 
+invalid entries. Unfortunately, it's not the same invalid entries. For 
+the police\_precincts field, there are 2,171,864 invalid entries 
 (35\%)., a consequential number of errors. Similarly for the 
@@
 \label{sec:negativeduration}
-	Let's look first at the created\_date and closed\_date fields. You might 
+For the created\_date and closed\_date fields, you might 
 expect that these two fields are automatically populated by the 
-	SR application software. In such a case, creating an SR would 
-	automatically trigger a created\_date based on a system clock. 
-	Thus the created\_date is associated with setting an SR status 
+SR application software. Thus the created\_date is associated with setting an SR status 
 to ``new'' or ``open''. And similarly when an SR's 	status is 
-	changed to ``closed'' then the closed\_date would be 
-	populated. Unfortunately, that is not always the case. There are
-	several anomalies with the various date fields.
+changed to ``closed''  the closed\_date would be 
+populated. Unfortunately, that does not appear to be the case as
+there are several anomalies with the various date fields.
 
@@
 the extremely large negative-duration SR removed. While there are few 
-	outlier points, the magnitude of the negative-durations is troubling and can 
+outliers, the magnitude of the negative-durations is troubling and can 
 produce bizarre analytical results.
@@
 \label{sec: zeroduration}		
-	An even more serious, and more prevalent problem with closed\_date(s) 
+A more serious and prevalent problem with closed\_date(s) 
 occurs when the closed\_date and created\_date are exactly the same -- to 
 the second -- which accordingly creates a \textbf{zero duration'}. Again 
-	this is nonsensical, but the presence of such zero durations can again 
+this is nonsensical, and the presence of such zero durations can again 
 severely distort analytics. There are 191,141 such SRs representing 3.1\% of 
-	all non-blank data; not an insignificant number. As shown in the chart 
-	below, 99\% of the zero duration SRs appear to predominately occur in five 
+all non-blank data. As shown in the chart 	below, 99\% of the 
+zero duration SRs appear to predominately occur in five 
 Agencies (Dept of Mental Health \&Hygiene, Dept of Transportation, 
@@
 \label{sec:midnightandnoon}
-	An additional problem discovered with the created\_date and 
-	closed\_date fields; there appears to be an unusually large number of SRs 
+Another problem discovered with the created\_date and 
+closed\_date fields - there appears to be an unusually large number of SRs 
 created or closed at exactly midnight (00:00:00) and exactly noon (12:00:00), 
@@
 work-day clock with many SRs created during day-light hours, and 
-	fewer SRs 	created at night and in the early hours of the morning. However, 
+fewer SRs 	created at night and the early hours of the morning. However, 
 here there appears to be significantly greater numbers of SR closed exactly 
 at midnight and exactly at noon, as well as a significant number of SRs 
-	created exactly at noon. The reason behind this anomaly is worth investigating as 
-	it affects the SR ``duration'' question. If SR durations are used to measure 
-	responsiveness for the delivery of services, and if those times are not an 
-	accurate reflection of the start/finish of the Service, then the use of duration 
-	as a service quality metric is called into question. Observations during 
-	this 2-year period include:
+created exactly at noon. Observations during this 2-year period include:
 
@@
 
-	There is, of course, a broad distribution of the time that SRs are 
-	created by day and hour, by night and day, by weekday and weekend.  
-	One would think that that distribution would be (mostly) evenly distributed 
-	by the minute and second. That is, if there are normally 100 SRs created 
-	between 1-2pm, that the distribution to the minute and second would be 
-	close to random. For example, there is no reason to suspect that a larger 
-	number of SRs would be created at 6 minutes past the hour rather 
-	than at 5 minutes past the hour. And certainly, if you look at the second 
-	that these SRs are created you would again expect a (more-or-less) 
-	random distribution; that there would be no reason to believe that a higher 
-	(or lower) number of SRs would arrive at 18 seconds past the minute 
-	versus at 39 seconds past the minute. Unfortunately, that turned out not to be true.
-
-	Let's look at the SR creation distribution. We discovered that there 
-	are a significant number of SRs created exactly at noon. To reveal this trend, 
-	we looked at the exact date-time stamp of the created\_date(s). We began 
-	by selected only SRs that were created exactly on the minute; at 09:00:00 
-	or 13:00:00 for example.  Here is a visualization of the number of SRs 
-	created and closed by hour-of-the-day:
-
-	\begin{figure}[tbp]
-		\centering
-		\includegraphics[width = \textwidth]{Created_Hourly_SR_count.pdf}
-		\caption{SRs by Hourly created\_date(s)}
-		\label{fig:hourly-created}
-	\end{figure}
-
-	\begin{figure}[tbp]
-		\centering
-		\includegraphics[width = \textwidth]{Closed_Hourly_SR_count.pdf}
-		\caption{SRs by Hourly closed\_date(s)}
-		\label{fig:hourly-closed}
-	\end{figure}
-
-	The created\_date shows a maximum value at 12:00. The closed\_date chart 
-	show significant spikes at both midnight and noon. The midnight (00:00) 
-	spike appears to be an anomaly especially as compared to the hours just 
-	before and after midnight where there is normally a pronounced slump 
-	in SR closures. And both the closed-at-noon and closed-at-midnight are 
-	approaching 2\textsigma's from the hourly average.
-
-	Let's drill down on each of these outlier points. First, we aggregate the 
-	created\_date over the 2-year period that fall exactly on the top-of-the-hour 
-	with the minute and second values equal to 0, e.g. 09:00:00, 14:00:00. 
-	This visualization reveals a clearer picture of where these spikes are:
-
-	\begin{figure}[tbp]
-		\centering
- 		\includegraphics[width = \textwidth]
- 		{2-year-trend-SR_created_by_top_of_hour.pdf}
-		\caption{SRs Created at Top-of-the-Hour}
-		\label{fig:tophourcreated}
-	\end{figure}
-
 One other way to visualize this anomaly is to look at the busiest day 
-	during this 2-year period (Friday, 2023-09-29). Here we will aggregated 
-	the created\_date(s) by minute (with seconds equal to zero). This 
-	provides a minute-by-minute look at SR creation on the busiest day 
-	of this study. Here we see a clear spike at exactly noon (12:00), well 
-	beyond the 3\textsigma line, and well above any other hour of the day.
+during this 2-year period (Friday, 2023-09-29). Then aggregate 
+the created\_date(s) by minute (with seconds equal to zero), providing
+a minute-by-minute look at SR creation. Here we see a clear 
+spike at exactly noon (12:00), well beyond the 3\textsigma line.
 
@@
 
-	Let's take a similar approach to the closed\_date distribution over the 
-	2-year period. We aggregate SRs by close\_date to find the top-of-the-hour, 
-	where minute and second values equal to 0, e.g. 20:00:00. Here we see 
-	spikes at both midnight (00:00) and at noon (12:00), with the midnight 
-	spike being well beyond the 3\textsigma line. Also note how much lower 
-	(and nearly equal) all the other top-of-the-hour values are, 
-	relative to the midnight and noon spikes.
-
-	\begin{figure}[tbp]
-		\centering
-		\includegraphics[width = \textwidth]
-		{2-year-trend-SR_closed_by_top_of_hour.pdf}
-		\caption{SRs Closed at Top-of-the-Hour}
-		\label{fig:tophourclosed}
-	\end{figure}
-
 As before, let's visualize this anomaly by looking at the closed\_date(s) 
@@
 We aggregate closed\_date(s) by the minute (with seconds equal to zero). 
-	This gives a minute-by-minute view of SR closure on the busiest day 
-	of this study. Clearly visual is the very large closure spikes occurring at 
-	both midnight and noon. The chart highlights both of these anomalies 
-	with spikes of 286 and 120 at 00:00:00 and 12:00:00 respectively. 
-	Again, note how these spikes are so very much above the median 
-	values for all other minutes.
+The chart highlights anomalies with spikes at 00:00:00 and 
+12:00:00 respectively.
 
@@
 of midnight and noon likely indicates the presence of a bulk create/close 
-	software process; a process that perhaps automatically ``closes'' 
+software process that perhaps automatically ``closes'' 
 (or ``creates'') a large number of SRs with a provided time-stamp of 
-	midnight (00:00:00) or noon (12:00:00), and does so in bulk. 
-	If so, then addressing this issue likely requires a closer look at those 
-	Agencies providing these suspect SR closed/created times. The desirable 
-	state is that the created\_date and closed\_date fields correctly 
-	reflect the true opening and closing of the SRs, such that an 
-	accurate ``duration'' can be computed for the SR life-cycle. 
-	Otherwise, we are likely seeing durations that are artificially 
-	manipulated by a software process performing bulk uploads 
-	to the 311 system.
-
-	Looking at the distribution by Agency for the ``closed-exactly-at-midnight'' 
+midnight (00:00:00) or noon (12:00:00). Again, such behavior will
+distort the duration of these SRs. Looking at the distribution 
+by Agency for the ``closed-exactly-at-midnight'' 
 shows that \textgreater90\% of these suspect SRs come from just two 
 Agencies - Dept. of Buildings (DOB) and Dept. of Sanitation (DSNY).
-	If you include HPD and DOT, that accounts for 99\% of the suspect 
-	SR closures. This Agency distribution does not follow the overall 
-	SR distribution and we can assume that it is Agency specific.
 
@@
 Agency, DSNY, is responsible for \textgreater99\% of the 
-	SRs ``closed-exactly-at-noon''. Accordingly, any effort to 
-	resolve this issue necessitates working with DSNY to fully understand 
-	how this is occurring.
+SRs ``closed-exactly-at-noon''.
 
@@
 
-	The 99,779 SRs created-exactly-at-noon also appears to 
-	originate with the Dept of Sanitation (DSNY) which is responsible 
-	for \textgreater99\% of the suspect SRs.
-
-	\begin{figure}[tbp]
-		\centering
-		\includegraphics[width = \textwidth]{created_at_noon_chart.pdf}
-		\caption{SRs Created Exactly at Noon by Agency}
-		\label{fig:noon-created}
-	\end{figure}
-
-	While it is difficult to ultimately know if these observations 
-	are indeed anomalies; they could be the result of normal, 
-	expected 311 SR activity. But at least by observation that 
-	outcome appears very unlikely.
-
 \subsection{resolution\_action\_update\_date}
 \label{sec: resolutionaction}
-	When an SR is updated, by any means on any field, the software 
+When an SR is updated, by any means on any field, the 311 software 
 automatically populates the resolution\_action\_update\_date. 
-	Analyzing that field it is apparent that some of the SR updates 
-	are happening long, long after the SR is closed. In this dataset 
-	there are eight SRs that have resolution\_action\_update\_date values 
-	of 44,601 days after the closed\_date. The problem lies not with 
-	the  resolution\_action\_update\_date, but rather with the closed\_date 
-	values for these 8 SRs, all of which are in the format of ``1900-01-01''. 
-	It has previously been discussed that these inaccurate dates may 
-	possibly be coming from an Excel spreadsheet as that is the minimum 
-	date which Excel can represent. Here is a sample of those SRs:
-
-	\begin{table}[tbp]
-	    \centering
-	    \caption{Extremely long Resolution Updates}
-		    \begin{tabular}{cccc}
-		        \hline
-		        Agency & Closed Date & Resolution Action Updated Date 
-		        & Post Closed Update Duration (days) \\
-		        \hline
-		        DHS & 1900-01-01 & 2022-02-12 00:58:43 & 44602 \\
-		        DHS & 1900-01-01 & 2022-02-11 14:28:59 & 44601 \\
-		        DHS & 1900-01-01 & 2022-02-11 14:17:41 & 44601 \\
-		        DHS & 1900-01-01 & 2022-02-11 13:45:59 & 44601 \\
-		        DHS & 1900-01-01 & 2022-02-11 13:22:12 & 44601 \\
-		        \hline
-		    \end{tabular}
-	    \label{tab:resolution-updates}
-	\end{table}
-
-	These obvious inaccurate date values are removed from the analysis. 
-	But even so, there are some very late updates to SRs, with some 
-	updates occurring almost two years after the SR is closed. There are a 
+Analyzing that field it becomes apparent that some of the SR updates 
+are happening long after the SR is closed. There are a 
 total of 7460 SRs that are updated \textgreater30 days and \textless{}730 
-	days after the closed\_date. The median of these late 
-	(\textless{}30 days) post-closed resolution\_action\_update\_date(s) is 
-	84 days. However, the mean resolution\_action\_update\_date for 
-	all of the SRs is only 0.39 days (9.3 hours) after the SR is closed. 
+days after the closed\_date (to avoid infeasible dates such as 1900-01-01).
 It is not known if this is normal behavior or an area that may 
 require further investigation.
-
-	\begin{table}[tbp]
-    	\centering
-     \caption{Long post-closed Resolution Updates}
-	    	\begin{tabular}{cccc}
-		        \hline
-		        Agency & Closed Date & Resolution Action Updated Date 
-		        & Post Closed Update Duration (days) \\
-		        \hline
-		        TLC & 2023-05-25 14:21:12 & 2023-10-19 10:57:02 & 147 \\
-		        TLC & 2023-07-12 11:03:46 & 2023-12-07 09:00:05 & 148 \\
-		        TLC & 2023-08-03 10:34:23 & 2024-01-29 10:00:37 & 179 \\
-		        TLC & 2022-04-04 10:54:39 & 2022-06-02 12:14:11 & 59 \\
-		        TLC & 2022-12-05 15:23:45 & 2023-02-02 09:06:45 & 59 \\
-		        \hline
-		\end{tabular}
-    	\label{tab:resolution}
-	\end{table}
-
-	Here is a look at the spread of the post-closed resolution\_action\_update\_date 
-	values those SRs 	that are \textgreater30 days and \textless{}730 days. 
-	There are some significant outliers:
 
@@
 
-	If we look at a breakout of these (late?) update dates by Agency, 
-	we can see that the vast majority li with two Agencies: Taxi \& Limo 
-	Commission (TLC) and Dept of Sanitation (DSNY) which account 
-	for 96\% of the SRs in question.  Is there a valid reason that TLC and 
-	DSNY would be updating an SR over month (or months) after 
-	the SR has been closed? Perhaps, but it seems to merit closer investigation.
-
-	\begin{figure}[tbp]
-		\centering
-		\includegraphics[width = \textwidth]{postClosedBarChart.pdf}
-		\caption{Post-Closed resolution\_action\_update\_date(s) 
-		\textgreater30 days by Agency}
-		\label{fig:resolution-by-agency}
-	\end{figure}
-
 \section{Accuracy and precision}
 \label{sec:precision}
-There is one area in particular where the question of precision vs. 
-accuracy immediately arises, and that is with the Latitude and Longitude 
-fields. Both the Longitude and Latitude fields are expresses as a 14-decimal 
-number, e.g. 40.86769186022511 (also the Location field which is a 
+An area where the question of precision vs. accuracy arises is
+with the Latitude and Longitude fields. Both the Longitude and 
+Latitude fields are expresses as a 14-decimal 
+number, e.g. 40.86769186022511. Also the Location field which is a 
 straight concatenation of latitude and longitude). Given that 1 
@@
 or 1/1,000,000,000 of a meter. For reference a DNA molecule is 
-approximately 2nm in width.
-
-Clearly the representation of the Latitude and Longitude fields are a 
-classic case of 14-digit precision, but limited accuracy. It is more likely 
-that the Lat/Long values are accurate to the 5\textsuperscript{th} decimal place, 
-about 3.64 feet; even that might be overstating things.
+approximately 2nm in width. Clearly the representation of 
+the Latitude and Longitude fields are a classic case of 14-digit 
+precision, but limited accuracy.
 
 \section{Redundant \& Duplicate fields}\label{sec:duplicates}
-During this analysis, several redundant fields were observed. Listed below 
-is a discussion for each of a pair of data fields that are redundant or 
-near-redundant. These fields should be examined further for 
-possible consolidation in the overall dataset.
+During this analysis, several redundant fields were observed. These 
+fields should be examined further for possible consolidation.
 
@@
 and longitude fields, with a comma and parenthesis 
-added. Example:
+added. The inclusion of the location field seems questionable 
+as the data is arguably more difficult to extract than the 
+two specific fields, especially for software programs. Example:
 
@@
 \end{itemize}
-
-The inclusion of the location field seems questionable as the data is 
-arguably more difficult to extract than the two specific fields, especially 
-for software programs.
 
@@
 translated and compared to the borough field, there is a 
-98.3\% match. Given the size of the dataset, that still leaves 
-110,715 non-matching values, not an insignificant number.
-
-And that, in a nutshell, is the problem with two ``near-duplicate'' 
-fields. It is like the  story of a man wearing two watches; he's never 
-quite sure what time it is. And with ``near-duplicate'' fields, which 
-field is correct when they disagree? If one field is blank, can you 
-use the other one? Which field can be used as the reference value. And 
-as we will see, there are several such examples in this dataset.
-
-We can tell where the non-matches occur. It is mostly in DOT, NYPD, 
-and DSNY which represent 90\% of the non-matching values.
-
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width = \textwidth]{non_matching_borough_boundaries_chart.pdf}
-	\caption{Non-matching borough and borough\_boundary by Agency}
-	\label{fig:borough-boundaries}
-\end{figure}
+98.3\% match. And that is the problem with two ``near-duplicate'' 
+fields; which field is correct when they disagree?
 
@@
 is used exclusively by the Taxi \& Limo Commission which governs 
-taxis and other cars for hire. There are a total of 3567 non-blank 
-entries in the taxi\_company\_borough field, with 99.94\% blank.
+taxis and other cars for hire.
 
@@
  USPS, while the companion zipcode field incident\_zip field had 
- an accuracy rate of 99.93\%. It is not surprising that only 0.58\% of 
- the two fields match (37,292 matches out of 6.4 million SRs.). While 
- an Agency breakdown can be determined, given the questionable 
- computation of the zip\_code field, such an analysis would seem 
- to provide any useful information.
+ an accuracy rate of 99.93\%.
 
@@
 fields not cited in the Data Dictionary. Their usage however, is easy to 
-discern. Are these two fields duplicates? Nearly so:  99.94\% of the entries 
-match. Since the source of the fields and the underlying computational 
+discern. These two fields are near-duplicates with 99.94\% of the entries 
+matching. Since the source of the fields and the underlying computational 
 process are not know, it remains a challenge to determine which 
@@
  field contains abbreviations for the City agencies such as NYPD, DOT, 
- HPD, DEP, etc. The agency\_name field contains the full name of 
+ HPD, etc. The agency\_name field contains the full name of 
  the various organizations: New York Police Department, Department 
  of Transportation, Department of Housing Preservation and 
- Development, Department of Environmental Protection, and so on. Given 
- how well known the NYC Agency abbreviations are it seems redundant 
- to include both in the dataset.
+ Development, and so on. It seems redundant to include both.
 
@@
 any noteworthy location, including but not limited to, parks, 
-hospitals, airports, sports facilities, performance spaces, etc.'' We 
+hospitals, airports, sports facilities, performance spaces, etc.'' Our analysis
 found that was not the case. To be sure many of the entries in 
 the landmark field do contain landmark names, e.g.``Pennsylvania 
-Station'', ``La Guardia Airport'', ``Gramercy Park'', the vast majority 
+Station'', ``La Guardia Airport'', etc., the vast majority 
 of the entries are street names, e.g. ``Fenton Avenue'', ``Steinway 
-Street'', ``Broadway''. These entries appear to be very similar to 
-those in the street\_name field. And in fact, the values of 
-street\_name and landmark exactly match 62\% of the time, and 
-while this is not a full duplicate, it is a precentage of matching 
-that is certainly indicative of duplicate usage. Even the non-matches 
+Street'', ``Broadway''. These entries are similar to those in the 
+street\_name field. And in fact, street\_name and landmark 
+match 62\% of the time. While a high enough percentage of matching 
+to be indicative of duplicate usage. Even the non-matches 
 (excluding blanks) would appear to be matches except for minor 
@@
 	        NINTH AVE & 9 AVE \\
-	        NORTH SIXTH ST & NORTH 6 ST \\
-	        SOUTH FOURTH ST & SOUTH 4 ST \\
 	        SIXTH AVE & 6 AVE \\
-	        THIRD AVE & 3 AVE \\
 	        EAST FIRST ST & EAST 1 ST \\
@@
 	        SAINT LAWRENCE AVE & ST LAWRENCE AVE \\
-	        SOUTH FIRST ST & SOUTH 1 ST \\
-	        WEST FIFTH ST & WEST 5 ST \\
 	        MT HOPE PL & MOUNT HOPE PL \\
@@
 
-The distribution by Agency for the non-matches somewhat follows 
-the overall SR distribution (NYPD, DPR, DSNY), but with some notable 
-exceptions such as Taxi \& Limo Commission, Dept of Homeless 
-Services (DHS), and Dept of Health \& Mental Hygiene (DOHMH).
-
-\begin{figure}[tbp]
-	\centering
-  	\includegraphics[width = \textwidth]
-  	{non-matchingstreet_nameandlandmark.pdf}
-  	\caption{Non-matching street\_name \& landmark by Agency}
-  	\label{fig:landmarkchart}
-\end{figure}
-
 \subsection{cross\_street\_1 \& intersection\_street\_1 and cross\_street\_2 
@@
 street address and the cross street, as in ``24 W 90th street between 
-Columbus and Amsterdam''. The Data Dictionary states that these 
-streets are ``based on the geo validated incident location''. It then 
-adds ``If the service request pertains to a location in the middle of a 
-block, cross street will be provided by geocoding...If the service 
-request pertains to a street corner, intersection will be provided by 
-geocoding.''  One might think that the cross\_street(s) would apply 
-for ``middle of a block'' incidents, while the intersection\_street(s) 
-would apply to those incidents that happen on a street corner; 
-that those two sets of fields would be exclusive to one another; 
-an either/or situation.
-
-Unfortunately, that is not the case. For the majority of data rows, both 
-pair of fields are populated and are duplicates. In fact, 88\% of the 
-time, cross\_street\_1 exactly matches intersection\_street\_1 and 
-cross\_street\_2 exactly matches intersection\_street\_2, Only 
-11.9\% are non-matches.
+Columbus and Amsterdam''. For the majority of data rows, both 
+pair of fields are populated and are duplicates; 88\% of 
+cross\_street\_1 matches intersection\_street\_1. Similarly with
+cross\_street\_2 exactly matches intersection\_street\_2.
 
 \textit{(Note: Address standardization was applied to the street 
-addresses using the R \emph{campfin} package. This was done so as 
-to prevent the software from generating a non-match between 
+addresses using the R \emph{campfin} package. This was done 
+to prevent generating a non-match between 
 streets such as ``240 E 69\textsuperscript{th} ST'' and ``240 E 
-69\textsuperscript{th} Street'' which are clearly the same address, 
-just slight formatting and spelling differences.)}
-
-The problem occurs when the two pairs of streets do not match, such as 
-when one is blank and the other is not. Which street is correct? If one 
-is blank, say cross\_street\_1, but  intersection\_street\_1 is not blank, should 
-you use the non-blank one?  Let's look at some examples:
-
-\begin{table}[tbp]
-	\centering
-     \caption{Matching/Non-Matching cross\_street\_1 \& intersection\_street\_1}
-     	\label{tab:streets1}
-	    	\begin{tabular}{l l l}
-	      	\toprule
-	      	\multicolumn{3}{c}
-	      	{\textbf{Matching cross\_street\_1 \& intersection\_street\_1}} \\
-	      	\midrule
-	      	\textbf{cross\_street\_1} & \textbf{intersection\_street\_1} & \textbf{agency} \\
-	      	\midrule
-	       	FORT HAMILTON PKWY & FORT HAMILTON PKWY & NYPD \\
-	        	87 ST              & 87 ST              & NYPD \\
-	        	EAST 169 ST        & EAST 169 ST        & NYPD \\
-	        	EAST 74 ST         & EAST 74 ST         & DOT  \\
-	        	CHERRY AVE         & CHERRY AVE         & NYPD \\
-	      	\midrule
-	      	\multicolumn{3}{c}
-	      	{\textbf{Non-matching cross\_street\_1 and intersection\_street\_1}} \\
-	      	\midrule
-	      	\textbf{cross\_street\_1} & \textbf{intersection\_street\_1} & \textbf{agency} \\
-	      	\midrule
-	        	KAPPOCK ST     & NETHERLAND AVE   & DOT \\
-	        	GREENE AVE     & CLINTON AVE      & DOT \\
-	        	LEWIS AVE      & FULTON ST        & DOT \\
-	        	GERARD AVE     & EAST 161 ST      & DOT \\
-	        	8 AVE          & WEST 136 ST      & DOT \\
-	      	\bottomrule
-	   	\end{tabular}
-\end{table}
-
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width = \textwidth]
-	{non-matchingcross_street_1andintersection_street_1.pdf}
-	\caption{Non-matching X\_Street\_1 \& Intersection\_Street\_1}
-	\label{fig:xstreet1}
-\end{figure}
-
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width = \textwidth]
-	{non-matchingcross_street_2andintersection_street_2.pdf}
-	\caption{Non-matching X\_Street\_2 \& Intersection\_Street\_2}
-	\label{fig:xstreet2}
-\end{figure}
+69\textsuperscript{th} Street'' which are clearly the same address
+but with formatting and spelling differences.)}
 
 In addition to the matching/non-matching values, we encountered a number 
-of what you might call \emph{near-matches}. There were numerous cases 
-where the cross\_street and intersection\_street would be close to 
-identical, e.g. HARMON DR and HARMON RD. To identify these 
-near-matches we applied a common method of determining matches 
-by measuring  the \href{https://en.wikipedia.org/wiki/Hamming_distance}
+of \emph{near-matches}, cases where the 
+cross\_street and intersection\_street are nearly identical, 
+e.g. HARMON DR and HARMON RD. To identify these 
+near-matches we measured the \href{https://en.wikipedia.org/wiki/Hamming_distance}
 {Hamming Distance} between the two fields. The Hamming Distance is a 
@@
 	        WEST 168 ST    & WEST 167 ST           & DOT    & 1 \\
-	        115 AVE        & 120 AVE               & DOT    & 2 \\
 	        105 AVE        & 107 AVE               & DOT    & 1 \\
-	        71 ST          & 80 ST                 & DOT    & 2 \\
 	        145 ST         & 150 ST                & DOT    & 2 \\
@@
 	        19 AVE         & 20 AVE                & DOT    & 2 \\
-	        25 AVE         & 30 AVE                & DOT    & 2 \\
 	        76 ST          & 79 ST                 & DOT    & 1 \\
@@
 	      \midrule
-	        19 AVE         & 18 AVE                & DOT    & 1 \\
 	        227 ST         & 225 ST                & DOT    & 1 \\
@@
 	        BEACH 110 ST   & BEACH 109 ST          & DOT    & 2 \\
-	        3 AVE          & 2 AVE                 & DOT    & 1 \\
 	        88 ST          & 87 ST                 & DOT    & 1 \\
 	        WEST 114 ST    & WEST 113 ST           & DOT    & 1 \\
-	        192 ST         & 189 ST                & DOT    & 2 \\
 	        5 AVE          & 4 AVE                 & DOT    & 1 \\
@@
 borough\_boundary field is only 98.3\% duplicate. Admittedly, 
-loosing 1.7\% of the data in those two fields is quite small, for 
-this large dataset that amounts to 110, 715 non-matching 
-occurrences. Similarly with police\_precincts and police\_precincts 
-where there is a 99.9\% match; close to 100\% but 
-not a complete match.
-
-Ultimately, it will require a deeper dive into the relationship 
-between these duplicate and near-duplicate fields to determine 
-if they can be removed or not. Pending that, the authors would propose 
-deleting the following fields from the overall data set, 
-which explanation provided.
+loosing 1.7\% of the data in those two fields is quite small, however
+for this large dataset that amounts to 110, 715 
+non-matching occurrences. The authors  propose 
+deleting the following fields from the data set.
 
@@
 	which is an abbreviation of the Agency's name. The agency\_name 
-	abbreviations are clear, simple, and well understood. The 
-	authors believe the agency\_name field could be eliminated 
-	without loss of data quality.
+	abbreviations are clear, simple, and well understood.
 
@@
 
-	\item location:  The location field is simply a concatenation of 
+	\item location:  The location field is a straight concatenation of 
 	the latitude and longitude fields, with a comma and parenthesis 
-	added. The authors feel this field actually hinders data analysis 
-	as the field must be re-split into the two components 
-	(latitude and longitude) to conduct geographic analysis. The data is 
-	a 100\% duplication of the latitude and longitude fields. Removal 
-	of this field will not result in any loss of data.
+	added. The data is a 100\% duplication of the latitude 
+	and longitude fields. Removal of this field will not 
+	result in any data loss.
 
@@
 	is more correct, but feel that removal of one of the precinct fields 
-	is prudent and would result in very minimal data loss. (We 
-	should note that both police\_precinct and police\_precincts 
-	fields each have 34.5\% invalid data.)
+	is prudent and would result in very minimal data loss.
 
 	\item borough\_boundaries (computed field): This field is one of 
-	the six computed fields. It has a 98.3 match with the borough 
-	field. We recommend deleting this fields despite the somewhat 
-	limited loss of data.
+	the six computed fields. It has a 98.3\% match with the borough 
+	field.
 
@@
 	as shown in the \ref{sec:case-study-zip-codes} can lead to 
-	dangerous mistakes when used for analytical purposes. We would 
-	recommend deleting this field.
+	dangerous mistakes when used for analytical purposes. 
+	We recommend deleting this field.
 \end{itemize}
@@
 in-fact useful, as the population of these fields is very, very scarce. For 
-example, the taxi\_company\_borough field is 99.94\% blank, and the 
-remaining entries match the borough field only 0.05\% field. A similar 
-situation exists for the road\_ramp field (99.75\% blank), the 
-vehicle\_type field (99.71\%blank), and possibly due\_date which 
-is 99.62\% blank. There are several others: bridge\_highway\_direction, 
-bridge\_highway\_name, bridge\_highway\_segment, and 
-taxi\_pick\_up\_location...all of which have \textgreater99\% 
-blanks. Most of those fields appear to be almost exclusively used by 
-the Taxi \& Limo Commission (TLC), which should be consulted 
-to determine if those fields are used and properly populated.
+example, the taxi\_company\_borough field is 99.94\% 
+blank. A similar situation exists for the road\_ramp field 
+(99.75\% blank), the vehicle\_type field (99.71\%blank), and 
+due\_date which is 99.62\% blank. There are several fields that are 
+mostly blank: bridge\_highway\_direction, bridge\_highway\_name, 
+bridge\_highway\_segment, and taxi\_pick\_up\_location...all 
+of which have \textgreater99\% blanks. Most of those fields 
+appear to used exclusively by the Taxi \& Limo Commission (TLC).
 
@@
 \label{sec:datadictionay}
-During this analysis, it became apparent that the published Data 
-Dictionary could use an update. We found many discrepancies 
-between the Data Dictionary and the actual data. While not actually, 
-incorrect or dirty data, it nonetheless can lead analytic efforts astray. In 
-data analysis, it's essential to accurately describe the nature and purpose 
-of each field to ensure proper handling and interpretation. Here 
-are some of the noted discrepancies:
+During this analysis, it became apparent that the  Data 
+Dictionary could use an update. We found several discrepancies 
+between the Data Dictionary and the actual data which can lead 
+analytic efforts astray.
+
+===========================================
+REVIEW BELOW HERE
+===========================================
 

### 2024-06-22 manuscript/manuscript.tex
Review of final section
Overall review of final sections to summarize content to most relevant points.

@@
 \section{Data Dictionary Observations} 
-\label{sec:datadictionay}
+\label{sec:datadictionary}
 During this analysis, it became apparent that the  Data 
@@
 
-===========================================
-REVIEW BELOW HERE
-===========================================
-
 \begin{itemize}
-	\item While the Open Data portal lists the basic data types of the 
+	\item While the Open Data portal lists the data types of the 
 	various fields, the Data Dictionary does not. Nor does it provide 
-	meaningful information on the specifics of various fields, such 
+	much information on the specifics of various fields, such 
 	as constraints or domain of legal values. For example, the 
 	incident\_zip field is specified as ``text''. That's probably not the 
-	best definition, perhaps	``categorical'' would be more 
-	appropriate with a description indicating that the data can 
+	best definition, perhaps	``categorical'' indicating that the data can 
 	contain only numeric characters and is not subjected to arithmetic 
-	operations. We found two instances with the zip\_code field 
-	contained character strings (``na'', ``N/A'').
-
-	\item Domain of legal values as contained in the Data Dictionary 
-	are incomplete. For example, the Data Dictionary indicates that 
-	the status field has values of assigned, canceled, closed, or pending. 
-	However, we found additional status values: in progress, started, 
-	and unspecified. Additionally, we discovered no SRs with a status of 
-	canceled. Similarly, the address\_type field is indicated to have 
-	values address, blockface, intersection, latlong, and placename. We 
-	discovered additional values of bbl and unrecognized. No values of 
-	latlong were observed. Other fields suffer the same 	inaccuracies 
-	(facility\_type, vehicle\_type, taxi\_pick\_up\_location, road\_ramp, city).
+	operations.
+
+	\item Many of the domain of legal values are incomplete or 
+	inaccurate. For example, the Data Dictionary indicates that 
+	the ``status'' field has values of \textit{assigned, canceled, closed}, 
+	or \textit{pending}. However, we found additional status 
+	values of: \textit{in progress, started}, and \textit{unspecified}. 
+	However, we found no SRs with a status of \textit{canceled}. 
+	Similarly, the address\_type field is indicated to have 
+	values \textit{address, blockface, intersection, latlong}, and 
+	\textit{placename}. We discovered the additional values of 
+	\textit{bbl} and \textit{unrecognized}. Additonally, no values of 
+	\textit{latlong} were observed. Other fields (facility\_type, 
+	vehicle\_type, taxi\_pick\_up\_location, road\_ramp, city) 
+	suffer similar inaccuracies in their doman of values.
+
+	\item As noted previously, there are six \textit{@computed} fields that
+	are present in the data, but not described in the Data Dictionary.
 \end{itemize}
@@
 \label{sec:recommendations}
-Having identified a number of data cleanliness related issues with the 
-311 Service Request (SR) dataset, what follows are a series of 
-recommendations as to how to resolve these issues. Below are 
-those recommendations;
+Here are recommendations as to how to resolve these issues.
 
@@
 	\item First, and most importantly, the Open Data Team will need 
-	to undertake a special effort to identify these data anomalies, confirm 
-	the veracity of the issue, and engage with the appropriate Agency 
-	Open Data Representatives to resolve these issues. That effort will 
-	not be easy, fast, or cheap as in some software changes will likely 
-	be required. Perhaps hardest will be working with the external City 
-	Agencies that feed data to the 311 SR dataset, but are not using 
-	the core 311 system. Integrations and API usage will need to be 
-	investigated, and in many cases the changes required to correct 
-	an issue will be on the source Agency side. Coordination and 
-	mutual approval of any changes will be key. Strong support from 
-	senior 	leadership will be a critical success factor.
-
-	\item Secondly, standards will be to be identified, agreed to, and 
-	commitment to support those standards must be obtained. For 
-	example, what are the domain of legal values for certain fields, 
-	especially those with identified issues?
-
-	\item Obviously, many of the data integrity issues can be 
+	to undertake a dedicated effort to identify these 
+	data anomalies, confirm the veracity of the issue, 
+	and engage with the appropriate Agency Open Data 
+	Representatives to resolve these issues. That effort will not 
+	be easy, fast, or cheap and in some cases software changes will likely 
+	be required. Coordination and mutual approval of any 
+	changes will be key, and solid support from senior leadership 
+	will be required.
+
+	\item Secondly, standards and protocols will need to be 
+	identified, agreed to, and commitment obtained. For example, 
+	what are the domain of legal values for certain fields?
+
+	\item Many of the data integrity issues could potentially be 
 	resolved through rule enforcement by software. For example, the 
-	software could easily prevent invalid values, such as zip codes, from 
-	entering the system. Similarly for an entire range of data issues. One 
-	key area is the created\_date and closed\_date which are frequently 
-	cited when determining responsiveness of various Agency services. The 
-	authors would argue that such dates should be driven by change 
-	in ``status'', not manually entered. For example, the closed\_date 
-	is captured when the SR ``status'' is changed to ``closed'', and 
-	not manually entered. Such a change, along with many others, could 
-	prevent errors such as negative and  zero durations which 
-	significantly affect measurements of responsiveness.
+	software could easily prevent invalid values, such as zip 
+	codes. Similarly for other data issues. One key area are is the 
+	created\_date and closed\_date which are frequently cited 
+	when determining responsiveness of various Agency services.
 
 	\item The large spikes in SR closure and creation occurring at 
-	midnight and noon, almost certainly point to an underlying automated 
-	process from an external-to-311 Agency. Such processes again distort 
-	the accurate capture of an SR duration. Data rule enforcement at the 
-	software integration or API level should be investigated.
-
-	\item One possible way to have software enforce certain data 
-	standards, would be to move more non-311 Agencies to the core 
-	311 software system. That system was completely upgraded in 
-	2019, and should be (relatively) easy to expand to other 
-	Agencies and usages.
+	midnight and noon, almost certainly point to an automated 
+	process from an external Agency integration. Such processes again distort 
+	the accurate capture of an SR duration.
+
+	\item Investigate moving additional Agencies to the core 
+	311 software system where enforcement of data rules would 
+	presumably be more consistent.
 
 	\item As noted, there are six ``computed'' fields in the data export 
-	that are not identified in the Data Dictionary. While these fields may 
-	be ``experimental'', they are nonetheless present in any data portal 
-	presentation or export. As noted in this study, a number of those 
-	computed fields suffer from data inaccuracies. It is difficult to know 
-	what to recommend for these six field without further understanding of 
-	their origin, computational method, or intended usage. We would 
+	that are not identified in the Data Dictionary. As noted, several of 
+	these computed fields suffer from data inaccuracies. We would 
 	recommend updating the Data Dictionary at a minimum and 
-	possibly hiding those fields from public display until such time as 
-	their usage and accuracy can be ascertained.
+	possibly hiding those fields from public usage until such time as 
+	their accuracy can be ascertained.
 
@@
 	their usage/utility should be carefully examined to determine if 
-	their inclusion in the 311 SR dataset is meaningful, adds value not found 
-	in other fields, and should be removed or continued to be collected. Initial 
-	analysis indicates that as much as 12\% size reductions are achievable 
-	with the removal of duplicate fields. A more thorough review and 
-	understanding of the fields would no doubt yield more savings.
+	their inclusion in the 311 SR dataset is meaningful and adds value not found 
+	in other fields. Initial analysis indicates that as much as 12\% 
+	size reductions are achievable.
 
 	\item Fields with extremely low data presence, such as taxi\_company\_borough 
-	which is populated only 0.06\% of the time. Similarly for several other 
-	fields evaluated in this study which have a high percentage of blank/unknown 
-	values. If the presence of a small percentage of SRs with these fields 
-	being populated is deemed useful and necessary, then so be it. But if 
-	the usage of these fields is so diminished, then perhaps its inclusion 
-	in the 311 SR dataset is not warranted.
+	which is populated only 0.06\% of the time, should be evaluated 
+	to determine if their inclusion is warranted.
 
@@
 	revision is indicated at 6 March 2023, well over a year ago. 	As noted 
-	in \textbackslash nameref: \nameref{sec:datadictionary} there are many 
-	problems with the Data Dictionary, including domain values, absence of 
+	in Section~\ref{sec:datadictionary} there are several problems 
+	with the Data Dictionary, including domain values, absence of 
 	the six ``computed'' fields, and inaccurate description of the usage of 
-	several fields. Some of the Data Dictionary entries can be undertaken 
-	immediately, which further revision should be addressed as part of 
-	a deeper data clean-up initiative.
+	several fields.
 
 	\item The excessive precision in the latitude, longitude, and location 
-	fields is simply nonsensical.  We would recommend no more than 
+	fields is  nonsensical.  We would recommend no more than 
 	four decimal places (0.000X) which represents a distance of approximately 
-	36.4 feet. Five decimals (0.0000X) represents 3.6 feet, which although 
-	possible, would be an extremely sophisticated and accurate 
-	geo-coding system. A 14 decimal number is simply a case of a lot 
-	of precision with limited accuracy, and should be avoided.
-
-	\item The 10 address-related address fields (incident\_address, 
+	36.4 feet. A 14 decimal number is simply a case of a great 
+	precision but limited accuracy.
+
+	\item The 11 address-related address fields (incident\_address, 
 	street\_name, city, intersection\_street\_1 \& 2, cross\_street\_1 \& 2,  
 	landmark, road\_ramp, bridge\_highway\_segment, and 
-	taxi\_pick\_up\_location should be in standard USPS format. There 
-	are several software packages which can standardize addresses to 
-	these standards. We believe such an effort should be automatically 
-	applied to these fields at data entry. This step would also greatly 
-	improve any underlying geocoding routines which 	may be used 
-	to determine data values such as BBL, latitude, longitude, and New 
-	York State X/Y-plane coordinates.
-
-	\item The use of the value ``UNKNOWN'' , ``N/A'', ``NA'', and ``na'' when 
-	entered directly as text into a field should be very carefully examined, 
-	and in all likelihood prohibited.  There are some indeed a few 
-	complaint\_type(s) where an ``N/A'' value is appropriate, such 
-	as a consumer\_complaint about certain non-geocentric issues, such 
-	as a travel agency or tax preparation service . But even in those cases 
-	a value of ``N/A'' should be entered only via a drop-down menu 
-	selection. During this analysis, there were far too many data 
-	representations of missing data including: <N/A>, ``na'', `` ``, UNKNOWN, 
-	and ``N/A''. It's challenging for analyst to have to account for so 
-	the many different methods of representing empty or null data fields. 
+	taxi\_pick\_up\_location) should be put into standard USPS format. There 
+	are several software packages which can normalize addresses to 
+	these standards.
+
+	\item There are multiple means by which missing data is 
+	represented. These include ``UNKNOWN'' , ``N/A'', ``NA'', ``na'', and 
+	a blank field, e.g. `` ''. Protocols should be established 
+	regarding how to represent missing data.
 \end{itemize}

### 2024-06-29 manuscript/manuscript.tex
Working to summarize paper to fewer pages
Small corrections to the R code (2 programs) to improve graphics and address color blind and B&W colors.  Manuscript review to reduce document length is about 1/4 complete

@@
 \usepackage[]{lineno}
+\usepackage{comment}
 \linenumbers*[1]
@@
 
-\jy{cut down to 6-8}
 \noindent
 \textit{Keywords:}
-	311 Service Request Data;
 	Data cleansing;
-	Data Consistency;
 	Data Curation;
-	Data Democratization;
-	Data Efficiency;
 	Data science;
-	Data Validity;
-	Government Data;
 	NYC Open Data;
-	Open Data Movement;
 	Open data;
-	Public Engagement;
-	Quality control;
-	Research Data Management;
 	Smart Cities;
@@
 \label{sec:intro}
-
 In the early 21st century,
@@
 
-\dt{
-A note on naming conventions: dataset field names are named 
-using ``snake\_case'' where each word in the variable name 
-is written in lowercase, and words are separated by 
-underscores, e.g. created\_date, complaint\_type, zip\_code, etc..   
-}
-
 \section{NYC 311 Service Request Data} 
 \label{sec:data}
-
 The NYC 311 service, a critical component of New York City's public
@@
 
+\begin{comment}
 \jy{FixME: this paragraph's location will be revisited later.}
@@
 data for analysis with custom written R programs.
-
-\jy{Every float (figure/table) needs to be referred to in the text.}
-
-\begin{figure}[tbp]
-	\centering
-  	\includegraphics[width=0.8\textwidth]{10-year-trend-monthly.pdf}
-  	\caption{10-year (2014-2023) Monthly SR Counts}
-  	\label{fig:10-yr-monthly}
-\end{figure}
-
-We also looked at a 10-year period (2014-2023) to
-observe system performance over a longer time-frame; see
+\end{comment}
+
+%\begin{figure}[tbp]
+%	\centering
+ % 	\includegraphics[width=0.8\textwidth]{10-year-trend-monthly.pdf}
+ %	\caption{10-year (2014-2023) Monthly SR Counts}
+ % 	\label{fig:10-yr-monthly}
+%\end{figure}
+
+To begin, we looked at a 10-year period (2014-2023) to
+observe the data over a longer time-frame; see
 Figure~\ref{fig:10-yr-monthly}. During that time-frame,
-the number of 311 SRs grew by 50\%. The summer and early fall months 
-make up all of the Top 5 SR counts by day.
-
+the number of 311 SRs grew by nearly 50\%. The summer months make
+up the highest days, and the winter months show the lowest activity.
+
+\begin{comment}
 \begin{table}[tbp]
@@
 \end{table}
-
-Table~\ref{tab:top-10-counts} summarizes the top 5 and bottom 5 days
-in counts of SRs. The highest daily SR count, 24,415, occurred on
+\end{comment}
+
+%Table~\ref{tab:top-10-counts} summarizes the top 5 and bottom 5 days
+i%n counts of SRs.
+
+The highest daily SR count, 24,415, occurred on
 August 4, 2020, coinciding with the aftermath of Tropical Storm
-Isaias, highlighting the system's critical role during crises. The
-second-highest count, 19,560, was the following day, reflecting the
-storm's continued impact. 
+Isaias, highlighting the system's critical role during crises.
+
+% The second-highest count of 19,560 was the following day, 
+%reflecting the storm's continued impact.
+
 Conversely, the lowest daily SR count, 2,965, was on Christmas Day,
 2014, with similar lows on other Christmas Days, indicating reduced
-activity during major holidays. The data also shows fewer requests on
-weekends, such as April 20, 2014 (Easter Sunday), and November 6, 2016
-(a non-holiday Sunday).
-
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width=0.8\textwidth]{SRs_by_Agency.pdf}
-  	\caption{SR counts by Agency with Cumulative Percentage}
-	\label{fig:SRcountbyAgency}
-\end{figure}
+activity during major holidays.
+
+% The data also shows fewer requests on weekends, such as 
+%April 20, 2014 (Easter Sunday), and November 6, 2016 (a non-holiday Sunday).
+
+%\begin{figure}[tbp]
+%	\centering
+%	\includegraphics[width=0.8\textwidth]{SRs_by_Agency.pdf}
+ % 	\caption{SR counts by Agency with Cumulative Percentage}
+%	\label{fig:SRcountbyAgency}
+%\end{figure}
 
@@
 cumulative percentage. Two important fields in this analysis are the
-responsible NYC Agency and the type of complaint. Given the size of
-the City of New York government, identifying the responsible agency
-and the type of complaint is essential for pinpointing the accountable
-party and troubleshooting discrepancies. This identification helps
-determine whether an issue is specific to an agency or indicative of a
-systemic problem. Although 311 SRs involve 16 different agencies, six
-agencies account for 90\% of the complaints. These are the New York
+responsible NYC Agency and the type of complaint. Identifying the 
+responsible agency and type of complaint is essential 
+for identifying the accountable party and troubleshooting discrepancies. 
+This identification also assists in determining whether an issue is 
+specific to an agency or indicative of a broader, systemic problem. 
+Six agencies account for 90\% of the complaints. These are the New York
 Police Department (NYPD), Housing Preservation and Development (HPD),
@@
 
+\begin{comment}
 \begin{figure}[tbp]
@@
 \end{table}
+\end{comment}
 
@@
 of all complaints. When consolidated, noise complaints emerge as the
-most frequently occurring type. Table~\ref{tab:noisecomplaints}
-provides a detailed breakdown of noise-related complaint types by
-count and responsible agency. The most common noise complaint is
-``Noise - Residential,'' with 675,502 complaints, accounting for 10.56\%
-of all service requests, and primarily handled by the NYPD. Other
-significant noise complaints include ``Noise - Street/Sidewalk"
-(4.70\%), ``Noise - Commercial'' (2.02\%), and ``Noise - Vehicle"
-(1.81\%), all managed by the NYPD. Additional categories such as
-"Noise - Helicopter'' (1.33\%) fall under the jurisdiction of the EDC,
-and ``Noise'' (1.57\%) is managed by the DEP. The prevalence of these
-complaints underscores the critical role of the NYPD and other
-agencies in addressing noise-related issues, reflecting the
-significant impact of urban noise on residents' quality of life. This
-data highlights the importance of efficient response and management
-strategies to mitigate noise disturbances in New York City.
+most frequently occurring complaint type.
+
+% Table~\ref{tab:noisecomplaints}
+%provides a detailed breakdown of noise-related complaint types by
+%count and responsible agency. The most common noise complaint is
+%``Noise - Residential,'' with 675,502 complaints, accounting for 10.56\%
+%of all service requests, and primarily handled by the NYPD. Other
+%significant noise complaints include ``Noise - Street/Sidewalk"
+(%4.70\%), ``Noise - Commercial'' (2.02\%), and ``Noise - Vehicle"
+(%1.81\%), all managed by the NYPD. Additional categories such as
+%"Noise - Helicopter'' (1.33\%) fall under the jurisdiction of the EDC,
+%and ``Noise'' (1.57\%) is managed by the DEP. The prevalence of these
+%complaints underscores the critical role of the NYPD and other
+%agencies in addressing noise-related issues, reflecting the
+%significant impact of urban noise on residents' quality of life. This
+%data highlights the importance of efficient response and management
+%strategies to mitigate noise disturbances in New York City.
 
@@
 \label{sec:structural}
-Structural issues in the context of data cleansing refers to issues 
-related to the how data is organized, formatted, or structured within a 
-dataset. Structural issues can make it difficult to analyze the 
-data effectively. Here are some characteristics of the 311 SR data set:
+Structural issues refers to how data is organized, formatted, 
+or structured within a dataset. Structural issues can make 
+it difficult to analyze the data effectively. Here are some 
+characteristics of the 311 SR data set:
 
@@
  identifies 41 data columns (fields) along with related information 
- for each column. However, when you download the data or 
- explore it via the online portal, you immediately note that there 
- are 47 columns (fields). The additional six fields are not addressed in the 
- Data Dictionary, but do and show up on the Column Manager widget
+ for each column. However, when downloaded the data contains 
+ six fields are not included in the Data Dictionary. However, these
+ fields do show up on the portal Column Manager widget
  as ``@computed\_region\_xxxx\_xxxx''. To some extent one can infer 
- what the these @computed\_fields are and how they are determined, but 
- it is not possible to know for certain. Most importantly, can these fields 
- be reliably used for analytical purposes?  These six fields are:
+ what the these @computed\_fields are and how they are determined but
+ not with certainty. Most importantly, can these fields be reliably 
+ used for analytical purposes?  These six fields are:
 
@@
 Fortunately, both the Data Dictionary and the portal column manager 
-do a good job of indicating the data type. The following fields were 
+clearly indicate the epected data type. The following fields were 
 checked for their specified data types:
@@
 	All are all in proper date format.
+
 	\item zip\_codes and incident\_zip: All are 5 numeric digits except 
 	for two non-numeric entries for incident\_zip (``na'' and ``N/A'').
+
 	\item x\_coordinate\_state\_plane \&  y\_coordinate\_state\_plane 
 	(a NY State geo-location system): All are all numeric.
+
 	\item latitude \& longitude: Both are numeric.
+
 	\item community\_districts, borough\_boundaries, 
 	city\_council\_district, police\_precinct, and police\_precincts: All are numeric.
+
 	\item Free-form text fields that contain a ``,'' (e.g. incident\_address 
@@
 \label{sec:blanks}
-Understanding the absence of data by field is an important factor 
+Understanding the absence of data by field is important 
 when undertaking analysis. For example, if you wanted to see if the SRs were
 closed before or after their due\_date, you would be challenged 
-as 99.57\% of the due\_date field is blank. Which calls into 
-question...are these missing values, or is the concept of a 
-due\_date simply not applicable for the vast majority of SRs?
-
-When counting the various fields for blank or N/A values, the fields appear 
+as 99.57\% of the due\_date field is blank. When counting 
+the various fields for blank or N/A values, the fields appear 
 to divide into three groups: \textbf{Mostly Empty}, \textbf{Partially Empty}, 
-and \textbf{Few/None Empty}. he Mostly Empty category ranges from 93-99.9\% blank. 
-It includes such fields as taxi\_company due\_date, pickup\_location, and 
+and \textbf{Few/None Empty}. The Mostly Empty category 
+ranges from 93-99.9\% blank and includes such fields as 
+taxi\_company due\_date, pickup\_location, and 
 landmark. The Partially Empty includes such fields as location\_type, borough, 
@@
 sense to inquiry as to why some fields are frequently blank. Here is a graphic 
-depiction of total empty (blank \& N/As) for each fields. You can see 
-the natural grouping into the Most, Some, and Few categories 
-in the stair-type visualization.
-
-\begin{figure}[tbp]
-	\centering
-  	\includegraphics[width=\textwidth]{BlankFields.pdf}
-	\caption{Number and Percentage of Empty/Blank Entries}
-	\label{fig:blank_fields}
-\end{figure}
+depiction of total empty (blank \& N/As) for each fields illustrating
+the grouping into the Most, Some, and Few categories.
+
+%\begin{figure}[tbp]
+%	\centering
+ % 	\includegraphics[width=\textwidth]{BlankFields.pdf}
+%	\caption{Number and Percentage of Empty/Blank Entries}
+%	\label{fig:blank_fields}
+%\end{figure}
 
@@
  \label{sec:domain}
- In this study, testing for invalid values revealed several issues 
- associated with select fields. Accordingly, an analyst needs to be 
- extra diligent if using these fields for analysis, and in all likelihood, 
- rows with invalid field values would  need to be removed from the 
- analysis. Here are some results.
+Any analytic effort needs to be aware of fields which 
+contain invalid values. Typically such fields would need to be
+identified and removed from the analysis. Here are some results
+of field validation.
 
@@
 	boundaries of New York City.
+
 	\item The unique\_key field was in fact unique.
-	\item Many fields have a domain of acceptable values, often 
-	determined from common usage or by examining historical 
-	datasets. Unfortunately, these domains of acceptable 
-	values are not specified in the Data Dictionary. These fields all tested as 
-	compliant within their domain:
+
+	\item Unfortunately, very few domains of acceptable 
+	values are specified in the Data Dictionary. The below 
+	fields were tested as complying with their domain as 
+	determined by public usage and historical datasets:
 	\begin{itemize}
@@
 
-\subsection{Issues with Zip Codes}
+%\subsection{Issues with Zip Codes}
 \label{sec:zipcodesissues}
- Unfortunately some fields proved to be  problematic when comparing 
- the values to a domain of legal values. For example, all zip codes 
+ \textbf{Issues with Zip Codes} Some fields proved to be problematic 
+ as to complying with a domain of legal values. For example, all zip codes 
  (two fields: zip\_codes and incident\_zip) should be valid as defined 
@@
  the computed field zip\_codes proved problematic with 
-58\% (3.6 million) of the field entries being invalid. Even the 
-field incident\_zip while having only .07\% invalid entries, exhibited
+58\% (3.6 million) of the field entries being invalid. And the 
+incident\_zip field, while having only .07\% invalid entries, still contains
 4163 such errors. The breakdown of the invalid entries in the zip\_code, sorted by 
-Agency shows that the distribution by percentage almost precisely 
-mirrors the overall breakdown of SRs by Agency, potentially indicating
-a systemic problem.
-
-\subsubsection{Case Study: Noise Complaints by Zip Code}
+Agency shows that the distribution by percentage mirrors the overall 
+breakdown of SRs by Agency, potentially indicating a systemic problem.
+
+%\subsubsection{Case Study: Noise Complaints by Zip Code}
 \label{sec:case-study-zip-codes}
-\textbf{Scenario:} The NYC Office of Nightlife (ONL) wants to 
-know ``What are the top 10 zip codes for Noise Complaints (all 8 types) 
-over the last two years?'' The intent is to assess the impact of the 
-recent NYC effort promoting a vibrant nightlife 
-in NYC while seeking to ease strained relations 
-between bar and club owners. Let's assume that this analysis 
-uses the zip\_codes field, one of the (six) 
-computed fields that has shown to have validity problems. 
-We then repeat the analysis with the incident\_zip field instead. Let's 
-subject these two data sets to validation against the USPS zip code database.
-
+\textbf{Case Study: Noise Complaints by Zip Code} Scenario: The NYC 
+Office of Nightlife (ONL) wants to know ``What are the top 10 
+zip codes for Noise Complaints over the last two years?'' ONL wants
+to assess the impact of the recent NYC effort promoting a vibrant nightlife 
+in NYC while seeking to ease strained relations between bar 
+and club owners. Let's assume that this analysis 
+uses the zip\_codes field, one of the (six) computed fields known to 
+have validity problems. We then repeat the analysis with the more 
+accurate incident\_zip field. We subject these two analysis effort 
+to validation against the USPS zip code database.
+
+\begin{comment}	 
 \begin{table}[tbp]
@@
 \end{table}
+\end{comment}
 
 As indicated, six out of ten zip\_codes are invalid, which corresponds closely 
-with what is observed in the overall dataset (58\%). Whereas the incident\_zip 
-field is completely valid, again in-line with the overall incident\_zip 
-validation percentage (99.04\%). If the analysis performed using the 
-zip\_codes field were to be 	presented to the Director, ONL, the conclusion 
-would be in error.
-
-\subsection{Issues with Police Precincts} 	
+with what is observed in the overall dataset (58\% invalid). Whereas 
+the incident\_zip field is completely valid, in-line with the 
+overall incident\_zip validation percentage of 99.04\%.
+
+%\subsection{Issues with Police Precincts} 	
 \label{sec:police-precincts}
-A curious case also exists when examining the two nearly identical 
-fields - police\_precincts and police\_precinct. Both of those fields 
-are among the ``computed'' fields in the dataset. Using 
+A curious case exists when examining two (nearly) identical 
+fields: police\_precincts and police\_precinct. Both of those fields 
+are among the ``computed'' fields previously mentioned. Using 
 \href{https://www.nyc.gov/site/nypd/bureaus/patrol/precincts-landing.page}
-{NYPD Precinct listings} it's possible to determine the valid 
-police precincts; they are generally numeric or have numeric 
-representations in the dataset. However, what we find is that 
-both fields police\_precinct and police\_precincts  have  35\% 
-invalid entries. Unfortunately, it's not the same invalid entries. For 
-the police\_precincts field, there are 2,171,864 invalid entries 
-(35\%)., a consequential number of errors. Similarly for the 
+{NYPD Precinct listings} it is possible to determine what are valid 
+police precincts. What we find is that both police\_precinct 
+and police\_precincts  have  35\% invalid entries. Unfortunately, 
+it's not the same invalid entries. For the police\_precincts field, 
+there are 2,171,864 invalid entries (35\%). Similarly for the 
 police\_precinct field, there are 2,171,778 invalid entries 
-(also 35\%).
-
-\subsection{Issues with Community Boards}
+(also 35\%), but not the same invalid entries.
+
+==========REVIEW BELOW HERE===========
+
+%\subsection{Issues with Community Boards}
 \label{sec:communityboards}
@@
 
-\subsection{Issues with Community Districts}
+%\subsection{Issues with Community Districts}
 \label{sec:communitydistrict}
@@
 
-\subsection{created\_date and closed\_date(s) -- Negative Duration}
+%\subsection{created\_date and closed\_date(s) -- Negative Duration}
 \label{sec:negativeduration}
@@
 
+\begin{comment}	
 \begin{table}[tbp]
@@
 
-\begin{figure}[tbp]
- 	 \centering
- 	 \includegraphics[width = \textwidth]{negative_duration_SR_barchart.pdf}
-	  \caption{Negative Duration SRs by Agency}
-	  \label{fig:negative-duration}
-\end{figure}
+%\begin{figure}[tbp]
+% 	 \centering
+% 	 \includegraphics[width = \textwidth]{negative_duration_SR_barchart.pdf}
+%	  \caption{Negative Duration SRs by Agency}
+%	  \label{fig:negative-duration}
+%\end{figure}
 
@@
 
-\begin{figure}[tbp]
- 	 \centering
- 	 \includegraphics[width = \textwidth]{negative_duration_SR_violin.pdf}
-	 \caption{Distribution of Negative Durations}
-	 \label{fig:negative-duration-violin}
-\end{figure}
-
-\subsubsection{Case Study: Homeless Person Assistance}
+%\begin{figure}[tbp]
+ %	 \centering
+ %	 \includegraphics[width = \textwidth]{negative_duration_SR_violin.pdf}
+%	 \caption{Distribution of Negative Durations}
+%	 \label{fig:negative-duration-violin}
+%\end{figure}
+
+%\subsubsection{Case Study: Homeless Person Assistance}
 \label{sec:homlessassistance}
@@
 
-\begin{enumerate}
+\begin{itemize}
     \item Select data from 2022-2023 and filter the data by 
@@
     \item Take an average of the ``duration'' field == \textbf{Answer:  -4.8 days}  
-\end{enumerate}
+\end{itemize}
 
@@
 
-\begin{figure}[tbp]
- 	 \centering
- 	 \includegraphics[width = \textwidth]{homeless_response_time_clean.pdf}
-	 \caption{Homeless Assistance SR Durations}
-	 \label{fig:homeless}
-\end{figure}
-
-\subsection{created\_date and closed\_date(s) --  Zero Duration}
+%\begin{figure}[tbp]
+% 	 \centering
+ %	 \includegraphics[width = \textwidth]{homeless_response_time_clean.pdf}
+%	 \caption{Homeless Assistance SR Durations}
+%	 \label{fig:homeless}
+%end{figure}
+
+%\subsection{created\_date and closed\_date(s) --  Zero Duration}
 \label{sec: zeroduration}		
@@
 
-\begin{figure}[tbp]
-	 \centering
-	 \includegraphics[width = \textwidth]{zero_duration_SR.pdf}
-	 \caption{SRs with Zero Durations by Agency}
-	 \label{fig:zero-duration}
-\end{figure}
-
-\subsection{created\_date and closed\_date(s) -- Midnight \&Noon}
+%\begin{figure}[tbp]
+%	 \centering
+%	 \includegraphics[width = \textwidth]{zero_duration_SR.pdf}
+%	 \caption{SRs with Zero Durations by Agency}
+%	 \label{fig:zero-duration}
+%\end{figure}
+
+%\subsection{created\_date and closed\_date(s) -- Midnight \&Noon}
 \label{sec:midnightandnoon}
@@
 
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width=\textwidth]
-	{2-year-trend-SR_created_by_minute_of_busiest_day.pdf}
-	\caption{SRs Created Minute-by-Minute on Busiest Day}
-	\label{fig:busiestcreated}
-\end{figure}
+%\begin{figure}[tbp]
+%	\centering
+%	\includegraphics[width=\textwidth]
+%	{2-year-trend-SR_created_by_minute_of_busiest_day.pdf}
+%	\caption{SRs Created Minute-by-Minute on Busiest Day}
+%	\label{fig:busiestcreated}
+%\end{figure}
 
@@
 
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width=\textwidth]
-	{2-year-trend-SR_closed_by_minute_of_busiest_day.pdf}
-	\caption{SRs Closed Minute-by-Minute on Busiest Day}
-	\label{fig:busiestclosed}
-\end{figure}
+%\begin{figure}[tbp]
+%	\centering
+%	\includegraphics[width=\textwidth]
+%	{2-year-trend-SR_closed_by_minute_of_busiest_day.pdf}
+%	\caption{SRs Closed Minute-by-Minute on Busiest Day}
+%	\label{fig:busiestclosed}
+%\end{figure}
 
@@
 
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width = \textwidth]{closed_at_midnight_chart.pdf}
-	\caption{SRs Closed Exactly at Midnight by Agency}
-	\label{fig:midnight-closed}
-\end{figure}
+%\begin{figure}[tbp]
+%	\centering
+%	\includegraphics[width = \textwidth]{closed_at_midnight_chart.pdf}
+%	\caption{SRs Closed Exactly at Midnight by Agency}
+%	\label{fig:midnight-closed}
+%end{figure}
 
@@
 
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width = \textwidth]
-	{closed_at_noon_chart.pdf}
-	\caption{SRs Closed Exactly at Noon by Agency}
-	\label{fig:noon-closed}
-\end{figure}
-
-\subsection{resolution\_action\_update\_date}
+%\begin{figure}[tbp]
+%	\centering
+%	\includegraphics[width = \textwidth]
+%	{closed_at_noon_chart.pdf}
+%	\caption{SRs Closed Exactly at Noon by Agency}
+%	\label{fig:noon-closed}
+%\end{figure}
+
+%\subsection{resolution\_action\_update\_date}
 \label{sec: resolutionaction}
@@
 
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width = \textwidth]{post_closed_violin.pdf}
-	\caption{Post-Closed resolution\_action\_update\_date(s) 
-	\textgreater30 days}
-	\label{fig:resolution-violin}
-\end{figure}
+%\begin{figure}[tbp]
+%	\centering
+%	\includegraphics[width = \textwidth]{post_closed_violin.pdf}
+%	\caption{Post-Closed resolution\_action\_update\_date(s) 
+%	\textgreater30 days}
+%	\label{fig:resolution-violin}
+%\end{figure}
 
@@
 
-\subsection{latitude \& longitude and the location fields}
+%\subsection{latitude \& longitude and the location fields}
 \label{sec:latlong}
@@
 
-\subsection{borough and park\_borough fields}
+%\subsection{borough and park\_borough fields}
 \label{sec:parkborough}
@@
 
-\subsection{borough and borough\_boundaries fields}
+%\subsection{borough and borough\_boundaries fields}
 \label{sec:boroughboundaries}
@@
 
-\subsection{borough and taxi\_company\_borough fields}
+%\subsection{borough and taxi\_company\_borough fields}
 \label{sec:taxicompanyborough}
@@
 
- \subsection{incident\_zip and zip\_codes fields}
+ %\subsection{incident\_zip and zip\_codes fields}
  \label{sec:zipcodes}
@@
 
- \subsection{police\_precinct and police\_precincts fields} 
+ %\subsection{police\_precinct and police\_precincts fields} 
  \label{sec:police} 
@@
 
- \subsection{agency and agency\_name fields}
+% \subsection{agency and agency\_name fields}
  \label{sec:agencyname}
@@
 
-\subsection{landmark and street\_name}
+%\subsection{landmark and street\_name}
 \label{sec:landmark}
@@
 
+\begin{comment}
 \begin{table}[ht]
@@
 \end{table}
-
-\subsection{cross\_street\_1 \& intersection\_street\_1 and cross\_street\_2 
+\end{comment{
+
+%\subsection{cross\_street\_1 \& intersection\_street\_1 and cross\_street\_2 
 \& intersection\_street\_2}
@@
 
+\begin{comment}
 \begin{table}[tbp]
@@
 \end{table}
-
- \subsection{Shrinking file size by removing duplicates}
+\end{comment}
+
+ %\subsection{Shrinking file size by removing duplicates}
  \label{sec:filesize}
@@
 
-\section{Protocol Suggestions} \label{sec:protocol}
+%\section{Protocol Suggestions} \label{sec:protocol}
 

### 2024-06-29 code/datacleansing.R

@@ -31,9 +31,12 @@ max_closed_date <- as.POSIXct("2024-01-31 23:59:59", format = "%Y-%m-%d %H:%M:%S
 
 #########################################################################
 # Create the bar chart
+#########################################################################
+# Create the bar chart
 create_bar_chart <- function(
     dataset,
     x_col,
+    y_col,
     chart_title,
     sub_title,
     x_axis_title,
@@ -42,7 +45,13 @@ create_bar_chart <- function(
     add_mean = FALSE,
     add_median = FALSE,
     add_sd = FALSE,
-    chart_file_name) {
+    add_trendline = FALSE,
+    add_maximum = FALSE,
+    add_minimum = FALSE,
+    add_second_maximum = FALSE,
+    extra_line = NULL,
+    chart_file_name) 
+{
   # Find the row with the maximum count
   max_row <- dataset[which.max(dataset[[y_axis_title]]), ]
   max_count <- max_row[[y_axis_title]]
@@ -89,10 +98,9 @@ create_bar_chart <- function(
     paste("Std Dev", print_out_title, ":"), format(median_value, big.mark = ",")
   )
 
-
   # Create the bar chart
   bar_chart <- ggplot(dataset, aes(x = .data[[x_col]], y = .data[[y_axis_title]])) +
-    geom_bar(stat = "identity", fill = "#009E73") +
+    geom_bar(stat = "identity", fill = "#117733") +
     theme(
       axis.title.x = element_text(vjust = 0, size = 11),
       axis.title.y = element_text(vjust = 1, size = 11),
@@ -104,56 +112,87 @@ create_bar_chart <- function(
     ) +
     geom_hline(
       yintercept = seq(starting_value, max_count, by = increment),
-      linetype = "dotted", color = "gray20", linewidth = 0.4
+      linetype = "dotted", color = "gray40", linewidth = 0.3
     ) +
     ggtitle(chart_title,
-      subtitle = paste(sub_title, total_count, sep = "")
+      subtitle = paste0(sub_title, format(total_count, big.mark = ","), sep = "")
     ) +
     labs(x = x_axis_title, y = NULL) +
-
-    # Add annotations for min and max values
-    annotate("text",
-      x = min_row[[x_col]], y = min_count,
-      label = "Min", size = 4, color = "gray20", vjust = 0.4, hjust = -0.6
-    ) +
-    annotate("text",
-      x = max_row[[x_col]], y = max_count,
[…]

### 2024-07-03 manuscript/manuscript.tex
Top to bottom review in attempt to summarize page count.
Edited and tables/charts removed.  Currently at 22 pages.

@@
 
-==========REVIEW BELOW HERE===========
-
 %\subsection{Issues with Community Boards}
 \label{sec:communityboards}
-Community Boards are an important aspect of NYC government. They are the 
-most local, grassroots form of City government, and a vital connection 
-between communities and elected officials and City agencies. Community 
-Boards are used as a measure of City services throughout the five 
+Community Boards are the most local, grassroots form of 
+City government; a vital connection between communities and 
+elected officials and City agencies. Community Boards are often
+used as a measure of equity of City services throughout the five 
 boroughs. In the 2022-2023 dataset there are 27,276 invalid community\_board 
@@
 of invalid Community Boards by Agency is not consistent with 
-the overall SR Agency distribution. This indicates that there are likely 
-specific issues at some key Agencies, in this case Taxi \& Limo 
+the overall SR Agency distribution. This distribution indicates that 
+there are likely issues at key Agencies, in this case Taxi \& Limo 
 Commission (TLC), Parks \& Recreation, etc.
@@
 Another of the ``computed'' fields is community\_districts. Community Districts 
-are the boundaries for the Community Boards, but unlike how 
-Community Boards are an arm of the government of New York City, 
-the Community District is used by the Department of City Planning 
-(DCP) for purposes of environmental, socio-economic, and demographic 
+are the boundaries for the Community Boards, but differ as they
+are used by the Department of City Planning (DCP) for purposes 
+of environmental, socio-economic, and demographic 
 purposes. It is a geographical division rather than a local 
@@
 expect that these two fields are automatically populated by the 
-SR application software. Thus the created\_date is associated with setting an SR status 
-to ``new'' or ``open''. And similarly when an SR's 	status is 
-changed to ``closed''  the closed\_date would be 
-populated. Unfortunately, that does not appear to be the case as
-there are several anomalies with the various date fields.
+SR application software; associated with setting an SR status 
+to ``new'' or ``closed''. Unfortunately, that does not appear 
+to be the case as there are several anomalies with the various date 
+fields, including:
 
@@
 	\item SRs with a closed\_date that occurs before the created\_date.
+
  	\item created\_date(s) and closed\_date(s) in the far distant past.
+
  	\item created\_date(s) and closed\_date(s) that match to the second.
+
 	\item A large number of SRs closed and created exactly at midnight 
-	or exactly at noon. 
-\end{itemize}
-
-These date fields are important because citizens, NYC Government Officials, 
-and NYC Agencies use these dates to measure the 
-responsiveness for providing services. The measure of ``duration'',	 
-of the response time of a 311 call is a closely 
-watched, carefully scrutinized metric, both in terms of overall performance 
-and with a geographical area. Duration, while not a field in the dataset, 
-it is easily computed by closed\_date - created\_date.
+	or exactly at noon, to the second
+\end{itemize}
+
+Citizens, NYC Government Officials, and NYC Agencies use 
+these dates to measure the responsiveness for providing 
+services. In particular, the of ``duration'' of the SR 
+is a carefully scrutinized metric used to measure the responsiveness
+of an Agency or service. Duration, while not directly present 
+in the dataset, is easily computed by (closed\_date - created\_date).
 
@@
 were created, thereby generating a nonsensical ``negative duration''. 
-While this is a small percentage overall (0.2\%) it can have significant impact 
-on response time analysis. Here is a sample:
+While this is a small percentage overall (0.2\%) these errors can 
+have significant impact on response time analysis. 
+Here is a sample:
 
@@
 
-The largest errors are shown ``excluding extreme negative values''. We found 
-eight SRs with extremely large negative durations (\textless{} -730), all containing 
-an entry of ``1900-01-01'' as the closed\_date, which generates extremely large 
-negative durations exceeding 44,601 days (122 yrs). That large of an anomaly 
-can skew statistical results, even though the number of occurrences is 
-small. According, these  SR rows are removed from the analysis. These 
-extremely large negative duration SRs are all from the 
-Department of Homeland Services (DHS) A visualization of the 
-negative-duration SRs by Agency shows that this negative-duration 
+As noted, the largest errors are shown ``excluding extreme negative values''. We found 
+eight SRs with extremely large negative durations (\textless{} -730), all originating
+with DHS and all containing an entry of ``1900-01-01'' as the closed\_date. This 
+generates extremely large negative durations exceeding 44,601 
+days (122 yrs). That large of an anomaly can significantly skew statistical 
+results, even though the number of occurrences is small. According, these 
+SR rows are removed from the analysis. Excluding these extreme negative
+values, a visualization Agency indicates that the negative-duration 
 issue is a DOT problem, where 95\% of these types of errors occur.
@@
 outliers, the magnitude of the negative-durations is troubling and can 
-produce bizarre analytical results.
+produce bizarre analytical results as will be shown below.
 
@@
 \label{sec:homlessassistance}
-\textbf{Scenario:} The Dept. of Homeless Services (DHS) wants to 
-know ``How quickly are 311 calls for ``Homeless Person Assistance
-resolved over the past two years?'' This is a typical request made by 
-both the public and City government. It's a performance metric along 
-the lines of ``How quickly is my Agency responding to ``critical'' 
-requests, and does that performance vary by Borough, Zip Code, etc. 
-As an analyst, this appears to be a routine effort:
-
-\begin{itemize}
-    \item Select data from 2022-2023 and filter the data by 
-    complaint\_type = ``Homeless Person Assistance'' (yields 55,000 SRs)
-    \item Compute a new field ``duration'' (closed\_date – created\_date)
+\textbf{Scenario:} DHS wants to know ``How quickly have 311 SRs
+for ``Homeless Person Assistance'' been resolved over the 
+past two years?'' This is a typical request made by 
+both the public and City government; analgous to ``How quickly 
+is my Agency responding to ``critical'' requests, and does 
+that performance vary by Borough, Zip Code, etc. Here is a
+sample analysis:
+
+\begin{itemize}
+    \item Select data from 2022-2023 and filter by complaint\_type = 
+    ``Homeless Person Assistance'' (this yields 55,000 SRs)
+
+    \item Compute the ``duration'' (closed\_date – created\_date)
+
     \item Take an average of the ``duration'' field == \textbf{Answer:  -4.8 days}  
@@
 in an absurd answer? The answer lies in the computation of the ``duration'' 
-field. As it turns out, there are eight DHS SRs that have a closed\_date 
+field. It turns out, there are eight DHS SRs that have a closed\_date 
 of ``1900-01-01''. Each of those SRs creates a negative duration of -44,602 
-days (-122 years). So just those eight SRs with extreme negative duration 
-values is enough to drive the average of the 75,000 
-Homeless Assistance SRs to a negative value, clearly incorrect.
-
-In this case, the median is a better measure of central tendency than the average. 
-And the median of the duration field is  0.2 days (approx 5 hrs).
+days (-122 years). Just those eight SRs is enough to drive the 
+average of the 75,000 Homeless Assistance SRs to a negative 
+value; clearly incorrect. In this case, the median is a better measure 
+of central tendency; the median is  0.2 days (approx 5 hrs).
 
@@
 \label{sec: zeroduration}		
-A more serious and prevalent problem with closed\_date(s) 
-occurs when the closed\_date and created\_date are exactly the same -- to 
-the second -- which accordingly creates a \textbf{zero duration'}. Again 
-this is nonsensical, and the presence of such zero durations can again 
-severely distort analytics. There are 191,141 such SRs representing 3.1\% of 
-all non-blank data. As shown in the chart 	below, 99\% of the 
-zero duration SRs appear to predominately occur in five 
-Agencies (Dept of Mental Health \&Hygiene, Dept of Transportation, 
-Dept of Business, Dept of Sanitation, and Dept of Environmental Protection.) 
-This is not in-line with the overall Agency distribution of SRs and indicates an 
-Agency-specific issue, and the solution likely lies at those Agencies.
+A more prevalent problem with closed\_date(s) occurs when 
+the closed\_date and created\_date are exactly the same -- to 
+the second. Accordingly, this creates a \textbf{zero duration'}, again 
+nonsensical. There are 191,141 such SRs representing 3.1\% of 
+all non-blank data. As shown below, 99\% of these zero duration 
+SRs occur in five Agencies:  (Dept of Mental Health \&Hygiene (DHMH), Dept 
+of Transportation (DOT), Dept of Business (DOB), Dept of Sanitation 
+(DSNY), and Dept of Environmental Protection (DEP). This distribution
+is does not mirror the overall Agency distribution of SRs indicating an 
+Agency-specific issue.
 
@@
 Another problem discovered with the created\_date and 
-closed\_date fields - there appears to be an unusually large number of SRs 
+closed\_date fields: there are an unusually large number of SRs 
 created or closed at exactly midnight (00:00:00) and exactly noon (12:00:00), 
-to the second. The distribution of SR creation and closure largely follows the 
-work-day clock with many SRs created during day-light hours, and 
-fewer SRs 	created at night and the early hours of the morning. However, 
-here there appears to be significantly greater numbers of SR closed exactly 
-at midnight and exactly at noon, as well as a significant number of SRs 
-created exactly at noon. Observations during this 2-year period include:
+to the second. Normally, distribution of SR creation and closure 
+largely follows the work-day clock with many SRs created during 
+day-light hours, and fewer SRs 	created at night and the early hours 
+of morning. However, there are significantly greater numbers 
+of SR closed exactly at midnight and noon, as well as a significant 
+number of SRs created exactly at noon. Observations include:
 
@@
 	 \item There were 99,779 SRs created exactly at noon (12:00:00)
+
 	 \item There were 235,347 SRs closed exactly at midnight (00:00:00)
+
 	\item There were 105,505 SRs closed exactly at noon. 
@@
 
-One other way to visualize this anomaly is to look at the busiest day 
+A unique way to visualize this anomaly is to examine the busiest day 
 during this 2-year period (Friday, 2023-09-29). Then aggregate 
@@
 As before, let's visualize this anomaly by looking at the closed\_date(s) 
-on the busiest day during this 2-year study period (Friday, 2023-09-29). 
-We aggregate closed\_date(s) by the minute (with seconds equal to zero). 
-The chart highlights anomalies with spikes at 00:00:00 and 
-12:00:00 respectively.
+on the busiest day (Friday, 2023-09-29). Aggregate closed\_date(s) by 
+the minute (with seconds equal to zero). This highlights spikes 
+at 00:00:00 and 12:00:00 respectively.
 
@@
 software process that perhaps automatically ``closes'' 
-(or ``creates'') a large number of SRs with a provided time-stamp of 
-midnight (00:00:00) or noon (12:00:00). Again, such behavior will
-distort the duration of these SRs. Looking at the distribution 
-by Agency for the ``closed-exactly-at-midnight'' 
-shows that \textgreater90\% of these suspect SRs come from just two 
-Agencies - Dept. of Buildings (DOB) and Dept. of Sanitation (DSNY).
+(or ``creates'') a large number of SRs with a generated time-stamp of 
+midnight (00:00:00) or noon (12:00:00). Again, such behavior distorts
+ the duration of these SRs. The distribution by Agency for 
+ the ``closed-exactly-at-midnight'' SRs indicates that \textgreater90\% of 
+ these suspect SRs come from just two Agencies: DOB and DSNY.
 
@@
 
-However, the ``SR closed at noon'' visualization shows that  a single 
-Agency, DSNY, is responsible for \textgreater99\% of the 
-SRs ``closed-exactly-at-noon''.
+A further visualization reveals that  a single Agency, DSNY, is 
+responsible for \textgreater99\% of the SRs ``closed-exactly-at-noon''.
 
@@
 \label{sec: resolutionaction}
-When an SR is updated, by any means on any field, the 311 software 
-automatically populates the resolution\_action\_update\_date. 
-Analyzing that field it becomes apparent that some of the SR updates 
-are happening long after the SR is closed. There are a 
-total of 7460 SRs that are updated \textgreater30 days and \textless{}730 
-days after the closed\_date (to avoid infeasible dates such as 1900-01-01).
-It is not known if this is normal behavior or an area that may 
-require further investigation.
+When an SR is updated, the 311 software automatically 
+populates the resolution\_action\_update\_date. Some of 
+those SR updates are happening long after the SR is 
+closed. There are a total of 7460 SRs that are updated \textgreater30 days 
+after the closed\_date (and \textless{}730 days to avoid infeasible 
+dates such as 1900-01-01). It is not known if this is normal behavior
+ or an area that requires further investigation.
 
@@
 \label{sec:precision}
-An area where the question of precision vs. accuracy arises is
-with the Latitude and Longitude fields. Both the Longitude and 
-Latitude fields are expresses as a 14-decimal 
-number, e.g. 40.86769186022511. Also the Location field which is a 
-straight concatenation of latitude and longitude). Given that 1 
-degree of latitude at the equator is equal to 111.044736 kilometers, 
-the ``1'' at the end of that number represents approximate 1.1104 nanometer 
-or 1/1,000,000,000 of a meter. For reference a DNA molecule is 
-approximately 2nm in width. Clearly the representation of 
-the Latitude and Longitude fields are a classic case of 14-digit 
-precision, but limited accuracy.
+An question of precision vs. accuracy arises with the Latitude 
+and Longitude fields. Both Latitude are 
+expressed as a 14-decimal number, e.g. 40.86769186022511. Given 
+that 1 degree of latitude at the equator is equal to 111.044736 
+kilometers, the ``1'' at the end of that number represents 
+approximately 1.1104 nanometers (1/1,000,000,000 of a meter). For 
+reference a DNA molecule is approximately 2nm in width. Clearly 
+the representation of the Latitude and Longitude fields is a 
+classic case of 14-digit precision, but limited accuracy.
 
 \section{Redundant \& Duplicate fields}\label{sec:duplicates}
-During this analysis, several redundant fields were observed. These 
-fields should be examined further for possible consolidation.
+During this analysis, several redundant fields were observed which
+should be examined further for possible consolidation.
 
@@
 The location field is a pure concatenation of the latitude 
-and longitude fields, with a comma and parenthesis 
-added. The inclusion of the location field seems questionable 
-as the data is arguably more difficult to extract than the 
-two specific fields, especially for software programs. Example:
+and longitude fields with a comma and parenthesis 
+added; 100\% duplicative. Thus the  location field seems questionable 
+as the data is arguably more difficult to extract and use than the 
+two individual fields. Example:
 
@@
 	\item  latitude: 40.768456429488
+
 	\item  longitude: -73.9575661888774
+
 	\item  location: (40.768456429488, -73.95756618887745)
@@
 \label{sec:parkborough}
-These two fields are 100\% matches; fully redundant.
+Borough and park\_borough: These two fields are 100\% matches; fully redundant.
 
@@
 \label{sec:boroughboundaries}
-The borough field is a text field for the five New York City boroughs (Manhattan, Queens, 
-Brooklyn, etc.). The borough\_boundary field is numeric with values 
-from 1-5 corresponding to the five boroughs (Staten Island, Brooklyn, 
-Queens, Manhattan, Bronx). When the borough\_boundary field is 
-translated and compared to the borough field, there is a 
-98.3\% match. And that is the problem with two ``near-duplicate'' 
+Borough and borough\_boundary fields are a 98.3\% match. This 
+is an example of the the problem with two ``near-duplicate'' 
 fields; which field is correct when they disagree?
@@
 \label{sec:taxicompanyborough}
-The two fields borough and taxi\_company\_borough might appear to be 
-related. In fact, despite the names, the two data fields are nearly 
-completely different with only 0.05\% matching. The mismatch is so 
-high that it suggests the two fields are used very differently. This field 
-is used exclusively by the Taxi \& Limo Commission which governs 
-taxis and other cars for hire.
+Borough and taxi\_company\_borough fields that despite their 
+names, are nearly completely different; only a 0.05\% match, suggesting
+that these two fields are used very differently. The taxi\_company\_borough
+ field is used exclusively by the TLC which governs all cars for hire.
 
@@
  The zip\_code field is one of the six computed fields and its validity 
- has been explored in Section \ref{sec:zip-codes} and found to be 
- lacking with 55.55\% of the entries invalid according to the 
- USPS, while the companion zipcode field incident\_zip field had 
- an accuracy rate of 99.93\%.
+ was explored in Section \ref{sec:zip-codes} and found to be 
+ lacking; 56\% of the entries are invalid. The companion 
+ field, incident\_zip, had an accuracy rate of 99.93\%. Our analysis
+ indicates that the computed zip\_code field should be eliminated
+ until such time as accuracy issues can be addressed.
 
@@
 Both the police\_precinct and police\_precincts are among the computed 
-fields not cited in the Data Dictionary. Their usage however, is easy to 
-discern. These two fields are near-duplicates with 99.94\% of the entries 
-matching. Since the source of the fields and the underlying computational 
-process are not know, it remains a challenge to determine which 
-field is the ``right'' field to use. (The validity of these fields is 
-covered in Section \ref{sec:police-precincts}.)
+fields not cited in the Data Dictionary. However, their usage is easy to 
+discern; they represent the NYPD precinct that corresponds to the
+incident address. These two fields are near-duplicates with 99.94\% of the entries 
+matching. (The validity of these fields is 
+covered in Section \ref{sec:police-precincts}.) Since the underlying 
+computational process is unknown, we are unable to determine which 
+field is ``more correct'' field. However having two fields with 99.9\% 
+duplicity indicates that only one police precinct field is necessary.
 
@@
  \label{sec:agencyname}
- While not duplicates, these fields have a 1:1 correspondence.  The Agency 
- field contains abbreviations for the City agencies such as NYPD, DOT, 
- HPD, etc. The agency\_name field contains the full name of 
- the various organizations: New York Police Department, Department 
- of Transportation, Department of Housing Preservation and 
- Development, and so on. It seems redundant to include both.
+ Agency and agency\_name fields: While not duplicates per se, these fields 
+ have a 1:1 correspondence.  The Agency field contains abbreviations 
+ for the City agencies such as NYPD, DOT, HPD, etc. The agency\_name 
+ field contains the full name of these organizations. It seems 
+ redundant to include both as the abbreviations are commonly used
+ and well understood.
 
@@
 \label{sec:landmark}
-The landmark field is listed in the Data Dictionary as ``Can refer to 
+Landmark and street\_name: The landmark field is listed in the Data Dictionary as ``Can refer to 
 any noteworthy location, including but not limited to, parks, 
@@
 the landmark field do contain landmark names, e.g.``Pennsylvania 
-Station'', ``La Guardia Airport'', etc., the vast majority 
+Station'', ``LaGuardia Airport'', etc., the vast majority 
 of the entries are street names, e.g. ``Fenton Avenue'', ``Steinway 
-Street'', ``Broadway''. These entries are similar to those in the 
-street\_name field. And in fact, street\_name and landmark 
-match 62\% of the time. While a high enough percentage of matching 
-to be indicative of duplicate usage. Even the non-matches 
-(excluding blanks) would appear to be matches except for minor 
-spelling and nomenclature changes, e.g. ``NINTH AVE'' \& ``9 AVE''.
+Street'', ``Broadway''. These entries are very similar to the 
+street\_name field with a 62\% match, enough to be indicative 
+of duplicate usage. Even the non-matches (excluding blanks) 
+appear to be matches except for minor spelling and 
+nomenclature changes, e.g. ``NINTH AVE'' \& ``9 AVE''.
 
@@
 \label{sec:street1}
-There are two sets of street pairs in the dataset in addition to 
-the complaint address:
-
-\begin{itemize}
-	\item cross\_street\_1
-	\item cross\_street\_2
-	\item intersection\_street\_1
-	\item intersection\_street\_2
-\end{itemize}
-
-These two pairs of streets are used to help identify the location of the 
-reported incident, as it is common in New York City to provide the 
-street address and the cross street, as in ``24 W 90th street between 
-Columbus and Amsterdam''. For the majority of data rows, both 
-pair of fields are populated and are duplicates; 88\% of 
-cross\_street\_1 matches intersection\_street\_1. Similarly with
-cross\_street\_2 exactly matches intersection\_street\_2.
+Cross\_street\_1 \& intersection\_street\_1 and cross\_street\_2 
+\& intersection\_street\_2: There are two sets of street pairs in addition to 
+the incident\address field. These pairs are:
+
+\begin{itemize}
+	\item cross\_street\_1 \& intersection\_street\_1
+
+	\item cross\_street\_2 \& intersection\_street\_2
+\end{itemize}
+
+These two street pairs used to assist in identifying the address, as 
+is common in New York City. These two street pairs
+have 88\% duplicates where cross\_street\_1 matches 
+intersection\_street\_1 and cross\_street\_2 
+matches intersection\_street\_2. We believe only one of these
+street pairs are necessary. However it continues to illustrate the 
+problem of two similar data fields that largely, but not exactly, match. In
+this case there are 12\% that do not match; which field to trust.
 
 \textit{(Note: Address standardization was applied to the street 
-addresses using the R \emph{campfin} package. This was done 
-to prevent generating a non-match between 
-streets such as ``240 E 69\textsuperscript{th} ST'' and ``240 E 
-69\textsuperscript{th} Street'' which are clearly the same address
-but with formatting and spelling differences.)}
-
-In addition to the matching/non-matching values, we encountered a number 
-of \emph{near-matches}, cases where the 
+addresses using the R ``campfin'' package. This prevents 
+generating a non-match between locations such as 
+``240 E 69\textsuperscript{th} ST'' and ``240 E 69\textsuperscript{th} Street'' 
+which are clearly the same address save for formatting and spelling differences.)}
+
+In addition to the matching values, we encountered a number 
+of \textit{near-matches}, situations where the 
 cross\_street and intersection\_street are nearly identical, 
@@
 near-matches we measured the \href{https://en.wikipedia.org/wiki/Hamming_distance}
-{Hamming Distance} between the two fields. The Hamming Distance is a 
-measure of how many letters have to be changed in order for the two 
+{Hamming Distance} between the two fields. Hamming Distance is a 
+measure of how many letters must be changed for the two 
 fields to match. For this analysis, we chose a Hamming Distance of 2. Here 
-are the results for both cross\_street and intersection\_street.
+are some sample results for both cross\_street and intersection\_street.
 
@@
  \label{sec:filesize}
-By removing duplicate and ``near-duplicate'' fields, it is possible to 
-shrink the file size by 12.3\% which for this dataset equates to a 
+Shrinking file size by removing duplicates: By removing duplicate 
+and ``near-duplicate'' fields, it is possible to 
+shrink the file size by 12.3\% which for this large dataset equates to a 
 reduction of 395 Mb. A smaller file size means faster downloads, 
@@
 some challenges with these proposed deletions, mostly with 
-the ``near duplicate'' fields.  So while the park\_borough is 
+the ``near duplicate'' fields.  For example, while the park\_borough is 
 a complete duplicate of the borough field, the 
@@
 loosing 1.7\% of the data in those two fields is quite small, however
-for this large dataset that amounts to 110, 715 
-non-matching occurrences. The authors  propose 
+for this large dataset that small percentage still amounts 
+to 110, 715 non-matching occurrences. The authors  propose 
 deleting the following fields from the data set.
@@
 Additionally, it would be worth investigating if certain data fields are 
-in-fact useful, as the population of these fields is very, very scarce. For 
+in-fact useful, as the population of these fields is very scarce. For 
 example, the taxi\_company\_borough field is 99.94\% 
@@
 (99.75\% blank), the vehicle\_type field (99.71\%blank), and 
-due\_date which is 99.62\% blank. There are several fields that are 
-mostly blank: bridge\_highway\_direction, bridge\_highway\_name, 
-bridge\_highway\_segment, and taxi\_pick\_up\_location...all 
-of which have \textgreater99\% blanks. Most of those fields 
-appear to used exclusively by the Taxi \& Limo Commission (TLC).
+due\_date which is 99.62\% blank. Several additional several fields
+are mostly blank: bridge\_highway\_direction, bridge\_highway\_name, 
+bridge\_highway\_segment, and taxi\_pick\_up\_location; all 
+have \textgreater99\% blanks. Most of these fields 
+appear to used exclusively by the TLC.
 
@@
 \label{sec:datadictionary}
-During this analysis, it became apparent that the  Data 
-Dictionary could use an update. We found several discrepancies 
-between the Data Dictionary and the actual data which can lead 
-analytic efforts astray.
+The Data Dictionary could use an update. We found several 
+discrepancies between the Data Dictionary and the 
+actual data which can potentially lead analytic efforts astray.
 
@@
 	various fields, the Data Dictionary does not. Nor does it provide 
-	much information on the specifics of various fields, such 
-	as constraints or domain of legal values. For example, the 
+	information on the specifics of various fields, such 
+	as constraints or the domain of legal values. For example, the 
 	incident\_zip field is specified as ``text''. That's probably not the 
-	best definition, perhaps	``categorical'' indicating that the data can 
-	contain only numeric characters and is not subjected to arithmetic 
-	operations.
+	best definition; perhaps	``categorical'' with notes indicating that 
+	the data can contain only numeric characters and is 
+	not subjected to arithmetic operations.
 
@@
 	values of: \textit{in progress, started}, and \textit{unspecified}. 
-	However, we found no SRs with a status of \textit{canceled}. 
+	Additionally, we found no SRs with a status of \textit{canceled}. 
 	Similarly, the address\_type field is indicated to have 
@@
 	\textit{placename}. We discovered the additional values of 
-	\textit{bbl} and \textit{unrecognized}. Additonally, no values of 
+	\textit{bbl} and \textit{unrecognized}. Additionally, no values of 
 	\textit{latlong} were observed. Other fields (facility\_type, 
 	vehicle\_type, taxi\_pick\_up\_location, road\_ramp, city) 
-	suffer similar inaccuracies in their doman of values.
+	suffer similar inaccuracies in their specified domain of values.
 
 	\item As noted previously, there are six \textit{@computed} fields that
-	are present in the data, but not described in the Data Dictionary.
+	are present in the data, but not included in the Data Dictionary.
 \end{itemize}
@@
 \label{sec:recommendations}
-Here are recommendations as to how to resolve these issues.
 
@@
 	Representatives to resolve these issues. That effort will not 
-	be easy, fast, or cheap and in some cases software changes will likely 
+	be easy or fast, and in some cases software changes will 
 	be required. Coordination and mutual approval of any 
-	changes will be key, and solid support from senior leadership 
-	will be required.
+	changes will be key, requiring solid support 
+	from senior leadership.
 
@@
 	identified, agreed to, and commitment obtained. For example, 
-	what are the domain of legal values for certain fields?
+	the domain of legal values, data accuracy, address normalization, etc..
 
 	\item Many of the data integrity issues could potentially be 
-	resolved through rule enforcement by software. For example, the 
+	resolved through enforcement by software. For example, the 
 	software could easily prevent invalid values, such as zip 
@@
 	created\_date and closed\_date which are frequently cited 
-	when determining responsiveness of various Agency services.
+	when determining responsiveness of various Agency services; software
+	could conceivably prevent illogical situations such as an SR closed 
+	before it is opened.
 
 	\item The large spikes in SR closure and creation occurring at 
-	midnight and noon, almost certainly point to an automated 
-	process from an external Agency integration. Such processes again distort 
-	the accurate capture of an SR duration.
+	midnight and noon almost certainly point to an automated 
+	process from an external Agency integration. Such processes distort 
+	the accurate capture of an SR duration and should be eliminated.
 
 	\item Investigate moving additional Agencies to the core 
-	311 software system where enforcement of data rules would 
-	presumably be more consistent.
+	311 software system where data rules would 
+	be more consistently enforced.
 
 	\item As noted, there are six ``computed'' fields in the data export 
-	that are not identified in the Data Dictionary. As noted, several of 
-	these computed fields suffer from data inaccuracies. We would 
+	that are not identified in the Data Dictionary. Several of 
+	these computed fields suffer from data inaccuracies. We  
 	recommend updating the Data Dictionary at a minimum and 
@@
 	other cases, the duplication is in the 80-90\% range. These fields and 
-	their usage/utility should be carefully examined to determine if 
-	their inclusion in the 311 SR dataset is meaningful and adds value not found 
-	in other fields. Initial analysis indicates that as much as 12\% 
-	size reductions are achievable.
+	their utility should be examined to determine if inclusion in the 
+	311 SR dataset is warranted. Our analysis indicates that as much as 12\% 
+	size reductions are achievable, likely more.
 
@@
 
-	\item The Data Dictionary is in need of an upgrade and update. The last 
-	revision is indicated at 6 March 2023, well over a year ago. 	As noted 
-	in Section~\ref{sec:datadictionary} there are several problems 
-	with the Data Dictionary, including domain values, absence of 
-	the six ``computed'' fields, and inaccurate description of the usage of 
-	several fields.
-
-	\item The excessive precision in the latitude, longitude, and location 
-	fields is  nonsensical.  We would recommend no more than 
-	four decimal places (0.000X) which represents a distance of approximately 
-	36.4 feet. A 14 decimal number is simply a case of a great 
-	precision but limited accuracy.
+	\item The Data Dictionary is dated March 2023. The issues addressed 
+	in Section~\ref{sec:datadictionary} should be the bases for an update of
+	the Data Dictionary.
+
+	\item Excessive precision in the latitude, longitude, and location 
+	fields is  nonsensical.  We recommend no more than 
+	four decimal places (0.000X), representing a distance of approximately 
+	36.4 feet. Further decimal points strain credibility.
 
@@
 	landmark, road\_ramp, bridge\_highway\_segment, and 
-	taxi\_pick\_up\_location) should be put into standard USPS format. There 
-	are several software packages which can normalize addresses to 
-	these standards.
+	taxi\_pick\_up\_location) should be standardized to USPS format. Software 
+	packages can perform this standardization.
 

### 2024-07-03 manuscript/manuscript.tex
Beginning 2nd pass for document summary.

@@
 2023. Today, the NYC 311 data system manages over 3 million service
-requests per year, covering issues such as street noise, illegal
-parking, heat and hot water, abandoned vehicles, and unsanitary
-conditions. This data is publicly accessible through the NYC Open Data
+requests per year. This data is publicly accessible through the NYC Open Data
 Portal, which provides tools for querying, grouping, aggregating,
@@
 
-Despite its success, the 311 system faces several challenges, such as
+Despite its success, the 311 system faces several challenges:
 data timeliness, accuracy, and consistency, difficulties in
-correlating data over long periods, handling of personally
+correlating data over long periods, excluding personally
 identifiable information (PII), integration with stand-alone systems
-at selected NYC agencies, and managing API usage and load for numerous
+at selected NYC agencies, and managing API usagefor numerous
 third-party users. These challenges are addressed by various agency
@@
   and Innovation (OTI)}, which provides technological support for the
-open data system. Data often needs to be exported and processed in
-external programs to address these issues, as observed in the
-examination of the large 311 dataset. Furthermore, the system must
-adapt to changes such as agency name changes and the introduction of
-new complaint types, which can complicate long-term data analysis.
+open data system.
 
@@
 routes based on analyses by the Taxi and Limousine Commission (TLC),
-improving responsiveness for city agencies like Parks and Recreation,
-the Department of Transportation, and the Department of Sanitation,
-and addressing homelessness through initiatives like the Homeless
-Outreach and Mobile Engagement Street Action Teams (HOME-STAT). These
-applications demonstrate the transformative potential of open data in
-enhancing urban living and fostering a more responsive and transparent
-city government.
-
-\begin{comment}
-\jy{FixME: this paragraph's location will be revisited later.}
-We extracted 311 SR data for the period of CY 2022--2023, a dataset of
-6.4 million SR records. This is a large dataset which enables 
-observation of some rarely seen events, data inaccuracies, 
-and logical inconsistencies. It provides enough data to 
-examine trends and data outlier analysis. We queried the data 
-directly from the NYC Open Data Portal and exported the 
-data for analysis with custom written R programs.
-\end{comment}
+improving responsiveness for city agencies, and addressing 
+homelessness through initiatives like the Homeless
+Outreach and Mobile Engagement Street Action Teams (HOME-STAT).
+
+To begin, we looked at a 10-year period (2014-2023) to
+observe the data over a longer time-frame; see
+Figure~\ref{fig:10-yr-monthly}. During that time-frame,
+the number of 311 SRs grew by nearly 50\%.
 
@@
 
-To begin, we looked at a 10-year period (2014-2023) to
-observe the data over a longer time-frame; see
-Figure~\ref{fig:10-yr-monthly}. During that time-frame,
-the number of 311 SRs grew by nearly 50\%. The summer months make
-up the highest days, and the winter months show the lowest activity.
-
-\begin{comment}
-\begin{table}[tbp]
-    \centering
-    \caption{Top 5 and Bottom 5 Days for SRs}
-    \label{tab:top-10-counts}
-    \begin{tabular}{@{}lll lll@{}}
-      \toprule
-      \multicolumn{2}{c}{Top 5 Days} & \multicolumn{2}{c}{Bottom 5 Days} \\
-      \cmidrule(r){1-2} \cmidrule(l){3-4}
-      Date & Count & Date & Count\\
-      \midrule
-      2020-08-04 & 24,415 & 2014-12-25 & 2965\\
-      2020-08-05 & 19,560 & 2015-12-25 & 3247\\
-      2023-09-29 & 17,962 & 2014-04-20 & 3287\\
-      2020-07-05 & 16,916 & 2015-12-26 & 3405\\
-      2020-06-21 & 15,883 & 2016-11-06 & 3419\\
-      \bottomrule
-    \end{tabular}
-\end{table}
-\end{comment}
-
-%Table~\ref{tab:top-10-counts} summarizes the top 5 and bottom 5 days
-i%n counts of SRs.
-
-The highest daily SR count, 24,415, occurred on
-August 4, 2020, coinciding with the aftermath of Tropical Storm
-Isaias, highlighting the system's critical role during crises.
-
-% The second-highest count of 19,560 was the following day, 
-%reflecting the storm's continued impact.
-
-Conversely, the lowest daily SR count, 2,965, was on Christmas Day,
-2014, with similar lows on other Christmas Days, indicating reduced
-activity during major holidays.
-
-% The data also shows fewer requests on weekends, such as 
-%April 20, 2014 (Easter Sunday), and November 6, 2016 (a non-holiday Sunday).
-
 %\begin{figure}[tbp]
@@
 
-Figure~\ref{fig:SRcountbyAgency} display the SR counts by agency with
-cumulative percentage. Two important fields in this analysis are the
-responsible NYC Agency and the type of complaint. Identifying the 
+Figure~\ref{fig:SRcountbyAgency} displays SR counts by Agency with
+cumulative percentage. Identifying the 
 responsible agency and type of complaint is essential 
-for identifying the accountable party and troubleshooting discrepancies. 
+for troubleshooting discrepancies. 
 This identification also assists in determining whether an issue is 
-specific to an agency or indicative of a broader, systemic problem. 
+specific to an Agency or indicative of a broader, systemic problem. 
 Six agencies account for 90\% of the complaints. These are the New York
@@
 
-The distribution of complaint types skews toward a few select
-categories, accompanied by a long tail. Figure~\ref{fig:SR_complaints}
-shows that the top 20 complaint types account for 70\% of the total
-220 different types. Notably, these top complaints include several
-variations of noise-related issues, which collectively represent 22\%
-of all complaints. When consolidated, noise complaints emerge as the
-most frequently occurring complaint type.
-
-% Table~\ref{tab:noisecomplaints}
-%provides a detailed breakdown of noise-related complaint types by
-%count and responsible agency. The most common noise complaint is
-%``Noise - Residential,'' with 675,502 complaints, accounting for 10.56\%
-%of all service requests, and primarily handled by the NYPD. Other
-%significant noise complaints include ``Noise - Street/Sidewalk"
-(%4.70\%), ``Noise - Commercial'' (2.02\%), and ``Noise - Vehicle"
-(%1.81\%), all managed by the NYPD. Additional categories such as
-%"Noise - Helicopter'' (1.33\%) fall under the jurisdiction of the EDC,
-%and ``Noise'' (1.57\%) is managed by the DEP. The prevalence of these
-%complaints underscores the critical role of the NYPD and other
-%agencies in addressing noise-related issues, reflecting the
-%significant impact of urban noise on residents' quality of life. This
-%data highlights the importance of efficient response and management
-%strategies to mitigate noise disturbances in New York City.
+Figure~\ref{fig:SR_complaints} shows that the top 20 complaint 
+types account for 70\% of the total 220 different types. Notably, 
+these top complaints include several variations of noise-related 
+issues, which collectively represent 22\% of all complaints, which
+when consolidated is the most frequently occurring complaint type.
 

### 2024-07-05 manuscript/manuscript.tex
comment on paper structure

@@
 quality and curation efficiency. Section~\ref{sec:structural} examines
+\jy{How about putting all these sections into one section on cass
+  study? They can be subsections in the merged section.
+  After that, we can have a section on recommendations for open
+  data curation based on what we learn from the case study. For each
+  recommended principle, we can link to the specifics of the case
+  study.}
 the technical dimensions of the dataset, the structural issues. 

### 2024-07-10 manuscript/manuscript.tex
Incorporating Jun Yan's comments
Reduced chapters for re-org

--- 2024-07-03 manuscript/manuscript.tex
@@
 
-\section{Introduction} 
+\section{INTRODUCTION} 
 \label{sec:intro}
-In the early 21st century,
-the open data movement began to take shape, driven by the
-fundamental belief that freely accessible data can transform 
-both societies and governments. This movement champions the principles
+In the early 21st century, the open data movement began 
+to take shape, driven by the fundamental belief that 
+freely accessible data can transform both societies and 
+governments. This movement champions the principles
 of transparency, innovation, and public engagement. 
@@
 
+\jy{How about putting all these sections into one section on case
+  study? They can be subsections in the merged section.
+  After that, we can have a section on recommendations for open
+  data curation based on what we learn from the case study. For each
+  recommended principle, we can link to the specifics of the case
+  study.}
+
 The paper is organized as follows:
-Section~\ref{sec:data} offers a brief review of the history of the 311
-system and the long-term trends presented via a 10-year analysis.
-Section~\ref{sec:issues} offers a 
-general discussion of data cleansing issues impacting data 
-quality and curation efficiency. Section~\ref{sec:structural} examines
-the technical dimensions of the dataset, the structural issues. 
-In Section~\ref{sec:datatypes} we examine the data fields for compliance 
-with the Data Dictionary. Section~\ref{sec:blanks} looks at the 
-data fields as regards missing, blank, or N/A entries. Section~\ref{sec:domain} 
-explores how data fields comply with a domain of legal 
-or acceptable values. Section~\ref{sec:inconsistencies} deals 
-with the important issues surrounding logical inconsistencies 
-and concerning patterns in the data. Section~\ref{sec:precision} 
-explores the issue of precision versus accuracy. Section~\ref{sec:duplicates}
-identifies duplicate and redundant data fields. Section~\ref{sec:improvements} provides 
-actionable suggestions for mitigating or resolving identified issues. 
-Following this, Section~\ref{sec:protocol} outlines a series of general 
-principles for the release of open data, drawing from our findings. The 
-paper concludes with a discussion and recommendations in 
-Section~\ref{sec:discussion}, encapsulating the key insights and 
-implications of our research.
+Section~\ref{sec:data} `\textit{NYC 31 Service Request Data} offers a 
+look at the history of the 311system. Long-term trends are 
+presented in a 10-year data analysis (2014-2023).
+Section~\ref{sec:issues} \textit{Data Cleansing Issues} offers a 
+discussion of specific data cleansing issues impacting data quality and 
+curation efficiency. These issues include structural issues, compliance 
+with the Data Dictionary, and missing, blank, or N/A entries. We also
+explore field compliance with a domain of legal or acceptable values.
+Next we deal with several important issues surrounding logical inconsistencies 
+and concerning patterns found in the data. We also explore an issue of precision 
+versus accuracy. This section concludes by identifying duplicate/redundant data fields
+and Data Dictionary observation. Section~\ref{sec:recommendations} 
+\textit{Recommendations} provides actionable suggestions for 
+mitigating or resolving the identified issues. Section~\ref{sec:discussion} ``Discussions'' 
+concludes with an encapsulation of key insights and implications of our research.
 
@@
 Portal, which provides tools for querying, grouping, aggregating,
-geo-mapping, visualizing, and exporting results.
-
+geo-mapping, visualizing, and exporting results. The
+\href{https://www.nyc.gov/content/oti/pages/}{NYC Office of Technology
+  and Innovation (OTI)}, which provides technological support for the
+open data system.
+
+\begin{comment}
 Despite its success, the 311 system faces several challenges:
@@
 third-party users. These challenges are addressed by various agency
-open data managers and the
-\href{https://www.nyc.gov/content/oti/pages/}{NYC Office of Technology
-  and Innovation (OTI)}, which provides technological support for the
-open data system.
+open data managers and 
+\end{comment}
 
@@
 routes based on analyses by the Taxi and Limousine Commission (TLC),
-improving responsiveness for city agencies, and addressing 
-homelessness through initiatives like the Homeless
-Outreach and Mobile Engagement Street Action Teams (HOME-STAT).
+and improving responsiveness across City Agencies
 
 To begin, we looked at a 10-year period (2014-2023) to
-observe the data over a longer time-frame; see
+observe a longer time-frame; see
 Figure~\ref{fig:10-yr-monthly}. During that time-frame,
@@
 
-%\begin{figure}[tbp]
-%	\centering
- % 	\includegraphics[width=0.8\textwidth]{10-year-trend-monthly.pdf}
- %	\caption{10-year (2014-2023) Monthly SR Counts}
- % 	\label{fig:10-yr-monthly}
-%\end{figure}
-
-%\begin{figure}[tbp]
-%	\centering
-%	\includegraphics[width=0.8\textwidth]{SRs_by_Agency.pdf}
- % 	\caption{SR counts by Agency with Cumulative Percentage}
-%	\label{fig:SRcountbyAgency}
-%\end{figure}
-
-Figure~\ref{fig:SRcountbyAgency} displays SR counts by Agency with
-cumulative percentage. Identifying the 
-responsible agency and type of complaint is essential 
-for troubleshooting discrepancies. 
-This identification also assists in determining whether an issue is 
-specific to an Agency or indicative of a broader, systemic problem. 
-Six agencies account for 90\% of the complaints. These are the New York
-Police Department (NYPD), Housing Preservation and Development (HPD),
+\begin{figure}[tbp]
+	\centering
+  	\includegraphics[width=0.8\textwidth]{10-year-trend_yearly.pdf}
+ 	\caption{10-year (2014-2023) Monthly SR Counts}
+  	\label{fig:10-yr-monthly}
+\end{figure}
+
+Identifying the responsible agency is essential 
+for troubleshooting discrepancies. This identification also assists in 
+determining whether an issue is specific to an Agency or indicative 
+of a broader, systemic problem. Six agencies account for 90\% of 
+the complaints: New York Police Department (NYPD), 
+Housing Preservation and Development (HPD),
 New York City Department of Sanitation (DSNY), Department of
@@
 
+\begin{figure}[tbp]
+	\centering
+	\includegraphics[width=0.8\textwidth]{SRs_by_Agency.pdf}
+  	\caption{SR counts by Agency with Cumulative Percentage}
+	\label{fig:SRcountbyAgency}
+\end{figure}
+
 \begin{comment}
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=0.8\textwidth]{SR_by_Complaint_Type.pdf} 
-  \caption{Top 20 complaint\_type(s) and Cumulative Percentage} 
-  \label{fig:SR_complaints}
-\end{figure}
-
 \begin{table}[tbp]
@@
 these top complaints include several variations of noise-related 
-issues, which collectively represent 22\% of all complaints, which
-when consolidated is the most frequently occurring complaint type.
+issues, which collectively represent 22\% of all complaints, the 
+most frequently occurring type of complaint.
+
+\begin{figure}[tbp]
+ \centering
+  \includegraphics[width=0.8\textwidth]{SR_by_Complaint_Type.pdf} 
+  \caption{Top 20 complaint\_type(s) and Cumulative Percentage} 
+  \label{fig:SR_complaints}
+\end{figure}
 
@@
 What is data cleansing?  Wikipedia 
-\href{https://en.wikipedia.org/wiki/Data_cleansing}{Data Cleansing} offers a good definition.
+\href{https://en.wikipedia.org/wiki/Data_cleansing}{Data Cleansing} 
+offers this definition.
 
@@
 
-Many quality criteria are required in order to process high-quality 
-data. These include:
-
+Many quality criteria are utilized in order to process high-quality 
+data, including:
+\begin{comment}
 \begin{itemize}
 	\item Data validation - this effort can span a number of criteria
+
 	\begin{itemize}
 		\item Mandatory fields: Certain data fields cannot be empty.
+
 		\item Data-types: Certain fields must be of the correct type, e.g. numeric, 
@@
 		in a Data Dictionary.
+
 		\item Domain compliance:  Many data fields must adhere to a specific 
 		domain of values, e.g. statuses, state names, zip codes, gender.
+
 	\end{itemize}   
@@
 	 all to indicate the absence of data.
+
 	\item Redundant, unnecessary, or irrelevant fields
+
 	\item Logical inconsistencies such as related fields that violate the 
 	nature of that relationship, e.g. a ``due date'' that is before the ``created date''
+
 	\item Accuracy versus precision
 \end{itemize}
-
-This analysis effort is to identify the presence of such errors; not to correct them. 
-That effort would be undertaken only after an investigation as to the why
-and how such errors came about, and a discussion as to whether or not it 
-is even an ``error''.
-
-\section{Structural Issues}
+\end{comment}
+
+This purpose of this analysis effort is to identify the presence of 
+errors; not necessarily to correct them. A correction effort should
+be undertaken only after an investigation as to the why and how 
+such errors came about, and a discussion as to whether or not it 
+even is an \textit{error}.
+
+\subsection{Structural Issues}
 \label{sec:structural}
 Structural issues refers to how data is organized, formatted, 
-or structured within a dataset. Structural issues can make 
+or presented within a dataset. Structural issues can make 
 it difficult to analyze the data effectively. Here are some 
@@
 	\item There are 47 columns of data for each row, exportable as a CSV file.
+
 	\item There are four date fields (created, closed, updated, due).
+
 	\item There are three borough fields; two of which appear to be duplicates.
+
 	\item Two zip code fields, but not duplicates
-	\item Seven street fields; one pair of which appear to be duplicates
+
+	\item Seven street fields; two pair of which appear to be duplicates
+
 	\item Two Police Precinct fields; not duplicates
-	\item In addition to incident\_address, there are five additional location fields: 
+
+	\item In addition to the incident\_address, there are five additional location fields: 
 	lat/long, street\_name, landmark, NY State plane, and Block \#
+
 	\item One free-form text field, resolution\_description, which 
@@
 
-\textbf{Issue: Fields not in the Data Dictionary} The 311 SR 
-\href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}
- identifies 41 data columns (fields) along with related information 
- for each column. However, when downloaded the data contains 
- six fields are not included in the Data Dictionary. However, these
- fields do show up on the portal Column Manager widget
- as ``@computed\_region\_xxxx\_xxxx''. To some extent one can infer 
- what the these @computed\_fields are and how they are determined but
- not with certainty. Most importantly, can these fields be reliably 
- used for analytical purposes?  These six fields are:
+The 311 SR \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}
+ identifies 41 data columns (fields) and related information 
+ for each column. However, when downloaded the data reveals 
+ six additional fields that are not included in the Data Dictionary. 
+(Note that these fields do show up on the portal Column Manager widget
+ as ``@computed\_region\_xxxx\_xxxx''.) To some extent one can infer 
+ what the these @computed\_fields are and how they are determined, but
+ not with certainty. Most importantly is the question, can these fields be reliably 
+ used for analytical purposes?  These six ``computed'' fields are:
 
@@
 	\item zip\_codes
+
 	\item community\_districts
+
 	\item borough\_boundaries
+
 	\item city\_council\_districts
+
 	\item police\_precincts
+
 	\item police\_precinct 
@@
 
-\section{Validating data types}
-\label{sec:datatypes}
-Fortunately, both the Data Dictionary and the portal column manager 
-clearly indicate the epected data type. The following fields were 
-checked for their specified data types:
-
-\begin{itemize}
-	\item created\_date, closed\_date, due\_date, and resolution\_action\_updated\_date: 
-	All are all in proper date format.
-
-	\item zip\_codes and incident\_zip: All are 5 numeric digits except 
-	for two non-numeric entries for incident\_zip (``na'' and ``N/A'').
-
-	\item x\_coordinate\_state\_plane \&  y\_coordinate\_state\_plane 
-	(a NY State geo-location system): All are all numeric.
-
-	\item latitude \& longitude: Both are numeric.
-
-	\item community\_districts, borough\_boundaries, 
-	city\_council\_district, police\_precinct, and police\_precincts: All are numeric.
-
-	\item Free-form text fields that contain a ``,'' (e.g. incident\_address 
-	\& resolution\_description) are properly enclosed in quotation marks.
-\end{itemize}
-
-\section{Blank \& N/A data}
+\subsection{Blank \& N/A data}
 \label{sec:blanks}
 Understanding the absence of data by field is important 
-when undertaking analysis. For example, if you wanted to see if the SRs were
-closed before or after their due\_date, you would be challenged 
-as 99.57\% of the due\_date field is blank. When counting 
-the various fields for blank or N/A values, the fields appear 
+when undertaking analysis. For example, if you wanted to 
+determine if SRs were closed before or after their due\_date, 
+you would be challenged as 99.6\% of the due\_date field is 
+blank. When counting fields for blank or N/A values, they appear 
 to divide into three groups: \textbf{Mostly Empty}, \textbf{Partially Empty}, 
-and \textbf{Few/None Empty}. The Mostly Empty category 
-ranges from 93-99.9\% blank and includes such fields as 
-taxi\_company due\_date, pickup\_location, and 
+and \textbf{Few/None Empty}.
+
+The Mostly Empty category ranges from 93-99.9\% blank 
+and includes such fields as taxi\_company due\_date, pickup\_location, and 
 landmark. The Partially Empty includes such fields as location\_type, borough, 
@@
 depiction of total empty (blank \& N/As) for each fields illustrating
-the grouping into the Most, Some, and Few categories.
-
-%\begin{figure}[tbp]
-%	\centering
- % 	\includegraphics[width=\textwidth]{BlankFields.pdf}
-%	\caption{Number and Percentage of Empty/Blank Entries}
-%	\label{fig:blank_fields}
-%\end{figure}
-
- \section{Validating Data for Acceptable Values}
+the grouping into the Most, Partial, and Few categories.
+
+\begin{figure}[tbp]
+	\centering
+  	\includegraphics[width=\textwidth]{BlankFields.pdf}
+	\caption{Number and Percentage of Empty/Blank Entries}
+	\label{fig:blank_fields}
+\end{figure}
+
+\begin{comment}
+ \subsection{Validating Data for Acceptable Values}
  \label{sec:domain}
 Any analytic effort needs to be aware of fields which 
-contain invalid values. Typically such fields would need to be
-identified and removed from the analysis. Here are some results
-of field validation.
-
-\begin{itemize}
-	\item Latitude and Longitude fields fell within the geographic 
+contain invalid values. Typically such fields would be
+identified and removed from the analysis. Here are the results
+of selected field validation.
+
+\begin{itemize}
+	\item Latitude and Longitude fields fall within the geographic 
 	boundaries of New York City.
 
-	\item The unique\_key field was in fact unique.
+	\item The unique\_key field was in-fact unique.
 
@@
 	determined by public usage and historical datasets:
+
 	\begin{itemize}
@@
 	\end{itemize}
-\end{itemize}
-
-%\subsection{Issues with Zip Codes}
+
+\end{itemize}
+\end{comment}
+
 \label{sec:zipcodesissues}
- \textbf{Issues with Zip Codes} Some fields proved to be problematic 
- as to complying with a domain of legal values. For example, all zip codes 
+ \textbf{Issues with Zip Codes:} Some fields proved to be problematic 
+ as to their compliance with a domain of legal values. For example, all zip codes 
  (two fields: zip\_codes and incident\_zip) should be valid as defined 
- by the USPS database which contains 37,946 valid zip codes. However,
- the computed field zip\_codes proved problematic with 
-58\% (3.6 million) of the field entries being invalid. And the 
-incident\_zip field, while having only .07\% invalid entries, still contains
-4163 such errors. The breakdown of the invalid entries in the zip\_code, sorted by 
-Agency shows that the distribution by percentage mirrors the overall 
-breakdown of SRs by Agency, potentially indicating a systemic problem.
-
-%\subsubsection{Case Study: Noise Complaints by Zip Code}
+ by the USPS database containing all 37,946 valid zip codes. However,
+ the computed field zip\_codes proved problematic with over 
+58\% (3.6 million) of the field entries being invalid. And while the more accurate 
+incident\_zip field has only .07\% invalid entries, it still contains
+4163 such errors; not insignificant. The breakdown of the invalid 
+entries in the zip\_code, sorted by Agency shows that the 
+distribution by percentage mirrors the overall breakdown of 
+SRs by Agency, potentially indicating a systemic problem.
+
 \label{sec:case-study-zip-codes}
-\textbf{Case Study: Noise Complaints by Zip Code} Scenario: The NYC 
+\textbf{Case Study: Noise Complaints by Zip Code.}  Scenario: The NYC 
 Office of Nightlife (ONL) wants to know ``What are the top 10 
 zip codes for Noise Complaints over the last two years?'' ONL wants
-to assess the impact of the recent NYC effort promoting a vibrant nightlife 
+to assess the impact of recent efforts promoting a vibrant nightlife 
 in NYC while seeking to ease strained relations between bar 
 and club owners. Let's assume that this analysis 
-uses the zip\_codes field, one of the (six) computed fields known to 
+uses the zip\_codes field, one of the computed fields known to 
 have validity problems. We then repeat the analysis with the more 
-accurate incident\_zip field. We subject these two analysis effort 
+accurate incident\_zip field. Finally, we subject these two analysis effort 
 to validation against the USPS zip code database.
 
-\begin{comment}	 
 \begin{table}[tbp]
@@
 \end{table}
-\end{comment}
 
@@
 the incident\_zip field is completely valid, in-line with the 
-overall incident\_zip validation percentage of 99.04\%.
-
-%\subsection{Issues with Police Precincts} 	
+overall incident\_zip validation percentage of 99\%.
+
 \label{sec:police-precincts}
-A curious case exists when examining two (nearly) identical 
-fields: police\_precincts and police\_precinct. Both of those fields 
-are among the ``computed'' fields previously mentioned. Using 
+\textbf{Issues with Police Precincts:} A curious case exists when 
+examining two (nearly) identical fields: police\_precincts and 
+police\_precinct. Both of those fields are among the computed fields 
+previously mentioned. Using 
 \href{https://www.nyc.gov/site/nypd/bureaus/patrol/precincts-landing.page}
-{NYPD Precinct listings} it is possible to determine what are valid 
+{NYPD Precinct listings} it is possible to determine the valid 
 police precincts. What we find is that both police\_precinct 
-and police\_precincts  have  35\% invalid entries. Unfortunately, 
-it's not the same invalid entries. For the police\_precincts field, 
-there are 2,171,864 invalid entries (35\%). Similarly for the 
-police\_precinct field, there are 2,171,778 invalid entries 
+and police\_precincts  have about 35\% invalid entries. Unfortunately, 
+it's not the same invalid entries. The police\_precincts field, 
+has 2,171,864 invalid entries (35\%). Similarly the 
+police\_precinct field has 2,171,778 invalid entries 
 (also 35\%), but not the same invalid entries.
 
-%\subsection{Issues with Community Boards}
 \label{sec:communityboards}
-Community Boards are the most local, grassroots form of 
-City government; a vital connection between communities and 
-elected officials and City agencies. Community Boards are often
-used as a measure of equity of City services throughout the five 
-boroughs. In the 2022-2023 dataset there are 27,276 invalid community\_board 
-entries which represents 0.43\% of non-blank data. The distribution 
-of invalid Community Boards by Agency is not consistent with 
-the overall SR Agency distribution. This distribution indicates that 
-there are likely issues at key Agencies, in this case Taxi \& Limo 
-Commission (TLC), Parks \& Recreation, etc.
-
-%\subsection{Issues with Community Districts}
+\textbf{Issues with Community Boards:} Community Boards are the most 
+local, grassroots form of City government; a vital connection 
+between communities, elected officials, and City agencies. Community 
+Boards often service as a measure of the equity of City services 
+throughout the five boroughs. In the 2022-2023 dataset there 
+are 27,276 invalid community\_board entries representing 0.43\% 
+of non-blank data. The distribution of invalid Community Boards 
+by Agency is not consistent with the overall SR Agency distribution,
+indicating that there are likely issues at key Agencies, in 
+this case Taxi \& Limo Commission (TLC), DPR, etc.
+
 \label{sec:communitydistrict}
-Another of the ``computed'' fields is community\_districts. Community Districts 
-are the boundaries for the Community Boards, but differ as they
-are used by the Department of City Planning (DCP) for purposes 
-of environmental, socio-economic, and demographic 
-purposes. It is a geographical division rather than a local 
-government division. Due to how the community\_district data 
-is formatted, it is not possible to establish validity. However, 
-it is possible to determine that the dataset contains 72 unique 
-entries, while there are only 59 valid Community Districts.
-
-%\subsection{created\_date and closed\_date(s) -- Negative Duration}
+\textbf{Issues with Community Districts:} Another of the computed fields 
+is community\_districts. Community Districts 
+are the geographic boundaries for the Community Boards, but differ as they
+are used by the Department of City Planning (DCP) for environmental, 
+socio-economic, and demographic purposes. It is a geographical 
+division rather than a City government division. Due to how the 
+community\_district data is formatted, it is not possible to 
+establish validity. However, we can determine that the dataset 
+contains 72 unique entries, while there are only 59 
+valid Community Districts.
+
 \label{sec:negativeduration}
-For the created\_date and closed\_date fields, you might 
+\textbf{Issues with created\_date and closed\_date - Negative Durations:} For 
+the created\_date and closed\_date fields, one might 
 expect that these two fields are automatically populated by the 
-SR application software; associated with setting an SR status 
+SR application software; associated with perhaps setting an SR status 
 to ``new'' or ``closed''. Unfortunately, that does not appear 
 to be the case as there are several anomalies with the various date 
-fields, including:
+fields:
 
@@
 
- 	\item created\_date(s) and closed\_date(s) in the far distant past.
+ 	\item created\_date(s) and closed\_date(s) in the far, far distant past.
 
@@
 
-	\item A large number of SRs closed and created exactly at midnight 
-	or exactly at noon, to the second
-\end{itemize}
-
-Citizens, NYC Government Officials, and NYC Agencies use 
-these dates to measure the responsiveness for providing 
-services. In particular, the of ``duration'' of the SR 
-is a carefully scrutinized metric used to measure the responsiveness
-of an Agency or service. Duration, while not directly present 
-in the dataset, is easily computed by (closed\_date - created\_date).
+	\item A large number of SRs closed and/or created exactly at midnight 
+	or exactly at noon, to the second.
+\end{itemize}
+
+Citizens, NYC Government Officials, and NYC Agencies especially use 
+the closed and created dates as a measure of the ``duration'' of the SR 
+and hence Agency responsiveness of a service. Duration, while 
+not directly present in the dataset, is easily 
+computed by (closed\_date - created\_date).
 
@@
 While this is a small percentage overall (0.2\%) these errors can 
-have significant impact on response time analysis. 
-Here is a sample:
-
-\begin{comment}	
+have significant impact on response time analysis. Here is a sample:
+
 \begin{table}[tbp]
     \centering
-    \caption{Largest and Smallest errors (days)}
+    \caption{Largest Negative Duration errors (days)}
 	    \begin{tabular}{l l l r l}
@@
 		        2023-12-18 03:13:00 & 2023-01-16 13:10:00 & -335.6 & DOT \\
-		        \midrule
-		        \multicolumn{5}{c}{\textbf{Smallest errors (days)}} \\
-		        \midrule
-		        \textbf{created\_date} & \textbf{closed\_date} & \textbf{duration} 
-		        & \textbf{agency} \\
-		        \midrule
-		        2023-06-28 07:07:31 & 2023-06-28 07:07:00 & -0.00036 & DOT \\
-		        2023-06-29 09:10:20 & 2023-06-29 09:10:00 & -0.00023 & DOT \\
-		        2023-01-12 06:50:13 & 2023-01-12 06:50:00 & -0.00015 & DOT \\
-		        2023-06-26 08:09:07 & 2023-06-26 08:09:00 & -0.00008 & DOT \\
-		        2023-01-12 06:51:01 & 2023-01-12 06:51:00 & -0.00001 & DOT \\
 		        \bottomrule
 	    \end{tabular}
-    \label{tab:combined_errors}
+    \label{tab:largest-errors}
 \end{table}
-\end{comment}
 
@@
 with DHS and all containing an entry of ``1900-01-01'' as the closed\_date. This 
-generates extremely large negative durations exceeding 44,601 
-days (122 yrs). That large of an anomaly can significantly skew statistical 
+generates extremely large negative durations exceeding -44,601 
+days (-122 yrs). Such a large of an anomaly can significantly skew statistical 
 results, even though the number of occurrences is small. According, these 
 SR rows are removed from the analysis. Excluding these extreme negative
-values, a visualization Agency indicates that the negative-duration 
-issue is a DOT problem, where 95\% of these types of errors occur.
-
-%\begin{figure}[tbp]
-% 	 \centering
-% 	 \includegraphics[width = \textwidth]{negative_duration_SR_barchart.pdf}
-%	  \caption{Negative Duration SRs by Agency}
-%	  \label{fig:negative-duration}
-%\end{figure}
+values, an analysis shows the negative-duration issue is a DOT problem, 
+where 95\% of these types of errors occur.
 
@@
 outliers, the magnitude of the negative-durations is troubling and can 
-produce bizarre analytical results as will be shown below.
-
-%\begin{figure}[tbp]
- %	 \centering
- %	 \includegraphics[width = \textwidth]{negative_duration_SR_violin.pdf}
-%	 \caption{Distribution of Negative Durations}
-%	 \label{fig:negative-duration-violin}
-%\end{figure}
-
-%\subsubsection{Case Study: Homeless Person Assistance}
+produce bizarre analytical results.
+
+\begin{figure}[tbp]
+	 \centering
+	 \includegraphics[width = \textwidth]{negative_duration_SR_violin.pdf}
+ \caption{Distribution of Negative Durations}
+ \label{fig:negative-duration-violin}
+\end{figure}
+
 \label{sec:homlessassistance}
-\textbf{Scenario:} DHS wants to know ``How quickly have 311 SRs
-for ``Homeless Person Assistance'' been resolved over the 
-past two years?'' This is a typical request made by 
-both the public and City government; analgous to ``How quickly 
+\textbf{Case Study: Homeless Person Assistance.} Scenario: DHS wants to know ``How 
+quickly have 311 SRs for ``Homeless Person Assistance'' been resolved 
+over the past two years?'' This is a typical request made by 
+both the public and City government; analogous to ``How quickly 
 is my Agency responding to ``critical'' requests, and does 
-that performance vary by Borough, Zip Code, etc. Here is a
-sample analysis:
-
-\begin{itemize}
-    \item Select data from 2022-2023 and filter by complaint\_type = 
+that performance vary by Borough, Zip Code, etc. To conduct such an 
+analysis, we can use the tools on the Open Data Portal. Here are the steps:
+
+\begin{itemize}
+    \item Select created\_date from 2022-2023 and filter by complaint\_type = 
     ``Homeless Person Assistance'' (this yields 55,000 SRs)
@@
 
-    \item Take an average of the ``duration'' field == \textbf{Answer:  -4.8 days}  
+    \item Take an average of the ``duration'' field  which yields: \textbf{Answer:  -4.8 days}  
 \end{itemize}
@@
 of ``1900-01-01''. Each of those SRs creates a negative duration of -44,602 
-days (-122 years). Just those eight SRs is enough to drive the 
-average of the 75,000 Homeless Assistance SRs to a negative 
+days (-122 years). Those eight SRs are enough to drive the 
+average duration of 55,000 Homeless Assistance SRs to a negative 
 value; clearly incorrect. In this case, the median is a better measure 
-of central tendency; the median is  0.2 days (approx 5 hrs).
+of central tendency. The median is  0.2 days (approx 5 hrs) which will
+surely please the DHS.
 
@@
 
-%\subsection{created\_date and closed\_date(s) --  Zero Duration}
 \label{sec: zeroduration}		
-A more prevalent problem with closed\_date(s) occurs when 
-the closed\_date and created\_date are exactly the same -- to 
-the second. Accordingly, this creates a \textbf{zero duration'}, again 
-nonsensical. There are 191,141 such SRs representing 3.1\% of 
-all non-blank data. As shown below, 99\% of these zero duration 
+\textbf{Issues with created\_date \& closed\_date - Zero Durations:} A more 
+prevalent problem occurs when the closed\_date and created\_date 
+are exactly the same -- to the second. Accordingly, this creates 
+a \textbf{zero duration'}, again nonsensical. There are 191,141 
+such SRs representing 3.1\% of all non-blank data. 99\% of these zero duration 
 SRs occur in five Agencies:  (Dept of Mental Health \&Hygiene (DHMH), Dept 
 of Transportation (DOT), Dept of Business (DOB), Dept of Sanitation 
-(DSNY), and Dept of Environmental Protection (DEP). This distribution
-is does not mirror the overall Agency distribution of SRs indicating an 
+(DSNY), and Dept of Environmental Protection (DEP). This Agency distribution
+is does not mirror the overall Agency distribution of SRs, likely indicating an 
 Agency-specific issue.
@@
 
-%\subsection{created\_date and closed\_date(s) -- Midnight \&Noon}
 \label{sec:midnightandnoon}
-Another problem discovered with the created\_date and 
-closed\_date fields: there are an unusually large number of SRs 
+\textbf{Issues with created\_date \& closed\_date(s) - Midnight \& Noon:} Another 
+problem discovered with the created\_date and 
+closed\_date fields - there are an unusually large number of SRs 
 created or closed at exactly midnight (00:00:00) and exactly noon (12:00:00), 
-to the second. Normally, distribution of SR creation and closure 
-largely follows the work-day clock with many SRs created during 
+to the second. Normal distribution of SR creation and closure 
+largely follows the work-day clock with most  SRs created during 
 day-light hours, and fewer SRs 	created at night and the early hours 
@@
 the created\_date(s) by minute (with seconds equal to zero), providing
-a minute-by-minute look at SR creation. Here we see a clear 
-spike at exactly noon (12:00), well beyond the 3\textsigma line.
-
-%\begin{figure}[tbp]
-%	\centering
-%	\includegraphics[width=\textwidth]
-%	{2-year-trend-SR_created_by_minute_of_busiest_day.pdf}
-%	\caption{SRs Created Minute-by-Minute on Busiest Day}
-%	\label{fig:busiestcreated}
-%\end{figure}
-
-As before, let's visualize this anomaly by looking at the closed\_date(s) 
-on the busiest day (Friday, 2023-09-29). Aggregate closed\_date(s) by 
-the minute (with seconds equal to zero). This highlights spikes 
+a minute-by-minute look at SR creation on that one day. Note the spike 
+at exactly noon (12:00), well beyond the 3\textsigma line.
+
+\begin{figure}[tbp]
+	\centering
+	\includegraphics[width=\textwidth]
+	{2-year-trend_SR_created_by_minute_of_busiest_day.pdf}
+	\caption{SRs Created Minute-by-Minute on Busiest Day}
+	\label{fig:busiestcreated}
+\end{figure}
+
+Repeat the same for the \# of SRs with closed\_date(s) 
+on the busiest day (Friday, 2023-09-29). Again, aggregated by 
+the minute (with seconds equal to zero). Note the spikes 
 at 00:00:00 and 12:00:00 respectively.
 
-%\begin{figure}[tbp]
-%	\centering
-%	\includegraphics[width=\textwidth]
-%	{2-year-trend-SR_closed_by_minute_of_busiest_day.pdf}
-%	\caption{SRs Closed Minute-by-Minute on Busiest Day}
-%	\label{fig:busiestclosed}
-%\end{figure}
+\begin{figure}[tbp]
+	\centering
+	\includegraphics[width=\textwidth]{2-year-trend-SR_closed_by_minute_of_busiest\_day.pdf}
+	\caption{SRs Closed Minute-by-Minute on Busiest Day}
+	\label{fig:busiestclosed}
+\end{figure}
 
@@
 software process that perhaps automatically ``closes'' 
-(or ``creates'') a large number of SRs with a generated time-stamp of 
-midnight (00:00:00) or noon (12:00:00). Again, such behavior distorts
- the duration of these SRs. The distribution by Agency for 
- the ``closed-exactly-at-midnight'' SRs indicates that \textgreater90\% of 
- these suspect SRs come from just two Agencies: DOB and DSNY.
-
-%\begin{figure}[tbp]
-%	\centering
-%	\includegraphics[width = \textwidth]{closed_at_midnight_chart.pdf}
-%	\caption{SRs Closed Exactly at Midnight by Agency}
-%	\label{fig:midnight-closed}
-%end{figure}
-
-A further visualization reveals that  a single Agency, DSNY, is 
-responsible for \textgreater99\% of the SRs ``closed-exactly-at-noon''.
-
-%\begin{figure}[tbp]
-%	\centering
-%	\includegraphics[width = \textwidth]
-%	{closed_at_noon_chart.pdf}
-%	\caption{SRs Closed Exactly at Noon by Agency}
-%	\label{fig:noon-closed}
-%\end{figure}
-
-%\subsection{resolution\_action\_update\_date}
+(or ``creates'') a large number of SRs with a generated time-stamp of exactly 
+midnight (00:00:00) or noon (12:00:00). Such behavior distorts
+ the duration of these SRs by providing inaccurate dates. The 
+ distribution by Agency for this anomaly indicates 
+ that \textgreater90\% of these ``created\_at\_noon'' SRs come from 
+ just two Agencies: DOB and DSNY.  Further investigation
+ reveals that  a single Agency, DSNY, is responsible 
+ for \textgreater99\% of the SRs ``closed-exactly-at-noon''.
+
 \label{sec: resolutionaction}
-When an SR is updated, the 311 software automatically 
+\textbf{Issues with resolution\_action\_update\_date:} When an 
+SR is updated, the 311 software automatically 
 populates the resolution\_action\_update\_date. Some of 
-those SR updates are happening long after the SR is 
+those  updates are happening long after the SR is 
 closed. There are a total of 7460 SRs that are updated \textgreater30 days 
@@
 
-%\begin{figure}[tbp]
-%	\centering
-%	\includegraphics[width = \textwidth]{post_closed_violin.pdf}
-%	\caption{Post-Closed resolution\_action\_update\_date(s) 
-%	\textgreater30 days}
-%	\label{fig:resolution-violin}
-%\end{figure}
-
-\section{Accuracy and precision}
+\begin{figure}[tbp]
+	\centering
+	\includegraphics[width = \textwidth]{post_closed_violin_chart.pdf}
+	\caption{Post-Closed resolution\_action\_update\_date(s) \textgreater30 days}
+	\label{fig:resolution-violin}
+\end{figure}
+
+\subsection{Accuracy and precision}
 \label{sec:precision}
 An question of precision vs. accuracy arises with the Latitude 
-and Longitude fields. Both Latitude are 
-expressed as a 14-decimal number, e.g. 40.86769186022511. Given 
+and Longitude fields. Both Latitude are expressed as 
+a 14-decimal number, e.g. 40.86769186022511. Given 
 that 1 degree of latitude at the equator is equal to 111.044736 
 kilometers, the ``1'' at the end of that number represents 
-approximately 1.1104 nanometers (1/1,000,000,000 of a meter). For 
+approximately 1.1 nanometers (1/1,000,000,000 of a meter). For 
 reference a DNA molecule is approximately 2nm in width. Clearly 
@@
 
-\section{Redundant \& Duplicate fields}\label{sec:duplicates}
+\subsection{Redundant \& Duplicate fields}
+\label{sec:duplicates}
 During this analysis, several redundant fields were observed which
@@
 
-%\subsection{latitude \& longitude and the location fields}
 \label{sec:latlong}
-The location field is a pure concatenation of the latitude 
+\textbf{latitude \& longitude and location:} The location 
+field is a 100\%pure concatenation of the latitude 
 and longitude fields with a comma and parenthesis 
-added; 100\% duplicative. Thus the  location field seems questionable 
-as the data is arguably more difficult to extract and use than the 
-two individual fields. Example:
+added. Thus the  location field seems questionable as the value, 
+and the data is arguably more difficult to 
+extract and use than the two individual fields. Example:
 
@@
 
-%\subsection{borough and park\_borough fields}
 \label{sec:parkborough}
-Borough and park\_borough: These two fields are 100\% matches; fully redundant.
-
-%\subsection{borough and borough\_boundaries fields}
+\textbf{borough \&  park\_borough fields:} Borough and 
+park\_borough: These two fields are 100\% matches; fully redundant.
+
 \label{sec:boroughboundaries}
-Borough and borough\_boundary fields are a 98.3\% match. This 
+\textbf{borough \& borough\_boundaries:} Borough and 
+borough\_boundary fields are a 98.3\% match. This 
 is an example of the the problem with two ``near-duplicate'' 
@@
 
-%\subsection{borough and taxi\_company\_borough fields}
 \label{sec:taxicompanyborough}
-Borough and taxi\_company\_borough fields that despite their 
-names, are nearly completely different; only a 0.05\% match, suggesting
+\textbf{borough \& taxi\_company\_borough:} Borough and 
+taxi\_company\_borough fields that despite their names, are 
+nearly completely different with only a 0.05\% match, suggesting 
 that these two fields are used very differently. The taxi\_company\_borough
- field is used exclusively by the TLC which governs all cars for hire.
-
- %\subsection{incident\_zip and zip\_codes fields}
+ field is used exclusively by the TLC which governs all cars for hire. Any
+ discussion of duplication with other borough-related fields 
+ should include a discussion with TLC as to the usage of these similar,
+ but non-matching, fields.
+
 \label{sec:zipcodes}
- The zip\_code field is one of the six computed fields and its validity 
- was explored in Section \ref{sec:zip-codes} and found to be 
+\textbf{incident\_zip \& zip\_codes:} The zip\_code field is one 
+of the six computed fields and its validity 
+ was previously explored and found to be 
  lacking; 56\% of the entries are invalid. The companion 
@@
  indicates that the computed zip\_code field should be eliminated
- until such time as accuracy issues can be addressed.
-
- %\subsection{police\_precinct and police\_precincts fields} 
+ or masked from the public until such time as accuracy 
+ issues can be addressed.
+
 \label{sec:police} 
-Both the police\_precinct and police\_precincts are among the computed 
+\textbf{police\_precinct \& police\_precincts:} Both the 
+police\_precinct and police\_precincts are among the computed 
 fields not cited in the Data Dictionary. However, their usage is easy to 
-discern; they represent the NYPD precinct that corresponds to the
-incident address. These two fields are near-duplicates with 99.94\% of the entries 
-matching. (The validity of these fields is 
-covered in Section \ref{sec:police-precincts}.) Since the underlying 
-computational process is unknown, we are unable to determine which 
+discern; they presumably represent the NYPD precinct 
+corresponding to the incident address. These two 
+fields are near-duplicates with 99.9\% of the entries 
+matching. Since the underlying computational process 
+is unknown, we are unable to determine which 
 field is ``more correct'' field. However having two fields with 99.9\% 
-duplicity indicates that only one police precinct field is necessary.
-
-% \subsection{agency and agency\_name fields}
+duplicity indicates that perhaps only one police 
+precinct field is necessary.
+
 \label{sec:agencyname}
- Agency and agency\_name fields: While not duplicates per se, these fields 
- have a 1:1 correspondence.  The Agency field contains abbreviations 
- for the City agencies such as NYPD, DOT, HPD, etc. The agency\_name 
- field contains the full name of these organizations. It seems 
- redundant to include both as the abbreviations are commonly used
- and well understood.
-
-%\subsection{landmark and street\_name}
+\textbf{agency \& agency\_name:} The agency and agency\_name 
+fields, while not duplicates per se, have a 1:1 correspondence. The 
+Agency field contains abbreviations for the City agencies, e.g. NYPD, 
+DOT, HPD, DSNY, etc. The agency\_name field contains the full name 
+of these organizations. It seems redundant to include both as 
+the abbreviations are commonly used and well understood.
+
 \label{sec:landmark}
-Landmark and street\_name: The landmark field is listed in the Data Dictionary as ``Can refer to 
-any noteworthy location, including but not limited to, parks, 
-hospitals, airports, sports facilities, performance spaces, etc.'' Our analysis
-found that was not the case. To be sure many of the entries in 
+\textbf{landmark \& street\_name:} The landmark field is listed 
+in the Data Dictionary as ``...any noteworthy location, 
+including but not limited to, parks, hospitals, airports, sports 
+facilities, performance spaces, etc.'' Our analysis
+found that was not the case. To be sure, many of the entries in 
 the landmark field do contain landmark names, e.g.``Pennsylvania 
-Station'', ``LaGuardia Airport'', etc., the vast majority 
+Station'', ``LaGuardia Airport'', etc.,  but the vast majority 
 of the entries are street names, e.g. ``Fenton Avenue'', ``Steinway 
-Street'', ``Broadway''. These entries are very similar to the 
+Street'', ``Broadway'', etc.. These entries are very similar to the 
 street\_name field with a 62\% match, enough to be indicative 
-of duplicate usage. Even the non-matches (excluding blanks) 
-appear to be matches except for minor spelling and 
-nomenclature changes, e.g. ``NINTH AVE'' \& ``9 AVE''.
-
-\begin{comment}
-\begin{table}[ht]
-    \centering
-    \caption{Non-matches between 'street\_name' and 'landmark' fields}
-	    \begin{tabular}{>{\raggedright\arraybackslash}p{5cm} >
-	 	{\raggedright\arraybackslash}p{5cm}}
-	 	\toprule
-	      \textbf{street\_name} & \textbf{landmark} \\
-	      \midrule
-	        MACDOUGAL ST & MAC DOUGAL ST \\
-	        NINTH AVE & 9 AVE \\
-	        SIXTH AVE & 6 AVE \\
-	        EAST FIRST ST & EAST 1 ST \\
-	        FOURTH AVE & 4 AVE \\
-	        SAINT LAWRENCE AVE & ST LAWRENCE AVE \\
-	        MT HOPE PL & MOUNT HOPE PL \\
-	        PENN STA & PENNSYLVANIA STA \\
-	      \bottomrule
-	    	\end{tabular}
-    \label{landmark}
-\end{table}
-\end{comment{
-
-%\subsection{cross\_street\_1 \& intersection\_street\_1 and cross\_street\_2 
-\& intersection\_street\_2}
-\label{sec:street1}
-Cross\_street\_1 \& intersection\_street\_1 and cross\_street\_2 
-\& intersection\_street\_2: There are two sets of street pairs in addition to 
-the incident\address field. These pairs are:
+of some level of duplicative usage. Even the non-matches 
+(excluding blanks) appear to be matches except for 
+minor spelling and nomenclature changes, 
+e.g. ``NINTH AVE'' \& ``9 AVE''. It's worth investigating what is going on
+with the usage of the landmark field.
+
+\label{sec:cross-street}
+\textbf{cross\_street\_1 \& intersection\_street\_1 and cross\_street\_2 
+\& intersection\_street\_2: } There are two sets of street pairs provided 
+in addition to the incident\_address field. These pairs are:
 
@@
 
-These two street pairs used to assist in identifying the address, as 
-is common in New York City. These two street pairs
-have 88\% duplicates where cross\_street\_1 matches 
+These two street pairs are used to assist in identifying the address, 
+as is common in New York City. These two street pairs
+have 88\% duplicates where both cross\_street\_1 matches 
 intersection\_street\_1 and cross\_street\_2 
 matches intersection\_street\_2. We believe only one of these
-street pairs are necessary. However it continues to illustrate the 
+street pairs are necessary, but which one? This illustrates the 
 problem of two similar data fields that largely, but not exactly, match. In
-this case there are 12\% that do not match; which field to trust.
-
-\textit{(Note: Address standardization was applied to the street 
-addresses using the R ``campfin'' package. This prevents 
-generating a non-match between locations such as 
+this case there are 12\% that do not match; which field to trust?
+
+\textit{(Note: Address standardization was applied to  street 
+addresses using the R \textit{campfin} package. This standardization 
+prevents generating a non-match between locations such as 
 ``240 E 69\textsuperscript{th} ST'' and ``240 E 69\textsuperscript{th} Street'' 
-which are clearly the same address save for formatting and spelling differences.)}
-
-In addition to the matching values, we encountered a number 
-of \textit{near-matches}, situations where the 
-cross\_street and intersection\_street are nearly identical, 
-e.g. HARMON DR and HARMON RD. To identify these 
-near-matches we measured the \href{https://en.wikipedia.org/wiki/Hamming_distance}
-{Hamming Distance} between the two fields. Hamming Distance is a 
+which are clearly the same address save for formatting 
+and slight spelling differences.)}
+
+\textbf{Near matches:} In addition to the matching values, we 
+encountered a number of \textit{near-matches}; situations where the 
+cross\_street and intersection\_street are nearly, but not exactly, identical. 
+For example, HARMON DR and HARMON RD. To identify these near-matches 
+we measured the \href{https://en.wikipedia.org/wiki/Hamming_distance}
+\textit{Hamming Distance} between the two fields. The Hamming Distance is a 
 measure of how many letters must be changed for the two 
-fields to match. For this analysis, we chose a Hamming Distance of 2. Here 
-are some sample results for both cross\_street and intersection\_street.
-
-\begin{comment}
+fields to match. For this analysis, we chose a Hamming Distance of two (2).
+ Here are some example results for near-matches for both 
+ cross\_street\_1 and intersection\_street\_1. 
+ (Cross\_street\_2 and intersection\_street\_2 show similar results.)
+
 \begin{table}[tbp]
@@
 \end{table}
-\begin{table}[tbp]
-    \centering
-    \caption{Summary of cross\_street\_2 and intersection\_street\_2}
-    \label{tab:summary_cross_intersection_2}
-	    \begin{tabular}{l r r}
-	        \toprule
-	        \textbf{category} & \textbf{count} & \textbf{percentage} \\
-	        \midrule
-	        Matching -- non-blank      & 4,012,683 & 62.75     \\
-	        Matching -- both blank     & 1,623,276 & 25.38     \\
-	        Non-matching                &   750,543 & 11.74     \\
-	        cross\_street\_2\_blank     &   222,788 & N/A       \\
-	        intersection\_street\_2\_blank &   517,431 & N/A       \\
-	        Near-match                  &     1,500 & 0.023456  \\
-	        \bottomrule
-	    \end{tabular}
-\end{table}
-
-\begin{table}[tbp]
-	\centering
-	\caption{Sample of near-matching cross\_street\_2 and 
-     	intersection\_street\_2 (both non-blank)}
-     	\label{tab:near_matching_cross_intersection_2}
-	     	\begin{tabular}{l l l r}
-	      \toprule
-	      \textbf{cross\_street\_2} & \textbf{intersection\_street\_2} 
-	      & \textbf{agency} & \textbf{hamming\_distance} \\
-	      \midrule
-	        227 ST         & 225 ST                & DOT    & 1 \\
-	        65 ST          & 65 PL                 & DOT    & 2 \\
-	        18 AVE         & 17 AVE                & DOT    & 1 \\
-	        BEACH 110 ST   & BEACH 109 ST          & DOT    & 2 \\
-	        88 ST          & 87 ST                 & DOT    & 1 \\
-	        WEST 114 ST    & WEST 113 ST           & DOT    & 1 \\
-	        5 AVE          & 4 AVE                 & DOT    & 1 \\
-	      \bottomrule
-	    	\end{tabular}
-\end{table}
-\end{comment}
-
- %\subsection{Shrinking file size by removing duplicates}
+
+\subsection{Shrinking file size by removing duplicates}
  \label{sec:filesize}
@@
 for this large dataset that small percentage still amounts 
-to 110, 715 non-matching occurrences. The authors  propose 
+to 110,715 non-matching occurrences. The authors  propose 
 deleting the following fields from the data set.
@@
 	\item park\_borough:  This field is a 100\% match with the 
-	borough field. Removal of this field will not result in any loss of 
-	data or data quality.
+	borough field, thus no data loss.
 
@@
 	added. The data is a 100\% duplication of the latitude 
-	and longitude fields. Removal of this field will not 
-	result in any data loss.
+	and longitude fields, thus no data loss.
 
@@
 	is more correct, but feel that removal of one of the precinct fields 
-	is prudent and would result in very minimal data loss.
+	is prudent and would result in minimal data loss.
 
@@
 	two pairs of fields each have an 88\% match. We would recommend 
-	deleting the two intersection\_street fields while acknowledging 
+	deleting the two intersection\_street (1 \& 2) fields while acknowledging 
 	some loss of data in doing so.
@@
 	as shown in the \ref{sec:case-study-zip-codes} can lead to 
-	dangerous mistakes when used for analytical purposes. 
-	We recommend deleting this field.
+	dangerous mistakes when used for analytical purposes. We 
+	recommend deleting this field.
 \end{itemize}
@@
 blank. A similar situation exists for the road\_ramp field 
-(99.75\% blank), the vehicle\_type field (99.71\%blank), and 
-due\_date which is 99.62\% blank. Several additional several fields
+(99.8\% blank), the vehicle\_type field (99.7\%blank), and 
+due\_date which is 99.6\% blank. Several additional several fields
 are mostly blank: bridge\_highway\_direction, bridge\_highway\_name, 
 bridge\_highway\_segment, and taxi\_pick\_up\_location; all 
-have \textgreater99\% blanks. Most of these fields 
+have \textgreater99\% blanks. Most of these latter fields 
 appear to used exclusively by the TLC.
 
-\section{Data Dictionary Observations} 
+\subsection{Data Dictionary Observations} 
 \label{sec:datadictionary}
-The Data Dictionary could use an update. We found several 
+The Data Dictionary could benefit from an update. We found several 
 discrepancies between the Data Dictionary and the 
@@
 	as constraints or the domain of legal values. For example, the 
-	incident\_zip field is specified as ``text''. That's probably not the 
-	best definition; perhaps	``categorical'' with notes indicating that 
+	incident\_zip field is specified as ``text''. That is probably not the 
+	best definition; perhaps	``categorical'' with a note indicating that 
 	the data can contain only numeric characters and is 
@@
 	inaccurate. For example, the Data Dictionary indicates that 
-	the ``status'' field has values of \textit{assigned, canceled, closed}, 
+	the status field has values of \textit{assigned, canceled, closed}, 
 	or \textit{pending}. However, we found additional status 
@@
 	vehicle\_type, taxi\_pick\_up\_location, road\_ramp, city) 
-	suffer similar inaccuracies in their specified domain of values.
+	suffer similar inaccuracies in their specified domain of values as
+	shown in the Data Dictionary.
 
@@
 \label{sec:recommendations}
-
-\begin{itemize}
-	\item First, and most importantly, the Open Data Team will need 
-	to undertake a dedicated effort to identify these 
+We propose the following recommendations by the Open Data Team to address
+the data cleanliness issues addressed in this paper.
+
+\begin{itemize}
+	\item First, the Open Data Team will need 
+	to undertake a dedicated effort to investigate these 
 	data anomalies, confirm the veracity of the issue, 
 	and engage with the appropriate Agency Open Data 
-	Representatives to resolve these issues. That effort will not 
-	be easy or fast, and in some cases software changes will 
+	Representatives to seek resolution. That effort will not 
+	be easy or fast, and in some cases software changes may  
 	be required. Coordination and mutual approval of any 
@@
 
-	\item Secondly, standards and protocols will need to be 
+	\item Secondly, standards and protocols need to be 
 	identified, agreed to, and commitment obtained. For example, 
-	the domain of legal values, data accuracy, address normalization, etc..
+	the domain of legal values, data accuracy, address normalization, 
+	data usage, data definitions, etc..
 
@@
 	resolved through enforcement by software. For example, the 
-	software could easily prevent invalid values, such as zip 
-	codes. Similarly for other data issues. One key area are is the 
+	software should be able to easily prevent invalid values, such as zip 
+	codes, and similarly for other data issues. One key area are is the 
 	created\_date and closed\_date which are frequently cited 
-	when determining responsiveness of various Agency services; software
-	could conceivably prevent illogical situations such as an SR closed 
-	before it is opened.
+	when determining responsiveness for various Agency services; software
+	could conceivably prevent illogical situations such as an SR being
+	``closed'' before it is ``opened''.
 
@@
 	process from an external Agency integration. Such processes distort 
-	the accurate capture of an SR duration and should be eliminated.
-
-	\item Investigate moving additional Agencies to the core 
-	311 software system where data rules would 
-	be more consistently enforced.
+	the accurate capture of an SR duration and should be eliminated in
+	favor of more accurate dates.
 
@@
 	that are not identified in the Data Dictionary. Several of 
-	these computed fields suffer from data inaccuracies. We  
+	these computed fields suffer from significant data inaccuracies. We  
 	recommend updating the Data Dictionary at a minimum and 
-	possibly hiding those fields from public usage until such time as 
+	possibly masking those fields from public usage until such time as 
 	their accuracy can be ascertained.
@@
 	as the borough and park\_borough fields, the duplication is 100\%. In 
-	other cases, the duplication is in the 80-90\% range. These fields and 
-	their utility should be examined to determine if inclusion in the 
+	other cases, the duplication is in the 80-90\% range. These near-duplicate 
+	fields and 	their utility should be examined to determine if inclusion in the 
 	311 SR dataset is warranted. Our analysis indicates that as much as 12\% 
-	size reductions are achievable, likely more.
+	size reductions are achievable with a first pass, possibly more upon deeper
+	analysis.
 
@@
 	landmark, road\_ramp, bridge\_highway\_segment, and 
-	taxi\_pick\_up\_location) should be standardized to USPS format. Software 
-	packages can perform this standardization.
+	taxi\_pick\_up\_location) should be standardized to USPS format. 
+	Software packages can easily perform this standardization.
 
@@
 	represented. These include ``UNKNOWN'' , ``N/A'', ``NA'', ``na'', and 
-	a blank field, e.g. `` ''. Protocols should be established 
-	regarding how to represent missing data.
-\end{itemize}
-
-%\section{Protocol Suggestions} \label{sec:protocol}
+	a blank field, e.g. `` ''. Protocols should be established regarding 
+	how to represent missing data.
+\end{itemize}
 

### 2024-07-12 manuscript/manuscript.tex
jds style

@@
-\documentclass[12pt, titlepage]{article}
-
-\usepackage[margin = 1in]{geometry}
-\usepackage{amsmath}
-\allowdisplaybreaks
-\usepackage{authblk}
-\usepackage{setspace}
-\usepackage{natbib}
-\usepackage{hyperref}\hypersetup{colorlinks=true, citecolor=blue}
-% \usepackage{parskip}
-\usepackage{textgreek}
-\usepackage{graphicx}
-\usepackage{booktabs} % For better looking tables
+\documentclass[linenumber]{jdsart}
+
+\volume{0}
+\issue{0}
+\pubyear{2022}
+\articletype{research-article}
+\doi{0000}
+
 \usepackage{siunitx} % For alignment of numbers
-\usepackage{placeins}
-\usepackage{caption}  % For custom captions
-\usepackage{float}   % Required for the H specifier
-\usepackage{adjustbox}
-\usepackage[]{lineno}
-\usepackage{comment}
-\linenumbers*[1]
-
-%% patches to make lineno work better with amsmath
-\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
-	\expandafter\let\csname old#1\expandafter\endcsname\csname
-	#1\endcsname
-	\expandafter\let\csname oldend#1\expandafter\endcsname\csname
-	end#1\endcsname
-	\renewenvironment{#1}%
-	{\linenomath\csname old#1\endcsname}%
-	{\csname oldend#1\endcsname\endlinenomath}}%
-\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
-	\patchAmsMathEnvironmentForLineno{#1}%
-	\patchAmsMathEnvironmentForLineno{#1*}}%
-\AtBeginDocument{%
-	\patchBothAmsMathEnvironmentsForLineno{equation}%
-	\patchBothAmsMathEnvironmentsForLineno{align}%
-	\patchBothAmsMathEnvironmentsForLineno{flalign}%
-	\patchBothAmsMathEnvironmentsForLineno{alignat}%
-	\patchBothAmsMathEnvironmentsForLineno{gather}%
-	\patchBothAmsMathEnvironmentsForLineno{multline}%
-}
-
 \sisetup{
@@
 
-% control floats
-\renewcommand\floatpagefraction{.9}
-\renewcommand\topfraction{.9}
-\renewcommand\bottomfraction{.9}
-\renewcommand\textfraction{.2}
-\setcounter{totalnumber}{50}
-\setcounter{topnumber}{50}
-\setcounter{bottomnumber}{50}
-
 \usepackage{xcolor}
@@
 
+\usepackage{comment}
+\usepackage{booktabs, textgreek}
+
+\begin{document}
+
+\begin{frontmatter}
+
 \title{Principles for Open Data Curation: A Case Study with the New
 York City 311 Service Request Data}
-
-\author[1]{David Tussey}
-\author[2]{Jun Yan}
-\affil[1]{Former Executive Director, NYC DoITT}
-\affil[2]{Department of Statistics, University of Connecticut}
-
-\begin{document}
-\maketitle
+\runtitle{Principles for Open Data Curation}
+
+\author[1]{\inits{D.}\fnms{David}~\snm{Tussey}}
+\author[2]{\inits{J.}\fnms{Jun}~\snm{Yan}}
+\address[1]{\institution{NYC DoITT}, \cny{USA}}
+\address[2]{Department of Statistics,
+  \institution{University of Connecticut}, \cny{USA}}
 
@@
 
-\bigskip
-
-\noindent
-\textit{Keywords:}
-	Data cleansing;
-	Data Curation;
-	Data science;
-	NYC Open Data;
-	Open data;
-	Smart Cities;
-	Transparency;
 \end{abstract}
 
-\doublespacing
+\begin{keywords}
+  \kwd{Data cleansing}
+  \kwd{Data Curation}
+  \kwd{Data science}
+  \kwd{NYC Open Data}
+  \kwd{Open data}
+  \kwd{Smart Cities}
+  \kwd{Transparency}
+\end{keywords}
+
+\end{frontmatter}
 
@@
 \label{sec:intro}
+
 In the early 21st century, the open data movement began 
@@
   \includegraphics[width=0.8\textwidth]{SR_by_Complaint_Type.pdf} 
-  \caption{Top 20 complaint\_type(s) and Cumulative Percentage} 
+  \caption{Top 20 complaint type(s) and Cumulative Percentage} 
   \label{fig:SR_complaints}
@@
 	\centering
-	\includegraphics[width=\textwidth]{2-year-trend-SR_closed_by_minute_of_busiest\_day.pdf}
+	\includegraphics[width=\textwidth]{2-year-trend-SR_closed_by_minute_of_busiest_day.pdf}
 	\caption{SRs Closed Minute-by-Minute on Busiest Day}
@@
 
-\bibliographystyle{asa}
+\bibliographystyle{jds}
 \bibliography{ref}

### 2024-09-08 manuscript/manuscript.tex
rephrased opening paragraph in cleansing sectoin

@@
 \label{sec:issues}
-What is data cleansing?  Wikipedia 
-\href{https://en.wikipedia.org/wiki/Data_cleansing}{Data Cleansing} 
-offers this definition.
-
-\begin{quote}\textit{Data cleansing or data cleaning is the process of detecting and 
-correcting (or removing) corrupt or inaccurate records from a record set, 
-table, or database and refers to identifying incomplete, incorrect, 
-inaccurate or irrelevant parts of the data and then replacing, 
-modifying, or deleting the dirty or coarse data.}
-\end{quote}
+
+Data cleansing refers to the process of identifying and rectifying
+errors, inconsistencies, and inaccuracies within datasets to ensure
+they are of high quality and reliable for analysis
+\citep{maletic2005data, hosseinzadeh2023data}. The process
+typically involves removing duplicate records, handling missing or
+incomplete data, correcting mislabeled or inaccurate entries, and
+standardizing data formats. In the context of open data, cleansing is
+especially important as open datasets often come from diverse,
+uncoordinated sources, leading to variations in data quality,
+completeness, and consistency. Without proper cleansing, the utility
+of open data can be severely limited, affecting its reliability for
+research, policy-making, and innovation. The main purpose of cleansing
+open data is to ensure that it is accurate, consistent, and usable
+across multiple platforms and by various stakeholders. This improves
+the trustworthiness of the data and enables better decision-making,
+more accurate analysis, and the integration of data into machine
+learning models or other systems.
+
+% What is data cleansing?  Wikipedia 
+% \href{https://en.wikipedia.org/wiki/Data_cleansing}{Data Cleansing} 
+% offers this definition.
+
+% \begin{quote}\textit{Data cleansing or data cleaning is the process of detecting and 
+% correcting (or removing) corrupt or inaccurate records from a record set, 
+% table, or database and refers to identifying incomplete, incorrect, 
+% inaccurate or irrelevant parts of the data and then replacing, 
+% modifying, or deleting the dirty or coarse data.}
+% \end{quote}
 

### 2024-09-08 manuscript/manuscript.tex
drafted discussion

@@
 \usepackage{booktabs, textgreek}
+
+%% float control
+\renewcommand\floatpagefraction{0.75}
+% \renewcommand\topfraction{.8}
+% \renewcommand\bottomfraction{.8}
+% \renewcommand\textfraction{.2}
+\setcounter{totalnumber}{50}
+\setcounter{topnumber}{50}
+\setcounter{bottomnumber}{50}
 
@@
 
+\jy{The 'title' in each figure could be removed because we have
+  captions. The fonts of the annotations in the figures need to be
+  increased. The code generating them need to be tidied for future
+  sharing.
+}
+
 \begin{figure}[tbp]
@@
 
+Our study highlights the vital importance of robust data curation in
+maintaining the reliability and utility of open datasets, with a close
+examination of the NYC 311 data. As these datasets are applied across
+diverse fields such as public services and academic research, issues
+like inconsistencies, missing values, and formatting errors can
+significantly undermine the quality of insights. Trustworthy machine
+learning systems, which rely heavily on high-quality data, are
+particularly vulnerable to errors when data is not properly
+curated. For example, biases or inconsistencies in input data can
+compromise both the accuracy and fairness of predictive models
+\citep{rahm2000data}.
+
+This research further emphasizes the necessity of harmonizing data
+from multiple city agencies, particularly in resolving discrepancies
+in field names, formats, and definitions. Trustworthy machine learning
+depends on not only accurate and clean data but also consistent data
+across sources. This challenge becomes more acute in datasets spanning
+long time periods, where changes in agency structures or reporting
+standards can introduce subtle biases. As \citet{rahm2000data} noted,
+ensuring long-term data consistency is crucial for longitudinal
+studies and predictive modeling, as even minor shifts in data can
+result in significant deviations in outcomes.
+
+Finally, the intersection of data curation and machine learning for
+public policy applications opens new opportunities for enhancing
+governance. By improving data quality, machine learning models can
+more effectively predict trends, allocate resources, and identify
+issues such as service delays. The COVID-19 pandemic demonstrated the
+critical role of real-time, trustworthy data in responding to public
+needs. Without proper data curation, the accuracy and timeliness of
+machine learning insights would have been compromised, affecting
+crucial decision-making processes. Future efforts could focus on
+automating data curation within real-time data pipelines, ensuring
+that the data feeding into machine learning models remains both clean
+and reliable \citep{chu2016data}.
+
 \bibliographystyle{jds}

### 2024-09-15 manuscript/manuscript.tex
added supp section
went over first two sections

@@
 
-\jy{How about putting all these sections into one section on case
-  study? They can be subsections in the merged section.
-  After that, we can have a section on recommendations for open
-  data curation based on what we learn from the case study. For each
-  recommended principle, we can link to the specifics of the case
-  study.}
-
-The paper is organized as follows:
-Section~\ref{sec:data} `\textit{NYC 31 Service Request Data} offers a 
-look at the history of the 311system. Long-term trends are 
-presented in a 10-year data analysis (2014-2023).
-Section~\ref{sec:issues} \textit{Data Cleansing Issues} offers a 
-discussion of specific data cleansing issues impacting data quality and 
-curation efficiency. These issues include structural issues, compliance 
-with the Data Dictionary, and missing, blank, or N/A entries. We also
-explore field compliance with a domain of legal or acceptable values.
-Next we deal with several important issues surrounding logical inconsistencies 
-and concerning patterns found in the data. We also explore an issue of precision 
-versus accuracy. This section concludes by identifying duplicate/redundant data fields
-and Data Dictionary observation. Section~\ref{sec:recommendations} 
-\textit{Recommendations} provides actionable suggestions for 
-mitigating or resolving the identified issues. Section~\ref{sec:discussion} ``Discussions'' 
-concludes with an encapsulation of key insights and implications of our research.
+The paper is organized as follows. Section~\ref{sec:data} provides 
+an overview of the history of the NYC 311 SR system and presents 
+a summary of SR counts over the 10-year period from 2014 to 2023. 
+Section~\ref{sec:issues} delves into specific data cleansing 
+challenges affecting data quality and curation efficiency, 
+including structural problems, adherence to the data dictionary, 
+and the presence of missing, blank, or N/A entries. This section 
+also investigates field compliance with legal or acceptable values, 
+highlights logical inconsistencies, and examines concerning patterns 
+in the data. We explore the balance between precision and accuracy 
+and identify duplicate or redundant fields, along with observations 
+on the Data Dictionary. Section~\ref{sec:recommendations} offers 
+practical recommendations for mitigating or resolving these issues, 
+while Section~\ref{sec:discussion} encapsulates key insights and 
+discusses the broader implications of our findings.
 
@@
 include the addition of online and mobile app channels in 2009, a
+\jy{should we mention the system upgrade just before the pandemic to
+  document this historical event?}
 record high of 348,463 monthly service requests in August 2020 due to
@@
 
-To begin, we looked at a 10-year period (2014-2023) to
-observe a longer time-frame; see
-Figure~\ref{fig:10-yr-monthly}. During that time-frame,
-the number of 311 SRs grew by nearly 50\%.
+\begin{figure}[tbp]
+	\centering
+  	\includegraphics[width=0.8\textwidth]{10-year-trend_yearly.pdf}
+ 	\caption{10-year (2014-2023) annual SR Counts.}
+  	\label{fig:10-yr}
+\end{figure}
+
+Our investigation focused on a 10-year period (2014--2023).
+Figure~\ref{fig:10-yr} shows the annual counts of NYC 311 SR over the
+10-year period from 2014 to 2023, revealing a clear upward trend in
+public engagement with the system. The most significant increase
+occurred in 2020, when the number of requests surged to 3.23 million
+due to the COVID-19 pandemic, illustrating the essential role of the
+311 system in supporting residents during times of crisis. Overall, SR
+counts increased by nearly 49.5\% between 2014 and 2023, reflecting
+growing public reliance on the system for addressing non-emergency
+issues. The steady rise in SRs can be attributed to the increased
+accessibility of the 311 system via online and mobile platforms, as
+well as heightened public awareness of the service. While the spike in
+2020 was exceptional, the broader trend indicates sustained growth in
+service request volume over the decade, highlighting the system's
+evolving role in managing both routine city operations and
+extraordinary events.
 
@@
 
-\begin{figure}[tbp]
-	\centering
-  	\includegraphics[width=0.8\textwidth]{10-year-trend_yearly.pdf}
- 	\caption{10-year (2014-2023) Monthly SR Counts}
-  	\label{fig:10-yr-monthly}
-\end{figure}
-
-Identifying the responsible agency is essential 
-for troubleshooting discrepancies. This identification also assists in 
-determining whether an issue is specific to an Agency or indicative 
-of a broader, systemic problem. Six agencies account for 90\% of 
-the complaints: New York Police Department (NYPD), 
+Figure~\ref{fig:SRcountbyAgency} provides a breakdown of SRs by
+agency, showing the cumulative percentage of SRs handled by each
+agency over a given time period. This figure highlights that the
+distribution of SRs is concentrated among a few key agencies. The six
+largest agencies are New York Police Department (NYPD), 
 Housing Preservation and Development (HPD),
@@
 and Department of Parks and Recreation (DPR).
+They collectively handle over 90\% of the total SR volume. This is
+indicative of the critical roles these agencies play in managing
+public concerns, ranging from noise complaints and housing issues to
+sanitation and transportation problems. The NYPD alone accounts for a
+substantial portion of the total SRs, underscoring its central role in
+addressing complaints related to public safety and order. The
+remaining 10\% distributed across a large number of smaller
+agencies. This concentration of requests points to the systemic
+reliance on a few key departments for resolving most non-emergency
+issues, further emphasizing the importance of optimizing data handling
+and curation processes within these high-volume agencies to ensure
+efficiency and responsiveness.
 
@@
 
-Figure~\ref{fig:SR_complaints} shows that the top 20 complaint 
-types account for 70\% of the total 220 different types. Notably, 
-these top complaints include several variations of noise-related 
-issues, which collectively represent 22\% of all complaints, the 
-most frequently occurring type of complaint.
-
 \begin{figure}[tbp]
@@
 \end{figure}
+
+Figure~\ref{fig:SR_complaints}
+illustrates the top 20 complaint types, accounting for 70\% 
+of all Service Requests (SRs), and their cumulative percentage. The 
+distribution of complaints is skewed, with a small number of issues 
+dominating the total SR volume. Noise-related complaints, including 
+residential, commercial, and street noise, make up 22\% of all 
+complaints, making them the most frequent issue in the NYC 311 system. 
+Other prominent categories include illegal parking, heat/hot water 
+complaints, and blocked driveways. The cumulative percentage curve 
+shows that after the top 20, the remaining complaint types are spread 
+thinly across the remaining 30\% of SRs. This concentration suggests 
+that improving responses to the most frequent complaints could have 
+a significant impact on overall service efficiency and resident 
+satisfaction. It also provides insights into the operational 
+pressures faced by the city agencies responsible for handling these 
+high-volume complaints.
 
@@
 
+\section*{Supplementary Material}
+The NYC 311 SR data and code used to produce the results reportred in
+this paper are available [FixMe].
+\jy{need to think: we need to keep a copy of the downloaded data;
+  if the data portal changes and our code may not work.}
+
 \bibliographystyle{jds}

### 2024-09-15 manuscript/manuscript.tex
edited section 3

@@
 
-\section{INTRODUCTION} 
+\section{Introduction} 
 \label{sec:intro}
@@
 	\centering
-  	\includegraphics[width=0.8\textwidth]{10-year-trend_yearly.pdf}
+  	\includegraphics[width=0.7\textwidth]{10-year-trend_yearly.pdf}
  	\caption{10-year (2014-2023) annual SR Counts.}
@@
 
-Our investigation focused on a 10-year period (2014--2023).
+Our investigation focused on a 10-year period (2014--2023), downloaded
+on \jy{Give date because the data is updated daily.}
+This version of the data is available as a supplement to facilitate
+reproducibility of the analysis presented in the sequel.
+\jy{How many total? file size? Somewhere we used only two years that
+  need to be clarified.}
 Figure~\ref{fig:10-yr} shows the annual counts of NYC 311 SR over the
@@
   sharing.
+}
+
+\jy{The plot sizes seem to be fixed in the generation code. That's why
+  the fonts are small when scaled. I'd generate them with approxiately
+  the expected sizes in the manuscript. I'd also use aspect ratio
+  0.618 (golden ratio).
 }
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{SRs_by_Agency.pdf}
+	\includegraphics[width=0.7\textwidth]{SRs_by_Agency.pdf}
   	\caption{SR counts by Agency with Cumulative Percentage}
@@
  \centering
-  \includegraphics[width=0.8\textwidth]{SR_by_Complaint_Type.pdf} 
+  \includegraphics[width=0.7\textwidth]{SR_by_Complaint_Type.pdf} 
   \caption{Top 20 complaint type(s) and Cumulative Percentage} 
@@
 
-Many quality criteria are utilized in order to process high-quality 
-data, including:
-\begin{comment}
-\begin{itemize}
-	\item Data validation - this effort can span a number of criteria
-
-	\begin{itemize}
-		\item Mandatory fields: Certain data fields cannot be empty.
-
-		\item Data-types: Certain fields must be of the correct type, e.g. numeric, 
-		character, date, 5 numeric digits, etc. Typically these are specified 
-		in a Data Dictionary.
-
-		\item Domain compliance:  Many data fields must adhere to a specific 
-		domain of values, e.g. statuses, state names, zip codes, gender.
-
-	\end{itemize}   
-	\item Structural errors to include naming conventions, 
-	 fields not in the Data Dictionary, or inconsistent data entry. For example 
-	 intermixing blanks, spaces, NA, N/A, and \textless{}NA\textgreater{} 
-	 all to indicate the absence of data.
-
-	\item Redundant, unnecessary, or irrelevant fields
-
-	\item Logical inconsistencies such as related fields that violate the 
-	nature of that relationship, e.g. a ``due date'' that is before the ``created date''
-
-	\item Accuracy versus precision
-\end{itemize}
-\end{comment}
-
-This purpose of this analysis effort is to identify the presence of 
-errors; not necessarily to correct them. A correction effort should
+Many quality criteria are employed to ensure high-quality data 
+processing. One of the primary efforts is data validation, which spans 
+several critical checks. For instance, mandatory fields must not be 
+left empty, ensuring that key information is always captured. 
+Additionally, certain fields must conform to specific data types, 
+such as numeric, character, or date formats, which are typically 
+outlined in a Data Dictionary. Another essential aspect of validation 
+is domain compliance, where fields must adhere to a predefined set of 
+values, such as statuses, state names, zip codes, or gender. Structural 
+errors also play a significant role in data quality, particularly when 
+naming conventions or data entries are inconsistent. Common issues 
+include fields that do not appear in the Data Dictionary or the 
+inconsistent use of blanks, spaces, NA, N/A, or ``<NA>'' to indicate 
+missing data. Furthermore, redundant or irrelevant fields can clutter 
+datasets, reducing efficiency. Logical inconsistencies are another 
+important consideration, such as related fields that violate expected 
+relationships, like a ``due date'' that precedes the ``created date.'' 
+Lastly, the balance between accuracy and precision is crucial, as both 
+must be carefully managed to ensure reliable and meaningful data.
+
+Here we identify the presence of issues in the NYC 311 SR open data;
+not necessarily to solve them. A solution effort should
 be undertaken only after an investigation as to the why and how 
-such errors came about, and a discussion as to whether or not it 
-even is an \textit{error}.
+such issues came about, and a discussion as to whether or not it 
+is an \textit{error}.
 
@@
 \label{sec:structural}
+
 Structural issues refers to how data is organized, formatted, 
@@
 
-The 311 SR \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}
- identifies 41 data columns (fields) and related information 
- for each column. However, when downloaded the data reveals 
- six additional fields that are not included in the Data Dictionary. 
-(Note that these fields do show up on the portal Column Manager widget
- as ``@computed\_region\_xxxx\_xxxx''.) To some extent one can infer 
- what the these @computed\_fields are and how they are determined, but
- not with certainty. Most importantly is the question, can these fields be reliably 
- used for analytical purposes?  These six ``computed'' fields are:
-
-\begin{itemize}
-	\item zip\_codes
-
-	\item community\_districts
-
-	\item borough\_boundaries
-
-	\item city\_council\_districts
-
-	\item police\_precincts
-
-	\item police\_precinct 
-\end{itemize}
-
-\subsection{Blank \& N/A data}
+The 311 SR 
+\href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}
+identifies 41 data columns (fields) and provides related information 
+for each. However, upon downloading the data, six additional fields 
+appear that are not included in the Data Dictionary. These fields are 
+visible in the portal's Column Manager widget, labeled with the prefix 
+``@computed\_region.'' The additional fields include \texttt{zip\_codes}, 
+\texttt{community\_districts}, \texttt{borough\_boundaries}, 
+\texttt{city\_council\_districts}, \texttt{police\_precincts}, and 
+\texttt{police\_precinct}. While one can infer the meaning and 
+derivation of these computed fields to some extent, uncertainty remains. 
+The critical question is whether these fields can be reliably used for 
+analytical purposes reliable documentation. In particular, which one
+of the two percinct fields should one use? What is the difference
+between the computed zip codes and the incident zipcode among the 41
+documented variables?
+
+\subsection{Missing Data}
 \label{sec:blanks}
@@
 when undertaking analysis. For example, if you wanted to 
-determine if SRs were closed before or after their due\_date, 
-you would be challenged as 99.6\% of the due\_date field is 
+determine if SRs were closed before or after their
+\texttt{due\_date}, you would be challenged as 99.6\% of the
+\texttt{due\_date} field is
 blank. When counting fields for blank or N/A values, they appear 
-to divide into three groups: \textbf{Mostly Empty}, \textbf{Partially Empty}, 
+to divide into three groups:
+\textbf{Mostly Empty}, \textbf{Partially Empty}, 
 and \textbf{Few/None Empty}. 
-
-The Mostly Empty category ranges from 93-99.9\% blank 
-and includes such fields as taxi\_company due\_date, pickup\_location, and 
-landmark. The Partially Empty includes such fields as location\_type, borough, 
-and cross\_street. And the Few/None Empty includes created\_date, 
-complaint\_type, agency, and status. In some cases it may make 
-sense to inquiry as to why some fields are frequently blank. Here is a graphic 
+The Mostly Empty category ranges from 93--99.9\% blank 
+and includes such fields as
+\texttt{taxi\_company due\_date},
+\texttt{pickup\_location}, and \texttt{landmark}.
+The Partially Empty includes such fields as
+\texttt{location\_type}, \texttt{borough}, 
+and \texttt{cross\_street}. And the Few/None Empty includes
+\texttt{created\_date}, \texttt{complaint\_type},
+\texttt{agency}, and \texttt{status}. In some cases it may make
+sense to inquiry as to why some fields are frequently blank.
+Figure~\ref{fig:blank_fields} presents a graphic 
 depiction of total empty (blank \& N/As) for each fields illustrating
 the grouping into the Most, Partial, and Few categories.
+
+\jy{Can we lighten the color of the bar to make the pertentages more
+  legible?}
 
@@
 
-\begin{comment}
 \subsection{Validating Data for Acceptable Values}
 \label{sec:domain}
-Any analytic effort needs to be aware of fields which 
-contain invalid values. Typically such fields would be
-identified and removed from the analysis. Here are the results
-of selected field validation.
-
-\begin{itemize}
-	\item Latitude and Longitude fields fall within the geographic 
-	boundaries of New York City.
-
-	\item The unique\_key field was in-fact unique.
-
-	\item Unfortunately, very few domains of acceptable 
-	values are specified in the Data Dictionary. The below 
-	fields were tested as complying with their domain as 
-	determined by public usage and historical datasets:
-
-	\begin{itemize}
-		\item address\_type
-		\item status
-		\item borough, borough\_boundaries, \& park\_borough 
-		\item data\_channel
-		\item vehicle\_type
-		\item city\_council\_district
-	\end{itemize}
-
-\end{itemize}
-\end{comment}
-
+Any analytic effort must ensure that fields containing invalid values 
+are identified and removed from the analysis. Below are the results of 
+selected field validation. The latitude and longitude fields were 
+found to fall within the geographic boundaries of New York City, and 
+the \texttt{unique\_key} field was indeed unique, as required. 
+Unfortunately, the Data Dictionary specifies very few domains of 
+acceptable values. However, the following fields were tested and found 
+to comply with their expected domains, as determined by public usage 
+and historical datasets: \texttt{address\_type}, \texttt{status}, 
+\texttt{borough}, \texttt{borough\_boundaries}, \texttt{park\_borough}, 
+\texttt{data\_channel}, \texttt{vehicle\_type}, and 
+\texttt{city\_council\_district}. Some fields proved to be problematic 
+as to their compliance with a domain of legal values.
+
+\paragraph{Zip Codes}
 \label{sec:zipcodesissues}
- \textbf{Issues with Zip Codes:} Some fields proved to be problematic 
- as to their compliance with a domain of legal values. For example, all zip codes 
- (two fields: zip\_codes and incident\_zip) should be valid as defined 
- by the USPS database containing all 37,946 valid zip codes. However,
- the computed field zip\_codes proved problematic with over 
-58\% (3.6 million) of the field entries being invalid. And while the more accurate 
-incident\_zip field has only .07\% invalid entries, it still contains
-4163 such errors; not insignificant. The breakdown of the invalid 
-entries in the zip\_code, sorted by Agency shows that the 
-distribution by percentage mirrors the overall breakdown of 
-SRs by Agency, potentially indicating a systemic problem.
-
-\label{sec:case-study-zip-codes}
-\textbf{Case Study: Noise Complaints by Zip Code.}  Scenario: The NYC 
-Office of Nightlife (ONL) wants to know ``What are the top 10 
-zip codes for Noise Complaints over the last two years?'' ONL wants
-to assess the impact of recent efforts promoting a vibrant nightlife 
-in NYC while seeking to ease strained relations between bar 
-and club owners. Let's assume that this analysis 
-uses the zip\_codes field, one of the computed fields known to 
-have validity problems. We then repeat the analysis with the more 
-accurate incident\_zip field. Finally, we subject these two analysis effort 
-to validation against the USPS zip code database.
+All zip codes (two fields: \texttt{zip\_codes} and \texttt{incident\_zip}) 
+should be valid as defined by the USPS database, which contains 
+37,946 valid zip codes. However, the computed field \texttt{zip\_codes} 
+proved problematic, with over 58\% (3.6 million) of its entries being 
+invalid. Although the more accurate \texttt{incident\_zip} field has 
+only 0.07\% invalid entries, this still results in 4,163 errors, which 
+is not insignificant. The breakdown of invalid entries in the 
+\texttt{zip\_codes} field, sorted by Agency, shows that the
+\jy{Which agency? Missing a table? We could prepare a supplement where
+  we can put additional tables/figures without limit.}
+distribution by percentage mirrors the overall breakdown of SRs by 
+Agency, potentially indicating a systemic problem.
 
@@
 
-As indicated, six out of ten zip\_codes are invalid, which corresponds closely 
-with what is observed in the overall dataset (58\% invalid). Whereas 
-the incident\_zip field is completely valid, in-line with the 
-overall incident\_zip validation percentage of 99\%.
-
+Consider a case study with the NYC Office of Nightlife (ONL). Suppose 
+that ONL wants to know, "What are the top 10 zip codes for Noise 
+Complaints over the last two years?" ONL seeks to assess the impact 
+of recent efforts promoting a vibrant nightlife in NYC while also 
+easing tensions between bar and club owners. Assume this analysis uses 
+the \texttt{zip\_codes} field, one of the computed fields known to have 
+validity issues. The analysis is then repeated using the more accurate 
+\texttt{incident\_zip} field. Finally, both analyses are validated 
+against the USPS zip code database. The top 10 zip codes and their 
+validation results are summarized in Table~\ref{tab:zipcodes}. Six out 
+of the ten \texttt{zip\_codes} entries are invalid, closely matching 
+the overall dataset's 58\% invalid rate. In contrast, the 
+\texttt{incident\_zip} field is fully valid, consistent with the 
+overall \texttt{incident\_zip} validation rate of 99\%.
+It is possible that the computed \texttt{zip\_codes} may mean the zip
+code of something differnent from the incident. Without proper
+documentation, it is confusing to have it in the released data.
+
+\paragraph{Police Precincts}
 \label{sec:police-precincts}
-\textbf{Issues with Police Precincts:} A curious case exists when 
-examining two (nearly) identical fields: police\_precincts and 
-police\_precinct. Both of those fields are among the computed fields 
-previously mentioned. Using 
+A curious case arises when examining two nearly identical fields: 
+\texttt{police\_precincts} and \texttt{police\_precinct}. Both fields 
+are among the computed fields previously mentioned. Using the 
 \href{https://www.nyc.gov/site/nypd/bureaus/patrol/precincts-landing.page}
-{NYPD Precinct listings} it is possible to determine the valid 
-police precincts. What we find is that both police\_precinct 
-and police\_precincts  have about 35\% invalid entries. Unfortunately, 
-it's not the same invalid entries. The police\_precincts field, 
-has 2,171,864 invalid entries (35\%). Similarly the 
-police\_precinct field has 2,171,778 invalid entries 
-(also 35\%), but not the same invalid entries.
-
+{NYPD Precinct listings}, it is possible to determine the valid police 
+precincts. Upon validation, both \texttt{police\_precincts} and 
+\texttt{police\_precinct} contain about 35\% invalid entries. However, 
+the invalid entries are not identical. The \texttt{police\_precincts} 
+field has 2,171,864 invalid entries (35\%), while the 
+\texttt{police\_precinct} field has 2,171,778 invalid entries (also 
+35\%), but they do not match.
+
+\paragraph{Community Boards}
 \label{sec:communityboards}
-\textbf{Issues with Community Boards:} Community Boards are the most 
-local, grassroots form of City government; a vital connection 
-between communities, elected officials, and City agencies. Community 
-Boards often service as a measure of the equity of City services 
-throughout the five boroughs. In the 2022-2023 dataset there 
-are 27,276 invalid community\_board entries representing 0.43\% 
-of non-blank data. The distribution of invalid Community Boards 
-by Agency is not consistent with the overall SR Agency distribution,
-indicating that there are likely issues at key Agencies, in 
-this case Taxi \& Limo Commission (TLC), DPR, etc.
-
+Community Boards are the most local, grassroots form of City government, 
+serving as a vital connection between communities, elected officials, 
+and City agencies. They often act as a measure of the equity of City 
+services across the five boroughs. In the 2022-2023 dataset, there are 
+27,276 invalid \texttt{community\_board} entries, representing 0.43\% 
+of non-blank data. The distribution of these invalid entries by Agency 
+is not consistent with the overall SR Agency 
+distribution. This suggests that issues may exist within specific key 
+Agencies, such as the Taxi \& Limousine Commission (TLC) and the 
+Department of Parks and Recreation (DPR).
+
+\paragraph{Community Districts}
 \label{sec:communitydistrict}
-\textbf{Issues with Community Districts:} Another of the computed fields 
-is community\_districts. Community Districts 
-are the geographic boundaries for the Community Boards, but differ as they
-are used by the Department of City Planning (DCP) for environmental, 
-socio-economic, and demographic purposes. It is a geographical 
-division rather than a City government division. Due to how the 
-community\_district data is formatted, it is not possible to 
-establish validity. However, we can determine that the dataset 
-contains 72 unique entries, while there are only 59 
-valid Community Districts.
-
+Another of the computed fields is \texttt{community\_districts}. 
+Community Districts define the geographic boundaries for the Community 
+Boards but differ in that they are used by the Department of City 
+Planning (DCP) for environmental, socio-economic, and demographic 
+purposes. They represent a geographical division rather than a City 
+government division. Due to the format of the \texttt{community\_district} 
+data, it is not possible to directly establish validity. However, the 
+dataset contains 72 unique entries, while only 59 valid Community 
+Districts exist.
+
+\paragraph{Date and Closed Date}
 \label{sec:negativeduration}
-\textbf{Issues with created\_date and closed\_date - Negative Durations:} For 
-the created\_date and closed\_date fields, one might 
-expect that these two fields are automatically populated by the 
-SR application software; associated with perhaps setting an SR status 
-to ``new'' or ``closed''. Unfortunately, that does not appear 
-to be the case as there are several anomalies with the various date 
-fields:
-
+For the \texttt{created\_date} and \texttt{closed\_date} fields, one 
+might expect that these fields are automatically populated by the SR 
+application software when setting an SR status to ``new'' or ``closed.'' 
+Unfortunately, this does not seem to be the case, as several anomalies 
+exist in the date fields:
 \begin{itemize}
-	\item SRs with a closed\_date that occurs before the created\_date.
-
- 	\item created\_date(s) and closed\_date(s) in the far, far distant past.
-
- 	\item created\_date(s) and closed\_date(s) that match to the second.
-
+    \item SRs with a \texttt{closed\_date} that occurs before the 
+    \texttt{created\_date}.
+    \item \texttt{created\_date(s)} and \texttt{closed\_date(s)} in 
+    the far distant past.
+    \item \texttt{created\_date(s)} and \texttt{closed\_date(s)} that 
+    match to the second.
     \item A large number of SRs closed and/or created exactly at midnight 
-	or exactly at noon, to the second.
+    or noon, to the second.
 \end{itemize}
 
-Citizens, NYC Government Officials, and NYC Agencies especially use 
-the closed and created dates as a measure of the ``duration'' of the SR 
-and hence Agency responsiveness of a service. Duration, while 
-not directly present in the dataset, is easily 
-computed by (closed\_date - created\_date).
-
-Closed-before-Created: There are 12,251 SRs that are closed before they 
-were created, thereby generating a nonsensical ``negative duration''. 
-While this is a small percentage overall (0.2\%) these errors can 
-have significant impact on response time analysis. Here is a sample:
+\subparagraph{Closed before Created: Negative Duration}
+Citizens, NYC Government Officials, and Agencies use the created and 
+closed dates to measure the ``duration'' of SRs, which reflects Agency 
+responsiveness. While duration is not directly present in the dataset, 
+it can be easily computed as the difference between
+\texttt{(closed\_date} and \texttt{created\_date}.
+There are 12,251 SRs where the \texttt{closed\_date} precedes the 
+\texttt{created\_date}, generating nonsensical ``negative durations.'' 
+Though this represents only 0.2\% of the dataset, these errors can 
+significantly impact response time analyses.
+Eight SRs with extremely large negative durations ($-$730 days), all 
+originating from the Department of Homeless Services (DHS), contain 
+an entry of ``1900-01-01'' as the \texttt{closed\_date}. This results 
+in negative durations exceeding $-$44,601 days (or 122 years). Such 
+anomalies, though rare, can significantly skew statistical results if 
+not addressed during data cleansing. As a result, these SR rows are 
+removed from the analysis.
 
@@
   \centering
-    \caption{Largest Negative Duration errors (days)}
+  \caption{Negative duraton of the largest magnitude (days) excluding
+    extreme negative values}
   \begin{tabular}{l l l r l}
     \toprule
-	        \multicolumn{5}{c}{\textbf{Largest errors (days) excluding 
-	        extreme negative values}} \\
-	        \midrule
-	        \textbf{created\_date} & \textbf{closed\_date} & \textbf{duration} 
+    {created\_date} & {closed\_date} & {duration} 
     & \textbf{agency} \\
     \midrule
-		        2023-01-27 14:40:00 & 2022-01-14 14:40:00 & -378.0 & DOT \\
-		        2023-01-18 10:06:00 & 2022-01-12 10:06:00 & -371.0 & DOT \\
-		        2023-01-27 14:36:00 & 2022-01-22 14:35:00 & -370.0 & DOT \\
-		        2023-01-11 11:10:00 & 2022-01-09 11:10:00 & -367.0 & DOT \\
-		        2023-12-18 03:13:00 & 2023-01-16 13:10:00 & -335.6 & DOT \\
+    2023-01-27 14:40:00 & 2022-01-14 14:40:00 & $-$378.0 & DOT \\
+    2023-01-18 10:06:00 & 2022-01-12 10:06:00 & $-$371.0 & DOT \\
+    2023-01-27 14:36:00 & 2022-01-22 14:35:00 & $-$370.0 & DOT \\
+    2023-01-11 11:10:00 & 2022-01-09 11:10:00 & $-$367.0 & DOT \\
+    2023-12-18 03:13:00 & 2023-01-16 13:10:00 & $-$335.6 & DOT \\
     \bottomrule
@@
 
-As noted, the largest errors are shown ``excluding extreme negative values''. We found 
-eight SRs with extremely large negative durations (\textless{} -730), all originating
-with DHS and all containing an entry of ``1900-01-01'' as the closed\_date. This 
-generates extremely large negative durations exceeding -44,601 
-days (-122 yrs). Such a large of an anomaly can significantly skew statistical 
-results, even though the number of occurrences is small. According, these 
-SR rows are removed from the analysis. Excluding these extreme negative
-values, an analysis shows the negative-duration issue is a DOT problem, 
-where 95\% of these types of errors occur.
-
-This violin chart shows the broad spread of negative-duration SRs, albeit with 
-the extremely large negative-duration SR removed. While there are few 
-outliers, the magnitude of the negative-durations is troubling and can 
-produce bizarre analytical results.
-
 \begin{figure}[tbp]
 	 \centering
- 	 \includegraphics[width = \textwidth]{negative_duration_SR_violin.pdf}
+ 	 \includegraphics[width = 0.7\textwidth]{negative_duration_SR_violin.pdf}
  \caption{Distribution of Negative Durations}
@@
 
-\label{sec:homlessassistance}
-\textbf{Case Study: Homeless Person Assistance.} Scenario: DHS wants to know ``How 
-quickly have 311 SRs for ``Homeless Person Assistance'' been resolved 
-over the past two years?'' This is a typical request made by 
-both the public and City government; analogous to ``How quickly 
-is my Agency responding to ``critical'' requests, and does 
-that performance vary by Borough, Zip Code, etc. To conduct such an 
-analysis, we can use the tools on the Open Data Portal. Here are the steps:
-
-\begin{itemize}
-    \item Select created\_date from 2022-2023 and filter by complaint\_type = 
-    ``Homeless Person Assistance'' (this yields 55,000 SRs)
-
-    \item Compute the ``duration'' (closed\_date – created\_date)
-
-    \item Take an average of the ``duration'' field  which yields: \textbf{Answer:  -4.8 days}  
-\end{itemize}
-
-Clearly that answer is nonsensical. How did such a simple task result 
-in an absurd answer? The answer lies in the computation of the ``duration'' 
-field. It turns out, there are eight DHS SRs that have a closed\_date 
-of ``1900-01-01''. Each of those SRs creates a negative duration of -44,602 
-days (-122 years). Those eight SRs are enough to drive the 
-average duration of 55,000 Homeless Assistance SRs to a negative 
-value; clearly incorrect. In this case, the median is a better measure 
-of central tendency. The median is  0.2 days (approx 5 hrs) which will
-surely please the DHS.
+Excluding the extreme negative values, 
+Figure~\ref{fig:negative-duration-violin} shows the broad spread of 
+negative-duration SRs. While there are few outliers, the magnitude 
+of the negative durations is troubling and can produce bizarre 
+analytical results. Table~\ref{tab:largest-errors} summarizes the 
+negative durations of the largest magnitude (in days). The data 
+suggests that the negative duration issue is predominantly a problem 
+within the Department of Transportation (DOT), where 95\% of these 
+errors occur.
+
+% For illustration, consider a case study of homeless person
+% assistance. Suppose that DHS wants to know ``How 
+% quickly have 311 SRs for ``Homeless Person Assistance'' been resolved 
+% over the past two years?'' This is a typical request made by 
+% both the public and City government; analogous to ``How quickly 
+% is my Agency responding to ``critical'' requests, and does 
+% that performance vary by Borough, Zip Code, etc. To conduct such an 
+% analysis, we can use the tools on the Open Data Portal. Here are the steps:
+
+% \begin{itemize}
+%     \item Select created\_date from 2022-2023 and filter by complaint\_type = 
+%     ``Homeless Person Assistance'' (this yields 55,000 SRs)
+
+%     \item Compute the ``duration'' (closed\_date – created\_date)
+
+%     \item Take an average of the ``duration'' field  which yields: \textbf{Answer:  -4.8 days}  
+% \end{itemize}
+
+% Clearly that answer is nonsensical. How did such a simple task result 
+% in an absurd answer? The answer lies in the computation of the ``duration'' 
+% field. It turns out, there are eight DHS SRs that have a closed\_date 
+% of ``1900-01-01''. Each of those SRs creates a negative duration of -44,602 
+% days (-122 years). Those eight SRs are enough to drive the 
+% average duration of 55,000 Homeless Assistance SRs to a negative 
+% value; clearly incorrect. In this case, the median is a better measure 
+% of central tendency. The median is  0.2 days (approx 5 hrs) which will
+% surely please the DHS.
 
@@
 
-\label{sec: zeroduration}		
-\textbf{Issues with created\_date \& closed\_date - Zero Durations:} A more 
-prevalent problem occurs when the closed\_date and created\_date 
-are exactly the same -- to the second. Accordingly, this creates 
-a \textbf{zero duration'}, again nonsensical. There are 191,141 
-such SRs representing 3.1\% of all non-blank data. 99\% of these zero duration 
-SRs occur in five Agencies:  (Dept of Mental Health \&Hygiene (DHMH), Dept 
-of Transportation (DOT), Dept of Business (DOB), Dept of Sanitation 
-(DSNY), and Dept of Environmental Protection (DEP). This Agency distribution
-is does not mirror the overall Agency distribution of SRs, likely indicating an 
-Agency-specific issue.
+\subparagraph{Resolution Action Update Date}
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width = 0.7\textwidth]{post_closed_violin_chart.pdf}
+  \caption{Post-closed resolution action update date(s) over 30 days.}
+  \label{fig:resolution-violin}
+\end{figure}
+
+\jy{Are these from 2-year or 10-year period? Need to clarify in
+  Section 2.}
+
+A similar issue occurs when resolution action update date is after the
+closing date. \jy{It might be possible, I think, if they allow updates
+  after a SR is closed.}
+When an SR is updated, the 311 software automatically populates the 
+\texttt{resolution\_action\_update\_date}. Some of these updates 
+occur long after the SR is closed. As shown in Figure~\ref{fig:resolution-violin}, 
+there are 7,460 SRs that were updated more than 30 days after the 
+\texttt{closed\_date}, but less than 730 days, to exclude infeasible 
+dates like 1900-01-01. The chart highlights the distribution of these 
+post-closure updates, with a notable concentration of updates within 
+the 30-to-90-day range. This pattern raises questions about whether 
+such delayed updates are standard operational practice or indicative 
+of a potential issue requiring further investigation. While the reason 
+for these delayed updates is unclear, they may impact the accuracy of 
+response time analyses if not properly accounted for.
+
+\subparagraph{Same Created Date \& Closed Date: Zero Durations}
+A more prevalent issue occurs when the \texttt{closed\_date} and 
+\texttt{created\_date} are exactly the same, down to the second. This 
+creates a \textbf{zero duration}, which is also nonsensical. There 
+are 191,141 such SRs, representing 3.1\% of all non-blank data. 
+Ninety-nine percent of these zero-duration SRs occur in five agencies:
+DHMH, DOT, DOB, DSNY, and DEP.. 
+This distribution does not mirror the overall SR distribution by 
+agency, suggesting an agency-specific issue.
 
@@
 
-\label{sec:midnightandnoon}
-\textbf{Issues with created\_date \& closed\_date(s) - Midnight \& Noon:} Another 
-problem discovered with the created\_date and 
-closed\_date fields - there are an unusually large number of SRs 
-created or closed at exactly midnight (00:00:00) and exactly noon (12:00:00), 
-to the second. Normal distribution of SR creation and closure 
-largely follows the work-day clock with most  SRs created during 
-day-light hours, and fewer SRs 	created at night and the early hours 
-of morning. However, there are significantly greater numbers 
-of SR closed exactly at midnight and noon, as well as a significant 
-number of SRs created exactly at noon. Observations include:
-
+\subparagraph{Created Date or Closed Date(s) at Midnight or Noon}
+Another issue found with the \texttt{created\_date} and 
+\texttt{closed\_date} fields is the unusually large number of SRs 
+created or closed at exactly midnight (00:00:00) or noon (12:00:00), 
+down to the second. Normally, SR creation and closure follow the 
+work-day schedule, with most SRs created during daylight hours and 
+fewer at night or in the early morning. However, there is a 
+significantly higher number of SRs closed exactly at midnight and noon, 
+as well as a large number created exactly at noon. Specific observations 
+include:
 \begin{itemize}
-	 \item There were 99,779 SRs created exactly at noon (12:00:00)
-
-	 \item There were 235,347 SRs closed exactly at midnight (00:00:00)
-
-	\item There were 105,505 SRs closed exactly at noon. 
+    \item 99,779 SRs were created exactly at noon (12:00:00).
+    \item 235,347 SRs were closed exactly at midnight (00:00:00).
+    \item 105,505 SRs were closed exactly at noon.
 \end{itemize}
-
-A unique way to visualize this anomaly is to examine the busiest day 
-during this 2-year period (Friday, 2023-09-29). Then aggregate 
-the created\_date(s) by minute (with seconds equal to zero), providing
-a minute-by-minute look at SR creation on that one day. Note the spike 
-at exactly noon (12:00), well beyond the 3\textsigma line.
 
@@
   \centering
-	\includegraphics[width=\textwidth]
-	{2-year-trend_SR_created_by_minute_of_busiest_day.pdf}
-	\caption{SRs Created Minute-by-Minute on Busiest Day}
-	\label{fig:busiestcreated}
+  \includegraphics[width=0.7\textwidth]{2-year-trend_SR_created_by_minute_of_busiest_day.pdf}
+  \includegraphics[width=0.7\textwidth]{2-year-trend-SR_closed_by_minute_of_busiest_day.pdf}
+  \caption{SRs Created and closed minute-by-minute on busiest day
+    during two year period \jy{give years}.}
+  \label{fig:busiest}
 \end{figure}
 
-Repeat the same for the \# of SRs with closed\_date(s) 
-on the busiest day (Friday, 2023-09-29). Again, aggregated by 
-the minute (with seconds equal to zero). Note the spikes 
-at 00:00:00 and 12:00:00 respectively.
-
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width=\textwidth]{2-year-trend-SR_closed_by_minute_of_busiest_day.pdf}
-	\caption{SRs Closed Minute-by-Minute on Busiest Day}
-	\label{fig:busiestclosed}
-\end{figure}
-
-These unusual patterns of created \& closed SRs at exactly the hours 
-of midnight and noon likely indicates the presence of a bulk create/close 
-software process that perhaps automatically ``closes'' 
-(or ``creates'') a large number of SRs with a generated time-stamp of exactly 
-midnight (00:00:00) or noon (12:00:00). Such behavior distorts
- the duration of these SRs by providing inaccurate dates. The 
- distribution by Agency for this anomaly indicates 
- that \textgreater90\% of these ``created\_at\_noon'' SRs come from 
- just two Agencies: DOB and DSNY.  Further investigation
- reveals that  a single Agency, DSNY, is responsible 
- for \textgreater99\% of the SRs ``closed-exactly-at-noon''.
-
-\label{sec: resolutionaction}
-\textbf{Issues with resolution\_action\_update\_date:} When an 
-SR is updated, the 311 software automatically 
-populates the resolution\_action\_update\_date. Some of 
-those  updates are happening long after the SR is 
-closed. There are a total of 7460 SRs that are updated \textgreater30 days 
-after the closed\_date (and \textless{}730 days to avoid infeasible 
-dates such as 1900-01-01). It is not known if this is normal behavior
- or an area that requires further investigation.
-
-\begin{figure}[tbp]
-	\centering
-	\includegraphics[width = \textwidth]{post_closed_violin_chart.pdf}
-	\caption{Post-Closed resolution\_action\_update\_date(s) \textgreater30 days}
-	\label{fig:resolution-violin}
-\end{figure}
+Figure~\ref{fig:busiest} visualizes this anomaly by examining the 
+busiest day during this 2-year period (Friday, 2023-09-29). The 
+\texttt{created\_date} values are aggregated by minute (with seconds 
+equal to zero), providing a minute-by-minute look at SR creation on 
+that day. Notably, there is a spike exactly at noon (12:00), well 
+beyond the 3\textsigma{} line. \jy{May not want to mention 3 sigma as
+  here we have highly skewed count data.}
+The same analysis was repeated for 
+\texttt{closed\_date} values on the busiest day, aggregated by minute 
+(with seconds equal to zero). Spikes were observed at both 00:00:00 
+and 12:00:00.
+
+These unusual patterns of SR creation and closure exactly at midnight 
+and noon suggest the presence of a bulk create/close software process 
+that automatically assigns time stamps of midnight (00:00:00) or noon 
+(12:00:00) to large batches of SRs. This behavior distorts the 
+calculated duration of these SRs by providing inaccurate dates. The 
+distribution by agency shows that over 90\% of the ``created-at-noon'' 
+SRs come from just two agencies: DOB and DSNY. Further investigation 
+reveals that a single agency, DSNY, is responsible for over 99\% of 
+the ``closed-at-noon'' SRs.
 
@@
 \label{sec:precision}
+
+\jy{This is caused by reading the number in as double
+  precision. Single precision would be good.}
+
 An question of precision vs. accuracy arises with the Latitude 
@@
 \label{sec:duplicates}
-During this analysis, several redundant fields were observed which
-should be examined further for possible consolidation.
-
-\label{sec:latlong}
-\textbf{latitude \& longitude and location:} The location 
-field is a 100\%pure concatenation of the latitude 
-and longitude fields with a comma and parenthesis 
-added. Thus the  location field seems questionable as the value, 
-and the data is arguably more difficult to 
-extract and use than the two individual fields. Example:
-
-\begin{itemize}
-	\item  latitude: 40.768456429488
-
-	\item  longitude: -73.9575661888774
-
-	\item  location: (40.768456429488, -73.95756618887745)
-\end{itemize}
-
-\label{sec:parkborough}
-\textbf{borough \&  park\_borough fields:} Borough and 
-park\_borough: These two fields are 100\% matches; fully redundant.
-
-\label{sec:boroughboundaries}
-\textbf{borough \& borough\_boundaries:} Borough and 
-borough\_boundary fields are a 98.3\% match. This 
-is an example of the the problem with two ``near-duplicate'' 
-fields; which field is correct when they disagree?
-
-\label{sec:taxicompanyborough}
-\textbf{borough \& taxi\_company\_borough:} Borough and 
-taxi\_company\_borough fields that despite their names, are 
-nearly completely different with only a 0.05\% match, suggesting 
-that these two fields are used very differently. The taxi\_company\_borough
- field is used exclusively by the TLC which governs all cars for hire. Any
- discussion of duplication with other borough-related fields 
- should include a discussion with TLC as to the usage of these similar,
- but non-matching, fields.
-
-\label{sec:zipcodes}
-\textbf{incident\_zip \& zip\_codes:} The zip\_code field is one 
-of the six computed fields and its validity 
- was previously explored and found to be 
- lacking; 56\% of the entries are invalid. The companion 
- field, incident\_zip, had an accuracy rate of 99.93\%. Our analysis
- indicates that the computed zip\_code field should be eliminated
- or masked from the public until such time as accuracy 
- issues can be addressed.
-
-\label{sec:police} 
-\textbf{police\_precinct \& police\_precincts:} Both the 
-police\_precinct and police\_precincts are among the computed 
-fields not cited in the Data Dictionary. However, their usage is easy to 
-discern; they presumably represent the NYPD precinct 
-corresponding to the incident address. These two 
-fields are near-duplicates with 99.9\% of the entries 
-matching. Since the underlying computational process 
-is unknown, we are unable to determine which 
-field is ``more correct'' field. However having two fields with 99.9\% 
-duplicity indicates that perhaps only one police 
-precinct field is necessary.
-
-\label{sec:agencyname}
-\textbf{agency \& agency\_name:} The agency and agency\_name 
-fields, while not duplicates per se, have a 1:1 correspondence. The 
-Agency field contains abbreviations for the City agencies, e.g. NYPD, 
-DOT, HPD, DSNY, etc. The agency\_name field contains the full name 
-of these organizations. It seems redundant to include both as 
-the abbreviations are commonly used and well understood.
-
-\label{sec:landmark}
-\textbf{landmark \& street\_name:} The landmark field is listed 
-in the Data Dictionary as ``...any noteworthy location, 
-including but not limited to, parks, hospitals, airports, sports 
-facilities, performance spaces, etc.'' Our analysis
-found that was not the case. To be sure, many of the entries in 
-the landmark field do contain landmark names, e.g.``Pennsylvania 
-Station'', ``LaGuardia Airport'', etc.,  but the vast majority 
-of the entries are street names, e.g. ``Fenton Avenue'', ``Steinway 
-Street'', ``Broadway'', etc.. These entries are very similar to the 
-street\_name field with a 62\% match, enough to be indicative 
-of some level of duplicative usage. Even the non-matches 
-(excluding blanks) appear to be matches except for 
-minor spelling and nomenclature changes, 
-e.g. ``NINTH AVE'' \& ``9 AVE''. It's worth investigating what is going on
-with the usage of the landmark field.
-
-\label{sec:cross-street}
+
+During this analysis, several redundant fields were observed and should 
+be examined for possible consolidation.
+
+\textbf{Latitude/Longitude and Location:} The \texttt{location} field 
+is a pure concatenation of the latitude and longitude fields with a 
+comma and parentheses added. This makes the \texttt{location} field 
+arguably harder to use than the individual fields. For example:  
+\texttt{latitude}: 40.768456429488, \texttt{longitude}: $-$73.9575661888774, 
+and \texttt{location}: (40.768456429488, $-$73.95756618887745).
+
+\textbf{borough \& park\_borough:} These two fields are fully redundant, 
+as they are 100\% matches.
+
+\textbf{borough \& borough\_boundaries:} These fields are 98.3\% matches. 
+This ``near-duplicate'' relationship creates a challenge in determining 
+which field is correct when discrepancies arise.
+
+\textbf{borough \& taxi\_company\_borough:} Despite their names, these 
+fields are almost entirely different, with only a 0.05\% match. The 
+\texttt{taxi\_company\_borough} field is used exclusively by the TLC, 
+indicating a need for consultation with TLC to understand the differing 
+uses.
+
+\textbf{incident\_zip \& zip\_codes:} The \texttt{zip\_codes} field, 
+one of the computed fields, was found to have 56\% invalid entries, 
+while the \texttt{incident\_zip} field has a 99.93\% accuracy rate. 
+We recommend eliminating or masking the \texttt{zip\_codes} field 
+until its accuracy can be improved.
+
+\textbf{police\_precinct \& police\_precincts:} These two computed 
+fields are 99.9\% matches. Since the computational process is unknown, 
+it is unclear which field is more accurate. However, having two fields 
+with near-duplicity suggests that only one is necessary.
+
+\textbf{agency \& agency\_name:} The \texttt{agency} field contains 
+abbreviations of City agencies (e.g., NYPD, DOT), while 
+\texttt{agency\_name} contains the full names. Including both seems 
+redundant, especially given the space taken up by full names in the 
+CSV file.
+
+\textbf{landmark \& street\_name:} While the \texttt{landmark} field is 
+intended for noteworthy locations (e.g., parks, airports), most entries 
+are actually street names, with a 62\% match to \texttt{street\_name}. 
+Even non-matches (excluding blanks) appear to be similar, differing 
+only in spelling or formatting (e.g., ``NINTH AVE'' vs. ``9 AVE''). This 
+indicates a possible misuse of the \texttt{landmark} field.
+
 \textbf{cross\_street\_1 \& intersection\_street\_1 and cross\_street\_2 
-\& intersection\_street\_2: } There are two sets of street pairs provided 
-in addition to the incident\_address field. These pairs are:
-
-\begin{itemize}
-	\item cross\_street\_1 \& intersection\_street\_1
-
-	\item cross\_street\_2 \& intersection\_street\_2
-\end{itemize}
-
-These two street pairs are used to assist in identifying the address, 
-as is common in New York City. These two street pairs
-have 88\% duplicates where both cross\_street\_1 matches 
-intersection\_street\_1 and cross\_street\_2 
-matches intersection\_street\_2. We believe only one of these
-street pairs are necessary, but which one? This illustrates the 
-problem of two similar data fields that largely, but not exactly, match. In
-this case there are 12\% that do not match; which field to trust?
-
-\textit{(Note: Address standardization was applied to  street 
-addresses using the R \textit{campfin} package. This standardization 
-prevents generating a non-match between locations such as 
-``240 E 69\textsuperscript{th} ST'' and ``240 E 69\textsuperscript{th} Street'' 
-which are clearly the same address save for formatting 
-and slight spelling differences.)}
-
-\textbf{Near matches:} In addition to the matching values, we 
-encountered a number of \textit{near-matches}; situations where the 
-cross\_street and intersection\_street are nearly, but not exactly, identical. 
-For example, HARMON DR and HARMON RD. To identify these near-matches 
-we measured the \href{https://en.wikipedia.org/wiki/Hamming_distance}
-\textit{Hamming Distance} between the two fields. The Hamming Distance is a 
-measure of how many letters must be changed for the two 
-fields to match. For this analysis, we chose a Hamming Distance of two (2).
- Here are some example results for near-matches for both 
- cross\_street\_1 and intersection\_street\_1. 
- (Cross\_street\_2 and intersection\_street\_2 show similar results.)
+\& intersection\_street\_2:} These two sets of street pairs are used to 
+help identify the incident address. After applying address 
+standardization using the R \textit{campfin} package, we found 88\% of 
+these pairs to be duplicates, while 12\% do not match. This discrepancy 
+raises the question of which field should be trusted. It is possible 
+that the two pairs originated from different agencies that used 
+different variable names to represent the same information, but no 
+documentation exists to clarify this relationship.
+
+\paragraph{Near Match}
+
+\jy{This investigation could be relegated to on online supplement
+  given the low percentage of near-match. It's taking too much spaces
+  here.
+}
 
@@
     \centering
-     \caption{Near-matches for cross\_street\_1, intersection\_street\_1}
+     \caption{Examples of near-matches for cross\_street\_1 and intersection\_street\_1.}
      \label{tab:x1nearmatches}
@@
 
-\subsection{Shrinking file size by removing duplicates}
+In addition to the matching values, we encountered a number of 
+\textit{near-matches}, situations where the \texttt{cross\_street} 
+and \texttt{intersection\_street} are nearly, but not exactly, identical. 
+An example of such a near-match is when one field reads ``HARMON DR'' 
+and the other reads ``HARMON RD.'' To identify these near-matches, 
+we applied the \href{https://en.wikipedia.org/wiki/Hamming_distance}
+{\textit{Hamming Distance}}, which measures the number of character 
+changes required to make the two fields identical. For this analysis, 
+we set the threshold for a near-match to a Hamming Distance of two (2).
+
+Table~\ref{tab:x1nearmatches} provides examples of near-matches for 
+\texttt{cross\_street\_1} and \texttt{intersection\_street\_1}. These 
+entries show slight variations in the street name or number, such as 
+``WEST 168 ST'' and ``WEST 167 ST,'' both processed by DOT, where only 
+one character differs. Although these discrepancies may seem small, 
+they can have implications for data accuracy, particularly when 
+geocoding or matching locations for analysis purposes.
+
+Table~\ref{tab:summary1} presents a summary of the match and 
+near-match results for \texttt{cross\_street\_1} and 
+\texttt{intersection\_street\_1}. As the table shows, 88.15\% of the 
+entries are exact matches, while 25.4\% are blank but matching. 
+However, 11.85\% of the entries are non-matching, and a small fraction 
+(0.002\%) fall into the near-match category. Although this percentage 
+seems negligible, even a small number of near-matches can lead to 
+confusion or errors when analyzing location-based data, especially 
+if critical decisions rely on accurate street data. Address 
+standardization or correction processes may be necessary to resolve 
+these inconsistencies and improve the reliability of the data.
+
+\subsection{Storage}
 \label{sec:filesize}
-Shrinking file size by removing duplicates: By removing duplicate 
-and ``near-duplicate'' fields, it is possible to 
-shrink the file size by 12.3\% which for this large dataset equates to a 
-reduction of 395 Mb. A smaller file size means faster downloads, 
-less storage impact, as well as simplifying data analysis efforts.
-
+
+By removing duplicate and ``near-duplicate'' fields, it is possible to 
+shrink the file size by 12.3\%, which for this large dataset equates to 
+a reduction of 395 Mb. A smaller file size means faster downloads, 
+less storage impact, and simpler data analysis efforts.
 Below is a list of duplicate and near-duplicate fields. There are 
-some challenges with these proposed deletions, mostly with 
-the ``near duplicate'' fields.  For example, while the park\_borough is 
-a complete duplicate of the borough field, the 
-borough\_boundary field is only 98.3\% duplicate. Admittedly, 
-loosing 1.7\% of the data in those two fields is quite small, however
-for this large dataset that small percentage still amounts 
-to 110,715 non-matching occurrences. The authors  propose 
-deleting the following fields from the data set.
+challenges with these proposed deletions, especially with the 
+``near-duplicate'' fields. For example, while the \texttt{park\_borough} 
+is a complete duplicate of the \texttt{borough} field, the 
+\texttt{borough\_boundary} field is only 98.3\% similar. Although losing 
+1.7\% of the data is small, it still amounts to 110,715 non-matching 
+occurrences. The authors propose deleting the following fields:
 
 \begin{itemize}
-	\item agency\_name: Each row of data contains the agency field, 
-	which is an abbreviation of the Agency's name. The agency\_name 
-	abbreviations are clear, simple, and well understood.
-
-	\item park\_borough:  This field is a 100\% match with the 
-	borough field, thus no data loss.
-
-	\item location:  The location field is a straight concatenation of 
-	the latitude and longitude fields, with a comma and parenthesis 
-	added. The data is a 100\% duplication of the latitude 
-	and longitude fields, thus no data loss.
-
-	\item police\_precinct: This field has a 99.95\% match with the 
-	police\_precincts fields. We are unable to determine which field 
-	is more correct, but feel that removal of one of the precinct fields 
-	is prudent and would result in minimal data loss.
-
-	\item borough\_boundaries (computed field): This field is one of 
-	the six computed fields. It has a 98.3\% match with the borough 
-	field.
-
-	\item cross\_street\_1 \& 2 and intersection\_street\_1 \& 2: These 
-	two pairs of fields each have an 88\% match. We would recommend 
-	deleting the two intersection\_street (1 \& 2) fields while acknowledging 
-	some loss of data in doing so.
-
-	\item zip\_codes(computed field):  This field is one of the six 
-	computed fields. However, this field has error rate of 58\% and 
-	as shown in the \ref{sec:case-study-zip-codes} can lead to 
-	dangerous mistakes when used for analytical purposes. We 
-	recommend deleting this field.
+    \item \textbf{agency\_name}: Each row contains the \texttt{agency} field, 
+    which uses abbreviations that are clear and well understood.
+
+    \item \textbf{park\_borough}: A 100\% match with the \texttt{borough} field, 
+    leading to no data loss.
+
+    \item \textbf{location}: A straight concatenation of latitude and longitude 
+    fields, offering no additional value beyond duplication.
+
+    \item \textbf{police\_precinct}: This field has a 99.95\% match with 
+    \texttt{police\_precincts}. Removing one would result in minimal data loss.
+
+    \item \textbf{borough\_boundaries (computed field)}: A 98.3\% match 
+    with the \texttt{borough} field.
+
+    \item \textbf{cross\_street\_1 \& 2 and intersection\_street\_1 \& 2}: 
+    An 88\% match between these pairs. We recommend deleting the 
+    \texttt{intersection\_street\_1 \& 2} fields while acknowledging some 
+    data loss.
+
+    \item \textbf{zip\_codes (computed field)}: This field has a 58\% error rate 
+    and, as discussed in \ref{sec:case-study-zip-codes}, can lead to significant 
+    analytical errors. We recommend deleting this field.
 \end{itemize}
 
-Additionally, it would be worth investigating if certain data fields are 
-in-fact useful, as the population of these fields is very scarce. For 
-example, the taxi\_company\_borough field is 99.94\% 
-blank. A similar situation exists for the road\_ramp field 
-(99.8\% blank), the vehicle\_type field (99.7\%blank), and 
-due\_date which is 99.6\% blank. Several additional several fields
-are mostly blank: bridge\_highway\_direction, bridge\_highway\_name, 
-bridge\_highway\_segment, and taxi\_pick\_up\_location; all 
-have \textgreater99\% blanks. Most of these latter fields 
-appear to used exclusively by the TLC.
-
-\subsection{Data Dictionary Observations} 
+Additionally, certain fields may not be useful due to their sparse population. 
+For example, the \texttt{taxi\_company\_borough} field is 99.94\% blank, 
+\texttt{road\_ramp} is 99.8\% blank, \texttt{vehicle\_type} is 99.7\% blank, 
+and \texttt{due\_date} is 99.6\% blank. Several other fields, such as 
+\texttt{bridge\_highway\_direction}, \texttt{bridge\_highway\_name}, 
+\texttt{bridge\_highway\_segment}, and \texttt{taxi\_pick\_up\_location}, 
+are over 99\% blank, with most used exclusively by the TLC.
+
+Further reduction is possible by coding categorical 
+variables, such as \texttt{complaint\_type}, through methods like dropdown 
+menus or encoded categories. By using a dropdown menu for 
+\texttt{complaint\_type}, data entry can be standardized, reducing 
+inconsistencies caused by manual input errors. Additionally, categorical 
+variables with a limited set of possible values can be stored more 
+efficiently by assigning unique codes or numerical representations for 
+each category. For instance, instead of repeatedly storing the full text 
+for every complaint type (e.g., ``Noise - Residential,'' ``Noise - 
+Commercial''), a single code could represent each type. This not only 
+reduces file size but also improves data processing speed and efficiency 
+during analysis. Moreover, standardizing categorical inputs prevents 
+the inclusion of misspellings or unrecognized entries, ensuring cleaner 
+and more reliable data. This approach is particularly useful in large 
+datasets, where frequently repeated categorical variables can 
+significantly increase storage requirements. Encoding such fields 
+as dropdown options or coded variables can lead to faster query times 
+and more consistent data handling.
+
+\subsection{Data Dictionary} 
 \label{sec:datadictionary}
+
 The Data Dictionary could benefit from an update. We found several 
-discrepancies between the Data Dictionary and the 
-actual data which can potentially lead analytic efforts astray.
+discrepancies between the Data Dictionary and the actual data, which 
+could potentially mislead analytic efforts.
 
 \begin{itemize}
-	\item While the Open Data portal lists the data types of the 
-	various fields, the Data Dictionary does not. Nor does it provide 
-	information on the specifics of various fields, such 
-	as constraints or the domain of legal values. For example, the 
-	incident\_zip field is specified as ``text''. That is probably not the 
-	best definition; perhaps	``categorical'' with a note indicating that 
-	the data can contain only numeric characters and is 
-	not subjected to arithmetic operations.
-
-	\item Many of the domain of legal values are incomplete or 
-	inaccurate. For example, the Data Dictionary indicates that 
-	the status field has values of \textit{assigned, canceled, closed}, 
-	or \textit{pending}. However, we found additional status 
-	values of: \textit{in progress, started}, and \textit{unspecified}. 
-	Additionally, we found no SRs with a status of \textit{canceled}. 
-	Similarly, the address\_type field is indicated to have 
-	values \textit{address, blockface, intersection, latlong}, and 
-	\textit{placename}. We discovered the additional values of 
-	\textit{bbl} and \textit{unrecognized}. Additionally, no values of 
-	\textit{latlong} were observed. Other fields (facility\_type, 
-	vehicle\_type, taxi\_pick\_up\_location, road\_ramp, city) 
-	suffer similar inaccuracies in their specified domain of values as
-	shown in the Data Dictionary.
-
-	\item As noted previously, there are six \textit{@computed} fields that
-	are present in the data, but not included in the Data Dictionary.
+    \item While the Open Data portal lists the data types of various 
+    fields, the Data Dictionary does not. Nor does it provide details 
+    on specific fields, such as constraints or the domain of legal 
+    values. For example, the \texttt{incident\_zip} field is specified 
+    as ``text.'' This may not be the most suitable definition; 
+    ``categorical,'' with a note that the data contains only numeric 
+    characters and is not subjected to arithmetic operations, might 
+    be a better description.
+
+    \item Many of the domains of legal values are incomplete or 
+    inaccurate. For example, the Data Dictionary indicates that the 
+    \texttt{status} field has values of \textit{assigned, canceled, 
+    closed}, or \textit{pending}, but we also found values such as 
+    \textit{in progress, started}, and \textit{unspecified}. 
+    Additionally, no SRs were found with a status of 
+    \textit{canceled}. The \texttt{address\_type} field similarly 
+    lists values of \textit{address, blockface, intersection, latlong}, 
+    and \textit{placename}, but we found additional values such as 
+    \textit{bbl} and \textit{unrecognized}, and no \textit{latlong} 
+    values were observed. Other fields, like \texttt{facility\_type}, 
+    \texttt{vehicle\_type}, \texttt{taxi\_pick\_up\_location}, 
+    \texttt{road\_ramp}, and \texttt{city}, exhibit similar inaccuracies 
+    in their specified domain values.
+
+  \item As mentioned earlier, six \texttt{@computed} fields that are
+    present in the data but not included in the Data Dictionary need
+    to be documented.
 \end{itemize}

### 2024-09-15 manuscript/manuscript.tex
smoothed section 4

@@
 {\textit{Hamming Distance}}, which measures the number of character
+\jy{needs a quick definition and reference}
 changes required to make the two fields identical. For this analysis, 
@@
 
-    \item \textbf{zip\_codes (computed field)}: This field has a 58\% error rate 
-    and, as discussed in \ref{sec:case-study-zip-codes}, can lead to significant 
+  \item \textbf{zip\_codes (computed field)}:
+    This field has a 58\% error rate and can lead to significant 
     analytical errors. We recommend deleting this field.
@@
 \label{sec:recommendations}
-We propose the following recommendations by the Open Data Team to address
-the data cleanliness issues addressed in this paper.
+We propose the following recommendations for the Open Data Team to 
+address the data cleanliness issues discussed in this paper.
 
 \begin{itemize}
-	\item First, the Open Data Team will need 
-	to undertake a dedicated effort to investigate these 
-	data anomalies, confirm the veracity of the issue, 
-	and engage with the appropriate Agency Open Data 
-	Representatives to seek resolution. That effort will not 
-	be easy or fast, and in some cases software changes may  
-	be required. Coordination and mutual approval of any 
-	changes will be key, requiring solid support 
-	from senior leadership.
-
-	\item Secondly, standards and protocols need to be 
-	identified, agreed to, and commitment obtained. For example, 
-	the domain of legal values, data accuracy, address normalization, 
-	data usage, data definitions, etc..
-
-	\item Many of the data integrity issues could potentially be 
-	resolved through enforcement by software. For example, the 
-	software should be able to easily prevent invalid values, such as zip 
-	codes, and similarly for other data issues. One key area are is the 
-	created\_date and closed\_date which are frequently cited 
-	when determining responsiveness for various Agency services; software
-	could conceivably prevent illogical situations such as an SR being
-	``closed'' before it is ``opened''.
-
-	\item The large spikes in SR closure and creation occurring at 
-	midnight and noon almost certainly point to an automated 
-	process from an external Agency integration. Such processes distort 
-	the accurate capture of an SR duration and should be eliminated in
-	favor of more accurate dates.
-
-	\item As noted, there are six ``computed'' fields in the data export 
-	that are not identified in the Data Dictionary. Several of 
-	these computed fields suffer from significant data inaccuracies. We  
-	recommend updating the Data Dictionary at a minimum and 
-	possibly masking those fields from public usage until such time as 
-	their accuracy can be ascertained.
-
-	\item Removal of duplicate fields should be examined. In some cases, such 
-	as the borough and park\_borough fields, the duplication is 100\%. In 
-	other cases, the duplication is in the 80-90\% range. These near-duplicate 
-	fields and 	their utility should be examined to determine if inclusion in the 
-	311 SR dataset is warranted. Our analysis indicates that as much as 12\% 
-	size reductions are achievable with a first pass, possibly more upon deeper
-	analysis.
-
-	\item Fields with extremely low data presence, such as taxi\_company\_borough 
-	which is populated only 0.06\% of the time, should be evaluated 
-	to determine if their inclusion is warranted.
-
-	\item The Data Dictionary is dated March 2023. The issues addressed 
-	in Section~\ref{sec:datadictionary} should be the bases for an update of
-	the Data Dictionary.
-
-	\item Excessive precision in the latitude, longitude, and location 
-	fields is nonsensical.  We recommend no more than 
-	four decimal places (0.000X), representing a distance of approximately 
-	36.4 feet. Further decimal points strain credibility.
-
-	\item The 11 address-related address fields (incident\_address, 
-	street\_name, city, intersection\_street\_1 \& 2, cross\_street\_1 \& 2,  
-	landmark, road\_ramp, bridge\_highway\_segment, and 
-	taxi\_pick\_up\_location) should be standardized to USPS format. 
-	Software packages can easily perform this standardization.
-
-	\item There are multiple means by which missing data is 
-	represented. These include ``UNKNOWN'' , ``N/A'', ``NA'', ``na'', and 
-	a blank field, e.g. `` ''. Protocols should be established regarding 
-	how to represent missing data.
+\item  Undertake a dedicated effort 
+    to investigate these data anomalies, confirm the veracity of the 
+    issues, and collaborate with the appropriate Agency Open Data 
+    Representatives to find resolutions. This effort may be complex 
+    and require software changes. Coordination and mutual approval 
+    of changes will be key, requiring strong support from senior 
+    leadership.
+
+    \item Standards and protocols should be established, agreed upon, 
+    and enforced. This includes defining the domain of legal values, 
+    ensuring data accuracy, implementing address normalization, 
+    clarifying data usage, and providing detailed data definitions.
+
+    \item Resolve many potential data integrity issues
+    through enforcement by software. For instance, software should 
+    prevent invalid entries like erroneous zip codes. Similarly, fields 
+    like \texttt{created\_date} and \texttt{closed\_date}, which 
+    measure responsiveness, should have logic checks to prevent 
+    illogical sequences, such as an SR being ``closed'' before 
+    being ``opened.''
+
+    \item The large spikes in SR creation and closure at midnight and 
+    noon likely point to automated processes from external Agency 
+    integrations. These distortions should be eliminated in favor of 
+    capturing accurate SR durations with precise timestamps.
+
+    \item As discussed, there are six ``computed'' fields that are not 
+    listed in the Data Dictionary. Some of these fields contain 
+    significant inaccuracies. We recommend updating the Data 
+    Dictionary and, if necessary, masking these fields from public use 
+    until their accuracy is improved.
+
+    \item Duplicate fields should be reviewed for removal. In some 
+    cases, such as \texttt{borough} and \texttt{park\_borough}, the 
+    duplication is 100\%. In others, the match is 80-90\%. 
+    These near-duplicates should be assessed to determine their 
+    relevance in the 311 SR dataset. Initial analysis suggests that 
+    removing these fields could reduce file size by up to 12\%, with 
+    further reductions possible upon deeper review.
+
+    \item Fields with very low data presence, such as 
+    \texttt{taxi\_company\_borough}, populated only 0.06\% of the time, 
+    should be evaluated to determine if they are necessary.
+
+    \item The Data Dictionary, last updated in March 2023, should be 
+    revised based on the issues discussed in Section~\ref{sec:datadictionary}.
+
+    \item The precision of latitude, longitude, and location fields 
+    should be reduced. We recommend limiting precision to four 
+    decimal places (0.000X), which represents approximately 36.4 feet. 
+    More decimal points strain credibility.
+
+    \item The 11 address-related fields (\texttt{incident\_address}, 
+    \texttt{street\_name}, \texttt{city}, \texttt{intersection\_street\_1 \& 2}, 
+    \texttt{cross\_street\_1 \& 2}, \texttt{landmark}, \texttt{road\_ramp}, 
+    \texttt{bridge\_highway\_segment}, and \texttt{taxi\_pick\_up\_location}) 
+    should be standardized to USPS format. Software can easily 
+    perform this standardization.
+
+    \item There are multiple ways missing data is represented, such as 
+    ``UNKNOWN'', ``N/A'', ``NA'', ``na'', and blanks (`` ''). Protocols 
+    should be established to standardize the representation of missing data.
 \end{itemize}

### 2024-09-15 manuscript/manuscript.tex
enhanced discussion

@@
 
-Our study highlights the vital importance of robust data curation in
-maintaining the reliability and utility of open datasets, with a close
-examination of the NYC 311 data. As these datasets are applied across
-diverse fields such as public services and academic research, issues
-like inconsistencies, missing values, and formatting errors can
-significantly undermine the quality of insights. Trustworthy machine
-learning systems, which rely heavily on high-quality data, are
-particularly vulnerable to errors when data is not properly
-curated. For example, biases or inconsistencies in input data can
-compromise both the accuracy and fairness of predictive models
-\citep{rahm2000data}.
+Our study highlights the critical role of robust data curation in
+ensuring the reliability and utility of open datasets, as demonstrated
+through the examination of the NYC 311 data. As these datasets are
+applied across diverse fields such as public services and academic
+research, inconsistencies, missing values, and formatting errors can
+significantly undermine the insights derived from them. Trustworthy
+machine learning systems, which rely heavily on high-quality data, are
+especially vulnerable to such issues. Errors in the underlying data,
+such as biases or inconsistencies, can compromise both the accuracy
+and fairness of predictive models, impacting real-world decisions in
+urban planning and policy \citep{rahm2000data, geiger2020garbage}.
 
 This research further emphasizes the necessity of harmonizing data
-from multiple city agencies, particularly in resolving discrepancies
-in field names, formats, and definitions. Trustworthy machine learning
-depends on not only accurate and clean data but also consistent data
-across sources. This challenge becomes more acute in datasets spanning
-long time periods, where changes in agency structures or reporting
-standards can introduce subtle biases. As \citet{rahm2000data} noted,
-ensuring long-term data consistency is crucial for longitudinal
-studies and predictive modeling, as even minor shifts in data can
-result in significant deviations in outcomes.
+across city agencies, particularly in resolving discrepancies in field
+names, formats, and definitions. Consistency in data is essential for
+trustworthy machine learning, where inconsistencies across sources can
+lead to distorted outcomes. This challenge is even more pronounced in
+datasets that span long time periods, as changes in agency structures
+or reporting standards can introduce subtle biases. Long-term data
+consistency is crucial for longitudinal studies and predictive
+modeling, as even minor data shifts can lead to significant deviations
+in model outcomes, as noted by \citet{rahm2000data} and
+\citet{borgman2012conundrum}.
 
 Finally, the intersection of data curation and machine learning for
-public policy applications opens new opportunities for enhancing
-governance. By improving data quality, machine learning models can
-more effectively predict trends, allocate resources, and identify
-issues such as service delays. The COVID-19 pandemic demonstrated the
-critical role of real-time, trustworthy data in responding to public
-needs. Without proper data curation, the accuracy and timeliness of
-machine learning insights would have been compromised, affecting
-crucial decision-making processes. Future efforts could focus on
-automating data curation within real-time data pipelines, ensuring
-that the data feeding into machine learning models remains both clean
-and reliable \citep{chu2016data}.
+public policy applications opens new avenues for improving governance.
+High-quality data enables machine learning models to better predict
+trends, allocate resources, and address service delivery issues, such
+as delays in responding to 311 complaints. The COVID-19 pandemic
+underscored the importance of real-time, trustworthy data in crisis
+management, where the accuracy and timeliness of data-driven insights
+were critical for public health responses. Without proper data
+curation, machine learning models used during the pandemic would have
+been compromised, affecting decisions on resource allocation and
+service provision \citep{wu2020subnational, kumar2020machine}.
+Future efforts could focus on automating data curation processes,
+particularly in real-time data pipelines, to ensure that the data used
+in machine learning models remains accurate, clean, and reliable
+\citep{chu2016data, zhang2019automating}.
 

### 2024-09-15 manuscript/manuscript.tex
corrected references in discussion

@@
 been compromised, affecting decisions on resource allocation and
-service provision \citep{wu2020subnational, kumar2020machine}.
+service provision \citep{worby2020face, khemasuwan2021applications}.
 Future efforts could focus on automating data curation processes,
@@
 in machine learning models remains accurate, clean, and reliable
-\citep{chu2016data, zhang2019automating}.
+\citep{chu2016data, hurbean2021open}.
 

### 2024-09-15 manuscript/manuscript.tex
connected to ML in intro

@@
 These initiatives represent significant strides in democratizing data, 
-in breaking barriers that once kept valuable information 
+breaking down barriers that once kept valuable information 
 on government performance in silos. Their collective impact 
-is profound, extending beyond mere data sharing to 
-fostering a culture of openness that benefits individuals, 
-communities, governments, and economies worldwide 
+extends beyond mere data sharing to fostering a culture of openness 
+that benefits individuals, communities, governments, and economies worldwide 
 \citep{barns2016mine, wang2016adoption}.
@@
 
-The civil and research applications of open data critically depend on
-its quality, making data curation fundamental in the open data
-ecosystem. Among the earliest discussions,
-\citet{witt2009constructing} focused on developing data curation
-profiles tailored to specific contexts, setting a precedent for
-targeted data management strategies. Addressing broader challenges in
-data sharing and management, \citet{borgman2012conundrum} highlighted
-the complexities of research data distribution, emphasizing the need
-for robust strategies. This is complemented by \citet{hart2016ten},
-who outlined essential principles for effective data management,
-particularly emphasizing meticulous curation practices. In
-collaborative data management, \citet{beheshti2019datasynapse}
-underscored the significance of cooperative environments for managing
-and sharing social data effectively. This aspect gains further
-relevance in \citet{mclure2014data}, which delved into the specific
-practices and needs within data curation communities. The practical
-implications of data curation are vividly illustrated in public health
-and global challenges. \citet{cantor2018facets} demonstrated the
-utility of curated open data in evaluating community health
-determinants. The COVID-19 pandemic served as a real-world example,
-with \citet{shankar2021data} observing the critical role of collective
-data curation efforts in managing and responding to the
-crisis. Collectively, these studies highlight the multifaceted nature
-of data curation and emphasize its indispensable role in enhancing the
-applicability and value of open data across various domains.
+Data curation, the process of organizing, maintaining, and ensuring
+the quality of datasets, plays a crucial role in maximizing the
+utility of open data. Proper curation ensures that datasets remain
+consistent, accurate, and useful for diverse applications. For example, 
+curating data ensures it can be leveraged effectively for machine learning 
+systems, which require high-quality data to produce reliable insights
+\citep{polyzotis2019data, jain2020overview}. Poor 
+curation may result in issues such as missing data, formatting errors, or 
+inconsistent values that can lead to biased or inaccurate outcomes
+\citep{geiger2020garbage}.
+This is particularly critical in domains where machine learning is applied 
+to sensitive tasks, such as public health or policy
+\citep{rahm2000data}.
+
+Research into data curation has explored these challenges in depth. 
+Among the earliest discussions, \citet{witt2009constructing} focused 
+on developing data curation profiles tailored to specific contexts, 
+setting a precedent for targeted data management strategies. 
+Addressing broader challenges in data sharing and management, 
+\citet{borgman2012conundrum} highlighted the complexities of 
+research data distribution, emphasizing the need for robust 
+strategies. This is complemented by \citet{hart2016ten}, who outlined 
+essential principles for effective data management, particularly 
+emphasizing meticulous curation practices. The utility of curated 
+open data is vividly illustrated in public health and global challenges, 
+where \citet{cantor2018facets} demonstrated the utility of curated 
+data in evaluating community health determinants, and 
+\citet{shankar2021data} observed its critical role during the 
+COVID-19 pandemic in managing collective responses.
 
@@
 the specifics of data curation challenges using the NYC 311 Service
-Request (SR) Data as a case study. This renowned and frequently viewed 
-dataset serves as a prime example for examining key issues in data curation, 
-including data validity, consistency, and curation efficiency. 
+Request (SR) Data as a case study. This dataset serves as a prime 
+example for examining key issues in data curation, including data 
+validity, consistency, and curation efficiency. 
 We illustrate these points with live examples drawn from our 

### 2024-09-15 manuscript/manuscript.tex
renamed ref.bib to refs.bib
rewrote recommendation section with principles and actionable steps

@@
 
-\section{NYC 311 Service Request Data} 
+\section{NYC 311 SR Data} 
 \label{sec:data}
@@
 
-\section{Data Cleansing Issues} 
+\section{NYC 311 SR Data Cleansing Issues} 
 \label{sec:issues}
@@
 
-\section{Recommendations} 
+\section{Government Open Data Curation Recommendations}
 \label{sec:recommendations}
-We propose the following recommendations for the Open Data Team to 
-address the data cleanliness issues discussed in this paper.
-
-\begin{itemize}
-\item  Undertake a dedicated effort 
-    to investigate these data anomalies, confirm the veracity of the 
-    issues, and collaborate with the appropriate Agency Open Data 
-    Representatives to find resolutions. This effort may be complex 
-    and require software changes. Coordination and mutual approval 
-    of changes will be key, requiring strong support from senior 
-    leadership.
-
-    \item Standards and protocols should be established, agreed upon, 
-    and enforced. This includes defining the domain of legal values, 
-    ensuring data accuracy, implementing address normalization, 
-    clarifying data usage, and providing detailed data definitions.
-
-    \item Resolve many potential data integrity issues
-    through enforcement by software. For instance, software should 
-    prevent invalid entries like erroneous zip codes. Similarly, fields 
-    like \texttt{created\_date} and \texttt{closed\_date}, which 
-    measure responsiveness, should have logic checks to prevent 
-    illogical sequences, such as an SR being ``closed'' before 
-    being ``opened.''
-
-    \item The large spikes in SR creation and closure at midnight and 
-    noon likely point to automated processes from external Agency 
-    integrations. These distortions should be eliminated in favor of 
-    capturing accurate SR durations with precise timestamps.
-
-    \item As discussed, there are six ``computed'' fields that are not 
-    listed in the Data Dictionary. Some of these fields contain 
-    significant inaccuracies. We recommend updating the Data 
-    Dictionary and, if necessary, masking these fields from public use 
-    until their accuracy is improved.
-
-    \item Duplicate fields should be reviewed for removal. In some 
-    cases, such as \texttt{borough} and \texttt{park\_borough}, the 
-    duplication is 100\%. In others, the match is 80-90\%. 
-    These near-duplicates should be assessed to determine their 
-    relevance in the 311 SR dataset. Initial analysis suggests that 
-    removing these fields could reduce file size by up to 12\%, with 
-    further reductions possible upon deeper review.
-
-    \item Fields with very low data presence, such as 
-    \texttt{taxi\_company\_borough}, populated only 0.06\% of the time, 
-    should be evaluated to determine if they are necessary.
-
-    \item The Data Dictionary, last updated in March 2023, should be 
-    revised based on the issues discussed in Section~\ref{sec:datadictionary}.
-
-    \item The precision of latitude, longitude, and location fields 
-    should be reduced. We recommend limiting precision to four 
-    decimal places (0.000X), which represents approximately 36.4 feet. 
-    More decimal points strain credibility.
-
-    \item The 11 address-related fields (\texttt{incident\_address}, 
-    \texttt{street\_name}, \texttt{city}, \texttt{intersection\_street\_1 \& 2}, 
-    \texttt{cross\_street\_1 \& 2}, \texttt{landmark}, \texttt{road\_ramp}, 
-    \texttt{bridge\_highway\_segment}, and \texttt{taxi\_pick\_up\_location}) 
-    should be standardized to USPS format. Software can easily 
-    perform this standardization.
-
-    \item There are multiple ways missing data is represented, such as 
-    ``UNKNOWN'', ``N/A'', ``NA'', ``na'', and blanks (`` ''). Protocols 
-    should be established to standardize the representation of missing data.
-\end{itemize}
+Building upon the insights gained from our case study of the 
+NYC 311 Service Request (SR) data, we propose a set of data 
+curation principles tailored for government-released open datasets. 
+These principles are designed to address the unique challenges 
+and requirements observed in the curation of such datasets, 
+ensuring that they remain reliable, consistent, and useful for 
+public services, research, and other applications.
+
+\subsection{Principles}
+\paragraph{Principle 1: Ensuring Consistency Across Agencies and Time}
+A key challenge in government-released datasets is the 
+harmonization of data fields across various agencies and 
+maintaining this consistency over time. As seen in the NYC 
+311 data, discrepancies in field names, formats, and definitions, 
+such as those between \texttt{police\_precinct} and 
+\texttt{police\_precincts}, can introduce significant 
+inconsistencies. To address this, agencies should adopt 
+standardized naming conventions and formats across all datasets. 
+Regular audits and data harmonization processes should be 
+implemented to ensure consistency across long-term datasets, 
+particularly as agency structures or reporting standards evolve. 
+For instance, ensuring that historical datasets reflect the same 
+field naming conventions over a span of years can prevent 
+inconsistent analyses.
+
+\paragraph{Principle 2: Guaranteeing Accuracy and Validity in Data Entry}
+Data accuracy and validity are fundamental to the utility of 
+open datasets. In the NYC 311 data, issues such as 
+invalid zip codes (e.g., 58\% invalid \texttt{zip\_codes} entries) and 
+nonsensical date entries (e.g., negative durations) highlight the need 
+for rigorous validation processes. Governments should establish 
+clear protocols for data entry, ensuring that fields are populated with 
+valid, legally acceptable values. For example, fields like 
+\texttt{created\_date} and \texttt{closed\_date} should be subject to 
+automated checks that prevent logical errors, such as closing an 
+SR before it was opened. Ensuring the integrity of critical fields 
+like geographic locations and agency names enhances the quality 
+and reliability of the data.
+
+\paragraph{Principle 3: Optimizing Storage and Representation Efficiency}
+Efficient storage and data representation are crucial for handling 
+large government datasets. By eliminating redundant fields and 
+encoding categorical variables, agencies can significantly reduce 
+file sizes and improve the efficiency of data analysis. In the NYC 
+311 dataset, for example, redundant fields such as \texttt{borough} 
+and \texttt{park\_borough} could be consolidated without any data 
+loss, saving space and reducing file size by up to 12.3\%. Another 
+example is the \texttt{taxi\_company\_borough} field, which is 99.94\% 
+blank and should be evaluated to determine if its inclusion is 
+necessary. Similarly, categorical variables like \texttt{complaint\_type} 
+could be encoded in a standardized format to optimize storage and 
+streamline analyses. Reducing file size not only saves storage space 
+but also enhances the performance of queries and analysis processes.
+
+\paragraph{Principle 4: Updating Data Dictionaries and Documentation}
+Clear and accurate documentation is essential for the effective use 
+of government-released datasets. The NYC 311 data revealed several 
+issues where the Data Dictionary did not accurately reflect the 
+actual data, such as missing field descriptions for computed fields. 
+Updating the Data Dictionary to include descriptions of all fields, 
+especially computed fields like \texttt{@computed\_region\_xxxx}, 
+would improve transparency and usability. Regular updates should 
+be made to ensure that any changes in data structure or content are 
+reflected. This practice enhances the ability of external users, 
+including researchers and developers, to understand and work 
+effectively with the data.
+
+\paragraph{Principle 5: Automating Quality Assurance Processes}
+Automating data validation and quality assurance processes can 
+significantly improve the reliability of government datasets. In the 
+NYC 311 case, issues such as large spikes in SR creation and closure 
+at exactly midnight or noon likely stem from automated processes that 
+distort the accurate capture of SR durations. By implementing 
+real-time validation tools, governments can prevent such anomalies 
+from entering the dataset. Automated systems can be designed to 
+flag inconsistencies, incorrect values, or illogical sequences, such 
+as invalid \texttt{incident\_zip} values or illogical date entries. 
+Automation can also facilitate faster detection and correction of 
+errors, reducing the need for manual intervention and ensuring 
+the data remains clean and reliable as it is continuously updated.
+
+\paragraph{Principle 6: Addressing Data Transparency and Accessibility}
+Transparency is a core tenet of open data, and governments must ensure 
+that datasets are not only accurate but also accessible and easily 
+understandable. This involves providing clear metadata and making 
+datasets available in user-friendly formats. For example, in the NYC 
+311 dataset, fields like \texttt{@computed\_region\_xxxx\_xxxx} and 
+the distinction between \texttt{incident\_zip} and \texttt{zip\_codes} 
+can be confusing for non-expert users. These fields should be clearly 
+documented, explaining how they are calculated and how they differ. 
+Additionally, governments should communicate transparently about 
+data quality issues, such as notifying the public when errors in SR 
+fields (e.g., erroneous closed dates) are discovered and explaining 
+how such issues will be resolved. By ensuring transparency and 
+keeping the public informed, governments foster trust and promote 
+wider use of datasets for research, civic engagement, and innovation.
+
+\subsection{Actionable Steps}
+While these principles are broad, actionable steps are required to 
+implement them effectively. Governments can begin by investing in 
+real-time validation systems, regular updates to data dictionaries, 
+and tools for encoding and standardizing data fields. For example, 
+systems that automate the enforcement of data validation rules 
+(e.g., flagging incorrect zip codes) can significantly improve data 
+quality. Collaboration with city agencies and stakeholders is crucial 
+to ensure that data standards, formats, and metadata meet the needs 
+of diverse users, including policymakers, researchers, and the 
+general public. Regular communication between agencies and external 
+users can help identify recurring issues and areas where improvements 
+are needed. By facilitating regular communication and collaboration 
+with various stakeholders, governments can ensure data remains relevant 
+and actionable for public use.
+
+Engagement with data scientists and statisticians is key to improving 
+data curation strategies. These experts can help design and refine 
+curation processes, such as creating algorithms for automatic data 
+cleaning, identifying inconsistent data entries, and ensuring the 
+dataset’s structure is maintained over time. Their expertise is also 
+critical in developing protocols for validating incoming data, correcting 
+errors in real-time, and ensuring consistency across multiple data sources. 
+Feedback from statisticians and data scientists, such as the issues reported 
+in this paper, needs a clear and accessible channel to reach the 311 SR data 
+team and other relevant city agencies. Establishing such feedback loops 
+ensures that recommendations for improving data curation and validation 
+are not only heard but implemented. This feedback mechanism would also allow 
+ongoing dialogue between data users and city officials, leading to more dynamic 
+and responsive data management practices.
+
+A particularly valuable avenue for engagement is through events such as NYC 
+Open Data Week, which fosters collaboration and innovation across sectors. 
+Our involvement in this project stemmed from such an event, where we were 
+inspired by the potential of open data to enhance public services. Governments 
+should continue supporting these kinds of initiatives, which not only promote 
+data literacy but also encourage community-driven improvements to datasets. 
+By leveraging the expertise of data scientists, statisticians, stakeholders, 
+and civic organizations, governments can ensure that their open datasets 
+remain accurate, transparent, and useful for a wide range of applications.
+
+% \begin{itemize}
+% \item  Undertake a dedicated effort 
+%     to investigate these data anomalies, confirm the veracity of the 
+%     issues, and collaborate with the appropriate Agency Open Data 
+%     Representatives to find resolutions. This effort may be complex 
+%     and require software changes. Coordination and mutual approval 
+%     of changes will be key, requiring strong support from senior 
+%     leadership.
+
+%     \item Standards and protocols should be established, agreed upon, 
+%     and enforced. This includes defining the domain of legal values, 
+%     ensuring data accuracy, implementing address normalization, 
+%     clarifying data usage, and providing detailed data definitions.
+
+%     \item Resolve many potential data integrity issues
+%     through enforcement by software. For instance, software should 
+%     prevent invalid entries like erroneous zip codes. Similarly, fields 
+%     like \texttt{created\_date} and \texttt{closed\_date}, which 
+%     measure responsiveness, should have logic checks to prevent 
+%     illogical sequences, such as an SR being ``closed'' before 
+%     being ``opened.''
+
+%     \item The large spikes in SR creation and closure at midnight and 
+%     noon likely point to automated processes from external Agency 
+%     integrations. These distortions should be eliminated in favor of 
+%     capturing accurate SR durations with precise timestamps.
+
+%     \item As discussed, there are six ``computed'' fields that are not 
+%     listed in the Data Dictionary. Some of these fields contain 
+%     significant inaccuracies. We recommend updating the Data 
+%     Dictionary and, if necessary, masking these fields from public use 
+%     until their accuracy is improved.
+
+%     \item Duplicate fields should be reviewed for removal. In some 
+%     cases, such as \texttt{borough} and \texttt{park\_borough}, the 
+%     duplication is 100\%. In others, the match is 80-90\%. 
+%     These near-duplicates should be assessed to determine their 
+%     relevance in the 311 SR dataset. Initial analysis suggests that 
+%     removing these fields could reduce file size by up to 12\%, with 
+%     further reductions possible upon deeper review.
+
+%     \item Fields with very low data presence, such as 
+%     \texttt{taxi\_company\_borough}, populated only 0.06\% of the time, 
+%     should be evaluated to determine if they are necessary.
+
+%     \item The Data Dictionary, last updated in March 2023, should be 
+%     revised based on the issues discussed in Section~\ref{sec:datadictionary}.
+
+%     \item The precision of latitude, longitude, and location fields 
+%     should be reduced. We recommend limiting precision to four 
+%     decimal places (0.000X), which represents approximately 36.4 feet. 
+%     More decimal points strain credibility.
+
+%     \item The 11 address-related fields (\texttt{incident\_address}, 
+%     \texttt{street\_name}, \texttt{city}, \texttt{intersection\_street\_1 \& 2}, 
+%     \texttt{cross\_street\_1 \& 2}, \texttt{landmark}, \texttt{road\_ramp}, 
+%     \texttt{bridge\_highway\_segment}, and \texttt{taxi\_pick\_up\_location}) 
+%     should be standardized to USPS format. Software can easily 
+%     perform this standardization.
+
+%     \item There are multiple ways missing data is represented, such as 
+%     ``UNKNOWN'', ``N/A'', ``NA'', ``na'', and blanks (`` ''). Protocols 
+%     should be established to standardize the representation of missing data.
+% \end{itemize}
 
@@
 \bibliographystyle{jds}
-\bibliography{ref}
+\bibliography{refs}
 

### 2024-09-23 manuscript/manuscript.tex

@@
 incomplete data, correcting mislabeled or inaccurate entries, and
-standardizing data formats. In the context of open data, cleansing is
+standardizing data formats \citep[e.g.,][]{cody2017cody,
+  van2018statistical}. In the context of open data, cleansing is
 especially important as open datasets often come from diverse,

### 2024-09-25 code/timelinecharts..R
Review of Section 1 and 2 of manuscript. Working changes to charts
Working on charts to remove title and expand annotations.

@@ -1,9 +1,9 @@
-# install.packages("ggplot2")
-# install.packages("scales")
-# install.packages("dplyr")
-# install.packages('zoo')
-# install.packages("ggpmisc")
-# install.packages("lubridate")
+install.packages("ggplot2")
+install.packages("scales")
+install.packages("dplyr")
+install.packages('zoo')
+install.packages("ggpmisc")
+install.packages("lubridate")
 
 library(ggplot2)
 library(scales)
@@ -32,7 +32,7 @@ year_digits <- substr(main_data_file, 1, 2)
 if (year_digits != "10") { year_digits <- "2"} 
   # Set path for the chart directory
 
-chart_prefix <- paste0(year_digits, "-year trends")
+chart_prefix <- paste0(year_digits, "-year")
 chart_directory_path <- file.path("..", "..", "charts", "2022-2023 study", chart_prefix)
 
 file_name_prefix <- chart_prefix #to name individual chart files
@@ -49,7 +49,7 @@ files <- list.files(functions_path, pattern = "\\.R$", full.names = TRUE)
 # Source each file
 lapply(files, source)
 
-sink(paste0("../../console_output/", year_digits, "-yr timeline_console_output.txt"))
+#sink(paste0("../../console_output/", year_digits, "-yr timeline_console_output.txt"))
 cat("\nExecution begins at:", formattedStartTime)
 
 #########################################################################
@@ -127,7 +127,6 @@ print_threshold <- nrow(sorted_by_agency)
 sorted_by_agency <- as.data.frame(sorted_by_agency)
 print(head(sorted_by_agency, print_threshold), row.names = FALSE, right = FALSE)
 
-#source("functions/create_combo_chart.R")
 chart_title <- "Complaints by Agency"
 chart_file_name <- "Complains_by_Agency.pdf"
 

### 2024-09-26 manuscript/manuscript.tex
Review of sections 1 and 2
Responding to comments from JY inserted into manuscript text.

@@
 \documentclass[linenumber]{jdsart}
+ \usepackage{setspace}
 
@@
 
+\doublespacing
 \begin{document}
@@
 societies and governments by promoting transparency,
-innovation, and public engagement. New York City (NYC) has been at
+innovation, and public engagement. The City of New York (NYC) has been at
 the forefront of this movement since the enactment of the Open 
-Data Law in 2012, which led to the creation of the NYC Open Data
+Data Law in 2012, leading to the creation of the NYC Open Data
 portal. This portal now hosts 2,700 datasets from 80 city agencies,
@@
 curation in ensuring the utility and reliability of open datasets.
+
 This paper examines the challenges of open data curation through a
-case study of the NYC 311 Service Request (SR), addressing issues 
+case study of the NYC 311 Service Request (SR) dataset, addressing issues 
 of data validity, consistency, and curation efficiency. Based on 
@@
 breaking down barriers that once kept valuable information 
-on government performance in silos. Their collective impact 
+regarding government performance in silos. Their collective impact 
 extends beyond mere data sharing to fostering a culture of openness 
@@
 
-New York City (NYC) has emerged as a leader in the open data movement,
+The City of New York (NYC) has emerged as a leader in the open data movement,
 marked by the enactment of the Open Data Law in 2012
@@
 
-Research into data curation has explored these challenges in depth. 
+Research into data curation has explored these challenges in\mbox{-}depth. 
 Among the earliest discussions, \citet{witt2009constructing} focused 
@@
 and the presence of missing, blank, or N/A entries. This section
-also investigates field compliance with legal or acceptable values, 
+utilizes 311 SR data over a a 2-year period from 2022-2023.  It 
+ investigates field compliance with reference or acceptable values, 
 highlights logical inconsistencies, and examines concerning patterns 
@@
 include the addition of online and mobile app channels in 2009, a
-\jy{should we mention the system upgrade just before the pandemic to
-  document this historical event?}
 record high of 348,463 monthly service requests in August 2020 due to
@@
 subway system, and a record 3.23 million service requests in
-2023. Today, the NYC 311 data system manages over 3 million service
-requests per year. This data is publicly accessible through the NYC Open Data
+2023.
+
+Today, the NYC 311 data system manages over 3 million service
+requests (SRs) per year. This data is publicly accessible through the NYC Open Data
 Portal, which provides tools for querying, grouping, aggregating,
@@
 \href{https://www.nyc.gov/content/oti/pages/}{NYC Office of Technology
-  and Innovation (OTI)}, which provides technological support for the
-open data system.
+  and Innovation (OTI)}, provides technological support for the
+open data system and data portal. Input data is sourced from 17 different
+NYC Agencies.
 
@@
 identifiable information (PII), integration with stand-alone systems
-at selected NYC agencies, and managing API usagefor numerous
+at selected NYC agencies, and managing API usage for numerous
 third-party users. These challenges are addressed by various agency
@@
 
-Our investigation focused on a 10-year period (2014--2023), downloaded
-on \jy{Give date because the data is updated daily.}
-This version of the data is available as a supplement to facilitate
-reproducibility of the analysis presented in the sequel.
-\jy{How many total? file size? Somewhere we used only two years that
-  need to be clarified.}
-Figure~\ref{fig:10-yr} shows the annual counts of NYC 311 SR over the
+Our investigation utilizes two 311 SR datasets covering 10-years and 2-years.
+The 10-year period (2014--2023), downloaded on 15 September 2024, 
+contains only three columns (created\_date, closed\_date, and 
+agency). This dataset is used to analyze long-term 
+trends and observe certain dating anomalies. The 10-year dataset is approximately 
+1.4 GB in size and contains over 27 million rows, each row representing one SR.
+ Instructions on downloading this file from the NYC Open Data Portal 
+ are available in the supplemental material.
+
+Additionally, this investigation uses a 2-year dataset covering the period
+2022--2023, also downloaded on 15 September 2024. This data set 
+contains the full 41 columns of 311 SR data available from the portal, 
+and is used to conduct detailed analysis in areas such as data consistency, 
+data validity, redundant fields, etc.. This 2-year dataset 
+is quite large at 3.7GB. It contains over 6 million rows of SRs. Instructions 
+on downloading this file from the NYC Open Data Portal are 
+available in the supplemental material.
+
+Figure~\ref{fig:10-yr} shows  annual counts of 311 SRs over the
 10-year period from 2014 to 2023, revealing a clear upward trend in
-public engagement with the system. The most significant increase
-occurred in 2020, when the number of requests surged to 3.23 million
-due to the COVID-19 pandemic, illustrating the essential role of the
+public engagement. As indicated, the most significant increase
+in activity occurred in 2020, when the number of requests surged to 3.23 million
+during the COVID-19 pandemic, illustrating the essential role of the
 311 system in supporting residents during times of crisis. Overall, SR
-counts increased by nearly 49.5\% between 2014 and 2023, reflecting
-growing public reliance on the system for addressing non-emergency
-issues. The steady rise in SRs can be attributed to the increased
-accessibility of the 311 system via online and mobile platforms, as
-well as heightened public awareness of the service. While the spike in
-2020 was exceptional, the broader trend indicates sustained growth in
-service request volume over the decade, highlighting the system's
-evolving role in managing both routine city operations and
-extraordinary events.
-
-\jy{The 'title' in each figure could be removed because we have
-  captions. The fonts of the annotations in the figures need to be
-  increased. The code generating them need to be tidied for future
-  sharing.
-}
+counts increased by 49.5\% between 2014 and 2023, reflecting
+growing public reliance on the system. (During this 10-year time 
+period, the population of NYC grew just 0.6\%.) The steady 
+rise in SRs can be attributed to the increased accessibility of 
+the 311 system via online and mobile platforms, as well as 
+heightened public awareness  and utilization of the service. 
+While the spike in 2020 was exceptional, the broader trend 
+indicates sustained growth in service request volume over 
+the decade, highlighting the system's evolving role in managing 
+both routine city operations and extraordinary events.
 
 \jy{The plot sizes seem to be fixed in the generation code. That's why
-  the fonts are small when scaled. I'd generate them with approxiately
+  the fonts are small when scaled. I'd generate them with approximately
   the expected sizes in the manuscript. I'd also use aspect ratio
@@
 agency, showing the cumulative percentage of SRs handled by each
-agency over a given time period. This figure highlights that the
+agency over a the 2-year time frame. Note that the
 distribution of SRs is concentrated among a few key agencies. The six
@@
 and Department of Parks and Recreation (DPR).
-They collectively handle over 90\% of the total SR volume. This is
+Collectively they handle over 90\% of the total SR volume. This is
 indicative of the critical roles these agencies play in managing
@@
 substantial portion of the total SRs, underscoring its central role in
-addressing complaints related to public safety and order. The
-remaining 10\% distributed across a large number of smaller
-agencies. This concentration of requests points to the systemic
-reliance on a few key departments for resolving most non-emergency
-issues, further emphasizing the importance of optimizing data handling
-and curation processes within these high-volume agencies to ensure
+addressing complaints related to public safety. The
+remaining 10\% is distributed across a number of other
+agencies. This concentration of SRs points to a
+reliance on a few key departments to resolve most non-emergency
+issues (e.g. non 911 calls). Accordingly, this underscores 
+the necessity for optimal data handling and curation processes 
+within these high-volume agencies in order to ensure 
 efficiency and responsiveness.
@@
 
-\begin{comment}
-\begin{table}[tbp]
-  \centering
-  \caption{Noise-related complaints\_type(s) by count with Agency}
-  \label{tab:noisecomplaints}
-  \begin{tabular}{@{}lS[table-format=7.0,round-mode=places,
-    round-precision=0]S[table-format=2.2,round-mode=places,
-    round-precision=2]l@{}} % 'l' for left-aligned, 'S' for siunitx number alignment
-    \toprule
-    \textbf{complaint\_type} & \textbf{Count} & \textbf{Percentage} & \textbf{Agency} \\ 
-    \midrule
-    Noise - Residential        & 675502 & 10.56 & NYPD  \\ 
-    Noise - Street/Sidewalk    & 300507 &  4.70 & NYPD  \\ 
-    Noise - Commercial         & 128892 &  2.02 & NYPD  \\ 
-    Noise - Vehicle            & 115956 &  1.81 & NYPD  \\ 
-    Noise                      & 100413 &  1.57 & DEP   \\ 
-    Noise - Helicopter         &  85345 &  1.33 & EDC   \\ 
-    Noise - Park               &  16620 &  0.26 & NYPD  \\ 
-    Noise - House of Worship &   2757 &  0.04 & NYPD  \\ 
-    \bottomrule
-  \end{tabular}
-\end{table}
-\end{comment}
-
 \begin{figure}[tbp]
@@
   \includegraphics[width=0.7\textwidth]{SR_by_Complaint_Type.pdf} 
-  \caption{Top 20 complaint type(s) and Cumulative Percentage} 
+  \caption{Top 20 complaint types and Cumulative Percentage} 
   \label{fig:SR_complaints}
@@
 illustrates the top 20 complaint types, accounting for 70\% 
-of all Service Requests (SRs), and their cumulative percentage. The 
+of all SRs and their cumulative percentage. Note that the 
 distribution of complaints is skewed, with a small number of issues 
-dominating the total SR volume. Noise-related complaints, including 
+dominating total SR volume. Noise-related complaints, including 
 residential, commercial, and street noise, make up 22\% of all 
@@
 uncoordinated sources, leading to variations in data quality,
-completeness, and consistency. Without proper cleansing, the utility
-of open data can be severely limited, affecting its reliability for
+completeness, and consistency. Without thorough data cleansing, the utility
+of open data can be quite limited, affecting its reliability for
 research, policy-making, and innovation. The main purpose of cleansing
@@
 learning models or other systems.
-
-% What is data cleansing?  Wikipedia 
-% \href{https://en.wikipedia.org/wiki/Data_cleansing}{Data Cleansing} 
-% offers this definition.
-
-% \begin{quote}\textit{Data cleansing or data cleaning is the process of detecting and 
-% correcting (or removing) corrupt or inaccurate records from a record set, 
-% table, or database and refers to identifying incomplete, incorrect, 
-% inaccurate or irrelevant parts of the data and then replacing, 
-% modifying, or deleting the dirty or coarse data.}
-% \end{quote}
 
@@
 important consideration, such as related fields that violate expected 
-relationships, like a ``due date'' that precedes the ``created date.'' 
+relationships, such as a ``due date'' that precedes the ``created date.'' 
 Lastly, the balance between accuracy and precision is crucial, as both 
@@
 Here we identify the presence of issues in the NYC 311 SR open data;
-not necessarily to solve them. A solution effort should
+not attempting to solve them. A solution effort should
 be undertaken only after an investigation as to the why and how 
 such issues came about, and a discussion as to whether or not it 
-is an \textit{error}.
+is an actual \textit{error} or holds some other status.
 
@@
 \begin{itemize}
-	\item There are 47 columns of data for each row, exportable as a CSV file.
+	\item There are 41 columns of data for each row, exportable as a CSV file.
+	Each row represents one Service Request.
 
@@
 
-	\item There are three borough fields; two of which appear to be duplicates.
-
-	\item Two zip code fields, but not duplicates
+	\item There are two borough fields; one of which appear to be a duplicate.
 
 	\item Seven street fields; two pair of which appear to be duplicates
-
-	\item Two Police Precinct fields; not duplicates
 
@@
 identifies 41 data columns (fields) and provides related information 
-for each. However, upon downloading the data, six additional fields 
-appear that are not included in the Data Dictionary. These fields are 
-visible in the portal's Column Manager widget, labeled with the prefix 
+for each. However, there are six additional fields that can be
+optionally downloaded. These fields are 
+visible in the portal's ``Column manager'' widget, labeled with the prefix 
 ``@computed\_region.'' The additional fields include \texttt{zip\_codes}, 
@@
 \texttt{city\_council\_districts}, \texttt{police\_precincts}, and 
-\texttt{police\_precinct}. While one can infer the meaning and 
-derivation of these computed fields to some extent, uncertainty remains. 
-The critical question is whether these fields can be reliably used for 
-analytical purposes reliable documentation. In particular, which one
-of the two percinct fields should one use? What is the difference
-between the computed zip codes and the incident zipcode among the 41
-documented variables?
+\texttt{police\_precinct}. While one can likely infer the meaning and 
+derivation of these computed fields, uncertainty remains as to
+their derivation and usage. These six fields are not included in
+the Data Dictionary. Our analysis has shown that these six 
+fields have numerous data validity issues. Liaison with the 
+NYC Open Data Teams leads us to treat these six fields 
+as experimental and not for suitable for normal use. Accordingly, 
+we will not address these fields in our analysis.
 
@@
 should be valid as defined by the USPS database, which contains 
-37,946 valid zip codes. However, the computed field \texttt{zip\_codes} 
-proved problematic, with over 58\% (3.6 million) of its entries being 
-invalid. Although the more accurate \texttt{incident\_zip} field has 
-only 0.07\% invalid entries, this still results in 4,163 errors, which 
-is not insignificant. The breakdown of invalid entries in the 
-\texttt{zip\_codes} field, sorted by Agency, shows that the
-\jy{Which agency? Missing a table? We could prepare a supplement where
-  we can put additional tables/figures without limit.}
-distribution by percentage mirrors the overall breakdown of SRs by 
-Agency, potentially indicating a systemic problem.
+37,946 valid zip codes. The \texttt{incident\_zip} field has 
+only 0.07\% invalid entries, this still results in 4,163 errors; 
+not insignificant. If you sort the invalid entries by Agency, 
+the distribution by percentage mirrors the \texttt{overall} 
+breakdown of SRs by Agency, potentially 
+indicating a systemic problem.
 
@@
 
-%\begin{figure}[tbp]
-%	 \centering
-%	 \includegraphics[width = \textwidth]{zero_duration_SR.pdf}
-%	 \caption{SRs with Zero Durations by Agency}
-%	 \label{fig:zero-duration}
-%\end{figure}
-
 \subparagraph{Created Date or Closed Date(s) at Midnight or Noon}
@@
 \end{itemize}
+
+A unique way to visualize this anomaly is to examine the busiest day 
+during this 2-year period (Friday, 2023-09-29). Then aggregate 
+the created\_date(s) by minute (with seconds equal to zero), providing
+a minute-by-minute look at SR creation on that one day. Note the spike 
+at exactly noon (12:00), well beyond the 3\textsigma line.
+
+\begin{figure}[tbp]
+    \centering
+    \begin{minipage}{0.5\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{2-year-trend_SR_created_by_minute_of_busiest_day.pdf}
+        \caption{SRs Created Minute-by-Minute on Busiest Day}
+        \label{fig:busiestcreated}
+    \end{minipage}%
+    \begin{minipage}{0.5\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{2-year-trend-SR_closed_by_minute_of_busiest_day.pdf}
+        \caption{SRs Closed Minute-by-Minute on Busiest Day}
+        \label{fig:busiestclosed}
+    \end{minipage}
+\end{figure}
 
@@
 
-Table~\ref{tab:x1nearmatches} provides examples of near-matches for 
+Table~\ref{tab:x1near-matches} provides examples of near-matches for 
 \texttt{cross\_street\_1} and \texttt{intersection\_street\_1}. These 
@@
 cleaning, identifying inconsistent data entries, and ensuring the 
-dataset’s structure is maintained over time. Their expertise is also 
+dataset(s) structure is maintained over time. Their expertise is also 
 critical in developing protocols for validating incoming data, correcting 
@@
 \section*{Supplementary Material}
-The NYC 311 SR data and code used to produce the results reportred in
+The NYC 311 SR data and code used to produce the results reported in
 this paper are available [FixMe].

### 2024-09-26 code/datacleansing.R

@@ -363,7 +363,7 @@ create_boxplot <- function(
     chart_file_name) {
   
     boxplot_chart <- ggplot(data = dataset, aes(x = duration, y = factor(1))) +
-      geom_jitter(color = "#2271B2", alpha = 0.4, size = 1.9, shape = 17) +
+      geom_jitter(color = "#0072B2", alpha = 0.4, size = 1.9, shape = 17) +
       geom_boxplot(width = 0.2, fill = "#E69F00", alpha = 0.7, color = "black") +
       theme(
         legend.position = "none", plot.title = element_text(hjust = 0.5),
@@ -387,7 +387,7 @@ create_violin_chart <- function(
   violin_chart <- ggplot(data = dataset, aes(x = !!rlang::sym(x_axis_field), y = factor(1))) +
     geom_jitter(width = 0.25, alpha = 0.4, color = "#0072B2", size = 1.9, shape = 17) +
     geom_violin(linewidth = 0.7, fill = "transparent", color = "black") +
-    geom_boxplot(width = 0.25, fill = "#D55E00", color = "black", alpha = 0.6, 
+    geom_boxplot(width = 0.25, fill = "#E69F00", color = "black", alpha = 0.6, 
                  outlier.colour = "black", outlier.size = 0.75) +
     labs(
       title = chart_title,
@@ -773,7 +773,7 @@ increment <- result$increment
 
 # Create the bar chart with vertical X-axis labels
 blank_chart <- ggplot(missingDataPerColumn, aes(x = reorder(field, -total_empty), y = total_empty)) +
-  geom_bar(stat = "identity", fill = "#117733") +
+  geom_bar(stat = "identity", fill = "#009E73") +
   theme(
     axis.title.x = element_text(vjust = 0, size = 11),
     axis.title.y = element_text(vjust = 1, size = 11),
@@ -792,7 +792,8 @@ blank_chart <- ggplot(missingDataPerColumn, aes(x = reorder(field, -total_empty)
     yintercept = seq(starting_value, max_count, by = increment),
     linetype = "dotted", color = "gray40"
   ) +
-  ggtitle("Number and % of blank/missing fields per column",
+#  ggtitle("Number and % of blank/missing fields per column",
+  ggtitle(NULL,
     subtitle = paste(chart_sub_title, format(num_rows_d311, big.mark = ","), sep = "")
   ) +
   labs(y = NULL, x = NULL)
[…]

### 2024-09-26 code/functions/create_bar_chart.R

@@ -65,21 +65,21 @@ create_bar_chart <- function(
   
   # Create the bar chart
   bar_chart <- ggplot(dataset, aes(x = .data[[x_col]], y = .data[[y_col]])) +
-    geom_bar(stat = "identity", fill = "#117733") +
+    geom_bar(stat = "identity", fill = "#009E73") +
     scale_x +
-    theme(
-      axis.title.x = element_text(vjust = 0, size = 11),
-      axis.title.y = element_text(vjust = 1, size = 11),
-      plot.title = element_text(hjust = 0.5, size = 13),
-      plot.subtitle = element_text(size = 9),
-      panel.background = element_rect(fill = "gray95", color = "gray95"),
-      axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, face = "bold"),
-      axis.text.y = element_text(face = "bold"),
-      legend.position = "none" # Remove legend
-    ) +
+    # theme(
+    #   axis.title.x = element_text(vjust = 0, size = 11),
+    #   axis.title.y = element_text(vjust = 1, size = 11),
+    #   plot.title = element_text(hjust = 0.5, size = 13),
+    #   plot.subtitle = element_text(size = 9),
+    #   panel.background = element_rect(fill = "gray95", color = "gray95"),
+    #   axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, face = "bold"),
+    #   axis.text.y = element_text(face = "bold"),
+    #   legend.position = "none" # Remove legend
+    # ) +
     geom_hline(
       yintercept = seq(starting_value, y_max_count, by = increment),
-      linetype = "dotted", color = "gray40", linewidth = 0.3
+      linetype = "dotted", color = "gray40", linewidth = 0.5
     ) +
     ggtitle(chart_title, subtitle = paste(sub_title, format(y_total_count, big.mark = ","), sep = "")) +
     labs(x = x_axis_title, y = y_axis_title)
@@ -89,7 +89,7 @@ create_bar_chart <- function(
       annotate("text",
                x = max_row[[x_col]], y = y_max_count,
                label = paste0("Max: ", format(y_max_count, big.mark = ",")),
-               size = 3.7, color = "black", vjust = -0.4, hjust = 0.5
+               size = 5, color = "black", vjust = -0.4, hjust = 0.5
       )
   }
   
@@ -103,7 +103,7 @@ create_bar_chart <- function(
       annotate("text",
[…]

### 2024-09-26 code/functions/create_combo_chart.R

@@ -55,7 +55,7 @@ create_combo_chart <- function(
   combo_chart <- ggplot(summary_df) +
     
     geom_bar( aes( x = reorder(agency, cumulative_percentage), y = count),
-              stat = "identity", fill = "#D55E00", width = 0.55 ) +
+              stat = "identity", fill = "#E69F00", width = 0.55 ) +
     
     geom_line( aes( x = reorder(agency, cumulative_percentage), y = cumulative_percentage*max_count,
                     group = 1,), colour = "black", linewidth = 1, linetype = "dotted", ) +
@@ -69,17 +69,17 @@ create_combo_chart <- function(
     geom_hline( yintercept = seq(starting_value, max_count, by = increment), 
                 linetype = "dotted", color = "gray40", linewidth = 0.45 ) +
     
-    theme(
-      axis.title.x = element_text(vjust = 0, size = 11),
-      axis.title.y.right = element_text(vjust = 1, size = 11, color = "black"),
-      axis.title.y.left = element_blank(),
-      axis.text.y.right = element_text(color = "black"),
-      axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1, face = "bold"),
-      plot.title = element_text(hjust = 0.5, size = 13),
-      plot.subtitle = element_text(size = 9),
-      legend.position = "none",
-      plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 5.5)
-    ) +
+    # theme(
+    #   axis.title.x = element_text(vjust = 0, size = 11),
+    #   axis.title.y.right = element_text(vjust = 1, size = 11, color = "black"),
+    #   axis.title.y.left = element_blank(),
+    #   axis.text.y.right = element_text(color = "black"),
+    #   axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1, face = "bold"),
+    #   plot.title = element_text(hjust = 0.5, size = 13),
+    #   plot.subtitle = element_text(size = 9),
+    #   legend.position = "none",
+    #   plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 5.5)
+    # ) +
     
     ggtitle(
       chart_title, subtitle = paste("(", earliest_title, "--", latest_title, ")",
@@ -94,7 +94,23 @@ create_combo_chart <- function(
     
     labs(x = NULL, y = NULL)
   
+  combo_chart <- combo_chart + theme(
[…]

### 2024-09-26 code/functions/create_special_bar_chart.R

@@ -8,7 +8,7 @@ create_special_bar_chart <- function(data, x_col, y_col, chart_title, subtitle_p
   
   # Create the bar chart with vertical X-axis labels
   chart <- ggplot(data, aes(x = !!sym(x_col), y = !!sym(y_col))) +
-    geom_bar(stat = "identity", fill = "#117733") +
+    geom_bar(stat = "identity", fill = "#009E73") +
     scale_x_discrete() +
     scale_y_continuous(labels = y_axis_labels) +
     theme(

### 2024-09-26 code/timelinecharts..R

@@ -1,9 +1,9 @@
-install.packages("ggplot2")
-install.packages("scales")
-install.packages("dplyr")
-install.packages('zoo')
-install.packages("ggpmisc")
-install.packages("lubridate")
+# install.packages("ggplot2")
+# install.packages("scales")
+# install.packages("dplyr")
+# install.packages('zoo')
+# install.packages("ggpmisc")
+# install.packages("lubridate")
 
 library(ggplot2)
 library(scales)
@@ -22,7 +22,7 @@ cat("\n***** Program initialization *****")
 setwd("C:/Users/david/OneDrive/Documents/datacleaningproject/nyc311clean/code")
 
 # Set path for the data file
-main_data_file <- "10-year 2014-2023.csv"
+main_data_file <- "311_Service_Requests_from_2022-2023_AS_OF_09-15-2024.csv"
 
 data1File <- file.path("..", "..", "data", main_data_file)
 
@@ -49,7 +49,7 @@ files <- list.files(functions_path, pattern = "\\.R$", full.names = TRUE)
 # Source each file
 lapply(files, source)
 
-#sink(paste0("../../console_output/", year_digits, "-yr timeline_console_output.txt"))
+sink(paste0("../../console_output/", year_digits, "-yr timeline_console_output.txt"))
 cat("\nExecution begins at:", formattedStartTime)
 
 #########################################################################
@@ -85,12 +85,12 @@ d311 <- consolidate_agencies((d311))
 # Remove any rows where the created_date is missing. This is a required field.
 d311 <- d311[!is.na(d311$created_date), ]
 
+# Filter out rows with NA values in the created_date column
+d311 <- d311[!is.na(d311$created_date), ]
+
 # Convert character date-time strings to datetime objects with America/New_York timezone
 d311$created_date <- as.POSIXct(d311$created_date, format = "%m/%d/%Y %I:%M:%S %p", tz = "America/New_York")
 d311$closed_date <- as.POSIXct(d311$closed_date, format = "%m/%d/%Y %I:%M:%S %p", tz = "America/New_York")
-
-# Filter out rows with NA values in the created_date column
-d311 <- d311[!is.na(d311$created_date), ]
 
 #########################################################################
[…]

### 2024-09-26 manuscript/manuscript.tex
added data cleaning to intro

@@
 consistent, accurate, and useful for diverse applications. For example, 
-curating data ensures it can be leveraged effectively for machine learning 
-systems, which require high-quality data to produce reliable insights
-\citep{polyzotis2019data, jain2020overview}. Poor 
-curation may result in issues such as missing data, formatting errors, or 
-inconsistent values that can lead to biased or inaccurate outcomes
-\citep{geiger2020garbage}.
-This is particularly critical in domains where machine learning is applied 
-to sensitive tasks, such as public health or policy
-\citep{rahm2000data}.
+well-curated data is essential for machine learning systems, which 
+require high-quality data to produce reliable insights 
+\citep{polyzotis2019data, jain2020overview}. A critical component of 
+data curation is data cleaning, which involves identifying and rectifying 
+inconsistencies, errors, and inaccuracies in datasets. While open data 
+initiatives have made vast amounts of data available, ensuring the 
+reliability of such data hinges on rigorous data cleaning processes. 
+Efficient usage of software tools can significantly streamline this 
+aspect of curation \citep[e.g.,][]{cody2017cody, van2018statistical}. Poor 
+curation may result in issues such as missing data, formatting errors, 
+or inconsistent values, leading to biased or inaccurate outcomes 
+\citep{geiger2020garbage}. This is particularly critical in domains 
+where machine learning is applied to sensitive tasks, such as public 
+health or policy \citep{rahm2000data}.
 
@@
 Building upon the insights gained from our case study of the 
-NYC 311 Service Request (SR) data, we propose a set of data 
+NYC 311 SR data, we propose a set of data 
 curation principles tailored for government-released open datasets. 

### 2024-09-27 manuscript/manuscript.tex
Manuscript update
Removing the computed fields reference and analysis.

@@
 	\centering
-  	\includegraphics[width=0.7\textwidth]{10-year-trend_yearly.pdf}
+  	\includegraphics[width = \textwidth]{10-year-trend_yearly.pdf}
  	\caption{10-year (2014-2023) annual SR Counts.}
@@
 both routine city operations and extraordinary events.
-
-\jy{The plot sizes seem to be fixed in the generation code. That's why
-  the fonts are small when scaled. I'd generate them with approximately
-  the expected sizes in the manuscript. I'd also use aspect ratio
-  0.618 (golden ratio).
-}
 
@@
 	\centering
-	\includegraphics[width=0.7\textwidth]{SRs_by_Agency.pdf}
+	\includegraphics[width = \textwidth]{SRs_by_Agency.pdf}
   	\caption{SR counts by Agency with Cumulative Percentage}
@@
  \centering
-  \includegraphics[width=0.7\textwidth]{SR_by_Complaint_Type.pdf} 
+  \includegraphics[width = \textwidth]{SR_by_Complaint_Type.pdf} 
   \caption{Top 20 complaint types and Cumulative Percentage} 
@@
 
-\jy{Can we lighten the color of the bar to make the pertentages more
-  legible?}
-
 \begin{figure}[tbp]
@@
 \label{sec:zipcodesissues}
-All zip codes (two fields: \texttt{zip\_codes} and \texttt{incident\_zip}) 
+All zip codes ( \texttt{incident\_zip}) 
 should be valid as defined by the USPS database, which contains 
-37,946 valid zip codes. The \texttt{incident\_zip} field has 
+44,173 valid zip codes. The \texttt{incident\_zip} field has 
 only 0.07\% invalid entries, this still results in 4,163 errors; 
@@
 indicating a systemic problem.
-
-\begin{table}[tbp]
-    \centering
-    \caption{Comparison of Top Ten Zip Codes Lists}
-	    \begin{tabular}{@{}lS[table-format=6.0,round-mode=places,
-	    round-precision=0]c@{\hskip 0.5cm}@{}lS[table-format=6.0,
-	    round-mode=places,round-precision=0]c@{}}
-		\toprule
-	 	\multicolumn{3}{c}{\textbf{zip\_codes}} & \multicolumn{3}{c}{\textbf{incident\_zip}} \\
-	      \cmidrule(r){1-3} \cmidrule(l){4-6}
-	      \textbf{Zip Code} & \textbf{Count} & \textbf{Valid?} 
-	      & \textbf{Zip Code} & \textbf{Count} & \textbf{Valid?} \\
-	      \midrule
-	        11275 & 104556 & FALSE & 10466 & 104562 & TRUE \\
-	        12420 & 27503 & TRUE & 10023 & 27972 & TRUE \\
-	        12428 & 26564 & TRUE & 10031 & 25548 & TRUE \\
-	        10935 & 25508 & FALSE & 10457 & 25066 & TRUE \\
-	        10934 & 23448 & FALSE & 10453 & 24752 & TRUE \\
-	        10931 & 22381 & TRUE & 10456 & 24751 & TRUE \\
-	        10930 & 22121 & TRUE & 10452 & 22527 & TRUE \\
-	        17613 & 21963 & FALSE & 10025 & 21705 & TRUE \\
-	        10936 & 21707 & FALSE & 10458 & 21689 & TRUE \\
-	        11606 & 21435 & FALSE & 10032 & 20622 & TRUE \\
-	      \bottomrule
-	    	\end{tabular}
- 	\label{tab:zipcodes}
-\end{table}
-
-Consider a case study with the NYC Office of Nightlife (ONL). Suppose 
-that ONL wants to know, "What are the top 10 zip codes for Noise 
-Complaints over the last two years?" ONL seeks to assess the impact 
-of recent efforts promoting a vibrant nightlife in NYC while also 
-easing tensions between bar and club owners. Assume this analysis uses 
-the \texttt{zip\_codes} field, one of the computed fields known to have 
-validity issues. The analysis is then repeated using the more accurate 
-\texttt{incident\_zip} field. Finally, both analyses are validated 
-against the USPS zip code database. The top 10 zip codes and their 
-validation results are summarized in Table~\ref{tab:zipcodes}. Six out 
-of the ten \texttt{zip\_codes} entries are invalid, closely matching 
-the overall dataset's 58\% invalid rate. In contrast, the 
-\texttt{incident\_zip} field is fully valid, consistent with the 
-overall \texttt{incident\_zip} validation rate of 99\%.
-It is possible that the computed \texttt{zip\_codes} may mean the zip
-code of something differnent from the incident. Without proper
-documentation, it is confusing to have it in the released data.
-
-\paragraph{Police Precincts}
-\label{sec:police-precincts}
-A curious case arises when examining two nearly identical fields: 
-\texttt{police\_precincts} and \texttt{police\_precinct}. Both fields 
-are among the computed fields previously mentioned. Using the 
-\href{https://www.nyc.gov/site/nypd/bureaus/patrol/precincts-landing.page}
-{NYPD Precinct listings}, it is possible to determine the valid police 
-precincts. Upon validation, both \texttt{police\_precincts} and 
-\texttt{police\_precinct} contain about 35\% invalid entries. However, 
-the invalid entries are not identical. The \texttt{police\_precincts} 
-field has 2,171,864 invalid entries (35\%), while the 
-\texttt{police\_precinct} field has 2,171,778 invalid entries (also 
-35\%), but they do not match.
-
-\paragraph{Community Boards}
-\label{sec:communityboards}
-Community Boards are the most local, grassroots form of City government, 
-serving as a vital connection between communities, elected officials, 
-and City agencies. They often act as a measure of the equity of City 
-services across the five boroughs. In the 2022-2023 dataset, there are 
-27,276 invalid \texttt{community\_board} entries, representing 0.43\% 
-of non-blank data. The distribution of these invalid entries by Agency 
-is not consistent with the overall SR Agency 
-distribution. This suggests that issues may exist within specific key 
-Agencies, such as the Taxi \& Limousine Commission (TLC) and the 
-Department of Parks and Recreation (DPR).
-
-\paragraph{Community Districts}
-\label{sec:communitydistrict}
-Another of the computed fields is \texttt{community\_districts}. 
-Community Districts define the geographic boundaries for the Community 
-Boards but differ in that they are used by the Department of City 
-Planning (DCP) for environmental, socio-economic, and demographic 
-purposes. They represent a geographical division rather than a City 
-government division. Due to the format of the \texttt{community\_district} 
-data, it is not possible to directly establish validity. However, the 
-dataset contains 72 unique entries, while only 59 valid Community 
-Districts exist.
 
@@
 	 \centering
- 	 \includegraphics[width = 0.7\textwidth]{negative_duration_SR_violin.pdf}
+ 	 \includegraphics[width=\textwidth]{negative_duration_SR_violin.pdf}
  \caption{Distribution of Negative Durations}
@@
   \centering
-  \includegraphics[width = 0.7\textwidth]{post_closed_violin_chart.pdf}
+  \includegraphics[width=\textwidth]{post_closed_violin_chart.pdf}
   \caption{Post-closed resolution action update date(s) over 30 days.}
@@
 \end{figure}
-
-\jy{Are these from 2-year or 10-year period? Need to clarify in
-  Section 2.}
 
@@
 
-\textbf{borough \& borough\_boundaries:} These fields are 98.3\% matches. 
-This ``near-duplicate'' relationship creates a challenge in determining 
-which field is correct when discrepancies arise.
-
 \textbf{borough \& taxi\_company\_borough:} Despite their names, these 
@@
 uses.
-
-\textbf{incident\_zip \& zip\_codes:} The \texttt{zip\_codes} field, 
-one of the computed fields, was found to have 56\% invalid entries, 
-while the \texttt{incident\_zip} field has a 99.93\% accuracy rate. 
-We recommend eliminating or masking the \texttt{zip\_codes} field 
-until its accuracy can be improved.
-
-\textbf{police\_precinct \& police\_precincts:} These two computed 
-fields are 99.9\% matches. Since the computational process is unknown, 
-it is unclear which field is more accurate. However, having two fields 
-with near-duplicity suggests that only one is necessary.
 
@@
 By removing duplicate and ``near-duplicate'' fields, it is possible to 
-shrink the file size by 12.3\%, which for this large dataset equates to 
-a reduction of 395 Mb. A smaller file size means faster downloads, 
+shrink the file size by 39.5\%, which for this large dataset equates to 
+a reduction of 1.4 Gb, which is significant. A smaller file size means faster downloads, 
 less storage impact, and simpler data analysis efforts.
-Below is a list of duplicate and near-duplicate fields. There are 
-challenges with these proposed deletions, especially with the 
-``near-duplicate'' fields. For example, while the \texttt{park\_borough} 
-is a complete duplicate of the \texttt{borough} field, the 
-\texttt{borough\_boundary} field is only 98.3\% similar. Although losing 
-1.7\% of the data is small, it still amounts to 110,715 non-matching 
-occurrences. The authors propose deleting the following fields:
+Below is a list of duplicate and near-duplicate fields.
 
@@
 
-    \item \textbf{police\_precinct}: This field has a 99.95\% match with 
-    \texttt{police\_precincts}. Removing one would result in minimal data loss.
-
-    \item \textbf{borough\_boundaries (computed field)}: A 98.3\% match 
-    with the \texttt{borough} field.
-
     \item \textbf{cross\_street\_1 \& 2 and intersection\_street\_1 \& 2}: 
     An 88\% match between these pairs. We recommend deleting the 
-    \texttt{intersection\_street\_1 \& 2} fields while acknowledging some 
-    data loss.
-
-  \item \textbf{zip\_codes (computed field)}:
-    This field has a 58\% error rate and can lead to significant 
-    analytical errors. We recommend deleting this field.
+    \texttt{intersection\_street\_1 \& 2} fields while acknowledging the
+    concomitant data loss.
+
 \end{itemize}
@@
     present in the data but not included in the Data Dictionary need
-    to be documented.
+    to be documented and perhaps labelled as ``experimental'' or
+    ``not for official use''.
 \end{itemize}
@@
 harmonization of data fields across various agencies and 
-maintaining this consistency over time. As seen in the NYC 
-311 data, discrepancies in field names, formats, and definitions, 
-such as those between \texttt{police\_precinct} and 
-\texttt{police\_precincts}, can introduce significant 
-inconsistencies. To address this, agencies should adopt 
+maintaining this consistency over time. To address this, agencies should adopt 
 standardized naming conventions and formats across all datasets. 
@@
 open datasets. In the NYC 311 data, issues such as 
-invalid zip codes (e.g., 58\% invalid \texttt{zip\_codes} entries) and 
+invalid zip codes and 
 nonsensical date entries (e.g., negative durations) highlight the need 
@@
 and \texttt{park\_borough} could be consolidated without any data 
-loss, saving space and reducing file size by up to 12.3\%. Another 
+loss, saving space and reducing file size. Another 
 example is the \texttt{taxi\_company\_borough} field, which is 99.94\% 
@@
 actual data, such as missing field descriptions for computed fields. 
-Updating the Data Dictionary to include descriptions of all fields, 
-especially computed fields like \texttt{@computed\_region\_xxxx}, 
+As mentioned, updating the Data Dictionary to include descriptions of all fields, 
+including the computed fields, e.g. \texttt{@computed\_region\_xxxx}, 
 would improve transparency and usability. Regular updates should 
@@
 from entering the dataset. Automated systems can be designed to 
-flag inconsistencies, incorrect values, or illogical sequences, such 
-as invalid \texttt{incident\_zip} values or illogical date entries. 
+flag inconsistencies, incorrect values, or illogical sequences, or 
+illogical date entries. 
 Automation can also facilitate faster detection and correction of 
@@
 understandable. This involves providing clear metadata and making 
-datasets available in user-friendly formats. For example, in the NYC 
-311 dataset, fields like \texttt{@computed\_region\_xxxx\_xxxx} and 
-the distinction between \texttt{incident\_zip} and \texttt{zip\_codes} 
-can be confusing for non-expert users. These fields should be clearly 
-documented, explaining how they are calculated and how they differ. 
-Additionally, governments should communicate transparently about 
+datasets available in user-friendly formats. Additionally, 
+governments should communicate transparently about 
 data quality issues, such as notifying the public when errors in SR 
@@
 this paper are available [FixMe].
-\jy{need to think: we need to keep a copy of the downloaded data;
-  if the data portal changes and our code may not work.}
 

### 2024-09-27 code/timelinecharts..R

@@ -541,11 +541,11 @@ cat("\nDay-of-the-Year Summary:")
 # Convert day_of_year to character
 day_counts_df$day_of_year <- as.character(day_counts_df$day_of_year)
 
-# Format the columns for better alignment
+# # Format the columns for better alignment
 formatted_day_counts_df <- data.frame(
-  day_of_year = format(day_counts_df$day_of_year, width = max(nchar(day_counts_df$day_of_year))),
-  count = format(day_counts_df$count, width = max(nchar(day_counts_df$count)))
-)
+   day_of_year = format(day_counts_df$day_of_year, width = max(nchar(day_counts_df$day_of_year))),
+   count = format(day_counts_df$count, width = max(nchar(day_counts_df$count)))
+ )
 
 day_counts_df <- day_counts_df[order(-day_counts_df$count), ]
 
@@ -941,7 +941,7 @@ SR_created_time_of_day <- create_bar_chart(
 # Overall closed time-of-day summary (0900, 1000, 1100, 1200, 1300, etc.)
 max_hour_of_the_day <- closed_hour_of_day_df[which.max(closed_hour_of_day_df$count), ]
 max_hour_of_the_day$closed_hour <- as.character(max_hour_of_the_day$closed_hour)
-max_count <- max_hour_of_the_day$count
+max_count <- formatted_day_counts_df- max_hour_of_the_day$count
 cat(
   "\nHour of the day with maximum 'closed' count is hour #:", max_hour_of_the_day$closed_hour,
   "with", format(max_hour_of_the_day$count, big.mark = ","), "SRs"

### 2024-09-28 manuscript/manuscript.tex
comments on presentations

@@
 \end{figure}
+
+\jy{If the 10-year data was only used for Figure~\ref{fig:10-yr}, we
+  could remove this figure and just describe it in text. This would
+  avoid the confusion of which dataset was used in the analyses in the
+  sequel.}
+
+\jy{If we keep the 10-year data, we might as well make some dynamic
+  visualization over time. For example, Figure 2/3 could have multiple
+  panels such that we can see several 2-year decompositions over
+  time.}
 
@@
 \end{figure}
+
+\jy{Currently we have 9 figures and 3 tables, which are too many for
+  an article. Some journal (e.g., Biometrics) requires no more than 6
+  tables/figures combines. We need to more selective in presenting
+  figures/tables, perhaps no more than 8 combined?
+}
 
@@
 \jy{This is caused by reading the number in as double
-  precision. Single precision would be good.}
+  precision. Single precision would be good. The raw csv file only
+  have 7 digits after the decimal point, e.g.,
+  $(40.769737, -73.91244)$.
+  }
 
@@
 uses.
+\jy{The taxi\_company\_borough may mean the borough of the location of
+  the taxi company, which can be irrelevant to the borough where the
+  incident was reported.. Such information could be useful only when
+  studying TLC related SRs. It'd be better to use a second table to
+  store such TLC dependent information.}
 
@@
 CSV file.
+
+\jy{In my Chance article, we found an agency
+  `DEPARTMENT OF CONSUMER AND WORKER PROTECTION', which occurred 361
+  times in May 2023. It was the only one that does not look like an
+  acronym. Did this occur in the two-year data?
+  I wonder if this has been fixed lately.}
+
+\jy{A related question is whether the data quality has improved over
+  the years}
 

### 2024-10-16 manuscript/manuscript.tex
New images. Review of about one-half of manuscript
Minor code changes. Updated charts in the image directory.  Editing Manuscript.

--- 2024-09-27 manuscript/manuscript.tex
@@
 \usepackage{booktabs, textgreek}
+\usepackage{subcaption}
 
@@
 an overview of the history of the NYC 311 SR system and presents 
-a summary of SR counts over the 10-year period from 2014 to 2023. 
+a summary of SR counts over a 2-year period from 2022 to 2023. 
 Section~\ref{sec:issues} delves into specific data cleansing 
@@
 and the presence of missing, blank, or N/A entries. This section
-utilizes 311 SR data over a a 2-year period from 2022-2023.  It 
  investigates field compliance with reference or acceptable values, 
@@
 the COVID-19 pandemic, the 2021 expansion to include the MTA's city
-subway system, and a record 3.23 million service requests in
-2023.
+subway system (the biggest expansion in their history), and a 
+record 3.23 million service requests in 2023.
 
@@
 third-party users. These challenges are addressed by various agency
-open data managers and 
+open data managers and OTI.
 \end{comment}
@@
 
-\begin{figure}[tbp]
-	\centering
-  	\includegraphics[width = \textwidth]{10-year-trend_yearly.pdf}
- 	\caption{10-year (2014-2023) annual SR Counts.}
-  	\label{fig:10-yr}
-\end{figure}
-
-Our investigation utilizes two 311 SR datasets covering 10-years and 2-years.
-The 10-year period (2014--2023), downloaded on 15 September 2024, 
-contains only three columns (created\_date, closed\_date, and 
-agency). This dataset is used to analyze long-term 
-trends and observe certain dating anomalies. The 10-year dataset is approximately 
-1.4 GB in size and contains over 27 million rows, each row representing one SR.
- Instructions on downloading this file from the NYC Open Data Portal 
- are available in the supplemental material.
-
-Additionally, this investigation uses a 2-year dataset covering the period
-2022--2023, also downloaded on 15 September 2024. This data set 
-contains the full 41 columns of 311 SR data available from the portal, 
-and is used to conduct detailed analysis in areas such as data consistency, 
-data validity, redundant fields, etc.. This 2-year dataset 
-is quite large at 3.7GB. It contains over 6 million rows of SRs. Instructions 
-on downloading this file from the NYC Open Data Portal are 
-available in the supplemental material.
-
-Figure~\ref{fig:10-yr} shows  annual counts of 311 SRs over the
-10-year period from 2014 to 2023, revealing a clear upward trend in
-public engagement. As indicated, the most significant increase
-in activity occurred in 2020, when the number of requests surged to 3.23 million
-during the COVID-19 pandemic, illustrating the essential role of the
-311 system in supporting residents during times of crisis. Overall, SR
-counts increased by 49.5\% between 2014 and 2023, reflecting
-growing public reliance on the system. (During this 10-year time 
-period, the population of NYC grew just 0.6\%.) The steady 
-rise in SRs can be attributed to the increased accessibility of 
-the 311 system via online and mobile platforms, as well as 
-heightened public awareness  and utilization of the service. 
+Our investigation utilizes an SR dataset covering  2-years (2022-2023).
+This dataset is approximately 3.7 GB in size and contains over 
+6.4 million rows with each row representing one SR. It contains 
+41 columns of data. It is used to conduct detailed analysis 
+in areas such as data consistency, data validity, redundant fields, 
+etc.. Instructions on downloading this file from the 
+NYC Open Data Portal are available in the supplemental material.
+
+During the 10-year period from 2014 to 2023,  311 experienced
+a 50\% in Service Requests, , reflecting a growing public reliance 
+on the system. (As a note, during this 10-year timeframe, the 
+population of NYC grew just 0.6\%.). The most significant increase
+in activity occurred in 2020, when the number of requests 
+surged to 3.23 million during the COVID-19 pandemic, 
+illustrating the essential role of the 311 system in supporting 
+NYC residents during times of crisis. The steady 
+rise in SRs can be at least partially attributed to the increased 
+accessibility of the 311 system via online and mobile 
+platforms, as well as heightened public awareness of the service. 
 While the spike in 2020 was exceptional, the broader trend 
-indicates sustained growth in service request volume over 
-the decade, highlighting the system's evolving role in managing 
+indicates sustained growth in SR volume in future years,  
+highlighting the system's evolving role in managing 
 both routine city operations and extraordinary events.
@@
 agency, showing the cumulative percentage of SRs handled by each
-agency over a the 2-year time frame. Note that the
-distribution of SRs is concentrated among a few key agencies. The six
-largest agencies are New York Police Department (NYPD), 
-Housing Preservation and Development (HPD),
-New York City Department of Sanitation (DSNY), Department of
-Transportation (DOT), Department of Environmental Protection (DEP),
-and Department of Parks and Recreation (DPR).
-Collectively they handle over 90\% of the total SR volume. This is
-indicative of the critical roles these agencies play in managing
+agency over a the 2-year timeframe. Note that the
+distribution of SRs is heavily concentrated among a few key 
+agencies. The six largest agencies are:
+
+\begin{itemize}
+    \item New York Police Department (NYPD) - 43\%
+    \item Housing Preservation and Development (HPD) - 21\%
+    \item New York City Department of Sanitation (DSNY) - 10\%
+    \item Department of Transportation (DOT) - 7\%
+    \item Department of Environmental Protection (DEP) - 5\%
+    \item Department of Parks and Recreation (DPR) - 4\%
+\end{itemize}
+
+Collectively these six City agnecies handle over 90\% of the total SR 
+volume, indicative of the critical role these agencies play in managing
 public concerns, ranging from noise complaints and housing issues to
 sanitation and transportation problems. The NYPD alone accounts for a
-substantial portion of the total SRs, underscoring its central role in
-addressing complaints related to public safety. The
-remaining 10\% is distributed across a number of other
-agencies. This concentration of SRs points to a
-reliance on a few key departments to resolve most non-emergency
-issues (e.g. non 911 calls). Accordingly, this underscores 
-the necessity for optimal data handling and curation processes 
+substantial portion of the total SRs, underscoring its leading role in
+addressing complaints related to public safety. The remaining
+10\% of SRs is distributed across 10 other
+agencies. This concentration of SRs to the `big six'' agencies 
+underscores the necessity for optimal data handling and curation processes 
 within these high-volume agencies in order to ensure 
@@
 Figure~\ref{fig:SR_complaints}
-illustrates the top 20 complaint types, accounting for 70\% 
-of all SRs and their cumulative percentage. Note that the 
-distribution of complaints is skewed, with a small number of issues 
-dominating total SR volume. Noise-related complaints, including 
-residential, commercial, and street noise, make up 22\% of all 
-complaints, making them the most frequent issue in the NYC 311 system. 
+illustrates how the top 20 complaint types, accounting for 70\% 
+of all SRs. Note that the distribution of complaints is skewed, 
+with a small number of issues dominating total SR volume. 
+Noise-related complaints, including residential, commercial, and 
+street noise, when combined make up 22\% of all 
+complaints; the most frequent issue in the NYC 311 system. 
 Other prominent categories include illegal parking, heat/hot water 
 complaints, and blocked driveways. The cumulative percentage curve 
-shows that after the top 20, the remaining complaint types are spread 
-thinly across the remaining 30\% of SRs. This concentration suggests 
-that improving responses to the most frequent complaints could have 
-a significant impact on overall service efficiency and resident 
-satisfaction. It also provides insights into the operational 
-pressures faced by the city agencies responsible for handling these 
-high-volume complaints.
+reveals that after the top 20 complaint types, the remaining 
+complaint types are spread thinly across the remaining 30\% of SRs,
+suggesting that improving responses to the most frequent 
+complaints could have an outsized impact on overall service 
+efficiency and resident satisfaction. This chart also provides 
+insights into the operational pressures faced by City agencies 
+responsible for handling these high-volume complaints.
 
@@
 completeness, and consistency. Without thorough data cleansing, the utility
-of open data can be quite limited, affecting its reliability for
-research, policy-making, and innovation. The main purpose of cleansing
-open data is to ensure that it is accurate, consistent, and usable
-across multiple platforms and by various stakeholders. This improves
-the trustworthiness of the data and enables better decision-making,
-more accurate analysis, and the integration of data into machine
+of open data can be limited and perhaps untrustworhty, affecting 
+its reliability for research, policy-making, and innovation. The 
+main purpose of cleansing open data is to ensure that it is 
+accurate, consistent, and usable across multiple platforms and by 
+various stakeholders with a myriad of purpose. Data cleansing improves
+the trustworthiness of the data and enables more accurate analysis, 
+better decision-making,and the integration of data into machine
 learning models or other systems.
@@
 Additionally, certain fields must conform to specific data types, 
-such as numeric, character, or date formats, which are typically 
-outlined in a Data Dictionary. Another essential aspect of validation 
+such as numeric, character, or date formats; typically 
+outlined in a Data Dictionary. An essential aspect of validation 
 is domain compliance, where fields must adhere to a predefined set of 
-values, such as statuses, state names, zip codes, or gender. Structural 
+values, such as statuses, state names, or zip codes. Structural 
 errors also play a significant role in data quality, particularly when 
@@
 Here we identify the presence of issues in the NYC 311 SR open data;
-not attempting to solve them. A solution effort should
-be undertaken only after an investigation as to the why and how 
-such issues came about, and a discussion as to whether or not it 
-is an actual \textit{error} or holds some other status.
+not attempting to solve them but rather to highlight their scope and
+magnitued. A solution effort should be undertaken only after further 
+investigation as to the why and how such issues came about, 
+and a discussion as to whether or not it is an actual \textit{error} or 
+holds some other status. Coordination with the originating Agency
+is paramount.
 
@@
 
-	\item There are four date fields (created, closed, updated, due).
+	\item There are four date fields (created, closed, action\_updated, due).
 
@@
 	\item In addition to the incident\_address, there are five additional location fields: 
-	lat/long, street\_name, landmark, NY State plane, and Block \#
-
-	\item One free-form text field, resolution\_description, which 
+	latitude/longitude, street\_name, landmark, USGS State Plane 
+	Coordinate System, and Block \#.
+
+	\item A free-form text field, resolution\_description, which 
 	supports 934 characters of input, including commas and special characters
@@
 visible in the portal's ``Column manager'' widget, labeled with the prefix 
-``@computed\_region.'' The additional fields include \texttt{zip\_codes}, 
-\texttt{community\_districts}, \texttt{borough\_boundaries}, 
-\texttt{city\_council\_districts}, \texttt{police\_precincts}, and 
-\texttt{police\_precinct}. While one can likely infer the meaning and 
+``@computed\_region.'' The additional fields include:
+
+\begin{itemize}
+    \item \texttt{zip\_codes}
+    \item \texttt{community\_districts}
+    \item \texttt{borough\_boundaries}
+    \item \texttt{city\_council\_districts}
+    \item \texttt{police\_precincts}
+    \item \texttt{police\_precinct}
+\end{itemize}
+
+While one can likely infer the meaning and 
 derivation of these computed fields, uncertainty remains as to
 their derivation and usage. These six fields are not included in
-the Data Dictionary. Our analysis has shown that these six 
-fields have numerous data validity issues. Liaison with the 
-NYC Open Data Teams leads us to treat these six fields 
+the Data Dictionary, which we have asked to be updated. Our 
+analysis shows that these six \texttt{computed} fields have 
+significant data validity issues, leading us to treat these six fields 
 as experimental and not for suitable for normal use. Accordingly, 
@@
 \texttt{due\_date}, you would be challenged as 99.6\% of the
-\texttt{due\_date} field is
-blank. When counting fields for blank or N/A values, they appear 
-to divide into three groups:
-\textbf{Mostly Empty}, \textbf{Partially Empty}, 
-and \textbf{Few/None Empty}. 
-The Mostly Empty category ranges from 93--99.9\% blank 
-and includes such fields as
+\texttt{due\_date} field is blank. When counting fields for 
+blank or N/A values, they appear to divide into three groups:
+
+\begin{itemize}
+    \item \textbf{Mostly Empty} -- 93-99.9\% blank 
+    \item \textbf{Partially Empty} -- 40-4\% blank
+    \item \textbf{Few/None Empty} -- 2-0\% blank
+\end{itemize}
+
+The Mostly Empty category includes such fields as
 \texttt{taxi\_company due\_date},
@@
 \texttt{created\_date}, \texttt{complaint\_type},
-\texttt{agency}, and \texttt{status}. In some cases it may make
-sense to inquiry as to why some fields are frequently blank.
+\texttt{agency}, and \texttt{status}. In some analysis efforts, it may be
+prudent to inquire as to why some fields are mostly or almost always blank.
 Figure~\ref{fig:blank_fields} presents a graphic 
@@
 Any analytic effort must ensure that fields containing invalid values 
-are identified and removed from the analysis. Below are the results of 
-selected field validation. The latitude and longitude fields were 
+are identified and isolated from the analysis. Below are the results of 
+selected field validations. The latitude and longitude fields were 
 found to fall within the geographic boundaries of New York City, and 
@@
 Unfortunately, the Data Dictionary specifies very few domains of 
-acceptable values. However, the following fields were tested and found 
+acceptable values. Nonetheless, the following fields were tested and found 
 to comply with their expected domains, as determined by public usage 
-and historical datasets: \texttt{address\_type}, \texttt{status}, 
-\texttt{borough}, \texttt{borough\_boundaries}, \texttt{park\_borough}, 
-\texttt{data\_channel}, \texttt{vehicle\_type}, and 
-\texttt{city\_council\_district}. Some fields proved to be problematic 
-as to their compliance with a domain of legal values.
+and historical datasets:
+
+\begin{itemize}
+    \item \texttt{address\_type}
+    \item \texttt{status}
+    \item \texttt{borough}
+    \item \texttt{borough\_boundaries}
+    \item \texttt{park\_borough}
+    \item \texttt{data\_channel}
+    \item \texttt{vehicle\_type}
+    \item \texttt{city\_council\_district}
+\end{itemize}
+
+But some fields proved problematic as to their 
+compliance with a domain of allowable values.
 
@@
 All zip codes ( \texttt{incident\_zip}) 
-should be valid as defined by the USPS database, which contains 
+should validate against the USPS database, which contains 
 44,173 valid zip codes. The \texttt{incident\_zip} field has 
-only 0.07\% invalid entries, this still results in 4,163 errors; 
-not insignificant. If you sort the invalid entries by Agency, 
-the distribution by percentage mirrors the \texttt{overall} 
+only 0.07\% invalid entries, this still results in 4,374 errors; 
+not insignificant. If you group the invalid entries by 
+percentage by Agency, it mirrors the \texttt{overall} 
 breakdown of SRs by Agency, potentially 
@@
 
-\paragraph{Date and Closed Date}
+\paragraph{Created and Closed Date}
 \label{sec:negativeduration}
@@
 Unfortunately, this does not seem to be the case, as several anomalies 
-exist in the date fields:
+exist in the created and closed date fields. Here are some of the errors:
+
 \begin{itemize}
     \item SRs with a \texttt{closed\_date} that occurs before the 
-    \texttt{created\_date}.
+    \texttt{created\_date} indicating a negative duration.
     \item \texttt{created\_date(s)} and \texttt{closed\_date(s)} in 
-    the far distant past.
+    the far distant past, e.g. 1900.
     \item \texttt{created\_date(s)} and \texttt{closed\_date(s)} that 
-    match to the second.
+    match to the second indicating a zero duration.
     \item A large number of SRs closed and/or created exactly at midnight 
@@
 it can be easily computed as the difference between
-\texttt{(closed\_date} and \texttt{created\_date}.
-There are 12,251 SRs where the \texttt{closed\_date} precedes the 
+\texttt{(closed\_date} and \texttt{created\_date}.  Duration is one of 
+most used fields that is constantly analyzed, such as determining if 
+one NYC borough receives faster response for ceratin 
+complaints. There are 12,450 SRs where the \texttt{closed\_date} precedes the 
 \texttt{created\_date}, generating nonsensical ``negative durations.'' 
 Though this represents only 0.2\% of the dataset, these errors can 
-significantly impact response time analyses.
-Eight SRs with extremely large negative durations ($-$730 days), all 
-originating from the Department of Homeless Services (DHS), contain 
-an entry of ``1900-01-01'' as the \texttt{closed\_date}. This results 
-in negative durations exceeding $-$44,601 days (or 122 years). Such 
+significantly impact response time analyses. Eight SRs with extremely 
+large negative durations ($-$730 days), all originating from the 
+Department of Homeless Services (DHS); all contain an entry 
+of ``1900-01-01'' as the \texttt{closed\_date}. This results 
+in negative durations exceeding $-$44,601 days (or 122 years). Many
+other SRs have negative durations in excess of 300+ days. Such 
 anomalies, though rare, can significantly skew statistical results if 
 not addressed during data cleansing. As a result, these SR rows are 
-removed from the analysis.
+removed from the analysis. Here is a sample:
 
@@
 
+Excluding the extreme negative values, 
+Figure~\ref{fig:negative-duration-violin} shows the broad spread of 
+negative-duration SRs. While there are few outliers, the magnitude 
+of the negative durations is troubling and can produce bizarre
+and unwanted analytical results.
+
 \begin{figure}[tbp]
@@
 
-Excluding the extreme negative values, 
-Figure~\ref{fig:negative-duration-violin} shows the broad spread of 
-negative-duration SRs. While there are few outliers, the magnitude 
-of the negative durations is troubling and can produce bizarre 
-analytical results. Table~\ref{tab:largest-errors} summarizes the 
+Table~\ref{tab:largest-errors} summarizes the 
 negative durations of the largest magnitude (in days). The data 
@@
 
-% For illustration, consider a case study of homeless person
-% assistance. Suppose that DHS wants to know ``How 
-% quickly have 311 SRs for ``Homeless Person Assistance'' been resolved 
-% over the past two years?'' This is a typical request made by 
-% both the public and City government; analogous to ``How quickly 
-% is my Agency responding to ``critical'' requests, and does 
-% that performance vary by Borough, Zip Code, etc. To conduct such an 
-% analysis, we can use the tools on the Open Data Portal. Here are the steps:
-
-% \begin{itemize}
-%     \item Select created\_date from 2022-2023 and filter by complaint\_type = 
-%     ``Homeless Person Assistance'' (this yields 55,000 SRs)
-
-%     \item Compute the ``duration'' (closed\_date – created\_date)
-
-%     \item Take an average of the ``duration'' field  which yields: \textbf{Answer:  -4.8 days}  
-% \end{itemize}
-
-% Clearly that answer is nonsensical. How did such a simple task result 
-% in an absurd answer? The answer lies in the computation of the ``duration'' 
-% field. It turns out, there are eight DHS SRs that have a closed\_date 
-% of ``1900-01-01''. Each of those SRs creates a negative duration of -44,602 
-% days (-122 years). Those eight SRs are enough to drive the 
-% average duration of 55,000 Homeless Assistance SRs to a negative 
-% value; clearly incorrect. In this case, the median is a better measure 
-% of central tendency. The median is  0.2 days (approx 5 hrs) which will
-% surely please the DHS.
-
-%\begin{figure}[tbp]
-% 	 \centering
- %	 \includegraphics[width = \textwidth]{homeless_response_time_clean.pdf}
-%	 \caption{Homeless Assistance SR Durations}
-%	 \label{fig:homeless}
-%end{figure}
-
 \subparagraph{Resolution Action Update Date}
+
+A similar issue occurs when resolution action update date is after the
+closing date.  When an SR is updated, the 311 software 
+automatically populates the \texttt{resolution\_action\_update\_date}. 
+While it is certainly possible to update an SR after it is closed, 
+some of these updates seem to be well beyond an expected 
+range, some over 2 years later. As shown in 
+Figure~\ref{fig:resolution-violin}, there are 10,265 SRs that 
+were updated more than 30 days after the 
+\texttt{closed\_date}, but less than 730 days, cutoff used to 
+exclude infeasible dates such as 1900-01-01. The chart 
+highlights the distribution of these post-closure updates, with 
+a notable concentration of updates within 
+the 30-to-90-day range. This pattern raises questions about whether 
+such delayed updates are standard operational practice or indicative 
+of a potential issue requiring further investigation.
 
@@
 
-A similar issue occurs when resolution action update date is after the
-closing date. \jy{It might be possible, I think, if they allow updates
-  after a SR is closed.}
-When an SR is updated, the 311 software automatically populates the 
-\texttt{resolution\_action\_update\_date}. Some of these updates 
-occur long after the SR is closed. As shown in Figure~\ref{fig:resolution-violin}, 
-there are 7,460 SRs that were updated more than 30 days after the 
-\texttt{closed\_date}, but less than 730 days, to exclude infeasible 
-dates like 1900-01-01. The chart highlights the distribution of these 
-post-closure updates, with a notable concentration of updates within 
-the 30-to-90-day range. This pattern raises questions about whether 
-such delayed updates are standard operational practice or indicative 
-of a potential issue requiring further investigation. While the reason 
-for these delayed updates is unclear, they may impact the accuracy of 
-response time analyses if not properly accounted for.
-
-\subparagraph{Same Created Date \& Closed Date: Zero Durations}
+\subparagraph{Identical Created Date \& Closed Date: Zero Durations}
 A more prevalent issue occurs when the \texttt{closed\_date} and 
 \texttt{created\_date} are exactly the same, down to the second. This 
-creates a \textbf{zero duration}, which is also nonsensical. There 
-are 191,141 such SRs, representing 3.1\% of all non-blank data. 
+creates a \textbf{zero duration}, which is nonsensical. There 
+are 163,720 such SRs, representing 2.6\% of all non-blank data. 
 Ninety-nine percent of these zero-duration SRs occur in five agencies:
-DHMH, DOT, DOB, DSNY, and DEP.. 
-This distribution does not mirror the overall SR distribution by 
-agency, suggesting an agency-specific issue.
+DHMH, DOT, DOB, DSNY, and DEP. This distribution does not 
+mirror the overall SR distribution by Agency, suggesting 
+an agency-specific issue.
 
@@
 created or closed at exactly midnight (00:00:00) or noon (12:00:00), 
-down to the second. Normally, SR creation and closure follow the 
+down to the second. Normally, SR creation and closure follows the 
 work-day schedule, with most SRs created during daylight hours and 
 fewer at night or in the early morning. However, there is a 
-significantly higher number of SRs closed exactly at midnight and noon, 
-as well as a large number created exactly at noon. Specific observations 
-include:
-\begin{itemize}
-    \item 99,779 SRs were created exactly at noon (12:00:00).
-    \item 235,347 SRs were closed exactly at midnight (00:00:00).
-    \item 105,505 SRs were closed exactly at noon.
-\end{itemize}
-
-A unique way to visualize this anomaly is to examine the busiest day 
-during this 2-year period (Friday, 2023-09-29). Then aggregate 
-the created\_date(s) by minute (with seconds equal to zero), providing
-a minute-by-minute look at SR creation on that one day. Note the spike 
-at exactly noon (12:00), well beyond the 3\textsigma line.
+significantly higher number of SRs closed exactly at midnight, 
+as well as a large number created exactly at midnight.
 
@@
     \centering
-    \begin{minipage}{0.5\textwidth}
+    \begin{subfigure}{\textwidth}
         \centering
-        \includegraphics[width=\textwidth]{2-year-trend_SR_created_by_minute_of_busiest_day.pdf}
-        \caption{SRs Created Minute-by-Minute on Busiest Day}
+        \includegraphics[width=\textwidth]{2-year-trend_SRs_created_on_the_hour.pdf}
+        \caption{SRs Created by Hour of the Day}
         \label{fig:busiestcreated}
-    \end{minipage}%
-    \begin{minipage}{0.5\textwidth}
+    \end{subfigure}
+    \par\medskip
+    \begin{subfigure}{\textwidth}
         \centering
-        \includegraphics[width=\textwidth]{2-year-trend-SR_closed_by_minute_of_busiest_day.pdf}
-        \caption{SRs Closed Minute-by-Minute on Busiest Day}
+        \includegraphics[width=\textwidth]{2-year-trend_SRs_closed_on_the_hour.pdf}
+        \caption{SRs Closed by Hour of the Day}
         \label{fig:busiestclosed}
-    \end{minipage}
+    \end{subfigure}
+    \caption{Comparison of SRs Created and Closed by Hour of the Day}
+    \label{fig:stacked}
 \end{figure}
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=0.7\textwidth]{2-year-trend_SR_created_by_minute_of_busiest_day.pdf}
-  \includegraphics[width=0.7\textwidth]{2-year-trend-SR_closed_by_minute_of_busiest_day.pdf}
-  \caption{SRs Created and closed minute-by-minute on busiest day
-    during two year period \jy{give years}.}
-  \label{fig:busiest}
-\end{figure}
-
-Figure~\ref{fig:busiest} visualizes this anomaly by examining the 
-busiest day during this 2-year period (Friday, 2023-09-29). The 
-\texttt{created\_date} values are aggregated by minute (with seconds 
-equal to zero), providing a minute-by-minute look at SR creation on 
-that day. Notably, there is a spike exactly at noon (12:00), well 
-beyond the 3\textsigma{} line. \jy{May not want to mention 3 sigma as
-  here we have highly skewed count data.}
-The same analysis was repeated for 
-\texttt{closed\_date} values on the busiest day, aggregated by minute 
-(with seconds equal to zero). Spikes were observed at both 00:00:00 
-and 12:00:00.
 
@@
 
-\jy{This is caused by reading the number in as double
-  precision. Single precision would be good.}
-
 An question of precision vs. accuracy arises with the Latitude 
@@
 a 14-decimal number, e.g. 40.86769186022511. Given 
-that 1 degree of latitude at the equator is equal to 111.044736 
+that 1 degree of latitude at the equator is equal to 111.04 
 kilometers, the ``1'' at the end of that number represents 
 approximately 1.1 nanometers (1/1,000,000,000 of a meter). For 
-reference a DNA molecule is approximately 2nm in width. Clearly 
-the representation of the Latitude and Longitude fields is a 
+reference a DNA molecule is approximately 2nm in width. The 
+representation of the Latitude and Longitude fields appears to be a 
 classic case of 14-digit precision, but limited accuracy.
+
+\dt{STOPPED EDITING HERE}
 

### 2024-10-16 code/datacleansing.R

@@ -163,10 +163,6 @@ d311[, columns_to_upper] <- lapply(d311[, columns_to_upper], toupper)
 cat("\n\n**********DATA SUMMARY**********\n")
 
 #########################################################################
-# Consolidate Agency names
-d311 <- consolidate_agencies((d311))
-
-#########################################################################
 # Filter out rows with NA values in the created_date column
 d311 <- d311[!is.na(d311$created_date), ]
 
@@ -297,26 +293,33 @@ missingDataPerColumn <- missingDataPerColumn %>%
   mutate(field = reorder(field, -total_empty))
 
 # Create the bar chart with vertical X-axis labels
-blank_chart <- ggplot(missingDataPerColumn, aes(x = reorder(field, -total_empty), y = total_empty)) +
+blank_chart <- ggplot(
+  
+  missingDataPerColumn, aes(x = reorder(field, -total_empty), y = total_empty)) +
+  
   geom_bar(stat = "identity", fill = "#009E73") +
+  
   theme(
-    axis.title.x = element_text(vjust = 0, size = 11),
-    axis.title.y = element_text(vjust = 1, size = 11),
+    axis.title.x = element_text(vjust = 0, size = 8),
+    axis.title.y = element_text(vjust = 1, size = 8),
     plot.title = element_text(hjust = 0.5, size = 13),
-    plot.subtitle = element_text(size = 9),
+    plot.subtitle = element_text(size = 7),
     panel.background = element_rect(fill = "gray95", color = "gray95"),
     axis.text.x = element_text(angle = 50, vjust = 1, hjust = 1, face = "bold"),
     axis.text.y = element_text(face = "bold"),
     legend.position = "none", # Remove all legends
     aspect.ratio = 0.618033 # Set aspect ratio
   ) +
+  
   geom_text(aes(
[…]

### 2024-10-16 code/functions/archive/create_boxplot.R

@@ -1,0 +1,32 @@
+# #########################################################################
+create_boxplot <- function(
+    dataset,
+    x_axis_title,
+    chart_title,
+    chart_file_name) {
+  
+  boxplot_chart <- ggplot(
+    data = dataset, 
+    
+    aes(x = duration, y = factor(1))) +
+    
+    geom_jitter(color = "#0072B2", alpha = 0.4, size = 1.9, shape = 17) +
+    
+    geom_boxplot(width = 0.2, fill = "#E69F00", alpha = 0.7, color = "black") +
+    
+    theme(
+      axis.text.x = element_text( face = "bold", size = 8 ),
+      axis.text.y = element_text(face = "bold", size = 8),
+      legend.position = "none", 
+      plot.title = element_text(hjust = 0.5),
+      plot.subtitle = element_text(size = 7)
+    ) +
+    
+    labs(
+      title = chart_title, x = x_axis_title, y = "",
+      subtitle = paste("(", earliest_title, "--", latest_title, ")", " count=", nrow(dataset), sep = "")
+    )
+  suppressMessages(print(boxplot_chart))
+}
+
+#########################################################################

### 2024-10-16 code/functions/base_bar_chart.R

@@ -47,9 +47,9 @@ base_bar_chart <- function(dataset, x_col, y_col, chart_title, sub_title,
     theme(
       axis.title = element_blank(),  # Remove x and y axis titles
       plot.title = element_text(hjust = 0.5, size = 12),
-      plot.subtitle = element_text(size = 9),
-      axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1, face = "bold"),
-      axis.text.y = element_text(face = "bold"),
+      plot.subtitle = element_text(size = 7),
+      axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1, face = "bold", size = 8),
+      axis.text.y = element_text(face = "bold", size = 8),
       aspect.ratio = 0.618033
     ) +
     

### 2024-10-16 code/functions/create_combo_chart.R

@@ -46,32 +46,39 @@ create_combo_chart <- function(
   result <- calculate_values(max_count)
   starting_value <- result$starting_value
   increment <- result$increment
-  
+
   # Step 1: Create the initial plot
   combo_chart <- ggplot(summary_df) +
+    
     geom_bar(aes(x = agency, y = count), stat = "identity", fill = "#E69F00", width = 0.55) +
+    
     geom_text(aes(label = count, x = agency, y = count), 
               colour = "black", hjust = 0.5, vjust = -0.5, size = 2.5) +
+    
     geom_text(aes(label = round(cumulative_percentage, 2), x = agency, y = max_count * cumulative_percentage),
               size = 3, colour = "black", hjust = 0.5, vjust = 1.7) +
+    
     scale_y_continuous(
       breaks = seq(starting_value, max_count, by = increment), 
-      sec.axis = sec_axis(~ . / max_count, name = "Cumulative Percentage")
+      sec.axis = sec_axis(~ . / max_count )
     ) +
+    
     ggtitle(chart_title, subtitle = paste("(", earliest_title, "--", latest_title, ")",
-                                          " total=", format(total_count, big.mark = ","), sep = "")) +
-    labs(x = NULL, y = "Count") +
+                                          " n=", format(total_count, big.mark = ","), sep = "")) +
+    labs(x = NULL, y = NULL) +
+    
     theme(
-      axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1, face = "bold", size = 7),
-      axis.text.y = element_text(face = "bold", size = 7),
-      axis.text.y.right = element_text(color = "black", face = "bold", size = 7),
+      axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1, face = "bold", size = 8),
+      axis.text.y = element_text(face = "bold", size = 8),
+      axis.text.y.right = element_text(color = "black", face = "bold", size = 8),
       plot.title = element_text(hjust = 0.5, size = 12),
-      plot.subtitle = element_text(size = 6),
+      plot.subtitle = element_text(size = 7),
       aspect.ratio = 0.618033
     )
   
   # Only add the geom_line if there's more than one agency
   if (nrow(summary_df) > 1) {
+    
     combo_chart <- combo_chart + 
       geom_line(aes(x = agency, y = cumulative_percentage * max_count, group = 1),
[…]

### 2024-10-16 code/functions/create_violin_chart.R

@@ -6,9 +6,16 @@ create_violin_chart <- function(
     chart_title,
     chart_file_name) {
   
-  violin_chart <- ggplot(data = dataset, aes(x = !!rlang::sym(x_axis_field), y = factor(1))) +
+  violin_chart <- ggplot(
+    
+    data = dataset, 
+    
+    aes(x = !!rlang::sym(x_axis_field), y = factor(1))) +
+    
     geom_jitter(width = 0.25, alpha = 0.4, color = "#0072B2", size = 1.9, shape = 17) +
+    
     geom_violin(linewidth = 0.7, fill = "transparent", color = "black") +
+    
     geom_boxplot(width = 0.25, fill = "#E69F00", color = "black", alpha = 0.6, 
                  outlier.colour = "black", outlier.size = 0.75) +
     labs(
@@ -18,9 +25,11 @@ create_violin_chart <- function(
       subtitle = paste("(", earliest_title, "--", latest_title, ")", " n=", 
                        format(nrow(dataset), big.mark = ","), sep = "")
     ) +
+    
     theme(
       plot.title = element_text(size = 13, hjust = 0.5),
-      plot.subtitle = element_text(size = 9)
+      plot.subtitle = element_text(size = 7),
+      axis.text.x = element_text(face = "bold", size = 8),
     )
   
   suppressMessages(print(violin_chart))

### 2024-10-16 code/functions/new_calculate_values.R

@@ -1,7 +1,6 @@
 #########################################################################
 
 calculate_values <- function(X, base = 25) {
-
   # Special handling for small values
   if (X <= 5) {
     return(list(starting_value = 5, increment = 1))
@@ -11,19 +10,21 @@ calculate_values <- function(X, base = 25) {
     return(list(starting_value = 15, increment = 3))
   } else if (X > 15 && X < base) {
     return(list(starting_value = X, increment = floor(X / 5)))
-  }  
-    
-    # Regular function for X >= base
-    tier_base <- base * 10^floor(log10(X/base))
-    multiplier <- ceiling(X / tier_base)
-    
-    starting_value <- tier_base * multiplier
-    increment <- starting_value / 5
-#    cat("\nResults", starting_value, increment)
-    
-    result <- list(starting_value = increment, increment = increment)
-
-    return(result)
   }
+  
+  # Regular handling for X >= base
+  tier_base <- base * 10^floor(log10(X / base))
+  multiplier <- ceiling(X / tier_base)
+  starting_value <- tier_base * multiplier
+  increment <- floor(starting_value / 5)
+  
+  # Ensure starting_value does not exceed X
+  if (starting_value > X) {
+    starting_value <- floor(X / 10) * 10  # Adjust starting value closer to X, rounded down
+    increment <- max(1, floor(starting_value / 5))  # Ensure increment is a positive integer
+  }
+  
+  return(list(starting_value = starting_value, increment = increment))
+}
 
 #########################################################################

### 2024-10-16 code/timelinecharts..R

@@ -20,7 +20,8 @@ cat("\nExecution begins at:", formattedStartTime)
 #########################################################################
 
 # Set path for the data file
-main_data_file <- "10-year 2014-2023.csv"
+#main_data_file <- "10-year 2014-2023.csv"
+main_data_file <- "311_Service_Requests_from_2022-2023_AS_OF_09-15-2024.csv"
 
 #########################################################################
 

### 2024-10-21 manuscript/manuscript.tex
Further editing

@@
 
-Collectively these six City agnecies handle over 90\% of the total SR 
+Collectively these six City agencies handle over 90\% of the total SR 
 volume, indicative of the critical role these agencies play in managing
@@
 completeness, and consistency. Without thorough data cleansing, the utility
-of open data can be limited and perhaps untrustworhty, affecting 
+of open data can be limited and perhaps untrustworthy, affecting 
 its reliability for research, policy-making, and innovation. The 
@@
 
-	\item There are four date fields (created, closed, action\_updated, due).
+	\item There are four date fields 
+	(\texttt{created\_date}, \texttt{closed\_date}, \texttt{resolution\_action\_updated\_date}, \& \texttt{due\_date}).
 
@@
 
-	\item In addition to the incident\_address, there are five additional location fields: 
-	latitude/longitude, street\_name, landmark, USGS State Plane 
-	Coordinate System, and Block \#.
-
-	\item A free-form text field, resolution\_description, which 
+	\item In addition to  \texttt{incident\_address}, there are five additional location fields: 
+	\texttt{latitude \& longitude}, \texttt{street\_name}, \texttt{landmark}, USGS State Plane 
+	Coordinate System, and \texttt{block} \#.
+
+	\item A free-form text field, \texttt{resolution\_description}, which 
 	supports 934 characters of input, including commas and special characters
@@
 \texttt{location\_type}, \texttt{borough}, 
-and \texttt{cross\_street}. And the Few/None Empty includes
+and \texttt{cross\_street\_1 \& \_2}. And the Few/None Empty includes
 \texttt{created\_date}, \texttt{complaint\_type},
@@
 not insignificant. If you group the invalid entries by 
-percentage by Agency, it mirrors the \texttt{overall} 
+percentage by Agency, it mirrors the \textit{overall} 
 breakdown of SRs by Agency, potentially 
@@
 For the \texttt{created\_date} and \texttt{closed\_date} fields, one 
-might expect that these fields are automatically populated by the SR 
+might expect that these fields to be automatically populated by the  
 application software when setting an SR status to ``new'' or ``closed.'' 
 Unfortunately, this does not seem to be the case, as several anomalies 
-exist in the created and closed date fields. Here are some of the errors:
+exist in the \texttt{created\_date} and \texttt{closed\_date} fields that
+an automated process would normally exclude, including:
 
@@
     \item SRs with a \texttt{closed\_date} that occurs before the 
-    \texttt{created\_date} indicating a negative duration.
+    \texttt{created\_date} creating a ``negative duration''.
     \item \texttt{created\_date(s)} and \texttt{closed\_date(s)} in 
-    the far distant past, e.g. 1900.
+    the far distant past, e.g. ``1900-01-01''.
     \item \texttt{created\_date(s)} and \texttt{closed\_date(s)} that 
-    match to the second indicating a zero duration.
-    \item A large number of SRs closed and/or created exactly at midnight 
-    or noon, to the second.
+    match \textbf{to the second} indicating a nonsensical ``zero duration''.
+    \item An unusually large number of SRs closed and/or created exactly at midnight 
+    and noon, to the second.
 \end{itemize}
@@
 Citizens, NYC Government Officials, and Agencies use the created and 
-closed dates to measure the ``duration'' of SRs, which reflects Agency 
+closed dates to measure the \textit{duration} of SRs, which often reflects an Agency's 
 responsiveness. While duration is not directly present in the dataset, 
@@
 \texttt{(closed\_date} and \texttt{created\_date}.  Duration is one of 
-most used fields that is constantly analyzed, such as determining if 
-one NYC borough receives faster response for ceratin 
+most frequently used 311metrics and is constantly analyzed, such as determining if 
+one NYC borough receives faster response than another for certain 
 complaints. There are 12,450 SRs where the \texttt{closed\_date} precedes the 
@@
   \centering
-  \caption{Negative duraton of the largest magnitude (days) excluding
+  \caption{Negative duration of the largest magnitude ($-$730 days) excluding
     extreme negative values}
@@
 
-\subparagraph{Resolution Action Update Date}
-
-A similar issue occurs when resolution action update date is after the
-closing date.  When an SR is updated, the 311 software 
-automatically populates the \texttt{resolution\_action\_update\_date}. 
-While it is certainly possible to update an SR after it is closed, 
-some of these updates seem to be well beyond an expected 
-range, some over 2 years later. As shown in 
-Figure~\ref{fig:resolution-violin}, there are 10,265 SRs that 
-were updated more than 30 days after the 
-\texttt{closed\_date}, but less than 730 days, cutoff used to 
-exclude infeasible dates such as 1900-01-01. The chart 
-highlights the distribution of these post-closure updates, with 
-a notable concentration of updates within 
-the 30-to-90-day range. This pattern raises questions about whether 
-such delayed updates are standard operational practice or indicative 
-of a potential issue requiring further investigation.
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=\textwidth]{post_closed_violin_chart.pdf}
-  \caption{Post-closed resolution action update date(s) over 30 days.}
-  \label{fig:resolution-violin}
-\end{figure}
-
 \subparagraph{Identical Created Date \& Closed Date: Zero Durations}
@@
 \texttt{created\_date} are exactly the same, down to the second. This 
-creates a \textbf{zero duration}, which is nonsensical. There 
+creates a \textbf{zero duration} for the SR, which is nonsensical. There 
 are 163,720 such SRs, representing 2.6\% of all non-blank data. 
-Ninety-nine percent of these zero-duration SRs occur in five agencies:
-DHMH, DOT, DOB, DSNY, and DEP. This distribution does not 
+99\% percent of these zero-duration SRs occur within five agencies:
+DHMH, DOT, DOB, DSNY, and DEP. This distribution \textbf{does not} 
 mirror the overall SR distribution by Agency, suggesting 
@@
 calculated duration of these SRs by providing inaccurate dates. The 
-distribution by agency shows that over 90\% of the ``created-at-noon'' 
-SRs come from just two agencies: DOB and DSNY. Further investigation 
-reveals that a single agency, DSNY, is responsible for over 99\% of 
-the ``closed-at-noon'' SRs.
+distribution by agency shows that over 92\% of the ``closed-exactly-at-midnight'' 
+SRs come from just two agencies: DOB (69\%) and DSNY  (23\%).
+
+\subparagraph{Resolution Action Update Date}
+A similar issue occurs when the \texttt{resolution\_action\_update\_date} occurs well after the
+\texttt{closed\_date}.  When an SR is updated, the 311 software 
+automatically populates the \texttt{resolution\_action\_update\_date}. 
+While it is certainly possible, and even routine, to update an SR after it is closed, 
+some of these updates seem to be well beyond an expected 
+date, with some ``updates'' occurring over 2 years later. As shown in 
+Figure~\ref{fig:resolution-violin}, there are 10,265 SRs that 
+were updated >30 days after the 
+\texttt{closed\_date} (but <730 days, an arbitrary cutoff used to 
+exclude infeasible dates such as ``1900-01-01''. The chart 
+highlights the distribution of these post-closure updates, with 
+a notable concentration of updates occurring within 
+the 30-to-90-day range. This pattern raises questions about whether 
+such delayed updates are standard operational practice or indicative 
+of a potential issue which might warrant further investigation.
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=\textwidth]{post_closed_violin_chart.pdf}
+  \caption{Post-closed \texttt{resolution\_action\_update\_date}(s) >30 days.}
+  \label{fig:resolution-violin}
+\end{figure}
 
@@
 
-An question of precision vs. accuracy arises with the Latitude 
-and Longitude fields. Both Latitude are expressed as 
+An question of precision vs. accuracy arises with the \texttt{latitude} 
+and \texttt{longitude} fields. Both Latitude are expressed as 
 a 14-decimal number, e.g. 40.86769186022511. Given 
@@
 reference a DNA molecule is approximately 2nm in width. The 
-representation of the Latitude and Longitude fields appears to be a 
-classic case of 14-digit precision, but limited accuracy.
-
-\dt{STOPPED EDITING HERE}
+representation of the \texttt{latitude} and \texttt{longitude} fields appears to be a 
+classic case of extreme precision, but limited accuracy.
 
@@
 
-\textbf{Latitude/Longitude and Location:} The \texttt{location} field 
-is a pure concatenation of the latitude and longitude fields with a 
+\texttt{Latitude/Longitude} and \texttt{location}: The \texttt{location} field 
+is a pure concatenation of the \texttt{latitude} and \texttt{longitude} fields except with a 
 comma and parentheses added. This makes the \texttt{location} field 
@@
 \texttt{latitude}: 40.768456429488, \texttt{longitude}: $-$73.9575661888774, 
-and \texttt{location}: (40.768456429488, $-$73.95756618887745).
-
-\textbf{borough \& park\_borough:} These two fields are fully redundant, 
+form the \texttt{location}: (40.768456429488, $-$73.95756618887745).
+
+\texttt{borough} \& \texttt{park\_borough}: These two fields are fully redundant, 
 as they are 100\% matches.
 
-\textbf{borough \& taxi\_company\_borough:} Despite their names, these 
+\texttt{borough} \& \texttt{taxi\_company\_borough}: Despite their names, these 
 fields are almost entirely different, with only a 0.05\% match. The 
@@
 indicating a need for consultation with TLC to understand the differing 
-uses.
-\jy{The taxi\_company\_borough may mean the borough of the location of
-  the taxi company, which can be irrelevant to the borough where the
-  incident was reported.. Such information could be useful only when
-  studying TLC related SRs. It'd be better to use a second table to
-  store such TLC dependent information.}
+uses. It is likely that that the usage of \texttt{taxi\_company\_borough}
+is quite different than that of \texttt{borough}. Such information could be 
+useful only when studying TLC related SRs. It might make more
+sense to store TLC data in a separate dataset.
+
+\dt{STOPPED EDITING HERE}
 

### 2024-10-23 manuscript/manuscript.tex
Update to date aggregation and review of manuscript
Made changes to date aggregations (monthly, daily, hourly, etc.) to improve performance. Altered two images in manuscript. Continued editing.

@@
 The Mostly Empty category includes such fields as
-\texttt{taxi\_company due\_date},
-\texttt{pickup\_location}, and \texttt{landmark}.
+\texttt{taxi\_company\_borough}, \texttt{due\_date},
+\texttt{road\_ramp}, and \texttt{bridge\_highway\_name}.
 The Partially Empty includes such fields as
-\texttt{location\_type}, \texttt{borough}, 
+\texttt{location\_type}, \texttt{landmark}, 
 and \texttt{cross\_street\_1 \& \_2}. And the Few/None Empty includes
@@
   \centering
-  \caption{Negative duration of the largest magnitude ($-$730 days) excluding
-    extreme negative values}
+  \caption{Sample  negative durations [excluding extreme  values ($-$730 days)]}
   \begin{tabular}{l l l r l}
@@
 
-\textbf{landmark \& street\_name:} While the \texttt{landmark} field is 
+\texttt{landmark} \& \texttt{street\_name:} While the \texttt{landmark} field is 
 intended for noteworthy locations (e.g., parks, airports), most entries 
@@
 only in spelling or formatting (e.g., ``NINTH AVE'' vs. ``9 AVE''). This 
-indicates a possible misuse of the \texttt{landmark} field.
-
-\textbf{cross\_street\_1 \& intersection\_street\_1 and cross\_street\_2 
-\& intersection\_street\_2:} These two sets of street pairs are used to 
-help identify the incident address. After applying address 
-standardization using the R \textit{campfin} package, we found 88\% of 
-these pairs to be duplicates, while 12\% do not match. This discrepancy 
-raises the question of which field should be trusted. It is possible 
-that the two pairs originated from different agencies that used 
+indicates a possible misuse of the \texttt{landmark} field as well as
+a high level of redundancy.
+
+\texttt{cross\_street\_1} \& \texttt{intersection\_street\_1} and \texttt{cross\_street\_2} 
+\& \texttt{intersection\_street\_2}: These two sets of street pairs are used to 
+help identify the incident address. We found 88\% of 
+these pairs to be duplicates. This duplication  
+raises the question of which field-pair should be trusted. It is possible 
+that the two street pairs originated from different agencies that used 
 different variable names to represent the same information, but no 
-documentation exists to clarify this relationship.
-
-\paragraph{Near Match}
-
-\jy{This investigation could be relegated to on online supplement
-  given the low percentage of near-match. It's taking too much spaces
-  here.
-}
-
-\begin{table}[tbp]
-    \centering
-     \caption{Examples of near-matches for cross\_street\_1 and intersection\_street\_1.}
-     \label{tab:x1nearmatches}
-		\begin{tabular}{l l l r}
-	        \toprule
-	        \textbf{cross\_street\_1} & \textbf{intersection\_street\_1} 
-	        & \textbf{agency} & \textbf{hamming\_distance} \\
-	        \midrule
-	        WEST 168 ST    & WEST 167 ST           & DOT    & 1 \\
-	        105 AVE        & 107 AVE               & DOT    & 1 \\
-	        145 ST         & 150 ST                & DOT    & 2 \\
-	        138 ST         & 139 ST                & DOT    & 1 \\
-	        19 AVE         & 20 AVE                & DOT    & 2 \\
-	        76 ST          & 79 ST                 & DOT    & 1 \\
-	        67 ST          & 68 ST                 & DOT    & 1 \\
-	        \bottomrule
-	    \end{tabular}
-\end{table}
+documentation exists to clarify this relationship. Without question
+these two street-pairs are highly redundant.
 
@@
 
-In addition to the matching values, we encountered a number of 
-\textit{near-matches}, situations where the \texttt{cross\_street} 
-and \texttt{intersection\_street} are nearly, but not exactly, identical. 
-An example of such a near-match is when one field reads ``HARMON DR'' 
-and the other reads ``HARMON RD.'' To identify these near-matches, 
-we applied the \href{https://en.wikipedia.org/wiki/Hamming_distance}
-{\textit{Hamming Distance}}, which measures the number of character
-\jy{needs a quick definition and reference}
-changes required to make the two fields identical. For this analysis, 
-we set the threshold for a near-match to a Hamming Distance of two (2).
-
-Table~\ref{tab:x1near-matches} provides examples of near-matches for 
-\texttt{cross\_street\_1} and \texttt{intersection\_street\_1}. These 
-entries show slight variations in the street name or number, such as 
-``WEST 168 ST'' and ``WEST 167 ST,'' both processed by DOT, where only 
-one character differs. Although these discrepancies may seem small, 
-they can have implications for data accuracy, particularly when 
-geocoding or matching locations for analysis purposes.
-
 Table~\ref{tab:summary1} presents a summary of the match and 
@@
 entries are exact matches, while 25.4\% are blank but matching.
-However, 11.85\% of the entries are non-matching, and a small fraction 
-(0.002\%) fall into the near-match category. Although this percentage 
-seems negligible, even a small number of near-matches can lead to 
-confusion or errors when analyzing location-based data, especially 
-if critical decisions rely on accurate street data. Address 
-standardization or correction processes may be necessary to resolve 
-these inconsistencies and improve the reliability of the data.
 

### 2024-10-23 code/functions/base_bar_chart.R

@@ -1,5 +1,104 @@
 #########################################################################
 
+# base_bar_chart <- function(dataset, x_col, y_col, chart_title, sub_title, 
+#                            y_axis_labels = scales::comma, 
+#                            add_maximum = FALSE, add_minimum = FALSE, 
+#                            add_mean = FALSE, add_median = FALSE, 
+#                            add_sd = FALSE, add_second_maximum = FALSE,
+#                            add_trendline = FALSE, extra_line = NULL,
+#                            horizontal_adjustment_max = 0.5,
+#                            vertical_adjustment_max = -1,
+#                            console_print_out_title = "Data Summary",
+#                            chart_file_name = NULL) {
+#   
+#   # Step 1: Print Data Summary
+#   cat("\n\n", console_print_out_title, " (first 20 rows):\n", sep = "")
+#   dataset <- as.data.frame(dataset)
+#   print(head(dataset[, c(x_col, y_col)], 20), row.names = FALSE)
+#   
+#   # Sort dataset by y_col (descending) and x_col based on type
+#   if (inherits(dataset[[x_col]], "Date") || inherits(dataset[[x_col]], "POSIXct")) {
+#     sorted_dataset <- dataset[order(-dataset[[y_col]], dataset[[x_col]]), ]
+#   } else if (is.factor(dataset[[x_col]])) {
+#     sorted_dataset <- dataset[order(-dataset[[y_col]]), ]
+#   } else {
+#     sorted_dataset <- dataset[order(-dataset[[y_col]], -dataset[[x_col]]), ]
+#   }
+#   
+#   # Calculate total count for y_col
+#   y_total_count <- sum(dataset[[y_col]], na.rm = TRUE)
+#   
+#   # Step 2: Set up x_scale based on x_col type
+#   # if (inherits(dataset[[x_col]], "Date")) {
+#   #   x_scale <- scale_x_date(expand = c(0.01, 0), labels = scales::date_format("%Y-%m"), breaks = scales::date_breaks("6 months"))
+#   # } else if (inherits(dataset[[x_col]], "POSIXct") || inherits(dataset[[x_col]], "POSIXt")) {
+#   #   x_scale <- scale_x_datetime(expand = c(0.01, 0), labels = scales::date_format("%Y-%m"), breaks = scales::date_breaks("6 months"))
[…]

### 2024-10-23 code/functions/create_bar_chart_categorical.R

@@ -1,22 +1,47 @@
 #########################################################################
 
-create_bar_chart_categorical <- function(dataset, x_col, y_col, chart_title = "Bar Chart", sub_title = "", 
-                                         y_axis_labels = scales::comma, 
-                                         add_maximum = FALSE, add_minimum = FALSE, 
-                                         add_mean = FALSE, add_median = FALSE, 
-                                         add_sd = FALSE, add_second_maximum = FALSE,
-                                         add_trendline = FALSE, extra_line = NULL,
-                                         horizontal_adjustment_max = 0.5,
-                                         vertical_adjustment_max = -1,
-                                         console_print_out_title = "Data Summary",
-                                         chart_file_name = NULL) {
+create_bar_chart_categorical <- function(
+    dataset, 
+    x_col, 
+    y_col, 
+    chart_title = "Bar Chart", 
+    sub_title = "",
+    y_axis_labels = scales::comma,
+    add_maximum = FALSE, 
+    add_minimum = FALSE,
+    add_mean = FALSE, 
+    add_median = FALSE,
+    add_sd = FALSE, 
+    add_second_maximum = FALSE,
+    add_trendline = FALSE, 
+    extra_line = NULL,
+    horizontal_adjustment_max = 0.5,
+    vertical_adjustment_max = -1,
+    console_print_out_title = "Data Summary",
+    chart_file_name = NULL) 
+{
   
   # Call base_bar_chart without x_scale
-  base_bar_chart(dataset, x_col, y_col, chart_title, sub_title, 
-                 y_axis_labels, add_maximum, add_minimum, add_mean, 
-                 add_median, add_sd, add_second_maximum, add_trendline, 
-                 extra_line, horizontal_adjustment_max, vertical_adjustment_max, 
-                 console_print_out_title, chart_file_name)
+  base_bar_chart(
+    dataset, 
+    x_col, 
+    y_col, 
+    chart_title, 
+    sub_title,
+    y_axis_labels, 
+    add_maximum, 
+    add_minimum, 
+    add_mean,
+    add_median, 
+    add_sd, 
+    add_second_maximum, 
+    add_trendline,
+    extra_line, 
+    horizontal_adjustment_max, 
+    vertical_adjustment_max,
+    console_print_out_title, 
+    chart_file_name
+  )
 }
 
 #########################################################################

### 2024-10-23 code/timelinecharts..R

@@ -1,892 +1,0 @@
-# install.packages("ggplot2")
-# install.packages("scales")
-# install.packages("dplyr")
-# install.packages('zoo')
-# install.packages("ggpmisc")
-# install.packages("lubridate")
-rm(list = ls(), envir = .GlobalEnv)
-
-library(ggplot2)
-library(scales)
-library(dplyr)
-library(zoo)
-library(ggpmisc)
-library(lubridate)
-
-programStart <- as.POSIXct(Sys.time())
-formattedStartTime <- format(programStart, "%Y-%m-%d %H:%M:%S")
-cat("\nExecution begins at:", formattedStartTime)
-
-#########################################################################
-
-# Set path for the data file
-main_data_file <- "10-year 2014-2023.csv"
-#main_data_file <- "311_Service_Requests_from_2022-2023_AS_OF_09-15-2024.csv"
-#main_data_file <- "JAN-SEP_2024_AS_OF_10-16-2024.csv"
-
-#########################################################################
-
-cat("\n***** Program initialization *****")
-
-setwd("C:/Users/david/OneDrive/Documents/datacleaningproject/nyc311clean/code")
-
-# Define the path to the directory containing your function scripts
-functions_path <- "functions"
-
-# Source all .R files in the directory
-files <- list.files(functions_path, pattern = "\\.R$", full.names = TRUE)
-
-# Source each file
-lapply(files, source)
-
-data1File <- file.path("..", "..", "data", main_data_file)
-
-# Extract the first two digits for file naming purposes
-year_digits <- substr(main_data_file, 1, 2)
-
-if (year_digits != "10") { year_digits <- "2"} 
-  # Set path for the chart directory
-
-chart_prefix <- paste0(year_digits, "-year")
-#chart_prefix <- "Jan-Sep_2024"
-chart_directory_path <- file.path("..", "..", "charts", "2022-2023 study", chart_prefix)
-
-file_name_prefix <- chart_prefix #to name individual chart files
-
-# Set scipen option to a large value to prevent scientific notation for numbers
-options(scipen = 10)
-
-
-sink(paste0("../../console_output/", year_digits, "-yr timeline_console_output.txt"))
-#sink(paste0("../../console_output/", "2024_console_output.txt"))
-
-
-cat("\nExecution begins at:", formattedStartTime)
-
-#########################################################################
[…]

### 2024-10-23 code/timelinecharts.R

@@ -1,0 +1,1527 @@
+# install.packages("ggplot2")
+# install.packages("scales")
+# install.packages("dplyr")
+# install.packages('zoo')
+# install.packages("ggpmisc")
+# install.packages("lubridate")
+rm(list = ls(), envir = .GlobalEnv)
+
+library(ggplot2)
+library(scales)
+library(dplyr)
+library(zoo)
+library(ggpmisc)
+library(lubridate)
+
+programStart <- as.POSIXct(Sys.time())
+formattedStartTime <- format(programStart, "%Y-%m-%d %H:%M:%S")
+cat("\nExecution begins at:", formattedStartTime)
+
+#########################################################################
+
+# Set path for the data file
+#main_data_file <- "10-year 2014-2023.csv"
+main_data_file <- "311_Service_Requests_from_2022-2023_AS_OF_09-15-2024.csv"
+#main_data_file <- "JAN-SEP_2024_AS_OF_10-16-2024.csv"
+#main_data_file <- "smaller_test_data.csv"
+# main_data_file <- "extra_small.csv"
+
+#########################################################################
+
+cat("\n***** Program initialization *****")
+
+setwd("C:/Users/david/OneDrive/Documents/datacleaningproject/nyc311clean/code")
+
+# Define the path to the directory containing your function scripts
+functions_path <- "functions"
+
+# Source all .R files in the directory
+files <- list.files(functions_path, pattern = "\\.R$", full.names = TRUE)
+
+# Source each file
+lapply(files, source)
+
+data1File <- file.path("..", "..", "data", main_data_file)
+
+# Extract the first two digits for file naming purposes
+year_digits <- substr(main_data_file, 1, 2)
+
+if (year_digits != "10") {
+  year_digits <- "2"
+}
+# Set path for the chart directory
+
+chart_prefix <- paste0(year_digits, "-year")
+# chart_prefix <- "Jan-Sep_2024"
+chart_directory_path <- file.path("..", "..", "charts", "2022-2023 study", chart_prefix)
+
+file_name_prefix <- chart_prefix # to name individual chart files
+
+# Set scipen option to a large value to prevent scientific notation for numbers
+options(scipen = 10)
+
+
+ sink(paste0("../../console_output/", year_digits, "-yr timeline_console_output.txt"))
+# sink(paste0("../../console_output/", "2024_console_output.txt"))
+
+
+cat("\nExecution begins at:", formattedStartTime)
+
[…]

### 2024-11-04 manuscript/manuscript.tex
Small cosmetic changes to R code. Review of last part of Manuscript

@@
 10\% of SRs is distributed across 10 other
-agencies. This concentration of SRs to the `big six'' agencies 
+agencies. This concentration of SRs to the ``big six'' agencies 
 underscores the necessity for optimal data handling and curation processes 
@@
         \includegraphics[width=\textwidth]{2-year-trend_SRs_created_on_the_hour.pdf}
-        \caption{SRs Created by Hour of the Day}
+        \caption{SRs Created Exactly on the Hour}
         \label{fig:busiestcreated}
@@
         \includegraphics[width=\textwidth]{2-year-trend_SRs_closed_on_the_hour.pdf}
-        \caption{SRs Closed by Hour of the Day}
+        \caption{SRs Closed Exactly on the Hour}
         \label{fig:busiestclosed}
@@
 
-\dt{STOPPED EDITING HERE}
-
-\textbf{agency \& agency\_name:} The \texttt{agency} field contains 
-abbreviations of City agencies (e.g., NYPD, DOT), while 
+\texttt{agency} \& \texttt{agency\_name}: The \texttt{agency} field contains 
+abbreviations of City agencies (e.g., NYPD, DOT, DSNY), while 
 \texttt{agency\_name} contains the full names. Including both seems 
 redundant, especially given the space taken up by full names in the 
-CSV file.
-
-\jy{In my Chance article, we found an agency
-  `DEPARTMENT OF CONSUMER AND WORKER PROTECTION', which occurred 361
-  times in May 2023. It was the only one that does not look like an
-  acronym. Did this occur in the two-year data?
-  I wonder if this has been fixed lately.}
-
-\jy{A related question is whether the data quality has improved over
-  the years}
-
-\texttt{landmark} \& \texttt{street\_name:} While the \texttt{landmark} field is 
-intended for noteworthy locations (e.g., parks, airports), most entries 
+CSV file. Additionally, the agency abbreviations are easily recognized.
+
+\texttt{landmark} \& \texttt{street\_name}: Per the Data Dictionary, the
+\texttt{landmark} field ``can refer to any noteworthy location, 
+including but not limited to, parks, hospitals, airports, sports 
+facilities, performance spaces, etc.'  However, most entries 
 are actually street names, with a 62\% match to \texttt{street\_name}. 
-Even non-matches (excluding blanks) appear to be similar, differing 
-only in spelling or formatting (e.g., ``NINTH AVE'' vs. ``9 AVE''). This 
-indicates a possible misuse of the \texttt{landmark} field as well as
-a high level of redundancy.
-
-\texttt{cross\_street\_1} \& \texttt{intersection\_street\_1} and \texttt{cross\_street\_2} 
-\& \texttt{intersection\_street\_2}: These two sets of street pairs are used to 
-help identify the incident address. We found 88\% of 
-these pairs to be duplicates. This duplication  
-raises the question of which field-pair should be trusted. It is possible 
-that the two street pairs originated from different agencies that used 
-different variable names to represent the same information, but no 
-documentation exists to clarify this relationship. Without question
-these two street-pairs are highly redundant.
+Even non-matches (excluding blanks) appear to technically be
+a match, differing only in spelling or formatting 
+(e.g., ``NINTH AVE'' vs. ``9 AVE''). This indicates a possible 
+misuse of the \texttt{landmark} field as well as a high level of redundancy.
+
+\texttt{cross\_street\_1/2} \& \texttt{intersection\_street\_1/2}: These 
+two sets of street pairs are used to help identify the 
+\texttt{incident address}. We found 88\% of these pairs to be 
+duplicates. This duplication raises the question of which field-pair 
+should be trusted; and if they are different, how should they be treated?
+It is possible that these two street pairs originated from different 
+agencies that used different variable names to represent the same 
+information, but no documentation exists to clarify this 
+relationship. These two street-pairs are highly redundant.
 
@@
 
-Table~\ref{tab:summary1} presents a summary of the match and 
-near-match results for \texttt{cross\_street\_1} and 
-\texttt{intersection\_street\_1}. As the table shows, 88.15\% of the 
-entries are exact matches, while 25.4\% are blank but matching.
+Table~\ref{tab:summary1} presents a summary of the matching results 
+for \texttt{cross\_street\_1} and \texttt{intersection\_street\_1}. As 
+the table shows, 88.15\% of the non-blank entries are exact matches.
 
@@
 less storage impact, and simpler data analysis efforts.
+
 Below is a list of duplicate and near-duplicate fields.
@@
 \begin{itemize}
-    \item \textbf{agency\_name}: Each row contains the \texttt{agency} field, 
-    which uses abbreviations that are clear and well understood.
-
-    \item \textbf{park\_borough}: A 100\% match with the \texttt{borough} field, 
-    leading to no data loss.
-
-    \item \textbf{location}: A straight concatenation of latitude and longitude 
-    fields, offering no additional value beyond duplication.
-
-    \item \textbf{cross\_street\_1 \& 2 and intersection\_street\_1 \& 2}: 
-    An 88\% match between these pairs. We recommend deleting the 
-    \texttt{intersection\_street\_1 \& 2} fields while acknowledging the
-    concomitant data loss.
+    \item \textbf{\texttt{agency\_name}}: Since each row contains the \texttt{agency} 
+    field, which uses abbreviations that are clear and well understood, we believe the
+    \texttt{agency\_name} can be removed with minimal data loss.
+
+    \item \textbf{\texttt{park\_borough}}: This field is a 100\% match with 
+    the \texttt{borough} field and can be deleted without data loss.
+
+    \item \textbf{\texttt{location}}: This field is a straight concatenation of 
+    the latitude and longitude fields. It can be deleted without data loss.
+
+    \item \textbf{\texttt{cross\_street\_1/2}} and \textbf{\texttt{intersection\_street\_1/2}}: 
+    These two stree pairs both exhibit an 88\% match. We recommend deleting the 
+    \texttt{intersection\_street\_1/2} fields while acknowledging 
+    the concomitant data loss.
 
@@
 
-Additionally, certain fields may not be useful due to their sparse population. 
-For example, the \texttt{taxi\_company\_borough} field is 99.94\% blank, 
-\texttt{road\_ramp} is 99.8\% blank, \texttt{vehicle\_type} is 99.7\% blank, 
-and \texttt{due\_date} is 99.6\% blank. Several other fields, such as 
-\texttt{bridge\_highway\_direction}, \texttt{bridge\_highway\_name}, 
-\texttt{bridge\_highway\_segment}, and \texttt{taxi\_pick\_up\_location}, 
-are over 99\% blank, with most used exclusively by the TLC.
-
-Further reduction is possible by coding categorical 
-variables, such as \texttt{complaint\_type}, through methods like dropdown 
-menus or encoded categories. By using a dropdown menu for 
-\texttt{complaint\_type}, data entry can be standardized, reducing 
-inconsistencies caused by manual input errors. Additionally, categorical 
-variables with a limited set of possible values can be stored more 
-efficiently by assigning unique codes or numerical representations for 
-each category. For instance, instead of repeatedly storing the full text 
-for every complaint type (e.g., ``Noise - Residential,'' ``Noise - 
-Commercial''), a single code could represent each type. This not only 
-reduces file size but also improves data processing speed and efficiency 
-during analysis. Moreover, standardizing categorical inputs prevents 
-the inclusion of misspellings or unrecognized entries, ensuring cleaner 
+Additionally, certain fields may not be useful for broad analysis
+due to their sparse data population. This especially applies to the fields below which
+are predominantly used by the TLC and are 99\% (or greater) blank. 
+\begin{quote}
+\textbf{Note:} \textit{Even a field that is 99\% blank still contains 64,000 
+non-blank rows in this large (6.4 million-row) dataset.}
+\end{quote}
+
+\begin{itemize}
+    \item \texttt{taxi\_company\_borough}
+    \item \texttt{road\_ramp}
+    \item \texttt{vehicle\_type}
+    \item \texttt{due\_date}
+    \item \texttt{bridge\_highway\_direction}
+    \item \texttt{bridge\_highway\_name}
+    \item \texttt{bridge\_highway\_segment}
+    \item \texttt{taxi\_pick\_up\_location}
+\end{itemize}
+
+Further reduction is possible by encoding certain categorical 
+variables Using \texttt{complaint\_type} as an example,
+instead of repeatedly storing the full tex of each complaint
+type, it would use a numeric representation. 
+(e.g., ``NOISE - STREET/SIDEWALK'')could be represented by
+, a numeric code (e.g. 126). In this datase the example complaint type
+occurs over 300,000 times. Such encoding not only reduces file 
+size, but also improves processing efficiency. Moreover, 
+standardizing categorical inputs prevents the inclusion of 
+misspellings or unrecognized/invalid entries, ensuring cleaner 
 and more reliable data. This approach is particularly useful in large 
-datasets, where frequently repeated categorical variables can 
-significantly increase storage requirements. Encoding such fields 
-as dropdown options or coded variables can lead to faster query times 
-and more consistent data handling.
+datasets, where frequently repeated text-based categorical variables can 
+significantly increase storage requirements.
 
@@
 
-The Data Dictionary could benefit from an update. We found several 
+The Data Dictionary would benefit from an update. We found several 
 discrepancies between the Data Dictionary and the actual data, which 
-could potentially mislead analytic efforts.
+could potentially mislead analytic efforts. Here are some examples:
 
@@
     fields, the Data Dictionary does not. Nor does it provide details 
-    on specific fields, such as constraints or the domain of legal 
-    values. For example, the \texttt{incident\_zip} field is specified 
-    as ``text.'' This may not be the most suitable definition; 
-    ``categorical,'' with a note that the data contains only numeric 
-    characters and is not subjected to arithmetic operations, might 
-    be a better description.
+    on specific fields such as constraints or the domain of legal 
+    values. For example, in the Data Dictionary, the \texttt{incident\_zip} 
+    field is specified as ``text.'' This may not be the most suitable 
+    definition. Perhaps ``categorical,'' with a note indicating that 
+    the data must contain only numeric values and is not subjected to 
+    arithmetic operations would be a better dictionary description.
 
@@
     \texttt{status} field has values of \textit{assigned, canceled, 
-    closed}, or \textit{pending}, but we also found values such as 
-    \textit{in progress, started}, and \textit{unspecified}. 
+    closed}, or \textit{pending}. However, we found additional
+    values such as \textit{in progress, started}, and \textit{unspecified}. 
     Additionally, no SRs were found with a status of 
-    \textit{canceled}. The \texttt{address\_type} field similarly 
+    \textit{canceled}.
+
+    \item The \texttt{address\_type} field similarly 
     lists values of \textit{address, blockface, intersection, latlong}, 
-    and \textit{placename}, but we found additional values such as 
-    \textit{bbl} and \textit{unrecognized}, and no \textit{latlong} 
-    values were observed. Other fields, like \texttt{facility\_type}, 
+    and \textit{placename}. However, we found additional values such as 
+    \textit{bbl} and \textit{unrecognized}. Additionally, no \textit{latlong} 
+    values were observed. Other fields, such as \texttt{facility\_type}, 
     \texttt{vehicle\_type}, \texttt{taxi\_pick\_up\_location}, 
@@
 clear protocols for data entry, ensuring that fields are populated with 
-valid, legally acceptable values. For example, fields like 
+valid, legally acceptable values. For example, fields such as 
 \texttt{created\_date} and \texttt{closed\_date} should be subject to 
@@
 like geographic locations and agency names enhances the quality 
-and reliability of the data.
+and reliability of the data. Additionally, a number of fields have 
+very defined domains, such as \texttt{incident\_zip}. Such fields
+should be subject to data entry validation, in this example
+validating against the USPS database at the data entry point.
 
@@
 Clear and accurate documentation is essential for the effective use 
-of government-released datasets. The NYC 311 data revealed several 
+of government-released datasets. Our study found several 
 issues where the Data Dictionary did not accurately reflect the 
-actual data, such as missing field descriptions for computed fields. 
+actual data, such as missing field descriptions for computed fields, and 
+inaccurate domain values for categorial fields such as \texttt{status}. 
 As mentioned, updating the Data Dictionary to include descriptions of all fields, 
@@
 would improve transparency and usability. Regular updates should 
-be made to ensure that any changes in data structure or content are 
-reflected. This practice enhances the ability of external users, 
+occur whenever there are changes in data structure, data usage, or 
+content. This practice enhances the ability of external users, 
 including researchers and developers, to understand and work 
@@
 NYC 311 case, issues such as large spikes in SR creation and closure 
-at exactly midnight or noon likely stem from automated processes that 
-distort the accurate capture of SR durations. By implementing 
-real-time validation tools, governments can prevent such anomalies 
-from entering the dataset. Automated systems can be designed to 
-flag inconsistencies, incorrect values, or illogical sequences, or 
-illogical date entries. 
+at exactly midnight (or noon) likely stem from a cron-job type 
+processes that distort the accurate capture of SR 
+durations. By implementing real-time validation tools, governments 
+can prevent such anomalies from entering the dataset in the first 
+place. Automated 
+systems can be designed to flag inconsistencies, incorrect values, 
+or illogical sequences, or illogical date entries. 
 Automation can also facilitate faster detection and correction of 
@@
 data quality issues, such as notifying the public when errors in SR 
-fields (e.g., erroneous closed dates) are discovered and explaining 
-how such issues will be resolved. By ensuring transparency and 
-keeping the public informed, governments foster trust and promote 
-wider use of datasets for research, civic engagement, and innovation.
+fields (e.g., erroneous \texttt{closed\_ date(s)} are discovered, including
+an explanation about how such issues will be resolved. By 
+ensuring transparency and keeping the public informed, 
+governments foster trust and promote wider use of datasets 
+for research, civic engagement, and innovation.
 
@@
 
-% \begin{itemize}
-% \item  Undertake a dedicated effort 
-%     to investigate these data anomalies, confirm the veracity of the 
-%     issues, and collaborate with the appropriate Agency Open Data 
-%     Representatives to find resolutions. This effort may be complex 
-%     and require software changes. Coordination and mutual approval 
-%     of changes will be key, requiring strong support from senior 
-%     leadership.
-
-%     \item Standards and protocols should be established, agreed upon, 
-%     and enforced. This includes defining the domain of legal values, 
-%     ensuring data accuracy, implementing address normalization, 
-%     clarifying data usage, and providing detailed data definitions.
-
-%     \item Resolve many potential data integrity issues
-%     through enforcement by software. For instance, software should 
-%     prevent invalid entries like erroneous zip codes. Similarly, fields 
-%     like \texttt{created\_date} and \texttt{closed\_date}, which 
-%     measure responsiveness, should have logic checks to prevent 
-%     illogical sequences, such as an SR being ``closed'' before 
-%     being ``opened.''
-
-%     \item The large spikes in SR creation and closure at midnight and 
-%     noon likely point to automated processes from external Agency 
-%     integrations. These distortions should be eliminated in favor of 
-%     capturing accurate SR durations with precise timestamps.
-
-%     \item As discussed, there are six ``computed'' fields that are not 
-%     listed in the Data Dictionary. Some of these fields contain 
-%     significant inaccuracies. We recommend updating the Data 
-%     Dictionary and, if necessary, masking these fields from public use 
-%     until their accuracy is improved.
-
-%     \item Duplicate fields should be reviewed for removal. In some 
-%     cases, such as \texttt{borough} and \texttt{park\_borough}, the 
-%     duplication is 100\%. In others, the match is 80-90\%. 
-%     These near-duplicates should be assessed to determine their 
-%     relevance in the 311 SR dataset. Initial analysis suggests that 
-%     removing these fields could reduce file size by up to 12\%, with 
-%     further reductions possible upon deeper review.
-
-%     \item Fields with very low data presence, such as 
-%     \texttt{taxi\_company\_borough}, populated only 0.06\% of the time, 
-%     should be evaluated to determine if they are necessary.
-
-%     \item The Data Dictionary, last updated in March 2023, should be 
-%     revised based on the issues discussed in Section~\ref{sec:datadictionary}.
-
-%     \item The precision of latitude, longitude, and location fields 
-%     should be reduced. We recommend limiting precision to four 
-%     decimal places (0.000X), which represents approximately 36.4 feet. 
-%     More decimal points strain credibility.
-
-%     \item The 11 address-related fields (\texttt{incident\_address}, 
-%     \texttt{street\_name}, \texttt{city}, \texttt{intersection\_street\_1 \& 2}, 
-%     \texttt{cross\_street\_1 \& 2}, \texttt{landmark}, \texttt{road\_ramp}, 
-%     \texttt{bridge\_highway\_segment}, and \texttt{taxi\_pick\_up\_location}) 
-%     should be standardized to USPS format. Software can easily 
-%     perform this standardization.
-
-%     \item There are multiple ways missing data is represented, such as 
-%     ``UNKNOWN'', ``N/A'', ``NA'', ``na'', and blanks (`` ''). Protocols 
-%     should be established to standardize the representation of missing data.
-% \end{itemize}
-
 \section{Discussion} \label{sec:discussion}
@@
 across city agencies, particularly in resolving discrepancies in field
-names, formats, and definitions. Consistency in data is essential for
+names, formats, field usage, and definitions. Consistency in data is essential for
 trustworthy machine learning, where inconsistencies across sources can

### 2024-11-04 code/datacleansing.R

@@ -63,7 +63,7 @@ lapply(files, source)
 options(scipen = 10)
 
 
-sink("../../console_output/core_console_output.txt")
+#sink("../../console_output/core_console_output.txt")
 #sink("../../console_output/2024_console_output.txt")
 
 cat("\nExecution begins at:", formattedStartTime)

### 2024-11-04 code/timelinecharts.R

@@ -20,8 +20,8 @@ cat("\nExecution begins at:", formattedStartTime)
 #########################################################################
 
 # Set path for the data file
-main_data_file <- "10-year 2014-2023.csv"
-#main_data_file <- "311_Service_Requests_from_2022-2023_AS_OF_09-15-2024.csv"
+#main_data_file <- "10-year 2014-2023.csv"
+main_data_file <- "311_Service_Requests_from_2022-2023_AS_OF_09-15-2024.csv"
 #main_data_file <- "JAN-SEP_2024_AS_OF_10-16-2024.csv"
 #main_data_file <- "smaller_test_data.csv"
 # main_data_file <- "extra_small.csv"

### 2024-11-04 manuscript/manuscript.tex
Adjustment to charts to better display in Manuscript
Adjusted PDF dimensions to enable better display of charts in Manuscript. Also continued editing Manuscript.

@@
 two sets of street pairs are used to help identify the 
-\texttt{incident address}. We found 88\% of these pairs to be 
+\texttt{incident address}. We found 88\% of these two street pairs to be 
 duplicates. This duplication raises the question of which field-pair 
@@
 relationship. These two street-pairs are highly redundant.
-
-\begin{table}[tbp]
-    \centering
-    \caption{Summary of cross\_street\_1 and intersection\_street\_1}
-    \label{tab:summary1}
-	    \begin{tabular}{l r r}
-	        \toprule
-	        \textbf{category} & \textbf{count} & \textbf{percentage} \\
-	        \midrule
-	        Matching                    & 5,637,182 & 88.15     \\
-	        Both blank -- Matching      & 1,624,499 & 25.4      \\
-	        Non-matching                &   757,726 & 11.85     \\
-	        cross\_street\_1\_blank     &   222,232 & N/A       \\
-	        intersection\_street\_1\_blank &   519,348 & N/A       \\
-	        Near-match                  &       128 & 0.002002  \\
-	        \bottomrule
-	    \end{tabular}
-\end{table}
-
-Table~\ref{tab:summary1} presents a summary of the matching results 
-for \texttt{cross\_street\_1} and \texttt{intersection\_street\_1}. As 
-the table shows, 88.15\% of the non-blank entries are exact matches.
 

### 2024-11-04 code/datacleansing.R

@@ -63,7 +63,7 @@ lapply(files, source)
 options(scipen = 10)
 
 
-#sink("../../console_output/core_console_output.txt")
+sink("../../console_output/core_console_output.txt")
 #sink("../../console_output/2024_console_output.txt")
 
 cat("\nExecution begins at:", formattedStartTime)
@@ -296,36 +296,89 @@ total_count <- sum(missingDataPerColumn$count)
 # Sort 'field' by 'total_empty' descending
 missingDataPerColumn <- missingDataPerColumn %>%
   mutate(field = reorder(field, -total_empty))
+# 
+# # Create the bar chart with vertical X-axis labels
+# blank_chart <- ggplot(
+#   
+#   missingDataPerColumn, aes(x = reorder(field, -total_empty), y = total_empty)) +
+#   
+#   geom_bar(stat = "identity", fill = "#009E73") +
+#   
+#   theme(
+#     axis.title.x = element_text(vjust = 0, size = 8),
+#     axis.title.y = element_text(vjust = 1, size = 8),
+#     plot.title = element_text(hjust = 0.5, size = 13),
+#     plot.subtitle = element_text(size = 7),
+#     panel.background = element_rect(fill = "gray95", color = "gray95"),
+#     axis.text.x = element_text(angle = 50, vjust = 1, hjust = 1, face = "bold"),
+#     axis.text.y = element_text(face = "bold"),
+#     legend.position = "none", # Remove all legends
+#     aspect.ratio = 0.618033 # Set aspect ratio
+#   ) +
+#   
+#   geom_text(aes(
+#     x = field, y = total_empty, label = pct_empty,
+#     angle = -70
+#   ), size = 2.3, color = "black") +
+#   
+#   ggtitle(NULL,
+#     subtitle = paste(chart_sub_title, format(num_rows_d311, big.mark = ","), sep = "")
+#   ) +
+#   
+#   labs(y = NULL, x = NULL)
+# 
+# # Build the plot to extract y-axis breaks
+# built_plot <- ggplot_build(blank_chart)
+# 
+# # Extract Y-axis breaks (try different possible locations)
+# y_breaks <- built_plot$layout$panel_params[[1]]$y$get_breaks()
+# # Filter out any NA values from y_breaks
+# y_breaks <- y_breaks[!is.na(y_breaks)]
+# 
+# # Add hlines using the Y-axis breaks
+# blank_chart <- blank_chart +
[…]

### 2024-11-04 code/functions/base_bar_chart.R

@@ -1,104 +1,5 @@
 #########################################################################
 
-# base_bar_chart <- function(dataset, x_col, y_col, chart_title, sub_title, 
-#                            y_axis_labels = scales::comma, 
-#                            add_maximum = FALSE, add_minimum = FALSE, 
-#                            add_mean = FALSE, add_median = FALSE, 
-#                            add_sd = FALSE, add_second_maximum = FALSE,
-#                            add_trendline = FALSE, extra_line = NULL,
-#                            horizontal_adjustment_max = 0.5,
-#                            vertical_adjustment_max = -1,
-#                            console_print_out_title = "Data Summary",
-#                            chart_file_name = NULL) {
-#   
-#   # Step 1: Print Data Summary
-#   cat("\n\n", console_print_out_title, " (first 20 rows):\n", sep = "")
-#   dataset <- as.data.frame(dataset)
-#   print(head(dataset[, c(x_col, y_col)], 20), row.names = FALSE)
-#   
-#   # Sort dataset by y_col (descending) and x_col based on type
-#   if (inherits(dataset[[x_col]], "Date") || inherits(dataset[[x_col]], "POSIXct")) {
-#     sorted_dataset <- dataset[order(-dataset[[y_col]], dataset[[x_col]]), ]
-#   } else if (is.factor(dataset[[x_col]])) {
-#     sorted_dataset <- dataset[order(-dataset[[y_col]]), ]
-#   } else {
-#     sorted_dataset <- dataset[order(-dataset[[y_col]], -dataset[[x_col]]), ]
-#   }
-#   
-#   # Calculate total count for y_col
-#   y_total_count <- sum(dataset[[y_col]], na.rm = TRUE)
-#   
-#   # Step 2: Set up x_scale based on x_col type
-#   # if (inherits(dataset[[x_col]], "Date")) {
-#   #   x_scale <- scale_x_date(expand = c(0.01, 0), labels = scales::date_format("%Y-%m"), breaks = scales::date_breaks("6 months"))
-#   # } else if (inherits(dataset[[x_col]], "POSIXct") || inherits(dataset[[x_col]], "POSIXt")) {
-#   #   x_scale <- scale_x_datetime(expand = c(0.01, 0), labels = scales::date_format("%Y-%m"), breaks = scales::date_breaks("6 months"))
[…]

### 2024-11-04 code/functions/create_combo_chart.R

@@ -68,7 +68,7 @@ create_combo_chart <- function(
     labs(x = NULL, y = NULL) +
     
     theme(
-      axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1, face = "bold", size = 8),
+      axis.text.x = element_text(angle = 70, vjust = 1, hjust = 1, face = "bold", size = 7),
       axis.text.y = element_text(face = "bold", size = 8),
       axis.text.y.right = element_text(color = "black", face = "bold", size = 8),
       plot.title = element_text(hjust = 0.5, size = 12),
@@ -100,8 +100,9 @@ create_combo_chart <- function(
   # Save the plot
   chart_path <- file.path(chart_directory_path, chart_file_name)
   chart_width <- 10
-  chart_height <- chart_width / 1.618
-  ggsave(chart_path, plot = combo_chart, width = chart_width, height = chart_height, dpi = 300)
+#  chart_height <- chart_width / 1.618
+  chart_height <- chart_width / 1.3
+    ggsave(chart_path, plot = combo_chart, width = chart_width, height = chart_height, dpi = 300)
 }
 
 #########################################################################

### 2024-11-06 manuscript/manuscript.tex
Top to bottom review. Mostly cleanup, spelling, grammar, etc.
Still need to work on the sections at the end. Part 4.

@@
 
-\hyphenpenalty=750
+\hyphenpenalty=900
 
@@
 insights from this case study, we propose a set of data curation 
-principles tailored for government-released open data. These principles 
+principles tailored for government\mbox{-}released open data. These principles 
 aim to enhance data management practices and ensure 
@@
 college enrollment statistics, jail inmate charges, and the location
-of city-wide free Internet access points. These datasets have been
+of city\mbox{-}wide free Internet access points. These datasets have been
 applied in civil life in various ways, such as mapping car crashes
@@
 consistent, accurate, and useful for diverse applications. For example, 
-well-curated data is essential for machine learning systems, which 
-require high-quality data to produce reliable insights 
+well\mbox{-}curated data is essential for machine learning systems, which 
+require high\mbox{-}quality data to produce reliable insights 
 \citep{polyzotis2019data, jain2020overview}. A critical component of 
@@
 \citet{shankar2021data} observed its critical role during the 
-COVID-19 pandemic in managing collective responses.
+COVID\mbox{-}19 pandemic in managing collective responses.
 
@@
 example for examining key issues in data curation, including data 
-validity, consistency, and curation efficiency. 
-We illustrate these points with live examples drawn from our 
+validity, consistency, and curation efficiency. We illustrate these 
+points with live examples drawn from our 
 processing of the 311 SR data. Secondly, building upon insights 
 gained from this case study, we propose a set of data curation 
-principles tailored for government-released open data. These 
+principles tailored for government\mbox{-}released open data. These 
 principles are designed to address the unique challenges 
@@
 an overview of the history of the NYC 311 SR system and presents 
-a summary of SR counts over a 2-year period from 2022 to 2023. 
+a summary of SR counts over a 2\mbox{-}year period from 2022 to 2023. 
 Section~\ref{sec:issues} delves into specific data cleansing 
@@
 and the presence of missing, blank, or N/A entries. This section
- investigates field compliance with reference or acceptable values, 
-highlights logical inconsistencies, and examines concerning patterns 
+ also investigates field compliance with reference or acceptable values, 
+highlights logical inconsistencies, and examines some concerning patterns 
 in the data. We explore the balance between precision and accuracy 
 and identify duplicate or redundant fields, along with observations 
-on the Data Dictionary. Section~\ref{sec:recommendations} offers 
+on the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}.
+
+Section~\ref{sec:recommendations} offers 
 practical recommendations for mitigating or resolving these issues, 
@@
 engagement and service response framework, serves as a centralized hub
-for non-emergency inquiries and requests. Introduced in 2003, the NYC
+for non\mbox{-}emergency inquiries and requests. Introduced in 2003, the NYC
 311 system was designed to streamline the city's response to
-non-emergency issues, ranging from noise complaints to street
-maintenance requests. Initially a phone-based call center, the system
+non\mbox{-}emergency issues, ranging from noise complaints to street
+maintenance requests. Initially a phone\mbox{-}based call center, the system
 evolved into a comprehensive data management platform handling
@@
 record high of 348,463 monthly service requests in August 2020 due to
-the COVID-19 pandemic, the 2021 expansion to include the MTA's city
+the COVID\mbox{-}19 pandemic, the 2021 expansion to include the MTA's city
 subway system (the biggest expansion in their history), and a 
@@
 Portal, which provides tools for querying, grouping, aggregating,
-geo-mapping, visualizing, and exporting results. The
+geo\mbox{-}mapping, visualizing, and exporting results. The
 \href{https://www.nyc.gov/content/oti/pages/}{NYC Office of Technology
   and Innovation (OTI)}, provides technological support for the
-open data system and data portal. Input data is sourced from 17 different
-NYC Agencies.
-
-\begin{comment}
-Despite its success, the 311 system faces several challenges:
-data timeliness, accuracy, and consistency, difficulties in
-correlating data over long periods, excluding personally
-identifiable information (PII), integration with stand-alone systems
-at selected NYC agencies, and managing API usage for numerous
-third-party users. These challenges are addressed by various agency
-open data managers and OTI.
-\end{comment}
+open data system and data portal. Input data for the 311 system 
+is sourced from 17 different NYC Agencies. Some of these Agencies
+utilize the core 311 software, while others transfer data to the
+311 Open Data Warehouse from in\mbox{-}house systems.
 
 The impact of NYC 311 data extends beyond operational efficiency; it
-has become instrumental in shaping city governance and community
+has become instrumental in shaping ity governance and community
 engagement. This open data not only ensures governmental transparency
@@
 public. The data has been pivotal in providing advice on shelters
-during emergencies, handling inquiries during the COVID-19 pandemic,
+during emergencies, handling inquiries during the COVID\mbox{-}19 pandemic,
 enforcing standards between landlords and tenants, reallocating taxi
@@
 
-Our investigation utilizes an SR dataset covering  2-years (2022-2023).
+Our investigation utilizes an SR dataset covering  2\mbox{-}years (2022-2023).
 This dataset is approximately 3.7 GB in size and contains over 
 6.4 million rows with each row representing one SR. It contains 
-41 columns of data. It is used to conduct detailed analysis 
+41 columns of data and was collected on 15 September 2023. The 
+dataset is used to conduct detailed analysis 
 in areas such as data consistency, data validity, redundant fields, 
@@
 
-During the 10-year period from 2014 to 2023,  311 experienced
-a 50\% in Service Requests, , reflecting a growing public reliance 
-on the system. (As a note, during this 10-year timeframe, the 
+During the 10\mbox{-}year period from 2014 to 2023,  311 experienced
+a 50\% increase in the number of Service Requests (SRs), 
+reflecting a growing public reliance and usage of 
+the 311 system. (During this 10\mbox{-}year time-frame, the 
 population of NYC grew just 0.6\%.). The most significant increase
-in activity occurred in 2020, when the number of requests 
-surged to 3.23 million during the COVID-19 pandemic, 
+in activity occurred in 2020, when the volume of requests 
+surged to 3.23 million during the COVID\mbox{-}19 pandemic, 
 illustrating the essential role of the 311 system in supporting 
-NYC residents during times of crisis. The steady 
+NYC residents during times of crisis. This steady 
 rise in SRs can be at least partially attributed to the increased 
@@
 indicates sustained growth in SR volume in future years,  
-highlighting the system's evolving role in managing 
-both routine city operations and extraordinary events.
+highlighting the system's expanding role in managing 
+both routine City operations and extraordinary events.
 
 Figure~\ref{fig:SRcountbyAgency} provides a breakdown of SRs by
-agency, showing the cumulative percentage of SRs handled by each
-agency over a the 2-year timeframe. Note that the
+the responsible Agency, showing the cumulative 
+percentage of SRs handled by each
+Agency over a the 2\mbox{-}year time-frame. Note that the
 distribution of SRs is heavily concentrated among a few key 
@@
 \begin{itemize}
-    \item New York Police Department (NYPD) - 43\%
+    \item New York Police Department (NYPD) - 43\% of all SRs
     \item Housing Preservation and Development (HPD) - 21\%
@@
 
-Collectively these six City agencies handle over 90\% of the total SR 
-volume, indicative of the critical role these agencies play in managing
+Collectively these six City Agencies handle over 90\% of the total SR 
+volume, indicative of the critical role these Agencies play in managing
 public concerns, ranging from noise complaints and housing issues to
-sanitation and transportation problems. The NYPD alone accounts for a
-substantial portion of the total SRs, underscoring its leading role in
-addressing complaints related to public safety. The remaining
-10\% of SRs is distributed across 10 other
-agencies. This concentration of SRs to the ``big six'' agencies 
-underscores the necessity for optimal data handling and curation processes 
-within these high-volume agencies in order to ensure 
-efficiency and responsiveness.
+sanitation and transportation problems. The remaining
+10\% of SRs is distributed across 10 additional Agencies. 
+This concentration of SRs to the ``big six'' Agencies 
+underscores the necessity for optimal data handling and 
+curation processes within these six high volume Agencies.
 
@@
 
+Figure~\ref{fig:SR_complaints}
+illustrates how the top 20 complaint types account for 70\% 
+of all SRs. Note that the distribution of complaints is skewed, 
+with a small number of issues dominating total SR volume. 
+Of special interest are noise\mbox{-}related complaints, of which there are eight different
+types including residential, commercial, and 
+street noise. When combined noise\mbox{-}related SRs make up 22\% of all 
+complaints; the most frequent issue in the NYC 311 system. 
+Other prominent categories include illegal parking, heat/hot water 
+complaints, and blocked driveways. The cumulative percentage curve 
+reveals that the top 20 complaint types comprise 70\% of SRs,
+while the remaining 190 complaint types are spread thinly 
+across the remaining 30\%. This suggests that improving 
+responses to the most frequent complaints could have an 
+outsized impact on overall service efficiency and resident 
+satisfaction. This chart also provides insights into the 
+operational pressures faced by City Agencies 
+responsible for handling these high-volume complaints.
+
 \begin{figure}[tbp]
@@
 
-Figure~\ref{fig:SR_complaints}
-illustrates how the top 20 complaint types, accounting for 70\% 
-of all SRs. Note that the distribution of complaints is skewed, 
-with a small number of issues dominating total SR volume. 
-Noise-related complaints, including residential, commercial, and 
-street noise, when combined make up 22\% of all 
-complaints; the most frequent issue in the NYC 311 system. 
-Other prominent categories include illegal parking, heat/hot water 
-complaints, and blocked driveways. The cumulative percentage curve 
-reveals that after the top 20 complaint types, the remaining 
-complaint types are spread thinly across the remaining 30\% of SRs,
-suggesting that improving responses to the most frequent 
-complaints could have an outsized impact on overall service 
-efficiency and resident satisfaction. This chart also provides 
-insights into the operational pressures faced by City agencies 
-responsible for handling these high-volume complaints.
-
 \section{NYC 311 SR Data Cleansing Issues} 
 \label{sec:issues}
-
 Data cleansing refers to the process of identifying and rectifying
@@
 especially important as open datasets often come from diverse,
-uncoordinated sources, leading to variations in data quality,
-completeness, and consistency. Without thorough data cleansing, the utility
-of open data can be limited and perhaps untrustworthy, affecting 
-its reliability for research, policy-making, and innovation. The 
-main purpose of cleansing open data is to ensure that it is 
-accurate, consistent, and usable across multiple platforms and by 
-various stakeholders with a myriad of purpose. Data cleansing improves
-the trustworthiness of the data and enables more accurate analysis, 
-better decision-making,and the integration of data into machine
-learning models or other systems.
-
-Many quality criteria are employed to ensure high-quality data 
+often uncoordinated sources, leading to variations in data quality,
+completeness, and consistency. Without thorough data cleansing, 
+the utility of open data can be limited and perhaps 
+untrustworthy, affecting its reliability for research, 
+policy making, and innovation. The main purpose of cleansing 
+open data is to ensure that it is accurate, consistent, and usable a
+cross multiple analytical platforms and by various stakeholders with a 
+myriad of purposes. Data cleansing improves the trustworthiness 
+of the data and enables more accurate analysis, 
+better decision making, and the improved integration of 
+data into machine learning models or other systems.
+
+Many quality criteria are employed to ensure high\mbox{-}quality data 
 processing. One of the primary efforts is data validation, which spans 
@@
 such as numeric, character, or date formats; typically 
-outlined in a Data Dictionary. An essential aspect of validation 
+outlined in a \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}.
+An
+essential aspect of validation 
 is domain compliance, where fields must adhere to a predefined set of 
-values, such as statuses, state names, or zip codes. Structural 
-errors also play a significant role in data quality, particularly when 
+values such as statuses, state names, or zip codes. Structural 
+errors can also play a significant role in data quality, particularly when 
 naming conventions or data entries are inconsistent. Common issues 
-include fields that do not appear in the Data Dictionary or the 
-inconsistent use of blanks, spaces, NA, N/A, or ``<NA>'' to indicate 
-missing data. Furthermore, redundant or irrelevant fields can clutter 
-datasets, reducing efficiency. Logical inconsistencies are another 
-important consideration, such as related fields that violate expected 
-relationships, such as a ``due date'' that precedes the ``created date.'' 
+include fields that do not appear in the Data Dictionary 
+and the inconsistent use of blanks, spaces, NA, N/A, 
+or ``<NA>'' to indicate missing data. Furthermore, redundant 
+or irrelevant fields can clutter datasets, reducing efficiency and 
+creating confusion or errors. Logical inconsistencies are another 
+consideration, such as related fields that violate expected 
+relationships, e.g. a ``due date'' that precedes the ``created date.'' 
 Lastly, the balance between accuracy and precision is crucial, as both 
-must be carefully managed to ensure reliable and meaningful data.
-
-Here we identify the presence of issues in the NYC 311 SR open data;
-not attempting to solve them but rather to highlight their scope and
-magnitued. A solution effort should be undertaken only after further 
-investigation as to the why and how such issues came about, 
-and a discussion as to whether or not it is an actual \textit{error} or 
+must be carefully managed to ensure meaningful data.
+
+Here we identify the presence of issues in the NYC 311 SR open dataset;
+not attempting to solve them, but rather to highlight their scope and
+magnitude. We would recommend that a solution effort 
+should be undertaken only after further investigation as to 
+the why and how such issues came about, including a
+discussion as to whether or not it is an actual \textit{error} or 
 holds some other status. Coordination with the originating Agency
-is paramount.
+is, of course, paramount.
 
@@
 \label{sec:structural}
-
 Structural issues refers to how data is organized, formatted, 
 or presented within a dataset. Structural issues can make 
-it difficult to analyze the data effectively. Here are some 
+it difficult to analyze the data effectively. Here are some overall 
 characteristics of the 311 SR data set:
@@
 	\item There are 41 columns of data for each row, exportable as a CSV file.
-	Each row represents one Service Request.
+	Each row represents a single Service Request (SR).
 
@@
 
-	\item Seven street fields; two pair of which appear to be duplicates
+	\item There are seven street fields; two pair of which appear to be duplicates
 
@@
 
-	\item A free-form text field, \texttt{resolution\_description}, which 
-	supports 934 characters of input, including commas and special characters
+	\item A free\mbox{-}form text field, \texttt{resolution\_description}, which 
+	supports 934 characters of input, including commas and special characters.
 \end{itemize}
@@
 visible in the portal's ``Column manager'' widget, labeled with the prefix 
-``@computed\_region.'' The additional fields include:
+``@computed\_region.'' and include:
 
@@
 
-While one can likely infer the meaning and 
-derivation of these computed fields, uncertainty remains as to
-their derivation and usage. These six fields are not included in
-the Data Dictionary, which we have asked to be updated. Our 
-analysis shows that these six \texttt{computed} fields have 
+While one can likely infer the meaning of these ``computed'' 
+fields, uncertainty remains as to their derivation and 
+usage. These six fields are not included in
+the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}.
+Our
+analysis shows that these six ``computed'' fields have 
 significant data validity issues, leading us to treat these six fields 
@@
 \texttt{due\_date}, you would be challenged as 99.6\% of the
-\texttt{due\_date} field is blank. When counting fields for 
+\texttt{due\_date} fields are blank. When sorting fields by
 blank or N/A values, they appear to divide into three groups:
@@
 \texttt{road\_ramp}, and \texttt{bridge\_highway\_name}.
-The Partially Empty includes such fields as
+The Partially Empty category includes such fields as
 \texttt{location\_type}, \texttt{landmark}, 
-and \texttt{cross\_street\_1 \& \_2}. And the Few/None Empty includes
-\texttt{created\_date}, \texttt{complaint\_type},
-\texttt{agency}, and \texttt{status}. In some analysis efforts, it may be
+and \texttt{cross\_street\_1 \& \_2}. And the Few/None Empty category
+ includes \texttt{created\_date}, \texttt{complaint\_type},
+\texttt{agency}, and \texttt{status}. For some analysis efforts, it may be
 prudent to inquire as to why some fields are mostly or almost always blank.
@@
 Any analytic effort must ensure that fields containing invalid values 
-are identified and isolated from the analysis. Below are the results of 
-selected field validations. The latitude and longitude fields were 
-found to fall within the geographic boundaries of New York City, and 
+are identified and isolated from the analysis. The \texttt{latitude} 
+and \texttt{longitude} fields were found to fall within the 
+geographic boundaries of New York City, and 
 the \texttt{unique\_key} field was indeed unique, as required. 
-Unfortunately, the Data Dictionary specifies very few domains of 
-acceptable values. Nonetheless, the following fields were tested and found 
-to comply with their expected domains, as determined by public usage 
-and historical datasets:
+Unfortunately, the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} specifies very few domains of
+acceptable values for other fields. Nonetheless, the following fields were tested and found to comply with their expected domains, as determined by public 
+usage and as evidenced in other historical datasets:
 
@@
 
-But some fields proved problematic as to their 
-compliance with a domain of allowable values.
-
-\paragraph{Zip Codes}
+However,  some fields proved problematic as to their 
+compliance with a domain of allowable values. These include:
+
+\paragraph{Zip Codes:}
 \label{sec:zipcodesissues}
-All zip codes ( \texttt{incident\_zip}) 
-should validate against the USPS database, which contains 
-44,173 valid zip codes. The \texttt{incident\_zip} field has 
+All \texttt{incident\_zip} zip codes should 
+validate against the USPS database, which contains 
+44,173 valid zip codes. And while the \texttt{incident\_zip} field has 
 only 0.07\% invalid entries, this still results in 4,374 errors; 
@@
 
-\paragraph{Created and Closed Date}
+\paragraph{Created and Closed Dates}
 \label{sec:negativeduration}
-For the \texttt{created\_date} and \texttt{closed\_date} fields, one 
-might expect that these fields to be automatically populated by the  
-application software when setting an SR status to ``new'' or ``closed.'' 
+With the \texttt{created\_date} and \texttt{closed\_date} fields, one 
+might expect that these fields would be automatically populated by the  
+application software. Thus when updating an SR status to ``new'' or ``closed''
+the appropriate software should automatically populate those fields. 
 Unfortunately, this does not seem to be the case, as several anomalies 
@@
     \item SRs with a \texttt{closed\_date} that occurs before the 
-    \texttt{created\_date} creating a ``negative duration''.
+    \texttt{created\_date} resulting a nonsensical ``negative duration''.
     \item \texttt{created\_date(s)} and \texttt{closed\_date(s)} in 
@@
     \item \texttt{created\_date(s)} and \texttt{closed\_date(s)} that 
-    match \textbf{to the second} indicating a nonsensical ``zero duration''.
+    match \textbf{to the second} creating an impossible ``zero duration''.
     \item An unusually large number of SRs closed and/or created exactly at midnight 
@@
 
-\subparagraph{Closed before Created: Negative Duration}
-Citizens, NYC Government Officials, and Agencies use the created and 
-closed dates to measure the \textit{duration} of SRs, which often reflects an Agency's 
-responsiveness. While duration is not directly present in the dataset, 
+\paragraph{Closed before Created: Negative Duration}
+Citizens, NYC Government Officials, and Agencies use the \texttt{created\_date}(s) and 
+\texttt{closed\_date}(s) to measure the \textit{duration} of SRs, 
+a useful surrogate to measure an Agency's responsiveness. 
+While duration is not directly present in the dataset, 
 it can be easily computed as the difference between
 \texttt{(closed\_date} and \texttt{created\_date}.  Duration is one of 
-most frequently used 311metrics and is constantly analyzed, such as determining if 
+most frequently used 311metrics and is frequently analyzed, such as determining if 
 one NYC borough receives faster response than another for certain 
@@
 \texttt{created\_date}, generating nonsensical ``negative durations.'' 
-Though this represents only 0.2\% of the dataset, these errors can 
+Although this represents only 0.2\% of the dataset, these errors can 
 significantly impact response time analyses. Eight SRs with extremely 
@@
 not addressed during data cleansing. As a result, these SR rows are 
-removed from the analysis. Here is a sample:
-
-\begin{table}[tbp]
-  \centering
-  \caption{Sample  negative durations [excluding extreme  values ($-$730 days)]}
-  \begin{tabular}{l l l r l}
-    \toprule
-    {created\_date} & {closed\_date} & {duration} 
-    & \textbf{agency} \\
-    \midrule
-    2023-01-27 14:40:00 & 2022-01-14 14:40:00 & $-$378.0 & DOT \\
-    2023-01-18 10:06:00 & 2022-01-12 10:06:00 & $-$371.0 & DOT \\
-    2023-01-27 14:36:00 & 2022-01-22 14:35:00 & $-$370.0 & DOT \\
-    2023-01-11 11:10:00 & 2022-01-09 11:10:00 & $-$367.0 & DOT \\
-    2023-12-18 03:13:00 & 2023-01-16 13:10:00 & $-$335.6 & DOT \\
-    \bottomrule
-  \end{tabular}
-  \label{tab:largest-errors}
-\end{table}
-
-Excluding the extreme negative values, 
+removed from the analysis.
+
+Excluding the extreme negative values ($-$730 days), 
 Figure~\ref{fig:negative-duration-violin} shows the broad spread of 
-negative-duration SRs. While there are few outliers, the magnitude 
-of the negative durations is troubling and can produce bizarre
-and unwanted analytical results.
+negative duration SRs. While there are few outliers, the magnitude 
+of these negative durations (300 days) is troubling and can produce 
+bizarre and unwanted analytical results.
 
@@
 
-\subparagraph{Identical Created Date \& Closed Date: Zero Durations}
+\begin{table}[tbp]
+  \centering
+  \caption{Sample  negative durations [excluding extreme  values that are$-$730 days]}
+   \label{tab:largest-errors}
+ 	\begin{tabular}{l l l r l}
+    \toprule
+    {created\_date} & {closed\_date} & {duration (days)}  & {agency} \\
+    \midrule
+    2023-01-27 14:40:00 & 2022-01-14 14:40:00 & $-$378 & DOT \\
+    2023-01-18 10:06:00 & 2022-01-12 10:06:00 & $-$371 & DOT \\
+    2023-01-27 14:36:00 & 2022-01-22 14:35:00 & $-$370 & DOT \\
+    2023-01-11 11:10:00 & 2022-01-09 11:10:00 & $-$367 & DOT \\
+    2023-12-18 03:13:00 & 2023-01-16 13:10:00 & $-$335 & DOT \\
+    \bottomrule
+    \end{tabular}
+ \end{table}
+
+\paragraph{Identical Created Date \& Closed Date: Zero Durations}
 A more prevalent issue occurs when the \texttt{closed\_date} and 
-\texttt{created\_date} are exactly the same, down to the second. This 
+\texttt{created\_date} are exactly the same, to the second. This 
 creates a \textbf{zero duration} for the SR, which is nonsensical. There 
-are 163,720 such SRs, representing 2.6\% of all non-blank data. 
-99\% percent of these zero-duration SRs occur within five agencies:
+are 163,720 zero duration SRs, representing 2.6\% of all non\mbox{-}blank data. 
+99\% percent of these zero duration SRs occur within five agencies:
 DHMH, DOT, DOB, DSNY, and DEP. This distribution \textbf{does not} 
 mirror the overall SR distribution by Agency, suggesting 
-an agency-specific issue.
-
-\subparagraph{Created Date or Closed Date(s) at Midnight or Noon}
-Another issue found with the \texttt{created\_date} and 
-\texttt{closed\_date} fields is the unusually large number of SRs 
-created or closed at exactly midnight (00:00:00) or noon (12:00:00), 
-down to the second. Normally, SR creation and closure follows the 
-work-day schedule, with most SRs created during daylight hours and 
+an Agency\mbox{-}specific issue.
+
+\paragraph{Created Date or Closed Date(s) at Midnight or Noon}
+Another issue involves the unusually large number of 
+SRs where the \texttt{created\_date} and \texttt{closed\_date} 
+indicate ``created'' or ``closed'' exactly at midnight (00:00:00) or 
+exactly at noon (12:00:00), down to the second. Normally, 
+SR creation and closure follows the work\mbox{-}day schedule, 
+with most SRs created during working hours and 
 fewer at night or in the early morning. However, there is a 
 significantly higher number of SRs closed exactly at midnight, 
-as well as a large number created exactly at midnight.
+as well as a large number created exactly at midnight. The magnitude
+of this anomaly is well outside the 3$\sigma$ level.
 
@@
 These unusual patterns of SR creation and closure exactly at midnight 
-and noon suggest the presence of a bulk create/close software process 
+suggest the presence of a bulk create/close software process, perhaps one
 that automatically assigns time stamps of midnight (00:00:00) or noon 
-(12:00:00) to large batches of SRs. This behavior distorts the 
-calculated duration of these SRs by providing inaccurate dates. The 
-distribution by agency shows that over 92\% of the ``closed-exactly-at-midnight'' 
-SRs come from just two agencies: DOB (69\%) and DSNY  (23\%).
-
-\subparagraph{Resolution Action Update Date}
-A similar issue occurs when the \texttt{resolution\_action\_update\_date} occurs well after the
-\texttt{closed\_date}.  When an SR is updated, the 311 software 
-automatically populates the \texttt{resolution\_action\_update\_date}. 
-While it is certainly possible, and even routine, to update an SR after it is closed, 
-some of these updates seem to be well beyond an expected 
-date, with some ``updates'' occurring over 2 years later. As shown in 
-Figure~\ref{fig:resolution-violin}, there are 10,265 SRs that 
-were updated >30 days after the 
-\texttt{closed\_date} (but <730 days, an arbitrary cutoff used to 
-exclude infeasible dates such as ``1900-01-01''. The chart 
-highlights the distribution of these post-closure updates, with 
-a notable concentration of updates occurring within 
-the 30-to-90-day range. This pattern raises questions about whether 
-such delayed updates are standard operational practice or indicative 
-of a potential issue which might warrant further investigation.
+(12:00:00) to large batches of SRs prior to sending the data to the 311 
+system. This behavior distorts the calculated duration of these 
+SRs by providing inaccurate dates. The distribution by agency 
+shows that over 92\% of the ``closed-exactly-at-midnight'' 
+SRs come from just two Agencies: DOB (69\%) and DSNY  (23\%).
+
+\paragraph{Resolution Action Update Date}
+A similar issue occurs when the \texttt{resolution\_action\_update\_date} 
+occurs well after the \texttt{closed\_date}.  When an SR is 
+updated, the 311 software automatically populates the \texttt{resolution\_action\_update\_date}. While it is certainly 
+possible, and even routine, to update an SR after it is closed, 
+some of these updates seem to be well beyond any normal or expected 
+timeframe, with some ``updates'' occurring over 2 years 
+later. As shown in Figure~\ref{fig:resolution-violin}, there 
+are 10,265 SRs that were updated >30 days after the 
+\texttt{closed\_date} (Note: The chart excludes values >730 days, 
+an arbitrary cutoff to exclude infeasible dates such as ``1900-01-01''.)
+ The chart highlights the distribution of these post-closure 
+ updates, with a notable concentration of updates occurring within 
+the 30\mbox{-}to\mbox{-}90 day range. This pattern raises 
+questions about whether such delayed updates are standard 
+operational practice or indicative of a potential issue 
+which might warrant further investigation. It should be noted
+that 84\% of these late updates are associated with the TLC agency
+and 12\% with DSNY.
 
@@
 \label{sec:precision}
-
 An question of precision vs. accuracy arises with the \texttt{latitude} 
-and \texttt{longitude} fields. Both Latitude are expressed as 
-a 14-decimal number, e.g. 40.86769186022511. Given 
+and \texttt{longitude} fields. as well as the duplicative 
+field \texttt{location}. These fields are expressed as 
+a 14\mbox{-}decimal number, e.g. 40.86769186022511. Given 
 that 1 degree of latitude at the equator is equal to 111.04 
 kilometers, the ``1'' at the end of that number represents 
-approximately 1.1 nanometers (1/1,000,000,000 of a meter). For 
-reference a DNA molecule is approximately 2nm in width. The 
-representation of the \texttt{latitude} and \texttt{longitude} fields appears to be a 
-classic case of extreme precision, but limited accuracy.
+approximately 1.1 nanometers (1/1,000,000,000 of a meter). (For 
+reference a DNA molecule is approximately 2nm in width.) The 
+representation of the \texttt{latitude} and \texttt{longitude} fields
+appears to be a classic case of extreme precision, but limited accuracy.
 
@@
 \label{sec:duplicates}
-
-During this analysis, several redundant fields were observed and should 
+During this analysis, several redundant fields were identified and should 
 be examined for possible consolidation.
 
-\texttt{Latitude/Longitude} and \texttt{location}: The \texttt{location} field 
-is a pure concatenation of the \texttt{latitude} and \texttt{longitude} fields except with a 
+\paragraph{\texttt{Latitude/Longitude} and \texttt{location}}: 
+The \texttt{location} field is a pure concatenation of 
+the \texttt{latitude} and \texttt{longitude} fields albeit with a 
 comma and parentheses added. This makes the \texttt{location} field 
@@
 
-\texttt{borough} \& \texttt{park\_borough}: These two fields are fully redundant, 
+\paragraph{\texttt{borough} \& \texttt{park\_borough}}: These two fields are fully redundant, 
 as they are 100\% matches.
 
-\texttt{borough} \& \texttt{taxi\_company\_borough}: Despite their names, these 
-fields are almost entirely different, with only a 0.05\% match. The 
-\texttt{taxi\_company\_borough} field is used exclusively by the TLC, 
-indicating a need for consultation with TLC to understand the differing 
-uses. It is likely that that the usage of \texttt{taxi\_company\_borough}
-is quite different than that of \texttt{borough}. Such information could be 
-useful only when studying TLC related SRs. It might make more
-sense to store TLC data in a separate dataset.
-
-\texttt{agency} \& \texttt{agency\_name}: The \texttt{agency} field contains 
-abbreviations of City agencies (e.g., NYPD, DOT, DSNY), while 
-\texttt{agency\_name} contains the full names. Including both seems 
+\paragraph{\texttt{borough} \& \texttt{taxi\_company\_borough}}: Despite 
+their names, these fields are almost entirely different, with only 
+a 0.05\% match. The \texttt{taxi\_company\_borough} field is 
+used exclusively by the TLC, indicating a need for consultation 
+with that Agency to understand the differing uses. It is 
+likely that that the usage of \texttt{taxi\_company\_borough} is 
+quite different than that of \texttt{borough}.
+
+\paragraph{\texttt{agency} \& \texttt{agency\_name}}: The \texttt{agency} 
+field contains the abbreviations of City agencies (e.g., NYPD, DOT, DSNY), 
+while \texttt{agency\_name} contains the full names. Including both seems 
 redundant, especially given the space taken up by full names in the 
-CSV file. Additionally, the agency abbreviations are easily recognized.
-
-\texttt{landmark} \& \texttt{street\_name}: Per the Data Dictionary, the
-\texttt{landmark} field ``can refer to any noteworthy location, 
-including but not limited to, parks, hospitals, airports, sports 
-facilities, performance spaces, etc.'  However, most entries 
-are actually street names, with a 62\% match to \texttt{street\_name}. 
-Even non-matches (excluding blanks) appear to technically be
-a match, differing only in spelling or formatting 
+CSV file. Additionally, the Agency abbreviations are easily recognized.
+
+\paragraph{\texttt{landmark} \& \texttt{street\_name}}: Per the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}, the \texttt{landmark} field ``can refer to any
+noteworthy location, including but not limited to, parks, 
+hospitals, airports, sports facilities, performance spaces, etc.'  
+However, most entries are actually street names, with a 62\% 
+match to \texttt{street\_name} field. Even the non\mbox{-}matches (excluding 
+blanks) appear to be a logical match, differing only in spelling or formatting 
 (e.g., ``NINTH AVE'' vs. ``9 AVE''). This indicates a possible 
@@
 
-\texttt{cross\_street\_1/2} \& \texttt{intersection\_street\_1/2}: These 
+\paragraph{\texttt{cross\_street\_1/2} \& \texttt{intersection\_street\_1/2}}: These 
 two sets of street pairs are used to help identify the 
-\texttt{incident address}. We found 88\% of these two street pairs to be 
-duplicates. This duplication raises the question of which field-pair 
-should be trusted; and if they are different, how should they be treated?
+\texttt{incident address}. We found 88\% of both of these two 
+street pairs to be duplicates. Such a high leve of duplication 
+raises the question of which field\mbox{-}pair should be trusted; 
+and if these street pairs are used differently, how should they be treated?
 It is possible that these two street pairs originated from different 
-agencies that used different variable names to represent the same 
+agencies that use different variable names to represent the same 
 information, but no documentation exists to clarify this 
-relationship. These two street-pairs are highly redundant.
+relationship. These two street\mbox{-}pairs are highly redundant.
 
@@
 \label{sec:filesize}
-
-By removing duplicate and ``near-duplicate'' fields, it is possible to 
+By removing duplicate and ``near\mbox{-}duplicate'' fields, it is possible to 
 shrink the file size by 39.5\%, which for this large dataset equates to 
-a reduction of 1.4 Gb, which is significant. A smaller file size means faster downloads, 
-less storage impact, and simpler data analysis efforts.
-
-Below is a list of duplicate and near-duplicate fields.
+a reduction of 1.4GB, a significant amount. A smaller file size 
+means faster downloads, less storage impact, and often 
+simpler data analysis efforts. Below is a list of duplicate 
+and near\mbox{-}duplicate fields which we believe could be
+removed from the dataset with minimal data loss.
 
@@
     \item \textbf{\texttt{agency\_name}}: Since each row contains the \texttt{agency} 
-    field, which uses abbreviations that are clear and well understood, we believe the
+    field, which uses clear and well understood abbreviations, we believe the
     \texttt{agency\_name} can be removed with minimal data loss.
@@
     \item \textbf{\texttt{location}}: This field is a straight concatenation of 
-    the latitude and longitude fields. It can be deleted without data loss.
+    the \texttt{latitude} and \texttt{longitude} fields. It can be 
+    deleted without data loss.
 
     \item \textbf{\texttt{cross\_street\_1/2}} and \textbf{\texttt{intersection\_street\_1/2}}: 
-    These two stree pairs both exhibit an 88\% match. We recommend deleting the 
+    These two street pairs both exhibit an 88\% match. We recommend deleting the 
     \texttt{intersection\_street\_1/2} fields while acknowledging 
     the concomitant data loss.
-
 \end{itemize}
@@
 Additionally, certain fields may not be useful for broad analysis
-due to their sparse data population. This especially applies to the fields below which
-are predominantly used by the TLC and are 99\% (or greater) blank. 
+due to their sparse data population. This especially applies to the fields below, 
+which are predominantly used by the TLC and are 99\% (or greater) blank.
+
 \begin{quote}
 \textbf{Note:} \textit{Even a field that is 99\% blank still contains 64,000 
-non-blank rows in this large (6.4 million-row) dataset.}
+non\mbox{-}blank rows in this large (6.4 million row) 
+dataset.} Non\mbox{-}trivial.
 \end{quote}
@@
 
-Further reduction is possible by encoding certain categorical 
-variables Using \texttt{complaint\_type} as an example,
-instead of repeatedly storing the full tex of each complaint
-type, it would use a numeric representation. 
-(e.g., ``NOISE - STREET/SIDEWALK'')could be represented by
-, a numeric code (e.g. 126). In this datase the example complaint type
-occurs over 300,000 times. Such encoding not only reduces file 
-size, but also improves processing efficiency. Moreover, 
-standardizing categorical inputs prevents the inclusion of 
-misspellings or unrecognized/invalid entries, ensuring cleaner 
-and more reliable data. This approach is particularly useful in large 
-datasets, where frequently repeated text-based categorical variables can 
+\paragraph{Field Encoding:} Further reduction is possible 
+by encoding selected categorical variables 
+Using \texttt{complaint\_type} as an example,
+instead of repeatedly storing the full text of each \texttt{complaint\_type}, 
+one could use a numeric representation. For example, 
+``NOISE - STREET/SIDEWALK'' could be represented by a 
+numeric code (e.g. ``126''). For this dataset the 
+``NOISE - STREET/SIDEWALK'' \texttt{complaint\_type} occurs 
+over 300,000 times. Such encoding not only reduces file 
+size, but would likely improve processing efficiency by using numeric
+values instead of text strings. Moreover, standardizing 
+categorical inputs prevents the inclusion of misspellings 
+or unrecognized/invalid entries, ensuring cleaner, more reliable 
+data. This approach is particularly useful in large datasets, where 
+frequently repeated text\mbox{-}based categorical variables can 
 significantly increase storage requirements.
@@
 \label{sec:datadictionary}
-
-The Data Dictionary would benefit from an update. We found several 
+The \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} would greatly benefit from an update.
+We found several
 discrepancies between the Data Dictionary and the actual data, which 
@@
 \begin{itemize}
-    \item While the Open Data portal lists the data types of various 
-    fields, the Data Dictionary does not. Nor does it provide details 
-    on specific fields such as constraints or the domain of legal 
-    values. For example, in the Data Dictionary, the \texttt{incident\_zip} 
+    \item While the NYC Open Data portal lists the data types of various 
+    fields, the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} does not.
+Nor does it provide details for specific fields such as constraints or the domain of legal
+    values.
+For example, in the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}, the \texttt{incident\_zip}
     field is specified as ``text.'' This may not be the most suitable 
@@
     \item Many of the domains of legal values are incomplete or 
-    inaccurate. For example, the Data Dictionary indicates that the 
+    inaccurate.
+For example, the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} indicates that the
     \texttt{status} field has values of \textit{assigned, canceled, 
@@
     values such as \textit{in progress, started}, and \textit{unspecified}. 
-    Additionally, no SRs were found with a status of 
-    \textit{canceled}.
+    Additionally, no SRs were found with a status of \textit{canceled}.
 
@@
 	\item As mentioned earlier, six \texttt{@computed} fields that are
-    present in the data but not included in the Data Dictionary need
-    to be documented and perhaps labelled as ``experimental'' or
+    present in the data but not included in the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} need
+    to be documented and perhaps labeled as ``experimental'' or
     ``not for official use''.
@@
 NYC 311 SR data, we propose a set of data 
-curation principles tailored for government-released open datasets. 
+curation principles tailored for government released open datasets. 
 These principles are designed to address the unique challenges 
@@
 \paragraph{Principle 1: Ensuring Consistency Across Agencies and Time}
-A key challenge in government-released datasets is the 
+A key challenge in government released datasets is the 
 harmonization of data fields across various agencies and 
@@
 Regular audits and data harmonization processes should be 
-implemented to ensure consistency across long-term datasets, 
+implemented to ensure consistency across long\mbox{-}term datasets, 
 particularly as agency structures or reporting standards evolve. 
@@
 Clear and accurate documentation is essential for the effective use 
-of government-released datasets. Our study found several 
+of government released datasets. Our study found several 
 issues where the Data Dictionary did not accurately reflect the 
@@
 NYC 311 case, issues such as large spikes in SR creation and closure 
-at exactly midnight (or noon) likely stem from a cron-job type 
+at exactly midnight (or noon) likely stem from a cron job type 
 processes that distort the accurate capture of SR 
-durations. By implementing real-time validation tools, governments 
+durations. By implementing real\mbox{-}time validation tools, governments 
 can prevent such anomalies from entering the dataset in the first 
@@
 understandable. This involves providing clear metadata and making 
-datasets available in user-friendly formats. Additionally, 
+datasets available in user\mbox{-}friendly formats. Additionally, 
 governments should communicate transparently about 
@@
 implement them effectively. Governments can begin by investing in 
-real-time validation systems, regular updates to data dictionaries, 
+real\mbox{-}time validation systems, regular updates to data dictionaries, 
 and tools for encoding and standardizing data fields. For example, 
@@
 critical in developing protocols for validating incoming data, correcting 
-errors in real-time, and ensuring consistency across multiple data sources. 
+errors in real\mbox{-}time, and ensuring consistency across multiple data sources. 
 Feedback from statisticians and data scientists, such as the issues reported 
@@
 should continue supporting these kinds of initiatives, which not only promote 
-data literacy but also encourage community-driven improvements to datasets. 
+data literacy but also encourage community\mbox{-}driven improvements to datasets. 
 By leveraging the expertise of data scientists, statisticians, stakeholders, 
@@
 
-\section{Discussion} \label{sec:discussion}
-
+\section{Discussion} 
+\label{sec:discussion}
 Our study highlights the critical role of robust data curation in
@@
 significantly undermine the insights derived from them. Trustworthy
-machine learning systems, which rely heavily on high-quality data, are
+machine learning systems, which rely heavily on high quality data, are
 especially vulnerable to such issues. Errors in the underlying data,
 such as biases or inconsistencies, can compromise both the accuracy
-and fairness of predictive models, impacting real-world decisions in
+and fairness of predictive models, impacting real\mbox{-}world decisions in
 urban planning and policy \citep{rahm2000data, geiger2020garbage}.
@@
 This research further emphasizes the necessity of harmonizing data
-across city agencies, particularly in resolving discrepancies in field
+across City Agencies, particularly concerning resolving discrepancies in field
 names, formats, field usage, and definitions. Consistency in data is essential for
@@
 lead to distorted outcomes. This challenge is even more pronounced in
-datasets that span long time periods, as changes in agency structures
-or reporting standards can introduce subtle biases. Long-term data
+datasets that span long time periods, as changes in Agency structures
+or reporting standards which can introduce subtle biases. Long\mbox{-}term data
 consistency is crucial for longitudinal studies and predictive
@@
 public policy applications opens new avenues for improving governance.
-High-quality data enables machine learning models to better predict
+High quality data enables machine learning models to better predict
 trends, allocate resources, and address service delivery issues, such
-as delays in responding to 311 complaints. The COVID-19 pandemic
-underscored the importance of real-time, trustworthy data in crisis
-management, where the accuracy and timeliness of data-driven insights
+as delays in responding to 311 complaints. The COVID\mbox{-}19 pandemic
+underscored the importance of real\mbox{-}time, trustworthy data in crisis
+management, where the accuracy and timeliness of data\mbox{-}driven insights
 were critical for public health responses. Without proper data
@@
 Future efforts could focus on automating data curation processes,
-particularly in real-time data pipelines, to ensure that the data used
+particularly in real\mbox{-}time data pipelines, to ensure that the data used
 in machine learning models remains accurate, clean, and reliable

### 2024-11-07 manuscript/manuscript.tex
Further edting
Added enumitem package and idented itemized lists. Rearranged a couple of paragraphs for better organization. Cosmetic changes to standardize presentation and improve readability.

@@
 \usepackage{subcaption}
+\usepackage{enumitem}
 
@@
 NYC Open Data Portal are available in the supplemental material.
+Here are some characteristics of the 311 SR data set:
+
+\begin{itemize}[left=1.5em]
+	\item There are 41 columns of data for each row, exportable as a CSV file.
+	Each row represents a single Service Request (SR).
+
+	\item There are four date fields 
+	(\texttt{created\_date}, \texttt{closed\_date}, \texttt{resolution\_action\_updated\_date}, \& \texttt{due\_date}).
+
+	\item There are two borough fields; one of which appear to be a duplicate.
+
+	\item There are seven street fields; two pair of which appear to be duplicates
+
+	\item In addition to  \texttt{incident\_address}, there are five additional location fields: 
+	\texttt{latitude \& longitude}, \texttt{street\_name}, \texttt{landmark}, USGS State Plane 
+	Coordinate System, and \texttt{block} \#.
+
+	\item A free\mbox{-}form text field, \texttt{resolution\_description}, which 
+	supports 934 characters of input, including commas and special characters.
+\end{itemize}
 
@@
 
-\begin{itemize}
-    \item New York Police Department (NYPD) - 43\% of all SRs
-    \item Housing Preservation and Development (HPD) - 21\%
-    \item New York City Department of Sanitation (DSNY) - 10\%
-    \item Department of Transportation (DOT) - 7\%
-    \item Department of Environmental Protection (DEP) - 5\%
-    \item Department of Parks and Recreation (DPR) - 4\%
+\begin{itemize}[left=1.5em]
+    \item New York Police Department (NYPD): 43\% of all SRs
+    \item Housing Preservation and Development (HPD): 21\%
+    \item New York City Department of Sanitation (DSNY): 10\%
+    \item Department of Transportation (DOT): 7\%
+    \item Department of Environmental Protection (DEP): 5\%
+    \item Department of Parks and Recreation (DPR): 4\%
 \end{itemize}
@@
 
-\section{NYC 311 SR Data Cleansing Issues} 
+\section{Data Cleansing Issues} 
 \label{sec:issues}
@@
 An
-essential aspect of validation 
-is domain compliance, where fields must adhere to a predefined set of 
+essential aspect of validation is domain compliance, where 
+fields must adhere to a predefined set of 
 values such as statuses, state names, or zip codes. Structural 
@@
 naming conventions or data entries are inconsistent. Common issues 
-include fields that do not appear in the Data Dictionary 
-and the inconsistent use of blanks, spaces, NA, N/A, 
-or ``<NA>'' to indicate missing data. Furthermore, redundant 
-or irrelevant fields can clutter datasets, reducing efficiency and 
-creating confusion or errors. Logical inconsistencies are another 
-consideration, such as related fields that violate expected 
+include fields that do not appear in the Data Dictionary,  
+inconsistent use of data fields. Furthermore, redundant 
+or irrelevant fields can clutter datasets, reducing efficiency, 
+creating confusion, and introducing errors. Logical inconsistencies 
+are another consideration, such as related fields that violate expected 
 relationships, e.g. a ``due date'' that precedes the ``created date.'' 
@@
 
-Here we identify the presence of issues in the NYC 311 SR open dataset;
-not attempting to solve them, but rather to highlight their scope and
-magnitude. We would recommend that a solution effort 
-should be undertaken only after further investigation as to 
-the why and how such issues came about, including a
-discussion as to whether or not it is an actual \textit{error} or 
-holds some other status. Coordination with the originating Agency
-is, of course, paramount.
+We identify the presence of issues in the NYC 311 SR open dataset;
+not attempting to solve them, but rather to highlight their scope, 
+magnitude, and potential impact. We would recommend 
+that a corrective solution should be undertaken only 
+after further investigation as to the why and how such 
+issues might come about, including  detailed discussions 
+with the originating Agency as to whether or not it is an 
+actual error, correct,  or holds some other status.
 
@@
 Structural issues refers to how data is organized, formatted, 
-or presented within a dataset. Structural issues can make 
-it difficult to analyze the data effectively. Here are some overall 
-characteristics of the 311 SR data set:
-
-\begin{itemize}
-	\item There are 41 columns of data for each row, exportable as a CSV file.
-	Each row represents a single Service Request (SR).
-
-	\item There are four date fields 
-	(\texttt{created\_date}, \texttt{closed\_date}, \texttt{resolution\_action\_updated\_date}, \& \texttt{due\_date}).
-
-	\item There are two borough fields; one of which appear to be a duplicate.
-
-	\item There are seven street fields; two pair of which appear to be duplicates
-
-	\item In addition to  \texttt{incident\_address}, there are five additional location fields: 
-	\texttt{latitude \& longitude}, \texttt{street\_name}, \texttt{landmark}, USGS State Plane 
-	Coordinate System, and \texttt{block} \#.
-
-	\item A free\mbox{-}form text field, \texttt{resolution\_description}, which 
-	supports 934 characters of input, including commas and special characters.
-\end{itemize}
-
-The 311 SR 
+or presented within a dataset. They can make 
+it difficult to analyze the data effectively and often require extensive 
+data cleaning efforts. Here are three structural issues we encountered.
+
+\paragraph{Undocumented Fields:} The 311 SR 
 \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}
@@
 
-\begin{itemize}
+\begin{itemize}[left=1.5em]
     \item \texttt{zip\_codes}
@@
 the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}.
-Our
-analysis shows that these six ``computed'' fields have 
+Additionally,
+our analysis shows that these six ``computed'' fields have 
 significant data validity issues, leading us to treat these six fields 
-as experimental and not for suitable for normal use. Accordingly, 
+as experimental and \textit{not for suitable for public use}. An update 
+to the Data Dictionary would seem appropriate. Accordingly, 
 we will not address these fields in our analysis.
 
-\subsection{Missing Data}
+\paragraph{Representation of Missing data:} We found that the 
+311 SR dataset exhibited inconsistent approaches when 
+presenting missing data. We discovered many different 
+representations of missing data: nulls, spaces, ``NA'', ``N/A'', 
+and ``<NA>''. This variety of representations complicates programming
+efforts, and increases the likelihood of introducing errors 
+in data analysis (typically by search for or excluding missing fields 
+and not aware of the various representations). A more consistent 
+approach to treating missing data should be employed to address this issue.
+
+\paragraph{Challenges with daylight saving time (DST):} The 311 SR 
+dataset captures local NYC times. Accordingly, twice each 
+year there are changes to the local time as a result of observing 
+DST, and this can introduce some unexpected results. For example, 
+on March 13, 2022 and March 12, 2023 there are no times 
+between 1:59AM and 2:59AM: the 2:00AM hour on those dates 
+is simply missing owing to the clocks moving forward; at 2:00AM 
+it instantly becomes 3:00AM. This creates a gap in the hourly 
+distribution of SRs on those days, and it exaggerates the 
+lifespan of SRs, albeit by a single hour. Similarly on 
+November 6, 2022 and November 5, 2023, the 
+hour of 1:00AM - 2:00AM is repeated twice. This can result 
+in a bizarre scenario where, for example, an SR can 
+be created at 1:59AM and then closed say 30 minutes later 
+at 1:29AM; seemingly nonsensical. Utilizing UTC time 
+for date\mbox{-}time stamps might help address this issue.
+
+\subsection{Missing Data by Field}
 \label{sec:blanks}
@@
 \texttt{due\_date}, you would be challenged as 99.6\% of the
-\texttt{due\_date} fields are blank. When sorting fields by
-blank or N/A values, they appear to divide into three groups:
-
-\begin{itemize}
-    \item \textbf{Mostly Empty} -- 93-99.9\% blank 
-    \item \textbf{Partially Empty} -- 40-4\% blank
-    \item \textbf{Few/None Empty} -- 2-0\% blank
+\texttt{due\_date} field are blank. When sorting fields by
+blank (or N/A values) they appear to divide into three groups:
+
+\begin{itemize}[left=1.5em]
+    \item \textbf{Mostly Empty}: 93-99.9\% blank 
+    \item \textbf{Partially Empty}: 40-4\% blank
+    \item \textbf{Few/None Empty}: 2-0\% blank
 \end{itemize}
 
-The Mostly Empty category includes such fields as
+The \textit{Mostly Empty} category includes fields such as
 \texttt{taxi\_company\_borough}, \texttt{due\_date},
 \texttt{road\_ramp}, and \texttt{bridge\_highway\_name}.
-The Partially Empty category includes such fields as
+The \textit{Partially Empty} category includes fields such as
 \texttt{location\_type}, \texttt{landmark}, 
-and \texttt{cross\_street\_1 \& \_2}. And the Few/None Empty category
- includes \texttt{created\_date}, \texttt{complaint\_type},
+and \texttt{cross\_street\_1 \& \_2}. And the \textit{Few/None Empty} 
+category includes \texttt{created\_date}, \texttt{complaint\_type},
 \texttt{agency}, and \texttt{status}. For some analysis efforts, it may be
-prudent to inquire as to why some fields are mostly or almost always blank.
-Figure~\ref{fig:blank_fields} presents a graphic 
-depiction of total empty (blank \& N/As) for each fields illustrating
+prudent to inquire further as to why some fields are 
+mostly or almost always blank. Figure~\ref{fig:blank_fields} is 
+a graphic depiction of total empty (blank \& N/As) for each field illustrating
 the grouping into the Most, Partial, and Few categories.
@@
 
-\begin{itemize}
+\begin{itemize}[left=1.5em]
     \item \texttt{address\_type}
@@
 
-\begin{itemize}
+\begin{itemize}[left=1.5em]
     \item SRs with a \texttt{closed\_date} that occurs before the 
@@
 
-\paragraph{\texttt{Latitude/Longitude} and \texttt{location}}: 
+\paragraph{\texttt{latitude/longitude} \& \texttt{location}:} 
 The \texttt{location} field is a pure concatenation of 
@@
 
-\paragraph{\texttt{borough} \& \texttt{park\_borough}}: These two fields are fully redundant, 
+\paragraph{\texttt{borough} \& \texttt{park\_borough}:} These two fields are fully redundant, 
 as they are 100\% matches.
 
-\paragraph{\texttt{borough} \& \texttt{taxi\_company\_borough}}: Despite 
+\paragraph{\texttt{borough} \& \texttt{taxi\_company\_borough}:} Despite 
 their names, these fields are almost entirely different, with only 
@@
 
-\paragraph{\texttt{agency} \& \texttt{agency\_name}}: The \texttt{agency} 
+\paragraph{\texttt{agency} \& \texttt{agency\_name}:} The \texttt{agency} 
 field contains the abbreviations of City agencies (e.g., NYPD, DOT, DSNY), 
@@
 
-\paragraph{\texttt{landmark} \& \texttt{street\_name}}: Per the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}, the \texttt{landmark} field ``can refer to any
-noteworthy location, including but not limited to, parks, 
-hospitals, airports, sports facilities, performance spaces, etc.'  
-However, most entries are actually street names, with a 62\% 
-match to \texttt{street\_name} field. Even the non\mbox{-}matches (excluding 
-blanks) appear to be a logical match, differing only in spelling or formatting 
+\paragraph{\texttt{landmark} \& \texttt{street\_name}:} Per the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary},
+the \texttt{landmark} field ``can refer to any noteworthy location, 
+including but not limited to, parks, hospitals, airports, sports facilities, 
+performance spaces, etc.'. However, most entries are actually 
+street names, with a 62\% match to \texttt{street\_name} field. 
+Even the non\mbox{-}matches (excluding blanks) appear to be 
+a logical match, differing only in spelling or formatting 
 (e.g., ``NINTH AVE'' vs. ``9 AVE''). This indicates a possible 
@@
 
-\paragraph{\texttt{cross\_street\_1/2} \& \texttt{intersection\_street\_1/2}}: These 
+\paragraph{\texttt{cross\_street\_1/2} \& \texttt{intersection\_street\_1/2}:} These 
 two sets of street pairs are used to help identify the 
 \texttt{incident address}. We found 88\% of both of these two 
-street pairs to be duplicates. Such a high leve of duplication 
+street pairs to be duplicates. Such a high level of duplication 
 raises the question of which field\mbox{-}pair should be trusted; 
@@
 
-\subsection{Storage}
+\subsection{Reducing File Size}
 \label{sec:filesize}
@@
 
-\begin{itemize}
+\begin{itemize}[left=1.5em]
     \item \textbf{\texttt{agency\_name}}: Since each row contains the \texttt{agency} 
@@
 \begin{quote}
-\textbf{Note:} \textit{Even a field that is 99\% blank still contains 64,000 
+\textbf{Note:} \textit{Even a field that is 99\% blank will contain 64,000 
 non\mbox{-}blank rows in this large (6.4 million row) 
@@
 
-\begin{itemize}
+\begin{itemize}[left=1.5em]
     \item \texttt{taxi\_company\_borough}
@@
 
-\begin{itemize}
+\begin{itemize}[left=1.5em]
     \item While the NYC Open Data portal lists the data types of various 

### 2024-11-11 manuscript/manuscript.tex
Completed editing
Minor grammatical and sentence structure correction. Ready for additional review.

@@
 Figure~\ref{fig:SRcountbyAgency} provides a breakdown of SRs by
-the responsible Agency, showing the cumulative 
+the responsible agency, showing the cumulative 
 percentage of SRs handled by each
-Agency over a the 2\mbox{-}year time-frame. Note that the
+agency over a the 2\mbox{-}year time-frame. Note that the
 distribution of SRs is heavily concentrated among a few key 
@@
 	\includegraphics[width = \textwidth]{SRs_by_Agency.pdf}
-  	\caption{SR counts by Agency with Cumulative Percentage}
+  	\caption{SR counts by agency with Cumulative Percentage}
 	\label{fig:SRcountbyAgency}
@@
 issues might come about, including  detailed discussions 
-with the originating Agency as to whether or not it is an 
+with the originating agency as to whether or not it is an 
 actual error, correct,  or holds some other status.
@@
 not insignificant. If you group the invalid entries by 
-percentage by Agency, it mirrors the \textit{overall} 
-breakdown of SRs by Agency, potentially 
+percentage by agency, it mirrors the \textit{overall} 
+breakdown of SRs by agency, potentially 
 indicating a systemic problem.
@@
 \texttt{closed\_date}(s) to measure the \textit{duration} of SRs, 
-a useful surrogate to measure an Agency's responsiveness. 
+a useful surrogate to measure an agency's responsiveness. 
 While duration is not directly present in the dataset, 
@@
 DHMH, DOT, DOB, DSNY, and DEP. This distribution \textbf{does not} 
-mirror the overall SR distribution by Agency, suggesting 
-an Agency\mbox{-}specific issue.
+mirror the overall SR distribution by agency, suggesting 
+an agency\mbox{-}specific issue.
 
@@
 used exclusively by the TLC, indicating a need for consultation 
-with that Agency to understand the differing uses. It is 
+with that agency to understand the differing uses. It is 
 likely that that the usage of \texttt{taxi\_company\_borough} is 
@@
 redundant, especially given the space taken up by full names in the 
-CSV file. Additionally, the Agency abbreviations are easily recognized.
+CSV file. Additionally, the agency abbreviations are easily recognized.
 
@@
 A key challenge in government released datasets is the 
-harmonization of data fields across various agencies and 
+harmonization of data fields across various Agencies and 
 maintaining this consistency over time. To address this, agencies should adopt 
@@
 field naming conventions over a span of years can prevent 
-inconsistent analyses.
-
-\paragraph{Principle 2: Guaranteeing Accuracy and Validity in Data Entry}
+inconsistent analyses. These standards should be promulgated 
+by the Open Data governing authority, following agreements 
+by the relevant parties.
+
+\paragraph{Principle 2: Guaranteeing Accuracy and Validity during Data Entry}
 Data accuracy and validity are fundamental to the utility of 
 open datasets. In the NYC 311 data, issues such as 
-invalid zip codes and 
-nonsensical date entries (e.g., negative durations) highlight the need 
-for rigorous validation processes. Governments should establish 
-clear protocols for data entry, ensuring that fields are populated with 
-valid, legally acceptable values. For example, fields such as 
-\texttt{created\_date} and \texttt{closed\_date} should be subject to 
-automated checks that prevent logical errors, such as closing an 
-SR before it was opened. Ensuring the integrity of critical fields 
-like geographic locations and agency names enhances the quality 
-and reliability of the data. Additionally, a number of fields have 
-very defined domains, such as \texttt{incident\_zip}. Such fields
-should be subject to data entry validation, in this example
-validating against the USPS database at the data entry point.
+invalid zip codes and nonsensical date entries 
+highlight the need for rigorous validation processes. Governments 
+should establish clear protocols for data entry, ensuring that 
+fields are populated with valid, legally acceptable values. To 
+the maximum extent possible,the software should assist in 
+ensuring data entries are accurate and valid. Some examples include:
+
+\begin{itemize}[left=1.5em]
+    \item Certain date fields should be promulgated by the software
+    system versus manual entry. Certain dates should be populate 
+    based upon an action in the system, such as a change in status.
+
+    \item Select fields should be subjected to automated checks 
+    that prevent logical errors, such as closing opening.
+
+    \item A number of fields have clearly defined domains of legal
+    values. These fields should be validated against a reference 
+    dataset upon entry. 
+ \end{itemize}
 
@@
 Efficient storage and data representation are crucial for handling 
-large government datasets. By eliminating redundant fields and 
-encoding categorical variables, agencies can significantly reduce 
-file sizes and improve the efficiency of data analysis. In the NYC 
-311 dataset, for example, redundant fields such as \texttt{borough} 
-and \texttt{park\_borough} could be consolidated without any data 
-loss, saving space and reducing file size. Another 
-example is the \texttt{taxi\_company\_borough} field, which is 99.94\% 
-blank and should be evaluated to determine if its inclusion is 
-necessary. Similarly, categorical variables like \texttt{complaint\_type} 
-could be encoded in a standardized format to optimize storage and 
-streamline analyses. Reducing file size not only saves storage space 
-but also enhances the performance of queries and analysis processes.
-
-\paragraph{Principle 4: Updating Data Dictionaries and Documentation}
-Clear and accurate documentation is essential for the effective use 
+large government datasets. Our research found several duplicate and
+near\mbox{-}duplicate fields in the 311 SR dataset. Eliminating 
+redundant fields and encoding categorical variables can 
+significantly reduce file sizes and improve the efficiency 
+of data analysis. Similarly, categorical variables could be 
+encoded in a standardized format to optimize storage and 
+streamline analyses. Reducing file size not only saves 
+storage space but also enhances the performance of queries, 
+downloads,  and analysis processes.
+
+\paragraph{Principle 4: Updated Data Dictionaries and Documentation}
+Clear and accurate data documentation is essential for the effective use 
 of government released datasets. Our study found several 
-issues where the Data Dictionary did not accurately reflect the 
-actual data, such as missing field descriptions for computed fields, and 
-inaccurate domain values for categorial fields such as \texttt{status}. 
-As mentioned, updating the Data Dictionary to include descriptions of all fields, 
-including the computed fields, e.g. \texttt{@computed\_region\_xxxx}, 
-would improve transparency and usability. Regular updates should 
+instances where the Data Dictionary did not accurately reflect the 
+actual data. Domains of legal/allowable values should be referenced
+in the Data Dictionary. Regular reviews and updates should 
 occur whenever there are changes in data structure, data usage, or 
@@
 Automating data validation and quality assurance processes can 
-significantly improve the reliability of government datasets. In the 
-NYC 311 case, issues such as large spikes in SR creation and closure 
-at exactly midnight (or noon) likely stem from a cron job type 
-processes that distort the accurate capture of SR 
-durations. By implementing real\mbox{-}time validation tools, governments 
-can prevent such anomalies from entering the dataset in the first 
-place. Automated 
+significantly improve the reliability of government datasets. Certain
+unusual trends can be identified, and irregularities highlighted. 
+In the NYC 311 case, issues such as large spikes in SR 
+creation and closure at exactly midnight (or noon) likely 
+stem from a cron job type processes that distort the 
+accurate capture of SR durations. By implementing 
+real\mbox{-}time and event\mbox{-}driven validation tools, governments 
+can both prevent such errors from entering the dataset in the first 
+place, as well as highlight anomalous trends. Automated 
 systems can be designed to flag inconsistencies, incorrect values, 
-or illogical sequences, or illogical date entries. 
-Automation can also facilitate faster detection and correction of 
-errors, reducing the need for manual intervention and ensuring 
-the data remains clean and reliable as it is continuously updated.
+illogical sequences, or illogical data entries. Automation can 
+also facilitate faster detection and correction of errors, reducing 
+the need for manual intervention and ensuring the data 
+remains clean and reliable as it is continuously updated.
 
@@
 Transparency is a core tenet of open data, and governments must ensure 
-that datasets are not only accurate but also accessible and easily 
-understandable. This involves providing clear metadata and making 
-datasets available in user\mbox{-}friendly formats. Additionally, 
-governments should communicate transparently about 
+that datasets are not only accurate but also easily accessible and 
+readily understandable. This involves providing clear 
+metadata and making datasets available in user\mbox{-}friendly 
+formats. Additionally, governments should communicate transparently about 
 data quality issues, such as notifying the public when errors in SR 
@@
 real\mbox{-}time validation systems, regular updates to data dictionaries, 
-and tools for encoding and standardizing data fields. For example, 
-systems that automate the enforcement of data validation rules 
-(e.g., flagging incorrect zip codes) can significantly improve data 
+and tools for encoding and standardizing data fields. This would include
+systems that automate enforcement of data validation rules 
+(e.g., flagging incorrect zip codes), significantly improving data 
 quality. Collaboration with city agencies and stakeholders is crucial 
@@
 Feedback from statisticians and data scientists, such as the issues reported 
-in this paper, needs a clear and accessible channel to reach the 311 SR data 
+in this paper, benefit from a clear and accessible channel to the 311 SR data 
 team and other relevant city agencies. Establishing such feedback loops 
-ensures that recommendations for improving data curation and validation 
-are not only heard but implemented. This feedback mechanism would also allow 
-ongoing dialogue between data users and city officials, leading to more dynamic 
+and other opportunities for public engagement, ensures that 
+recommendations for improving data curation and validation 
+are not only heard and discussed, but implemented. This feedback 
+mechanism would also allow ongoing dialogue between 
+data users and city officials, leading to more dynamic 
 and responsive data management practices.
@@
 inspired by the potential of open data to enhance public services. Governments 
-should continue supporting these kinds of initiatives, which not only promote 
-data literacy but also encourage community\mbox{-}driven improvements to datasets. 
+should continue supporting these types of initiatives, which not only promote 
+data literacy, but encourage community\mbox{-}driven improvements to datasets. 
 By leveraging the expertise of data scientists, statisticians, stakeholders, 
@@
 This research further emphasizes the necessity of harmonizing data
-across City Agencies, particularly concerning resolving discrepancies in field
-names, formats, field usage, and definitions. Consistency in data is essential for
+across City agencies, particularly resolving discrepancies in field
+names, formats, usage, and definitions. Consistency in data is essential for
 trustworthy machine learning, where inconsistencies across sources can
 lead to distorted outcomes. This challenge is even more pronounced in
-datasets that span long time periods, as changes in Agency structures
+datasets that span long time periods, as changes in agency structures
 or reporting standards which can introduce subtle biases. Long\mbox{-}term data

### 2024-11-19 manuscript/manuscript.tex
Top to bottom read through
A number of minor edits, formatting, etc. Nearly done.

@@
 
-\hyphenpenalty=900
+\hyphenpenalty=950
 
@@
 Data Law in 2012, leading to the creation of the NYC Open Data
-portal. This portal now hosts 2,700 datasets from 80 city agencies,
+portal. This portal currently hosts 2,700 datasets,
 serving as a crucial resource for research across various domains, 
 including health, urban development, and transportation. The 
-success of these initiatives highlights the importance of data 
+success of this initiative highlights the importance of data 
 curation in ensuring the utility and reliability of open datasets.
@@
 \href{https://www.data.gov}{Data.gov} portal in 2009, a pioneering
-platform in making government data widely accessible. Shortly after,
+platform in making government data widely accessible. Shortly afterwards,
 the European Union followed suit, unveiling its
@@
 These initiatives represent significant strides in democratizing data, 
-breaking down barriers that once kept valuable information 
+removing barriers that once kept valuable information 
 regarding government performance in silos. Their collective impact 
@@
   portal}, which today hosts an impressive array of 2,700 datasets
-from 80 different city agencies. This resource has become invaluable
+across 80 different city agencies. This resource has become invaluable
 for researchers across various fields and has significantly enhanced
@@
 college enrollment statistics, jail inmate charges, and the location
-of city\mbox{-}wide free Internet access points. These datasets have been
+of City\mbox{-}wide free Internet access points. These datasets have been
 applied in civil life in various ways, such as mapping car crashes
@@
 initiatives have made vast amounts of data available, ensuring the 
-reliability of such data hinges on rigorous data cleaning processes. 
+reliability and utility of such data hinges on rigorous data cleaning processes. 
 Efficient usage of software tools can significantly streamline this 
@@
 and the presence of missing, blank, or N/A entries. This section
- also investigates field compliance with reference or acceptable values, 
+ also investigates data field compliance with reference or acceptable values, 
 highlights logical inconsistencies, and examines some concerning patterns 
@@
 
-\section{NYC 311 SR Data} 
+\section{NYC 311 Service Request (SR) Data} 
 \label{sec:data}
@@
 include the addition of online and mobile app channels in 2009, a
-record high of 348,463 monthly service requests in August 2020 due to
-the COVID\mbox{-}19 pandemic, the 2021 expansion to include the MTA's city
-subway system (the biggest expansion in their history), and a 
-record 3.23 million service requests in 2023.
+record high of 348,463 monthly service requests in August 2020 associated 
+with the the COVID\mbox{-}19 pandemic, the 2021 expansion 
+to include the MTA's city subway system (the biggest 
+expansion in 311  history), and a record 3.23 million 
+service requests in 2023.
 
@@
 \href{https://www.nyc.gov/content/oti/pages/}{NYC Office of Technology 
-  and Innovation (OTI)}, provides technological support for the
-open data system and data portal. Input data for the 311 system 
-is sourced from 17 different NYC Agencies. Some of these Agencies
+and Innovation (OTI)}, provides technical support for 
+the 311 application, the open data infrastructure, mobile \& web applications, 
+and the public\mbox{-}facing data portal. Input data for the 311 system 
+is sourced from 16 different NYC Agencies. Some of these Agencies
 utilize the core 311 software, while others transfer data to the
-311 Open Data Warehouse from in\mbox{-}house systems.
+311 Open Data Warehouse from their in\mbox{-}house systems.
 
 The impact of NYC 311 data extends beyond operational efficiency; it
-has become instrumental in shaping ity governance and community
-engagement. This open data not only ensures governmental transparency
+has become instrumental in shaping City governance and community
+engagement. Open data not only ensures governmental transparency
 but also empowers researchers, civic developers, and the general
@@
 routes based on analyses by the Taxi and Limousine Commission (TLC),
-and improving responsiveness across City Agencies
-
-Our investigation utilizes an SR dataset covering  2\mbox{-}years (2022-2023).
-This dataset is approximately 3.7 GB in size and contains over 
-6.4 million rows with each row representing one SR. It contains 
-41 columns of data and was collected on 15 September 2023. The 
-dataset is used to conduct detailed analysis 
-in areas such as data consistency, data validity, redundant fields, 
-etc.. Instructions on downloading this file from the 
-NYC Open Data Portal are available in the supplemental material.
-Here are some characteristics of the 311 SR data set:
+and improving responsiveness across City Agencies.
+
+Our investigation utilizes a 311 SR dataset covering  2\mbox{-}years (2022-2023).
+ This sample was collected on 15 September 2024, and is used 
+ to conduct detailed analysis in areas such as data 
+ consistency, data validity, redundant fields, etc.. Instructions 
+ on downloading this file from the NYC Open Data Portal 
+ are available in the supplemental material. Here are some 
+ characteristics of the 311 SR dataset used in this analysis:
 
 \begin{itemize}[left=1.5em]
-	\item There are 41 columns of data for each row, exportable as a CSV file.
-	Each row represents a single Service Request (SR).
-
-	\item There are four date fields 
-	(\texttt{created\_date}, \texttt{closed\_date}, \texttt{resolution\_action\_updated\_date}, \& \texttt{due\_date}).
-
-	\item There are two borough fields; one of which appear to be a duplicate.
+	\item The dataset is approximately 3.7 GB in size and contains 
+	over 6.4 million rows. It was exported as a CSV file.
+
+	\item There are 41 columns of data for each row, with each
+	row representing a single Service Request (SR).
+
+	\item There are four date fields: \texttt{created\_date}, \texttt{closed\_date}, 
+	\texttt{resolution\_action\_updated\_date}, \& \texttt{due\_date}.
+
+	\item There are two borough fields; one of which appears to be a duplicate.
 
@@
 	\item In addition to  \texttt{incident\_address}, there are five additional location fields:
+
 	\texttt{latitude \& longitude}, \texttt{street\_name}, \texttt{landmark}, USGS State Plane 
-	Coordinate System, and \texttt{block} \#.
+	Coordinate System, and \texttt{block} \textbf{\#}.
 
 	\item A free\mbox{-}form text field, \texttt{resolution\_description}, which 
-	supports 934 characters of input, including commas and special characters.
+	supports 930 characters of input, including commas and 
+	special characters  which can be problematic for automated processing.
 \end{itemize}
 
-During the 10\mbox{-}year period from 2014 to 2023,  311 experienced
-a 50\% increase in the number of Service Requests (SRs), 
-reflecting a growing public reliance and usage of 
+During the 10\mbox{-}year period from 2014 to 2023,  NYC's 311 system
+experienced a 50\% increase in the number of SRs, 
+reflecting a growing public reliance on and usage of 
 the 311 system. (During this 10\mbox{-}year time-frame, the 
@@
 Figure~\ref{fig:SRcountbyAgency} provides a breakdown of SRs by
-the responsible agency, showing the cumulative 
+the responsible Agency, showing the cumulative 
 percentage of SRs handled by each
-agency over a the 2\mbox{-}year time-frame. Note that the
+agency over the 2\mbox{-}year time-frame. Note that the
 distribution of SRs is heavily concentrated among a few key 
@@
 
-\begin{itemize}[left=1.5em]
+\begin{enumerate}[left=1.5em]
     \item New York Police Department (NYPD): 43\% of all SRs
@@
     \item Department of Parks and Recreation (DPR): 4\%
-\end{itemize}
+\end{enumerate}
 
@@
 underscores the necessity for optimal data handling and 
-curation processes within these six high volume Agencies.
+curation processes within these  high volume Agencies.
 
@@
 	\includegraphics[width = \textwidth]{SRs_by_Agency.pdf}
-  	\caption{SR counts by agency with Cumulative Percentage}
+  	\caption{SR Counts by Agency with Cumulative Percentage}
 	\label{fig:SRcountbyAgency}
@@
 
-Figure~\ref{fig:SR_complaints}
-illustrates how the top 20 complaint types account for 70\% 
-of all SRs. Note that the distribution of complaints is skewed, 
+The dataset contains 210 different \texttt{complaint\_types}. 
+Figure~\ref{fig:SR_complaints} illustrates how the Top 20 
+\texttt{complaint\_type(s)} account for 70\% 
+of all SRs. The distribution of complaints is skewed, 
 with a small number of issues dominating total SR volume. 
-Of special interest are noise\mbox{-}related complaints, of which there are eight different
-types including residential, commercial, and 
-street noise. When combined noise\mbox{-}related SRs make up 22\% of all 
+Of special interest are noise\mbox{-}related complaints, of 
+which there are eight different types including vehicle, 
+residential, commercial, and street noise. When 
+combined, noise\mbox{-}related SRs make up 22\% of all 
 complaints; the most frequent issue in the NYC 311 system. 
@@
 complaints, and blocked driveways. The cumulative percentage curve 
-reveals that the top 20 complaint types comprise 70\% of SRs,
-while the remaining 190 complaint types are spread thinly 
+reveals that the Top 20 \texttt{complaint\_type(s)} comprise 70\% of SRs,
+while the remaining 190 other \texttt{complaint\_type(s)}  are spread thinly 
 across the remaining 30\%. This suggests that improving 
@@
 satisfaction. This chart also provides insights into the 
-operational pressures faced by City Agencies 
+operational pressures faced by those City Agencies 
 responsible for handling these high-volume complaints.
@@
 Many quality criteria are employed to ensure high\mbox{-}quality data 
-processing. One of the primary efforts is data validation, which spans 
+processing. One of the primary efforts is data validation, which includes 
 several critical checks. For instance, mandatory fields must not be 
@@
 naming conventions or data entries are inconsistent. Common issues 
-include fields that do not appear in the Data Dictionary,  
-inconsistent use of data fields. Furthermore, redundant 
-or irrelevant fields can clutter datasets, reducing efficiency, 
+include fields that do not appear in the Data Dictionary and  
+inconsistent usage of data fields. Furthermore, redundant 
+(or irrelevant) fields can clutter datasets, reducing efficiency, 
 creating confusion, and introducing errors. Logical inconsistencies 
@@
 Lastly, the balance between accuracy and precision is crucial, as both 
-must be carefully managed to ensure meaningful data.
+must be managed to ensure meaningful data.
 
@@
 issues might come about, including  detailed discussions 
-with the originating agency as to whether or not it is an 
+with the originating agency as to whether or not the issue is an 
 actual error, correct,  or holds some other status.
@@
 for each. However, there are six additional fields that can be
-optionally downloaded. These fields are 
+optionally downloaded if manually selected. These fields are 
 visible in the portal's ``Column manager'' widget, labeled with the prefix 
@@
 our analysis shows that these six ``computed'' fields have 
-significant data validity issues, leading us to treat these six fields 
-as experimental and \textit{not for suitable for public use}. An update 
+significant data validity issues, leading us to categorize these six fields 
+as experimental and \textit{not for public use}. An update 
 to the Data Dictionary would seem appropriate. Accordingly, 
@@
 in data analysis (typically by search for or excluding missing fields 
-and not aware of the various representations). A more consistent 
+and not aware of the various representations of such). A more consistent 
 approach to treating missing data should be employed to address this issue.
@@
 year there are changes to the local time as a result of observing 
-DST, and this can introduce some unexpected results. For example, 
-on March 13, 2022 and March 12, 2023 there are no times 
-between 1:59AM and 2:59AM: the 2:00AM hour on those dates 
+DST. This can introduce some unexpected results. For example, 
+on both March 13, 2022 and March 12, 2023 there are no times 
+between 1:59AM and 2:59AM. The 2:00AM hour on those dates 
 is simply missing owing to the clocks moving forward; at 2:00AM 
 it instantly becomes 3:00AM. This creates a gap in the hourly 
-distribution of SRs on those days, and it exaggerates the 
-lifespan of SRs, albeit by a single hour. Similarly on 
+distribution of SRs on those days, and exaggerates the 
+lifespan of SRs, albeit by a single hour. Similarly on both
 November 6, 2022 and November 5, 2023, the 
@@
 in a bizarre scenario where, for example, an SR can 
-be created at 1:59AM and then closed say 30 minutes later 
+be \textit{created} at 1:59AM and then be \textit{closed} 30 minutes later 
 at 1:29AM; seemingly nonsensical. Utilizing UTC time 
-for date\mbox{-}time stamps might help address this issue.
+for date\mbox{-}time stamps would help address this issue.
 
@@
 \texttt{due\_date}, you would be challenged as 99.6\% of the
-\texttt{due\_date} field are blank. When sorting fields by
+\texttt{due\_date} fields are blank. When sorting fields by
 blank (or N/A values) they appear to divide into three groups:
@@
 mostly or almost always blank. Figure~\ref{fig:blank_fields} is 
-a graphic depiction of total empty (blank \& N/As) for each field illustrating
-the grouping into the Most, Partial, and Few categories.
+a graphic depiction of total empty (blank \& N/As) for each 
+field illustrating the grouping into 
+the \textbf{Most}, \textbf{Partial}, and \textbf{Few/None} categories.
 
@@
 Any analytic effort must ensure that fields containing invalid values 
-are identified and isolated from the analysis. The \texttt{latitude} 
-and \texttt{longitude} fields were found to fall within the 
-geographic boundaries of New York City, and 
+are identified and isolated from the analysis. For example, the \texttt{latitude} 
+and \texttt{longitude} fields were found to all fall within the 
+geographic boundaries of New York City. Likewise, 
 the \texttt{unique\_key} field was indeed unique, as required. 
@@
 validate against the USPS database, which contains 
-44,173 valid zip codes. And while the \texttt{incident\_zip} field has 
-only 0.07\% invalid entries, this still results in 4,374 errors; 
-not insignificant. If you group the invalid entries by 
-percentage by agency, it mirrors the \textit{overall} 
-breakdown of SRs by agency, potentially 
+44,173 valid zip codes. We discovered that the \texttt{incident\_zip} 
+field has 0.07\% invalid entries, resulting in 4,374 errors. This includes
+zip codes such as ``12345'', ``10000'', and ``99999'. If you group 
+the invalid zip code entries by percentage by agency, it 
+mirrors the \textit{overall} breakdown of SRs by Agency, potentially 
 indicating a systemic problem.
 
-\paragraph{Created and Closed Dates}
+\paragraph{Created and Closed Dates:}
 \label{sec:negativeduration}
@@
 might expect that these fields would be automatically populated by the  
-application software. Thus when updating an SR status to ``new'' or ``closed''
-the appropriate software should automatically populate those fields. 
+application software. Thus when creating  or closing an SR 
+the the software would automatically populate these fields. 
 Unfortunately, this does not seem to be the case, as several anomalies 
@@
     \item SRs with a \texttt{closed\_date} that occurs before the 
-    \texttt{created\_date} resulting a nonsensical ``negative duration''.
+    \texttt{created\_date} resulting a nonsensical ``negative duration''. (12,450)
     \item \texttt{created\_date(s)} and \texttt{closed\_date(s)} in 
-    the far distant past, e.g. ``1900-01-01''.
+    the far distant past, e.g. ``1900-01-01''. (8)
     \item \texttt{created\_date(s)} and \texttt{closed\_date(s)} that 
-    match \textbf{to the second} creating an impossible ``zero duration''.
-    \item An unusually large number of SRs closed and/or created exactly at midnight 
-    and noon, to the second.
+    match \textbf{to the second} creating an impossible ``zero duration''. (163,720)
+    \item An unusually large number of SRs closed and/or created 
+    exactly at midnight, to the second. (247,000)
 \end{itemize}
 
-\paragraph{Closed before Created: Negative Duration}
+\paragraph{Closed before Created (Negative Duration):}
 Citizens, NYC Government Officials, and Agencies use the \texttt{created\_date}(s) and 
@@
 While duration is not directly present in the dataset, 
-it can be easily computed as the difference between
-\texttt{(closed\_date} and \texttt{created\_date}.  Duration is one of 
-most frequently used 311metrics and is frequently analyzed, such as determining if 
-one NYC borough receives faster response than another for certain 
-complaints. There are 12,450 SRs where the \texttt{closed\_date} precedes the 
+it is easily computed as the difference between
+\texttt{closed\_date} and \texttt{created\_date}.  Duration is one of 
+most frequently used 311 performance metrics 
+and is frequently analyzed, such as determining if 
+one NYC borough receives faster Agency responses than 
+another borough for certain complaints. There are 
+12,450 SRs where the \texttt{closed\_date} precedes the 
 \texttt{created\_date}, generating nonsensical ``negative durations.'' 
@@
 significantly impact response time analyses. Eight SRs with extremely 
-large negative durations ($-$730 days), all originating from the 
+large negative durations, all originating from the 
 Department of Homeless Services (DHS); all contain an entry 
-of ``1900-01-01'' as the \texttt{closed\_date}. This results 
-in negative durations exceeding $-$44,601 days (or 122 years). Many
-other SRs have negative durations in excess of 300+ days. Such 
-anomalies, though rare, can significantly skew statistical results if 
-not addressed during data cleansing. As a result, these SR rows are 
-removed from the analysis.
-
-Excluding the extreme negative values ($-$730 days), 
+of \textit{1900-01-01} as the \texttt{closed\_date}. This results 
+in very large negative durations exceeding $-$44,601 
+days (or 122 years). Many other SRs have negative durations 
+in excess of $-$300+ days. Such anomalies, though rare, can 
+significantly skew statistical results if not addressed during 
+data cleansing. Accordingly, these SR rows are removed from our analysis.
+
+Excluding the extreme negative values (any that are <$-$730 days), 
 Figure~\ref{fig:negative-duration-violin} shows the broad spread of 
 negative duration SRs. While there are few outliers, the magnitude 
-of these negative durations (300 days) is troubling and can produce 
-bizarre and unwanted analytical results.
+of these negative durations (some <$-$300 days) is troubling 
+and can produce bizarre analytical results such as greatly 
+distorting the mean response time for a specific complaint category.
 
@@
   \centering
-  \caption{Sample  negative durations [excluding extreme  values that are$-$730 days]}
+  \caption{Sample  Negative Durations [excluding extreme values that are <$-$730 days]}
    \label{tab:largest-errors}
@@
 
-\paragraph{Identical Created Date \& Closed Date: Zero Durations}
+\paragraph{Identical Created \& Closed Dates (Zero Durations):}
 A more prevalent issue occurs when the \texttt{closed\_date} and 
-\texttt{created\_date} are exactly the same, to the second. This 
-creates a \textbf{zero duration} for the SR, which is nonsensical. There 
-are 163,720 zero duration SRs, representing 2.6\% of all non\mbox{-}blank data. 
-99\% percent of these zero duration SRs occur within five agencies:
-DHMH, DOT, DOB, DSNY, and DEP. This distribution \textbf{does not} 
+\texttt{created\_date} have exactly the same date\mbox{-}time stamp, 
+to the second. This creates a \textbf{zero duration} for the 
+SR, which is nonsensical. There are 163,720 such zero duration 
+SRs, representing 2.6\% of all non\mbox{-}blank data. 
+Additionally, 99\% percent of these zero duration SRs 
+occur within five agencies: Department of Health \& Mental 
+Hygiene (DHMH), DOT, Department of Buildings (DOB), DSNY, and 
+DEP. This distribution \textbf{does not} 
 mirror the overall SR distribution by agency, suggesting 
-an agency\mbox{-}specific issue.
-
-\paragraph{Created Date or Closed Date(s) at Midnight or Noon}
+an Agency\mbox{-}specific issue.
+
+\paragraph{Created or Closed at Midnight:}
 Another issue involves the unusually large number of 
 SRs where the \texttt{created\_date} and \texttt{closed\_date} 
-indicate ``created'' or ``closed'' exactly at midnight (00:00:00) or 
-exactly at noon (12:00:00), down to the second. Normally, 
-SR creation and closure follows the work\mbox{-}day schedule, 
-with most SRs created during working hours and 
-fewer at night or in the early morning. However, there is a 
-significantly higher number of SRs closed exactly at midnight, 
-as well as a large number created exactly at midnight. The magnitude
-of this anomaly is well outside the 3$\sigma$ level.
+indicate ``created'' or ``closed'' exactly at midnight (00:00:00), 
+to the second. Normally, SR creation and closure follows 
+the work\mbox{-}day schedule, with most SRs created 
+during working hours and fewer at night or in the 
+early morning. However, there is a significantly higher 
+number of SRs closed exactly at midnight, as well 
+as a large number created exactly at midnight. The magnitude 
+of this anomaly is well outside the 3$\sigma$ level as shown 
+in Figure~\ref{fig:stacked}
 
@@
 suggest the presence of a bulk create/close software process, perhaps one
-that automatically assigns time stamps of midnight (00:00:00) or noon 
-(12:00:00) to large batches of SRs prior to sending the data to the 311 
-system. This behavior distorts the calculated duration of these 
-SRs by providing inaccurate dates. The distribution by agency 
-shows that over 92\% of the ``closed-exactly-at-midnight'' 
-SRs come from just two Agencies: DOB (69\%) and DSNY  (23\%).
+that automatically assigns time stamps of midnight (00:00:00) to 
+large batches of SRs prior to sending the data to the 311 
+data warehouse. This behavior distorts the calculated duration of these 
+SRs as well as overall system usage by providing inaccurate 
+dates. Further analysis of the distribution by Agency shows that over 92\% of 
+the ``closed-exactly-at-midnight'' SRs come from just two 
+Agencies: DOB (69\%) and DSNY  (23\%).
 
 \paragraph{Resolution Action Update Date}
-A similar issue occurs when the \texttt{resolution\_action\_update\_date} 
-occurs well after the \texttt{closed\_date}.  When an SR is 
+A similar issue occurs when an SR is modified (and the \texttt{resolution\_action\_update\_date} is altered) well 
+after the \texttt{closed\_date}.  When an SR is 
 updated, the 311 software automatically populates the \texttt{resolution\_action\_update\_date}. While it is certainly 
 possible, and even routine, to update an SR after it is closed, 
-some of these updates seem to be well beyond any normal or expected 
-timeframe, with some ``updates'' occurring over 2 years 
-later. As shown in Figure~\ref{fig:resolution-violin}, there 
-are 10,265 SRs that were updated >30 days after the 
-\texttt{closed\_date} (Note: The chart excludes values >730 days, 
-an arbitrary cutoff to exclude infeasible dates such as ``1900-01-01''.)
- The chart highlights the distribution of these post-closure 
+some of these updates seem to be well beyond any 
+normal or expected timeframe, with some ``updates'' 
+occurring over 2 years later. As shown in Figure~\ref{fig:resolution-violin}, 
+there are 10,265 SRs that were updated >30 days after the 
+\texttt{closed\_date} (Not that Figure~\ref(fig:resolution-violin) excludes 
+values >730 days, an arbitrary cutoff to exclude 
+infeasible dates such as \textit{1900-01-01}. The chart 
+highlights the distribution of these post-closure 
  updates, with a notable concentration of updates occurring within 
-the 30\mbox{-}to\mbox{-}90 day range. This pattern raises 
+the 30\mbox{-}to\mbox{-}200 day range. This pattern raises 
 questions about whether such delayed updates are standard 
 operational practice or indicative of a potential issue 
-which might warrant further investigation. It should be noted
+which warrants further investigation. It should be noted
 that 84\% of these late updates are associated with the TLC agency
@@
 
-\subsection{Accuracy and precision}
+\subsection{Accuracy and Precision:}
 \label{sec:precision}
 An question of precision vs. accuracy arises with the \texttt{latitude} 
-and \texttt{longitude} fields. as well as the duplicative 
-field \texttt{location}. These fields are expressed as 
-a 14\mbox{-}decimal number, e.g. 40.86769186022511. Given 
-that 1 degree of latitude at the equator is equal to 111.04 
-kilometers, the ``1'' at the end of that number represents 
+and \texttt{longitude} fields. as well as with the duplicative 
+\texttt{location} field. These fields are expressed as 
+a 14\mbox{-}decimal number, e.g. \texttt{latitude} 40.86769186022511. 
+Given that 1 degree of latitude at the equator is equal to 111.04 
+kilometers, the ``1'' at the end of the decimal number represents 
 approximately 1.1 nanometers (1/1,000,000,000 of a meter). (For 
-reference a DNA molecule is approximately 2nm in width.) The 
+reference a DNA molecule is approximately 2 nanometers in width.) The 
 representation of the \texttt{latitude} and \texttt{longitude} fields
-appears to be a classic case of extreme precision, but limited accuracy.
+appears to be a classic case of extreme precision, but limited 
+accuracy. All the \texttt{latitude}, \texttt{longitude}, and \texttt{location}
+fields display this characteristic.
 
@@
 the \texttt{latitude} and \texttt{longitude} fields albeit with a 
-comma and parentheses added. This makes the \texttt{location} field 
-arguably harder to use than the individual fields. For example:  
-\texttt{latitude}: 40.768456429488, \texttt{longitude}: $-$73.9575661888774, 
-form the \texttt{location}: (40.768456429488, $-$73.95756618887745).
+comma and parentheses added. Arguably, this makes the \texttt{location} field 
+more difficult to use than the individual fields. For example:  
+\texttt{latitude}: ``40.768456429488'', \texttt{longitude}: ``$-$73.9575661888774'', 
+form the \texttt{location}: ``(40.768456429488, $-$73.9575661888774)''.
 
@@
 redundant, especially given the space taken up by full names in the 
-CSV file. Additionally, the agency abbreviations are easily recognized.
+CSV file and the fact that the agency abbreviations are 
+easily recognized and well understood.
 
 \paragraph{\texttt{landmark} \& \texttt{street\_name}:} Per the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary},
-the \texttt{landmark} field ``can refer to any noteworthy location, 
+the \texttt{landmark} field \textit{can refer to any noteworthy location, 
 including but not limited to, parks, hospitals, airports, sports facilities, 
-performance spaces, etc.'. However, most entries are actually 
+performance spaces, etc..} However, most entries are actually 
 street names, with a 62\% match to \texttt{street\_name} field. 
-Even the non\mbox{-}matches (excluding blanks) appear to be 
+Even many of the non\mbox{-}matches (excluding blanks) appear to be 
 a logical match, differing only in spelling or formatting 
@@
 
-\paragraph{\texttt{cross\_street\_1/2} \& \texttt{intersection\_street\_1/2}:} These 
-two sets of street pairs are used to help identify the 
+\paragraph{\texttt{cross\_street\_1 \& 2} \& \texttt{intersection\_street\_1 \& 2}:} These 
+two street pairs are used to help identify the 
 \texttt{incident address}. We found 88\% of both of these two 
-street pairs to be duplicates. Such a high level of duplication 
+street pairs ( 1 \& 2) to be duplicates. Such a high level of duplication 
 raises the question of which field\mbox{-}pair should be trusted; 
@@
 information, but no documentation exists to clarify this 
-relationship. These two street\mbox{-}pairs are highly redundant.
+relationship. Nonetheless, these two street\mbox{-}pairs are highly redundant.
 
@@
     \item \textbf{\texttt{cross\_street\_1/2}} and \textbf{\texttt{intersection\_street\_1/2}}: 
-    These two street pairs both exhibit an 88\% match. We recommend deleting the 
+    These two street pairs both exhibit an 88\% match. While we cannot 
+    confirm the accuracy of either of these street pairs, we recommend deleting the 
     \texttt{intersection\_street\_1/2} fields while acknowledging 
@@
 \begin{quote}
-\textbf{Note:} \textit{Even a field that is 99\% blank will contain 64,000 
-non\mbox{-}blank rows in this large (6.4 million row) 
-dataset.} Non\mbox{-}trivial.
+\textit{[We should note that even a field that is 99\% 
+blank will still contain 64,000 non\mbox{-}blank rows in this 
+large (6.4 million row) dataset.}; non\mbox{-}trivial.]
 \end{quote}
@@
 The \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} would greatly benefit from an update.
-We found several
-discrepancies between the Data Dictionary and the actual data, which 
-could potentially mislead analytic efforts. Here are some examples:
-
-\begin{itemize}[left=1.5em]
-    \item While the NYC Open Data portal lists the data types of various 
-    fields, the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} does not.
-Nor does it provide details for specific fields such as constraints or the domain of legal
-    values.
-For example, in the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}, the \texttt{incident\_zip}
-    field is specified as ``text.'' This may not be the most suitable 
-    definition. Perhaps ``categorical,'' with a note indicating that 
-    the data must contain only numeric values and is not subjected to 
-    arithmetic operations would be a better dictionary description.
-
-    \item Many of the domains of legal values are incomplete or 
-    inaccurate.
-For example, the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} indicates that the
-    \texttt{status} field has values of \textit{assigned, canceled, 
-    closed}, or \textit{pending}. However, we found additional
-    values such as \textit{in progress, started}, and \textit{unspecified}. 
-    Additionally, no SRs were found with a status of \textit{canceled}.
-
-    \item The \texttt{address\_type} field similarly 
-    lists values of \textit{address, blockface, intersection, latlong}, 
-    and \textit{placename}. However, we found additional values such as 
-    \textit{bbl} and \textit{unrecognized}. Additionally, no \textit{latlong} 
-    values were observed. Other fields, such as \texttt{facility\_type}, 
-    \texttt{vehicle\_type}, \texttt{taxi\_pick\_up\_location}, 
-    \texttt{road\_ramp}, and \texttt{city}, exhibit similar inaccuracies 
+We found several discrepancies between the
+Data Dictionary and the actual data, which could potentially 
+mislead analytic efforts. Here are some examples:
+
+\paragraph{Data typing:} While the NYC Open Data portal displays the data types of 
+various fields (aka ``columns''), the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} does
+not.
+For example, in the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary},
+the \texttt{incident\_zip} field is specified as ``text.'' This may 
+not be the most suitable data definition. Perhaps a better 
+dictionary description would be  ``categorical'' with a note 
+indicating that the data must contain only numeric values 
+and is not subjected to arithmetic operations.
+
+\paragraph{Domain of Legal Values:} Many of the data fields have
+domains of legal values that are missing, incomplete, or inaccurate. For 
+example, the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} indicates
+that the \texttt{status} field takes on values of \textit{assigned, canceled, 
+closed}, or \textit{pending}. However, we observed additional values 
+such as \textit{in progress, started}, and \textit{unspecified}. Additionally, 
+no SRs were found with a status of \textit{canceled}. Similarly, 
+the Data Dictionary for the \texttt{address\_type} field lists as acceptable
+values \textit{address, blockface, intersection, latlong}, and 
+\textit{placename}. However, we found additional values such 
+as \textit{bbl} and \textit{unrecognized}. Additionally, no 
+\textit{latlong} values were observed. Other fields, such 
+as \texttt{facility\_type}, \texttt{vehicle\_type}, \texttt{taxi\_pick\_up\_location},
+texttt{road\_ramp}, and \texttt{city}, exhibit similar inaccuracies 
 in their specified domain values.
 
-	\item As mentioned earlier, six \texttt{@computed} fields that are
-    present in the data but not included in the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} need
-    to be documented and perhaps labeled as ``experimental'' or
-    ``not for official use''.
-\end{itemize}
+\paragraph{Missing data fields:} As mentioned earlier, six \texttt{@computed} 
+fields that are present in the data but not included in 
+the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} need
+to be documented and perhaps labeled as ``experimental'' 
+or ``not for official use''.
 
@@
 maintaining this consistency over time. To address this, agencies should adopt 
-standardized naming conventions and formats across all datasets. 
-Regular audits and data harmonization processes should be 
-implemented to ensure consistency across long\mbox{-}term datasets, 
-particularly as agency structures or reporting standards evolve. 
-For instance, ensuring that historical datasets reflect the same 
-field naming conventions over a span of years can prevent 
-inconsistent analyses. These standards should be promulgated 
-by the Open Data governing authority, following agreements 
-by the relevant parties.
+standardized naming conventions, value domains, usages, and 
+formats across all datasets. Regular audits and data harmonization 
+processes should be implemented to ensure consistency 
+across long\mbox{-}term datasets, particularly as agency structures 
+or reporting standards evolve. For instance, ensuring that 
+historical datasets reflect the same field naming conventions 
+over a span of years can prevent inconsistent analyses. 
+These standards should be promulgated by the Open Data 
+governing authority, following agreements by the relevant parties.
 
@@
     \item Certain date fields should be promulgated by the software
-    system versus manual entry. Certain dates should be populate 
+    system versus manual entry. Critical dates should be populated
     based upon an action in the system, such as a change in status.
@@
     \item Select fields should be subjected to automated checks 
-    that prevent logical errors, such as closing opening.
+    that prevent logical errors, such as closing and opening.
 
@@
 creation and closure at exactly midnight (or noon) likely 
-stem from a cron job type processes that distort the 
+stem from a cron script process that distorts the 
 accurate capture of SR durations. By implementing 
@@
 
+Automated Quality Assurance techniques should be employed to 
+monitor the quality of the data, as measured by agreed upon
+standards, such as the \% of invalid zip codes. Such tools could
+capture and display the quality of the data sets as presented
+on appropriate dashboards. Deviations in quality would be
+automatically highlighted, alerting appropriate corrective actions.
+
 Engagement with data scientists and statisticians is key to improving 
@@
 in this paper, benefit from a clear and accessible channel to the 311 SR data 
-team and other relevant city agencies. Establishing such feedback loops 
+team and other relevant City Agencies. Establishing such feedback loops 
 and other opportunities for public engagement, ensures that 
@@
 mechanism would also allow ongoing dialogue between 
-data users and city officials, leading to more dynamic 
+data users and City officials, leading to more dynamic 
 and responsive data management practices.
@@
 datasets that span long time periods, as changes in agency structures
-or reporting standards which can introduce subtle biases. Long\mbox{-}term data
-consistency is crucial for longitudinal studies and predictive
-modeling, as even minor data shifts can lead to significant deviations
-in model outcomes, as noted by \citet{rahm2000data} and
-\citet{borgman2012conundrum}.
+or reporting standards which can introduce subtle biases and data 
+distortions. Long\mbox{-}term data consistency is crucial for 
+longitudinal studies and predictive modeling, as even minor 
+data shifts can lead to significant deviations in model outcomes, 
+as noted by \citet{rahm2000data} and\citet{borgman2012conundrum}.
 

### 2024-11-19 manuscript/manuscript.tex
quick comments to intro and data sections

@@
 
-\doublespacing
 \begin{document}
+
+% \doublespacing
+
+% \tableofcontents % Optional: Table of Contents
+% \listoffigures % List of Figures
+% \listoftables % List of Tables
 
@@
   \institution{University of Connecticut}, \cny{USA}}
-
- \tableofcontents % Optional: Table of Contents
- \listoffigures % List of Figures
- \listoftables % List of Tables
 
@@
 serving as a crucial resource for research across various domains, 
-including health, urban development, and transportation. The 
-success of this initiative highlights the importance of data 
-curation in ensuring the utility and reliability of open datasets.
-
-This paper examines the challenges of open data curation through a
-case study of the NYC 311 Service Request (SR) dataset, addressing issues 
+including health, urban development, and transportation. While the
+open data movement has made vase amounts of data accessible, the
+challenge of ensuring data quality and usability through effective
+curation remain insufficiently addressed in the literature..
+This paper examines these challenges through a
+case study of the NYC 311 Service Request dataset, focusing on issues 
 of data validity, consistency, and curation efficiency. Based on 
 insights from this case study, we propose a set of data curation 
-principles tailored for government\mbox{-}released open data. These principles 
-aim to enhance data management practices and ensure 
-the ongoing utility of open data. The paper concludes with 
-actionable recommendations for enhancing data curation and outlines
-general principles for the effective release of open data.
-
+principles tailored for government-released open data aiming
+to enhance data management practices and ensure 
+the ongoing utility of open data.
 \end{abstract}
 
-\begin{keywords}
+\begin{keywords} % no repeating those in title
   \kwd{Data cleansing}
-  \kwd{Data Curation}
   \kwd{Data science}
   \kwd{NYC Open Data}
-  \kwd{Open data}
   \kwd{Smart Cities}
@@
 \citet{shankar2021data} observed its critical role during the 
-COVID\mbox{-}19 pandemic in managing collective responses.
+COVID-19 pandemic in managing collective responses.
+\jy{The gap in the literature needs to be further highlighted to set 
+  the stage for our contributions.}
 
@@
 gained from this case study, we propose a set of data curation 
-principles tailored for government\mbox{-}released open data. These 
+principles tailored for government-released open data. These 
 principles are designed to address the unique challenges 
@@
 on the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}.
-
 Section~\ref{sec:recommendations} offers 
@@
 \label{sec:data}
+
 The NYC 311 service, a critical component of New York City's public
@@
 with the the COVID\mbox{-}19 pandemic, the 2021 expansion 
+\jy{use full name of MTA because it is used only once.}
 to include the MTA's city subway system (the biggest 
@@
 and improving responsiveness across City Agencies.
+\jy{need to expand the importance of 311 SR data to the research
+  community: such as
+  resource allocation/response \citep{zha2014profiling, raj2021swift},
+  social equity \citep{kontokosta2021bias},
+  street flood \citep{agonafir2022understanding}, etc.}
 
@@
 there are 10,265 SRs that were updated >30 days after the 
-\texttt{closed\_date} (Not that Figure~\ref(fig:resolution-violin) excludes 
+\texttt{closed\_date} (Not that Figure~\ref{fig:resolution-violin} excludes 
 values >730 days, an arbitrary cutoff to exclude 
@@
  updates, with a notable concentration of updates occurring within 
-the 30\mbox{-}to\mbox{-}200 day range. This pattern raises 
+the 30--200 day range. This pattern raises 
 questions about whether such delayed updates are standard 
@@
 \texttt{location} field. These fields are expressed as
+\jy{This is a artifact of R reading them in as doubles. Single
+  precision should be sufficient as illustrated in my Chance manuscript.}
 a 14\mbox{-}decimal number, e.g. \texttt{latitude} 40.86769186022511. 

### 2024-11-19 manuscript/manuscript.tex
sorted and reformatted bib

@@
 but also empowers researchers, civic developers, and the general
+\jy{back up with some references:
+  \citep{minkoff2016nyc, o2017uncharted, kontokosta2021bias}}
 public. The data has been pivotal in providing advice on shelters
@@
   resource allocation/response \citep{zha2014profiling, raj2021swift},
-  social equity \citep{kontokosta2021bias},
+  social equity \citep{white2018promises, kontokosta2021bias},
+  noise \citep{dove2022sounds},
   street flood \citep{agonafir2022understanding}, etc.}
@@
 characteristics of the 311 SR dataset used in this analysis:
-
+\jy{need to clarify that we've made the column names to lower cases
+  and replaced spaces with underscores.}
 \begin{itemize}[left=1.5em]
@@
 \item In addition to  \texttt{incident\_address}, there are five additional location fields: 
-
   \texttt{latitude \& longitude}, \texttt{street\_name}, \texttt{landmark}, USGS State Plane 
   Coordinate System, and \texttt{block} \textbf{\#}.
+  \jy{use raw names from the data for consistency.}
 
@@
 
-During the 10\mbox{-}year period from 2014 to 2023,  NYC's 311 system
+During the 10-year period from 2014 to 2023,  NYC's 311 system
 experienced a 50\% increase in the number of SRs, 
@@
 
+\jy{Remove the text above each figure. They should be included in the
+  caption.}
+
 Figure~\ref{fig:SRcountbyAgency} provides a breakdown of SRs by
@@
 agencies. The six largest agencies are:
-
 \begin{enumerate}[left=1.5em]
@@
 \texttt{taxi\_company\_borough}, \texttt{due\_date},
+\jy{These variables seem only relevant to certain agencies, which
+  could be made available in a separate table.}
 \texttt{road\_ramp}, and \texttt{bridge\_highway\_name}.
@@
     \texttt{created\_date} resulting a nonsensical ``negative duration''. (12,450)
+\jy{what's the number in the parentheses?}
     \item \texttt{created\_date(s)} and \texttt{closed\_date(s)} in 
@@
     \toprule
-    {created\_date} & {closed\_date} & {duration (days)}  & {agency} \\
+    \texttt{created\_date} & \texttt{closed\_date} & Duration (days)}  & {Agency} \\
     \midrule
@@
 \texttt{created\_date} have exactly the same date\mbox{-}time stamp, 
-to the second. This creates a \textbf{zero duration} for the 
+to the second. This creates a \emph{zero duration} for the 
 SR, which is nonsensical. There are 163,720 such zero duration 

### 2024-11-19 manuscript/manuscript.tex
comments on some figures; added chance article to bib

@@
 \label{sec:intro}
+
+\jy{Need to avoid url links in the text. Could cite some as references
+  if needed.}
+
+\jy{No need for \mbox{-} in text.  one - is a hyphen; two -- is
+  en-dash;   three --- is an em-dash.}
 
@@
 
+\jy{Figures 4-6 seems taking more spaces than necessary. Fig 4 and 6
+  could be on one with ggplot facet. Figure 5 too. Both using two
+  panels on one row.}
+
 \begin{figure}[tbp]
@@
     \toprule
-    \texttt{created\_date} & \texttt{closed\_date} & Duration (days)}  & {Agency} \\
+    \texttt{created\_date} & \texttt{closed\_date} & {Duration (days)}  & {Agency} \\
     \midrule
@@
 \jy{This is a artifact of R reading them in as doubles. Single
-  precision should be sufficient as illustrated in my Chance manuscript.}
+  precision should be sufficient as illustrated in \citet{bates2024csv}.}
 a 14\mbox{-}decimal number, e.g. \texttt{latitude} 40.86769186022511. 
@@
 
+\jy{Could refer to cross-plotform data format such as Arrow
+  \citep{bates2024csv}.}
+
 \subsection{Data Dictionary} 
 \label{sec:datadictionary}
+
 The \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} would greatly benefit from an update.
@@
 The NYC 311 SR data and code used to produce the results reported in
-this paper are available [FixMe].
+this paper are available at \url{https://github.com/jun-yan/nyc311clean}.
 

### 2024-11-21 manuscript/manuscript.tex
Initial editing on Manuscript
Responding to JY's comments.

@@
 \newcommand{\dt}[1]{\textcolor{purple}{DT: (#1)}}
-\newcommand{\jy}[1]{\textcolor{cyan}{JY: (#1)}}
+\newcommand{\jy}[1]{\textcolor{red}{JY: (#1)}}
 
@@
 
-\jy{No need for \mbox{-} in text.  one - is a hyphen; two -- is
-  en-dash;   three --- is an em-dash.}
-
 In the early 21st century, the open data movement began 
@@
 college enrollment statistics, jail inmate charges, and the location
-of City\mbox{-}wide free Internet access points. These datasets have been
+of City-wide free Internet access points. These datasets have been
 applied in civil life in various ways, such as mapping car crashes
@@
 consistent, accurate, and useful for diverse applications. For example, 
-well\mbox{-}curated data is essential for machine learning systems, which 
-require high\mbox{-}quality data to produce reliable insights 
+well-curated data is essential for machine learning systems, which 
+require high-quality data to produce reliable insights 
 \citep{polyzotis2019data, jain2020overview}. A critical component of 
@@
 
-Research into data curation has explored these challenges in\mbox{-}depth. 
+Research into data curation has explored these challenges in-depth. 
 Among the earliest discussions, \citet{witt2009constructing} focused 
@@
 an overview of the history of the NYC 311 SR system and presents 
-a summary of SR counts over a 2\mbox{-}year period from 2022 to 2023. 
+a summary of SR counts over a 2-year period from 2022 to 2023. 
 Section~\ref{sec:issues} delves into specific data cleansing 
@@
 engagement and service response framework, serves as a centralized hub
-for non\mbox{-}emergency inquiries and requests. Introduced in 2003, the NYC
+for non-emergency inquiries and requests. Introduced in 2003, the NYC
 311 system was designed to streamline the city's response to
-non\mbox{-}emergency issues, ranging from noise complaints to street
-maintenance requests. Initially a phone\mbox{-}based call center, the system
+non-emergency issues, ranging from noise complaints to street
+maintenance requests. Initially a phone-based call center, the system
 evolved into a comprehensive data management platform handling
 millions of requests annually. Key milestones since its launch in 2003
-include the addition of online and mobile app channels in 2009, a
-record high of 348,463 monthly service requests in August 2020 associated 
-with the the COVID\mbox{-}19 pandemic, the 2021 expansion 
-\jy{use full name of MTA because it is used only once.}
-to include the MTA's city subway system (the biggest 
-expansion in 311  history), and a record 3.23 million 
-service requests in 2023.
+include the addition of online and mobile channels in 2009, a
+record monhtly high of 348,463 Service Requests in 
+August 2020 (associated with the the COVID-19 
+pandemic), the 2021 expansion to include the 
+Metropolitan Transit Authority (MTA) NYC subway 
+system (the biggest expansion in 311 history), 
+and a yearly record of 3.23 million SRs in 2023.
 
@@
 Portal, which provides tools for querying, grouping, aggregating,
-geo\mbox{-}mapping, visualizing, and exporting results. The
+geo-mapping, visualizing, and exporting results. The
 \href{https://www.nyc.gov/content/oti/pages/}{NYC Office of Technology 
@@
 the 311 application, the open data infrastructure, mobile \& web applications, 
-and the public\mbox{-}facing data portal. Input data for the 311 system 
+and the public-facing data portal. Input data for the 311 system 
 is sourced from 16 different NYC Agencies. Some of these Agencies
 utilize the core 311 software, while others transfer data to the
-311 Open Data Warehouse from their in\mbox{-}house systems.
+311 Open Data Warehouse from their in-house systems.
 
@@
 public. The data has been pivotal in providing advice on shelters
-during emergencies, handling inquiries during the COVID\mbox{-}19 pandemic,
+during emergencies, handling inquiries during the COVID-19 pandemic,
 enforcing standards between landlords and tenants, reallocating taxi
@@
 
-Our investigation utilizes a 311 SR dataset covering  2\mbox{-}years (2022-2023).
+Our investigation utilizes a 311 SR dataset covering  2-years (2022-2023).
 This sample was collected on 15 September 2024, and is used 
@@
 characteristics of the 311 SR dataset used in this analysis:
-\jy{need to clarify that we've made the column names to lower cases
-  and replaced spaces with underscores.}
+
 \begin{itemize}[left=1.5em]
@@
 
-\item There are 41 columns of data for each row, with each
-  row representing a single Service Request (SR).
-
-\item There are four date fields: \texttt{created\_date}, \texttt{closed\_date}, 
-  \texttt{resolution\_action\_updated\_date}, \& \texttt{due\_date}.
-
-\item There are two borough fields; one of which appears to be a duplicate.
-
-\item There are seven street fields; two pair of which appear to be duplicates
-
-\item In addition to  \texttt{incident\_address}, there are five additional location fields: 
-  \texttt{latitude \& longitude}, \texttt{street\_name}, \texttt{landmark}, USGS State Plane 
-  Coordinate System, and \texttt{block} \textbf{\#}.
-  \jy{use raw names from the data for consistency.}
-
-\item A free\mbox{-}form text field, \texttt{resolution\_description}, which 
-  supports 930 characters of input, including commas and 
-  special characters  which can be problematic for automated processing.
+\item There are 41 columns (fields) of data for each row, with each
+  row representing a single Service Request (SR). To facilitate 
+  processing, column names were made lower case and 
+  spaces were replaced with underscores, e.g. \texttt{complaint\_type}. 
+  This naming convention is used throught this paper.
+
+\item Each SR has four date fields: 
+	\begin{itemize}	
+		\item{\texttt{created\_date}}
+		\item{\texttt{closed\_date}}
+		\item{\texttt{resolution\_action\_updated\_date}}
+		\item{\texttt{due\_date}}
 	\end{itemize}
+
+\item There are two borough fields, \texttt{borough} \& \texttt{park\_borough}, 
+		which appear to be duplicates.
+
+\item There are seven street fields; two pair of which appear to be duplicates:
+	\begin{itemize}
+		\item{\texttt{incident\_address}}
+		\item{\texttt{street\_name}}
+		\item{\texttt{cross\_street\_1}}
+		\item{\texttt{cross\_street\_2}}
+		\item{\texttt{intersection\_street\_1}}
+		\item{\texttt{intersection\_street\_2}}
+		\item{\texttt{landmark}}
+	\end{itemize}
+
+\item In addtion to \texttt{incident\_address}, each SR has six
+	additional geographic fields:
+	\begin{itemize}
+		\item{\texttt{latitude} \& \texttt{longitude}}
+		\item{\texttt{location}}
+		\item{\texttt{street\_name}}
+		\item{\texttt{landmark}}
+		\item{\texttt{block}}
+		\item{\texttt{x\_coordinate\_state\_plane} \& \texttt{y\_coordinate\_state\_plane}}.
+ 	\end{itemize}
+
+\item A free-form text field, \texttt{resolution\_description}, which 
+  supports 930 characters including commas and 
+  special characters; often problematic for automated processing.
+\end{itemize}
+\medskip
 
@@
 reflecting a growing public reliance on and usage of 
-the 311 system. (During this 10\mbox{-}year time-frame, the 
+the 311 system. (During this 10-year time-frame, the 
 population of NYC grew just 0.6\%.). The most significant increase
 in activity occurred in 2020, when the volume of requests 
-surged to 3.23 million during the COVID\mbox{-}19 pandemic, 
+surged to 3.23 million during the COVID-19 pandemic, 
 illustrating the essential role of the 311 system in supporting 
@@
 percentage of SRs handled by each
-agency over the 2\mbox{-}year time-frame. Note that the
+agency over the 2-year time-frame. Note that the
 distribution of SRs is heavily concentrated among a few key 
@@
 with a small number of issues dominating total SR volume. 
-Of special interest are noise\mbox{-}related complaints, of 
+Of special interest are noise-related complaints, of 
 which there are eight different types including vehicle, 
 residential, commercial, and street noise. When 
-combined, noise\mbox{-}related SRs make up 22\% of all 
+combined, noise-related SRs make up 22\% of all 
 complaints; the most frequent issue in the NYC 311 system. 
@@
 
-Many quality criteria are employed to ensure high\mbox{-}quality data 
+Many quality criteria are employed to ensure high-quality data 
 processing. One of the primary efforts is data validation, which includes 
@@
 at 1:29AM; seemingly nonsensical. Utilizing UTC time 
-for date\mbox{-}time stamps would help address this issue.
+for date-time stamps would help address this issue.
 
@@
 \begin{itemize}[left=1.5em]
-    \item SRs with a \texttt{closed\_date} that occurs before the 
-    \texttt{created\_date} resulting a nonsensical ``negative duration''. (12,450)
-\jy{what's the number in the parentheses?}
-    \item \texttt{created\_date(s)} and \texttt{closed\_date(s)} in 
-    the far distant past, e.g. ``1900-01-01''. (8)
-    \item \texttt{created\_date(s)} and \texttt{closed\_date(s)} that 
-    match \textbf{to the second} creating an impossible ``zero duration''. (163,720)
-    \item An unusually large number of SRs closed and/or created 
-    exactly at midnight, to the second. (247,000)
+    \item 12,450 SRs with a \texttt{closed\_date} that occurs before the 
+    \texttt{created\_date} resulting a nonsensical ``negative duration''. 
+    \item Eight SRs with \texttt{created\_date(s)} or \texttt{closed\_date(s)} in 
+    the distant past, specifically ``1900-01-01''. Despite the small number of 
+    these entries, such extreme values can greatly distort analytical conclusions.
+    \item 163,720 SRs with identical \texttt{created\_date(s)} and \texttt{closed\_date(s)} 
+    to the second, creating a nonsensical ``zero duration''. 
+    \item 247,000 SRs closed and/or created exactly at midnight, to the second.
 \end{itemize}
@@
 A more prevalent issue occurs when the \texttt{closed\_date} and 
-\texttt{created\_date} have exactly the same date\mbox{-}time stamp, 
+\texttt{created\_date} have exactly the same date-time stamp, 
 to the second. This creates a \emph{zero duration} for the 
 SR, which is nonsensical. There are 163,720 such zero duration 
-SRs, representing 2.6\% of all non\mbox{-}blank data. 
+SRs, representing 2.6\% of all non-blank data. 
 Additionally, 99\% percent of these zero duration SRs 
@@
 mirror the overall SR distribution by agency, suggesting 
-an Agency\mbox{-}specific issue.
+an Agency-specific issue.
 
@@
 to the second. Normally, SR creation and closure follows 
-the work\mbox{-}day schedule, with most SRs created 
+the work-day schedule, with most SRs created 
 during working hours and fewer at night or in the 
@@
   precision should be sufficient as illustrated in \citet{bates2024csv}.}
-a 14\mbox{-}decimal number, e.g. \texttt{latitude} 40.86769186022511. 
+a 14-decimal number, e.g. \texttt{latitude} 40.86769186022511. 
 Given that 1 degree of latitude at the equator is equal to 111.04 
@@
 street names, with a 62\% match to \texttt{street\_name} field. 
-Even many of the non\mbox{-}matches (excluding blanks) appear to be 
+Even many of the non-matches (excluding blanks) appear to be 
 a logical match, differing only in spelling or formatting 
@@
 street pairs ( 1 \& 2) to be duplicates. Such a high level of duplication 
-raises the question of which field\mbox{-}pair should be trusted; 
+raises the question of which field-pair should be trusted; 
 and if these street pairs are used differently, how should they be treated?
@@
 information, but no documentation exists to clarify this 
-relationship. Nonetheless, these two street\mbox{-}pairs are highly redundant.
+relationship. Nonetheless, these two street-pairs are highly redundant.
 
@@
 \label{sec:filesize}
-By removing duplicate and ``near\mbox{-}duplicate'' fields, it is possible to 
+By removing duplicate and ``near-duplicate'' fields, it is possible to 
 shrink the file size by 39.5\%, which for this large dataset equates to 
@@
 simpler data analysis efforts. Below is a list of duplicate 
-and near\mbox{-}duplicate fields which we believe could be
+and near-duplicate fields which we believe could be
 removed from the dataset with minimal data loss.
@@
 \textit{[We should note that even a field that is 99\% 
-blank will still contain 64,000 non\mbox{-}blank rows in this 
-large (6.4 million row) dataset.}; non\mbox{-}trivial.]
+blank will still contain 64,000 non-blank rows in this 
+large (6.4 million row) dataset.}; non-trivial.]
 \end{quote}
@@
 data. This approach is particularly useful in large datasets, where 
-frequently repeated text\mbox{-}based categorical variables can 
+frequently repeated text-based categorical variables can 
 significantly increase storage requirements.
@@
 processes should be implemented to ensure consistency 
-across long\mbox{-}term datasets, particularly as agency structures 
+across long-term datasets, particularly as agency structures 
 or reporting standards evolve. For instance, ensuring that 
@@
 large government datasets. Our research found several duplicate and
-near\mbox{-}duplicate fields in the 311 SR dataset. Eliminating 
+near-duplicate fields in the 311 SR dataset. Eliminating 
 redundant fields and encoding categorical variables can 
@@
 accurate capture of SR durations. By implementing 
-real\mbox{-}time and event\mbox{-}driven validation tools, governments 
+real-time and event-driven validation tools, governments 
 can both prevent such errors from entering the dataset in the first 
@@
 readily understandable. This involves providing clear 
-metadata and making datasets available in user\mbox{-}friendly 
+metadata and making datasets available in user-friendly 
 formats. Additionally, governments should communicate transparently about 
@@
 implement them effectively. Governments can begin by investing in 
-real\mbox{-}time validation systems, regular updates to data dictionaries, 
+real-time validation systems, regular updates to data dictionaries, 
 and tools for encoding and standardizing data fields. This would include
@@
 critical in developing protocols for validating incoming data, correcting 
-errors in real\mbox{-}time, and ensuring consistency across multiple data sources. 
+errors in real-time, and ensuring consistency across multiple data sources. 
 Feedback from statisticians and data scientists, such as the issues reported 
@@
 should continue supporting these types of initiatives, which not only promote 
-data literacy, but encourage community\mbox{-}driven improvements to datasets. 
+data literacy, but encourage community-driven improvements to datasets. 
 By leveraging the expertise of data scientists, statisticians, stakeholders, 
@@
 such as biases or inconsistencies, can compromise both the accuracy
-and fairness of predictive models, impacting real\mbox{-}world decisions in
+and fairness of predictive models, impacting real-world decisions in
 urban planning and policy \citep{rahm2000data, geiger2020garbage}.
@@
 or reporting standards which can introduce subtle biases and data 
-distortions. Long\mbox{-}term data consistency is crucial for 
+distortions. Long-term data consistency is crucial for 
 longitudinal studies and predictive modeling, as even minor 
@@
 trends, allocate resources, and address service delivery issues, such
-as delays in responding to 311 complaints. The COVID\mbox{-}19 pandemic
-underscored the importance of real\mbox{-}time, trustworthy data in crisis
-management, where the accuracy and timeliness of data\mbox{-}driven insights
+as delays in responding to 311 complaints. The COVID-19 pandemic
+underscored the importance of real-time, trustworthy data in crisis
+management, where the accuracy and timeliness of data-driven insights
 were critical for public health responses. Without proper data
@@
 Future efforts could focus on automating data curation processes,
-particularly in real\mbox{-}time data pipelines, to ensure that the data used
+particularly in real-time data pipelines, to ensure that the data used
 in machine learning models remains accurate, clean, and reliable

### 2024-11-21 code/datacleansing.R

@@ -68,7 +68,7 @@ sink("../../console_output/core_console_output.txt")
 cat("\nExecution begins at:", formattedStartTime)
 cat("\n***** Program initialization *****")
 
-options(digits = 14) # Set the number of decimal places to 14
+options(digits = 15) # Set the number of decimal places to 14
 
 #########################################################################
 # File contains column names in the "header" line.

### 2024-11-22 manuscript/manuscript.tex
Response to JY's comments
Removed subtitles from top of charts. Updating manuscript and ref.bib to show URLs.

@@
 A landmark in this journey was the launch of the United States'
-\href{https://www.data.gov}{Data.gov} portal in 2009, a pioneering
+Data.gov portal in 2009 \citep{dataGov}, a pioneering
 platform in making government data widely accessible. Shortly afterwards,
@@
 \citep{chu2016data, hurbean2021open}.
+x`
 

### 2024-11-22 code/functions/base_bar_chart.R

@@ -52,14 +52,14 @@ base_bar_chart <- function(dataset, x_col, y_col, chart_title, sub_title,
     theme(
       axis.title = element_blank(),  # Remove x and y axis titles
       plot.title = element_text(hjust = 0.5, size = 12),
-      plot.subtitle = element_text(size = 7),
+#      plot.subtitle = element_text(size = 7),
       axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1, face = "bold", size = 8),
       axis.text.y = element_text(face = "bold", size = 8),
       aspect.ratio = 0.618033
     ) +
     
     # Set the subtitle with total count included
-    ggtitle(chart_title, subtitle = paste(sub_title, format(y_total_count, big.mark = ","), sep = " ")) +
+#    ggtitle(chart_title, subtitle = paste(sub_title, format(y_total_count, big.mark = ","), sep = " ")) +
     
     # Remove x and y axis labels
     labs(x = NULL, y = NULL)  

### 2024-11-22 code/functions/create_violin_chart.R

@@ -22,8 +22,8 @@ create_violin_chart <- function(
       title = chart_title,
       x = x_axis_title,
       y = "",
-      subtitle = paste("(", earliest_title, "--", latest_title, ")", " n=", 
-                       format(nrow(dataset), big.mark = ","), sep = "")
+      # subtitle = paste("(", earliest_title, "--", latest_title, ")", " n=", 
+      #                  format(nrow(dataset), big.mark = ","), sep = "")
     ) +
     
     theme(

### 2024-11-24 manuscript/manuscript.tex
Addressing JY's comments
Working on addressing JY's comments. Making small changes to R code to accommodate. Still need to update images, but making good progress.  Removed -47 R code file. Adding URLs to ref.bib.

@@
 \usepackage{enumitem}
+\usepackage{tikz} % For creating diagrams
+\usepackage{hyperref}   % For clickable links and breaking long URLs
+\usepackage{url}        % Optional if hyperref is not loaded
+\usetikzlibrary{positioning} % Required for relative positioning (e.g., "of" keyword)
 
@@
 
-\jy{Need to avoid url links in the text. Could cite some as references
-  if needed.}
-
 In the early 21st century, the open data movement began 
@@
 platform in making government data widely accessible. Shortly afterwards,
-the European Union followed suit, unveiling its
-\href{https://data.europa.eu/euodp}{Open Data Portal} in 2012, further
-cementing the movement's global reach. Furthermore, the World Bank's Open
+the European Union followed suit, unveiling its Open Data Portal 
+\citep{dataEU} in 2012, further cementing the movement's 
+global reach. Furthermore, the World Bank's Open
 Data initiative, initiated in 2010, stands out as a comprehensive
 repository for global development data, available at
-\href{https://data.worldbank.org}{World Bank Open Data}. 
+World Bank Open Data \citep{dataWorldBank}. 
 These initiatives represent significant strides in democratizing data, 
@@
 \citep{zuiderwijk2014open}. This landmark legislation led to the
-creation of the \href{https://opendata.cityofnewyork.us}{NYC Open Data
-  portal}, which today hosts an impressive array of 2,700 datasets
+creation of the NYC Open Data Portal \citep{dataNYC}, which 
+today hosts an impressive array of 2,700 datasets
 across 80 different city agencies. This resource has become invaluable
@@
 COVID-19 pandemic in managing collective responses.
+
 \jy{The gap in the literature needs to be further highlighted to set 
@@
 and identify duplicate or redundant fields, along with observations 
-on the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}.
+on the Data Dictionary \citep{datadictionaryNYC}.
 Section~\ref{sec:recommendations} offers 
@@
 Today, the NYC 311 data system manages over 3 million service
-requests (SRs) per year. This data is publicly accessible through the NYC Open Data
-Portal, which provides tools for querying, grouping, aggregating,
+requests (SRs) per year. This data is publicly accessible 
+through the NYC Open Data Portal, which provides 
+tools for querying, grouping, aggregating,
 geo-mapping, visualizing, and exporting results. The
-\href{https://www.nyc.gov/content/oti/pages/}{NYC Office of Technology 
-and Innovation (OTI)}, provides technical support for 
+NYC Office of Technology and Innovation (OTI) provides technical support for 
 the 311 application, the open data infrastructure, mobile \& web applications, 
@@
 but also empowers researchers, civic developers, and the general
-\jy{back up with some references:
-  \citep{minkoff2016nyc, o2017uncharted, kontokosta2021bias}}
+
+\jy{back up with some references:}
+
+  \citep{minkoff2016nyc, o2017uncharted, kontokosta2021bias}
 public. The data has been pivotal in providing advice on shelters
@@
 and improving responsiveness across City Agencies.
+
 \jy{need to expand the importance of 311 SR data to the research
@@
 
-\jy{Remove the text above each figure. They should be included in the
-  caption.}
-
 Figure~\ref{fig:SRcountbyAgency} provides a breakdown of SRs by
@@
 agencies. The six largest agencies are:
+
 \begin{enumerate}[left=1.5em]
@@
 	\includegraphics[width = \textwidth]{SRs_by_Agency.pdf}
-  	\caption{SR Counts by Agency with Cumulative Percentage}
+  	\caption{SR Counts by Agency for the time period 2022-2023}
 	\label{fig:SRcountbyAgency}
@@
   \includegraphics[width = \textwidth]{SR_by_Complaint_Type.pdf} 
-  \caption{Top 20 complaint types and Cumulative Percentage} 
+  \caption{Top 20 Complaint Types for the time period 2022-2023} 
   \label{fig:SR_complaints}
@@
 such as numeric, character, or date formats; typically 
-outlined in a \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}.
-An
-essential aspect of validation is domain compliance, where 
+outlined in a Data Dictionary. An essential aspect of 
+validation is domain compliance, where 
 fields must adhere to a predefined set of 
@@
 \paragraph{Undocumented Fields:} The 311 SR 
-\href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}
-identifies 41 data columns (fields) and provides related information 
+The 311 Service Request Data Dictionary identifies 
+41 data columns (fields) and provides related information 
 for each. However, there are six additional fields that can be
@@
 usage. These six fields are not included in
-the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}.
-Additionally,
-our analysis shows that these six ``computed'' fields have 
+the 311 Service Request Data Dictionary. Additionally, our analysis 
+revealed that these six ``computed'' fields have 
 significant data validity issues, leading us to categorize these six fields 
@@
 \begin{itemize}[left=1.5em]
-    \item \textbf{Mostly Empty}: 93-99.9\% blank 
-    \item \textbf{Partially Empty}: 40-4\% blank
-    \item \textbf{Few/None Empty}: 2-0\% blank
+    \item Mostly Empty: 93-99.9\% blank 
+    \item Partially Empty: 40-4\% blank
+    \item Few/None Empty: 2-0\% blank
 \end{itemize}
@@
 field illustrating the grouping into 
-the \textbf{Most}, \textbf{Partial}, and \textbf{Few/None} categories.
+the Most, Partial, and Few/None categories.
 
@@
   	\includegraphics[width=\textwidth]{BlankFields.pdf}
-	\caption{Number and Percentage of Empty/Blank Entries}
+	\caption{Number and Percentage of Empty/Blank Entries for the time period 2022-2023}
 	\label{fig:blank_fields}
@@
 the \texttt{unique\_key} field was indeed unique, as required. 
-Unfortunately, the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} specifies very few domains of
+Unfortunately, the 311 Service Request Data Dictionary specifies very few domains of 
 acceptable values for other fields. Nonetheless, the following fields were tested and found to comply with their expected domains, as determined by public 
@@
 
-\jy{Figures 4-6 seems taking more spaces than necessary. Fig 4 and 6
-  could be on one with ggplot facet. Figure 5 too. Both using two
-  panels on one row.}
-
 \begin{figure}[tbp]
   \centering
+  % First figure
+  \begin{subfigure}[t]{0.495\textwidth} % Adjust width as needed
+    \centering
+    \includegraphics[width=\textwidth]{post_closed_violin_chart.pdf}
+    \caption{Post-closed \texttt{resolution\_action\_update\_date}(s) >30 days}
+    \label{fig:resolution-violin}
+  \end{subfigure}
+  \hfill % Horizontal space between subfigures
+  % Second figure
+  \begin{subfigure}[t]{0.495\textwidth} % Adjust width as needed
+    \centering
     \includegraphics[width=\textwidth]{negative_duration_SR_violin.pdf}
- \caption{Distribution of Negative Durations}
+    \caption{Distribution of Negative Durations.}
     \label{fig:negative-duration-violin}
+  \end{subfigure}
+  % Main figure caption
+  \caption{Comparison of Violin Plots: (a) Post-Closed Resolution and (b) Negative Durations for the time period 2022-2023}
+  \label{fig:violin-plots}
 \end{figure}
@@
 Hygiene (DHMH), DOT, Department of Buildings (DOB), DSNY, and 
-DEP. This distribution \textbf{does not} 
-mirror the overall SR distribution by agency, suggesting 
-an Agency-specific issue.
+DEP. This distribution does not mirror the overall SR 
+distribution by agency, suggesting an Agency-specific issue.
 
@@
     \centering
-    \begin{subfigure}{\textwidth}
+    % First subfigure
+    \begin{subfigure}[t]{0.495\textwidth} % Adjust width to fit side by side
         \centering
@@
     \end{subfigure}
-    \par\medskip
-    \begin{subfigure}{\textwidth}
+    \hfill % Add horizontal space between subfigures
+    % Second subfigure
+    \begin{subfigure}[t]{0.495\textwidth} % Adjust width to fit side by side
         \centering
@@
     \end{subfigure}
-    \caption{Comparison of SRs Created and Closed by Hour of the Day}
-    \label{fig:stacked}
+    % Main caption
+    \caption{Comparison of SRs Created \& Closed by Hour-of-the-Day for the time period 2022-2023}
+    \label{fig:comparison}
 \end{figure}
@@
 \texttt{location} field. These fields are expressed as
-\jy{This is a artifact of R reading them in as doubles. Single
-  precision should be sufficient as illustrated in \citet{bates2024csv}.}
 a 14-decimal number, e.g. \texttt{latitude} 40.86769186022511. 
@@
 
-\subsection{Redundant \& Duplicate fields}
+\subsection{Redundant Fields}
 \label{sec:duplicates}
@@
 more difficult to use than the individual fields. For example:
-\texttt{latitude}: ``40.768456429488'', \texttt{longitude}: ``$-$73.9575661888774'', 
-form the \texttt{location}: ``(40.768456429488, $-$73.9575661888774)''.
+
+\vspace{.25cm}
+\begin{tikzpicture}[node distance=2cm and 1.5cm, >=stealth]
+
+    % Nodes for latitude, longitude, and location (centered between margins)
+    \node (latitude) at (-1.5, 0.9) {40.768456429488};
+    \node (longitude) at (1.8, 0.95) {$-$73.9575661888774};
+    \node (location) at (0.285, 0) {(40.768456429488, $-$73.9575661888774)};
+
+    % Labels anchored to the left margin
+    \node[anchor=west] at (-8.5, 0.9) {\texttt{latitude} \& \texttt{longitude} fields};
+    \node[anchor=west] at (-8.5, 0) {\texttt{location} field};
+
+    % Arrows connecting specific parts of the nodes
+    \draw[->] (latitude.south) -- ([xshift=-1.77cm]location.north);
+    \draw[->] (longitude.south) -- ([xshift=1.525cm]location.north);
+
+\end{tikzpicture}
 
@@
 
-\paragraph{\texttt{landmark} \& \texttt{street\_name}:} Per the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary},
-the \texttt{landmark} field \textit{can refer to any noteworthy location, 
+\paragraph{\texttt{landmark} \& \texttt{street\_name}:} Per the 
+311 Service Request Data Dictionary, the \texttt{landmark} 
+field \textit{can refer to any noteworthy location, 
 including but not limited to, parks, hospitals, airports, sports facilities, 
@@
 \begin{itemize}[left=1.5em]
-    \item \textbf{\texttt{agency\_name}}: Since each row contains the \texttt{agency} 
+    \item \texttt{agency\_name}: Since each row contains the \texttt{agency} 
     field, which uses clear and well understood abbreviations, we believe the
@@
 
-    \item \textbf{\texttt{park\_borough}}: This field is a 100\% match with 
+    \item \texttt{park\_borough}: This field is a 100\% match with 
     the \texttt{borough} field and can be deleted without data loss.
 
-    \item \textbf{\texttt{location}}: This field is a straight concatenation of 
+    \item \texttt{location}: This field is a straight concatenation of 
     the \texttt{latitude} and \texttt{longitude} fields. It can be 
@@
 
-    \item \textbf{\texttt{cross\_street\_1/2}} and \textbf{\texttt{intersection\_street\_1/2}}: 
+    \item \texttt{cross\_street\_1/2} and \texttt{intersection\_street\_1/2}: 
     These two street pairs both exhibit an 88\% match. While we cannot 
@@
 
-The \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} would greatly benefit from an update.
-We found several discrepancies between the
-Data Dictionary and the actual data, which could potentially 
-mislead analytic efforts. Here are some examples:
+The 311 Service Request Data Dictionary would greatly benefit from an update.
+We found several discrepancies between the Data Dictionary and the actual data, which could potentially mislead analytic efforts.
+Here are some examples:
 
 \paragraph{Data typing:} While the NYC Open Data portal displays the data types of 
-various fields (aka ``columns''), the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} does
-not.
-For example, in the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary},
+various fields (aka ``columns''), the Data Dictionary does 
+not. For example, in the Data Dictionary, 
 the \texttt{incident\_zip} field is specified as ``text.'' This may 
@@
 domains of legal values that are missing, incomplete, or inaccurate. For 
-example, the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} indicates
+example, the Data Dictionary indicates 
 that the \texttt{status} field takes on values of \textit{assigned, canceled, 
@@
 fields that are present in the data but not included in 
-the \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} need
-to be documented and perhaps labeled as ``experimental'' 
-or ``not for official use''.
+the Data Dictionary need to be documented and perhaps 
+labeled as ``experimental'' or ``not for official use''.
 
@@
 \subsection{Principles}
-\paragraph{Principle 1: Ensuring Consistency Across Agencies and Time}
+\paragraph{Principle 1: Establishing Cross-Orginazational Consistency}
 A key challenge in government released datasets is the 
@@
 
-\paragraph{Principle 2: Guaranteeing Accuracy and Validity during Data Entry}
+\paragraph{Principle 2: Ensuring Data Entry Accuracy}
 Data accuracy and validity are fundamental to the utility of 
@@
 
-\paragraph{Principle 3: Optimizing Storage and Representation Efficiency}
+\paragraph{Principle 3: Optimizing Storage Efficiency}
 Efficient storage and data representation are crucial for handling 
@@
 
-\paragraph{Principle 4: Updated Data Dictionaries and Documentation}
+\paragraph{Principle 4: Updating Data Dictionaries}
 Clear and accurate data documentation is essential for the effective use 
@@
 
-\paragraph{Principle 6: Addressing Data Transparency and Accessibility}
+\paragraph{Principle 6: Addressing Data Transparency}
 Transparency is a core tenet of open data, and governments must ensure 
@@
 \citep{chu2016data, hurbean2021open}.
-x`
 

### 2024-11-24 code/datacleansing-47 columns.R

@@ -1,2236 +1,0 @@
-#########################################################################
-# install.packages("ggplot2")
-# install.packages("campfin")
-# install.packages("stringr")
-# install.packages("stringdist")
-# install.packages("dply(raw data")
-# install.packages("styler")
-# install.packages("ggpmisc")
-# install.packages("lubridate")
-#install.packages("data.table")
-
-library(ggplot2)
-library(campfin)
-library(stringr)
-library(stringdist)
-library(dplyr)
-library(scales)
-library(ggpmisc)
-library(lubridate)
-library(data.table)
-
-#########################################################################
-programStart <- as.POSIXct(Sys.time())
-formattedStartTime <- format(programStart, "%Y-%m-%d %H:%M:%S")
-cat("\nExecution begins at:", formattedStartTime)
-cat("\n***** Program initialization *****")
-
-# Set the working directory to the "nyc311clean/code" directory to enable relative codes.
-# Alter this line of code to match your particular machine
-setwd("C:/Users/david/OneDrive/Documents/datacleaningproject/nyc311clean/code")
-
-# Set path for the data file
-main_data_file <- "47-columns.csv"
-data1File <- file.path("..", "..", "data", main_data_file)
-#data1File <- file.path("..", "..", "data", "test_data.csv")
-
-# Hard code the max_closed_date to be midnight of the date of the data export from NYC Open Data
-max_closed_date <- as.POSIXct("2024-08-31 23:59:59", format = "%Y-%m-%d %H:%M:%S")
-
-chart_directory_path <- file.path("..", "..", "charts", "2022-2023 study", "Core charts-47")
-
-writeFilePath <- file.path("..", "..", "data", "smaller.csv")
-
-# Define the path to the directory containing your function scripts
-functions_path <- "functions"
-
-# Source all .R files in the directory
-files <- list.files(functions_path, pattern = "\\.R$", full.names = TRUE)
-# Source each file
-lapply(files, source)
- 
-# Set scipen option to a large value
-options(scipen = 10)
-
-sink("../../console_output/47_core_console_output.txt")
-cat("\nExecution begins at:", formattedStartTime)
-cat("\n***** Program initialization *****")
-
-options(digits = 14) # Set the number of decimal places to 14
-
-#########################################################################
[…]

### 2024-11-24 code/datacleansing.R

@@ -26,7 +26,6 @@ library(sf)
 #########################################################################
 
 main_data_file <- "311_Service_Requests_from_2022-2023_AS_OF_09-15-2024.CSV"
-#main_data_file <- "JAN-SEP_2024_AS_OF_10-16-2024.CSV
 #main_data_file <- "smaller_test_data.csv"
 #main_data_file <- "extra_small.csv"
 
@@ -39,36 +38,73 @@ formattedStartTime <- format(programStart, "%Y-%m-%d %H:%M:%S")
 cat("\nExecution begins at:", formattedStartTime)
 cat("\n***** Program initialization *****")
 
-# Set the working directory to the "nyc311clean/code" directory to enable relative codes.
-# Alter this line of code to match your particular machine
-setwd("C:/Users/david/OneDrive/Documents/datacleaningproject/nyc311clean/code")
+# Prior to running program, ensure the R code is located in the working directory, 
+# and that the R functions are located in a sub-directory named "functions".
+# Data files should be placed in a sub-directory named "data".
 
-# Set path for the data file
-data1File <- file.path("..", "..", "data", main_data_file)
+# Create the sub-directories used during program execution.
+# Get the current working directory
+working_dir <- getwd()
 
-chart_directory_path <- file.path("..", "..", "charts", "2022-2023 study", "core charts")
+# Define the path for the main data file (CSV file)
+data_file <- file.path(working_dir, "data")
 
-writeFilePath <- file.path("..", "..", "data", "smaller.csv")
+# Create the main data directory if it doesn't already exist, then halt program execution.
+if (!dir.exists(data_file)) {
+  dir.create(data_file)
+  cat("\nCreating the data directory below the working directory. \nPlace program data there.")
+  stop("\n\nProgram terminated because data directory does not exist.")
+}
 
 # Define the path to the directory containing your function scripts
-functions_path <- "functions"
+functions_path <- file.path(working_dir, "functions")
 
-# Source all .R files in the directory
-files <- list.files(functions_path, pattern = "\\.R$", full.names = TRUE)
+# Create the directory if it doesn't already exist, and then halt program execution.
+if (!dir.exists(functions_path)) {
+  cat("\nCreate the 'functions' directory below the working directory. 
[…]

### 2024-11-24 code/functions/base_bar_chart.R

@@ -53,8 +53,9 @@ base_bar_chart <- function(dataset, x_col, y_col, chart_title, sub_title,
       axis.title = element_blank(),  # Remove x and y axis titles
       plot.title = element_text(hjust = 0.5, size = 12),
 #      plot.subtitle = element_text(size = 7),
-      axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1, face = "bold", size = 8),
+      axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1, face = "bold", size = 9),
       axis.text.y = element_text(face = "bold", size = 8),
+      panel.background = element_rect(fill = "gray100", color = "gray100"),
       aspect.ratio = 0.618033
     ) +
     

### 2024-11-24 code/functions/create_combo_chart.R

@@ -50,7 +50,7 @@ create_combo_chart <- function(
   # Step 1: Create the initial plot without specifying y-axis breaks
   combo_chart <- ggplot(summary_df) +
     
-    geom_bar(aes(x = agency, y = count), stat = "identity", fill = "#E69F00", width = 0.55) +
+    geom_bar(aes(x = agency, y = count), stat = "identity", fill = "#44AA99", width = 0.55) +
     
     geom_text(aes(label = count, x = agency, y = count), 
               colour = "black", hjust = 0.5, vjust = -0.5, size = 2.5) +
@@ -68,12 +68,13 @@ create_combo_chart <- function(
     labs(x = NULL, y = NULL) +
     
     theme(
-      axis.text.x = element_text(angle = 70, vjust = 1, hjust = 1, face = "bold", size = 7),
+      axis.text.x = element_text(angle = 70, vjust = 1, hjust = 1, face = "bold", size = 8.5),
       axis.text.y = element_text(face = "bold", size = 8),
       axis.text.y.right = element_text(color = "black", face = "bold", size = 8),
       plot.title = element_text(hjust = 0.5, size = 12),
       plot.subtitle = element_text(size = 7),
-      aspect.ratio = 0.618033
+      panel.background = element_rect(fill = "gray98", color = "gray98"),
+            aspect.ratio = 0.618033
     )
   
   # Only add the geom_line if there's more than one agency

### 2024-11-24 code/functions/create_violin_chart.R

@@ -8,15 +8,15 @@ create_violin_chart <- function(
   
   violin_chart <- ggplot(
     
-    data = dataset, 
+    data = dataset,
     
     aes(x = !!rlang::sym(x_axis_field), y = factor(1))) +
     
-    geom_jitter(width = 0.25, alpha = 0.4, color = "#0072B2", size = 1.9, shape = 17) +
+    geom_jitter(width = 0.2, alpha = 0.85, color = "#0072B2", size = 2, shape = 17) +
     
     geom_violin(linewidth = 0.7, fill = "transparent", color = "black") +
     
-    geom_boxplot(width = 0.25, fill = "#E69F00", color = "black", alpha = 0.6, 
+    geom_boxplot(width = 0.25, fill = "#E69F00", color = "black", alpha = 0.65, 
                  outlier.colour = "black", outlier.size = 0.75) +
     labs(
       title = chart_title,
@@ -29,8 +29,9 @@ create_violin_chart <- function(
     theme(
       plot.title = element_text(size = 13, hjust = 0.5),
       plot.subtitle = element_text(size = 7),
-      axis.text.x = element_text(face = "bold", size = 8),
-    )
+      axis.text.x = element_text(face = "bold", size = 9),
+      panel.background = element_rect(fill = "gray98", color = "gray98")
+      )
   
   suppressMessages(print(violin_chart))
   chart_path <- file.path(chart_directory_path, chart_file_name)

### 2024-11-24 code/functions/make_column_names_user_friendly.R

@@ -1,25 +1,66 @@
 #########################################################################
+  # names(dataset) <- gsub(
+  #   x = names(dataset),
+  #   pattern = "(\\.|\\s)+",
+  #   replacement = "_"
+  # )
+  # 
+  # ## Drop the trailing "."s
+  # names(dataset) <- gsub(
+  #   x = names(dataset),
+  #   pattern = "(_)+$",
+  #   replacement = ""
+  # )
+  # 
+  # ## Convert to lower case.
+  # names(dataset) <- tolower(names(dataset))
+  # 
+  # ## Return the revised column names
+  # return(dataset)
+  #}
 
-make_column_names_user_friendly<- function(dataset) {
+# Define the make_column_names_user_friendly function
+make_column_names_user_friendly <- function(dataset) {
   
+  # Replace dots and spaces with underscores
   names(dataset) <- gsub(
     x = names(dataset),
     pattern = "(\\.|\\s)+",
     replacement = "_"
   )
   
-  ## Drop the trailing "."s
+  # Remove all parentheses
   names(dataset) <- gsub(
     x = names(dataset),
-    pattern = "(_)+$",
+    pattern = "\\(|\\)",
     replacement = ""
   )
   
-  ## Convert to lower case.
+  # Drop trailing underscores
+  names(dataset) <- gsub(
+    x = names(dataset),
+    pattern = "_+$",
+    replacement = ""
+  )
+  
+  # Convert to lower case
   names(dataset) <- tolower(names(dataset))
   
-  ## Return the revised column names
+  # Return the dataset with modified column names
   return(dataset)
 }
+
+# Apply the function
+d311 <- make_column_names_user_friendly(d311)
+
+# Step 2: Reattach parentheses to the "location" field
+if ("location" %in% names(d311)) {
+  d311$location <- paste0("(", d311$location, ")")
+}
+
+# Print the resulting column names
+colnames(d311)
+
+
 
 #########################################################################

### 2024-11-24 code/functions/rank_by_agency.R

@@ -2,6 +2,7 @@
 
 # Take a dataset and sort it by agency. Then add percentage and cumulative percentage columns.
 # Dataset should contain agency as a fields
+
 rank_by_agency <- function(dataset) {
   sorted_dataset <- dataset %>%
     group_by(agency) %>%

### 2024-11-24 code/datacleansing.R
Code cleanup for images
Refactored code for manuscript image production.

@@ -92,8 +92,7 @@ if (!dir.exists(output_dir)) {
 }
 
 # Start directing console output to the file
-sink(output_file)
-#sink("../../console_output/2024_console_output.txt")
+#sink(output_file)
 
 cat("\nExecution begins at:", formattedStartTime)
 cat("\n***** Program initialization *****")
@@ -123,16 +122,7 @@ USPSzipcodes <- as.data.frame(fread(
   colClasses = "character"
 ))
 
-
-# USPSzipcodes <-
-#   read.csv(USPS_zipcode_reference_file,
-#     header = TRUE,
-#     colClasses = "character"
-#   )
-
-
 USPSzipcodes <- as.data.frame(USPSzipcodes)
-
 USPSzipcodes <- make_column_names_user_friendly(USPSzipcodes)
 
 # extract the 'delivery_zipcode' field
@@ -141,12 +131,6 @@ USPSzipcodesOnly <- USPSzipcodes[, "delivery_zipcode", drop = FALSE]
 #########################################################################
 # Load the USPS zipcode file
 USPS_abbreviations_reference_file <- file.path( data_file, "USPSabb.csv")
-# USPSabbreviations <-
-#   read.csv(USPS_abbreviations_reference_file,
-#     header = TRUE,
-#     colClasses = rep("character")
-#   )
-# 
 
 USPSabbreviations <- as.data.frame(fread(
   USPS_abbreviations_reference_file,
@@ -159,10 +143,10 @@ names(USPSabbreviations) <- c("full", "abb")
 
 #########################################################################
 # Load the main 311 SR data file. Set the read & write paths.
-main_code_file_path <- file.path( working_dir, "data", main_data_file)
+main_data_file_path <- file.path( working_dir, "data", main_data_file)
 
 d311 <- as.data.frame(fread(
-  main_code_file_path,
+  main_data_file_path,
   colClasses = "character"
 ))
 
@@ -267,7 +251,7 @@ latest_date_formatted <- format(latest_date, format = "%Y-%m-%d %H:%M:%S")
 earliest_title <- format(as.Date(earliest_date_formatted), format = "%Y-%m-%d")
 latest_title <- format(as.Date(latest_date_formatted), format = "%Y-%m-%d")
 
-chart_sub_title <- paste("(", earliest_title, "--", latest_title, ") total=", sep = "")
+#chart_sub_title <- paste("(", earliest_title, "--", latest_title, ") total=", sep = "")
 
[…]

### 2024-11-24 code/functions/areAllNumbers.R

@@ -1,27 +1,20 @@
+
 # #########################################################################
 # Validate that number fields are all numeric
 areAllNumbers <- function(numberField) {
   
-  # Identify the non-blank values using nzchar()
-  non_blank_indices <- which(nzchar(numberField))
+  # Remove blank and NA values
+  non_blank <- numberField[!is.na(numberField) & nzchar(numberField)]
   
-  # Subset the vector to keep only the non-blank values
-  numberField <- numberField[non_blank_indices]
-  
-  # remove blank and NAs
-  allNumbers <-
-    suppressWarnings(!is.na(as.numeric(numberField[numberField != ""])))
+  # Check if all remaining values are numeric
+  allNumbers <- suppressWarnings(!is.na(as.numeric(non_blank)))
   
   if (!all(allNumbers)) {
-    # find indices of values that are not numeric
-    non_numeric_values <-
-      numberField[grepl("[^[:digit:]]", numberField)]
+    # Extract the non-numeric values
+    non_numeric_values <- non_blank[!allNumbers]
     
-    cat(
-      "Non-numeric values in the vector: ",
-      non_numeric_values,
-      "\n"
-    )
+    cat("Non-numeric values in the vector:\n")
+    print(unique(non_numeric_values)) # Use unique to avoid redundant entries
   }
   return(all(allNumbers))
 }

### 2024-11-24 code/functions/are_valid_values.R

@@ -1,26 +1,30 @@
+
 # #########################################################################
 # Validate that all fields are in the list of allowable values
+# This function works on two lists: dataset and list_of_valid_values
 are_valid_values <- function(
     dataset,
-    listValidValues,
+    list_of_valid_values,
     field_name) {
+  
+  # Count the number of blank entires. Use later when computing percentage.
+  number_of_blank_entries <- sum(!nzchar(dataset) | is.na(dataset))
+  
   # Subset the vector to keep only the non-blank values
   dataset <- dataset[nzchar(dataset) & !is.na(dataset) & dataset != ""]
+  dataset_length <- length(dataset)
   
-  # determine valid values
-  check_list <- (dataset %in% listValidValues[, 1])
-  inList <- dataset[check_list]
-  notInList <- dataset[!check_list]
-  num_invalid <- length(notInList)
+  in_list <- dataset %in% list_of_valid_values
+  not_in_list <- dataset[!in_list]
+  
+  num_invalid <- length(not_in_list)
+  num_valid <- length(in_list)
   
   if (num_invalid > 0) {
-    invalid_d311_rows <- d311[(d311[[field_name]] %in% notInList), ]
+    invalid_rows <- dataset[dataset %in% not_in_list]
+    percentage_invalid <- round((num_invalid/(dataset_length - number_of_blank_entries)) * 100, 2)
+    unique_invalid <- length(unique(not_in_list))
     
-    # retrieve the number of blank fields for calculation purposes
-    number_of_blank_entries <-
-      missingDataPerColumn$blanks[missingDataPerColumn$field == field_name]
-    percentage_invalid <- round((num_invalid / (num_rows_d311 - number_of_blank_entries)) * 100, 2)
-    unique_invalid <- length(unique(notInList))
     cat(
       "\n\nThere are",
       format(num_invalid, big.mark = ","),
@@ -35,29 +39,25 @@ are_valid_values <- function(
     )
     
     # Sort the table in descending order
-    sorted_invalid_table <-
-      sort(table(notInList), decreasing = TRUE)
+    sorted_invalid_table <- sort(table(not_in_list), decreasing = TRUE)
     
     # Convert the table to a data frame and calculate the percentage column
-    invalid_df <-
-      data.frame(
-        invlaid_names = names(sorted_invalid_table),
-        count = as.numeric(sorted_invalid_table)
-      )
+    invalid_df <- data.frame(
[…]

### 2024-11-24 code/functions/base_bar_chart.R

@@ -9,7 +9,8 @@ base_bar_chart <- function(dataset, x_col, y_col, chart_title, sub_title,
                            horizontal_adjustment_max = 0.5,
                            vertical_adjustment_max = -1,
                            console_print_out_title = "Data Summary",
-                           chart_file_name = NULL) {
+                           chart_file_name = NULL,
+                           chart_directory) {
   
   # Step 1: Print Data Summary
   cat("\n\n", console_print_out_title, " (first 20 rows):\n", sep = "")
@@ -130,9 +131,11 @@ base_bar_chart <- function(dataset, x_col, y_col, chart_title, sub_title,
   print(bar_chart)
   if (!is.null(chart_file_name)) {
     chart_width <- 10
-    chart_width <- chart_width / 1.618  # Golden ratio
-    chart_path <- file.path(chart_directory_path, chart_file_name)
-    ggsave(chart_path, plot = bar_chart, width = chart_width, height = chart_width, dpi = 300)
+#    chart_height <- chart_width / 1.618  # Golden ratio
+    chart_height <- chart_width / 1.2  # Golden ratio
+    chart_path <- file.path(chart_directory, chart_file_name)
+    ggsave(chart_path, plot = bar_chart, width = chart_width, 
+           height = chart_height, dpi = 300)
   }
 }
 

### 2024-11-24 code/functions/check_date_format.R

@@ -1,18 +1,0 @@
-# #########################################################################
-# Function to check if a column is a valid date format
-check_date_format <- function(column_name) {
-  date_column <- d311[[column_name]]
-  date_check <- try(as.Date(date_column, format = "%m/%d/%Y %I:%M:%S %p"))
-  
-  if (inherits(date_check, "try-error")) {
-    invalid_dates <- date_column[!grepl("^\\d{2}/\\d{2}/\\d{4} \\d{2}:\\d{2}:\\d{2} [APMapm]{2}$", date_column)]
-    if (length(invalid_dates) > 0) {
-      print(paste("Invalid dates in", column_name, ":", invalid_dates, sep = ""))
-    }
-    return(paste("Field '", column_name, "' is not in a valid date format", sep = ""))
-  } else {
-    return(paste("Field '", column_name, "' is in a valid date format", sep = ""))
-  }
-}
-
-#########################################################################

### 2024-11-24 code/functions/create_bar_chart_categorical.R

@@ -18,7 +18,8 @@ create_bar_chart_categorical <- function(
     horizontal_adjustment_max = 0.5,
     vertical_adjustment_max = -1,
     console_print_out_title = "Data Summary",
-    chart_file_name = NULL) 
+    chart_file_name = NULL,
+    chart_directory) 
 {
   
   # Call base_bar_chart without x_scale
@@ -40,7 +41,8 @@ create_bar_chart_categorical <- function(
     horizontal_adjustment_max, 
     vertical_adjustment_max,
     console_print_out_title, 
-    chart_file_name
+    chart_file_name,
+    chart_directory
   )
 }
 

### 2024-11-24 code/functions/create_bar_chart_numeric.R

@@ -9,14 +9,15 @@ create_bar_chart_numeric <- function(dataset, x_col, y_col, chart_title = "Bar C
                                      horizontal_adjustment_max = 0.5,
                                      vertical_adjustment_max = -1,
                                      console_print_out_title = "Data Summary",
-                                     chart_file_name = NULL) {
-  
+                                     chart_file_name = NULL,
+                                     chart_directory)
+ {
   # Call to base_bar_chart without x_scale
   base_bar_chart(dataset, x_col, y_col, chart_title, sub_title, 
                  y_axis_labels, add_maximum, add_minimum, add_mean, 
                  add_median, add_sd, add_second_maximum, add_trendline, 
                  extra_line, horizontal_adjustment_max, vertical_adjustment_max, 
-                 console_print_out_title, chart_file_name)
+                 console_print_out_title, chart_file_name, chart_directory)
 }
 
 

### 2024-11-24 code/functions/create_combo_chart.R

@@ -4,7 +4,8 @@ create_combo_chart <- function(
     dataset,
     chart_title = NULL,
     chart_file_name = NULL,
-    console_print_out_title = "Data Summary"
+    console_print_out_title = "Data Summary",
+    chart_directory
 ) {
   
   # Step 1: Create summary_df sorted by count and calculate cumulative percentage
@@ -38,10 +39,10 @@ create_combo_chart <- function(
   max_count <- max(summary_df$count)
   total_count <- sum(summary_df$count)
   
-  earliest_date <- min(d311$created_date, na.rm = TRUE)
-  latest_date <- max(d311$created_date, na.rm = TRUE)
-  earliest_title <- format(as.Date(earliest_date), format = "%Y-%m-%d")
-  latest_title <- format(as.Date(latest_date), format = "%Y-%m-%d")
+  # earliest_date <- min(dataset$created_date, na.rm = TRUE)
+  # latest_date <- max(dataset$created_date, na.rm = TRUE)
+  # earliest_title <- format(as.Date(earliest_date), format = "%Y-%m-%d")
+  # latest_title <- format(as.Date(latest_date), format = "%Y-%m-%d")
   
   # result <- calculate_values(max_count)
   # starting_value <- result$starting_value
@@ -99,10 +100,10 @@ create_combo_chart <- function(
   
   
   # Save the plot
-  chart_path <- file.path(chart_directory_path, chart_file_name)
+  chart_path <- file.path(chart_directory, chart_file_name)
   chart_width <- 10
 #  chart_height <- chart_width / 1.618
-  chart_height <- chart_width / 1.3
+  chart_height <- chart_width / 1.2
     ggsave(chart_path, plot = combo_chart, width = chart_width, height = chart_height, dpi = 300)
 }
 

### 2024-11-24 code/functions/create_violin_chart.R

@@ -12,7 +12,7 @@ create_violin_chart <- function(
     
     aes(x = !!rlang::sym(x_axis_field), y = factor(1))) +
     
-    geom_jitter(width = 0.2, alpha = 0.85, color = "#0072B2", size = 2, shape = 17) +
+    geom_jitter(width = 0.2, height = 0.4, alpha = 0.85, color = "#0072B2", size = 2, shape = 17) +
     
     geom_violin(linewidth = 0.7, fill = "transparent", color = "black") +
     
@@ -35,7 +35,10 @@ create_violin_chart <- function(
   
   suppressMessages(print(violin_chart))
   chart_path <- file.path(chart_directory_path, chart_file_name)
-  suppressMessages(ggsave(chart_path, plot = violin_chart, width = 10, height = 8))
+  chart_width <- 10
+  chart_height <- chart_width / 1.2
+  ggsave(chart_path, plot = violin_chart, dpi = 300,
+                          width = chart_width, height = chart_height)
 }
 
 #########################################################################

### 2024-11-24 code/functions/cross_street_analysis.R

@@ -1,8 +1,13 @@
-# #########################################################################
+
+##########################################################################
 cross_street_analysis <- function(
     dataset,
     cross_street,
-    intersection_street) {
+    intersection_street,
+    chart_directory) {
+  
+  num_rows_dataset <- nrow(dataset)
+  
   # Remove extra blanks
   dataset <- dataset %>%
     mutate(
@@ -121,6 +126,7 @@ cross_street_analysis <- function(
       non_dup_streets,
       chart_title,
       chart_file_name,
+      chart_directory,
       console_print_out_title = paste0("Summary of non-Matching '", cross_street, "' and '", intersection_street, "'by Agency")
     )
   }
@@ -132,7 +138,7 @@ cross_street_analysis <- function(
   cat(
     "\nThere are", format(num_rows_cross_street_blank, big.mark = ","),
     "occurrences where'", cross_street, "'is blank, \nbut '", intersection_street, "' is not blank representing",
-    round((num_rows_cross_street_blank / num_rows_d311 * 100), 2), "% of total rows."
+    round((num_rows_cross_street_blank / num_rows_dataset * 100), 2), "% of total rows."
   )
   
   cat(
@@ -149,7 +155,7 @@ cross_street_analysis <- function(
   cat(
     "\nThere are", format(num_rows_intersection_street_blank, big.mark = ","),
     "occurrences where'", intersection_street, "'is blank, \nbut '", cross_street, "' is not blank representing",
-    round(num_rows_intersection_street_blank / num_rows_d311 * 100, 2),
+    round(num_rows_intersection_street_blank / num_rows_dataset * 100, 2),
     "% of total rows."
   )
   
@@ -221,12 +227,12 @@ cross_street_analysis <- function(
       )
     )
   
-  summary_street$percentage[1] <- round(((num_rows_equal - num_rows_both_blank)/num_rows_d311) * 100, 2)
-  summary_street$percentage[2] <- round((num_rows_both_blank / num_rows_d311) * 100, 2)
-  summary_street$percentage[3] <- round((num_rows_non_dup / num_rows_d311) * 100, 2)
+  summary_street$percentage[1] <- round(((num_rows_equal - num_rows_both_blank)/num_rows_dataset) * 100, 2)
+  summary_street$percentage[2] <- round((num_rows_both_blank / num_rows_dataset) * 100, 2)
[…]

### 2024-11-24 code/functions/detect_duplicates.R

@@ -9,17 +9,19 @@ detect_duplicates <- function(
   
   rows_condition_1 <- dataset[[reference_field]] != dataset[[duplicate_field]]
   
+  num_rows_dataset <- nrow(dataset)
+  
   all_non_matching_rows <- rows_condition_1
   non_matching_fields <- dataset[all_non_matching_rows, ]
   num_non_matching_fields <- nrow(non_matching_fields)
-  non_matching_percentage <- round((num_non_matching_fields / num_rows_d311) * 100, 2)
+  non_matching_percentage <- round((num_non_matching_fields / num_rows_dataset) * 100, 2)
   non_matching_fields <- non_matching_fields %>%
     select("unique_key", all_of(reference_field), all_of(duplicate_field), "agency")
   
   all_matching_rows <- rows_condition_0
   matching_fields <- dataset[all_matching_rows, ]
   num_matching_fields <- nrow(matching_fields)
-  matching_percentage <- round((num_matching_fields / num_rows_d311) * 100, 2)
+  matching_percentage <- round((num_matching_fields / num_rows_dataset) * 100, 2)
   matching_fields <- matching_fields %>%
     select("unique_key", all_of(reference_field), all_of(duplicate_field), "agency")
   

### 2024-11-24 code/functions/make_column_names_user_friendly.R

@@ -49,17 +49,17 @@ make_column_names_user_friendly <- function(dataset) {
   # Return the dataset with modified column names
   return(dataset)
 }
-
-# Apply the function
-d311 <- make_column_names_user_friendly(d311)
-
-# Step 2: Reattach parentheses to the "location" field
-if ("location" %in% names(d311)) {
-  d311$location <- paste0("(", d311$location, ")")
-}
-
-# Print the resulting column names
-colnames(d311)
+# 
+# # Apply the function
+# d311 <- make_column_names_user_friendly(dataset)
+# 
+# # Step 2: Reattach parentheses to the "location" field
+# if ("location" %in% names(dataset)) {
+#   dataset$location <- paste0("(", d311$location, ")")
+# }
+# 
+# # Print the resulting column names
+# colnames(d311)
 
 
 

### 2024-11-24 code/timelinecharts.R

@@ -4,7 +4,7 @@
 # install.packages('zoo')
 # install.packages("ggpmisc")
 # install.packages("lubridate")
-rm(list = ls(), envir = .GlobalEnv)
+#rm(list = ls(), envir = .GlobalEnv)
 
 library(ggplot2)
 library(scales)
@@ -20,9 +20,8 @@ cat("\nExecution begins at:", formattedStartTime)
 #########################################################################
 
 # Set path for the data file
-main_data_file <- "10-year 2014-2023.csv"
-#main_data_file <- "311_Service_Requests_from_2022-2023_AS_OF_09-15-2024.csv"
-#main_data_file <- "JAN-SEP_2024_AS_OF_10-16-2024.csv"
+#main_data_file <- "10-year 2014-2023.csv"
+main_data_file <- "311_Service_Requests_from_2022-2023_AS_OF_09-15-2024.csv"
 #main_data_file <- "smaller_test_data.csv"
 #main_data_file <- "extra_small.csv"
 
@@ -32,16 +31,9 @@ cat("\n***** Program initialization *****")
 
 setwd("C:/Users/david/OneDrive/Documents/datacleaningproject/nyc311clean/code")
 
-# Define the path to the directory containing your function scripts
-functions_path <- "functions"
+working_dir <- getwd()
 
-# Source all .R files in the directory
-files <- list.files(functions_path, pattern = "\\.R$", full.names = TRUE)
-
-# Source each file
-lapply(files, source)
-
-data1File <- file.path("..", "..", "data", main_data_file)
+data_directory <- file.path(working_dir, "data")
 
 # Extract the first two digits for file naming purposes
 year_digits <- substr(main_data_file, 1, 2)
@@ -49,21 +41,43 @@ year_digits <- substr(main_data_file, 1, 2)
 if (year_digits != "10") {
   year_digits <- "2"
 }
+
 # Set path for the chart directory
+# Create the directory if it doesn't already exist
+if (!dir.exists(chart_directory_path)) {
+  dir.create(chart_directory_path)
+}
 
 chart_prefix <- paste0(year_digits, "-year")
-# chart_prefix <- "Jan-Sep_2024"
-chart_directory_path <- file.path("..", "..", "charts", "2022-2023 study", chart_prefix)
+
+# Define the path for the charts
+chart_directory_path <- file.path(working_dir, "charts")
 
 file_name_prefix <- chart_prefix # to name individual chart files
+
[…]

### 2024-11-24 manuscript/manuscript.tex
Code cleanup for images
Refactored code for manuscript image production.
minor edits/fixes

@@
 include the addition of online and mobile channels in 2009, a
-record monhtly high of 348,463 Service Requests in 
+record monhtly high of 348,463 SRs in 
 August 2020 (associated with the the COVID-19 
@@
 \item There are 41 columns (fields) of data for each row, with each
-  row representing a single Service Request (SR). To facilitate 
+  row representing a single SR. To facilitate 
   processing, column names were made lower case and 
@@
 	\centering
-	\includegraphics[width = \textwidth]{SRs_by_Agency.pdf}
-  	\caption{SR Counts by Agency for the time period 2022-2023}
+	\includegraphics[width = \textwidth]{SRs_by_Agency-crop.pdf}
+  	\caption{311 SR counts by agency for the time period 2022--2023.}
 	\label{fig:SRcountbyAgency}
@@
  \centering
-  \includegraphics[width = \textwidth]{SR_by_Complaint_Type.pdf} 
-  \caption{Top 20 Complaint Types for the time period 2022-2023} 
+  \includegraphics[width = \textwidth]{SR_by_Complaint_Type-crop.pdf} 
+  \caption{Top 20 complaint types for the 311 SR during 2022--2023.} 
   \label{fig:SR_complaints}
@@
 \paragraph{Undocumented Fields:} The 311 SR 
-The 311 Service Request Data Dictionary identifies 
+The 311 SR Data Dictionary identifies 
 41 data columns (fields) and provides related information 
@@
 usage. These six fields are not included in
-the 311 Service Request Data Dictionary. Additionally, our analysis 
+the 311 SR Data Dictionary. Additionally, our analysis 
 revealed that these six ``computed'' fields have 
@@
 	\centering
-  	\includegraphics[width=\textwidth]{BlankFields.pdf}
-	\caption{Number and Percentage of Empty/Blank Entries for the time period 2022-2023}
+  	\includegraphics[width=\textwidth]{BlankFields-cro.pdf}
+	\caption{Number and percentage of empty/blank entries for the
+          311 SR data during 2022--2023.}
 	\label{fig:blank_fields}
@@
 the \texttt{unique\_key} field was indeed unique, as required. 
-Unfortunately, the 311 Service Request Data Dictionary specifies very few domains of 
-acceptable values for other fields. Nonetheless, the following fields were tested and found to comply with their expected domains, as determined by public 
+Unfortunately, the 311 SR Data Dictionary specifies very few domains
+of acceptable values for other fields. Nonetheless, the following
+fields were tested and found to comply with their expected
+domains, as determined by public
 usage and as evidenced in other historical datasets:
@@
     \centering
-    \includegraphics[width=\textwidth]{post_closed_violin_chart.pdf}
-    \caption{Post-closed \texttt{resolution\_action\_update\_date}(s) >30 days}
-    \label{fig:resolution-violin}
+    \includegraphics[width=\textwidth]{negative_duration_SR_violin.pdf}
+    \caption{Distribution of negative response times.}
+    \label{fig:negative-duration-violin}
   \end{subfigure}
@@
     \centering
-    \includegraphics[width=\textwidth]{negative_duration_SR_violin.pdf}
-    \caption{Distribution of Negative Durations.}
-    \label{fig:negative-duration-violin}
+    \includegraphics[width=\textwidth]{post_closed_violin_chart.pdf}
+    \caption{Post-closed \texttt{resolution\_action\_update\_date}(s) >30 days}
+    \label{fig:resolution-violin}
   \end{subfigure}
   % Main figure caption
-  \caption{Comparison of Violin Plots: (a) Post-Closed Resolution and (b) Negative Durations for the time period 2022-2023}
+  \caption{Violin plots for negative response times and post-closed
+    resolution during 2022--2023.}
   \label{fig:violin-plots}
@@
   \centering
-  \caption{Sample  Negative Durations [excluding extreme values that are <$-$730 days]}
+  \caption{Sample  negative durations [excluding extreme values that are <$-$730 days]}
    \label{tab:largest-errors}
@@
 of this anomaly is well outside the 3$\sigma$ level as shown 
-in Figure~\ref{fig:stacked}
+in Figure~\ref{fig:exacthours}
 
@@
         \includegraphics[width=\textwidth]{2-year-trend_SRs_created_on_the_hour.pdf}
-        \caption{SRs Created Exactly on the Hour}
+        \caption{SRs created exactly on the hour}
         \label{fig:busiestcreated}
@@
         \includegraphics[width=\textwidth]{2-year-trend_SRs_closed_on_the_hour.pdf}
-        \caption{SRs Closed Exactly on the Hour}
+        \caption{SRs closed exactly on the hour}
         \label{fig:busiestclosed}
@@
     % Main caption
-    \caption{Comparison of SRs Created \& Closed by Hour-of-the-Day for the time period 2022-2023}
-    \label{fig:comparison}
+    \caption{Comparison of 311 SRs created \& closed by hour-of-the-day during 2022--2023.}
+    \label{fig:exacthours}
 \end{figure}
@@
 
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=\textwidth]{post_closed_violin_chart.pdf}
-  \caption{Post-closed \texttt{resolution\_action\_update\_date}(s) >30 days.}
-  \label{fig:resolution-violin}
-\end{figure}
-
 \subsection{Accuracy and Precision:}
@@
 \paragraph{\texttt{landmark} \& \texttt{street\_name}:} Per the 
-311 Service Request Data Dictionary, the \texttt{landmark} 
+311 SR Data Dictionary, the \texttt{landmark} 
 field \textit{can refer to any noteworthy location, 
@@
 
-The 311 Service Request Data Dictionary would greatly benefit from an update.
-We found several discrepancies between the Data Dictionary and the actual data, which could potentially mislead analytic efforts.
-Here are some examples:
+The 311 SR Data Dictionary would greatly benefit from an update. We
+found several discrepancies between the Data Dictionary and the actual
+data, which could potentially mislead analytic efforts. Here are some
+examples:
 

### 2024-12-05 manuscript/manuscript.tex

@@
 	\centering
-  	\includegraphics[width=\textwidth]{BlankFields-cro.pdf}
+  	\includegraphics[width=\textwidth]{BlankFields-crop.pdf}
 	\caption{Number and percentage of empty/blank entries for the
@@
 
+\jy{The font sizes of annotations in Figure 4 and 5 need to be
+  increased. In addition, the size of the pdf figure could be made
+  arguments to the R function that generated them (instead of being
+  fixed in within the function).}
+
 \begin{figure}[tbp]
@@
 occurring over 2 years later. As shown in Figure~\ref{fig:resolution-violin}, 
-there are 10,265 SRs that were updated >30 days after the 
+there are 10,265 SRs that were updated $>30$ days after the 
 \texttt{closed\_date} (Not that Figure~\ref{fig:resolution-violin} excludes 
-values >730 days, an arbitrary cutoff to exclude 
+values $>730$ days, an arbitrary cutoff to exclude 
 infeasible dates such as \textit{1900-01-01}. The chart 

### 2024-12-12 manuscript/manuscript.tex
Spell check and correction
Working on JY's comments.

@@
   spaces were replaced with underscores, e.g. \texttt{complaint\_type}. 
-  This naming convention is used throught this paper.
+  This naming convention is used throughout this paper.
 
@@
 reflecting a growing public reliance on and usage of 
-the 311 system. (During this 10-year time-frame, the 
+the 311 system. (During this 10-year time frame, the 
 population of NYC grew just 0.6\%.). The most significant increase
@@
 percentage of SRs handled by each
-agency over the 2-year time-frame. Note that the
+agency over the 2-year time frame. Note that the
 distribution of SRs is heavily concentrated among a few key 
@@
 The \textit{Mostly Empty} category includes fields such as
-\texttt{taxi\_company\_borough}, \texttt{due\_date},
-\jy{These variables seem only relevant to certain agencies, which
-  could be made available in a separate table.}
+\texttt{taxi\_company\_borough}, \texttt{due\_date}
 \texttt{road\_ramp}, and \texttt{bridge\_highway\_name}.
+
+Of particular note is that several of the \textit{Mostly Empty} fields
+are used by only one or two Agencies. Example:
+
 The \textit{Partially Empty} category includes fields such as
 \texttt{location\_type}, \texttt{landmark}, 
-and \texttt{cross\_street\_1 \& \_2}. And the \textit{Few/None Empty} 
+and \texttt{cross\_street\_1 \& \_2}.
+
+And the \textit{Few/None Empty} 
 category includes \texttt{created\_date}, \texttt{complaint\_type},
-\texttt{agency}, and \texttt{status}. For some analysis efforts, it may be
-prudent to inquire further as to why some fields are 
-mostly or almost always blank. Figure~\ref{fig:blank_fields} is 
+\texttt{agency}, and \texttt{status}.
+
+For some analysis efforts, it may be prudent to inquire 
+further as to why some fields are mostly or almost 
+always blank. Figure~\ref{fig:blank_fields} is 
 a graphic depiction of total empty (blank \& N/As) for each 
-field illustrating the grouping into 
-the Most, Partial, and Few/None categories.
+field illustrating the grouping into the Most, 
+Partial, and Few/None categories.
 
@@
 
-\jy{Could refer to cross-plotform data format such as Arrow
+\jy{Could refer to cross-platform data format such as Arrow
   \citep{bates2024csv}.}

### 2024-12-16 manuscript/manuscript.tex
Revised chart margin and height/width sizing to eliminate blank space
Incorporating JY's suggestions on adding parameters to chart sizing, including height, width, and margins (top, left, bottom, right). Removed the "-crop" PDF files and replaced with new charts generated by the R programs.

@@
 	\centering
-	\includegraphics[width = \textwidth]{SRs_by_Agency-crop.pdf}
+	\includegraphics[width = \textwidth]{SRs_by_Agency.pdf}
   	\caption{311 SR counts by agency for the time period 2022--2023.}
@@
  \centering
-  \includegraphics[width = \textwidth]{SR_by_Complaint_Type-crop.pdf} 
+  \includegraphics[width = \textwidth]{SR_by_Complaint_Type.pdf} 
   \caption{Top 20 complaint types for the 311 SR during 2022--2023.} 
@@
 
-The \textit{Mostly Empty} category includes fields such as
+The Mostly Empty category includes:
 \texttt{taxi\_company\_borough}, \texttt{due\_date}
-\texttt{road\_ramp}, and \texttt{bridge\_highway\_name}.
-
-Of particular note is that several of the \textit{Mostly Empty} fields
-are used by only one or two Agencies. Example:
-
-The \textit{Partially Empty} category includes fields such as
+\texttt{road\_ramp}, and \texttt{bridge\_highway\_name}. The 
+Partially Empty category includes:
 \texttt{location\_type}, \texttt{landmark}, 
-and \texttt{cross\_street\_1 \& \_2}.
-
-And the \textit{Few/None Empty} 
+and \texttt{cross\_street\_1 \& \_2}. And the Few/None Empty 
 category includes \texttt{created\_date}, \texttt{complaint\_type},
-\texttt{agency}, and \texttt{status}.
+\texttt{agency}, and \texttt{status}. Of particular note is that 
+several of the Mostly Empty fields are used by only one or 
+two Agencies (See: \autoref{tab:field-usage-summary}).
+
+\begin{table}[htbp]
+    \centering
+    \caption{Fields used by only one or two Agencies with limited non-blank entires}
+    \label{tab:field-usage-summary}
+    \begin{tabular}{l r r p{4cm}}
+        \toprule
+        \textbf{Field} & \textbf{\# of entries} & \textbf{\% non-blank} & \textbf{Agencies using} \\
+        \midrule
+        \texttt{taxi\_company\_borough} & 3,567  & 0.06\% & TLC \\
+        \texttt{vehicle\_type}           & 18,642 & 0.29\% & TLC, NYPD \\
+        \texttt{due\_date}               & 24,414 & 0.38\% & DSNY \\
+        \texttt{taxi\_pick\_up\_location} & 64,954 & 1.02\% & TLC \\
+        \bottomrule
+    \end{tabular}
+\end{table}
 
@@
 a graphic depiction of total empty (blank \& N/As) for each 
-field illustrating the grouping into the Most, 
-Partial, and Few/None categories.
+field illustrating the grouping into the Mostly Empty, 
+Partially Empty, and Few/None Empty categories.
 
@@
 	\centering
-  	\includegraphics[width=\textwidth]{BlankFields-crop.pdf}
+  	\includegraphics[width=\textwidth]{BlankFields.pdf}
 	\caption{Number and percentage of empty/blank entries for the
@@
     \includegraphics[width=\textwidth]{negative_duration_SR_violin.pdf}
-    \caption{Distribution of negative response times.}
+    \caption{Distribution of SR durations.}
     \label{fig:negative-duration-violin}
@@
   % Main figure caption
-  \caption{Violin plots for negative response times and post-closed
-    resolution during 2022--2023.}
+  \caption{Violin plots for negative SR durations and post-closed
+    resolution updates (2022--2023.}
   \label{fig:violin-plots}
@@
 in Figure~\ref{fig:exacthours}
-
-\jy{The font sizes of annotations in Figure 4 and 5 need to be
-  increased. In addition, the size of the pdf figure could be made
-  arguments to the R function that generated them (instead of being
-  fixed in within the function).}
 
@@
 
-\jy{Could refer to cross-platform data format such as Arrow
-  \citep{bates2024csv}.}
+One promising alternative to reducing data storage requirements 
+for this large 311 SR dataset (3.7GB) would be employment of 
+the cross-platform Apache Arrow package. Apache Arrow (arrow.apache.org) 
+is an innovative, cross-language development platform designed for 
+in-memory data. It defines a standard for representing columnar data, 
+enabling efficient analytics and real-time data processing. As such,
+it offers dramatic improvements in file storage size, as well as 
+efficient analysis and cross-platform support for R, Python, 
+and Julia programming languages. Remarkable storage size 
+reduction (within several orders of magnitude) is possible. \citep{bates2024csv}.
 

### 2024-12-16 code/datacleansing.R

@@ -1,15 +1,15 @@
 #########################################################################
-# install.packages("ggplot2")
-# install.packages("campfin")
-# install.packages("stringr")
-# install.packages("stringdist")
-# install.packages("dply(raw data")
-# install.packages("styler")
-# install.packages("ggpmisc")
-# install.packages("lubridate")
-# install.packages("data.table")
-# install.packages("sf")
-# install.packages("tidyverse")
+install.packages("ggplot2")
+install.packages("campfin")
+install.packages("stringr")
+install.packages("stringdist")
+install.packages("dply(raw data")
+install.packages("styler")
+install.packages("ggpmisc")
+install.packages("lubridate")
+install.packages("data.table")
+install.packages("sf")
+install.packages("tidyverse")
 
 library(tidyr)
 library(ggplot2)
@@ -21,9 +21,10 @@ library(scales)
 library(ggpmisc)
 library(lubridate)
 library(data.table)
-library(sf)
+library(sf) 
 
 #########################################################################
+rm(list = ls())
 
 main_data_file <- "311_Service_Requests_from_2022-2023_AS_OF_09-15-2024.CSV"
 #main_data_file <- "smaller_test_data.csv"
@@ -35,16 +36,10 @@ max_closed_date <- as.POSIXct("2024-09-15 23:59:59", format = "%Y-%m-%d %H:%M:%S
 #########################################################################
 programStart <- as.POSIXct(Sys.time())
 formattedStartTime <- format(programStart, "%Y-%m-%d %H:%M:%S")
-cat("\nExecution begins at:", formattedStartTime)
 cat("\n***** Program initialization *****")
-
-# Prior to running program, ensure the R code is located in the working directory, 
-# and that the R functions are located in a sub-directory named "functions".
-# Data files should be placed in a sub-directory named "data".
 
 setwd("C:\\Users\\David\\OneDrive\\Documents\\datacleaningproject\\nyc311clean")
 
-# Create the sub-directories used during program execution.
 # Get the current working directory
 working_dir <- getwd()
 
@@ -58,28 +53,36 @@ output_file <- file.path(output_dir, "core_console_output.txt")
 # Define the path to the directory containing your function scripts
 functions_path <- file.path(base_dir, "functions")
 
+# Get all .R files in the "functions" sub-directory
[…]

### 2024-12-16 code/functions/base_bar_chart.R

@@ -10,7 +10,10 @@ base_bar_chart <- function(dataset, x_col, y_col, chart_title, sub_title,
                            vertical_adjustment_max = -1,
                            console_print_out_title = "Data Summary",
                            chart_file_name = NULL,
-                           chart_directory) {
+                           chart_directory = ".",
+                           chart_width = 10,
+                           chart_height = 7,
+                           annotation_size = 5) {
   
   # Step 1: Print Data Summary
   cat("\n\n", console_print_out_title, " (first 20 rows):\n", sep = "")
@@ -33,12 +36,7 @@ base_bar_chart <- function(dataset, x_col, y_col, chart_title, sub_title,
   if (inherits(dataset[[x_col]], "Date")) {
     x_scale <- scale_x_date(expand = c(0.01, 0), labels = scales::date_format("%Y-%m"), breaks = scales::date_breaks("6 months"))
   } else if (inherits(dataset[[x_col]], "POSIXct") || inherits(dataset[[x_col]], "POSIXt")) {
-    # Default to "6 months" break for POSIXct, unless finer granularity is detected
-    if (any(diff(as.numeric(dataset[[x_col]])) < 86400)) { # 86400 seconds = 1 day
-      x_scale <- scale_x_datetime(expand = c(0.01, 0), labels = scales::date_format("%H:%M"), breaks = scales::date_breaks("2 hours"))
-    } else {
-      x_scale <- scale_x_datetime(expand = c(0.01, 0), labels = scales::date_format("%Y-%m"), breaks = scales::date_breaks("6 months"))
-    }
+    x_scale <- scale_x_datetime(expand = c(0.01, 0), labels = scales::date_format("%Y-%m"), breaks = scales::date_breaks("6 months"))
   } else if (is.numeric(dataset[[x_col]])) {
     x_scale <- scale_x_continuous(expand = c(0.01, 0))
   } else {
@@ -51,94 +49,51 @@ base_bar_chart <- function(dataset, x_col, y_col, chart_title, sub_title,
     x_scale +
     scale_y_continuous(labels = y_axis_labels) +
     theme(
-      axis.title = element_blank(),  # Remove x and y axis titles
+      axis.title = element_blank(),
       plot.title = element_text(hjust = 0.5, size = 12),
-#      plot.subtitle = element_text(size = 7),
       axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1, face = "bold", size = 9),
[…]

### 2024-12-16 code/functions/create_combo_chart.R

@@ -5,7 +5,12 @@ create_combo_chart <- function(
     chart_title = NULL,
     chart_file_name = NULL,
     console_print_out_title = "Data Summary",
-    chart_directory
+    chart_directory,
+    chart_width = 10,
+    chart_height = 7,
+    annotation_size = 4.5,
+    num_x_labels = 20,          # New parameter for when to skip labels
+    skip_frequency = 2          # New parameter for how often to skip labels
 ) {
   
   # Step 1: Create summary_df sorted by count and calculate cumulative percentage
@@ -18,7 +23,6 @@ create_combo_chart <- function(
       cumulative_percentage = cumsum(percentage)
     ) %>%
     slice_head(n = 20)
-  
   
   # Step 2: Reorder the agency factor based on the descending count
   summary_df <- summary_df %>%
@@ -36,52 +40,59 @@ create_combo_chart <- function(
   
   print(format(summary_df, justify = "left"), row.names = FALSE)
   
+  # Count the number of unique x-axis data points
+  total_x_labels <- length(unique(summary_df$agency))
+  
+  # Utility function to blank every 'skip_frequency' label if total x labels > num_x_labels
+  blank_alternate_labels <- function(values) {
+    if (total_x_labels > num_x_labels) {
+      ifelse(seq_along(values) %% skip_frequency == 0, "", values)
+    } else {
+      values
+    }
+  }
+  
+  # Apply the label blanking logic to the count and cumulative percentage labels
+  count_labels <- blank_alternate_labels(summary_df$count)
+  cumulative_labels <- blank_alternate_labels(round(summary_df$cumulative_percentage, 2))
+  
   # Chart creation
   max_count <- max(summary_df$count)
   total_count <- sum(summary_df$count)
   
-  # earliest_date <- min(dataset$created_date, na.rm = TRUE)
-  # latest_date <- max(dataset$created_date, na.rm = TRUE)
-  # earliest_title <- format(as.Date(earliest_date), format = "%Y-%m-%d")
-  # latest_title <- format(as.Date(latest_date), format = "%Y-%m-%d")
-  
-  # result <- calculate_values(max_count)
-  # starting_value <- result$starting_value
-  # increment <- result$increment
-
   # Step 1: Create the initial plot without specifying y-axis breaks
   combo_chart <- ggplot(summary_df) +
     
     geom_bar(aes(x = agency, y = count), stat = "identity", fill = "#44AA99", width = 0.55) +
     
[…]

### 2024-12-16 code/functions/create_violin_chart.R

@@ -1,16 +1,22 @@
-# #########################################################################
+##########################################################################
 create_violin_chart <- function(
     dataset,
     x_axis_title,
     x_axis_field,
     chart_title,
     chart_file_name,
-    chart_directory) {
+    chart_directory,
+    chart_width = 10,
+    chart_height = 7,
+    margin_top = 1,     # Default margin values (units in pts)
+    margin_right = 2,
+    margin_bottom = 1,
+    margin_left = 2
+) {
   
+  # Create the violin chart
   violin_chart <- ggplot(
-    
     data = dataset,
-    
     aes(x = !!rlang::sym(x_axis_field), y = factor(1))) +
     
     geom_jitter(width = 0.2, height = 0.4, alpha = 0.85, color = "#0072B2", size = 2, shape = 17) +
@@ -19,27 +25,29 @@ create_violin_chart <- function(
     
     geom_boxplot(width = 0.25, fill = "#E69F00", color = "black", alpha = 0.65, 
                  outlier.colour = "black", outlier.size = 0.75) +
+    
+    scale_y_discrete(expand = c(0, 0)) +  # Remove vertical padding
     labs(
       title = chart_title,
       x = x_axis_title,
-      y = "",
-      # subtitle = paste("(", earliest_title, "--", latest_title, ")", " n=", 
-      #                  format(nrow(dataset), big.mark = ","), sep = "")
+      y = NULL
     ) +
     
     theme(
       plot.title = element_text(size = 13, hjust = 0.5),
-      plot.subtitle = element_text(size = 7),
       axis.text.x = element_text(face = "bold", size = 9),
-      panel.background = element_rect(fill = "gray98", color = "gray98")
-      )
+      panel.background = element_rect(fill = "gray96", color = "gray96"),
+      plot.margin = margin(t = margin_top, r = margin_right, 
+                           b = margin_bottom, l = margin_left, unit = "pt")
+    )
   
+  
+  # Print the chart (suppress unnecessary warnings)
   suppressMessages(print(violin_chart))
+  
+  # Save the chart to the specified path
   chart_path <- file.path(chart_directory, chart_file_name)
-  chart_width <- 10
-  chart_height <- chart_width / 1.2
   ggsave(chart_path, plot = violin_chart, dpi = 300,
-                          width = chart_width, height = chart_height)
+         width = chart_width, height = chart_height)
 }
-
 #########################################################################

### 2024-12-16 code/timelinecharts.R

@@ -15,9 +15,11 @@ library(ggpmisc)
 library(lubridate)
 library(data.table)
 
+#########################################################################
+rm(list = ls())
+
 programStart <- as.POSIXct(Sys.time())
 formattedStartTime <- format(programStart, "%Y-%m-%d %H:%M:%S")
-cat("\nExecution begins at:", formattedStartTime)
 
 #########################################################################
 
@@ -30,10 +32,6 @@ main_data_file <- "311_Service_Requests_from_2022-2023_AS_OF_09-15-2024.csv"
 
 cat("\n***** Program initialization *****")
 
-# Prior to running program, ensure the R code is located in the working directory, 
-# and that the R functions are located in a sub-directory named "functions".
-# Data files should be placed in a sub-directory named "data".
-
 setwd("C:\\Users\\David\\OneDrive\\Documents\\datacleaningproject\\nyc311clean")
 
 # Create the sub-directories used during program execution.
@@ -44,27 +42,6 @@ working_dir <- getwd()
 # Set the base directory under the working directory
 base_dir <- file.path(working_dir, "code")
 
-# Define the subdirectories
-sub_dirs <- c("charts", "functions", "data", "console_output")
-
-# Check if the base directory exists, and create it if it doesn't
-if (!dir.exists(base_dir)) {
-  dir.create(base_dir)
-  cat("Base directory '", base_dir, "' created.\n", sep = "")
-}
-
-# Loop through the subdirectories and create them if they don't exist
-for (sub_dir in sub_dirs) {
-  dir_path <- file.path(base_dir, sub_dir)
-  if (!dir.exists(dir_path)) {
-    dir.create(dir_path)
-    cat("Subdirectory '", dir_path, "' created.\n", sep = "")
-}
-}
-
-# Define the path to the directory containing your function scripts
-functions_path <- file.path(base_dir, "functions")
-
 # Define the path for the main data file (CSV file)
 data_file <- file.path(base_dir, "data")
 
@@ -74,22 +51,33 @@ chart_directory_path <- file.path(base_dir, "charts")
 # Create the directory for the reduced size file following shrinkage code.
 writeFilePath <- file.path(base_dir, "data")
 
-# Start directing console output to the file
 # Define the console output directory and file name.
 output_dir <- file.path(base_dir, "console_output")
[…]

### 2024-12-24 manuscript/manuscript.tex

@@
 COVID-19 pandemic in managing collective responses.
-
-\jy{The gap in the literature needs to be further highlighted to set 
-  the stage for our contributions.}
-
+However, the literature remains limited in providing specific and
+actionable frameworks for handling large-scale municipal datasets.
+
+By addressing this gap, this study transitions from theory to
+practical insights, aiming to establish principles that enhance both
+the reliability and utility of open government data.
 The contributions of this paper are twofold. First, we delve into
@@
 engagement. Open data not only ensures governmental transparency
-but also empowers researchers, civic developers, and the general
-
-\jy{back up with some references:}
-
-  \citep{minkoff2016nyc, o2017uncharted, kontokosta2021bias}
-public. The data has been pivotal in providing advice on shelters
+but also empowers civic developers, the general public, and
+policy-makers \citep{minkoff2016nyc, o2017uncharted,
+  kontokosta2021bias}.
+The data has been pivotal in providing advice on shelters
 during emergencies, handling inquiries during the COVID-19 pandemic,
@@
 and improving responsiveness across City Agencies.
-
-\jy{need to expand the importance of 311 SR data to the research
-  community: such as
-  resource allocation/response \citep{zha2014profiling, raj2021swift},
-  social equity \citep{white2018promises, kontokosta2021bias},
-  noise \citep{dove2022sounds},
-  street flood \citep{agonafir2022understanding}, etc.}
+It also supports a wide range of studies in resource allocation and
+emergency response strategies \citep{zha2014profiling, raj2021swift},
+social equity in service provision \citep{white2018promises,
+  kontokosta2021bias}, and urban challenges such as noise pollution
+\citep{dove2022sounds} and street flooding
+\citep{agonafir2022understanding}.
 
@@
 \end{figure}
+\jy{Figure 4 and 5: Increase the fonts of the axis ticks. This might
+  be caused by the size of the pdf figure being too big relative to
+  the size shown in the compiled manuscript. So, keep the target size
+  in mind may help, avoiding overstretching. In addition, the x-axis
+  ticks could be made upright.}
 
@@
 
-\vspace{.25cm}
+\begin{center}
 \begin{tikzpicture}[node distance=2cm and 1.5cm, >=stealth]
@@
 \end{tikzpicture}
+\end{center}
 
@@
 and Julia programming languages. Remarkable storage size 
-reduction (within several orders of magnitude) is possible. \citep{bates2024csv}.
+reduction (within several orders of magnitude) is possible \citep{bates2024csv}.
 

### 2025-01-01 code/datacleansing.R
Updates responding to JY's comments - mostly chart formatting
Updated control on charts to allow for dynamic sizing of x-axis labels and tick marks. Required changes to all the charting functions as well as the two core programs (datacleansing.R and timeline.R). Simplified the basic chart program to accommodate both numeric and categorical data.  Moved revised images to manuscript image file.

@@ -1,15 +1,46 @@
 #########################################################################
 # -------------------------------------------------------------
-# 📁 Set Working Directory for the Project
+# 📦 INSTALL AND LOAD REQUIRED PACKAGES
+# -------------------------------------------------------------
+required_packages <- c(
+  "ggplot2", 
+  "scales", 
+  "dplyr", 
+  "zoo", 
+  "ggpmisc", 
+  "lubridate", 
+  "data.table",
+  "renv", 
+  "sf", 
+  
+  "stringdist", 
+  "styler", 
+  "tidyverse", 
+  "rlang", 
+  "httr"
+)
+
+# Check and install missing packages
+missing_packages <- required_packages[!(required_packages %in% installed.packages()[, "Package"])]
+if (length(missing_packages)) {
+  install.packages(missing_packages)
+  cat("📦 Installed missing packages:", paste(missing_packages, collapse = ", "), "\n")
+}
+
+# Load the required packages
+lapply(required_packages, library, character.only = TRUE)
+
+# -------------------------------------------------------------
+# 📁 Set Working Directory for the Project.
 # -------------------------------------------------------------
 # Set working directory to the location of the initialization script
 setwd("C:/Users/David/OneDrive/Documents/datacleaningproject/nyc311clean/code")
 # 
 # # Source the initialization.R script.
-source("run_this_program_first.R")
+#source("run_this_program_first.R")
 
 # Include the working directory for the R programs and sub-directories, passed as a parameter
-setwd("C:\\Users\\David\\OneDrive\\Documents\\datacleaningproject\\download_code")
+#setwd("C:\\Users\\David\\OneDrive\\Documents\\datacleaningproject\\code")
 
 #########################################################################
 rm(list = ls())
@@ -452,7 +483,9 @@ create_combo_chart(
   chart_directory = chart_directory_path,
   console_print_out_title = "Summary of Complaint Type",
   num_x_labels = 10,
-  annotation_size = 4
+  annotation_size = 4,
+  x_axis_tick_size = 10, 
+  x_axis_label_angle = 60
 )
 
 # Restore column names
@@ -875,36 +908,16 @@ if (num_rows_closedBeforeOpened > 1) {
     chart_directory = chart_directory_path
   )
 
-  # Create boxplot of the (negative) duration values
-  negativeDurationChart <- ggplot(
-    
-    data = large_neg_duration, 
[…]

### 2025-01-01 code/functions/adjust_feb_29_to_28.R

@@ -2,34 +2,31 @@
 
 # Define a function to adjust February 29 to February 28 for specified date columns
 adjust_feb_29_to_28 <- function(data, date_cols) {
+  
   for (col in date_cols) {
-    if (col %in% names(data)) {
-      # Ensure the column is in POSIXct format; if not, attempt to convert it
-      if (!inherits(data[[col]], "POSIXct")) {
-        data[[col]] <- as.POSIXct(data[[col]], format = "%Y-%m-%d %H:%M:%S", tz = "UTC")
-      }
-      
-      # Handle invalid conversion results (NA)
-      valid_indices <- !is.na(data[[col]])
-      
-      # Identify rows where the date occurs on February 29th among valid rows
-      feb_29_indices <- which(valid_indices & format(data[[col]], "%m-%d") == "02-29")
+  
+      if (col %in% names(data)) {
+    
+          # Identify rows where the date occurs on February 29
+      feb_29_indices <- which(!is.na(data[[col]]) & format(data[[col]], "%m-%d") == "02-29")
       num_29Feb_rows <- length(feb_29_indices)
-      
+
       if (num_29Feb_rows > 0) {
         cat("\nAdjusted", num_29Feb_rows, "rows in the", col, "column from 29 February to 28 February\n")
-        head(num_29Feb_rows, 50)
+
         # Replace February 29 dates with February 28 of the same year
         data[[col]][feb_29_indices] <- as.POSIXct(
-          paste(format(data[[col]][feb_29_indices], "%Y-02-28"),
-                format(data[[col]][feb_29_indices], "%H:%M:%S")),
-          tz = "America/New_York"
+          paste(
+            format(data[[col]][feb_29_indices], "%Y-02-28"),
+            format(data[[col]][feb_29_indices], "%H:%M:%S")
+          ),
+          tz = "UTC"
         )
       }
     }
   }
+  
   return(data)
 }
-
 
 #########################################################################

### 2025-01-01 code/functions/base_bar_chart.R

@@ -1,26 +1,39 @@
-#########################################################################
-
-base_bar_chart <- function(dataset, x_col, y_col, chart_title, sub_title, 
-                           y_axis_labels = scales::comma, 
-                           add_maximum = FALSE, add_minimum = FALSE, 
-                           add_mean = FALSE, add_median = FALSE, 
-                           add_sd = FALSE, add_second_maximum = FALSE,
-                           add_trendline = FALSE, extra_line = NULL,
-                           horizontal_adjustment_max = 0.5,
-                           vertical_adjustment_max = -1,
-                           console_print_out_title = "Data Summary",
-                           chart_file_name = NULL,
-                           chart_directory = ".",
-                           chart_width = 10,
-                           chart_height = 7,
-                           annotation_size = 5) {
+base_bar_chart <- function(
+    dataset, 
+    x_col, 
+    y_col, 
+    chart_title = "Bar Chart", 
+    sub_title = "",
+    y_axis_labels = scales::comma,
+    add_maximum = FALSE, 
+    add_minimum = FALSE,
+    add_mean = FALSE, 
+    add_median = FALSE, 
+    add_sd = FALSE,
+    add_second_maximum = FALSE,
+    add_trendline = FALSE,
+    extra_line = NULL,
+    horizontal_adjustment_max = 0.5,
+    vertical_adjustment_max = -1,
+    console_print_out_title = "Data Summary",
+    chart_file_name = NULL,
+    chart_directory = ".",
+    chart_width = 10,
+    chart_height = 7,
+    annotation_size = 5,
+    x_axis_tick_size = 12,  # Size of x-axis ticks
+    x_axis_label_angle = 90,  # Angle of x-axis labels
+    x_axis_tick_length = unit(0.3, "cm")  # Length of x-axis tick marks
+) {
+  
+  cat("x_col type:", class(dataset[[x_col]]), "\n")
   
   # Step 1: Print Data Summary
   cat("\n\n", console_print_out_title, " (first 20 rows):\n", sep = "")
   dataset <- as.data.frame(dataset)
   print(head(dataset[, c(x_col, y_col)], 20), row.names = FALSE)
   
-  # Sort dataset by y_col (descending) and x_col based on type
[…]

### 2025-01-01 code/functions/create_bar_chart_categorical.R

@@ -1,49 +1,0 @@
-#########################################################################
-
-create_bar_chart_categorical <- function(
-    dataset, 
-    x_col, 
-    y_col, 
-    chart_title = "Bar Chart", 
-    sub_title = "",
-    y_axis_labels = scales::comma,
-    add_maximum = FALSE, 
-    add_minimum = FALSE,
-    add_mean = FALSE, 
-    add_median = FALSE,
-    add_sd = FALSE, 
-    add_second_maximum = FALSE,
-    add_trendline = FALSE, 
-    extra_line = NULL,
-    horizontal_adjustment_max = 0.5,
-    vertical_adjustment_max = -1,
-    console_print_out_title = "Data Summary",
-    chart_file_name = NULL,
-    chart_directory) 
-{
-  
-  # Call base_bar_chart without x_scale
-  base_bar_chart(
-    dataset, 
-    x_col, 
-    y_col, 
-    chart_title, 
-    sub_title,
-    y_axis_labels, 
-    add_maximum, 
-    add_minimum, 
-    add_mean,
-    add_median, 
-    add_sd, 
-    add_second_maximum, 
-    add_trendline,
-    extra_line, 
-    horizontal_adjustment_max, 
-    vertical_adjustment_max,
-    console_print_out_title, 
-    chart_file_name,
-    chart_directory
-  )
-}
-
-#########################################################################

### 2025-01-01 code/functions/create_bar_chart_numeric.R

@@ -1,24 +1,0 @@
-#########################################################################
-
-create_bar_chart_numeric <- function(dataset, x_col, y_col, chart_title = "Bar Chart", sub_title = "", 
-                                     y_axis_labels = scales::comma, 
-                                     add_maximum = FALSE, add_minimum = FALSE, 
-                                     add_mean = FALSE, add_median = FALSE, 
-                                     add_sd = FALSE, add_second_maximum = FALSE,
-                                     add_trendline = FALSE, extra_line = NULL,
-                                     horizontal_adjustment_max = 0.5,
-                                     vertical_adjustment_max = -1,
-                                     console_print_out_title = "Data Summary",
-                                     chart_file_name = NULL,
-                                     chart_directory)
- {
-  # Call to base_bar_chart without x_scale
-  base_bar_chart(dataset, x_col, y_col, chart_title, sub_title, 
-                 y_axis_labels, add_maximum, add_minimum, add_mean, 
-                 add_median, add_sd, add_second_maximum, add_trendline, 
-                 extra_line, horizontal_adjustment_max, vertical_adjustment_max, 
-                 console_print_out_title, chart_file_name, chart_directory)
-}
-
-
-#########################################################################

### 2025-01-01 code/functions/create_boxplot.R

@@ -1,0 +1,164 @@
+# Create a reusable function for generating boxplots
+create_boxplot <- function(
+    dataset, 
+    x_axis_field,
+    chart_title, 
+    x_axis_title = "", 
+    output_file_name, 
+    chart_directory, 
+    chart_width = 10, 
+    chart_height = 7, 
+    x_axis_tick_size = 15,  # Font size for x-axis labels
+    x_axis_label_angle = 90,  # Angle for x-axis labels
+    plot_title_size = 16,  # Font size for plot title
+    box_width = 0.5,  # Default width for the box
+    jitter_size = 2,  # Size of jitter points
+    jitter_alpha = 0.85,  # Opacity of jitter points
+    tick_length = unit(0.3, "cm")  # Length of tick marks
+) {
+  
+  # Generate the boxplot
+  boxplot_chart <- ggplot(
+    data = dataset, 
+    aes(x = .data[[x_axis_field]], y = factor(1))
+  ) +
+    geom_jitter(
+      color = "#0072B2", 
+      size = jitter_size, 
+      shape = 17, 
+      alpha = jitter_alpha
+    ) +
+    geom_boxplot(
+      outlier.colour = "black", 
+      outlier.shape = 16, 
+      linewidth = 0.7,
+      fill = "#E69F00", 
+      size = 1, 
+      color = "black", 
+      alpha = 0.65,
+      width = box_width
+    ) +
+    theme(
+      legend.position = "none",
+      plot.title = element_text(hjust = 0.5, size = plot_title_size),
+      axis.text.x = element_text(
+        angle = x_axis_label_angle, 
+        vjust = 1, 
+        hjust = 1, 
+        face = "bold", 
+        size = x_axis_tick_size
+      ),
+      axis.text.y = element_blank(),  # Remove y-axis labels
+      axis.ticks.y = element_blank(), # Remove y-axis tick marks
+      axis.ticks.length = tick_length,  # Set tick mark length
+      panel.background = element_rect(fill = "gray96", color = "gray96")
+    ) +
+    labs(
+      title = chart_title,
+      x = x_axis_title,
+      y = NULL
+    )
+  
+  # Print and save the chart
+  print(boxplot_chart)
+  
+  # Save the chart
+  chart_path <- file.path(chart_directory, output_file_name)
+  ggsave(
+    chart_path, 
+    plot = boxplot_chart, 
+    width = chart_width, 
[…]

### 2025-01-01 code/functions/create_violin_chart.R

@@ -11,7 +11,10 @@ create_violin_chart <- function(
     margin_top = 1,     # Default margin values (units in pts)
     margin_right = 2,
     margin_bottom = 1,
-    margin_left = 2
+    margin_left = 2,
+    x_axis_tick_size = 14,       # New parameter for x-axis tick size
+    x_axis_label_size = 14,      # New parameter for x-axis label font size
+    x_axis_tick_length = unit(0.3, "cm")  # New parameter for x-axis tick length
 ) {
   
   # Create the violin chart
@@ -35,12 +38,15 @@ create_violin_chart <- function(
     
     theme(
       plot.title = element_text(size = 13, hjust = 0.5),
-      axis.text.x = element_text(face = "bold", size = 9),
+      axis.text.x = element_text(face = "bold", size = x_axis_tick_size),  # Adjust tick font size
+      axis.title.x = element_text(size = x_axis_label_size, face = "bold"), # Adjust axis label font size
+      axis.text.y = element_blank(),  # Remove y-axis labels
+      axis.ticks.y = element_blank(), # Remove y-axis tick marks
+      axis.ticks.length = x_axis_tick_length,  # Adjust tick mark length
       panel.background = element_rect(fill = "gray96", color = "gray96"),
       plot.margin = margin(t = margin_top, r = margin_right, 
                            b = margin_bottom, l = margin_left, unit = "pt")
     )
-  
   
   # Print the chart (suppress unnecessary warnings)
   suppressMessages(print(violin_chart))
@@ -50,4 +56,4 @@ create_violin_chart <- function(
   ggsave(chart_path, plot = violin_chart, dpi = 300,
          width = chart_width, height = chart_height)
 }
-#########################################################################
+##########################################################################

### 2025-01-01 code/functions/cross_street_analysis.R

@@ -85,26 +85,6 @@ cross_street_analysis <- function(
   
   non_dup_streets_non_blank <- non_dup_streets[non_dup_streets[[cross_street]] != "" &
                                                  non_dup_streets[[intersection_street]] != "", ]
-  
-  # # Define the relative path to go up two directories and then into "data"
-  # relative_path <- file.path("..", "..", "data")
-  # 
-  # # Ensure the directory exists (optional)
-  # if (!dir.exists(relative_path)) {
-  #   dir.create(relative_path, recursive = TRUE)
-  # }
-  # 
-  # # Select only the cross_street and intersection_street columns
-  # subset_data <- non_dup_streets_non_blank[, c(cross_street, intersection_street)]
-  # 
-  # # Construct the filename with dynamic column names and place it in the relative path
-  # filename <- file.path(relative_path, paste0("non_dup_streets_non_blank_", cross_street, "_", intersection_street, ".csv"))
-  # 
-  # # Save the subset dataframe to disk at the specified relative path
-  # write.csv(subset_data, file = filename, row.names = FALSE)
-  # 
-  # cat("Data saved to:", filename, "\n")  # Optional printout for confirmation
-  # 
   num_rows_non_dup_non_blank <- nrow(non_dup_streets_non_blank)
   
   cat("\nSample of matching",cross_street, "&", intersection_street, ":\n")

### 2025-01-01 code/run_this_program_first.R

@@ -5,8 +5,9 @@
 # -------------------------------------------------------------
 # 📂 SET THE BASE WORKING DIRECTORY
 # -------------------------------------------------------------
-# Customize this path as needed
-working_dir <- "C:/Users/David/OneDrive/Documents/datacleaningproject/download_code"
+# Customize this path as needed/desired
+working_dir <- "C:/Users/David/OneDrive/Documents/datacleaningproject/initialization_test_directory"
+# -------------------------------------------------------------
 
 # If the working directory does not exist, create it
 if (!dir.exists(working_dir)) {
@@ -17,7 +18,6 @@ if (!dir.exists(working_dir)) {
 setwd(working_dir)
 base_dir <- working_dir  # Store as 'base_dir' for use in subdirectories
 cat("📍 Working directory set to:", getwd(), "\n")
-
 
 # -------------------------------------------------------------
 # 📦 INSTALL AND LOAD REQUIRED PACKAGES
@@ -49,7 +49,6 @@ if (length(missing_packages)) {
 # Load the required packages
 lapply(required_packages, library, character.only = TRUE)
 
-
 # -------------------------------------------------------------
 # 📁 CREATE SUBDIRECTORIES UNDER THE WORKING DIRECTORY
 # -------------------------------------------------------------
@@ -63,7 +62,6 @@ for (sub_dir in sub_dirs) {
 }
 
 cat("\n✅ Project directory structure complete.")
-
 
 # -------------------------------------------------------------
 # 📦 DOWNLOAD AND EXTRACT DATASETS FROM FIGSHARE
@@ -133,7 +131,6 @@ for (i in seq_along(github_urls)) {
   }
 }
 
-
 # -------------------------------------------------------------
 # 📂 DOWNLOAD AND SOURCE ALL R FUNCTION FILES FROM GITHUB
 # -------------------------------------------------------------
@@ -191,7 +188,6 @@ if (length(function_files) > 0) {
 } else {
   cat("\n❌ No function files found to source in:", functions_dir, "\n")
 }
-
 
 # -------------------------------------------------------------
 # 🏁 FINAL MESSAGE TO THE USER

### 2025-01-01 code/timelinecharts.R

@@ -1,14 +1,44 @@
 # -------------------------------------------------------------
+# 📦 INSTALL AND LOAD REQUIRED PACKAGES
+# -------------------------------------------------------------
+required_packages <- c(
+  "ggplot2", 
+  "scales", 
+  "dplyr", 
+  "zoo", 
+  "ggpmisc", 
+  "lubridate", 
+  "data.table",
+  "renv", 
+  "sf", 
+  "stringdist", 
+  "styler", 
+  "tidyverse", 
+  "rlang", 
+  "httr"
+)
+
+# Check and install missing packages
+missing_packages <- required_packages[!(required_packages %in% installed.packages()[, "Package"])]
+if (length(missing_packages)) {
+  install.packages(missing_packages)
+  cat("📦 Installed missing packages:", paste(missing_packages, collapse = ", "), "\n")
+}
+
+# Load the required packages
+lapply(required_packages, library, character.only = TRUE)
+
+# -------------------------------------------------------------
 # 📁 Set Working Directory for the Project
 # -------------------------------------------------------------
 # Set working directory to the location of the initialization script
 setwd("C:/Users/David/OneDrive/Documents/datacleaningproject/nyc311clean/code")
 # 
 # # Source the initialization.R script.
-source("run_this_program_first.R")
+#source("run_this_program_first.R")
 
 # Include the working directory for the R programs and sub-directories, passed as a parameter
-setwd("C:\\Users\\David\\OneDrive\\Documents\\datacleaningproject\\download_code")
+#setwd("C:\\Users\\David\\OneDrive\\Documents\\datacleaningproject\\download_code")
 
 #########################################################################
 rm(list = ls())
@@ -332,27 +362,27 @@ if (nrow(yearly_df) > 2) { # skip if <2 years. Not enough data to be meaningful.
     label = paste0(year_digits, "-yr growth: ", percentage_growth, "%", sep = ""),
     size = 3.7, color = "#661100", vjust = -1, hjust = 0.1
   )
-
-  SR_yearly <- create_bar_chart_numeric(
+  
+  SR_yearly <- base_bar_chart(
     dataset = yearly_df,
-    x_col = "Year",
-    y_col = "count",
-    chart_title = NULL,
-    sub_title = chart_sub_title,
-    console_print_out_title = "Yearly SR count",
-    add_mean = TRUE,
-    add_median = FALSE,
-    add_sd = FALSE,
-    add_trendline = TRUE,
-    add_maximum = FALSE,
-    add_minimum = FALSE,
[…]

### 2025-01-08 manuscript/manuscript.tex
Updates responding to JY's comments - mostly chart formatting
Updated control on charts to allow for dynamic sizing of x-axis labels and tick marks. Required changes to all the charting functions as well as the two core programs (datacleansing.R and timeline.R). Simplified the basic chart program to accommodate both numeric and categorical data.  Moved revised images to manuscript image file.
Simple edits
Included note on  the attempt to standardize street names, which failed to increase "matching" of cross_street & intersection_street by any meaningful amount.

@@
 include the addition of online and mobile channels in 2009, a
-record monhtly high of 348,463 SRs in 
+record monthly high of 348,463 SRs in 
 August 2020 (associated with the the COVID-19 
@@
 
-Our investigation utilizes a 311 SR dataset covering  2-years (2022-2023).
-This sample was collected on 15 September 2024, and is used 
-to conduct detailed analysis in areas such as data 
-consistency, data validity, redundant fields, etc.. Instructions 
-on downloading this file from the NYC Open Data Portal 
-are available in the supplemental material. Here are some 
-characteristics of the 311 SR dataset used in this analysis:
+Our investigation uses a 311 SR dataset covering  2-years (2022-2023), 
+and was collected on 15 September 2024. It is to conduct 
+analysis in areas such as data consistency, data validity, 
+redundant fields, etc.. Instructions on downloading this file are 
+available in the supplemental material. Characteristics of the 311 SR dataset include:
 
 \begin{itemize}[left=1.5em]
-\item The dataset is approximately 3.7 GB in size and contains 
+\item The dataset is 3.7 GB in size and contains 
   over 6.4 million rows. It was exported as a CSV file.
@@
 Collectively these six City Agencies handle over 90\% of the total SR 
-volume, indicative of the critical role these Agencies play in managing
-public concerns, ranging from noise complaints and housing issues to
-sanitation and transportation problems. The remaining
-10\% of SRs is distributed across 10 additional Agencies. 
-This concentration of SRs to the ``big six'' Agencies 
-underscores the necessity for optimal data handling and 
-curation processes within these  high volume Agencies.
+volume, indicative of the critical role they play in managing
+public concerns; ranging from noise complaints and housing issues to
+sanitation and transportation issues. The remaining
+10\% of SRs are distributed across 10 additional Agencies. 
+This concentration of SRs to these ``big six'' Agencies 
+underscores the necessity for optimal data curation processes 
+within these  high volume Agencies.
 
@@
 Structural issues refers to how data is organized, formatted, 
-or presented within a dataset. They can make 
-it difficult to analyze the data effectively and often require extensive 
-data cleaning efforts. Here are three structural issues we encountered.
+or presented within a dataset. Structural issues can cause difficulty in 
+ analyzing the data effectively, often necessitating extensive 
+data cleaning. We encountered three structural issues.
 
@@
 \label{sec:blanks}
-Understanding the absence of data by field is important 
-when undertaking analysis. For example, if you wanted to 
-determine if SRs were closed before or after their
-\texttt{due\_date}, you would be challenged as 99.6\% of the
-\texttt{due\_date} fields are blank. When sorting fields by
-blank (or N/A values) they appear to divide into three groups:
+Understanding the absence of data by field greatly aids when 
+conducting analysis. For example, determining if SRs were closed 
+before or after their \texttt{due\_date}, you would be challenged as 99.6\% of the
+\texttt{due\_date} fields are blank, and are only used by one Agency, DSNY. Blank 
+(or N/A values) they appear to divide into three groups:
 
@@
 For some analysis efforts, it may be prudent to inquire 
-further as to why some fields are mostly or almost 
-always blank. Figure~\ref{fig:blank_fields} is 
-a graphic depiction of total empty (blank \& N/As) for each 
-field illustrating the grouping into the Mostly Empty, 
-Partially Empty, and Few/None Empty categories.
+ as to why some fields are mostly/almost always blank. 
+Figure~\ref{fig:blank_fields} is a graphic depiction of the \%
+empty (blank \& N/As) for each field. It illustrates the grouping into 
+Mostly Empty, Partially Empty, and Few/None Empty segments.
 
@@
 
-However,  some fields proved problematic as to their 
-compliance with a domain of allowable values. These include:
+However, some fields lacked compliance
+to a domain of allowable values, including:
 
@@
 SRs where the \texttt{created\_date} and \texttt{closed\_date} 
-indicate ``created'' or ``closed'' exactly at midnight (00:00:00), 
+indicate action exactly at midnight (00:00:00), 
 to the second. Normally, SR creation and closure follows 
@@
 \end{figure}
-\jy{Figure 4 and 5: Increase the fonts of the axis ticks. This might
-  be caused by the size of the pdf figure being too big relative to
-  the size shown in the compiled manuscript. So, keep the target size
-  in mind may help, avoiding overstretching. In addition, the x-axis
-  ticks could be made upright.}
-
-These unusual patterns of SR creation and closure exactly at midnight 
+
+These unusual patterns of SR creation/closure at midnight 
 suggest the presence of a bulk create/close software process, perhaps one
 that automatically assigns time stamps of midnight (00:00:00) to 
-large batches of SRs prior to sending the data to the 311 
-data warehouse. This behavior distorts the calculated duration of these 
+ batches of SRs prior to sending the data to the 311 
+database. This behavior distorts the calculated duration of these 
 SRs as well as overall system usage by providing inaccurate 
@@
 \label{sec:precision}
-An question of precision vs. accuracy arises with the \texttt{latitude} 
+The question of precision vs. accuracy arises with the \texttt{latitude} 
 and \texttt{longitude} fields. as well as with the duplicative 
@@
 Given that 1 degree of latitude at the equator is equal to 111.04 
-kilometers, the ``1'' at the end of the decimal number represents 
+kilometers, the ``1'' at the end of that decimal number represents 
 approximately 1.1 nanometers (1/1,000,000,000 of a meter). (For 
-reference a DNA molecule is approximately 2 nanometers in width.) The 
-representation of the \texttt{latitude} and \texttt{longitude} fields
-appears to be a classic case of extreme precision, but limited 
+reference a DNA molecule is approximately 2 nanometers in width.) This 
+representation appears to be a classic case of extreme precision, albeit with limited 
 accuracy. All the \texttt{latitude}, \texttt{longitude}, and \texttt{location}
@@
 \label{sec:duplicates}
-During this analysis, several redundant fields were identified and should 
-be examined for possible consolidation.
+This analysis revealed several redundant fields, indicting a potential need 
+for consolidation.
 
@@
 
+\begin{quote}
+\textit{Note: The authors attempted extensive standardization of street names  
+in an attempt to match what were obvious identical streets, deemed unequal 
+due only to spelling and naming conventions. Those efforts were intended to match, for example, 
+"KANE PL" and "KANE PLACE" as well as "FLATLANDS 7 ST" and "FLATLANDS 7 STREET". However, these 
+efforts were  abandoned after they increased street name matches by less than 1\%.)}
+\end{quote}
+
 \subsection{Reducing File Size}
 \label{sec:filesize}
-By removing duplicate and ``near-duplicate'' fields, it is possible to 
-shrink the file size by 39.5\%, which for this large dataset equates to 
-a reduction of 1.4GB, a significant amount. A smaller file size 
-means faster downloads, less storage impact, and often 
-simpler data analysis efforts. Below is a list of duplicate 
-and near-duplicate fields which we believe could be
+By removing duplicate and near-duplicate fields, one can reduce file size by 39.5\%, 
+equating to a reduction of 1.4GB, a significant amount. A smaller file size 
+means faster downloads, reduced storage impact, and often simpler data analysis. 
+Below is a list of duplicate and near-duplicate fields which we believe could be
 removed from the dataset with minimal data loss.
@@
     \item \texttt{park\_borough}: This field is a 100\% match with 
-    the \texttt{borough} field and can be deleted without data loss.
+    the \texttt{borough} field. It can be deleted without data loss.
 
@@
     \item \texttt{cross\_street\_1/2} and \texttt{intersection\_street\_1/2}: 
-    These two street pairs both exhibit an 88\% match. While we cannot 
+    These two street pairs exhibit an 88\% match. Although we cannot 
     confirm the accuracy of either of these street pairs, we recommend deleting the 
@@
 
-\paragraph{Field Encoding:} Further reduction is possible 
+\paragraph{Field Encoding:} Further size reduction is possible 
 by encoding selected categorical variables 
@@
 ``NOISE - STREET/SIDEWALK'' \texttt{complaint\_type} occurs 
-over 300,000 times. Such encoding not only reduces file 
+over 300,000 times. Encoding not only reduces file 
 size, but would likely improve processing efficiency by using numeric
-values instead of text strings. Moreover, standardizing 
-categorical inputs prevents the inclusion of misspellings 
-or unrecognized/invalid entries, ensuring cleaner, more reliable 
+values instead of text strings. Moreover, standardized 
+categorical inputs often prevents misspellings 
+and invalid entries, ensuring cleaner, more reliable 
 data. This approach is particularly useful in large datasets, where 
-frequently repeated text-based categorical variables can 
-significantly increase storage requirements.
-
-One promising alternative to reducing data storage requirements 
+frequently repeated text-based categorical variables significantly 
+increases storage requirements.
+
+One promising alternative to reducing data storage  
 for this large 311 SR dataset (3.7GB) would be employment of 
@@
 found several discrepancies between the Data Dictionary and the actual
-data, which could potentially mislead analytic efforts. Here are some
-examples:
+data. Such errors could potentially mislead analytic efforts. Examples include:
 
@@
 \label{sec:recommendations}
-Building upon the insights gained from our case study of the 
+Building upon  insights from our case study of the 
 NYC 311 SR data, we propose a set of data 
-curation principles tailored for government released open datasets. 
+curation principles tailored for government open datasets. 
 These principles are designed to address the unique challenges 
-and requirements observed in the curation of such datasets, 
-ensuring that they remain reliable, consistent, and useful for 
+ observed in the curation of such datasets, 
+ensuring they remain reliable, consistent, and useful for 
 public services, research, and other applications.
@@
 \subsection{Principles}
-\paragraph{Principle 1: Establishing Cross-Orginazational Consistency}
+\paragraph{Principle 1: Establishing Cross-Organizational Consistency}
 A key challenge in government released datasets is the 
-harmonization of data fields across various Agencies and 
-maintaining this consistency over time. To address this, agencies should adopt 
+harmonization of data fields across various Agencies, and 
+maintaining that consistency over time. To address this, Agencies should adopt 
 standardized naming conventions, value domains, usages, and 
@@
 processes should be implemented to ensure consistency 
-across long-term datasets, particularly as agency structures 
-or reporting standards evolve. For instance, ensuring that 
+across long-term datasets, particularly as Agency structures 
+and reporting standards evolve. For instance, ensuring that 
 historical datasets reflect the same field naming conventions 
@@
 \begin{itemize}[left=1.5em]
-    \item Certain date fields should be promulgated by the software
-    system versus manual entry. Critical dates should be populated
+    \item Certain date fields should be promulgated by the application software
+     versus manual entry. Dates should be populated
     based upon an action in the system, such as a change in status.
@@
     \item Select fields should be subjected to automated checks 
-    that prevent logical errors, such as closing and opening.
+    that prevent logical errors, such as a closing date occurring before the creation date.
 
@@
 Clear and accurate data documentation is essential for the effective use 
-of government released datasets. Our study found several 
+of government datasets. Our study found several 
 instances where the Data Dictionary did not accurately reflect the 
@@
 content. This practice enhances the ability of external users, 
-including researchers and developers, to understand and work 
+including researchers and developers, to work 
 effectively with the data.
@@
 errors in real-time, and ensuring consistency across multiple data sources. 
-Feedback from statisticians and data scientists, such as the issues reported 
+Feedback, such as the issues reported 
 in this paper, benefit from a clear and accessible channel to the 311 SR data 
@@
 
-A particularly valuable avenue for engagement is through events such as NYC 
-Open Data Week, which fosters collaboration and innovation across sectors. 
+An especially valuable engagement avenue is through events such as NYC 
+Open Data Week, which foster collaboration and innovation across sectors. 
 Our involvement in this project stemmed from such an event, where we were 
 inspired by the potential of open data to enhance public services. Governments 
-should continue supporting these types of initiatives, which not only promote 
+should continue these types of initiatives, which not only promote 
 data literacy, but encourage community-driven improvements to datasets. 
 By leveraging the expertise of data scientists, statisticians, stakeholders, 
-and civic organizations, governments can ensure that their open datasets 
+and civic organizations, governments can ensure their open datasets 
 remain accurate, transparent, and useful for a wide range of applications.

### 2025-01-10 manuscript/manuscript.tex
Minor edits for readability. Add "field usage by Agency" table

@@
 This paper examines these challenges through a
-case study of the NYC 311 Service Request dataset, focusing on issues 
-of data validity, consistency, and curation efficiency. Based on 
+case study of the NYC 311 Service Request dataset, focusing on  
+data validity, consistency, and curation efficiency. Based on 
 insights from this case study, we propose a set of data curation 
@@
 tools for querying, grouping, aggregating,
-geo-mapping, visualizing, and exporting results. The
-NYC Office of Technology and Innovation (OTI) provides technical support for 
-the 311 application, the open data infrastructure, mobile \& web applications, 
+visualizing, and exporting results. The NYC Office of Technology and Innovation (OTI) 
+provides technical support for the 311 application, 
+the open data infrastructure, mobile \& web applications, 
 and the public-facing data portal. Input data for the 311 system 
@@
 \item The dataset is 3.7 GB in size and contains 
-  over 6.4 million rows. It was exported as a CSV file.
+  over 6.4 million rows. It exports as a CSV file.
 
@@
 
-While one can likely infer the meaning of these ``computed'' 
+While one can likely infer the meaning of these \textit{computed} 
 fields, uncertainty remains as to their derivation and 
@@
 to the Data Dictionary would seem appropriate. Accordingly, 
-we will not address these fields in our analysis.
+we omit these fields in our analysis.
 
@@
 311 SR dataset exhibited inconsistent approaches when 
-presenting missing data. We discovered many different 
+presenting missing data. We discovered different 
 representations of missing data: nulls, spaces, ``NA'', ``N/A'', 
 and ``<NA>''. This variety of representations complicates programming
-efforts, and increases the likelihood of introducing errors 
-in data analysis (typically by search for or excluding missing fields 
-and not aware of the various representations of such). A more consistent 
+efforts and increases the likelihood of introducing errors 
+in data analysis (typically when selecting data for blank, NA, or missing fields). A more consistent 
 approach to treating missing data should be employed to address this issue.
@@
 \label{sec:blanks}
-Understanding the absence of data by field greatly aids when 
-conducting analysis. For example, determining if SRs were closed 
-before or after their \texttt{due\_date}, you would be challenged as 99.6\% of the
-\texttt{due\_date} fields are blank, and are only used by one Agency, DSNY. Blank 
-(or N/A values) they appear to divide into three groups:
+Understanding the absence of data by field greatly aids analysis. 
+For example, determining if SRs were closed 
+before their \texttt{due\_date}, would be challenging as 99.6\% of the
+\texttt{due\_date} entries are blank as only DSNY appears to use this field.
+Blank (or N/A values) divide into three groups:
 
@@
 
+\vspace{\baselineskip}
+\noindent
 The Mostly Empty category includes:
-\texttt{taxi\_company\_borough}, \texttt{due\_date}
-\texttt{road\_ramp}, and \texttt{bridge\_highway\_name}. The 
-Partially Empty category includes:
-\texttt{location\_type}, \texttt{landmark}, 
-and \texttt{cross\_street\_1 \& \_2}. And the Few/None Empty 
-category includes \texttt{created\_date}, \texttt{complaint\_type},
-\texttt{agency}, and \texttt{status}. Of particular note is that 
-several of the Mostly Empty fields are used by only one or 
+\texttt{taxi\_company\_borough}, 
+\texttt{bridge\_highway\_name},
+\texttt{due\_date}, and \texttt{road\_ramp}. 
+The Partially Empty category includes:
+\texttt{cross\_street\_1 \& \_2},
+\texttt{location\_type}, and texttt{landmark}. And 
+the Few/None Empty category includes \texttt{created\_date}, 
+\texttt{complaint\_type},
+\texttt{agency}, 
+and \texttt{status}. 
+Of particular note is that several of the Mostly Empty fields are used by only one or 
 two Agencies (See: \autoref{tab:field-usage-summary}).
 
-\begin{table}[htbp]
+\begin{table}[tbp]
     \centering
@@
     \label{tab:field-usage-summary}
+    \small
     \begin{tabular}{l r r p{4cm}}
@@
 
+% Force a blank line and remove indentation
+\vspace{\baselineskip}
+\noindent
 For some analysis efforts, it may be prudent to inquire 
@@
 \end{figure}
+
+\vspace{\baselineskip}
+\noindent
+The accomanying R program also produces a CSV file showing Field usage by Agency. 
+Table~\ref{tab:field-usage-by-agency} contains a sample of Field usage by select Agencies.
+
+\begin{table}[tbp]
+\centering
+\caption{Sample Field Usage by select Agencies}
+\label{tab:field-usage-by-agency}
+\small
+\begin{tabular}{lrrrrrrr}
+\toprule
+\textbf{\textbf{Field}} & \textbf{DEP} & \textbf{DHS} & \textbf{DOB} & \textbf{DOT} & \textbf{DPR} & \textbf{DSNY} & \textbf{TLC} \\
+\midrule
+\texttt{landmark}                  & 2935     & 70185    & 282      & 135150   & 254218   & 424284   & 63133  \\
+\texttt{facility\_type}            & 175549   & 0        & 51575    & 159667   & 0        & 66012    & 0      \\
+\texttt{due\_date}                 & 0        & 0        & 0        & 0        & 0        & 32097    & 0      \\
+\texttt{community\_board}          & 361715   & 99613    & 203715   & 401143   & 266201   & 608188   & 73601  \\
+\texttt{vehicle\_type}             & 0        & 0        & 0        & 0        & 0        & 0        & 1696   \\
+\texttt{taxi\_company\_borough}    & 0        & 0        & 0        & 0        & 0        & 0        & 4025   \\
+\texttt{taxi\_pick\_up\_location}  & 0        & 0        & 0        & 0        & 0        & 0        & 72558  \\
+\texttt{bridge\_highway\_name}     & 15       & 21012    & 0        & 7451     & 0        & 613      & 977    \\
+\texttt{bridge\_highway\_direction} & 15      & 8973     & 0        & 7450     & 0        & 614      & 1006   \\
+\texttt{road\_ramp}                & 15       & 2588     & 0        & 7450     & 0        & 613      & 999    \\
+\texttt{bridge\_highway\_segment}  & 15       & 21013    & 0        & 7450     & 0        & 613      & 987    \\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
+\vspace{\baselineskip}
+\noindent
 However, some fields lacked compliance
@@
 \begin{itemize}[left=1.5em]
-    \item 12,450 SRs with a \texttt{closed\_date} that occurs before the 
-    \texttt{created\_date} resulting a nonsensical ``negative duration''. 
+    \item 12,450 SRs with a \texttt{closed\_date} before \texttt{created\_date} creating a nonsensical ``negative duration''. 
     \item Eight SRs with \texttt{created\_date(s)} or \texttt{closed\_date(s)} in 
-    the distant past, specifically ``1900-01-01''. Despite the small number of 
-    these entries, such extreme values can greatly distort analytical conclusions.
+    the distant past, specifically  \textit{1900-01-01}. Despite few of 
+    these entries, such extreme values can greatly distort analytical efforts. 
     \item 163,720 SRs with identical \texttt{created\_date(s)} and \texttt{closed\_date(s)} 
@@
     % Main caption
-    \caption{Comparison of 311 SRs created \& closed by hour-of-the-day during 2022--2023.}
+    \caption{311 SRs Created \& Closed exactly on the hour (hh:00:00) for 2022--2023.}
     \label{fig:exacthours}
@@
 \begin{quote}
-\textit{Note: The authors attempted extensive standardization of street names  
+\textit{Note: The authors attempted standardization of street names  
 in an attempt to match what were obvious identical streets, deemed unequal 
 due only to spelling and naming conventions. Those efforts were intended to match, for example, 
-"KANE PL" and "KANE PLACE" as well as "FLATLANDS 7 ST" and "FLATLANDS 7 STREET". However, these 
-efforts were  abandoned after they increased street name matches by less than 1\%.)}
+\textit{KANE PL} and \textit{KANE PLACE} as well as \textit{AVE P} and \textit{AVENUE P}. However, these 
+efforts were  abandoned after they resulted in an increase in street name matches of less than 1\%.)}
 \end{quote}
@@
     These two street pairs exhibit an 88\% match. Although we cannot 
-    confirm the accuracy of either of these street pairs, we recommend deleting the 
-    \texttt{intersection\_street\_1/2} fields while acknowledging 
+    confirm the accuracy of either of these pairs, we recommend deleting the 
+    \texttt{intersection\_street\_1/2} fields, acknowledging 
     the concomitant data loss.
@@
 
+\vspace{\baselineskip}
+\noindent
 Additionally, certain fields may not be useful for broad analysis
-due to their sparse data population. This especially applies to the fields below, 
-which are predominantly used by the TLC and are 99\% (or greater) blank.
-
+due to their sparse data population. This caveat applies to the fields 
+below which are 99\% (or greater) blank.
+
+\vspace{\baselineskip}
+\noindent
 \begin{quote}
-\textit{[We should note that even a field that is 99\% 
+\textit{We should note that even a field that is 99\% 
 blank will still contain 64,000 non-blank rows in this 
-large (6.4 million row) dataset.}; non-trivial.]
+large (6.4 million row) dataset.; certainly a non-trivia amount of data.}
 \end{quote}
@@
 
-The 311 SR Data Dictionary would greatly benefit from an update. We
-found several discrepancies between the Data Dictionary and the actual
-data. Such errors could potentially mislead analytic efforts. Examples include:
+The 311 SR Data Dictionary needs updating. We found several 
+discrepancies between the Data Dictionary and the actual
+data. Such errors could potentially mislead analytic efforts. Examples::
 
@@
 \section*{Supplementary Material}
-The NYC 311 SR data and code used to produce the results reported in
-this paper are available at \url{https://github.com/jun-yan/nyc311clean}.
+The NYC 311 R code used to produce these results, along 
+with additional analytical findings, are available 
+at \url{https://github.com/jun-yan/nyc311clean}. Reference 
+data files (311 SR data, USPS zip codes) are available 
+at \url{https://doi.org/10.6084/m9.figshare.c.7617311}.
 

### 2025-01-12 manuscript/manuscript.tex
touch up

@@
 
-\item In addtion to \texttt{incident\_address}, each SR has six
+\item In addtion to \texttt{incident\_address}, each SR has eight
 	additional geographic fields:
@@
 Blank (or N/A values) divide into three groups:
-
 \begin{itemize}[left=1.5em]
-    \item Mostly Empty: 93-99.9\% blank 
-    \item Partially Empty: 40-4\% blank
-    \item Few/None Empty: 2-0\% blank
+    \item Mostly Empty: 93-99.9\% blank;
+    \item Partially Empty: 40-4\% blank;
+    \item Few/None Empty: 2-0\% blank.
 \end{itemize}
 
-\vspace{\baselineskip}
-\noindent
 The Mostly Empty category includes:
@@
 
-% Force a blank line and remove indentation
-\vspace{\baselineskip}
-\noindent
-For some analysis efforts, it may be prudent to inquire 
- as to why some fields are mostly/almost always blank. 
-Figure~\ref{fig:blank_fields} is a graphic depiction of the \%
-empty (blank \& N/As) for each field. It illustrates the grouping into 
-Mostly Empty, Partially Empty, and Few/None Empty segments.
-
 \begin{figure}[tbp]
@@
 
-\vspace{\baselineskip}
-\noindent
-The accomanying R program also produces a CSV file showing Field usage by Agency. 
-Table~\ref{tab:field-usage-by-agency} contains a sample of Field usage by select Agencies.
+For some analysis efforts, it may be prudent to inquire 
+as to why some fields are mostly/almost always blank. 
+Figure~\ref{fig:blank_fields} is a graphic depiction of the \%
+empty (blank \& N/As) for each field. It illustrates the grouping into 
+Mostly Empty, Partially Empty, and Few/None Empty segments.
+
+Table~\ref{tab:field-usage-by-agency} highlights the varied usage of
+certain fields across NYC agencies, revealing both specialization and
+redundancy. For instance, \texttt{landmark} and
+\texttt{facility\_type} are widely used, indicating their importance
+in classifying SRs. In contrast, fields like
+\texttt{taxi\_company\_borough} and \texttt{vehicle\_type} are used
+exclusively by TLC, suggesting that such fields, which serve a narrow
+purpose, could be better stored in separate, agency-specific tables.
+This approach would reduce the size and complexity of the primary
+dataset, making it easier to maintain and analyze while preserving
+the ability to perform specialized analyses when necessary.
+Streamlining through harmonization and validation would further
+enhance consistency and optimize cross-agency analyses.
 
@@
 \centering
-\caption{Sample Field Usage by select Agencies}
+\caption{Sample field usage by select agencies among the 6.4 million
+  SRs in 2022--2023. }
 \label{tab:field-usage-by-agency}
@@
 usage and as evidenced in other historical datasets:
-
 \begin{itemize}[left=1.5em]
@@
 \end{itemize}
-
-\vspace{\baselineskip}
-\noindent
 However, some fields lacked compliance
-to a domain of allowable values, including:
+to a domain of allowable values, as outlined next.
 
@@
 an automated process would normally exclude, including:
-
 \begin{itemize}[left=1.5em]
@@
 311 SR Data Dictionary, the \texttt{landmark} 
-field \textit{can refer to any noteworthy location, 
-including but not limited to, parks, hospitals, airports, sports facilities, 
-performance spaces, etc..} However, most entries are actually 
-street names, with a 62\% match to \texttt{street\_name} field. 
-Even many of the non-matches (excluding blanks) appear to be 
-a logical match, differing only in spelling or formatting 
+field is intended to refer to any noteworthy location, 
+including but not limited to parks, hospitals, airports, sports facilities, 
+and performance spaces. However, most entries are actually 
+street names, with a 62\% match to the \texttt{street\_name} field. 
+Many of the non-matches (excluding blanks) also appear to be 
+logical matches, differing only in spelling or formatting 
 (e.g., ``NINTH AVE'' vs. ``9 AVE''). This indicates a possible 
-misuse of the \texttt{landmark} field as well as a high level of redundancy.
+misuse of the \texttt{landmark} field and highlights significant
+redundancy.
 
@@
 
-\begin{quote}
-\textit{Note: The authors attempted standardization of street names  
-in an attempt to match what were obvious identical streets, deemed unequal 
-due only to spelling and naming conventions. Those efforts were intended to match, for example, 
-\textit{KANE PL} and \textit{KANE PLACE} as well as \textit{AVE P} and \textit{AVENUE P}. However, these 
-efforts were  abandoned after they resulted in an increase in street name matches of less than 1\%.)}
-\end{quote}
+We attempted to standardize street names to reconcile entries that
+were clearly identical but differed due to spelling or naming
+conventions. For example, we sought to match entries such as
+\textit{KANE PL} and \textit{KANE PLACE} or \textit{AVE P} and
+\textit{AVENUE P}. However, this standardization effort was
+ultimately abandoned, as it resulted in less than a 1\% increase in
+matched street names.
 
@@
 
-\vspace{\baselineskip}
-\noindent
 Additionally, certain fields may not be useful for broad analysis
-due to their sparse data population. This caveat applies to the fields 
+due to their sparse data population. This applies to the fields 
 below which are 99\% (or greater) blank. 
-
-\vspace{\baselineskip}
-\noindent
-\begin{quote}
-\textit{We should note that even a field that is 99\% 
-blank will still contain 64,000 non-blank rows in this 
-large (6.4 million row) dataset.; certainly a non-trivia amount of data.}
-\end{quote}
-
 \begin{itemize}[left=1.5em]
@@
 \end{itemize}
+While such sparsity may limit their use in 
+broad analyses, these fields can still contain valuable insights 
+for specific agency-level studies, given that even a 1\% 
+non-blank rate equates to approximately 64,000 rows in this 
+6.4 million-row dataset. Proper handling of sparse fields, 
+such as storing them in separate tables or using them in 
+targeted analyses, could improve data accessibility and 
+efficiency without compromising the dataset’s broader utility.
 

### 2025-01-12 manuscript/manuscript.tex
touched abstract and intro

@@
 serving as a crucial resource for research across various domains, 
-including health, urban development, and transportation. While the
-open data movement has made vase amounts of data accessible, the
-challenge of ensuring data quality and usability through effective
-curation remain insufficiently addressed in the literature..
-This paper examines these challenges through a
-case study of the NYC 311 Service Request dataset, focusing on  
-data validity, consistency, and curation efficiency. Based on 
-insights from this case study, we propose a set of data curation 
-principles tailored for government-released open data aiming
-to enhance data management practices and ensure 
-the ongoing utility of open data.
+including health, urban development, and transportation. However, the
+effective use of open data relies heavily on data quality and
+usability, challenges that remain insufficiently addressed in the
+literature. This paper examines these challenges through a
+case study of the NYC 311 Service Request dataset,  identifying key
+issues in data validity, consistency, and curation efficiency. We
+propose a set of data curation principles, tailored for
+government-released open data, to address these challenges.
+Our findings highlight the importance of harmonized field definitions,
+streamlined storage, and automated quality checks, offering practical
+guidelines for improving the reliability and utility of open datasets.
 \end{abstract}
@@
   \kwd{NYC Open Data}
-  \kwd{Smart Cities}
+  \kwd{Smart city}
   \kwd{Transparency}
@@
 \citep{barns2016mine, wang2016adoption}.
+Despite these global efforts, the effective use of open data hinges on
+addressing critical challenges in data curation, such as ensuring
+quality, consistency, and usability.
 
@@
 today hosts an impressive array of 2,700 datasets
-across 80 different city agencies. This resource has become invaluable
-for researchers across various fields and has significantly enhanced
-local government transparency. Popular datasets include information on
+across 80 different city agencies. This wealth of data serves as a
+powerful tool for researchers and policymakers, significantly
+enhancing local government transparency.
+Popular datasets include information on
 restaurant health inspection violations, car crashes, high school and
@@
 understanding and addressing of complex urban challenges.
-
+For example, NYC’s 311 data has been used to optimize resource
+allocation in urban planning and improve emergency response times,
+showcasing its practical relevance in real-time decision-making.
+
+However, the open data presents substantial curation challenges.
 Data curation, the process of organizing, maintaining, and ensuring
@@
 COVID-19 pandemic in managing collective responses.
-However, the literature remains limited in providing specific and
-actionable frameworks for handling large-scale municipal datasets.
+Despite these advances, there remains a notable gap in providing
+actionable frameworks for managing large-scale municipal datasets.
 

### 2025-01-13 manuscript/manuscript.tex
minor edits to section 2
Minor edits. Grammar.

@@
 the forefront of this movement since the enactment of the Open 
-Data Law in 2012, leading to the creation of the NYC Open Data
-portal. This portal currently hosts 2,700 datasets,
+Data Law in 2012, creating the NYC Open Data
+portal. The portal currently hosts 2,700 datasets,
 serving as a crucial resource for research across various domains, 
@@
 usability, challenges that remain insufficiently addressed in the
-literature. This paper examines these challenges through a
+literature. This paper examines these challenges via a
 case study of the NYC 311 Service Request dataset,  identifying key
@@
 redundant fields, etc.. Instructions on downloading this file are 
-available in the supplemental material. Characteristics of the 311 SR dataset include:
+available in the supplemental material. Characteristics of this
+two-year 311 SR dataset are as follows.
 
@@
 \end{itemize}
-\medskip
 
@@
 \texttt{taxi\_company\_borough}, 
-\texttt{bridge\_highway\_name},
-\texttt{due\_date}, and \texttt{road\_ramp}. 
+\texttt{due\_date}, 
+\texttt{road\_ramp},
+and \texttt{bridge\_highway\_name}. 
 The Partially Empty category includes:
-\texttt{cross\_street\_1 \& \_2},
-\texttt{location\_type}, and texttt{landmark}. And 
-the Few/None Empty category includes \texttt{created\_date}, 
+\texttt{location\_type}, 
+\texttt{landmark} and \texttt{cross\_street\_1 \& \_2}.
+The Few/None Empty category includes \texttt{created\_date}, 
 \texttt{complaint\_type},
@@
 and \texttt{status}. 
-Of particular note is that several of the Mostly Empty fields are used by only one or 
+Note that several of the Mostly Empty fields are used by only one or 
 two Agencies (See: \autoref{tab:field-usage-summary}).
@@
 \texttt{due\_date}                 & 0        & 0        & 0        & 0        & 0        & 32097    & 0      \\
-\texttt{community\_board}          & 361715   & 99613    & 203715   & 401143   & 266201   & 608188   & 73601  \\
 \texttt{vehicle\_type}             & 0        & 0        & 0        & 0        & 0        & 0        & 1696   \\
@@
 \texttt{bridge\_highway\_name}     & 15       & 21012    & 0        & 7451     & 0        & 613      & 977    \\
-\texttt{bridge\_highway\_direction} & 15      & 8973     & 0        & 7450     & 0        & 614      & 1006   \\
 \texttt{road\_ramp}                & 15       & 2588     & 0        & 7450     & 0        & 613      & 999    \\
-\texttt{bridge\_highway\_segment}  & 15       & 21013    & 0        & 7450     & 0        & 613      & 987    \\
 \bottomrule

### 2025-01-14 manuscript/manuscript.tex
minor edits

@@
 analyzing the data effectively, often necessitating extensive 
-data cleaning. We encountered three structural issues.
+data cleaning. We identified three primary structural issues in the
+NYC 311 dataset: undocumented fields, inconsistent representations of
+missing data, and daylight saving time anomalies.
 
@@
 
-\subsection{Accuracy and Precision:}
+\subsection{Accuracy and Precision}
 \label{sec:precision}
@@
 public services, research, and other applications.
+To facilitate implementation, the principles are prioritized based on
+their impact and feasibility, offering a clear roadmap for improving
+data curation practices.
 
 \subsection{Principles}
+
 \paragraph{Principle 1: Establishing Cross-Organizational Consistency}
@@
 harmonization of data fields across various Agencies, and 
-maintaining that consistency over time. To address this, Agencies should adopt 
+maintaining that consistency over time. Inconsistent field
+definitions, data formats, or value domains can lead to erroneous
+analyses and hinder cross-agency collaboration.
+To address this, Agencies should adopt 
 standardized naming conventions, value domains, usages, and 
@@
 should establish clear protocols for data entry, ensuring that 
-fields are populated with valid, legally acceptable values. To 
+fields are populated with valid, legally acceptable values. Accurate
+data entry reduces the risk of introducing systemic errors,
+particularly in high-volume datasets. To
 the maximum extent possible,the software should assist in 
@@
 large government datasets. Our research found several duplicate and
-near-duplicate fields in the 311 SR dataset. Eliminating 
+near-duplicate fields in the 311 SR dataset. Removing these
+redundancies could reduce the dataset size by 39.5\%, improving
+performance without compromising analytical value. Eliminating 
 redundant fields and encoding categorical variables can 
@@
 instances where the Data Dictionary did not accurately reflect the 
-actual data. Domains of legal/allowable values should be referenced
+actual data. For example, undocumented fields and outdated domain
+definitions in the 311 SR dataset introduced ambiguity, complicating
+analysis. Domains of legal/allowable values should be referenced
 in the Data Dictionary. Regular reviews and updates should 
@@
 stem from a cron script process that distorts the 
-accurate capture of SR durations. By implementing 
+accurate capture of SR durations. Real-time validation systems could
+prevent such errors, ensuring data integrity at the point of entry. By
+implementing
 real-time and event-driven validation tools, governments 
@@
 
+Transparency is a core tenet of open data, and governments must ensure that datasets are not only accurate but also easily accessible and readily understandable.
+**This includes publishing clear metadata, which enables users to understand the dataset’s structure, limitations, and intended use cases.** Additionally, governments should communicate transparently about data quality issues, such as notifying the public when errors in SR fields (e.g., erroneous \texttt{closed\_date(s)}) are discovered, including an explanation about how such issues will be resolved.
+By ensuring transparency and keeping the public informed, governments foster trust and promote wider use of datasets for research, civic engagement, and innovation.
+
 \paragraph{Principle 6: Addressing Data Transparency}
@@
 metadata and making datasets available in user-friendly 
-formats. Additionally, governments should communicate transparently about 
-data quality issues, such as notifying the public when errors in SR 
+formats. This includes publishing clear metadata, which enables users
+to understand the dataset’s structure, limitations, and intended use
+cases. Additionally, governments should communicate transparently
+about data quality issues, such as notifying the public when errors in SR 
 fields (e.g., erroneous \texttt{closed\_ date(s)} are discovered, including
@@
 While these principles are broad, actionable steps are required to 
-implement them effectively. Governments can begin by investing in 
+implement them effectively. Prioritizing these steps will help
+governments address the most critical data curation challenges
+first. Governments can begin by investing in
 real-time validation systems, regular updates to data dictionaries, 
@@
 (e.g., flagging incorrect zip codes), significantly improving data 
-quality. Collaboration with city agencies and stakeholders is crucial 
+quality. Automated systems can proactively detect and correct common
+issues, reducing the manual workload.
+Collaboration with city agencies and stakeholders is crucial
 to ensure that data standards, formats, and metadata meet the needs 
@@
 capture and display the quality of the data sets as presented
-on appropriate dashboards. Deviations in quality would be
+on appropriate dashboards. These dashboards provide a real-time
+snapshot of data quality, making deviations immediately visible and
+actionable. Deviations in quality would be
 automatically highlighted, alerting appropriate corrective actions.
@@
 cleaning, identifying inconsistent data entries, and ensuring the 
-dataset(s) structure is maintained over time. Their expertise is also 
-critical in developing protocols for validating incoming data, correcting 
-errors in real-time, and ensuring consistency across multiple data sources. 
+dataset(s) structure is maintained over time. Their involvement also
+helps to establish best practices for validating incoming data and
+maintaining consistency across data sources.
 Feedback, such as the issues reported 
@@
 recommendations for improving data curation and validation 
-are not only heard and discussed, but implemented. This feedback 
-mechanism would also allow ongoing dialogue between 
-data users and City officials, leading to more dynamic 
-and responsive data management practices.
+are not only heard and discussed, but implemented. An iterative
+feedback system enables timely resolution of issues and fosters
+long-term improvements in data management.
 
@@
 should continue these types of initiatives, which not only promote 
-data literacy, but encourage community-driven improvements to datasets. 
+data literacy, but encourage community-driven improvements to
+datasets. These events can strengthen public engagement and
+demonstrate the societal value of open data initiative
 By leveraging the expertise of data scientists, statisticians, stakeholders, 
@@
 
+While the findings here are specific to NYC 311 data, the principles 
+proposed have broader applicability. Datasets in healthcare, 
+transportation, and environmental monitoring can similarly benefit 
+from these curation strategies, enabling improved outcomes across 
+domains.
 This research further emphasizes the necessity of harmonizing data
@@
 at \url{https://github.com/jun-yan/nyc311clean}. Reference 
-data files (311 SR data, USPS zip codes) are available 
+data files (311 SR data of 2022--2023, USPS zip codes) are available 
 at \url{https://doi.org/10.6084/m9.figshare.c.7617311}.

### 2025-02-07 data_anomalies/code/data_anomaly_presentation.qmd
Establishing a new folder for the 2025 NYC Open Data Week
Initial updload to Git for the work associated with the 2025 NYC Open Data Week presentations.  Expect to delete many of the files in the "function" folder over time.

---
title: "Identifying Data Anomalies"
author: 
  - "David Tussey (formerly w/DoITT)"
  - "Dr. Jun Yan (Dept of Statistics, UConn)"
format: revealjs
---

## Introduction

-   NYC 311 Service Requests
-   October through December 2024
-   Over 900,000 service requests
-   41 fields of information

## What is Data Hygiene?

Key aspects we'll examine:

-   Missing Values
-   Invalid Values
-   Inconsistent Representations
-   Structural Issues

## Missing Values Analysis {background-color="#f8f9fa"}

Is our data missing any values?

Let's analyze key fields:

::: callout-tip
## Interactive Analysis

[Launch Missing Values Check](http://127.0.0.1:4749)
:::

## Interpreting Missing Values

What we found:

-   `incident_zip`: About 1% blank values
-   `descriptor`: Mix of blank and "Missing" values
-   `resolution_description`: N/A values common
-   Several fields with multiple missing value types

## Next Analysis: ZIP Code Validation {background-color="#f8f9fa"}

Are all ZIP codes valid for NYC?

::: callout-tip
## Interactive Analysis

[Launch ZIP Code Check](http://127.0.0.1:6353)
:::

## Future Analyses

Additional checks we can run:

-   Date field consistency
-   Borough validation
-   Geographic coordinate validation
-   Category standardization

## Recommendations

Based on our findings:

1.  Standardize how missing values are recorded
2.  Add input validation for ZIP codes
3.  Review data collection process
4.  Document data quality issues found

### 2025-02-07 code/datacleansing.R

@@ -44,12 +44,16 @@ setwd("C:/Users/David/OneDrive/Documents/datacleaningproject/nyc311clean/code")
 #########################################################################
 rm(list = ls())
 
-main_data_file <- "311_Service_Requests_from_2023-2024_AS_OF_01-05-2025.CSV"
+#main_data_file <- "311_Service_Requests_from_2023-2024_AS_OF_01-05-2025.CSV"
+main_data_file <- "3-month_311SR_10-01-2024_thru_12-31-2024_AS_OF_02-02-2025.csv"
 #main_data_file <- "smaller_test_data.csv"
 #main_data_file <- "extra_small.csv"
 
-# Hard code the max_closed_date to be midnight of the date of the data export from NYC Open Data
-max_closed_date <- as.POSIXct("2024-09-15 23:59:59", format = "%Y-%m-%d %H:%M:%S")
+# Extract the 10-character date after "AS_OF_"
+max_closed_date <- sub(".*AS_OF_([0-9-]+)\\.csv$", "\\1", main_data_file)
+
+# Convert to POSIXct format
+max_closed_date <- as.POSIXct(max_closed_date, format = "%m-%d-%Y", tz = "UTC") + (23*3600 + 59*60 + 59)
 
 #########################################################################
 programStart <- as.POSIXct(Sys.time())
@@ -648,7 +652,7 @@ points_df <- data.frame(
 )
 
 # Convert to an sf object and apply the State Plane projection.
-points_sf <- st_as_sf(points_df, coords = c("lon", "lat"), crs = 4326) # WGS84 lat/long
+points_sf <- st_as_sf(points_df, coords = c("lon", "lat"), crs = 4326) # WGS84 lat/long.
 points_sp <- st_transform(points_sf, crs = 2263) # Convert to State Plane
 
 # Change the x/y_coordinate_state_plane fields to "numeric" to enable comparison.

### 2025-02-07 data_anomalies/code/data_anamoly_setup.R

@@ -1,0 +1,238 @@
+#########################################################################
+# -------------------------------------------------------------
+# 📦 INSTALL AND LOAD REQUIRED PACKAGES
+# -------------------------------------------------------------
+required_packages <- c(
+  "arrow",
+  "bslib",
+  "DT",
+  "data.table",
+  "dplyr",
+  "ggplot2",
+  "ggpmisc",
+  "httr",
+  "lubridate",
+  "renv",
+  "rlang",
+  "scales",
+  "sf",
+  "shiny",
+  "stringdist",
+  "stringr",
+  "styler",
+  "tidyverse",
+  "zoo"
+)
+
+# Check and install missing packages
+missing_packages <- required_packages[!(required_packages %in% installed.packages()[, "Package"])]
+if (length(missing_packages)) {
+  install.packages(missing_packages)
+  cat("📦 Installed missing packages:", paste(missing_packages, collapse = ", "), "\n")
+}
+
+# Load the required packages
+lapply(required_packages, library, character.only = TRUE)
+
+# -------------------------------------------------------------
+# 📁 Set Working Directory for the Project.
+# -------------------------------------------------------------
+# Set working directory to the location of the initialization script
+setwd("C:/Users/David/OneDrive/Documents/datacleaningproject/data_anomalies/code")
+ 
+#########################################################################
+rm(list = ls())
+
+main_data_file <- "3-month_311SR_10-01-2024_thru_12-31-2024_AS_OF_02-02-2025.csv"
+
+# Extract the 10-character date after "AS_OF_"
+max_closed_date <- sub(".*AS_OF_([0-9-]+)\\.csv$", "\\1", main_data_file)
+
+# Convert to POSIXct format
+max_closed_date <- as.POSIXct(max_closed_date, format = "%m-%d-%Y", tz = "UTC") + (23*3600 + 59*60 + 59)
+
+#########################################################################
+programStart <- as.POSIXct(Sys.time())
+formattedStartTime <- format(programStart, "%Y-%m-%d %H:%M:%S")
+
+cat("\n***** Program initialization *****")
+
+# Set the base directory under the working directorybase_dir <- getwd()
+base_dir <- getwd()
+
+# Define the console output directory and file name.
+output_dir <- file.path(base_dir, "console_output")
+output_file <- file.path(output_dir, "core_console_output.txt")
+
[…]

### 2025-02-07 data_anomalies/code/functions/adjust_feb_29_to_28.R

@@ -1,0 +1,30 @@
+# Define a function to adjust February 29 to February 28 for specified date columns
+adjust_feb_29_to_28 <- function(data, date_cols) {
+  
+  for (col in date_cols) {
+    
+    if (col %in% names(data)) {
+      
+      # Identify rows where the date occurs on February 29
+      feb_29_indices <- which(!is.na(data[[col]]) & format(data[[col]], "%m-%d") == "02-29")
+      num_29Feb_rows <- length(feb_29_indices)
+      
+      if (num_29Feb_rows > 0) {
+        cat("\nAdjusted", num_29Feb_rows, "rows in the", col, "column from 29 February to 28 February\n")
+        
+        # Replace February 29 dates with February 28 of the same year
+        data[[col]][feb_29_indices] <- as.POSIXct(
+          paste(
+            format(data[[col]][feb_29_indices], "%Y-02-28"),
+            format(data[[col]][feb_29_indices], "%H:%M:%S")
+          ),
+          tz = "UTC"
+        )
+      } else {
+        cat("\nNo rows to adjust in the", col, "column. No February 29 dates found.\n")
+      }
+    }
+  }
+  
+  return(data)
+}

### 2025-02-07 data_anomalies/code/functions/are_valid_values.R

@@ -1,0 +1,65 @@
+
+# #########################################################################
+# Validate that all fields are in the list of allowable values
+# This function works on two lists: dataset and list_of_valid_values
+are_valid_values <- function(
+    dataset,
+    list_of_valid_values,
+    field_name) {
+  
+  # Count the number of blank entires. Use later when computing percentage.
+  number_of_blank_entries <- sum(!nzchar(dataset) | is.na(dataset))
+  
+  # Subset the vector to keep only the non-blank values
+  dataset <- dataset[nzchar(dataset) & !is.na(dataset) & dataset != ""]
+  dataset_length <- length(dataset)
+  
+  in_list <- dataset %in% list_of_valid_values
+  not_in_list <- dataset[!in_list]
+  
+  num_invalid <- length(not_in_list)
+  num_valid <- length(in_list)
+  
+  if (num_invalid > 0) {
+    invalid_rows <- dataset[dataset %in% not_in_list]
+    percentage_invalid <- round((num_invalid/(dataset_length - number_of_blank_entries)) * 100, 2)
+    unique_invalid <- length(unique(not_in_list))
+    
+    cat(
+      "\n\nThere are",
+      format(num_invalid, big.mark = ","),
+      "invalid", field_name, "entries \nrepresenting",
+      percentage_invalid,
+      "% of non-blank data,\n"
+    )
+    cat(
+      "comprised of",
+      unique_invalid,
+      "different", field_name, "entries.\n"
+    )
+    
+    # Sort the table in descending order
+    sorted_invalid_table <- sort(table(not_in_list), decreasing = TRUE)
+    
+    # Convert the table to a data frame and calculate the percentage column
+    invalid_df <- data.frame(
+      invalid_names = names(sorted_invalid_table),
+      count = as.numeric(sorted_invalid_table)
+    )
+    invalid_df$percentage <- round(prop.table(invalid_df$count) * 100, 2)
+    invalid_df$cumulative_percentage <- cumsum(invalid_df$percentage)
+    
+    # Print the top 10 values
+    cat("\nTop Ten invalid '", field_name, "':\n")
+    print(head(invalid_df, 10), right = FALSE)
+
+    results <- list(valid = all(in_list), invalid = not_in_list)
+    return(results)
+  } else {
+    cat("\n\nAll values of '", field_name, "' are valid.")
+    results <- list(all(in_list), not_in_list)
+    return(results)
+  }
+}
+
+#########################################################################

### 2025-02-07 data_anomalies/code/functions/detect_duplicates.R

@@ -1,0 +1,50 @@
+#########################################################################
+detect_duplicates <- function(
+    dataset,
+    reference_field,
+    duplicate_field) {
+  
+  #Identify matching fields including blanks and NA values
+  rows_condition_0 <- dataset[[reference_field]] == dataset[[duplicate_field]]
+  
+  rows_condition_1 <- dataset[[reference_field]] != dataset[[duplicate_field]]
+  
+  num_rows_dataset <- nrow(dataset)
+  
+  all_non_matching_rows <- rows_condition_1
+  non_matching_fields <- dataset[all_non_matching_rows, ]
+  num_non_matching_fields <- nrow(non_matching_fields)
+  non_matching_percentage <- round((num_non_matching_fields / num_rows_dataset) * 100, 2)
+  non_matching_fields <- non_matching_fields %>%
+    select("unique_key", all_of(reference_field), all_of(duplicate_field), "agency")
+  
+  all_matching_rows <- rows_condition_0
+  matching_fields <- dataset[all_matching_rows, ]
+  num_matching_fields <- nrow(matching_fields)
+  matching_percentage <- round((num_matching_fields / num_rows_dataset) * 100, 2)
+  matching_fields <- matching_fields %>%
+    select("unique_key", all_of(reference_field), all_of(duplicate_field), "agency")
+  
+  if (num_non_matching_fields > 0) {
+    
+    if (num_matching_fields > 0) {
+      cat(
+        "\n\nThere are",
+        format(num_matching_fields, big.mark = ","),
+        "matches between '", reference_field, "' and '", duplicate_field, "'\nrepresenting",
+        matching_percentage, "% of data.\n")
+    }
+    cat(
+      "\nThere are",
+      format(num_non_matching_fields, big.mark = ","),
+      "non-matches between '", reference_field, "' and '", duplicate_field, "'\nrepresenting",
+      non_matching_percentage, "% of data.\n")
+    
+    return(non_matching_fields)
+  }
+  else {
+    cat("\nThere is a 100% match between '", reference_field, "' and '", duplicate_field)
+  }
+}
+
+#########################################################################

### 2025-02-07 data_anomalies/code/shiny_apps/anomaly_dashboard_app.R

@@ -1,0 +1,160 @@
+##################################################################################
+library(shiny)
+library(shinydashboard)
+library(bslib)
+library(scales)  # For comma formatting
+
+##################################################################################
+# Load the data
+data_path <- "C:/Users/David/OneDrive/Documents/datacleaningproject/data_anomalies/code/data/3-month_311SR_10-01-2024_thru_12-31-2024_AS_OF_02-02-2025.rds"
+cleaned_data <- readRDS(data_path)
+
+# Specify the port for the dashboard
+options(shiny.port = 4001)
+
+##############################################################################################
+ui <- dashboardPage(
+  dashboardHeader(title = "Data Anomaly Explorer"),
+  
+  dashboardSidebar(
+    sidebarMenu(
+      menuItem("Dashboard Overview", tabName = "home", icon = icon("house")),
+      menuItem("Identify Missing Values", tabName = "missing_values", icon = icon("circle-exclamation")),
+      menuItem("Validate against Data Schema", tabName = "schema_validator", icon = icon("check-circle")),
+      menuItem("Validate Zipcodes", tabName = "zip_validator", icon = icon("map-pin")),
+      menuItem("Evaluate Date Field Logic", tabName = "date_validator", icon = icon("calendar-check")),
+      menuItem("Detect Duplicate Fields", tabName = "field_duplication", icon = icon("copy")),
+      menuItem("Detect Dupes via Fuzzy Matching", tabName = "fuzzy_matcing", icon = icon("copy")),
+      menuItem("Project Details", tabName = "about", icon = icon("info-circle"))
+    )
+  ),
+  
+  dashboardBody(
+    tabItems(
+      # Home/Dashboard Overview Tab
+      tabItem(tabName = "home",
+              fluidRow(
+                box(
+                  title = "NYC 311 Service Requests Data Anomaly Explorer",
+                  width = 12,
+                  status = "primary",
+                  solidHeader = TRUE,
+                  HTML("
+              <p>This dashboard provides tools to analyze and validate 311 Service Request data.</p>
+              <ul>
+                <li>Detect and explore missing values</li>
+                <li>Validate data schema and constraints</li>
+                <li>Check for Zipcode accuracy</li>
+                <li>Evaluate date fields for logical consistency</li>
+                <li>Detect duplicate fields</li>
+                <li>Detect duplicates fields using fuzzy matchings</li>
+              </ul>
+            ")
+                )
+              ),
+              fluidRow(
[…]

### 2025-02-07 data_anomalies/code/shiny_apps/duplicate_fields_app.R

@@ -1,0 +1,197 @@
+##############################################################################################
+library(bslib)
+library(data.table)
+library(DT)
+library(shiny)
+library(shinycssloaders)
+library(stringr)
+
+# Configure options
+options(shiny.port = 4006)
+
+# Load and prepare data
+data_path <- "C:/Users/David/OneDrive/Documents/datacleaningproject/data_anomalies/code/data/3-month_311SR_10-01-2024_thru_12-31-2024_AS_OF_02-02-2025.rds"
+cleaned_data <- readRDS(data_path)
+setDT(cleaned_data)
+
+##############################################################################################
+parse_location_field <- function(location_string) {
+  # More robust coordinate extraction
+  # This will work with various formats like (40.640, -73.973), 40.640, -73.973, etc.
+  coords <- stringr::str_extract_all(location_string, "-?\\d+\\.\\d+")[[1]]
+  
+  if(length(coords) == 2) {
+    return(list(
+      lat = as.numeric(coords[1]),
+      lon = as.numeric(coords[2])
+    ))
+  } else if(length(coords) == 4) {
+    # Handle cases with 4 numbers (might include parentheses)
+    return(list(
+      lat = as.numeric(coords[3]),
+      lon = as.numeric(coords[4])
+    ))
+  }
+  
+  return(list(lat = NA_real_, lon = NA_real_))
+}
+
+##############################################################################################
+evaluate_coordinate_matches <- function(data) {
+
+  # Prepare data for comparison - exclude both NA and empty string
+  valid_cases <- data[!is.na(latitude) & !is.na(longitude) & 
+                        !is.na(location) & location != ""]
+  
+  # Extract coordinates from location field
+  matches <- regmatches(valid_cases$location, gregexpr("-?\\d+\\.\\d+", valid_cases$location))
+  extracted_lat <- as.numeric(sapply(matches, `[`, 1))
+  extracted_lon <- as.numeric(sapply(matches, `[`, 2))
+  
+  total_pairs <- length(extracted_lat)
+  
+  # Compare extracted coordinates with existing latitude and longitude
+  coordinate_matches <- valid_cases[
+    abs(valid_cases$latitude - extracted_lat) < 1e-5 & 
+      abs(valid_cases$longitude - extracted_lon) < 1e-5
+  ]
+  
+  results <- data.table(
+    Field_Pair = "Lat/Long - Location",
+    Count = nrow(coordinate_matches),
[…]

### 2025-02-07 data_anomalies/code/shiny_apps/fuzzy_matching_app.R

@@ -1,0 +1,961 @@
+##################################################################################
+# fuzzy_matching_app.R
+library(shiny)
+library(data.table)
+library(stringdist)
+library(DT)
+library(shinycssloaders)
+library(shinyjs)
+library(phonics)
+library(stringdist)
+library(future)
+library(promises)
+library(later)
+
+# Specify the port (choose a unique port number)
+options(shiny.port = 4007)
+
+##################################################################################
+# Load data
+data_path <- "C:/Users/David/OneDrive/Documents/datacleaningproject/data_anomalies/code/data/3-month_311SR_10-01-2024_thru_12-31-2024_AS_OF_02-02-2025.rds"
+cleaned_data <- readRDS(data_path)
+setDT(cleaned_data)
+
+# Initialize clean_already flag
+clean_already <- FALSE
+
+# Set normalize_count for tracking normalizatio
+normalize_count <- 0  
+
+##################################################################################
+# Create a vector of street suffix mappings
+street_suffix_map <- c(
+  "AVENUE" = "AVE",
+  "STREET" = "ST",
+  "ROAD" = "RD",
+  "BOULEVARD" = "BLVD",
+  "DRIVE" = "DR",
+  "BEND" = "BND",
+  "PARKWAY" = "PKWY",
+  "COURT" = "CT",
+  "LANE" = "LN",
+  "EXPRESSWAY" = "EXPY",
+  "TERRACE" = "TER",
+  "PATH" = "PATH",
+  "HIGHWAY" = "HWY",
+  "WAY" = "WAY",
+  "ALLEY" = "ALY",
+  "WALK" = "WALK",
+  "LOOP" = "LOOP",
+  "CIRCLE" = "CIR",
+  "PLAZA" = "PLZ",
+  "CRESCENT" = "CRES",
+  "BRIDGE" = "BRG",
+  "VIADUCT" = "VIA",
+  "TURNPIKE" = "TPKE",
+  "OVERPASS" = "OPAS",
+  "SQUARE" = "SQ",
+  "PLACE" = "PL"
+)
+
+##################################################################################
+street_prefix_map <- c(
+  "AVENUE" = "AVE",
+  "EAST" = "E",
+  "WEST" = "W",
+  "NORTH" = "N",
+  "SOUTH" = "S",
+  "SAINT" = "ST"
+)
+
+##################################################################################
+# Create pattern lists for faster matching
+street_patterns <- list(
+  suffix = mapply(
[…]

### 2025-02-07 data_anomalies/code/shiny_apps/illogical_dates_app.R

@@ -1,0 +1,320 @@
+##############################################################################################
+library(shiny)
+library(shinycssloaders)
+library(DT)
+library(data.table)
+library(lubridate)
+
+# Configure options
+options(shiny.port = 4005)
+options(digits.secs = 3)
+
+# Load and prepare data
+data_path <- "C:/Users/David/OneDrive/Documents/datacleaningproject/data_anomalies/code/data/3-month_311SR_10-01-2024_thru_12-31-2024_AS_OF_02-02-2025.rds"
+cleaned_data <- readRDS(data_path)
+setDT(cleaned_data)
+
+##############################################################################################
+# Process date columns
+date_cols <- c("created_date", "closed_date", "due_date", "resolution_action_updated_date")
+for(col in date_cols) {
+  if(col %in% names(cleaned_data)) {
+    cleaned_data[, (col) := as.POSIXct(get(col), origin="1970-01-01", tz="UTC")]
+  }
+}
+
+##############################################################################################
+# Helper function for getting unique samples
+get_unique_samples <- function(data, date1_col, date2_col = NULL, n = 5) {
+  if (is.null(date2_col)) {
+    # For single date samples
+    samples <- unique(data[order(get(date1_col))][, .(
+      date = format(get(date1_col), "%Y-%m-%d %H:%M:%S")
+    )])[1:n]
+    return(paste(samples$date, collapse = "; "))
+  } else {
+    # For date pair samples
+    samples <- unique(data[order(get(date2_col))][, .(
+      first_date = format(get(date1_col), "%Y-%m-%d %H:%M:%S"),
+      second_date = format(get(date2_col), "%Y-%m-%d %H:%M:%S")
+    )])[1:n]
+    return(paste(mapply(function(d1, d2) paste0(d1, " -> ", d2),
+                        samples$first_date, 
+                        samples$second_date), 
+                 collapse = "; "))
+  }
+}
+
+##############################################################################################
+validate_date_logic <- function(data) {
+  results <- data.table(
+    Rule = character(),
+    Error_Type = character(),
+    Count = numeric(),
+    Percentage = numeric(),
+    Sample_Values = character()
+  )
+  
+  total_rows <- nrow(data)
+  current_date <- Sys.Date()
+  
+  # 1. Check for negative duration (closed_date before created_date)
+  invalid_closure <- data[!is.na(closed_date) & 
+                            !is.na(created_date) & 
[…]

### 2025-02-07 data_anomalies/code/shiny_apps/missing_values_app.R

@@ -1,0 +1,125 @@
+library(shiny)
+library(shinycssloaders)
+library(DT)
+library(data.table)
+
+# Specify the port (choose a unique port number)
+options(shiny.port = 4002)
+
+##############################################################################################
+# Load the data
+data_path <- "C:/Users/David/OneDrive/Documents/datacleaningproject/data_anomalies/code/data/3-month_311SR_10-01-2024_thru_12-31-2024_AS_OF_02-02-2025.rds"
+cleaned_data <- readRDS(data_path)
+setDT(cleaned_data)  # Convert to data.table for better performance
+
+##############################################################################################
+ui <- fluidPage(
+  theme = bslib::bs_theme(version = 5, bootswatch = "flatly"),
+  
+  titlePanel("Missing Values Analysis"),
+  
+  fluidRow(
+    column(12,
+           actionButton("analyze", "Analyze Missing Values", class = "btn-primary btn-lg"),
+           br(), br(),
+           shinycssloaders::withSpinner(DTOutput("results_table"), type = 4)
+    )
+  )
+)
+
+##############################################################################################
+server <- function(input, output, session) {
+  missing_results <- reactiveVal(NULL)
+  is_loading <- reactiveVal(FALSE)
+  
+  observeEvent(input$analyze, {
+    is_loading(TRUE)
+    missing_results(NULL)
+    
+    results <- local({
+      # Pre-allocate results table
+      max_rows <- length(names(cleaned_data)) * 2
+      results <- data.table(
+        Field = character(max_rows),
+        Missing_Type = character(max_rows),
+        Count = numeric(max_rows),
+        Percentage = numeric(max_rows)
+      )
+      
+      row_counter <- 0L
+      total_rows <- nrow(cleaned_data)
+      
+      # Process each column
+      for(col in names(cleaned_data)) {
+        col_type <- class(cleaned_data[[col]])[1]
+        has_missing <- FALSE  # Flag to track if any missing values found
+        
+        # Check for actual NAs (NULL values)
+        na_count <- sum(is.na(cleaned_data[[col]]))
+        if(na_count > 0) {
+          has_missing <- TRUE
+          row_counter <- row_counter + 1L
+          set(results, i = row_counter, j = 1L:4L,
+              value = list(col, "NULL", na_count,
+                           round(na_count/total_rows, 4)))
+        }
+        
+        # For character columns, check for empty strings and N/A text
+        if(col_type == "character") {
[…]

### 2025-02-07 data_anomalies/code/shiny_apps/run_apps/run_anomaly_dashboard_app.R

@@ -1,0 +1,7 @@
+################################################################################
+
+# Run the duplicate_fields_app.R
+library(shiny)
+runApp("C:\\Users\\David\\OneDrive\\Documents\\datacleaningproject\\data_anomalies\\code\\shiny_apps\\anomaly_dashboard_app.R", port = 4001)
+
+################################################################################

### 2025-02-07 data_anomalies/code/shiny_apps/run_apps/run_duplicate_fields_app.R

@@ -1,0 +1,7 @@
+################################################################################
+
+# Run the duplicate_fields_app.R
+library(shiny)
+runApp("C:\\Users\\David\\OneDrive\\Documents\\datacleaningproject\\data_anomalies\\code\\shiny_apps\\duplicate_fields_app.R", port = 4006)
+
+################################################################################

### 2025-02-07 data_anomalies/code/shiny_apps/run_apps/run_fuzzy_matching_app.R

@@ -1,0 +1,7 @@
+################################################################################
+
+# Run the fuzzy_matching_app.R
+library(shiny)
+runApp("C:\\Users\\David\\OneDrive\\Documents\\datacleaningproject\\data_anomalies\\code\\shiny_apps\\fuzzy_matching_app.R", port = 4007)
+
+################################################################################

### 2025-02-07 data_anomalies/code/shiny_apps/run_apps/run_illogical_dates_app.R

@@ -1,0 +1,7 @@
+################################################################################
+
+#App to run illogical_dates_app.R shiny app
+library(shiny)
+runApp("C:\\Users\\David\\OneDrive\\Documents\\datacleaningproject\\data_anomalies\\code\\shiny_apps\\illogical_dates_app.R", port = 4005)
+
+################################################################################

### 2025-02-07 data_anomalies/code/shiny_apps/run_apps/run_missing_values_app.R

@@ -1,0 +1,7 @@
+################################################################################
+
+# Run the missing_values_app.R
+library(shiny)
+runApp("C:\\Users\\David\\OneDrive\\Documents\\datacleaningproject\\data_anomalies\\code\\shiny_apps\\missing_values_app.R", port = 4002)
+
+################################################################################

### 2025-02-07 data_anomalies/code/shiny_apps/run_apps/run_schema_validator_app.R

@@ -1,0 +1,7 @@
+################################################################################
+
+# App to run schema_validator.R shiny app
+library(shiny)
+runApp("C:\\Users\\David\\OneDrive\\Documents\\datacleaningproject\\data_anomalies\\code\\shiny_apps\\schema_validator_app.R", port = 4003)
+
+################################################################################

### 2025-02-07 data_anomalies/code/shiny_apps/schema_validator_app.R

@@ -1,0 +1,573 @@
+##############################################################################################
+library(shiny)
+library(shinycssloaders)
+library(DT)
+library(data.table)
+library(stringr)
+library(bslib)
+
+##############################################################################################
+# Configuration items
+# Specify the port (choose a unique port number)
+options(shiny.port = 4003)
+options(digits.secs = 3)  # Ensure subsecond precision is maintained
+
+# Load and prepare data
+data_path <- "C:/Users/David/OneDrive/Documents/datacleaningproject/data_anomalies/code/data/3-month_311SR_10-01-2024_thru_12-31-2024_AS_OF_02-02-2025.rds"
+cleaned_data <- readRDS(data_path)
+setDT(cleaned_data)
+
+# Process date columns
+date_cols <- c("created_date", "closed_date", "due_date", "resolution_action_updated_date")
+for(col in date_cols) {
+  if(col %in% names(cleaned_data)) {
+    cleaned_data[, (col) := as.POSIXct(get(col), origin="1970-01-01", tz="UTC")]
+  }
+}
+
+##############################################################################################
+schema <- list(
+  unique_key = list(
+    type = "integer",
+    allow_empty = FALSE,
+    unique = TRUE,
+    description = "Unique identifier for each service request"
+  ),
+  created_date = list(
+    type = "POSIXct",
+    allow_empty = FALSE,
+    description = "Timestamp when the service request was created"
+  ),
+  closed_date = list(
+    type = "POSIXct",
+    allow_empty = TRUE,
+    description = "Timestamp when the service request was closed (can be empty for ongoing requests)"
+  ),
+  agency = list(
+    type = "character",
+    allow_empty = FALSE,
+    description = "Agency code or abbreviation"
+  ),
+  agency_name = list(
+    type = "character",
+    allow_empty = FALSE,
+    valid_values = c(
+      "NEW YORK CITY POLICE DEPARTMENT",
+      "DEPARTMENT OF HOUSING PRESERVATION AND DEVELOPMENT",
+      "DEPARTMENT OF HEALTH AND MENTAL HYGIENE",
+      "DEPARTMENT OF ENVIRONMENTAL PROTECTION", 
+      "DEPARTMENT OF TRANSPORTATION",
+      "DEPARTMENT OF PARKS AND RECREATION",
[…]

### 2025-02-07 data_anomalies/code/shiny_apps/zip_validator_app.R

@@ -1,0 +1,109 @@
+##############################################################################################
+
+library(shiny)
+library(shinycssloaders)
+library(DT)
+library(data.table)  # Adding for better performance
+
+# Specify the port (choose a unique port number)
+options(shiny.port = 4004)
+
+##############################################################################################
+# Load the data
+data_path <- "C:/Users/David/OneDrive/Documents/datacleaningproject/data_anomalies/code/data/3-month_311SR_10-01-2024_thru_12-31-2024_AS_OF_02-02-2025.rds"
+zip_path <- "C:/Users/David/OneDrive/Documents/datacleaningproject/data_anomalies/code/data/USPS_zipcodes.rds"
+cleaned_data <- readRDS(data_path)
+zip_codes <- readRDS(zip_path)
+
+# Convert to data.table for better performance
+setDT(cleaned_data)
+setDT(zip_codes)
+
+##############################################################################################
+# Function to validate zip codes
+are_valid_values <- function(check_values, valid_values) {
+  # Remove any leading/trailing whitespace
+  check_values <- trimws(check_values)
+  valid_values <- trimws(valid_values)
+  
+  # Filter out NA, empty strings, and blank values
+  check_values[is.na(check_values) | check_values == "" | check_values == " "] <- NA
+  
+  # Convert valid_values to a set for faster lookup
+  valid_set <- unique(valid_values)
+  
+  # Check if each value exists in the valid set
+  check_values %in% valid_set
+}
+
+##############################################################################################
+ui <- fluidPage(
+  theme = bslib::bs_theme(version = 5, bootswatch = "flatly"),
+  
+  titlePanel("ZIP Code Validation Analysis"),
+  
+  fluidRow(
+    column(12,
+           actionButton("analyze", "Detect Invalid ZIP Codes", class = "btn-primary btn-lg"),
+           br(), br(),
+           shinycssloaders::withSpinner(DTOutput("results_table"), type = 4)
+    )
+  )
+)
+
+##############################################################################################
+server <- function(input, output, session) {
+  zip_analysis_results <- reactiveVal(NULL)
+  is_loading <- reactiveVal(FALSE)
+  
+  observeEvent(input$analyze, {
+    is_loading(TRUE)
+    zip_analysis_results(NULL)
+    
+    results <- local({
+      invalid_zips <- !are_valid_values(cleaned_data$incident_zip, zip_codes$delivery_zipcode)
+      
+      invalid_zips_filtered <- cleaned_data$incident_zip[invalid_zips]
[…]

### 2025-02-07 data_anomalies/code/functions/archived/make_column_names_user_friendly.R
Initial setup of Github folder for ODW 2025 data quality over time presentation

@@ -1,0 +1,28 @@
+#########################################################################
+  
+# Define the make_column_names_user_friendly function for a character vector
+make_column_names_user_friendly <- function(column_names) {
+  
+  # Replace spaces with underscores
+  column_names <- gsub(
+    x = column_names,
+    pattern = "\\s+",
+    replacement = "_"
+  )
+  
+  # Remove all parentheses
+  column_names <- gsub(
+    x = column_names,
+    pattern = "\\(|\\)",
+    replacement = ""
+  )
+  
+  # Convert to lowercase
+  column_names <- tolower(column_names)
+  
+  # Return the updated column names
+  return(column_names)
+}
+
+
+#########################################################################

### 2025-02-07 data_anomalies/code/shiny_apps/zip_validator.R

@@ -1,0 +1,55 @@
+library(shiny)
+library(DT)
+
+# Load the data
+data_path <- "C:/Users/David/OneDrive/Documents/datacleaningproject/data_hygiene/code/data/3-month_311SR_10-01-2024_thru_12-31-2024.rds"
+zip_path <- "C:/Users/David/OneDrive/Documents/datacleaningproject/data_hygiene/code/data/USPS_zipcodes.rds"
+
+cleaned_data <- readRDS(data_path)
+zip_codes <- readRDS(zip_path)
+
+ui <- fluidPage(
+  titlePanel("ZIP Code Validation Analysis"),
+  
+  fluidRow(
+    column(12,
+           actionButton("analyze", "Analyze ZIP Codes", class = "btn-primary"),
+           br(), br(),
+           DTOutput("results_table")
+    )
+  )
+)
+
+server <- function(input, output, session) {
+  output$results_table <- renderDT({
+    req(input$analyze)
+    
+    # Get invalid ZIPs using the are_valid_values function
+    invalid_zips <- !are_valid_values(cleaned_data$incident_zip, zip_codes$delivery_zipcode)
+    
+    # Create summary of invalid values
+    invalid_summary <- data.frame(
+      ZIP_Code = unique(cleaned_data$incident_zip[invalid_zips]),
+      Count = as.numeric(table(cleaned_data$incident_zip[invalid_zips])),
+      stringsAsFactors = FALSE
+    )
+    
+    # Add percentage
+    total_records <- nrow(cleaned_data)
+    invalid_summary$Percentage <- round(invalid_summary$Count / total_records * 100, 2)
+    
+    # Sort by frequency
+    invalid_summary <- invalid_summary[order(-invalid_summary$Count),]
+    
+    # Display results
+    datatable(invalid_summary,
+              options = list(
+                pageLength = 10,
+                order = list(1, 'desc')  # Sort by Count by default
+              )) %>%
+      formatCurrency("Count", currency = "", interval = 3, mark = ",", digits = 0) %>%
+      formatRound("Percentage", digits = 2)
+  })
+}
+
+shinyApp(ui, server)

### 2025-02-07 measuring_data_quality_over_time/code/functions/archived/hygiene_make_column_names_user_friendly.R

@@ -1,0 +1,31 @@
+#########################################################################
+  
+# Define the make_column_names_user_friendly function for a character vector
+hygiene_make_column_names_user_friendly <- function(column_names) {
+  # Check if the input is a character vector
+  if (!is.character(column_names)) {
+    stop("Input must be a character vector.")
+  }
+  
+  # Replace spaces with underscores
+  column_names <- gsub(
+    pattern = "\\s+",
+    replacement = "_",
+    x = column_names
+  )
+  
+  # Remove all parentheses
+  column_names <- gsub(
+    pattern = "\\(|\\)",
+    replacement = "",
+    x = column_names
+  )
+  
+  # Convert to lowercase
+  column_names <- tolower(column_names)
+  
+  # Return the updated column names
+  return(column_names)
+}
+
+#########################################################################

### 2025-10-22 Journal_of_Data_Science/JDS_manuscript/JDS_manuscript.tex
Latest versions
Near final code repository.

\documentclass[linenumber]{jdsart}
\usepackage{setspace}

\volume{0}
\issue{0}
\pubyear{2022}
\articletype{research-article}
\doi{0000}

\usepackage{siunitx} % For alignment of numbers
\sisetup{
    group-separator = {,},
    round-mode = places,
    round-precision = 2,
    output-decimal-marker = {.},
    table-number-alignment = center,
    table-figures-integer = 6,
    table-figures-decimal = 2,
    table-figures-uncertainty = 2
}

% image path
\graphicspath{{.}{./images}}

\usepackage{xcolor}
\newcommand{\dt}[1]{\textcolor{purple}{\textbf{DT: (#1)}}}

\let\proglang=\textsf
%% \newcommand{\pkg}[1]{{\fontseries{m}\selectfont #1}}
%% \newcommand\code[2][black]{\textcolor{#1}{\texttt{#2}}}

\usepackage{comment}
\usepackage{subcaption}
\usepackage{booktabs, textgreek}
\usepackage{enumitem}
\usepackage{tikz} % For creating diagrams
\usepackage{hyperref}   % For clickable links and breaking long URLs
\usepackage{url}        % Optional if hyperref is not loaded
\usepackage{fontawesome}
\usetikzlibrary{positioning} % Required for relative positioning (e.g., "of" keyword)
\captionsetup[subfigure]{justification=centering, font=small}

%% float control
\renewcommand\floatpagefraction{0.75}
% \renewcommand\topfraction{.8}
% \renewcommand\bottomfraction{.8}
% \renewcommand\textfraction{.2}
\setcounter{totalnumber}{50}
\setcounter{topnumber}{50}
\setcounter{bottomnumber}{50}

\begin{document}

%\doublespacing

%\tableofcontents % Optional: Table of Contents
%\listoffigures % List of Figures
%\listoftables % List of Tables

\begin{frontmatter}

\title{Detecting Data Anomalies in Open Data: A Case Study with the New
York City 311 Service Request Dataset}
\runtitle{Searching for Data Anomalies}

\author[1]{\inits{D.}\fnms{David}~\snm{Tussey}}
\address[1]{\institution{NYC \textsc{DoITT}}, \cny{USA}}

\hyphenpenalty=950

\begin{abstract}
The open data movement has transformed governance by promoting transparency, 
innovation, and public engagement. Since the 2012 Open Data Law, the City of 
New York (NYC) has been a leader, publishing over 2,400 datasets through its 
Open Data portal. Yet the value of open data depends on its quality—a topic 
often overlooked in the literature.

This paper presents a case study of anomaly detection within the popular NYC 311 
Service Request (\textsc{SR}) dataset. The analysis reveals that 
uncritical reliance on open datasets 
can yield misleading results. Detection of these anomalies required custom software and iterative 
exploration to uncover patterns not immediately evident.

Rather than cleaning the data, this study characterizes them, while emphasizing 
that some anomalies—such as a lifecycle of only ten seconds—cannot be judged 
invalid without context. The findings highlight the analytical rigor 
and interpretive judgment necessary to ensure credible insights from open data.
\end{abstract}

\begin{keywords} % no repeating those in title
  \kwd{Data anomalies}
  \kwd{Anomaly detection}
  \kwd{Data quality}
  \kwd{Data validation}
  \kwd{Data curation}
  \kwd{Data cleansing}
  \kwd{Data science}
  \kwd{NYC Open Data}
  \kwd{Smart city}
  \kwd{Transparency}
  \kwd{Government data}
\end{keywords}

\end{frontmatter}

%-------------------------------------------------------------------------------
%	Section: Introduction
%-------------------------------------------------------------------------------

\section{Introduction}
\label{sec:intro}
The open data movement, emerging in the early twenty-first century, rests on 
the premise that freely accessible information can transform governance and 
society by fostering transparency, innovation, and civic engagement. 
A major milestone was the launch of the U.S.\ \textsc{Data.gov} portal in 
2009\,\citep{dataGov}, followed by the European Union’s 
\textsc{Open Data Portal} in 2012\,\citep{dataEU} and the World Bank’s 
\textsc{Open Data} initiative in 2010\,\citep{dataWorldBank}. 
Together, these efforts democratized access to government information and 
encouraged data-driven inquiry across disciplines 
\citep{barns2016mine,wang2016adoption}.  
Yet the effectiveness of open data depends not merely on access but on 
quality, consistency, and usability—issues that remain only partly resolved.

New York City (NYC) has been a municipal leader in this arena since the 
\textsc{Open Data Law} of 2012\,\citep{zuiderwijk2014open}, which established 
the \textsc{NYC Open Data Portal}\,\citep{dataNYC}. 
Hosting more than 2{,}400 datasets across roughly ninety agencies, the portal 
has enhanced governmental transparency and supported extensive research in 
fields such as public health\,\citep{cantor2018facets,shankar2021data}, urban 
development\,\citep{neves2020impacts}, and transportation\,\citep{gerte2019understanding}. 
Datasets on restaurant inspections, traffic collisions, and educational 
enrollment, among others, have informed both policy design and civic 
applications.  In particular, NYC’s 311~Service Request (\textsc{SR}) data have 
been used to optimize resource allocation and improve emergency response, 
demonstrating their operational value.

The size and diversity of these datasets, however, pose persistent challenges 
for data curation—the processes that ensure accuracy, completeness, and 
coherence across time and sources.  High-quality curation underpins effective 
machine-learning models and statistical analyses 
\citep{polyzotis2019data,jain2020overview}, while inadequate practices can 
introduce bias and undermine decision-making 
\citep{geiger2020garbage,rahm2000data}.  
Although open data initiatives emphasize dissemination, the burden of 
verifying and cleaning data often falls to end users 
\citep{cody2017cody,van2018statistical}.

Research on data curation has advanced understanding of these challenges.  
\citet{witt2009constructing} introduced data-curation profiles for specific 
research contexts; \citet{borgman2012conundrum} examined governance and 
sustainability issues; and \citet{hart2016ten} articulated general principles 
for effective management.  Empirical studies further illustrate the value of 
well-curated data in addressing public-health and crisis-response problems 
\citep{cantor2018facets,shankar2021data}.  Yet large-scale municipal datasets 
remain under-analyzed with respect to systematic quality assessment.

The study addresses this gap through a case analysis of the 
\textsc{NYC 311 SR} dataset.  Rather than attempting to cleanse or correct 
records, this investigation focuses on detecting and characterizing anomalies 
that may compromise analytical validity. This approach 
required custom software, iterative exploration, and 
methodological rigor to uncover patterns not readily visible through routine 
inspection.  Building on these results, we propose a set of practical 
principles for anomaly detection and data assessment tailored to 
government-released open data, with the broader goal of enhancing reliability, 
reproducibility, and interpretability in research and policy contexts.

Section~\ref{sec:service} reviews the history of the NYC~311~system and its 
role as a data source. Section~\ref{sec:data} describes the dataset analyzed. 
Section~\ref{sec:anomalies} details the principal data-quality issues, 
while Section~\ref{sec:principles} outlines recommended 
curation principles.  Section~\ref{sec:conclusion} concludes with key 
findings and implications for future open-data initiatives.

%-------------------------------------------------------------------------------
%	Section: NYC 311 Service Request (SR) Data}
%-------------------------------------------------------------------------------

\section{The New York City 311 Non-Emergency Service} 
\label{sec:service}
The NYC 311 service, a cornerstone of the City’s 
service response framework, functions as a hub for 
non-emergency inquiries, complaints, and requests. Launched in 2003, the system was 
designed to streamline responses to issues ranging from noise complaints 
to street maintenance. Initially limited to phone calls, the service 
expanded to include a mobile application, web portal, text 
messaging, social media, and chat support. Over two decades, 
NYC 311 has evolved into a comprehensive data management platform processing over 3.4 million 
annual requests.

Beyond its operational function, \textsc{NYC 311} data are
instrumental in shaping City governance and community engagement. 
Open data enhance transparency while empowering civic developers, 
policymakers, and the public\,\citep{minkoff2016nyc,o2017uncharted,kontokosta2021bias}. 
These data have informed crisis management 
(e.g., directing residents to shelters during emergencies, extreme weather events,
 or coordinating COVID-19 responses), improved everyday urban governance 
(e.g., landlord--tenant enforcement, taxi route realignment by the 
Taxi and Limousine Commission (TLC), rodent control, and agency responsiveness), 
Additionally the data has supported a range of academic studies on urban issues, 
including resource allocation\,\citep{zha2014profiling,raj2021swift}, 
social equity in service provision\,\citep{white2018promises,kontokosta2021bias}, 
and environmental challenges \,\citep{dove2022sounds} 
and street flooding\,\citep{agonafir2022understanding}.

%-------------------------------------------------------------------------------
%	Section: 311 SR Dataset Description
%-------------------------------------------------------------------------------

\section{Dataset Description} 
\label{sec:data}
This paper uses a five-year dataset of NYC 311 Service Requests (SRs), 
downloaded from the NYC Open Data Portal on 10 October 2025. 
The dataset main characteristics are:

\begin{itemize}[left=1.5em]
  	\item The raw dataset consists of 16 million records 
  	(9 GB) in CSV format. For efficiency 
	using the R programming language, the file was 
 	converted to the RDS format. Each record 
	has 41 fields, with each row 
	representing a single SR (e.g., complaint or request).

	 \item Four date fields are expressed in \texttt{YYYY-MM-DD hh:mm:ss} format. 
  	(Not all agencies track times to the second; some record times only to the minute.)
    		\begin{itemize}
      			\item \texttt{created\_date}
      			\item \texttt{closed\_date}
      			\item \texttt{resolution\_action\_updated\_date} (also \texttt{updated\_date})
      			\item \texttt{due\_date}
    		\end{itemize}

  	\item Two borough fields, \texttt{borough} and \texttt{park\_borough}, appear 
  	to duplicate one another.

  	\item Seven street-related fields are included, with two pairs of apparent duplicates:
    		\begin{itemize}
      			\item \texttt{incident\_address}
      			\item \texttt{street\_name}
     			 \item \texttt{landmark}
      			\item \texttt{cross\_street\_1}, \texttt{cross\_street\_2}
      			\item \texttt{intersection\_street\_1}, \texttt{intersection\_street\_2}
    		\end{itemize}

	  \item Eight incident location fields are included (two of which are geo-coded):
		 \begin{itemize}
    			\item \texttt{incident\_address}
    			\item \texttt{latitude} and \texttt{longitude}
    			\item \texttt{location} (a concatenation of latitude and longitude)
    			\item \texttt{street\_name}
    			\item \texttt{landmark}
    			\item \texttt{block} (NYC tax-related mapping system)
    			\item \texttt{x\_coordinate\_state\_plane} and \texttt{y\_coordinate\_state\_plane} 
         		(State Plane Coordinate System, a NOAA-developed surveying framework)
 	 	\end{itemize}
\end{itemize}

\begin{table}[tbp]
  \centering
  \caption{Top six agencies by share of Service Requests (2020--2024).}
  \label{tab:big-six-agencies}
  \begin{tabular}{lrr}
    \toprule
    \textbf{Agency} & \textbf{Abbreviation} & \textbf{Share of SRs} \\
    \midrule
    New York Police Department & NYPD & 43\% \\
    Housing Preservation and Development & HPD & 19\% \\
    Department of Sanitation & DSNY & 12\% \\
    Department of Transportation & DOT & 7\% \\
    Department of Environmental Protection & DEP & 5\% \\
    Department of Parks and Recreation & DPR & 4\% \\
    \midrule
    \textbf{Total (Top Six)} & & \textbf{90\%} \\
    \bottomrule
  \end{tabular}
\end{table}

The dataset contains 252 distinct \texttt{complaint\_type} categories.
However, the top twenty complaints 
account for nearly 70\% of all SRs. The complaint distribution is 
highly skewed, with a small number of issues dominating 
overall volume and a long tail of low-frequency complaints.

Noise-related complaints are especially prominent, represented 
by eight separate categories, including vehicle, residential, commercial, 
and street noise. Taken together, these account for nearly 
one-quarter (23.4\%) of all complaints, making \emph{noise} by 
far the most common issue. Other leading categories 
include illegal parking, heat and hot water complaints, and blocked driveways.

Anomalies in the dataset are typically associated with specific city agencies, 
which facilitates targeted corrective actions. As shown in 
Table~\ref{tab:big-six-agencies}, the six largest agencies account for
nearly ninety percent of all SRs submitted between 2020 and 2024.

%-------------------------------------------------------------------------------
%	Section: Data Cleansing Issues and Anomalies
%-------------------------------------------------------------------------------

\section{Data Cleansing and Anomaly Identification} 
\label{sec:anomalies}
Data cleansing is the process of identifying and correcting errors, 
inconsistencies, anomalies, and inaccuracies to ensure that a dataset is 
sufficiently reliable for analysis\,\citep{maletic2005data,hosseinzadeh2023data}. 
Typical operations include removing duplicates, handling missing values, 
correcting mislabeled entries, and standardizing formats\,
\citep[e.g.,][]{cody2017cody,van2018statistical}. 
For open data, these challenges are amplified by the diversity of sources 
and the lack of coordination across contributing agencies, leading to 
uneven quality and consistency.
When data cleansing is achieved, it enhances trust, supports reliable analysis, and 
facilitates integration into applications such as machine learning. 
This study does not attempt to cleanse the \textsc{NYC 311} dataset but 
instead focuses on the preceding step—systematically detecting and 
characterizing anomalies that may compromise its reliability.

%-------------------------------------------------------------------------------
%	Sub-section: Pre-analysis Data Modifications
%-------------------------------------------------------------------------------

\subsection{Pre-analysis Data Modifications}
\label{subsec:premodifications}
Prior to analysis, the raw dataset undergoes a series of preparatory 
modifications to facilitate efficient processing. These steps 
include both structural adjustments and content standardization:

\begin{itemize}[left=1.5em]
  \item \textbf{Standardization of structure:} 
  Column names are normalized for ease of reference, and the raw 
  \texttt{.csv} file is converted into the more efficient 
  \texttt{.rds} format to improve performance.

  \item \textbf{Date handling:} 
  All date fields are converted to the \texttt{POSIXct} class and aligned 
  to the \textit{America/New\_York} time zone, enabling accurate 
  treatment of Daylight Saving Time (DST) effects 
  (discussed in Section~\ref{par:dst}).

  \item \textbf{Agency harmonization:} 
  Several agencies were renamed during the study period; records were 
  consolidated under their current designations for consistency.

  \item \textbf{Validation of completeness:} 
  Rows are screened for mandatory fields. 
  Missing or absent values—recorded in multiple raw formats—are 
  standardized to \texttt{NA} to facilitate easy in processing.
  Notably, approximately 27\% of the dataset contains missing entries.

  \item \textbf{Field formatting:} 
  Text fields are converted to uppercase for consistent matching, and 
  numeric fields are explicitly cast to the appropriate type.
\end{itemize}

%-------------------------------------------------------------------------------
%	Sub-section: Structural Issues
%-------------------------------------------------------------------------------

\subsection{Structural Issues}
\label{subsec:structural}
Structural issues concern the way data are organized, formatted, or 
encoded---that is, how the dataset is \emph{structured} rather than 
what it \emph{contains}. Structural issues can hinder efficient analysis and often require 
substantial preprocessing before the data can be reliably used. 
In the \textsc{NYC 311} dataset, three notable structural issues were identified:

\begin{itemize}[left=1.5em]
  \item Inconsistent representation of missing-data values
  \item Daylight Saving Time (DST) anomalies
  \item Precision versus accuracy issues in the \texttt{latitude}, \texttt{longitude}, and \texttt{location} fields
\end{itemize}

\paragraph{Inconsistent Representation of Missing Data}
\label{par:missingdata}
The \textsc{NYC 311} dataset exhibits inconsistent conventions for representing 
missing values. Observed forms include nulls, blank spaces, 
``\texttt{NA}'', ``\texttt{N/A}'', and ``\texttt{<NA>}''. 
Such variation complicates programming and analysis, increasing the likelihood 
of errors when filtering or aggregating records. To address this 
issue, all missing values were standardized to \texttt{NA}.

\paragraph{Daylight Saving Time Issues}
\label{par:dst}
The \textsc{NYC 311} dataset records timestamps in New York City local time, 
which observes Daylight Saving Time (DST). 
Twice each year, this practice introduces temporal discontinuities that affect 
the calculation of SR durations, defined as the difference 
between \texttt{closed\_date} and \texttt{created\_date}.

During the spring transition, local clocks advance from 01:59:59 directly to 
03:00:00, eliminating the entire hour between 02:00:00 and 02:59:59. 
As a result, durations of SRs created before the transition and closed shortly 
afterward are exaggerated by up to one hour. 
While this distortion may appear minor, it is consequential for 
time-sensitive requests such as noise complaints or homelessness-related cases.

During the preprocessing steps several date fields were found to 
violate the DST spring-forward rule, with timestamps recorded during the 
nonexistent hour between 02:00:00 and 02:59:59 
(e.g., \texttt{03/10/2024 02:00:46 AM}, \texttt{03/08/2020 02:10:00 AM}). 
Such values are invalid, as local time advances directly from 01:59:59 to 
03:00:00 on those dates. 
These anomalies were corrected by shifting the affected timestamps forward 
to valid times 
(e.g., \texttt{03/10/2024 03:00:46 AM}, \texttt{03/08/2020 03:10:00 AM}). 
In total, nineteen such corrections were made in the dataset.

The fall DST adjustment introduces a more serious systemic issue by producing 
SRs with \emph{negative} durations, where the 
\texttt{closed\_date} precedes the \texttt{created\_date}. 
Such cases are logically impossible and can distort agency-level performance 
measures, particularly for time-sensitive SRs. 
The following example illustrates the problem:

\begin{itemize}[left=1.5em]
  \item A residential noise complaint SR is created at 01{:}50 on Sunday, March~10,~2024, 
  the date of the DST fall-back transition.

  \item The \textsc{NYPD} responds promptly, resolving the issue in 
  forty minutes. 
  Without the DST adjustment, the resolution would normally be recorded 
  at 02{:}30.

  \item However, during this interval the local clock is set back one hour, 
  so the resolution is time-stamped as 01{:}30.

  \item As a result, the dataset records the SR as closed 
  twenty minutes \emph{before} it was created, yielding a nonsensical
  negative duration of $-20$~minutes.  
\end{itemize}

On average, approximately eighty-six SRs per year exhibit this systemic issue, 
with a mean negative duration of $-42$~minutes.

\paragraph{Precision versus Accuracy in Latitude/Longitude Values}
\label{par:precision}
The \texttt{latitude}, \texttt{longitude}, and the duplicative 
\texttt{location} fields in the raw dataset are recorded with 
fourteen decimal places 
(e.g.,~\texttt{latitude}~$=~40.6409147797767$, 
\texttt{longitude}~$=-73.9736421630642$, 
\texttt{location}~$=(40.640914779776715,\,-73.97364216306418)$). 
Accordingly, 14 decimal points imply positional accuracy at 
the molecular level—an implausible representation. 
This is a case of extreme numerical precision rather than meaningful spatial 
accuracy. For comparison, the official New York City Borough Boundary Water Included
dataset (\textsc{NYBBWI}) published by the Department of City Planning provides 
coordinates to six decimal places, with practical accuracy on the order 
of a few meters. 
(\href{https://s-media.nyc.gov/agencies/dcp/assets/files/pdf/data-tools/bytes/nybbwi_metadata.pdf}{NYC~DCP,~2025}). 
In the present analysis, the original coordinate values were preserved 
without rounding and compared to the official city boundary extents 
(\textsc{NYBBWI}) using a tolerance of approximately \SI{100}{\meter}.

%-------------------------------------------------------------------------------
%	Sub-section: Missing Data by Field
%-------------------------------------------------------------------------------

\subsection{Missing Data by Field}
\label{subsec:blanks}
Understanding the extent and distribution of missing data across fields 
is essential for evaluating the overall usability of the dataset. 
For example, assessing whether SRs were closed before 
their \texttt{due\_date} is largely infeasible, as more than 
\SI{99}{\percent} of entries in this field are blank.

Additionally, field usage  varies considerably across agencies.
For instance, \texttt{landmark} and \texttt{facility\_type} are 
widely used across agencies, underscoring their importance in 
classifying service requests (SRs). 
By contrast, fields such as \texttt{taxi\_pick\_up\_location} are used 
exclusively by the Taxi and Limousine Commission (\textsc{TLC}).

Table~\ref{tab:missing-data} summarizes the extent of missingness across 
the dataset, grouping fields by the proportion of blank entries and 
providing representative examples within each category.

\begin{table}[tbp]
  \centering
  \caption{Categories of missing data by field in the NYC~311 SR dataset.}
  \label{tab:missing-data}
  \small
  \begin{tabular}{lll}
    \toprule
    \textbf{Category} & \textbf{Percentage blank} & \textbf{Representative fields} \\
    \midrule
    Mostly empty      & 94--99.9\%   & \texttt{taxi\_company\_borough}, \texttt{due\_date}, \\
                      &              & \texttt{road\_ramp}, \texttt{bridge\_highway\_name} \\
    Partially empty   & 56--93\%      & \texttt{location\_type}, \texttt{address\_type}, \\
                      &              & \texttt{landmark}, \texttt{cross\_street\_1/2} \\
    Few or none empty & 0--55 \%       & \texttt{created\_date}, \texttt{complaint\_type}, \\
                      &              & \texttt{agency} \\
    \bottomrule
  \end{tabular}
\end{table}

Figure~\ref{fig:blank_fields} depicts the proportion of missing values 
across fields, illustrating their distribution into Mostly Empty, 
Partially Empty, and Few/None Empty categories.

\begin{figure}[tbp]
	\centering
  	\includegraphics[width=0.7\textwidth, height=0.33\textheight, keepaspectratio]{data_completeness_by_field_bar_chart.pdf}
	\caption{Number and percentage of missing entries for the
          311 SR data (2020--2024)}
	\label{fig:blank_fields}
\end{figure}

%-------------------------------------------------------------------------------
%	Sub-section: Validating Data for Compliance with Allowable Values
%-------------------------------------------------------------------------------

\subsection{Validating Data for Compliance with Allowable Values}
\label{subsec:domain}

\paragraph{Validating Data Type}
\label{par:validatingdatatype}
The 311~\textsc{SR} Data Dictionary, last updated in~2023, 
provides only limited information regarding the data type or class of each 
field, although most can be reasonably inferred. 
The NYC~Open~Data Portal supplements this documentation by indicating field 
types with icons: 
\textbf{T}{\scriptsize\textbf{T}} for text, 
\emph{\textbf{\#}} for numeric, 
\faCalendar\ for date or time fields, 
and \faMapMarker\ for geospatial data.

In most cases, these indicators are sufficient for validation. 
However, the \texttt{incident\_zip} field is labeled with the text emblem 
(\textbf{T}{\scriptsize\textbf{T}}), even though ZIP codes are better 
regarded as fixed--length, numeric but non-computational identifiers.

All other fields validated properly against their expected data types, as 
determined from the 311~\textsc{SR} Data Dictionary and 
the NYC~Open~Data Portal descriptions. 
More generally, analytic work requires that fields contain values that are 
valid, bounded, and compliant with a domain of known or legal values. 
For some fields, the 311~\textsc{SR} Data Dictionary provides explicit 
definitions; for others, valid domains must be inferred from public usage, 
additional research, or comparison with external datasets. All 
fields examined here were compliant with their known value domains.

The following fields were validated against known domains and found to be 
compliant. They can be grouped into three categories:

\begin{itemize}[left=1.5em]
  \item \textbf{Organizational fields:} 
        \texttt{agency}, \texttt{status}, 
        \texttt{open\_data\_channel\_type}, and \texttt{vehicle\_type}.
  \item \textbf{Geographic fields:} 
        \texttt{borough}, \texttt{park\_borough}, 
        \texttt{city\_council\_district}, and \texttt{taxi\_company\_borough}.
  \item \textbf{Address-related fields:} 
        \texttt{address\_type}.
\end{itemize}

\paragraph{Zip Codes}
\label{par:zipcodesissues}
Validation of the \texttt{incident\_zip} field is relatively straightforward, 
as the U.S.~Postal Service (\textsc{USPS}) maintains a definitive database 
of 42{,}735 valid ZIP~codes. 
Within the NYC~311 dataset, however, 3{,}274 entries were found to be invalid, 
including clearly erroneous values such as \texttt{10000}, \texttt{00083}, 
and \texttt{11111}. 
Although these represent only \SI{0.02}{\percent} of all records, 
the presence of invalid ZIP~codes is notable because this field is among the 
easiest to validate against an authoritative reference. 
Such anomalies highlight the importance of incorporating automated domain 
checks, particularly for fields governed by well-defined external standards.

\paragraph{Community Boards:}
\label{sec:communityboardissues}
The \texttt{community\_board} field captures one of 
New~York~City’s subdivisions of local governance. 
Validation should be straightforward, as there is a publicly 
available list of community boards. 
Nonetheless, this analysis identified 56{,}914 invalid entries, 
representing values outside the accepted set. 
Such anomalies underscore the risks that arise when categorical fields 
linked to well-defined administrative structures are not systematically 
validated.

%-------------------------------------------------------------------------------
%	Sub-section: Date Field Issues
%-------------------------------------------------------------------------------

\subsection{Date Field Issues:}
\label{subsec:datefieldissues}
In addition to the DST-related structural anomalies discussed in 
Section~\ref{par:dst}, the dataset exhibits a range of problems 
specific to its date fields. 
Dates should follow a logical temporal sequence—for example, the 
\texttt{created\_date} should precede the \texttt{closed\_date}. 
The NYC~311 dataset includes four such fields: 
\texttt{created\_date}, \texttt{due\_date}, \texttt{updated\_date}, 
and \texttt{closed\_date}, which are presumably populated in that order. 
In practice, however, numerous records violate these expected relationships, 
introducing inconsistencies that complicate both analysis and interpretation.

\subsubsection{Created Date }
\label{subsubsec:createddate}
All \texttt{created\_date} values were checked to ensure that none were set 
in the future or before the inception of the 311~system (circa~2003); all 
records were compliant. 
However, the analysis identified a notable clustering of timestamps at 
exactly midnight (14{,}409~cases) and exactly noon (31{,}872~cases), 
an anomaly that will be examined further in 
Section~\ref{subsubsec:midnightnoonissues}.

\subsubsection{Due Date Anomalies:}
\label{subsubsec:duedate}
The \texttt{due\_date} field is rarely populated, appearing almost 
exclusively in Department of Sanitation (\textsc{DSNY}) records and 
only about \SI{0.3}{\percent} of the time. 
All values passed validation checks (no dates in the past or implausible 
future), and no unusual temporal patterns were observed.

Sixteen cases were identified in which the \texttt{due\_date} precedes the 
\texttt{created\_date}---a nonsensical condition. These premature 
dates ranged from just over one year early to more than 
two years early. 
Any analysis incorporating such values would produce distorted results, 
particularly when evaluating timeliness against stated deadlines. 
Given that this field is rarely used (the third least populated in the 
dataset, at only \SI{0.3}{\percent} overall), such anomalies are unlikely 
to have a major impact on aggregate analyses but should nonetheless be 
excluded when encountered.

\subsubsection{Closed Date }
\label{subsubsec:closeddate}
There are substantial issues associated with the 
\texttt{closed\_date} field. 
All values were checked to ensure that none were set in the future or 
before the inception of the 311~system (circa~2003); all records were 
compliant in this respect. 
However, the analysis identified a notable clustering of timestamps at 
exactly midnight (14{,}409~cases) and exactly noon (31{,}872~cases), 
an anomaly that will be examined further 
in Section~\ref{subsubsec:midnightnoonissues}

\subsubsection{Resolution Action Updated Date Anomalies:}
\label{subsubsec:resolutionupdatedate}
The \texttt{resolution\_action\_updated\_date} field records the 
datetime when an \textsc{SR} is modified or updated. 
No values were found to lie in the implausible past or future. 
However, 376{,}155~entries (\SI{2.35}{\percent}) occur 
\emph{before} the corresponding \texttt{created\_date}. 
Nearly \SI{80}{\percent} of these early updates fall within a 
single day of creation—many within minutes or even seconds—suggesting 
a potential system-level issue in which the field is populated 
prematurely.

A more substantive irregularity arises in post-closure updates. 
While such updates are relatively common 
(\SI{54}{\percent} of all updates occur after the 
\texttt{closed\_date}), 27{,}810~records were found in which the update 
occurred more than 30~days—and in some cases up to 1{,}593~days 
(\textgreater4~years)—after closure. 
A Pareto analysis by agency indicates that over \SI{80}{\percent} of these 
late updates are concentrated within the 
Taxi~\&~Limousine Commission (\textsc{TLC}), 
Department of Parks~\&~Recreation (\textsc{DPR}), 
and Department of Health~\&~Mental~Hygiene (\textsc{DOHMH}). 
This clustering raises the question of whether such delays represent 
routine administrative practices or potential data quality issues.

Finally, 91~records exhibit \emph{extremely} late post-closure updates,
linked to invalid \texttt{closed\_date} values set to 1899 or~1900.
These cases likely stem from the well-documented SQL
default-date error in handling blank fields.
Such artifacts should be identified and excluded to prevent
severe distortions in any downstream analysis
(as illustrated in Figure~\ref{fig:extreme-positive-boxplot}).

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{boxplot_extreme_positive_days_by_agency.pdf}
  \caption{Boxplot of extremely large positive durations (in days) by agency.}
  \label{fig:extreme-positive-boxplot}
\end{figure}

\subsubsection{Daylight Saving Time (DST) Anomalies:}
\label{subsubsec:dst}
The American practice of observing Daylight Saving Time (DST) 
introduces systemic challenges in measuring 
\textsc{SR} durations. 
As noted in Section~\ref{par:dst}, the fall--back adjustment 
imparts nonsensical \emph{negative} durations in roughly 
86~\textsc{SR}s per year, when \texttt{closed\_date} 
precedes \texttt{created\_date}.

The spring--forward adjustment, by contrast, inflates 
\textsc{SR} durations. At 02{:}00~local time, clocks are 
advanced directly to 03{:}00, creating an artificial one-hour gap. 
Any \textsc{SR}s active (e.g., with status 
\texttt{OPEN}, \texttt{PENDING}, \texttt{IN~PROGRESS}, or 
\texttt{STARTED}) before the transition and subsequently closed 
appear one hour longer in duration than in reality. 
This analysis found that approximately 55{,}000~\textsc{SR}s 
annually are affected by this error.

Although discrepancies of $+1$~hour or $-42$~minutes may seem minor, 
they can meaningfully distort performance assessments for 
time-sensitive issues. Examples include 
\texttt{Noise--Residential}, \texttt{Animal--Abuse}, 
\texttt{Homeless~Person~Assistance}, and 
\texttt{Drug~Activity}, where response time is a critical measure 
of agency effectiveness. 
These systemic anomalies underscore the need to explicitly account 
for DST when analyzing \textsc{SR} durations.

\subsubsection{Midnight and Noon: Creation and Closure Issues}
\label{subsubsec:midnightnoonissues}
Overall, 311~\textsc{SR} activity follows a typical
daily cycle, with most requests created between 08{:}00 and 23{:}00
(see Figure~\ref{fig:created-hour-dist}).
Closer inspection of the \texttt{created\_date} and
\texttt{closed\_date} fields, however, reveals a distinct temporal
artifact in which unusually large numbers of records occur precisely at
midnight (00{:}00{:}00) and noon (12{:}00{:}00). Such perfect alignment is highly
unlikely to result from human activity and instead suggests that
automated system routines are creating or closing batches of
\textsc{SR}s at these times. While this behavior may not represent an
error per~se, it reflects a systemic process artifact that can distort
analyses of hourly workload, duration distributions, and agency
performance metrics.

As shown in Figures~\ref{fig:exacthours}\subref*{fig:created_top_hour}~and~%
\subref*{fig:closed_top_hour}, both the creation and closure events display
a pronounced concentration at exactly midnight and noon, far exceeding the
expected random frequency and well above \( +3\sigma. \)

\begin{figure}[tbp]
  \centering
  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\textwidth]{created_top_of_hour_distribution.pdf}
    \caption{SRs created at the top of the hour (hh{:}00{:}00).}
    \label{fig:created_top_hour}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\textwidth]{closed_top_of_hour_distribution.pdf}
    \caption{SRs closed at the top of the hour (hh{:}00{:}00).}
    \label{fig:closed_top_hour}
  \end{subfigure}
  \caption{311~(\textsc{SR}s) created and closed exactly
  at the top of the hour (hh{:}00{:}00).}
  \label{fig:exacthours}
\end{figure}

 \subsection{Duration Issues}
\label{subsec:duration}
The life span of an \textsc{SR}—its 
\textit{duration}—is widely used as a performance metric for 
New~York~City (\textsc{NYC}) agencies. 
Duration is frequently employed to assess equity in service provision 
across neighborhoods, communities, and boroughs, and is often cited in 
City reports as a proxy for agency responsiveness. 
For these reasons, \textsc{SR} duration warrants careful analysis.

In principle, duration is the elapsed time between the 
\texttt{created\_date} and \texttt{closed\_date} fields. 
Given the prevalence of automated systems, one might expect these fields 
to be populated directly by application software in response to user actions, 
thereby minimizing human error and enforcing basic temporal logic. 
However, this is not consistently the case. 
The dataset contains several anomalies in the 
\texttt{created\_date} and \texttt{closed\_date} fields that would 
ordinarily be prevented by well-designed automation. 
These anomalies raise concerns about the reliability of duration as a 
performance indicator and underscore the need for closer scrutiny.

\subsubsection{Negative Durations:}
\label{subsubsec:negativedurations}
A total of 45{,}296~service requests (\textsc{SR}s) exhibit a 
\texttt{closed\_date} that precedes the \texttt{created\_date}, 
producing nonsensical negative durations. 
More than \SI{98}{\percent} of these anomalies originate from a single 
agency—the Department of Transportation (\textsc{DOT}). 
Across all affected \textsc{SR}s, negative durations average $-119$~days, 
a nontrivial distortion.

\begin{figure}[tbp]
    \centering
    \includegraphics[width=\textwidth, height=0.33\textheight, keepaspectratio]{boxplot_negative_days_by_agency.pdf}
    \caption{311~SRs with negative durations (days) by agency. 
    \textit{Note: pseudo-log scale.}}
    \label{fig:negativedurations}
\end{figure}

Within this group, 115~\textsc{SR}s show extreme negative durations due to 
\texttt{closed\_date} values in the distant past, specifically 
\texttt{1899-12-31} and \texttt{1900-01-01}. 
These records imply durations exceeding $-122$~years. 
Although rare, such outliers can dramatically skew statistical summaries; 
for example, mean calculations become unreliable if not adjusted. 
{\small\emph{(Note: These extreme values likely stem from the well-documented 
SQL default-date anomaly, in which blank or null date fields are coerced to 
the placeholder value \texttt{1900-01-01}. A similar behavior occurs 
in Microsoft~Excel., whose internal date system also begins on January~1,~1900.)}}

Table~\ref{tab:largest-errors} provides a summary of 
\textsc{SR}s with negative durations (in hours). 
Consistent with the broader distribution, the overwhelming majority of 
these cases originate within the Department of Transportation, which 
accounts for approximately \SI{98}{\percent} of all observed errors. 
This concentration reinforces the conclusion that the negative-duration 
anomaly is not evenly distributed across agencies, but rather reflects 
systemic or procedural issues specific to \textsc{DOT}.

\begin{table}[tbp]
    \centering
    \caption{Sample negative durations [excluding extreme values that 
    are lower than $-730$~days]}
    \label{tab:largest-errors}
    \small
    \begin{tabular}{l l r l}
        \toprule
        \texttt{created\_date} & \texttt{closed\_date} & 
        {Duration (hours)} & {Agency} \\
        \midrule
        2024-02-22 13:49:00 & 2023-05-30 13:48:00 & $-6433.02$ & DOT \\
        2020-05-18 15:24:00 & 2020-05-07 15:24:00 & $-264.00$  & DOT \\
        2022-01-10 14:20:00 & 2022-01-06 14:20:00 & $-96.00$   & DOT \\
        2020-05-01 11:33:00 & 2020-04-28 11:33:00 & $-72.00$   & DOT \\
        2021-06-01 11:04:00 & 2021-05-31 10:55:00 & $-24.15$   & DOT \\
        \bottomrule
    \end{tabular}
\end{table}

 \subsubsection{Zero and 1-second Durations:}
\label{subsubsec:zerodurations}
A more prevalent anomaly arises when the \texttt{closed\_date} and 
\texttt{created\_date} share the exact same timestamp to the second. 
This results in a \emph{zero-duration} \textsc{SR}, 
which is inherently nonsensical. 
A total of 387{,}595~zero-duration \textsc{SR}s were identified, 
representing \SI{2.5}{\percent} of all non-blank records. 
Notably, \SI{98}{\percent} of these cases originate from just five agencies—%
the Department of Health~and~Mental~Hygiene (\textsc{DOHMH}), 
Transportation (\textsc{DOT}), Buildings (\textsc{DOB}), 
Environmental~Protection (\textsc{DEP}), and 
Sanitation (\textsc{DSNY}). 
This concentration does not mirror the overall \textsc{SR} distribution 
by agency, suggesting agency-specific data entry or system-level issues.

In addition, another 3{,}460~\textsc{SR}s are recorded as closing within 
one second of their creation. 
While not strictly illogical like negative durations, such 
near-instantaneous closures are highly improbable and represent a 
suspicious pattern, meriting further investigation.

\subsubsection{Suspicious Durations}
\label{subsubsec:suspiciousdurations}
\textbf{Rationale.}  
Service Request (\textsc{SR}) durations are highly right-skewed, 
with some extending months or years. 
Such extremes distort summary measures and hinder comparisons across agencies, 
time periods, or complaint types. 
To ensure interpretable results, a screening procedure was applied 
to identify these \textit{suspicious} durations.

\textbf{Methodology.}  
Durations were truncated to plausible limits: 
2~seconds (excluding zero- and one-second artifacts) to 
172{,}800~seconds (48~hours). 
This reduced the dataset from 15.3~to~10.0~million records 
(\SI{64.9}{\percent} retained) and greatly reduced the long tail. 
Within this subset, five complementary outlier tests were evaluated—
Standard Deviation (SD), Median Absolute Deviation (MAD), 
log-normal thresholds, percentile cutoffs (1–10\%), 
and Interquartile Range (IQR).

\textbf{Findings.}  
Truncation lowered the mean-to-median ratio from~53.7~to~5.2. 
SD, MAD, and IQR tests flagged no further anomalies, 
while log-normal and percentile approaches produced 
operationally reasonable limits. 
The adopted rule—log-normal three–SD cutoff 
($\approx$\,28~seconds)—balanced rigor and interpretability, 
identifying only \SI{0.07}{\percent} of records as outliers. 
This threshold provides a defensible boundary 
between routine and atypical \textsc{SR}s.

\textbf{Summary.}  
A total of 405{,}796~\textsc{SR}s—including 
14{,}741~newly flagged cases—exhibited statistically suspicious or 
operationally implausible durations.

\subsubsection{Positive Durations:}
\label{subsubsec:positivedurations}
While a positive duration is not inherently erroneous, a subset of 
service requests (\textsc{SR}s) exhibit unusually long lifespans. 
five~years (1{,}826~days) were designated as \emph{extreme}.

A total of 138{,}318~\textsc{SR}s fall into the \emph{large} category (1--2~years). 
These cases are concentrated in three agencies—the Department of 
Parks~and~Recreation (\textsc{DPR}), the Department of 
Health~and~Mental~Hygiene (\textsc{DOHMH}), and the 
Economic~Development~Corporation (\textsc{EDC})—which together account for 
\SI{94}{\percent} of all large-duration \textsc{SR}s. 
The average length of these cases was 1{,}099~days.

At the more severe end, 1{,}211~\textsc{SR}s qualify as 
\emph{extreme} positive durations ($>$\,5~years). 
Nearly all of these (\SI{94}{\percent}) occurred within \textsc{DPR}. 
The average duration in this group was 1{,}875~days, 
with the maximum recorded at 2{,}009~days.

\begin{figure}[tbp]
    \centering
    \includegraphics[width=\textwidth, height=0.33\textheight, keepaspectratio]{boxplot_extreme_positive_days_by_agency.pdf}
    \caption{311~service requests (\textsc{SR}s) with extremely large 
    positive durations.}
    \label{fig:extremepositive}
\end{figure}

\subsection{Duplicate \& Redundant Fields}
\label{subsec:redundant}
This analysis revealed several redundant fields, indicating a potential 
need for consolidation and simplification.

\paragraph{\texttt{latitude/longitude} vs.~\texttt{location}}  
The \texttt{location} field is merely a concatenation of the 
\texttt{latitude} and \texttt{longitude} values, enclosed in parentheses 
and separated by a comma. Functionally, it adds no new information yet introduces unnecessary parsing 
complexity.

While this duplication is relatively benign, it inflates dataset size 
and complicates downstream processing. 
Consolidating these fields—or removing \texttt{location} in favor of the 
atomic \texttt{latitude} and \texttt{longitude} columns—would simplify 
analysis, reduce redundancy, and improve structural coherence.

\paragraph{\texttt{borough} vs.~\texttt{park\_borough}}  
These two fields are entirely redundant, with values matching 
\SI{100}{\percent} across all records. Retaining both 
only increases duplication and introduces potential 
confusion for users.

\paragraph{\texttt{borough} vs.~\texttt{taxi\_company\_borough}}  
Despite their similar names, these fields capture almost entirely distinct 
information, sharing only \SI{0.29}{\percent} of overlapping values. 
The \texttt{taxi\_company\_borough} field appears to be used exclusively 
by the Taxi~and~Limousine~Commission (\textsc{TLC}), suggesting that it 
serves a specialized purpose distinct from the general 
\texttt{borough} designation. This example illustrates 
how field naming conventions can obscure 
specialized agency usage, complicating cross-agency data harmonization.

\paragraph{\texttt{agency} vs.~\texttt{agency\_name}}  
The \texttt{agency} field stores abbreviated identifiers for City agencies 
(e.g.,~\textsc{NYPD}, \textsc{DOT}, \textsc{DSNY}), 
while \texttt{agency\_name} provides their full titles. 
Because the abbreviations are widely recognized and sufficient for most analyses, 
maintaining both fields is, for the most part, redundant. 
Retaining only the abbreviated form would reduce file size and improve efficiency 
without diminishing interpretability.

\paragraph{\texttt{landmark} vs.~\texttt{street\_name}}  
According to the 311~Service~Request (\textsc{SR}) Data Dictionary, 
the \texttt{landmark} field is intended to denote noteworthy sites such as 
parks, hospitals, airports, sports facilities, or performance venues. 
In practice, however, the majority of entries are street names, with 
approximately \SI{60}{\percent} matching the \texttt{street\_name} field. 
Many of the remaining non-matches (excluding blanks) appear to be 
near-equivalents, differing only in spelling or formatting 
(e.g.,~``NINTH~AVE'' vs.~``9~AVE''). 
This pattern suggests frequent misuse of the \texttt{landmark} field and 
underscores its functional redundancy with \texttt{street\_name}.

\paragraph{\texttt{cross\_street\_1/2} vs.~\texttt{intersection\_street\_1/2}}  
These two street-pair fields are intended to help identify the 
\texttt{incident\_address}. 
Analysis revealed that approximately \SI{84}{\percent} of entries are direct 
duplicates across the two pairs, raising the question of which set should be 
considered authoritative. If the fields are used differently across agencies, 
the lack of documentation leaves their relationship unclear. 
Regardless, the high level of duplication indicates substantial redundancy.

%-------------------------------------------------------------------------------
%	Sub-section: Reducing File Size
%-------------------------------------------------------------------------------

\subsection{Reducing File Size}
\label{subsec:filesize}
Eliminating duplicate and near-duplicate fields could reduce the 
dataset size by approximately \SI{39.5}{\percent}—a substantial gain. 
The analysis revealed that the following fields could be safely 
removed with minimal loss of information:

\begin{itemize}[left=1.5em]
  \item \texttt{agency\_name}: Redundant with \texttt{agency}, which already uses 
        widely recognized abbreviations.  
  \item \texttt{park\_borough}: Identical to \texttt{borough} and therefore unnecessary.  
  \item \texttt{location}: A concatenation of \texttt{latitude} and \texttt{longitude}, 
        offering no additional analytical value.  
  \item \texttt{cross\_street\_1/2} and \texttt{intersection\_street\_1/2}: 
        With an \SI{84}{\percent} match rate, these fields are highly redundant. 
        Deleting \texttt{intersection\_street\_1/2} is recommended, though this may 
        entail limited information loss.  
\end{itemize}

In addition, certain fields are of limited analytical value because they are sparsely 
populated (\textgreater\SI{99}{\percent} blank). 
These include \texttt{taxi\_company\_borough}, \texttt{road\_ramp}, \texttt{vehicle\_type}, 
\texttt{due\_date}, \texttt{bridge\_highway\_direction}, \texttt{bridge\_highway\_name}, 
\texttt{bridge\_highway\_segment}, and \texttt{taxi\_pick\_up\_location}.  
Although sparse overall, even a \SI{1}{\percent} fill rate corresponds to 
roughly 160{,}000 records in this 16~million-row dataset, suggesting that 
agency-specific analyses may still benefit from these fields. 
A practical approach would be to store such variables in separate, 
agency-level tables.

\paragraph{Field Encoding.}  
Further size reductions can be achieved through categorical encoding. 
For example, rather than storing the text string 
``\texttt{NOISE~--~STREET/SIDEWALK}'' more than 300{,}000~times, 
a numeric or symbolic code could be used instead. 
This approach reduces storage overhead, accelerates processing, and 
improves data quality by eliminating misspellings and other invalid entries.

\paragraph{Alternative Storage Formats.}  
Beyond structural adjustments, storage efficiency can be substantially improved 
through the adoption of modern columnar data formats. 
Apache~Arrow, a cross-platform, in-memory data framework, enables efficient analytics 
and real-time processing while supporting interoperability across programming languages 
such as~R, Python, and Julia. 
Replacing CSV files with Arrow-based storage can reduce dataset size by orders of magnitude 
while simultaneously improving query performance and computational throughput 
\citep{bates2024csv}.
The integration of columnar storage into open-data publishing pipelines would markedly 
enhance accessibility for analysts while reducing both storage and transmission costs.

%-------------------------------------------------------------------------------
%	Section: Principles
%-------------------------------------------------------------------------------

\section{Principles for Government Open Data Curation}
\label{sec:principles}

Building on the insights gained from the case study of the 
\textsc{NYC 311 SR} dataset, this section articulates a set 
of guiding principles for improving the curation and management of 
government open datasets. 
Derived from the challenges identified in the 311~case study, 
these principles are broadly applicable to other large-scale data 
environments. 
Each principle aims to enhance reliability, efficiency, and usability, 
while promoting transparency and public trust in open-data initiatives. 
Together, they provide a coherent framework for sustainable, 
high-quality open-data governance across both organizational 
and technical boundaries.

\subsection{Principle~1: Establish Cross-Organizational Consistency}
\label{subsec:principle1}

A persistent challenge in government open data is the lack of harmonization 
among agencies in field definitions, formats, and value domains.  
Such inconsistencies distort analyses and impede cross-agency integration.  
To address this, standardized conventions for field naming, categorical 
domains, and formatting must be defined, applied uniformly, and audited 
regularly to ensure continuity as reporting practices evolve.

Equally important is accountability.  
Currently, no entity oversees the quality of NYC’s~311~\textsc{SR} data across 
agencies; each submits independently, allowing anomalies to persist.  
A coordinating governance function—central or distributed—should define, 
monitor, and enforce quality standards.  
Without such oversight, even advanced validation tools risk underuse, 
undermining the credibility and consistency of open data.

\subsection{Principle~2: Ensure Data Entry Accuracy}
\label{subsec:principle2}
Data accuracy and validity are prerequisites for any useful open dataset.  
In the \textsc{NYC~311} case, anomalies such as invalid ZIP codes and 
illogical dates highlight the need for validation at the point of entry.  
Fields should accept only values drawn from authorized domains, with 
software safeguards minimizing human error.  
Key measures include:

\begin{itemize}[left=1.5em]
  \item Auto-generating critical date fields from system events (e.g., status changes);  
  \item Blocking records with implausible sequences, such as a closing date preceding creation;  
  \item Validating structured fields (e.g., agency or ZIP codes) against authoritative references.  
\end{itemize}

Embedding such checks directly in submission workflows reduces systemic 
errors in high-volume datasets and turns quality assurance from a reactive 
activity into a routine operational safeguard.

\subsection{Principle~3: Optimize Storage Efficiency}
\label{subsec:principle3}
Efficient storage and representation are vital for large-scale government 
datasets.  
In the \textsc{NYC~311} case, duplicate or near-duplicate fields inflated 
file size without adding analytic value; removing them could reduce storage 
needs by nearly 40\% while maintaining utility.

Encoding categorical variables in standardized or numeric formats further 
reduces space, prevents inconsistencies, and simplifies analysis.  
Such optimization lowers storage and transmission costs and improves the 
speed, reliability, and reproducibility of workflows—especially important 
when datasets are widely shared for public use.

\subsection{Principle~4: Maintain and Update Data Dictionaries}
\label{subsec:principle4}
Accurate metadata are vital to the usability of government datasets.  
In the \textsc{NYC~311} case, the Data Dictionary contained undocumented 
fields and outdated domain definitions, creating ambiguity and complicating 
analysis.

Data Dictionaries should specify data types, valid value domains, and clear 
field definitions, and be reviewed regularly to reflect structural or 
semantic changes.  
Consistent maintenance enables researchers, policymakers, and developers to 
interpret data correctly and supports transparency, reproducibility, and 
trust in open-data resources.

\subsection{Principle~5: Automate Quality Assurance Processes}
\label{subsec:principle5}
Automated validation and quality assurance markedly enhance the reliability 
of government datasets by detecting anomalies in real time.  
In the \textsc{NYC~311} case, spikes in request creation and closure at 
midnight or noon likely reflect batch processes that distort duration 
measures.  Real-time, event-driven validation can prevent such errors at the 
point of entry and surface irregular patterns as they occur.

Automated routines should flag inconsistencies, illogical sequences, and 
implausible values.  Increasingly, artificial-intelligence (AI) methods—%
including anomaly-detection algorithms for numeric data and natural-language 
models for text fields—complement rule-based checks by identifying subtle or 
rare anomalies that manual review might miss.  
Beyond operational benefits, AI-assisted monitoring strengthens the 
foundation for downstream analytics and machine-learning applications, 
ensuring that training data remain fair, accurate, and actionable.

\subsection{Principle~6: Ensure Data Transparency}
\label{subsec:principle6}
Transparency is a core tenet of open data, requiring that datasets be 
not only accurate but also accessible and comprehensible to a wide 
range of users. This entails publishing clear and complete metadata 
that describe dataset structure, limitations, and intended uses.  
Equally important is open communication about data-quality issues—%
for example, notifying users when systemic anomalies (such as erroneous 
\texttt{closed\_date} values) are identified and explaining how such 
issues will be addressed.

Proactive transparency fosters trust, promotes accountability, and 
encourages broader civic and research use of datasets.  
By making quality-related information public, agencies transform open 
data from a static disclosure into a participatory process, one in 
which users, analysts, and policymakers collectively contribute to 
continuous improvement and shared understanding.

\subsection{Principle~7: Foster Interoperability and Scalable Access}
\label{subsec:principle7}
Government open data should be distributed in formats that balance public 
accessibility with computational efficiency.  
While flat \texttt{CSV} files remain the universal standard, they become 
inefficient as datasets scale to tens of millions of rows.

To improve scalability, agencies should offer complementary access options—%
such as structured APIs and columnar formats like Apache Parquet or Arrow—%
while retaining \texttt{CSV} for compatibility.  
Because most agencies already manage data in enterprise systems capable of 
exporting multiple formats, leveraging these infrastructures can serve both 
casual users and high-volume analysts without duplicating effort or risking 
inconsistency.

Interoperability across formats and platforms ensures that open data remain 
accessible, efficient, and sustainable for long-term public use.

\subsection{Principle~8: Embed Ethical and Privacy Safeguards}
\label{subsec:principle8}

Open-data publication must balance transparency with the protection of 
individual and community privacy.  
Even without direct identifiers, modern data-linkage techniques can 
re-identify individuals by combining open data with other public sources—%
a risk demonstrated in recent studies and government cases.

To reduce such risks, agencies should apply disclosure-avoidance methods 
such as aggregation, suppression, masking, or differential privacy, and 
conduct ethical reviews for releases containing detailed geospatial or 
temporal information linked to sensitive service categories.

Integrating privacy safeguards into curation workflows keeps open-data 
programs ethically responsible, legally compliant, and publicly trusted.  
Transparency and privacy are not opposing values but complementary 
pillars of responsible data governance.

\subsection{Synthesis, Outlook, and Actionable Steps}
\label{subsec:actions}
Together, these principles outline a framework for sustaining the reliability 
and integrity of government open data across the full lifecycle of stewardship—%
from harmonization and validation at entry to automation, transparency, and 
ethical accountability.  Although grounded in the \textsc{NYC~311 SR} case, the 
lessons apply broadly wherever open data support public accountability and 
civic innovation.

To move from principle to practice, governments should prioritize 
real-time validation, regular updates to data dictionaries, and standardized 
encoding and storage formats.  Automated checks can flag errors such as 
invalid ZIP codes, detect anomalies as they occur, and reduce reliance on 
manual review.  These measures both improve quality and free staff for 
higher-value analytical work.

Cross-agency and external collaboration is equally vital.  Shared standards, 
metadata, and feedback channels allow issues to be identified and resolved 
transparently while aligning datasets with user needs.  Quality-assurance 
dashboards can continuously monitor key indicators (e.g., missing coordinates 
or invalid codes) and trigger corrective action, reinforcing accountability 
and trust.

Partnerships with data scientists, statisticians, and civic technologists can 
further enhance scalability and methodological rigor, ensuring that validation 
processes remain adaptive and future-proof.  Public-facing initiatives such as 
\textsc{NYC~Open~Data~Week} complement these efforts by promoting data 
literacy, fostering civic engagement, and maintaining feedback loops between 
agencies and the communities they serve.

By embedding these technical and participatory mechanisms, governments can 
transform open-data programs from static repositories into dynamic, trusted 
infrastructures that evolve in quality, transparency, and public value.

%-------------------------------------------------------------------------------
%	Section: Conclusion
%-------------------------------------------------------------------------------

\section{Conclusion}
\label{sec:conclusion}
This study highlights the essential role of systematic anomaly detection and 
robust curation in ensuring the reliability of government open data, using 
New York City’s \textsc{311 SR} dataset as a case study. 
The analysis revealed diverse anomalies—erroneous or missing dates, redundant 
fields, and implausible durations—that, though often subtle, can distort 
results and erode confidence in downstream analyses.

From these findings, we outlined practical principles for improving open-data 
quality: cross-agency consistency, rigorous validation, efficient storage, 
updated documentation, automated quality checks, and transparency. 
Together, these measures provide a roadmap for maintaining the integrity and 
utility of large public datasets.

Although centered on \textsc{NYC 311}, the implications extend to other 
domains such as health, transportation, and environmental monitoring, where 
data inconsistency and incompleteness persist. 
Sound curation is particularly critical for machine-learning applications, 
where biased inputs can compromise accuracy and fairness 
\citep{rahm2000data,geiger2020garbage}. 
The COVID-19 pandemic further underscored the societal need for accurate, 
timely, and trustworthy data 
\citep{worby2020face,khemasuwan2021applications}.

Looking forward, automated validation pipelines and feedback mechanisms will 
be key to sustaining quality at scale. 
Embedding these processes into technical and organizational workflows will 
help transform administrative records into reliable public assets that 
support transparent governance, sound research, and civic innovation.

Ultimately, well-curated data underpin both analytical rigor and public 
trust—without them, the promise of open data cannot be fully realized.

%-------------------------------------------------------------------------------
%	Section: Supplementary Material
%-------------------------------------------------------------------------------

\section*{Supplementary Material}
Extensive supplementary analyses and charts are provided to support 
and extend the findings presented in this paper.  
The full \textsc{R} source code used for dataset prepraration and anomaly detection, 
along with instructions on how to duplicate this analysis—is available at 
\url{https://github.com/jun-yan/nyc311clean}.  
Readers are encouraged to explore this repository for reproducible examples 
and implementation details that complement the discussion in the main text.

Reference data files, including the 2022--2023 NYC~311~Service Request 
records and supporting USPS ZIP code tables, are available at 
\url{https://doi.org/10.6084/m9.figshare.c.7617311}.  
Together, these materials provide a comprehensive foundation for replicating, 
validating, and extending the analyses described in this study.

\bibliographystyle{jds}
\bibliography{refs}

\end{document}

### 2025-10-22 Journal_of_Data_Science/code/datacleansing_for_jds.R

@@ -1,0 +1,2041 @@
+######################################################################### 
+######################################################################### 
+
+main_data_file <-  
+  "5-year_311SR_01-01-2020_thru_12-31-2024_AS_OF_10-10-2025.rds"
+ 
+# Boolean flag. TRUE to redirect console output to text file
+# FALSE to display console output on the screen
+enable_sink <- FALSE    
+
+#The "as of" date in "YYYY-MM-DD" format
+projection_date <- "2025-10-20"   
+
+#Number of SRs for the year through the projection_date  
+projection_SR_count <- 2779203  
+
+# Okabe-Ito palette for colorblind safe 
+palette(c("#E69F00", "#56B4E9", "#009E73", "#F0E442", 
+          "#0072B2", "#D55E00", "#CC79A7", "#999999"))
+
+#########################################################################
+# ------------------------------------------------------ -------
+# 📦 INSTALL AND LOAD REQUIRED PACKAGES
+# -------------------------------------------------------------
+load_required_packages <- function(verbose = TRUE) {
+  # Ordered to avoid common masking issues
+  required_packages <- c(
+    "data.table",      # Load first - commonly masked functions
+    "arrow",
+    "fasttime",
+    "lubridate",
+    "here",
+    "zoo",
+    "ggplot2",         # Core plotting (scales is a dependency, auto-loaded)
+    "ggpmisc",
+    "ggpattern",
+    "ggrastr",
+    "qcc",
+    "qicharts2",
+    "grid",
+    "gridExtra",
+    "sf",              # Spatial - can mask some dplyr functions
+    "stringr",         # Load before dplyr/tidyverse
+    "stringdist",
+    "dplyr",           # Load before tidyverse
+    "tidyverse",       # Load late - masks many functions
+    "scales",          # Load AFTER tidyverse to mask purrr::discard and readr::col_factor
+    "bslib",
+    "shiny",
+    "DT",              # Load after dplyr to avoid confusion
+    "gt",
+    "styler",
+    "rlang",
+    "renv",
+    "remotes"
+  )
+  
+  for (pkg in required_packages) {
+    if (!requireNamespace(pkg, quietly = TRUE)) {
+      if (verbose) message(sprintf("📦 Installing missing package: %s", pkg))
+      tryCatch(
+        install.packages(pkg),
[…]

### 2025-10-22 Journal_of_Data_Science/code/jds_data_quality_over_time.R

@@ -4,63 +4,102 @@
 
 ################################################################################
 # Set the base directory under the working directory base_dir <- getwd()
-working_dir <- getwd()
+base_dir <- file.path(
+  "C:",
+  "Users",
+  "David",
+  "OneDrive",
+  "Documents", 
+  "datacleaningproject", 
+  "nyc311clean",
+  "Journal_of_Data_Science"
+)
+cat("Running in STANDALONE mode\n")
+cat("Base directory:", base_dir, "\n")
+#}
 
-base_dir <- file.path(working_dir, "datacleaningproject", "nyc311clean",  
-                      "data_quality_over_time" )
+# Define all paths relative to base_dir (works in both modes)
+data_dir      <- file.path(base_dir, "data")
+chart_dir     <- file.path(base_dir, "charts", "quality")
+functions_dir <- file.path(base_dir, "code", "functions")
+console_dir   <- file.path(base_dir, "console_output")
+write_dir     <- file.path(base_dir, "data")
 
-# Define the path for the main data file (CSV file)
-data_dir <- file.path(base_dir, "data")
- 
-# Define the path for the charts
-chart_dir <- file.path(base_dir, "charts")
+# Create directories if they don't exist
+dirs_to_create <- c(data_dir, chart_dir, console_dir)
+for (dir in dirs_to_create) {
+  if (!dir.exists(dir)) {
+    dir.create(dir, recursive = TRUE)
+    cat("Created directory:", dir, "\n")
+  }
+}
 
-# Define the path to the directory containing your function scripts
-functions_path <- file.path(base_dir, "code", "functions")
+cat("\nDirectory paths set:\n")
+cat("  Data:", data_dir, "\n")
+cat("  Charts:", chart_dir, "\n")
+cat("  Functions:", functions_dir, "\n")
+cat("  Console output:", console_dir, "\n")
+cat("  Write:", write_dir, "\n")
 
+# Okabe-Ito palette for colorblind safe
+palette(c("#E69F00", "#56B4E9", "#009E73", "#F0E442", 
+          "#0072B2", "#D55E00", "#CC79A7", "#999999"))
+
+################################################################################
+# Get all .R files in the "functions" sub-directory
 ######### Commence directing console output to the file ##########
[…]

### 2025-10-22 code/datacleansing.R

@@ -41,7 +41,7 @@ setwd("C:/Users/David/OneDrive/Documents/datacleaningproject/nyc311clean/code")
 #########################################################################
 rm(list = ls())
 
-main_data_file <- "3-month_311SR_10-01-2024_thru_12-31-2024_AS_OF_02-02-2025.csv"
+main_data_file <- "5-year_311SR_01-01-2020_thru_12-31-2024_AS_OF_09-23-2025.csv"
 
 # Extract the 10-character date after "AS_OF_"
 max_closed_date <- sub(".*AS_OF_([0-9-]+)\\.csv$", "\\1", main_data_file)

### 2025-10-22 code/functions/base_bar_chart.R

@@ -16,17 +16,17 @@ base_bar_chart <- function(
     horizontal_adjustment_max = 0.5,
     vertical_adjustment_max = -1,
     console_print_out_title = "Data Summary",
-    rows_to_print = 20,  # New parameter for number of rows to print
+    rows_to_print = 20,
     chart_file_name = NULL,
     chart_directory = ".",
     chart_width = 10,
     chart_height = 7,
     annotation_size = 5,
-    x_axis_tick_size = 12,  # Size of x-axis ticks
-    x_axis_label_angle = 90,  # Angle of x-axis labels
-    x_axis_tick_length = unit(0.3, "cm")  # Length of x-axis tick marks
+    x_axis_tick_size = 12,
+    x_axis_label_angle = 90,
+    x_axis_tick_length = unit(0.3, "cm"),
+    x_label_every = 1  # NEW PARAMETER: show every nth x-axis label
 ) {
-  
   
   # Step 0: Add percentage and cumulative percentage columns with 2 decimal points
   dataset <- dataset %>%
@@ -42,8 +42,8 @@ base_bar_chart <- function(
   dataset <- as.data.frame(dataset)
   
   # Temporarily increase max.print to handle larger output
-  old_max_print <- getOption("max.print")  # Save the current max.print value
-  options(max.print = max(1000, rows_to_print * 4))  # Adjust based on expected rows and columns
+  old_max_print <- getOption("max.print")
+  options(max.print = max(1000, rows_to_print * 4))
   
   # Print the first 'rows_to_print' rows of all relevant columns
   print(dataset[1:rows_to_print, c(x_col, y_col, "percentage", "cumulative_percentage")], row.names = FALSE)
@@ -64,23 +64,44 @@ base_bar_chart <- function(
     sorted_dataset <- dataset[order(-dataset[[y_col]]), ]
   }
   
-  # Step 3: Determine x_scale based on x_col type
-  x_scale <- if (inherits(dataset[[x_col]], "Date")) {
-    scale_x_date(
+  # Step 3: Determine x_scale based on x_col type with label spacing
+  if (inherits(dataset[[x_col]], "Date")) {
+    # For dates, use date_breaks with multiplier
+    break_interval <- paste(x_label_every, "days")
+    x_scale <- scale_x_date(
       expand = c(0.01, 0), 
-      labels = scales::date_format("%Y-%m"), 
-      breaks = scales::date_breaks("1 months")
[…]

### 2025-10-22 code/timelinecharts.R

@@ -45,7 +45,8 @@ formattedStartTime <- format(programStart, "%Y-%m-%d %H:%M:%S")
 
 #########################################################################
 # Set path for the data file
-main_data_file <- "311_Service_Requests_from_2022-2023_AS_OF_09-15-2024.csv"
+main_data_file <- "5-year_311SR_01-01-2020_thru_12-31-2024_AS_OF_09-23-2025.csv"
+#main_data_file <- "311_Service_Requests_from_2022-2023_AS_OF_09-15-2024.csv"
 #main_data_file <- "smaller_test_data.csv"
 #main_data_file <- "extra_small.csv"
 
@@ -62,7 +63,7 @@ base_dir <- working_dir
 data_file <- file.path(base_dir, "data")
 
 # Define the path for the charts
-chart_directory_path <- file.path(base_dir, "charts")
+chart_directory_path <- file.path(base_dir, "charts", "5-year_charts")
 
 # Create the directory for the reduced size file following shrinkage code.
 writeFilePath <- file.path(base_dir, "data")
@@ -133,8 +134,8 @@ d311[columns_to_upper] <- lapply(d311[columns_to_upper], toupper)
 d311 <- consolidate_agencies((d311))
 
 # Convert each date field to POSIXct format in UTC
-d311$created_date <- as.POSIXct(d311$created_date, format = "%m/%d/%Y %I:%M:%S %p", tz = "UTC")
-d311$closed_date <- as.POSIXct(d311$closed_date, format = "%m/%d/%Y %I:%M:%S %p", tz = "UTC")
+d311$created_date <- as.POSIXct(d311$created_date, format = "%m/%d/%Y %I:%M:%S %p", tz = "America/New_York")
+d311$closed_date <- as.POSIXct(d311$closed_date, format = "%m/%d/%Y %I:%M:%S %p", tz = "America/New_York")
 
 # Call the function on your dataframe d311
 date_columns <- c("created_date", "closed_date")
@@ -148,8 +149,8 @@ num_years <- unique(years)
 
 cat("\nTotal rows:", format(num_rows, big.mark = ","), "covering", length(num_years), "years")
 
-year_digits <- 2
-file_name_prefix <- "2-year"
+year_digits <- 5
+file_name_prefix <- "5-year"
 #########################################################################
 # Calculate the earliest and latest dates directly
 earliest_date <- min(d311$created_date, na.rm = TRUE)
@@ -408,6 +409,7 @@ extra_line <- annotate("text",
 
 SR_monthly <- base_bar_chart(
[…]

### 2025-10-22 data_quality_over_time/code/functions/standardize_missing_chars.R

@@ -1,0 +1,56 @@
+standardize_missing_chars <- function(DT,
+                                      tokens = c("na","<na>","n/a","null"),
+                                      report_zeroes = FALSE) {
+  stopifnot(is.data.table(DT))
+  
+  labels <- c("<blank>", tokens)                 # what we will count
+  token_counts <- setNames(integer(length(labels)), labels)
+  total_to_na <- 0L
+  
+  cat("\n[ Normalizing missing-like values (character) -> NA ]\n")
+  
+  char_cols <- names(DT)[vapply(DT, is.character, logical(1))]
+  for (col in char_cols) {
+    x <- DT[[col]]
+    x_trim <- trimws(x)
+    
+    # Key used for matching tokens (exact, after trim); NA stays NA
+    keys <- ifelse(is.na(x_trim), NA_character_,
+                   ifelse(x_trim == "", "<blank>", tolower(x_trim)))
+    
+    to_na   <- !is.na(keys) & (keys == "<blank>" | keys %in% tokens)
+    to_trim <- !is.na(x) & (x != x_trim) & !to_na
+    
+    # Count tokens converted in this column
+    if (any(to_na)) {
+      tab <- table(keys[to_na])
+      for (k in names(tab)) token_counts[k] <- token_counts[k] + as.integer(tab[[k]])
+    }
+    
+    # Apply changes
+    if (any(to_na))   set(DT, which(to_na), col, NA_character_)
+    if (any(to_trim)) set(DT, which(to_trim), col, x_trim[to_trim])
+    
+    added <- sum(to_na)
+    total_to_na <- total_to_na + added
+    if (added > 0L || any(to_trim)) {
+      cat(sprintf(" - %-25s: %6d → NA%s\n",
+                  col, added, if (any(to_trim)) " (whitespace trimmed)" else ""))
+    }
+  }
+  
+  # Summary
+  cat("\nSummary of values converted to NA by token:\n")
+  idx <- if (report_zeroes) seq_along(token_counts) else which(token_counts > 0L)
+  if (length(idx) == 0L) {
+    cat("   (none)\n")
+  } else {
+    ord <- order(token_counts[idx], decreasing = TRUE)
+    for (j in idx[ord]) cat(sprintf("   %-10s : %d\n", names(token_counts)[j], token_counts[j]))
+  }
+  
+  cat(sprintf("\nTotal values normalized to NA: %d\n", total_to_na))
+  cat("Done.\n")
+  
[…]

### 2025-11-28 data_anomalies/README.Rmd

---
title: "README"
output: html_document
date: "2025-02-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# NYC 311 Data Analysis Apps

## Setup
1. Clone/download this repository
2. Install required R packages:
   ```R
   install.packages(c("shiny", "here", "other_packages_you_use"))

### 2025-11-28 manuscript/manuscript.tex

@@
 
-While one can likely infer the meaning of these \textit{computed} 
+While one can likely infer the meaning of these \emph{computed} 
 fields, uncertainty remains as to their derivation and 
@@
 significant data validity issues, leading us to categorize these six fields 
-as experimental and \textit{not for public use}. An update 
+as experimental and \emph{not for public use}. An update 
 to the Data Dictionary would seem appropriate. Accordingly, 
@@
 presenting missing data. We discovered different 
-representations of missing data: nulls, spaces, ``NA'', ``N/A'', 
-and ``<NA>''. This variety of representations complicates programming
+representations of missing data: nulls, spaces, ``\texttt{NA}'', ``\texttt{N/A}'', 
+and ``\texttt{<NA>}''. This variety of representations complicates programming
 efforts and increases the likelihood of introducing errors 
@@
 
-\paragraph{Challenges with daylight saving time (DST):} The 311 SR 
+\paragraph{Challenges with Daylight Saving Time (DST):} The 311 SR 
 dataset captures local NYC times. Accordingly, twice each 
@@
 in a bizarre scenario where, for example, an SR can 
-be \textit{created} at 1:59AM and then be \textit{closed} 30 minutes later 
+be \emph{created} at 1:59AM and then be \emph{closed} 30 minutes later 
 at 1:29AM; seemingly nonsensical. Utilizing UTC time 
@@
 as to why some fields are mostly/almost always blank. 
-Figure~\ref{fig:blank_fields} is a graphic depiction of the \%
+Figure~\ref{fig:blank_fields} is a graphic depiction of the percentage of
 empty (blank \& N/As) for each field. It illustrates the grouping into 
@@
 field has 0.07\% invalid entries, resulting in 4,374 errors. This includes
-zip codes such as ``12345'', ``10000'', and ``99999'. If you group 
+zip codes such as ``12345'', ``10000'', and ``99999''. If you group 
 the invalid zip code entries by percentage by agency, it 
-mirrors the \textit{overall} breakdown of SRs by Agency, potentially 
+mirrors the \emph{overall} breakdown of SRs by Agency, potentially 
 indicating a systemic problem.
@@
 \begin{itemize}[left=1.5em]
-    \item 12,450 SRs with a \texttt{closed\_date} before \texttt{created\_date} creating a nonsensical ``negative duration''. 
+\item 12,450 SRs with a \texttt{closed\_date} before
+  \texttt{created\_date} creating a nonsensical ``negative duration''. 
 \item Eight SRs with \texttt{created\_date(s)} or \texttt{closed\_date(s)} in 
-    the distant past, specifically  \textit{1900-01-01}. Despite few of 
+  the distant past, specifically  \texttt{1900-01-01}. Despite few of 
   these entries, such extreme values can greatly distort analytical efforts. 
@@
 Citizens, NYC Government Officials, and Agencies use the \texttt{created\_date}(s) and 
-\texttt{closed\_date}(s) to measure the \textit{duration} of SRs, 
+\texttt{closed\_date}(s) to measure the \emph{duration} of SRs, 
 a useful surrogate to measure an agency's responsiveness. 
@@
 Department of Homeless Services (DHS); all contain an entry 
-of \textit{1900-01-01} as the \texttt{closed\_date}. This results 
+of \texttt{1900-01-01} as the \texttt{closed\_date}. This results 
 in very large negative durations exceeding $-$44,601 
@@
 
-Excluding the extreme negative values (any that are <$-$730 days), 
+Excluding the extreme negative values (any that are lower than $-$730 days), 
 Figure~\ref{fig:negative-duration-violin} shows the broad spread of 
 negative duration SRs. While there are few outliers, the magnitude 
-of these negative durations (some <$-$300 days) is troubling 
+of these negative durations (some lower than $-$300 days) is troubling 
 and can produce bizarre analytical results such as greatly 
@@
   \centering
-  \caption{Sample  negative durations [excluding extreme values that are <$-$730 days]}
+  \caption{Sample  negative durations [excluding extreme values that
+    are lower than $-$730 days]}
    \label{tab:largest-errors}
@@
 values $>730$ days, an arbitrary cutoff to exclude 
-infeasible dates such as \textit{1900-01-01}. The chart 
+infeasible dates such as \texttt{1900-01-01}. The chart 
 highlights the distribution of these post-closure 
@@
 \texttt{incident address}. We found 88\% of both of these two 
-street pairs ( 1 \& 2) to be duplicates. Such a high level of duplication 
+street pairs (1 \& 2) to be duplicates. Such a high level of duplication 
 raises the question of which field-pair should be trusted; 
@@
 conventions. For example, we sought to match entries such as
-\textit{KANE PL} and \textit{KANE PLACE} or \textit{AVE P} and
-\textit{AVENUE P}. However, this standardization effort was
+\texttt{KANE PL} and \texttt{KANE PLACE} or \texttt{AVE P} and
+\texttt{AVENUE P}. However, this standardization effort was
 ultimately abandoned, as it resulted in less than a 1\% increase in
@@
 Using \texttt{complaint\_type} as an example,
-instead of repeatedly storing the full text of each \texttt{complaint\_type}, 
+instead of repeatedly storing the full text of each possible value,
 one could use a numeric representation. For example, 
-``NOISE - STREET/SIDEWALK'' could be represented by a 
-numeric code (e.g. ``126''). For this dataset the 
-``NOISE - STREET/SIDEWALK'' \texttt{complaint\_type} occurs 
+``\texttt{NOISE - STREET/SIDEWALK}'' could be represented by a 
+numeric code. For this dataset this code occurs 
 over 300,000 times. Encoding not only reduces file 
@@
 example, the Data Dictionary indicates 
-that the \texttt{status} field takes on values of \textit{assigned, canceled, 
-closed}, or \textit{pending}. However, we observed additional values 
-such as \textit{in progress, started}, and \textit{unspecified}. Additionally, 
-no SRs were found with a status of \textit{canceled}. Similarly, 
+that the \texttt{status} field takes on values of \texttt{assigned, canceled, 
+closed}, or \texttt{pending}. However, we observed additional values 
+such as \texttt{in progress, started}, and \texttt{unspecified}. Additionally, 
+no SRs were found with a status of \texttt{canceled}. Similarly, 
 the Data Dictionary for the \texttt{address\_type} field lists as acceptable
-values \textit{address, blockface, intersection, latlong}, and 
-\textit{placename}. However, we found additional values such 
-as \textit{bbl} and \textit{unrecognized}. Additionally, no 
-\textit{latlong} values were observed. Other fields, such 
+values \texttt{address, blockface, intersection, latlong}, and 
+\texttt{placename}. However, we found additional values such 
+as \texttt{bbl} and \texttt{unrecognized}. Additionally, no 
+\texttt{latlong} values were observed. Other fields, such 
 as \texttt{facility\_type}, \texttt{vehicle\_type}, \texttt{taxi\_pick\_up\_location},
@@
 remains clean and reliable as it is continuously updated.
-
-Transparency is a core tenet of open data, and governments must ensure that datasets are not only accurate but also easily accessible and readily understandable.
-**This includes publishing clear metadata, which enables users to understand the dataset’s structure, limitations, and intended use cases.** Additionally, governments should communicate transparently about data quality issues, such as notifying the public when errors in SR fields (e.g., erroneous \texttt{closed\_date(s)}) are discovered, including an explanation about how such issues will be resolved.
-By ensuring transparency and keeping the public informed, governments foster trust and promote wider use of datasets for research, civic engagement, and innovation.
 
@@
 monitor the quality of the data, as measured by agreed upon
-standards, such as the \% of invalid zip codes. Such tools could
+standards, such as the percentage of invalid zip codes. Such tools could
 capture and display the quality of the data sets as presented
